### 2025-02-05 brms_meeg.qmd
initial commit

---
title: "Modelling M/EEG data with Bayesian nonparametric multilevel models"
shorttitle: "Bayesian M/EEG modelling"
author:
  - name: Ladislas Nalborczyk
    corresponding: true
    orcid: 0000-0002-7419-9855
    email: ladislas.nalborczyk@cnrs.fr
    url: https://lnalborczyk.github.io
    affiliations:
      - id: id1
        name: "Aix Marseille Univ, CNRS, LPL"
        address: "5 avenue Pasteur"
        city: "13100 Aix-en-Provence"
        country: France
  - name: Paul Bürkner
    orcid: 0000-0001-5765-8995
    url: https://paulbuerkner.com
    affiliations:
      - name: "TU Dortmund University, Department of Statistics"
author-note:
  disclosures:
    conflict-of-interest: "The authors have no conflicts of interest to disclose."
    # financial-support: "This study was supported by Grant [Grant Number] from [Funding Source]."
abstract: "Time-resolved electrophysoiological measurements such as those offered by magneto- or electro-encephalography (M/EEG) provide a unique window onto neural activity underlying cognitive process and how they unfold over time. Typically, we are interested in testing whether such measures differ across conditions and/or groups. The conventional approach consists in conducting mass-univariate statistics followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster-based inference. However, these cluster-based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the onset and offset of effects (e.g., differences across conditions). Here, we introduce a novel *model-based approch* for analysing one-dimensional M/EEG timeseries such as ERPs or decoding timecourses and their differences across conditions. This approach relies on Bayesian nonparametric multilevel modelling (multilevel generalised additive models or Gaussian processes), which outputs posterior probabililities of the effect being above chance at every timestep/voxel, while naturally taking into account the temporal dependencies and between-subject variability present in such data."
keywords: [EEG, MEG, generalised additive models, gaussian processes, mixed-effects models, Bayesian statistics, brms]
floatsintext: true
numbered-lines: true
bibliography: meeg_modelling.bib
suppress-title-page: false
link-citations: true
# see https://www.overleaf.com/learn/latex/Font_typefaces
# and systemfonts::system_fonts()
# pdf-engine: xelatex
# monofont: Roboto
mainfont: CMU Serif
fontsize: 10pt
draftfirst: false
draftall: false
a4paper: true
lang: en
toc: true
language:
  citation-last-author-separator: "&"
  email: "email"
  title-block-author-note: "Author note"
  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
format:
  # apaquarto-html: default
  apaquarto-pdf:
    documentmode: doc # man, jou, doc, stu
    keep-tex: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# loading packages
library(changepoint)
library(tidyverse)
library(tidybayes)
library(patchwork)
library(MetBrewer)
library(permuco)
library(scales)
library(scico)
library(knitr)
library(brms)

# setting knitr options
knitr::opts_chunk$set(
    cache = TRUE,
    eval = TRUE,
    echo = TRUE,
    warning = FALSE,
    message = FALSE,
    fig.align = "center",
    out.width = "100%",
    fig.asp = 0.75,
    fig.pos = "!htb",
    dev = "cairo_pdf"
    )

# setting a default ggplot2 theme
# theme_set(theme_light(base_size = 10, base_family = "CMU Sans Serif") )
theme_set(theme_light(base_size = 10, base_family = "Open Sans") )
```

# Introduction (in progress)

Here are some useful references to be discussed [@combrisson_exceeding_2015; @ehinger_unfold_2019; @hayasaka_validating_2003; @luck_how_2017; @pedersen_hierarchical_2019; @riutort-mayol_practical_2023; @rousselet_using_2025; @frossard2021; @frossard2022; @gramfort2013; @maris2007; @pernet2015]... See also [@maris2011]... and [@rosenblatt2018] (history of cluster-based approaches and using a data split?)... Cluster failure [@eklund2016]...

In the following, we consider two approaches to modelling the non-linear timecourse of M/EEG ERPs or decoding performance: i) generalised additive models (GAMs) with thin-plate smoothing splines [@wood2003; @wood2004] and ii) Gaussian processes (GPs) with a smooth covariance kernel [@rasmussen2005] and low-rank approximation [@riutort-mayol_practical_2023].

## Previous modelling work

Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]... Using Bayes factors [@teichmann2022]...

## Generalised additive models

In generalised additive models (GAMs), the functional relationship between predictors and response variable is decomposed into a sum of low-dimensional non-parametric functions. A typical GAM has the following form:

$$
\begin{aligned} 
y_{i} &\sim \mathrm{EF}\left(\mu_{i}, \phi\right)\\
g\left(\mu_i\right) &= \eta_{i} = A_{i} + \mathbf{X}_{i} \gamma + \sum_{j=1}^{J} f_{j}\left(x_{ij}\right)
\end{aligned}
$$

where $g(\cdot)$ is the link function, $A$ is an offset, $\mathbf{X}_{i}$ is the $i$ th row of a parametric model matrix, $\gamma$ is a vector of parameters for the parametric terms, $f_{j}$ is a smooth function of covariate $x_{j}$, $y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)$ denotes that the observations $y_{i}$ are distributed as some member of the exponential family of distributions (e.g., Gaussian, Gamma, Beta, Poisson) with mean $\mu_{i}$ and scale parameter $\phi$.

The smooth functions $f_{j}$ are represented in the model via penalised splines basis expansions of the covariates, that are a weighted sum of basis functions:

$$
f_{j}\left(x_{i j}\right) = \sum_{k=1}^K \beta_{jk} b_{jk}\left(x_{ij}\right)
$$

where $\beta_{jk}$ is the weight (coefficient) associated with the $k$ th basis function $b_{j k}()$ evaluated at the covariate value $x_{ij}$ for the $j$ th smooth function $f_{j}$...

## Gaussian process regression

A Gaussian process (GP) is a stochastic process that defines the distribution over a collection of random variables indexed by a continuous variable, that is $\{f(t): t \in \mathcal{T}\}$ for some index set $\mathcal{T}$ [@riutort-mayol_practical_2023; @rasmussen2005].

Note that nonparametric does not mean there aren't parameters, it means that there are infinitely many parameters.

From [brms documentation](https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp): A GP is a stochastic process, which describes the relation between one or more predictors $x=\left(x_1, \ldots, x_d\right)$ and a response $f(x)$, where $d$ is the number of predictors. A GP is the generalization of the multivariate normal distribution to an infinite number of dimensions. Thus, it can be interpreted as a prior over functions. The values of $f()$ at any finite set of locations are jointly multivariate normal, with a covariance matrix defined by the covariance kernel $k_p\left(x_i, x_j\right)$, where $p$ is the vector of parameters of the GP:

$$
\left(f\left(x_1\right), \ldots f\left(x_n\right) \sim \operatorname{MVN}\left(0,\left(k_p\left(x_i, x_j\right)\right)_{i, j=1}^n\right)\right.
$$

The smoothness and general behaviour of the function $f$ depends only on the choice of covariance kernel, see <https://mc-stan.org/docs/functions-reference/matrix_operations.html#gaussian-process-covariance-functions>. It ensures that values that are close together in the input space will be mapped to similar output values... From this perspective, $f$ is a realisation of an infinite dimensional normal distribution:

$$
f \sim \mathrm{Normal}(0, C(\lambda))
$$

where $C$ is a covariance kernel with hyperparameters $\lambda$ that defines the covariance between two function values $f\left(t_1\right)$ and $f\left(t_2\right)$ for two time points $t_1$ and $t_2$ [@rasmussen2005]. Similar to the different choices of the basis function for splines, different choices of the covariance kernel lead to different GPs. In this article, we consider the squared-exponential (a.k.a. Gaussian or radial basis function) kernel, which computes the squared distance between points and converts it into a measure of similarity. It is defined as:

$$
C(\lambda):=C\left(t_1, t_2, \sigma, \gamma\right):=\sigma^2 \exp \left(-\frac{\left(t_1-t_2\right)^2}{2 \gamma^2}\right)
$$

with hyperparameters $\lambda=(\sigma, \gamma)$, expressing the overall scale of GP and the length-scale, respectively [@rasmussen2005]. The advantages of this kernel are that it is computationally efficient and (infinitely) smooth making it a reasonable choice for the purposes of the present article. Here again, $\lambda$ hyperparameters are estimated from the data, along with all other model parameters.

Low-rank approximate Gaussian processes are of main interest in machine learning and statistics due to the high computational demands of exact Gaussian process models [@riutort-mayol_practical_2023]...

## Objectives

...

# Methods

## M/EEG data simulation

Following the approach used by @sassenhagen2019 and @rousselet_using_2025, we simulated EEG data stemming from two conditions, one with noise only, and the other with noise + signal. As in previous studies, the noise was generated by superimposing 50 sinusoids at different frequencies, following an EEG-like spectrum [see details and code in @yeung2004]...

For simulating EEG data, see <https://github.com/GRousselet/onsetsim> and <https://github.com/unfoldtoolbox/UnfoldSim.jl>...

Copy-pasted from @rousselet_using_2025: "The signal was a truncated Gaussian defining an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms. It was created by getting the probability density function of a standard normal with 93 points from x= 1.5 to 1.5, rescaled in amplitude between 0 and 1 and padded with 79 zeros to the left and to the right, to form a time series of 251 points, from 0 to 500 ms, in steps of 2 ms (500 Hz sampling rate)."

```{r eeg-group, echo = FALSE, fig.width = 9, fig.cap = "Some ERPs in two conditions with 50 trials each, for a group of 20 participants."}
# importing R version of Matlab code from Yeung et al. (2004)
source("code/eeg_noise.R")

# importing the ERP template with true onset = 160 ms, F=81, and max at F=126
source("code/erp_template.R")

# to use with the eeg_noise function
meanpower <- unlist(read.table("code/meanpower.txt") )

# defining simulation parameters
n_trials <- 50 # number of trials
n_ppt <- 20 # number of participants
outvar <- 1 # noise variance
srate <- 500 # sampling rate in Hz
ronset <- seq(150, 170, 2) # random onset for each participant

for (P in 1:n_ppt) { # for each participant

    # get random onset
    ponset <- sample(ronset, 1)

    # find starting point
    st <- which(Xf == ponset)

    # pad vector
    temp2 <- c(rep(0, st - 2), erp, rep(0, Nf - st - length(erp) + 2) )

    # initialising empty conditions
    cond1 <- matrix(0, nrow = n_trials, ncol = Nf)
    cond2 <- matrix(0, nrow = n_trials, ncol = Nf)

    for (T in 1:n_trials) { # for each trial

        cond1[T, ] <- temp1 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
        cond2[T, ] <- temp2 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)

    }

    # converting results to dataframe
    temp_df <- data.frame(
        x = rep(Xf, 2*nrow(cond1) ),
        y = c(c(t(cond1) ), c(t(cond2) ) ),
        trial = c(rep(1:n_trials, each = length(Xf) ), rep(1:n_trials, each = length(Xf) ) ),
        condition = factor(rep(c("cond1", "cond2"), each = Nf*n_trials) ),
        participant = paste0("participant_", sprintf("%02d", P) )
        ) %>%
        select(participant, condition, trial, x, y)

    # and appending it to previous participants
    if (exists("raw_df") ) {

        raw_df <- bind_rows(raw_df, temp_df)

    } else {

        raw_df <- temp_df

    }

}

# plotting the data
raw_df %>%
    summarise(y = mean(y), .by = c(participant, condition, x) ) %>%
    ggplot(aes(x, y, colour = condition) ) +
    geom_hline(yintercept = 0, linetype = 2) +
    geom_vline(xintercept = true_onset, linetype = 2) +
    geom_line(show.legend = FALSE) +
    facet_wrap(~participant) +
    labs(x = "Time (ms)", y = "EEG signal (a.u.)")
```

```{r erp-diff, out.width = "75%", fig.cap = "Group-level average difference between conditions (mean +/- standard error of the mean). The 'true' (population value of the) onset is indicated by the vertical dashed line."}
# averaging across participants
group_df <- raw_df %>%
    pivot_wider(names_from = condition, values_from = y, values_fn = mean) %>%
    mutate(y_diff = cond2 - cond1) %>%
    summarise(
        y_mean = mean(y_diff),
        y_sem = sd(y_diff) / sqrt(n() ),
        .by = x
        ) %>%
    mutate(lower = y_mean - y_sem, upper = y_mean + y_sem)

# plotting the data
group_df %>%
    ggplot(aes(x = x, y = y_mean) ) +
    geom_hline(yintercept = 0, linetype = 2) +
    geom_vline(xintercept = true_onset, linetype = 2) +
    geom_ribbon(
        aes(ymin = lower, ymax = upper, colour = NULL),
        alpha = 0.25, show.legend = FALSE
        ) +
    geom_line(show.legend = FALSE) +
    labs(x = "Time (ms)", y = "EEG difference (a.u.)")
```

## Model fitting

Models were fitted using the `brms` package [@burkner_brms_2017; @burkner_advanced_2018; @nalborczyk2019]...

```{r gam, message = FALSE}
# defining a contrast for condition
contrasts(group_df$condition) <- c(-0.5, 0.5)

# fitting the GAM
gam <- brm(
    # cubic regression splines with k-1 basis functions
    # then we should try s(x, participant, bs = "fs")
    # that is, random factor smooth interactions...
    y ~ condition + s(x, bs = "cr", k = 20, by = condition),
    data = group_df,
    family = gaussian(),
    iter = 5000,
    chains = 4,
    cores = 4,
    file = "models/gam.rds"
    )
```

Now we fit the GP...

```{r gp, message = FALSE}
gp_model <- brm(
    # k refers to the number of basis functions for
    # computing Hilbert-space approximate GPs
    y ~ gp(x, k = 20, by = condition),
    # if k = NA (default), exact GPs are computed
    # y ~ gp(x, by = condition),
    data = group_df,
    family = gaussian(),
    control = list(adapt_delta = 0.99),
    iter = 5000,
    chains = 4,
    cores = 4,
    file = "models/gp.rds"
    )
```

And we plot the posterior predictions...

```{r gam-preds, echo = FALSE}
# defining a function to plot posterior predictions
plot_post_preds <- function (model, data) {

    post_preds <- conditional_effects(
        x = model,
        effect = "x:condition",
        method = "posterior_epred",
        re_formula = NULL,
        prob = 0.99
        )[[1]]

    post_preds %>%
        ggplot(aes(x = x, y = y, colour = condition, fill = condition) ) +
        geom_hline(yintercept = 0, linetype = 2) +
        geom_point(
            data = data %>% summarise(y = mean(y), .by = c(x, condition) ),
            aes(fill = NULL),
            pch = 21, show.legend = FALSE
            ) +
        geom_line(aes(y = estimate__), show.legend = FALSE) +
        geom_ribbon(
            aes(ymin = lower__, ymax = upper__, colour = NULL),
            alpha = 0.25, show.legend = FALSE
            ) +
        labs(x = "Time (ms)", y = "EEG signal (a.u.)")

}
```

```{r gam-gp-preds, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior predictions of the GAM (left) and GP (right) models."}
# plotting the posterior predictions
plot_post_preds(model = gam, data = group_df) +
    plot_post_preds(model = gp_model, data = group_df)
```

## Posterior probability of difference above 0

We can retrieve the posterior probability of the slope for `condition` at each timestep.

```{r, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Slope for the difference between conditions according to the GAM (left) or the GP (right)."}
# defining a function to plot posterior slope
plot_post_slope <- function (model, data, eps = 0.1) {

  # defining a sequence of x values to make predictions
  x_seq <- crossing(
      x = seq(min(data$x), max(data$x), length.out = 100),
      condition = c("cond1", "cond2")
      ) %>%
      arrange(condition)

  # retrieving posterior samples
  posterior_samples <- posterior_epred(object = model, newdata = x_seq)

  # computing the difference between cond2 and cond1
  posterior_samples <- posterior_samples[, 101:200] - posterior_samples[, 1:100]

  # computing the probability that y is above 0
  prob_y_above_0 <- data.frame(
      x = seq(min(data$x), max(data$x), length.out = 100)
      ) %>%
      mutate(
          m = colMeans(posterior_samples),
          lower = apply(
              X = posterior_samples,
              MARGIN = 2, FUN = quantile, probs = 0.025
              ),
          upper = apply(
              X = posterior_samples,
              MARGIN = 2, FUN = quantile, probs = 0.975
              )
          )

  # plotting it
  prob_y_above_0 %>%
        ggplot(aes(x = x, y = m) ) +
        geom_hline(yintercept = 0, linetype = 2) +
        geom_line() +
        geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
        # theme_light(base_size = 12) +
        labs(x = "Time (ms)", y = expression(Pr(beta~"|"~data, model) ) )

}

# using the function with model
plot_post_slope(model = gam, data = group_df, eps = 0.1) +
    plot_post_slope(model = gp_model, data = group_df, eps = 0.1)
```

We can also compute the posterior probability of the slope for `condition` being above 0 (@fig-post-prob-test).

```{r fig-post-prob-test, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior probability of the EEG signal being above 0+eps according to the GAM (left) and the GP (right)."}
# defining a function to plot posterior prob of slope above 0
plot_post_test <- function (model, data, eps = 0.1) {

    # defining a sequence of x values to make predictions
    x_seq <- crossing(
        x = seq(min(data$x), max(data$x), length.out = 100),
        condition = c("cond1", "cond2")
        ) %>%
        arrange(condition)

    # retrieving posterior samples
    posterior_samples <- posterior_epred(object = model, newdata = x_seq)

    # computing the difference between cond2 and cond1
    posterior_samples <- posterior_samples[, 101:200] - posterior_samples[, 1:100]

    # computing the probability that y is above 0
    prob_y_above_0 <- data.frame(
        x = x_seq$x,
        m = colMeans(posterior_samples > (0 + eps) ),
        lower = apply(
            X = posterior_samples > (0 + eps),
            MARGIN = 2, FUN = quantile, probs = 0.025
            ),
        upper = apply(
            X = posterior_samples > (0 + eps),
            MARGIN = 2, FUN = quantile, probs = 0.975
            )
        )

    # plotting it
    prob_y_above_0 %>%
    ggplot(aes(x = x, y = m) ) +
    geom_hline(yintercept = 0.5, linetype = 2) +
    geom_line() +
    # geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
    # theme_light(base_size = 12) +
    # labs(x = "Time (ms)", y = "Pr(EEG>0)|data)")
    labs(x = "Time (ms)", y = expression(Pr(beta>0~"|"~data, model) ) )

}

# using the function with the two models
plot_post_test(model = gam, data = group_df, eps = 0.1) +
    plot_post_test(model = gp_model, data = group_df, eps = 0.1)
```

We can also express this as the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional BF thresholds (@fig-post-prob-ratio).

```{r fig-post-prob-ratio, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Ratio of posterior probability according to the GAM (left) and the GP (right). Timesteps above threshold (>10) are highlighted in green."}
# defining a sequence of x values to make predictions
n_approx <- 1000
x_seq <- crossing(
    x = seq(min(group_df$x), max(group_df$x), length.out = n_approx),
    condition = c("cond1", "cond2")
    ) %>%
    arrange(condition)

# retrieving posterior samples
posterior_samples <- posterior_epred(object = gam, newdata = x_seq)

# computing the difference between cond2 and cond1
posterior_samples <- posterior_samples[, (ncol(posterior_samples)/2+1):ncol(posterior_samples)] -
    posterior_samples[, 1:(ncol(posterior_samples)/2)]

# sanity checks
# ts.plot(colMeans(posterior_samples) )
# ts.plot(apply(X = posterior_samples, MARGIN = 2, FUN = mean) )
# ts.plot(apply(X = posterior_samples, MARGIN = 2, FUN = sd) )
# mean(apply(X = posterior_samples, MARGIN = 2, FUN = sd) )

# computing the probability that y is above 0
eps <- 0.05
prob_y_above_0 <- data.frame(
    x = seq(min(group_df$x), max(group_df$x), length.out = n_approx)
    ) %>%
    mutate(post_prob = colMeans(posterior_samples) ) %>%
    mutate(m = colMeans(posterior_samples > eps) ) %>%
    mutate(prob_ratio = m / (1 - m) )

# printing the identified clusters
# the signal was a truncated Gaussian defining an objective onset at 160 ms,
# a maximum at 250 ms, and an offset at 342 ms.
threshold <- 10
exceeding_times_gam <- prob_y_above_0 %>%
    dplyr::filter(prob_ratio > threshold) %>%
    summarise(cluster_onset = min(x), cluster_offset = max(x) ) %>%
    mutate(
        cluster_peak = prob_y_above_0 %>%
            dplyr::filter(post_prob == max(post_prob) ) %>%
            pull(x)
        ) %>%
    mutate(true_onset = 160, true_peak = 250, true_offset = 342) %>%
    mutate(model = "GAM")

# plotting it
gam_ratio_plot <- prob_y_above_0 %>%
    mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
    ggplot(aes(x = x, y = prob_ratio) ) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    geom_hline(yintercept = 3, linetype = "dashed", color = "darkred") +
    geom_hline(yintercept = 1/3, linetype = "dashed", color = "darkred") +
    geom_hline(yintercept = 10, linetype = "dashed", color = "red") +
    geom_hline(yintercept = 1/10, linetype = "dashed", color = "red") +
    geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
    geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
    geom_line() +
    geom_point(
        data = . %>% dplyr::filter(!is.na(above_thres) ),
        aes(y = threshold),
        colour = "darkgreen",
        shape = 15,
        size = 3,
        na.rm = TRUE
        ) +
    scale_y_log10(labels = label_log(digits = 2), limits = c(NA, 1e4) ) +
    labs(
        x = "Time (ms)",
        y = expression(Pr(beta>0~"|"~data, model) / (1 - Pr(beta>0~"|"~data, model) ) )
        )

# retrieving posterior samples
posterior_samples <- posterior_epred(object = gp_model, newdata = x_seq)

# computing the difference between cond2 and cond1
posterior_samples <- posterior_samples[, (ncol(posterior_samples)/2+1):ncol(posterior_samples)] -
    posterior_samples[, 1:(ncol(posterior_samples)/2)]

# computing the probability that y is above 0
eps <- 0.05
prob_y_above_0 <- data.frame(
    x = seq(min(group_df$x), max(group_df$x), length.out = n_approx)
    ) %>%
    mutate(post_prob = colMeans(posterior_samples) ) %>%
    mutate(m = colMeans(posterior_samples > eps) ) %>%
    mutate(prob_ratio = m / (1 - m) )

# printing the identified clusters
# the signal was a truncated Gaussian defining an objective onset at 160 ms,
# a maximum at 250 ms, and an offset at 342 ms.
threshold <- 10
exceeding_times_gp <- prob_y_above_0 %>%
    dplyr::filter(prob_ratio > threshold) %>%
    summarise(cluster_onset = min(x), cluster_offset = max(x) ) %>%
    mutate(
        cluster_peak = prob_y_above_0 %>%
            dplyr::filter(post_prob == max(post_prob) ) %>%
            pull(x)
        ) %>%
    mutate(true_onset = 160, true_peak = 250, true_offset = 342) %>%
    mutate(model = "GP")

# plotting it
gp_ratio_plot <- prob_y_above_0 %>%
    mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
    ggplot(aes(x = x, y = prob_ratio) ) +
    geom_hline(yintercept = 1, linetype = "dashed") +
    geom_hline(yintercept = 3, linetype = "dashed", color = "darkred") +
    geom_hline(yintercept = 1/3, linetype = "dashed", color = "darkred") +
    geom_hline(yintercept = 10, linetype = "dashed", color = "red") +
    geom_hline(yintercept = 1/10, linetype = "dashed", color = "red") +
    geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
    geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
    geom_line() +
    geom_point(
        data = . %>% dplyr::filter(!is.na(above_thres) ),
        aes(y = threshold),
        colour = "darkgreen",
        shape = 15,
        size = 3,
        na.rm = TRUE
        ) +
    scale_y_log10(labels = label_log(digits = 2), limits = c(NA, 1e4) ) +
    labs(
        x = "Time (ms)",
        y = expression(Pr(beta>0~"|"~data, model) / (1 - Pr(beta>0~"|"~data, model) ) )
        )

# plotting results for the two models
gam_ratio_plot + gp_ratio_plot

# printing the results
# bind_rows(exceeding_times_gam, exceeding_times_gp) %>%
#     mutate(across(.cols = where(is.numeric), .fns = ~round(.x, 2) ) ) %>%
#     relocate(model, .before = cluster_onset)
```

\newpage

## Error properties of the proposed approach

Below we are plotting the error function (where the error is computed as $|\hat{\theta}-\theta|$) for both the onset and offset values of the ERP difference, according to various `eps` and `threshold` values. Remember that the signal is generated from a truncated Gaussian defining an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms. @fig-onset-error shows that the GP model can *exactly* recover the true onset and offset values, given some reasonable choice of `eps` and `threshold` values.

```{r fig-onset-error, echo = FALSE, dpi = 300, fig.width = 8, fig.asp = 0.5, fig.cap = "Error function of onset (left) and offset (right) estimation according to various eps and threshold values (according to the GP model). Minimum error values are indicated by red crosses."}
# defining values of eps and threshold to test
eps_values <- seq(from = 0, to = 0.2, by = 0.01)
threshold_values <- seq(from = 1, to = 50, by = 1)
n_approx <- n_distinct(group_df$x)

# initialising results dataframe
results <- crossing(eps = eps_values, threshold = threshold_values) %>%
    mutate(
        estimated_onset = NA, estimated_peak = NA, estimated_offset = NA,
        error_onset = NA, error_peak = NA, error_offset = NA
        )

# generating a grid of values to compute predictions for
x_seq <- crossing(
    x = seq(min(group_df$x), max(group_df$x), length.out = n_approx*2-1),
    condition = c("cond1", "cond2")
    ) %>%
    arrange(condition)

# retrieving posterior samples
posterior_samples <- posterior_epred(object = gp_model, newdata = x_seq)

# computing the difference between cond2 and cond1
posterior_samples <- posterior_samples[, (ncol(posterior_samples)/2+1):ncol(posterior_samples)] -
    posterior_samples[, 1:(ncol(posterior_samples)/2)]

# looping over different values of eps and threshold
for (i in seq_len(nrow(results) ) ) {

    # retrieving current eps and threshold values
    eps <- results$eps[i]
    threshold <- results$threshold[i]

    # computing probability metrics with the current eps value
    prob_y_above_0 <- data.frame(
        x = seq(min(group_df$x), max(group_df$x), length.out = n_approx*2-1)
        ) %>%
    mutate(post_prob = colMeans(posterior_samples) ) %>%
    mutate(m = colMeans(posterior_samples > eps) ) %>%
    mutate(prob_ratio = m / (1 - m) )

    # finding onset, offset, and peak for the current threshold
    exceeding_times <- prob_y_above_0 %>%
        dplyr::filter(prob_ratio > threshold) %>%
        summarise(
            cluster_onset = min(x, na.rm = TRUE),
            cluster_offset = max(x, na.rm = TRUE)
            ) %>%
        mutate(
            cluster_peak = prob_y_above_0 %>%
            dplyr::filter(post_prob == max(post_prob, na.rm = TRUE) ) %>%
            pull(x)
            ) %>%
        mutate(true_onset = 160, true_peak = 250, true_offset = 342)

    # storing the results in the dataframe
    if (nrow(exceeding_times) > 0) {

        results$estimated_onset[i] <- exceeding_times$cluster_onset
        results$estimated_peak[i] <- exceeding_times$cluster_peak
        results$estimated_offset[i] <- exceeding_times$cluster_offset

        # computing errors
        results$error_onset[i] <- abs(
            exceeding_times$cluster_onset - exceeding_times$true_onset
            )
        results$error_peak[i] <- abs(
            exceeding_times$cluster_peak - exceeding_times$true_peak
            )
        results$error_offset[i] <- abs(
            exceeding_times$cluster_offset - exceeding_times$true_offset
            )

    }

}

# plotting the results
onset_plot <- results %>%
    ggplot(aes(x = eps, y = threshold, fill = error_onset) ) +
    geom_tile(show.legend = FALSE) +
    geom_point(
        data = results %>% dplyr::filter(error_onset == min(error_onset, na.rm = TRUE) ),
        aes(x = eps, y = threshold),
        color = "orangered", size = 2, shape = 4,
        show.legend = FALSE
        ) +
    scale_x_continuous(expand = c(0, 0) ) +
    scale_y_continuous(expand = c(0, 0) ) +
    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
    # scale_fill_viridis_c() +
    # scale_y_log10(labels = label_log(digits = 2) ) +
    # theme_light(base_size = 10) +
    labs(
        title = "Onset error",
        x = "eps",
        y = "Threshold",
        fill = "Error"
        )

offset_plot <- results %>%
    ggplot(aes(x = eps, y = threshold, fill = error_offset) ) +
    geom_tile(show.legend = FALSE) +
    geom_point(
        data = results %>% dplyr::filter(error_offset == min(error_offset, na.rm = TRUE) ),
        aes(x = eps, y = threshold),
        color = "orangered", size = 2, shape = 4,
        show.legend = FALSE
        ) +
    scale_x_continuous(expand = c(0, 0) ) +
    scale_y_continuous(expand = c(0, 0) ) +
    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
    # theme_light(base_size = 10) +
    labs(
        title = "Offset error",
        x = "eps",
        y = "Threshold",
        fill = "Error"
        )

# combining the plots
onset_plot + offset_plot

# displaying best parameters values (according to onset error)
# library(flextable)
results %>%
    dplyr::select(-estimated_peak, -error_peak) %>%
    arrange(error_onset) %>%
    data.frame() %>%
    head()

# displaying best parameters values (according to offset error)
# library(flextable)
results %>%
    dplyr::select(-estimated_peak, -error_peak) %>%
    arrange(error_offset) %>%
    data.frame() %>%
    head()
```

\newpage

## Using summary statistics of ERPs

Next we fit a hierarchical GAM using summary statistics of ERPs (mean and SD) at the participant level (similar to what is done in meta-analysis).

```{r meta-gam, message = FALSE}
# averaging across participants
summary_df <- raw_df %>%
    summarise(
        eeg = mean(y),
        eeg_sd = sd(y),
        .by = c(participant, condition, x)
        )

# defining a contrast for condition
contrasts(summary_df$condition) <- c(-0.5, 0.5)

# fitting the GAM
meta_gam <- brm(
    # using by-participant SD of ERPs across trials
    eeg | se(eeg_sd) ~
        condition + s(x, bs = "cr", k = 10, by = condition) +
        (1 | participant),
    data = summary_df,
    family = gaussian(),
    iter = 5000,
    chains = 4,
    cores = 4,
    file = "models/meta_gam.rds"
    )
```

```{r meta-gam-preds, fig.cap = "Hierarchical GAM posterior predictions."}
# plotting the posterior predictions
plot(
    conditional_effects(x = meta_gam, effect = "x:condition"),
    points = FALSE, theme = theme_light(), plot = FALSE
    )[[1]] +
    labs(x = "Time (ms)", y = "EEG signal (a.u.)")
```

```{r meta-gp, eval = FALSE, message = FALSE}
# fitting the GP
meta_gp <- brm(
    # using by-participant SD of ERPs across trials
    eeg | se(eeg_sd) ~
        condition + gp(x, k = 20, by = condition) +
        (1 | participant),
    data = summary_df,
    family = gaussian(),
    control = list(adapt_delta = 0.95, max_treedepth = 20),
    iter = 5000,
    chains = 4,
    cores = 4,
    file = "models/meta_gp.rds"
    )
```

```{r meta-gp-preds, eval = FALSE, fig.cap = "Hierarchical GP posterior predictions."}
# plotting the posterior predictions
plot(
    conditional_effects(x = meta_gp, effect = "x:condition"),
    points = FALSE, theme = theme_light(), plot = FALSE
    )[[1]] +
    labs(x = "Time (ms)", y = "EEG signal (a.u.)")
```

## Application to 1D decoding results (accuracy over time)

Assume we have M/EEG data and we conducted time-resolved multivariate pattern analysis (MVPA), also known as decoding. As a result, we have a timecourse of decoding accuracies (e.g., ROC AUC), bounded between 0 and 1, per participant (@fig-sim-decoding).

```{r fig-sim-decoding, echo = FALSE, out.width = "75%", fig.cap = "Exemplary (simulated) group-level average timecourse of binary decoding accuracy (ROC AUC)."}
# simulating decoding accuracies (ROC AUC)
n_participants <- 20
n_timepoints <- 240
timepoints <- seq(from = 0.1, to = 1.3, length.out = n_timepoints)

# defining a function to generate AUC using a Beta distribution
simulate_auc <- function (
        timepoints, onset = 0.1,
        peak_time = 0.4, sigma = 0.5, amplitude = 30
        ) {

    # ensuring timepoints - onset > 0
    adjusted_time <- timepoints - onset + 0.001

    # log-normal peak function with right skew
    log_peak <- log(peak_time - onset + 0.001)
    log_t <- log(adjusted_time)

    # computing right-skewed alpha with log-normal peak
    alpha_t <- 10 + amplitude * exp(-((log_t - log_peak)^2) / (2 * sigma^2) )

    # keeping baseline at chance level (0.5)
    beta_t <- 10

    # simulate Beta-distributed AUC values
    auc <- rbeta(length(timepoints), shape1 = alpha_t, shape2 = beta_t)

    # returning the AUC
    return (auc)

}

# simulating data for all participants
decoding_data <- data.frame(
    participant = rep(1:n_participants, each = n_timepoints),
    time = rep(timepoints, times = n_participants),
    auc = unlist(lapply(1:n_participants, function (i) {
        simulate_auc(
            timepoints,
            onset = 0.1,
            # variable peak time
            peak_time = runif(1, 0.6, 0.7),
            # variable spread
            sigma = runif(1, 0.1, 0.3),
            # variable peak amplitude
            amplitude = runif(1, 30, 40)
            )})
        )
    ) %>%
    mutate(time = time - 0.3)

# computing group mean and 95% quantile interval
group_data <- decoding_data %>%
    group_by(time) %>%
    summarise(
        mean_auc = mean(auc),
        lower = quantile(auc, 0.025),
        upper = quantile(auc, 0.975)
        )

# plotting it
group_data %>%
    ggplot(aes(x = time, y = mean_auc) ) +
    geom_hline(yintercept = 0.5, linetype = "dashed") +
    geom_vline(xintercept = 0.0, linetype = "dashed") +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "steelblue", alpha = 0.2) +
    geom_line(color = "steelblue", size = 1) +
    ylim(0, 1) +
    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
```

Now, we want to *test* whether the group-level average decoding accuracy is above chance (i.e., 0.5) at each timestep. We use a similar GAM/GP as previously, but we replace the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one.

```{r gam-decoding, message = FALSE}
# fitting the GAM
decoding_gam <- brm(
    auc ~ s(time, bs = "cr", k = 10),
    data = decoding_data,
    family = Beta(),
    iter = 5000,
    chains = 4,
    cores = 4,
    file = "models/decoding_gam.rds"
    )
```

```{r gp-decoding, message = FALSE}
# fitting the GP
decoding_gp <- brm(
    auc ~ gp(time, k = 20),
    data = decoding_data,
    family = Beta(),
    control = list(adapt_delta = 0.9),
    iter = 5000,
    chains = 4,
    cores = 4,
    file = "models/decoding_gp.rds"
    )
```

```{r decoding-preds, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Posterior predictions of the GAM (left) and GP (right) fitted on decoding accuracy over time."}
# plotting the posterior predictions
gam_decoding_preds <- plot(
    conditional_effects(x = decoding_gam, prob = 0.99),
    points = TRUE,
    point_args = list(size = 1, pch = 21, colour = "grey"),
    line_args = list(colour = "black"),
    theme = theme_light(), plot = FALSE
    )[[1]] +
    geom_hline(yintercept = 0.5, linetype = 2) +
    geom_vline(xintercept = 0.0, linetype = 2) +
    ylim(0, 1) +
    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")

gp_decoding_preds <- plot(
    conditional_effects(x = decoding_gp, prob = 0.99),
    points = TRUE,
    point_args = list(size = 1, pch = 21, colour = "grey"),
    line_args = list(colour = "black"),
    theme = theme_light(), plot = FALSE
    )[[1]] +
    geom_hline(yintercept = 0.5, linetype = 2) +
    geom_vline(xintercept = 0.0, linetype = 2) +
    ylim(0, 1) +
    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")

# combining the two plots
gam_decoding_preds + gp_decoding_preds
```

Next, we plot the posterior probability of decoding accuracy being above chance level (plus some epsilon) (@fig-decoding-post).

```{r fig-decoding-post, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Posterior probability of decoding accuracy being above chance level according to the GAM (left) or the GP (right)."}
# defining a function to plot posterior prob of slope above 0
plot_post_decoding <- function (model, data, chance_level = 0.5, eps = 0.1) {

    # defining a sequence of x values to make predictions
    x_seq <- crossing(time = seq(min(data$time), max(data$time), length.out = 100) )

    # retrieving posterior samples
    posterior_samples <- posterior_epred(object = model, newdata = x_seq)

    # computing the probability that y is above 0
    prob_y_above_0 <- data.frame(
        time = x_seq$time,
        m = colMeans(posterior_samples > (chance_level + eps) ),
        lower = apply(
            X = posterior_samples > (chance_level + eps),
            MARGIN = 2, FUN = quantile, probs = 0.025
            ),
        upper = apply(
            X = posterior_samples > (chance_level + eps),
            MARGIN = 2, FUN = quantile, probs = 0.975
            )
        )

    # plotting it
    prob_y_above_0 %>%
    ggplot(aes(x = time, y = m) ) +
    geom_hline(yintercept = chance_level, linetype = 2) +
    geom_vline(xintercept = 0, linetype = 2) +
    geom_line() +
    # geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
    labs(x = "Time (s)", y = expression(Pr(AUC>chance~"|"~data, model) ) )

}

# using the function with the two models
plot_post_decoding(
    model = decoding_gam, data = decoding_data,
    chance_level = 0.5, eps = 0.1
    ) +
    plot_post_decoding(
        model = decoding_gp, data = decoding_data,
        chance_level = 0.5, eps = 0.1
        ) 
```

```{r fig-decoding-ratio, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Ratio of posterior probabilities of decoding accuracy being above chance level according to the GAM (left) or the GP (right)."}
# defining a function to plot posterior prob of slope above 0
plot_post_decoding_ratio <- function (model, data, chance_level = 0.5, eps = 0.1) {

    # defining a sequence of x values to make predictions
    # x_seq <- crossing(time = seq(min(data$time), max(data$time), length.out = 100) )
    x_seq <- data.frame(time = data$time)

    # retrieving posterior samples
    posterior_samples <- posterior_epred(object = model, newdata = x_seq)

    # computing the probability that y is above 0
    prob_y_above_0 <- data.frame(time = data$time) %>%
        mutate(m = colMeans(posterior_samples > (chance_level + eps) ) ) %>%
        mutate(prob_ratio = m / (1 - m) )

    # printing the identified clusters
    threshold <- 10
    exceeding_times <- prob_y_above_0 %>%
        dplyr::filter(prob_ratio > threshold) %>%
        summarise(cluster_onset = min(time), cluster_offset = max(time) )
    print(exceeding_times)

    # plotting it
    prob_y_above_0 %>%
        mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
        ggplot(aes(x = time, y = prob_ratio) ) +
        geom_hline(yintercept = 1, linetype = "dashed") +
        geom_hline(yintercept = 3, linetype = "dashed", color = "darkred") +
        geom_hline(yintercept = 1/3, linetype = "dashed", color = "darkred") +
        geom_hline(yintercept = 10, linetype = "dashed", color = "red") +
        geom_hline(yintercept = 1/10, linetype = "dashed", color = "red") +
        geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
        geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
        geom_vline(xintercept = 0, linetype = 2) +
        geom_line() +
        geom_point(
            data = . %>% dplyr::filter(!is.na(above_thres) ),
            aes(y = threshold),
            colour = "darkgreen",
            shape = 15,
            size = 3,
            na.rm = TRUE
            ) +
        scale_y_log10(labels = label_log(digits = 2), limits = c(NA, 1e4) ) +
        labs(
            x = "Time (s)",
            y = expression(Pr(AUC>chance~"|"~data, model) / (1 - Pr(AUC>chance~"|"~data, model) ) )
            )

}

# using the function with the two models
plot_post_decoding_ratio(
    model = decoding_gam, data = decoding_data,
    chance_level = 0.5, eps = 0.05
    ) +
    plot_post_decoding_ratio(
        model = decoding_gp, data = decoding_data,
        chance_level = 0.5, eps = 0.05
        )
```

\newpage

## Application to 2D decoding results (cross-temporal generalisation)

Assume we have M/EEG data and we have conducted cross-temporal generalisation analyses [@king2014]. As a result, we have a 2D matrix where each element contains the decoding accuracy (e.g., ROC AUC) of a classifier trained at timestep $\text{training}_{i}$ and tested at timestep $\text{testing}_{j}$ (@fig-sim-timegen).

```{r fig-sim-timegen, echo = FALSE, dpi = 300, fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
# number of timepoints
n_participants <- 20
# n_timepoints <- 240 # (1.2s at 200Hz)
n_timepoints <- 60 # (1.2s at 50Hz)
timepoints <- seq(from = -0.2, to = 1, length.out = n_timepoints)

# smooth onset and offset parameters
decode_onset <- 0.2
decode_offset <- 0.3

# smoothness of onset/offset
sigma_smooth <- 0.05

# generating participant-specific cross-temporal generalisation matrices
timegen_data <- tibble()

# for each participant
for (p in 1:n_participants) {

    # adding variability in peak amplitude and width per participant
    peak_variability <- runif(1, 30, 50) # random peak strength (higher = stronger decoding)
    smooth_variability <- runif(1, 0.04, 0.06) # random smoothness variation
    onset_variability <- runif(1, -0.05, 0.05) # small jitter in onset time
    offset_variability <- runif(1, -0.05, 0.05) # small jitter in offset time

    # computing smooth transition function (soft onset and offset with participant variation)
    decode_window <- outer(timepoints, timepoints, function (t1, t2) {

        # smooth onset
        onset_factor <- 1 /
            (1 + exp(-(t1 - (decode_onset + onset_variability) ) / sigma_smooth) )

        # smooth offset
        offset_factor <- 1 /
            (1 + exp((t1 - (decode_offset + offset_variability)) / sigma_smooth) )

        # diagonal Gaussian
        onset_factor * offset_factor * exp(-((t1 - t2)^2) / (2 * smooth_variability^2) )

    })

    # defining smooth alpha and beta parameters for Beta distribution
    base_alpha <- 10
    peak_alpha <- peak_variability
    alpha_matrix <- base_alpha + peak_alpha * decode_window
    beta_matrix <- base_alpha

    # generating AUC values using the Beta distribution
    auc_matrix_beta <- matrix(
        rbeta(n_timepoints^2, shape1 = alpha_matrix, shape2 = beta_matrix),
        nrow = n_timepoints, ncol = n_timepoints
        )

    # converting to long format for plotting
    temp_data <- expand.grid(
        train_time = timepoints, test_time = timepoints
        ) %>%
        mutate(auc = as.vector(auc_matrix_beta), participant = p)

    # storing results
    timegen_data <- bind_rows(timegen_data, temp_data)

}

# plotting cross-temporal generalization matrix
timegen_data %>%
    summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>%
    # pull(auc) %>% hist()
    ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
    geom_tile() +
    geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
    geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
    geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
    scale_x_continuous(expand = c(0, 0) ) +
    scale_y_continuous(expand = c(0, 0) ) + 
    coord_fixed() +
    scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
    # theme_light() +
    labs(x = "Testing time (s)", y = "Training time (s)")
```

Now, we want to test whether and when decoding performance is above chance level (0.5 for a binary decoding task). This one is computationally very intense...

```{r gam-timegen, message = FALSE}
# fitting a GAM with a 2D smooth function
timegen_gam <- brm(
    # 2D thin-plate spline (tp) to model smooth interactions between training and testing time
    auc ~ s(train_time, test_time, bs = "tp", k = 10),
    data = timegen_data,
    family = Beta(),
    iter = 5000,
    chains = 4,
    cores = 4,
    file = "models/timegen_gam.rds"
    )
```

```{r gam-timegen-preds, eval = FALSE, echo = FALSE, fig.cap = "Posterior predictions of the GAM fitted on decoding accuracy over time."}
# plotting the posterior predictions
plot(
    conditional_effects(x = timegen_gam, prob = 0.95),
    points = FALSE,
    # point_args = list(size = 1, pch = 21, colour = "grey"),
    # line_args = list(colour = "black"),
    method = "posterior_epred",
    surface = TRUE,
    theme = theme_light(),
    plot = TRUE
    )#[[1]] +
    # geom_hline(yintercept = 0.5, linetype = 2) +
    # geom_vline(xintercept = 0.0, linetype = 2) +
    # ylim(0, 1) +
    # labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")

# posterior predictive checks
pp_check(timegen_gam)
```

```{r gam-timegen-post-preds, echo = FALSE, fig.width = 8, fig.asp = 0.6, fig.cap = "Posterior probability of decoding accuracy being above chance level (2D GAM)."}
# defining a function to plot posterior prob of slope above 0
plot_post_timegen <- function (model, data, chance_level = 0.5, eps = 0.1) {

    # defining a sequence of x values to make predictions
    preds_conds <- crossing(
        train_time = unique(data$train_time)[c(TRUE, FALSE)],
        test_time = unique(data$test_time)[c(TRUE, FALSE)]
        )

    # retrieving posterior predictions
    post_preds <- conditional_effects(
        x = model,
        conditions = preds_conds,
        method = "posterior_epred",
        prob = 0.95,
        ndraws = 100
        )[[1]]

    # plotting cross-temporal generalization matrix
    post_preds_plot <- post_preds %>%
        summarise(auc = mean(estimate__), .by = c(train_time, test_time) ) %>%
        ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
        geom_tile() +
        geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
        geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
        geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
        scale_x_continuous(expand = c(0, 0) ) +
        scale_y_continuous(expand = c(0, 0) ) + 
        coord_fixed() +
        scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
        # theme_light() +
        labs(x = "Testing time (s)", y = "Training time (s)")

    # retrieving posterior samples
    posterior_samples <- posterior_epred(object = model, newdata = preds_conds)

    # computing the probability that y is above 0
    prob_y_above_0 <- preds_conds %>%
        mutate(post_prob = colMeans(posterior_samples) ) %>%
        mutate(m = colMeans(posterior_samples > (chance_level + eps) ) ) %>%
        mutate(prob_ratio = m / (1 - m) )

    # plotting it
    # prob_y_above_0 %>% filter(prob_ratio != Inf) %>% pull(prob_ratio) %>% range()
    prob_above_chance_plot <- prob_y_above_0 %>%
        mutate(prob_ratio = ifelse(prob_ratio == Inf, 2000, prob_ratio) ) %>%
        mutate(prob_ratio = ifelse(prob_ratio == 0, 0.0005, prob_ratio) ) %>%
        ggplot(aes(x = train_time, y = test_time, fill = prob_ratio) ) +
        geom_tile() +
        geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
        geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
        geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
        scale_x_continuous(expand = c(0, 0) ) +
        scale_y_continuous(expand = c(0, 0) ) + 
        coord_fixed() +
        scale_fill_scico(palette = "vik", midpoint = 0, name = "p/(1-p)", trans = "log10") +
        # scale_fill_scico(palette = "vik", midpoint = 1, name = "p/(1-p)") +
        # theme_light() +
        labs(x = "Testing time (s)", y = "Training time (s)")

    # combining the two plots
    post_preds_plot + prob_above_chance_plot

    # some tests
    # post_preds_plot +
    #     geom_contour(
    #         data = prob_y_above_0,
    #         aes(x = train_time, y = test_time, z = prob_ratio),
    #         breaks = 10,
    #         color = "red",
    #         linewidth = 1,
    #         inherit.aes = FALSE
    #         )

}

# using the function with the decoding GAM
plot_post_timegen(model = timegen_gam, data = timegen_data, chance_level = 0.5, eps = 0.01)
```

```{r gp-timegen, eval = FALSE, message = FALSE}
# fitting a GP with two temporal dimensions
timegen_gp <- brm(
    auc ~ gp(train_time, test_time, k = 20),
    data = timegen_data,
    family = Beta(),
    control = list(adapt_delta = 0.95, max_treedepth = 20),
    iter = 2000,
    chains = 4,
    cores = 4,
    file = "models/timegen_gp.rds"
    )
```

```{r gp-timegen-preds, eval = FALSE, echo = FALSE, fig.cap = "Posterior predictions of the GP fitted on decoding accuracy over time."}
# plotting the posterior predictions
plot(
    conditional_effects(x = timegen_gp, prob = 0.95),
    points = FALSE,
    method = "posterior_epred",
    surface = TRUE,
    theme = theme_light(),
    plot = TRUE
    )

# posterior predictive checks
pp_check(timegen_gp)

# checking timecourse of timegen_data's diagonal
timegen_data %>%
    filter(train_time == test_time) %>%
    summarise(auc = mean(auc), .by = train_time) %>%
    ggplot(aes(x = train_time, y = auc) ) +
    geom_hline(yintercept = 0.5, linetype = 2) +
    geom_line() +
    # theme_light() +
    labs(x = "Time (s)", y = "AUC")
```

## Application to actual M/EEG data

Assessing the reliability of the proposed approach (in comparison to other methods) using some sort of split-half reliability [@rosenblatt2018]?

## Comparing the identified onsets to other approaches

Comparing performance with uncorrected massive test, classical correction procedures (e.g., Bonferroni, FDR, FWER), cluster-based permutation test, cluster-depth, change point, [TFCE](https://jaromilfrossard.github.io/permuco4brain/articles/tfce.html)... We used the `permuco` package [@frossard2021]...

```{r permutations-1d, echo = FALSE, fig.cap = "T-values timecourse with onset identified using the permuco package (clustermass method)."}
# centering AUC relative to chance level (0.5)
decoding_data_perm <- decoding_data %>%
    mutate(auc_centered = auc - 0.5) %>%
    dplyr::select(-auc)

# formatting the AUC data to be used with permuco
dv <- decoding_data_perm %>%
    pivot_wider(names_from = time, values_from = auc_centered) %>%
    dplyr::select(-participant) %>%
    data.frame()

# formatting the intercept to be used with permuco
iv <- decoding_data %>%
    dplyr::select(participant) %>%
    distinct() %>%
    mutate(participant = as.factor(participant) ) %>%
    data.frame() %>%
    mutate(intercept = 1)

# performing cluster-based permutation test
cluster_results <- clusterlm(
    formula = dv ~ intercept,
    data = iv,
    # data = decoding_data_perm,
    # formula = auc_centered ~ participant,
    # number of permutations (increase for accuracy)
    # np = 2000,
    # type of transformation
    type = "signflip"
    # using a t-test
    # test = "t"
    # multiple comparison correction
    # multcomp = "clustermass"
    # multcomp = "clusterdepth_head",
    # multcomp = "tfce"
    )

# printing results
print(cluster_results)

# plotting significant clusters
plot(cluster_results)

# retrieving t and p-values
f_values <- cluster_results$multiple_comparison$intercept$clustermass$main[, 1]
p_values <- cluster_results$multiple_comparison$intercept$clustermass$main[, 2]
```

```{r changepoint, echo = FALSE, fig.cap = "T-values timecourse with true onset and onset identified using the changepoint package (binary segmentation method)."}
# computing the timecourse of t-values^2
t_values <- raw_df %>%
    summarise(y = mean(y), .by = c(x, condition, participant) ) %>%
    pivot_wider(names_from = condition, values_from = y) %>%
    # head()
    group_by(x) %>%
    summarise(tvalues = t.test(cond1, cond2)$statistic^2 ) %>%
    ungroup()

# using the changepoint package to identify onsets and offsets
res <- cpt.meanvar(data = t_values$tvalues, method = "BinSeg", Q = 2)

# plotting the results
true_onset <- 160
t_values %>%
    ggplot(aes(x = x, y = tvalues) ) +
    geom_line(linewidth = 1) +
    geom_vline(xintercept = true_onset) +
    geom_vline(xintercept = Xf[res@cpts[1]], linetype = "dotted") +
    labs(x = "Time (ms)", y = bquote(t^2) ) +
    # theme_light() +
    ggtitle(paste0("Change point onset: ", Xf[res@cpts[1]], "ms, true onset: ", true_onset, "ms") )
```

Now cluster-based permutations in 2D...

```{r permutations-2d, fig.cap = "Blah blah..."}
# See https://github.com/jaromilfrossard/permuco4brain
# devtools::install_github("jaromilfrossard/permuco4brain", build_vignettes = TRUE)
library(permuco4brain)
```

```{python, echo = FALSE, eval = FALSE}
# or using the MNE package in Python
import mne
```

# Results

...

# Discussion

...

## Summary of the proposed approach

...

## Increasing potential usage

TO-DO: Prepare a wrapper `R` package and show how to call it in `Python` and integrate it with `MNE-Python` [@gramfort2013] pipelines...

## Limitations and future directions

...

## Conclusions

...

\newpage

# References

::: {#refs}
:::

\newpage

# Appendix

````{=html}
<!--

# Generalised additive models

In brief, a GAM is a GLM that includes a sum of smooth functions of predictor(s)... The general formula for a GAM is:

$$
g(\mathbb{E}[Y]) = \alpha + \sum_{j=1}^p s_j\left(X_j\right)
$$

where:

- $Y$ is the response variable,
- $g()$ is the link function that relates the expected value of $Y(E(Y)$ to the linear predictor.
- $\alpha$ is the intercept term.
- $s_j\left(X_j\right)$ are smooth functions, such as splines, applied to the predictor variable $X_j$.
- $p$ is the number of predictor variables.

Explain GAMs with code and figures here...

```{r gam-tutorial, echo = TRUE, eval = FALSE}
# see https://gavinsimpson.github.io/gratia/articles/posterior-simulation.html
# see https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10b-GAMs.html
# and https://m-clark.github.io/generalized-additive-models/technical.html
# create a set of /k/ splines for vector /x/
get_basis <- function(x, k = 20){
  # fit a dummy GAM
  fit_gam <- mgcv::gam(
    formula = y ~ s(times, bs = "cr", k = 20),
    data    = tibble(y = 1, times = x))
  # extract model matrix
  model_matrix <- model.matrix(fit_gam)
  # wrangle to long tibble
  some_curves <- model_matrix |> 
    as_tibble() |> 
    mutate(x = x) |> 
    pivot_longer(cols = -x) |> 
    mutate(
      name = str_replace(name, "s\\(times\\).", "curve_"))
  return(some_curves)
}

make_a_wiggle <- function(some_curves, weights, data = NULL) {

  your_wiggly_line <- some_curves |> 
    mutate(
      weight = rep(weights, n_x),
      weighted_value = value * weight
      ) |> 
    group_by(x) |> 
    summarize(wiggly_line = sum(weighted_value)) |> 
    ungroup()

  your_wiggle_plot <- your_wiggly_line |> 
    ggplot(aes(x = x, y = wiggly_line)) + 
    geom_line(color = project_colors[1], size=1.5) + 
    ggtitle("your wiggly line") +
    xlab("x") + ylab("")

  your_wiggle_plot
}

weights <- c(0, 1, 10, rep(0, times = 17) )
make_a_wiggle(some_curves, weights)
```

# Gaussian processes

Explain GPs with code and figures here...

```{r gp-tutorial, echo = TRUE, eval = FALSE}
# see https://m-clark.github.io/models-by-example/gaussian-process.html
# see https://distill.pub/2019/visual-exploration-gaussian-processes/
# and http://www.infinitecuriosity.org/vizgp/
# defining the mean-centered function
gp_mu <- function (x) {

    map_dbl(x, function (x) x = 0)

}

# defining the covariance function (squared exponential kernel)
gp_K <- function (x, l = 1, sigma_f = 1) {

    sigma_f * exp( -(1/(2 * l^2)) * as.matrix(dist(x, upper = TRUE, diag = TRUE) ^ 2) )

}

# uniform prior
x_prior <- seq(-5, 5, 0.2)

# multivariate draws
y_prior = MASS::mvrnorm(
    n  = n_prior,
    mu = gp_mu(x_prior),
    Sigma = gp_K(x_prior, l = l, sigma_f = sigma_f)
    )

# plotting it
prior_data = data.frame(
    x  = x_prior,
    y  = t(y_prior),
    sd = apply(y_prior, 2, sd)
    ) %>% 
    pivot_longer(-c(x, sd), names_to = "variable")

prior_data %>%
    ggplot(aes(x = x, y = value) ) +
    geom_line(aes(group = variable), color = "#FF550080", alpha = 0.5) +
    labs(title = "Prior")
```

-->
````

# Mathematical formulation of the bivariate GAM

```{=html}
<!--

See https://bookdown.org/epeterson_2010/bios526_book/Module_5_3.html#bivariate-splines

-->
```

To model cross-temporal generalisation matrices of decoding performance (ROC AUC), we extended the initial (decoding) GAM to take into account the bivariate temporal distribution of AUC values, thus producing naturally smoothed estimates (timecourses) of AUC values and posterior probabilities. This model can be written as follows:

$$
\begin{aligned}
\text{AUC}_{i} &\sim \mathrm{Beta}(\mu_{i}, \phi)\\
g(\mu_{i}) &= f \left(\text{train}_{i}, \text{test}_{i} \right)\\
\end{aligned}
$$

where we assume that AUC values come from a $\mathrm{Beta}$ distribution with two parameters $\mu$ and $\phi$. We can think of $f \left(\text{train}_{i}, \text{test}_{i} \right)$ as a surface (a smooth function of two variables) that we can model using a 2-dimensional splines. Let $\mathbf{s}_{i} = \left(\text{train}_{i}, \text{test}_{i} \right)$ be some pair of training and testing samples, and let $\mathbf{k}_{m} = \left(\text{train}_{m}, \text{test}_{m} \right)$ denote the $m^{\text{th}}$ knot in the domain of $\text{train}_{i}$ and $\text{test}_{i}$. We can then express the smooth function as:

$$
f \left(\text{train}_{i}, \text{test}_{i} \right) = \alpha + \sum_{m=1}^M \beta_{m} b_{m} \left(\tilde{s}_{i}, \tilde{k}_{m} \right)
$$

Note that $b_{m}(,)$ is a basis function that maps $R \times R \rightarrow R$. A popular bivariate basis function uses *thin-plate splines*, which extend to $\mathbf{s}_{i} \in \mathbb{R}^{d}$ and $\partial l_{g}$ penalties. These splines are designed to interpolate and approximate smooth surfaces over two dimensions (hence the "bivariate" term). For $d=2$ dimensions and $l=2$ (smoothness penalty involving second order derivative):

$$
f \left(\tilde{s}_{i} \right) = \alpha + \beta_{1} x_{i} + \beta_{2} z_{i} +\sum_{m=1}^{M} \beta_{2+m} b_m\left(\tilde{s}_i, \tilde{k}_m\right)
$$

using the the radial basis function given by:

$$
b_m\left(\tilde{s}_i, \tilde{k}_m\right)=\left\|\tilde{s}_i-\tilde{k}_m\right\|^2 \log \left\|\tilde{s}_i-\tilde{k}_m\right\|
$$

where $\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|$ is the Euclidean distance between the covariate $\mathbf{s}_{i}$ and the knot location $\mathbf{k}_{m}$.

\newpage

# Threshold-free cluster enhancement

Cluster-based permutation approaches require defining a cluster-forming threshold (e.g., a t- or f-value) as the initial step of the algorithm. As different cluster-forming thresholds lead to clusters with different spatial or temporal extent, this threshold modulates the sensitivity of the subsequent permutation test. The threshold-free cluster enhancement method (TFCE) was introduced by @smith2009 to overcome this arbitrary threshold.

In brief, the TFCE method works as follows. Instead of picking an arbitrary cluster-forming threshold (e.g., $t=2$), we try all (or many) possible thresholds in a given range and check whether a given timestep/voxel belongs to a significant cluster under any of the set of thresholds... Then, instead of using cluster mass, we use a weighted average between the cluster extend ($e$, how broad is the cluster, that is, how many connected samples it contains) and the cluster height ($h$, how high is the cluster, that is, how large is the test statistic) according to the formula:

$$
\text{TFCE} = \int_{h} e(h)^{E} h^{H} dh
$$

Where... Then, we can compute p-values for each timestep.... we permute conditions, compute the TFCE score for the permuted set, and take the max(TFCE)... But see @sassenhagen2019...

### 2025-02-05 brms_meeg.tex
initial commit

\documentclass[
  doc,
  floatsintext,
  longtable,
  a4paper,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\geometry{inner=1in, outer=1in}
\fancyhfoffset[LE,RO]{0cm}

\usepackage[bidi=default]{babel}
\babelprovide[main,import]{english}

\babelfont{rm}[,RawFeature={fallback=mainfontfallback}]{CMU Serif}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

%\newlength\Oldarrayrulewidth
%\newlength\Oldtabcolsep

\usepackage{hyperref}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[nolongtablepatch]{lineno}
\linenumbers

\usepackage{fontspec}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

  \setmainfont[,RawFeature={fallback=mainfontfallback}]{CMU Serif}

\title{Modelling M/EEG data with Bayesian nonparametric multilevel
models}

\shorttitle{Bayesian M/EEG modelling}

\usepackage{etoolbox}

\authorsnames[{1},{2}]{Ladislas Nalborczyk,Paul Bürkner}

\authorsaffiliations{
{Aix Marseille Univ, CNRS, LPL},{TU Dortmund University, Department of
Statistics}}

\leftheader{Nalborczyk and Bürkner}

\abstract{Time-resolved electrophysoiological measurements such as those
offered by magneto- or electro-encephalography (M/EEG) provide a unique
window onto neural activity underlying cognitive process and how they
unfold over time. Typically, we are interested in testing whether such
measures differ across conditions and/or groups. The conventional
approach consists in conducting mass-univariate statistics followed by
some form of multiplicity correction (e.g., FDR, FWER) or cluster-based
inference. However, these cluster-based methods have an important
downside: they shift the focus of inference from the timepoint to the
cluster level, thus preventing any conclusion to be made about the onset
and offset of effects (e.g., differences across conditions). Here, we
introduce a novel \emph{model-based approch} for analysing
one-dimensional M/EEG timeseries such as ERPs or decoding timecourses
and their differences across conditions. This approach relies on
Bayesian nonparametric multilevel modelling (multilevel generalised
additive models or Gaussian processes), which outputs posterior
probabililities of the effect being above chance at every
timestep/voxel, while naturally taking into account the temporal
dependencies and between-subject variability present in such data.}

\keywords{EEG, MEG, generalised additive models, gaussian
processes, mixed-effects models, Bayesian statistics, brms}

\authornote{\par{\addORCIDlink{Ladislas
Nalborczyk}{0000-0002-7419-9855}}\par{\addORCIDlink{Paul
Bürkner}{0000-0001-5765-8995}}

\par{   The authors have no conflicts of interest to disclose.    }
\par{Correspondence concerning this article should be addressed
to Ladislas Nalborczyk, Aix Marseille Univ, CNRS, LPL, 5 avenue
Pasteur, 13100
Aix-en-Provence, France, email: ladislas.nalborczyk@cnrs.fr}
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother

\urlstyle{same}

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

%% Disable longtable counter
%% https://tex.stackexchange.com/a/248395/211326

\usepackage{etoolbox}

\makeatletter
\patchcmd{\LT@caption}
  {\bgroup}
  {\bgroup\global\LTpatch@captiontrue}
  {}{}
\patchcmd{\longtable}
  {\par}
  {\par\global\LTpatch@captionfalse}
  {}{}
\apptocmd{\endlongtable}
  {\ifLTpatch@caption\else\addtocounter{table}{-1}\fi}
  {}{}
\newif\ifLTpatch@caption
\makeatother

\begin{document}

\maketitle

\hypertarget{toc}{}
\tableofcontents
\newpage
\section[Introduction]{Modelling M/EEG data with Bayesian nonparametric
multilevel models}

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\resetlinenumber[1]

\section{Introduction (in progress)}\label{introduction-in-progress}

Here are some useful references to be discussed
(\citeproc{ref-combrisson_exceeding_2015}{Combrisson \& Jerbi, 2015};
\citeproc{ref-ehinger_unfold_2019}{Ehinger \& Dimigen, 2019};
\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021},
\citeproc{ref-frossard2022}{2022}; \citeproc{ref-gramfort2013}{Gramfort,
2013}; \citeproc{ref-hayasaka_validating_2003}{Hayasaka, 2003};
\citeproc{ref-luck_how_2017}{Luck \& Gaspelin, 2017};
\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
\citeproc{ref-pedersen_hierarchical_2019}{Pedersen et al., 2019};
\citeproc{ref-pernet2015}{Pernet et al., 2015};
\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al., 2023};
\citeproc{ref-rousselet_using_2025}{Rousselet, 2025})\ldots{} See also
(\citeproc{ref-maris2011}{Maris, 2011})\ldots{} and
(\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}) (history of
cluster-based approaches and using a data split?)\ldots{} Cluster
failure (\citeproc{ref-eklund2016}{Eklund et al., 2016})\ldots{}

In the following, we consider two approaches to modelling the non-linear
timecourse of M/EEG ERPs or decoding performance: i) generalised
additive models (GAMs) with thin-plate smoothing splines
(\citeproc{ref-wood2003}{Wood, 2003}; \citeproc{ref-wood2004}{Wood,
2004}) and ii) Gaussian processes (GPs) with a smooth covariance kernel
(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}) and low-rank
approximation (\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol
et al., 2023}).

\subsection{Previous modelling work}\label{previous-modelling-work}

Disentangling overlapping processes
(\citeproc{ref-skukies_brain_2024}{Skukies et al., 2024};
\citeproc{ref-skukies_modelling_2021}{Skukies \& Ehinger, 2021})\ldots{}
Using Bayes factors (\citeproc{ref-teichmann2022}{Teichmann,
2022})\ldots{}

\subsection{Generalised additive
models}\label{generalised-additive-models}

In generalised additive models (GAMs), the functional relationship
between predictors and response variable is decomposed into a sum of
low-dimensional non-parametric functions. A typical GAM has the
following form:

\[
\begin{aligned} 
y_{i} &\sim \mathrm{EF}\left(\mu_{i}, \phi\right)\\
g\left(\mu_i\right) &= \eta_{i} = A_{i} + \mathbf{X}_{i} \gamma + \sum_{j=1}^{J} f_{j}\left(x_{ij}\right)
\end{aligned}
\]

where \(g(\cdot)\) is the link function, \(A\) is an offset,
\(\mathbf{X}_{i}\) is the \(i\) th row of a parametric model matrix,
\(\gamma\) is a vector of parameters for the parametric terms, \(f_{j}\)
is a smooth function of covariate \(x_{j}\),
\(y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)\) denotes that the
observations \(y_{i}\) are distributed as some member of the exponential
family of distributions (e.g., Gaussian, Gamma, Beta, Poisson) with mean
\(\mu_{i}\) and scale parameter \(\phi\).

The smooth functions \(f_{j}\) are represented in the model via
penalised splines basis expansions of the covariates, that are a
weighted sum of basis functions:

\[
f_{j}\left(x_{i j}\right) = \sum_{k=1}^K \beta_{jk} b_{jk}\left(x_{ij}\right)
\]

where \(\beta_{jk}\) is the weight (coefficient) associated with the
\(k\) th basis function \(b_{j k}()\) evaluated at the covariate value
\(x_{ij}\) for the \(j\) th smooth function \(f_{j}\)\ldots{}

\subsection{Gaussian process
regression}\label{gaussian-process-regression}

A Gaussian process (GP) is a stochastic process that defines the
distribution over a collection of random variables indexed by a
continuous variable, that is \(\{f(t): t \in \mathcal{T}\}\) for some
index set \(\mathcal{T}\) (\citeproc{ref-rasmussen2005}{Rasmussen \&
Williams, 2005};
\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
2023}).

Note that nonparametric does not mean there aren't parameters, it means
that there are infinitely many parameters.

From
\href{https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp}{brms
documentation}: A GP is a stochastic process, which describes the
relation between one or more predictors
\(x=\left(x_1, \ldots, x_d\right)\) and a response \(f(x)\), where \(d\)
is the number of predictors. A GP is the generalization of the
multivariate normal distribution to an infinite number of dimensions.
Thus, it can be interpreted as a prior over functions. The values of
\(f()\) at any finite set of locations are jointly multivariate normal,
with a covariance matrix defined by the covariance kernel
\(k_p\left(x_i, x_j\right)\), where \(p\) is the vector of parameters of
the GP:

\[
\left(f\left(x_1\right), \ldots f\left(x_n\right) \sim \operatorname{MVN}\left(0,\left(k_p\left(x_i, x_j\right)\right)_{i, j=1}^n\right)\right.
\]

The smoothness and general behaviour of the function \(f\) depends only
on the choice of covariance kernel, see
\url{https://mc-stan.org/docs/functions-reference/matrix_operations.html\#gaussian-process-covariance-functions}.
It ensures that values that are close together in the input space will
be mapped to similar output values\ldots{} From this perspective, \(f\)
is a realisation of an infinite dimensional normal distribution:

\[
f \sim \mathrm{Normal}(0, C(\lambda))
\]

where \(C\) is a covariance kernel with hyperparameters \(\lambda\) that
defines the covariance between two function values \(f\left(t_1\right)\)
and \(f\left(t_2\right)\) for two time points \(t_1\) and \(t_2\)
(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). Similar to
the different choices of the basis function for splines, different
choices of the covariance kernel lead to different GPs. In this article,
we consider the squared-exponential (a.k.a. Gaussian or radial basis
function) kernel, which computes the squared distance between points and
converts it into a measure of similarity. It is defined as:

\[
C(\lambda):=C\left(t_1, t_2, \sigma, \gamma\right):=\sigma^2 \exp \left(-\frac{\left(t_1-t_2\right)^2}{2 \gamma^2}\right)
\]

with hyperparameters \(\lambda=(\sigma, \gamma)\), expressing the
overall scale of GP and the length-scale, respectively
(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). The
advantages of this kernel are that it is computationally efficient and
(infinitely) smooth making it a reasonable choice for the purposes of
the present article. Here again, \(\lambda\) hyperparameters are
estimated from the data, along with all other model parameters.

Low-rank approximate Gaussian processes are of main interest in machine
learning and statistics due to the high computational demands of exact
Gaussian process models
(\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
2023})\ldots{}

\subsection{Objectives}\label{objectives}

\ldots{}

\section{Methods}\label{methods}

\subsection{M/EEG data simulation}\label{meeg-data-simulation}

Following the approach used by Sassenhagen \& Draschkow
(\citeproc{ref-sassenhagen2019}{2019}) and Rousselet
(\citeproc{ref-rousselet_using_2025}{2025}), we simulated EEG data
stemming from two conditions, one with noise only, and the other with
noise + signal. As in previous studies, the noise was generated by
superimposing 50 sinusoids at different frequencies, following an
EEG-like spectrum (see details and code in
\citeproc{ref-yeung2004}{Yeung et al., 2004})\ldots{}

For simulating EEG data, see
\url{https://github.com/GRousselet/onsetsim} and
\url{https://github.com/unfoldtoolbox/UnfoldSim.jl}\ldots{}

Copy-pasted from Rousselet (\citeproc{ref-rousselet_using_2025}{2025}):
``The signal was a truncated Gaussian defining an objective onset at 160
ms, a maximum at 250 ms, and an offset at 342 ms. It was created by
getting the probability density function of a standard normal with 93
points from x= 1.5 to 1.5, rescaled in amplitude between 0 and 1 and
padded with 79 zeros to the left and to the right, to form a time series
of 251 points, from 0 to 500 ms, in steps of 2 ms (500 Hz sampling
rate).''

\begin{figure}[H]

\caption{Some ERPs in two conditions with 50 trials each, for a group of
20 participants.}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/eeg-group-1.pdf}

}

\end{figure}%

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# averaging across participants}
\NormalTok{group\_df }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ condition, }\AttributeTok{values\_from =}\NormalTok{ y, }\AttributeTok{values\_fn =}\NormalTok{ mean) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_diff =}\NormalTok{ cond2 }\SpecialCharTok{{-}}\NormalTok{ cond1) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{summarise}\NormalTok{(}
        \AttributeTok{y\_mean =} \FunctionTok{mean}\NormalTok{(y\_diff),}
        \AttributeTok{y\_sem =} \FunctionTok{sd}\NormalTok{(y\_diff) }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(}\FunctionTok{n}\NormalTok{() ),}
        \AttributeTok{.by =}\NormalTok{ x}
\NormalTok{        ) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lower =}\NormalTok{ y\_mean }\SpecialCharTok{{-}}\NormalTok{ y\_sem, }\AttributeTok{upper =}\NormalTok{ y\_mean }\SpecialCharTok{+}\NormalTok{ y\_sem)}

\CommentTok{\# plotting the data}
\NormalTok{group\_df }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y\_mean) ) }\SpecialCharTok{+}
    \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =}\NormalTok{ true\_onset, }\AttributeTok{linetype =} \DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{geom\_ribbon}\NormalTok{(}
        \FunctionTok{aes}\NormalTok{(}\AttributeTok{ymin =}\NormalTok{ lower, }\AttributeTok{ymax =}\NormalTok{ upper, }\AttributeTok{colour =} \ConstantTok{NULL}\NormalTok{),}
        \AttributeTok{alpha =} \FloatTok{0.25}\NormalTok{, }\AttributeTok{show.legend =} \ConstantTok{FALSE}
\NormalTok{        ) }\SpecialCharTok{+}
    \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{show.legend =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (ms)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG difference (a.u.)"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{Group-level average difference between conditions (mean +/-
standard error of the mean). The `true' (population value of the) onset
is indicated by the vertical dashed line.}

{\centering \includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/erp-diff-1.pdf}

}

\end{figure}%

\subsection{Model fitting}\label{model-fitting}

Models were fitted using the \texttt{brms} package
(\citeproc{ref-burkner_brms_2017}{Bürkner, 2017},
\citeproc{ref-burkner_advanced_2018}{2018};
\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019})\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# defining a contrast for condition}
\FunctionTok{contrasts}\NormalTok{(group\_df}\SpecialCharTok{$}\NormalTok{condition) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{)}

\CommentTok{\# fitting the GAM}
\NormalTok{gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# cubic regression splines with k{-}1 basis functions}
    \CommentTok{\# then we should try s(x, participant, bs = "fs")}
    \CommentTok{\# that is, random factor smooth interactions...}
\NormalTok{    y }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(x, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
    \AttributeTok{data =}\NormalTok{ group\_df,}
    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/gam.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

Now we fit the GP\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{gp\_model }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# k refers to the number of basis functions for}
    \CommentTok{\# computing Hilbert{-}space approximate GPs}
\NormalTok{    y }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gp}\NormalTok{(x, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
    \CommentTok{\# if k = NA (default), exact GPs are computed}
    \CommentTok{\# y \textasciitilde{} gp(x, by = condition),}
    \AttributeTok{data =}\NormalTok{ group\_df,}
    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/gp.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

And we plot the posterior predictions\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# plotting the posterior predictions}
\FunctionTok{plot\_post\_preds}\NormalTok{(}\AttributeTok{model =}\NormalTok{ gam, }\AttributeTok{data =}\NormalTok{ group\_df) }\SpecialCharTok{+}
    \FunctionTok{plot\_post\_preds}\NormalTok{(}\AttributeTok{model =}\NormalTok{ gp\_model, }\AttributeTok{data =}\NormalTok{ group\_df)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{Posterior predictions of the GAM (left) and GP (right) models.}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-gp-preds-1.pdf}

}

\end{figure}%

\subsection{Posterior probability of difference above
0}\label{posterior-probability-of-difference-above-0}

We can retrieve the posterior probability of the slope for
\texttt{condition} at each timestep.

\begin{figure}[H]

\caption{Slope for the difference between conditions according to the
GAM (left) or the GP (right).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/unnamed-chunk-1-1.pdf}

}

\end{figure}%

We can also compute the posterior probability of the slope for
\texttt{condition} being above 0 (Figure~\ref{fig-post-prob-test}).

\begin{figure}[!htb]

\caption{\label{fig-post-prob-test}Posterior probability of the EEG
signal being above 0+eps according to the GAM (left) and the GP
(right).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-test-1.pdf}

}

\end{figure}%

We can also express this as the ratio of posterior probabilities (i.e.,
\(p/(1-p)\)) and visualise the timecourse of this ratio superimposed
with the conventional BF thresholds (Figure~\ref{fig-post-prob-ratio}).

\begin{figure}[!htb]

\caption{\label{fig-post-prob-ratio}Ratio of posterior probability
according to the GAM (left) and the GP (right). Timesteps above
threshold (\textgreater10) are highlighted in green.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.pdf}

}

\end{figure}%

\newpage

\subsection{Error properties of the proposed
approach}\label{error-properties-of-the-proposed-approach}

Below we are plotting the error function (where the error is computed as
\(|\hat{\theta}-\theta|\)) for both the onset and offset values of the
ERP difference, according to various \texttt{eps} and \texttt{threshold}
values. Remember that the signal is generated from a truncated Gaussian
defining an objective onset at 160 ms, a maximum at 250 ms, and an
offset at 342 ms. Figure~\ref{fig-onset-error} shows that the GP model
can \emph{exactly} recover the true onset and offset values, given some
reasonable choice of \texttt{eps} and \texttt{threshold} values.

\begin{verbatim}
   eps threshold estimated_onset estimated_offset error_onset error_offset
1 0.06        15             160              342           0            0
2 0.06        16             160              342           0            0
3 0.06        17             160              342           0            0
4 0.06        18             160              342           0            0
5 0.06        19             160              342           0            0
6 0.06        20             160              342           0            0
\end{verbatim}

\begin{verbatim}
   eps threshold estimated_onset estimated_offset error_onset error_offset
1 0.06         3             159              342           1            0
2 0.06         4             159              342           1            0
3 0.06         5             159              342           1            0
4 0.06         6             159              342           1            0
5 0.06         7             159              342           1            0
6 0.06         8             159              342           1            0
\end{verbatim}

\begin{figure}[!htb]

\caption{\label{fig-onset-error}Error function of onset (left) and
offset (right) estimation according to various eps and threshold values
(according to the GP model). Minimum error values are indicated by red
crosses.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-error-1.pdf}

}

\end{figure}%

\newpage

\subsection{Using summary statistics of
ERPs}\label{using-summary-statistics-of-erps}

Next we fit a hierarchical GAM using summary statistics of ERPs (mean
and SD) at the participant level (similar to what is done in
meta-analysis).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# averaging across participants}
\NormalTok{summary\_df }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{summarise}\NormalTok{(}
        \AttributeTok{eeg =} \FunctionTok{mean}\NormalTok{(y),}
        \AttributeTok{eeg\_sd =} \FunctionTok{sd}\NormalTok{(y),}
        \AttributeTok{.by =} \FunctionTok{c}\NormalTok{(participant, condition, x)}
\NormalTok{        )}

\CommentTok{\# defining a contrast for condition}
\FunctionTok{contrasts}\NormalTok{(summary\_df}\SpecialCharTok{$}\NormalTok{condition) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{)}

\CommentTok{\# fitting the GAM}
\NormalTok{meta\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# using by{-}participant SD of ERPs across trials}
\NormalTok{    eeg }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(x, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
\NormalTok{        (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ participant),}
    \AttributeTok{data =}\NormalTok{ summary\_df,}
    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/meta\_gam.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# plotting the posterior predictions}
\FunctionTok{plot}\NormalTok{(}
    \FunctionTok{conditional\_effects}\NormalTok{(}\AttributeTok{x =}\NormalTok{ meta\_gam, }\AttributeTok{effect =} \StringTok{"x:condition"}\NormalTok{),}
    \AttributeTok{points =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{theme =} \FunctionTok{theme\_light}\NormalTok{(), }\AttributeTok{plot =} \ConstantTok{FALSE}
\NormalTok{    )[[}\DecValTok{1}\NormalTok{]] }\SpecialCharTok{+}
    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (ms)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG signal (a.u.)"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{Hierarchical GAM posterior predictions.}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/meta-gam-preds-1.pdf}

}

\end{figure}%

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting the GP}
\NormalTok{meta\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# using by{-}participant SD of ERPs across trials}
\NormalTok{    eeg }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{gp}\NormalTok{(x, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
\NormalTok{        (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ participant),}
    \AttributeTok{data =}\NormalTok{ summary\_df,}
    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{20}\NormalTok{),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/meta\_gp.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# plotting the posterior predictions}
\FunctionTok{plot}\NormalTok{(}
    \FunctionTok{conditional\_effects}\NormalTok{(}\AttributeTok{x =}\NormalTok{ meta\_gp, }\AttributeTok{effect =} \StringTok{"x:condition"}\NormalTok{),}
    \AttributeTok{points =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{theme =} \FunctionTok{theme\_light}\NormalTok{(), }\AttributeTok{plot =} \ConstantTok{FALSE}
\NormalTok{    )[[}\DecValTok{1}\NormalTok{]] }\SpecialCharTok{+}
    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (ms)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG signal (a.u.)"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\subsection{Application to 1D decoding results (accuracy over
time)}\label{application-to-1d-decoding-results-accuracy-over-time}

Assume we have M/EEG data and we conducted time-resolved multivariate
pattern analysis (MVPA), also known as decoding. As a result, we have a
timecourse of decoding accuracies (e.g., ROC AUC), bounded between 0 and
1, per participant (Figure~\ref{fig-sim-decoding}).

\begin{figure}[!htb]

\caption{\label{fig-sim-decoding}Exemplary (simulated) group-level
average timecourse of binary decoding accuracy (ROC AUC).}

\centering{

\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-decoding-1.pdf}

}

\end{figure}%

Now, we want to \emph{test} whether the group-level average decoding
accuracy is above chance (i.e., 0.5) at each timestep. We use a similar
GAM/GP as previously, but we replace the \(\mathrm{Normal}\) likelihood
function by a \(\mathrm{Beta}\) one.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting the GAM}
\NormalTok{decoding\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
    \AttributeTok{data =}\NormalTok{ decoding\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/decoding\_gam.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting the GP}
\NormalTok{decoding\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gp}\NormalTok{(time, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
    \AttributeTok{data =}\NormalTok{ decoding\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.9}\NormalTok{),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/decoding\_gp.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{Posterior predictions of the GAM (left) and GP (right) fitted
on decoding accuracy over time.}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/decoding-preds-1.pdf}

}

\end{figure}%

Next, we plot the posterior probability of decoding accuracy being above
chance level (plus some epsilon) (Figure~\ref{fig-decoding-post}).

\begin{figure}[!htb]

\caption{\label{fig-decoding-post}Posterior probability of decoding
accuracy being above chance level according to the GAM (left) or the GP
(right).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-post-1.pdf}

}

\end{figure}%

\begin{verbatim}
  cluster_onset cluster_offset
1     0.1263598      0.6937238
  cluster_onset cluster_offset
1     0.1615063      0.6786611
\end{verbatim}

\begin{figure}[!htb]

\caption{\label{fig-decoding-ratio}Ratio of posterior probabilities of
decoding accuracy being above chance level according to the GAM (left)
or the GP (right).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-ratio-1.pdf}

}

\end{figure}%

\newpage

\subsection{Application to 2D decoding results (cross-temporal
generalisation)}\label{application-to-2d-decoding-results-cross-temporal-generalisation}

Assume we have M/EEG data and we have conducted cross-temporal
generalisation analyses (\citeproc{ref-king2014}{King \& Dehaene,
2014}). As a result, we have a 2D matrix where each element contains the
decoding accuracy (e.g., ROC AUC) of a classifier trained at timestep
\(\text{training}_{i}\) and tested at timestep \(\text{testing}_{j}\)
(Figure~\ref{fig-sim-timegen}).

\begin{figure}[!htb]

\caption{\label{fig-sim-timegen}Exemplary (simulated) group-level
average cross-temporal generalisation matrix of decoding performance
(ROC AUC).}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-timegen-1.pdf}

}

\end{figure}%

Now, we want to test whether and when decoding performance is above
chance level (0.5 for a binary decoding task). This one is
computationally very intense\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting a GAM with a 2D smooth function}
\NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
    \CommentTok{\# 2D thin{-}plate spline (tp) to model smooth interactions between training and testing time}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{s}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
    \AttributeTok{data =}\NormalTok{ timegen\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/timegen\_gam.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

\caption{Posterior probability of decoding accuracy being above chance
level (2D GAM).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-timegen-post-preds-1.pdf}

}

\end{figure}%

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# fitting a GP with two temporal dimensions}
\NormalTok{timegen\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gp}\NormalTok{(train\_time, test\_time, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
    \AttributeTok{data =}\NormalTok{ timegen\_data,}
    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
    \AttributeTok{file =} \StringTok{"models/timegen\_gp.rds"}
\NormalTok{    )}
\end{Highlighting}
\end{Shaded}

\subsection{Application to actual M/EEG
data}\label{application-to-actual-meeg-data}

Assessing the reliability of the proposed approach (in comparison to
other methods) using some sort of split-half reliability
(\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018})?

\subsection{Comparing the identified onsets to other
approaches}\label{comparing-the-identified-onsets-to-other-approaches}

Comparing performance with uncorrected massive test, classical
correction procedures (e.g., Bonferroni, FDR, FWER), cluster-based
permutation test, cluster-depth, change point,
\href{https://jaromilfrossard.github.io/permuco4brain/articles/tfce.html}{TFCE}\ldots{}
We used the \texttt{permuco} package
(\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021})\ldots{}

\begin{verbatim}
Effect: intercept.
Alternative Hypothesis: two.sided.
Statistic: fisher(1, 19).
Resampling Method: freedman_lane.
Type of Resampling: signflip.
Number of Dependant Variables: 240.
Number of Resamples: 5000.
Multiple Comparisons Procedure: clustermass.
Threshold: 4.38075.
Mass Function: the sum.
Table of clusters.

  start end cluster mass P(>mass)
1    25  25     4.883149   1.0000
2   127 127     6.183416   0.4212
3   218 218     5.080433   1.0000
\end{verbatim}

\begin{figure}[H]

\caption{T-values timecourse with onset identified using the permuco
package (clustermass method).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/permutations-1d-1.pdf}

}

\end{figure}%

\begin{figure}[H]

\caption{T-values timecourse with true onset and onset identified using
the changepoint package (binary segmentation method).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/changepoint-1.pdf}

}

\end{figure}%

Now cluster-based permutations in 2D\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# See https://github.com/jaromilfrossard/permuco4brain}
\CommentTok{\# devtools::install\_github("jaromilfrossard/permuco4brain", build\_vignettes = TRUE)}
\FunctionTok{library}\NormalTok{(permuco4brain)}
\end{Highlighting}
\end{Shaded}

\section{Results}\label{results}

\ldots{}

\section{Discussion}\label{discussion}

\ldots{}

\subsection{Summary of the proposed
approach}\label{summary-of-the-proposed-approach}

\ldots{}

\subsection{Increasing potential
usage}\label{increasing-potential-usage}

TO-DO: Prepare a wrapper \texttt{R} package and show how to call it in
\texttt{Python} and integrate it with \texttt{MNE-Python}
(\citeproc{ref-gramfort2013}{Gramfort, 2013}) pipelines\ldots{}

\subsection{Limitations and future
directions}\label{limitations-and-future-directions}

\ldots{}

\subsection{Conclusions}\label{conclusions}

\ldots{}

\newpage

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-burkner_brms_2017}
Bürkner, P.-C. (2017). \textbf{Brms} : {An} {{\emph{R}}} {Package} for
{Bayesian Multilevel Models Using} {{\emph{Stan}}}. \emph{Journal of
Statistical Software}, \emph{80}(1).
\url{https://doi.org/10.18637/jss.v080.i01}

\bibitem[\citeproctext]{ref-burkner_advanced_2018}
Bürkner, P.-C. (2018). Advanced {Bayesian Multilevel Modeling} with the
{R Package} brms. \emph{The R Journal}, \emph{10}(1), 395.
\url{https://doi.org/10.32614/RJ-2018-017}

\bibitem[\citeproctext]{ref-combrisson_exceeding_2015}
Combrisson, E., \& Jerbi, K. (2015). Exceeding chance level by chance:
{The} caveat of theoretical chance levels in brain signal classification
and statistical assessment of decoding accuracy. \emph{Journal of
Neuroscience Methods}, \emph{250}, 126--136.
\url{https://doi.org/10.1016/j.jneumeth.2015.01.010}

\bibitem[\citeproctext]{ref-ehinger_unfold_2019}
Ehinger, B. V., \& Dimigen, O. (2019). Unfold: An integrated toolbox for
overlap correction, non-linear modeling, and regression-based {EEG}
analysis. \emph{PeerJ}, \emph{7}, e7838.
\url{https://doi.org/10.7717/peerj.7838}

\bibitem[\citeproctext]{ref-eklund2016}
Eklund, A., Nichols, T. E., \& Knutsson, H. (2016). Cluster failure: Why
fMRI inferences for spatial extent have inflated false-positive rates.
\emph{Proceedings of the National Academy of Sciences}, \emph{113}(28),
7900--7905. \url{https://doi.org/10.1073/pnas.1602413113}

\bibitem[\citeproctext]{ref-frossard2021}
Frossard, J., \& Renaud, O. (2021). Permutation Tests for Regression,
ANOVA, and Comparison of Signals: The {\textbf{permuco}} Package.
\emph{Journal of Statistical Software}, \emph{99}(15).
\url{https://doi.org/10.18637/jss.v099.i15}

\bibitem[\citeproctext]{ref-frossard2022}
Frossard, J., \& Renaud, O. (2022). The cluster depth tests: Toward
point-wise strong control of the family-wise error rate in massively
univariate tests with application to M/EEG. \emph{NeuroImage},
\emph{247}, 118824.
\url{https://doi.org/10.1016/j.neuroimage.2021.118824}

\bibitem[\citeproctext]{ref-gramfort2013}
Gramfort, A. (2013). MEG and EEG data analysis with MNE-python.
\emph{Frontiers in Neuroscience}, \emph{7}.
\url{https://doi.org/10.3389/fnins.2013.00267}

\bibitem[\citeproctext]{ref-hayasaka_validating_2003}
Hayasaka, S. (2003). Validating cluster size inference: Random field and
permutation methods. \emph{NeuroImage}, \emph{20}(4), 2343--2356.
\url{https://doi.org/10.1016/j.neuroimage.2003.08.003}

\bibitem[\citeproctext]{ref-king2014}
King, J.-R., \& Dehaene, S. (2014). Characterizing the dynamics of
mental representations: the temporal generalization method. \emph{Trends
in Cognitive Sciences}, \emph{18}(4), 203--210.
\url{https://doi.org/10.1016/j.tics.2014.01.002}

\bibitem[\citeproctext]{ref-luck_how_2017}
Luck, S. J., \& Gaspelin, N. (2017). How to get statistically
significant effects in any {ERP} experiment (and why you shouldn't).
\emph{Psychophysiology}, \emph{54}(1), 146--157.
\url{https://doi.org/10.1111/psyp.12639}

\bibitem[\citeproctext]{ref-maris2011}
Maris, E. (2011). Statistical testing in electrophysiological studies.
\emph{Psychophysiology}, \emph{49}(4), 549--565.
\url{https://doi.org/10.1111/j.1469-8986.2011.01320.x}

\bibitem[\citeproctext]{ref-maris2007}
Maris, E., \& Oostenveld, R. (2007). Nonparametric statistical testing
of EEG- and MEG-data. \emph{Journal of Neuroscience Methods},
\emph{164}(1), 177--190.
\url{https://doi.org/10.1016/j.jneumeth.2007.03.024}

\bibitem[\citeproctext]{ref-nalborczyk2019}
Nalborczyk, L., Batailler, C., Lœvenbruck, H., Vilain, A., \& Bürkner,
P.-C. (2019). An Introduction to Bayesian Multilevel Models Using brms:
A Case Study of Gender Effects on Vowel Variability in Standard
Indonesian. \emph{Journal of Speech, Language, and Hearing Research},
\emph{62}(5), 1225--1242.
\url{https://doi.org/10.1044/2018_jslhr-s-18-0006}

\bibitem[\citeproctext]{ref-pedersen_hierarchical_2019}
Pedersen, E. J., Miller, D. L., Simpson, G. L., \& Ross, N. (2019).
Hierarchical generalized additive models in ecology: An introduction
with mgcv. \emph{PeerJ}, \emph{7}, e6876.
\url{https://doi.org/10.7717/peerj.6876}

\bibitem[\citeproctext]{ref-pernet2015}
Pernet, C. R., Latinus, M., Nichols, T. E., \& Rousselet, G. A. (2015).
Cluster-based computational methods for mass univariate analyses of
event-related brain potentials/fields: A simulation study. \emph{Journal
of Neuroscience Methods}, \emph{250}, 85--93.
\url{https://doi.org/10.1016/j.jneumeth.2014.08.003}

\bibitem[\citeproctext]{ref-rasmussen2005}
Rasmussen, C. E., \& Williams, C. K. I. (2005). \emph{Gaussian Processes
for Machine Learning}.
\url{https://doi.org/10.7551/mitpress/3206.001.0001}

\bibitem[\citeproctext]{ref-riutort-mayol_practical_2023}
Riutort-Mayol, G., Bürkner, P.-C., Andersen, M. R., Solin, A., \&
Vehtari, A. (2023). Practical {Hilbert} space approximate {Bayesian
Gaussian} processes for probabilistic programming. \emph{Statistics and
Computing}, \emph{33}(1), 17.
\url{https://doi.org/10.1007/s11222-022-10167-2}

\bibitem[\citeproctext]{ref-rosenblatt2018}
Rosenblatt, J. D., Finos, L., Weeda, W. D., Solari, A., \& Goeman, J. J.
(2018). All-Resolutions Inference for brain imaging. \emph{NeuroImage},
\emph{181}, 786--796.
\url{https://doi.org/10.1016/j.neuroimage.2018.07.060}

\bibitem[\citeproctext]{ref-rousselet_using_2025}
Rousselet, G. A. (2025). Using cluster-based permutation tests to
estimate {MEG}/{EEG} onsets: {How} bad is it? \emph{European Journal of
Neuroscience}, \emph{61}(1), e16618.
\url{https://doi.org/10.1111/ejn.16618}

\bibitem[\citeproctext]{ref-sassenhagen2019}
Sassenhagen, J., \& Draschkow, D. (2019). Cluster{-}based permutation
tests of MEG/EEG data do not establish significance of effect latency or
location. \emph{Psychophysiology}, \emph{56}(6).
\url{https://doi.org/10.1111/psyp.13335}

\bibitem[\citeproctext]{ref-skukies_modelling_2021}
Skukies, R., \& Ehinger, B. (2021). Modelling event duration and overlap
during {EEG} analysis. \emph{Journal of Vision}, \emph{21}(9), 2037.
\url{https://doi.org/10.1167/jov.21.9.2037}

\bibitem[\citeproctext]{ref-skukies_brain_2024}
Skukies, R., Schepers, J., \& Ehinger, B. (2024, December 9).
\emph{Brain responses vary in duration - modeling strategies and
challenges}. \url{https://doi.org/10.1101/2024.12.05.626938}

\bibitem[\citeproctext]{ref-smith2009}
Smith, S., \& Nichols, T. (2009). Threshold-free cluster enhancement:
Addressing problems of smoothing, threshold dependence and localisation
in cluster inference. \emph{NeuroImage}, \emph{44}(1), 83--98.
\url{https://doi.org/10.1016/j.neuroimage.2008.03.061}

\bibitem[\citeproctext]{ref-teichmann2022}
Teichmann, L. (2022). An empirically driven guide on using bayes factors
for m/EEG decoding. \emph{Aperture Neuro}, \emph{2}, 1--10.
\url{https://doi.org/10.52294/apertureneuro.2022.2.maoc6465}

\bibitem[\citeproctext]{ref-wood2003}
Wood, S. N. (2003). Thin Plate Regression Splines. \emph{Journal of the
Royal Statistical Society Series B: Statistical Methodology},
\emph{65}(1), 95--114. \url{https://doi.org/10.1111/1467-9868.00374}

\bibitem[\citeproctext]{ref-wood2004}
Wood, S. N. (2004). Stable and Efficient Multiple Smoothing Parameter
Estimation for Generalized Additive Models. \emph{Journal of the
American Statistical Association}, \emph{99}(467), 673--686.
\url{https://doi.org/10.1198/016214504000000980}

\bibitem[\citeproctext]{ref-yeung2004}
Yeung, N., Bogacz, R., Holroyd, C. B., \& Cohen, J. D. (2004). Detection
of synchronized oscillations in the electroencephalogram: An evaluation
of methods. \emph{Psychophysiology}, \emph{41}(6), 822--832.
\url{https://doi.org/10.1111/j.1469-8986.2004.00239.x}

\end{CSLReferences}

\newpage

\appendix

\section{Mathematical formulation of the bivariate
GAM}\label{mathematical-formulation-of-the-bivariate-gam}

To model cross-temporal generalisation matrices of decoding performance
(ROC AUC), we extended the initial (decoding) GAM to take into account
the bivariate temporal distribution of AUC values, thus producing
naturally smoothed estimates (timecourses) of AUC values and posterior
probabilities. This model can be written as follows:

\[
\begin{aligned}
\text{AUC}_{i} &\sim \mathrm{Beta}(\mu_{i}, \phi)\\
g(\mu_{i}) &= f \left(\text{train}_{i}, \text{test}_{i} \right)\\
\end{aligned}
\]

where we assume that AUC values come from a \(\mathrm{Beta}\)
distribution with two parameters \(\mu\) and \(\phi\). We can think of
\(f \left(\text{train}_{i}, \text{test}_{i} \right)\) as a surface (a
smooth function of two variables) that we can model using a
2-dimensional splines. Let
\(\mathbf{s}_{i} = \left(\text{train}_{i}, \text{test}_{i} \right)\) be
some pair of training and testing samples, and let
\(\mathbf{k}_{m} = \left(\text{train}_{m}, \text{test}_{m} \right)\)
denote the \(m^{\text{th}}\) knot in the domain of \(\text{train}_{i}\)
and \(\text{test}_{i}\). We can then express the smooth function as:

\[
f \left(\text{train}_{i}, \text{test}_{i} \right) = \alpha + \sum_{m=1}^M \beta_{m} b_{m} \left(\tilde{s}_{i}, \tilde{k}_{m} \right)
\]

Note that \(b_{m}(,)\) is a basis function that maps
\(R \times R \rightarrow R\). A popular bivariate basis function uses
\emph{thin-plate splines}, which extend to
\(\mathbf{s}_{i} \in \mathbb{R}^{d}\) and \(\partial l_{g}\) penalties.
These splines are designed to interpolate and approximate smooth
surfaces over two dimensions (hence the ``bivariate'' term). For \(d=2\)
dimensions and \(l=2\) (smoothness penalty involving second order
derivative):

\[
f \left(\tilde{s}_{i} \right) = \alpha + \beta_{1} x_{i} + \beta_{2} z_{i} +\sum_{m=1}^{M} \beta_{2+m} b_m\left(\tilde{s}_i, \tilde{k}_m\right)
\]

using the the radial basis function given by:

\[
b_m\left(\tilde{s}_i, \tilde{k}_m\right)=\left\|\tilde{s}_i-\tilde{k}_m\right\|^2 \log \left\|\tilde{s}_i-\tilde{k}_m\right\|
\]

where \(\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|\) is the Euclidean
distance between the covariate \(\mathbf{s}_{i}\) and the knot location
\(\mathbf{k}_{m}\).

\newpage

\section{Threshold-free cluster
enhancement}\label{threshold-free-cluster-enhancement}

Cluster-based permutation approaches require defining a cluster-forming
threshold (e.g., a t- or f-value) as the initial step of the algorithm.
As different cluster-forming thresholds lead to clusters with different
spatial or temporal extent, this threshold modulates the sensitivity of
the subsequent permutation test. The threshold-free cluster enhancement
method (TFCE) was introduced by Smith \& Nichols
(\citeproc{ref-smith2009}{2009}) to overcome this arbitrary threshold.

In brief, the TFCE method works as follows. Instead of picking an
arbitrary cluster-forming threshold (e.g., \(t=2\)), we try all (or
many) possible thresholds in a given range and check whether a given
timestep/voxel belongs to a significant cluster under any of the set of
thresholds\ldots{} Then, instead of using cluster mass, we use a
weighted average between the cluster extend (\(e\), how broad is the
cluster, that is, how many connected samples it contains) and the
cluster height (\(h\), how high is the cluster, that is, how large is
the test statistic) according to the formula:

\[
\text{TFCE} = \int_{h} e(h)^{E} h^{H} dh
\]

Where\ldots{} Then, we can compute p-values for each timestep\ldots. we
permute conditions, compute the TFCE score for the permuted set, and
take the max(TFCE)\ldots{} But see Sassenhagen \& Draschkow
(\citeproc{ref-sassenhagen2019}{2019})\ldots{}

\end{document}

### 2025-02-05 code/eeg_noise.R

@@ -1,0 +1,40 @@
+##############################################################################
+# Generate noise with the power spectrum of human EEG
+# R version + control variance: GAR, University of Glasgow, Oct 2023
+# From https://github.com/GRousselet/onsetsim/blob/main/code/eeg_noise.R
+# Implemented by: Rafal Bogacz and Nick Yeung, Princeton Univesity, December 2002
+# Adapted from Matlab code:
+# https://data.mrc.ox.ac.uk/data-set/simulated-eeg-data-generator
+############################################################################
+
+##############################################################################
+# Input:
+#   frames - number of signal frames per each trial
+#   epochs - number of simulated trials
+#   srate - sampling rate of simulated signal
+# Output:
+#   signal - simulated EEG signal; vector: 1 by frames*epochs containing concatenated trials
+##############################################################################
+
+eeg_noise <- function (frames = 51, srate = 100, outvar = 1, meanpower) {
+    
+    sumsig <- 50 # number of sinusoids from which each simulated signal is composed of
+    out <- vector(mode = "numeric", length = frames)
+    freq <- 0
+    
+    for (i in 1:sumsig) {
+        
+        freq <- freq + (4 * runif(1) )
+        freqamp <- meanpower[min(ceiling(freq), 125)] / meanpower[1]
+        phase <- runif(1) * 2 * pi
+        out <- out + sin((1:frames)/srate * 2 * pi * freq + phase) * freqamp
+        
+    }
+    
+    out <- out - mean(out)
+    out <- out / sd(out)
+    out <- out * sqrt(outvar)
+    
+    return(out)
+    
+}

### 2025-02-05 code/erp_template.R

@@ -1,0 +1,13 @@
+##############################################################################
+# from https://github.com/GRousselet/onsetsim/blob/main/code/erp_template.R
+# true onset = 160 ms, F = 81, max at F = 126
+##############################################################################
+
+true_onset <- 160
+Xf <- seq(0, 500, 2)
+Nf <- length(Xf)
+temp1 <- vector(mode = "numeric", length = Nf)
+erp <- dnorm(seq(-1.5, 1.5, length.out = 93), 0, 1)
+erp <- erp - min(erp)
+erp <- erp / max(erp)
+temp2 <- c(rep(0, 79), erp, rep(0, 79) )

### 2025-02-05 code/functions.R

@@ -1,0 +1,202 @@
+##############################################################################
+# # Functions from https://github.com/GRousselet/onsetsim/blob/main/code/functions.R
+# # Generate permutation distributions of t values for independent groups
+# ## cond1 and cond2 are Nt trials x Nf time frames matrices
+# permtdist <- function(cond1, cond2, Nt, Nf, nboot = 2000){
+#   
+#   st2 <- Nt+1 # index to start condition 2 
+#   en2 <- Nt+Nt # index to end condition 2
+#   all <- rbind(cond1, cond2)
+#   
+#   perm.tvals <- matrix(data = 0, nrow = nboot, ncol = Nf)
+#   # perm.pvals <- matrix(data = 0, nrow = nboot, ncol = Nf)
+#   
+#   for(B in 1:nboot){
+#     
+#     permdata <- all[sample(1:en2, size = en2, replace = FALSE),]
+#     permcond1 <- permdata[1:Nt,]
+#     permcond2 <- permdata[st2:en2,]
+#     
+#     perm.tvals[B,] <- Rfast::ttests(permcond1, permcond2, paired=FALSE)[,1]    
+#     # for(F in 1:Nf){ # for each time frame:
+#     #   perm.tvals[B,F] <- t.test(permcond1[,F], permcond2[,F])$statistic
+#     # }
+#   }
+#   # list(tvals = perm.tvals, pvals = perm.pvals)
+#   perm.tvals
+# }
+# 
+# # Generate permutation distributions of t values for one group
+# ## diff is a matrix of pairwise differences Np participants x Nf time frames matrices
+# permtdist.onesamp <- function(diff, Np, Nf, nboot = 2000){
+#   
+#   perm.tvals <- matrix(data = 0, nrow = nboot, ncol = Nf)
+#   permdiff <- diff
+#   
+#   for(B in 1:nboot){
+#     
+#     for(P in 1:Np){
+#       newsign <- sample(c(-1,1),1) # flip a coin
+#       permdiff[P,] <- diff[P,] * newsign 
+#     }
+#     
+#     for(F in 1:Nf){ # for each time frame:
+#       perm.tvals[B,F] <- t.test(permdiff[,F])$statistic
+#     }
+#     
+#   }
+#   perm.tvals
+# }
+# 
+# # Cluster correction
+# 
[…]

### 2025-02-06 brms_meeg.qmd
now modelling actual decoding results (instead of simulated ones)

@@
 numbered-lines: true
-bibliography: meeg_modelling.bib
+bibliography:
+  - meeg_modelling.bib
+  - grateful-refs.bib
 suppress-title-page: false
@@
     keep-tex: true
+    include-in-header:
+      - text: |
+          \usepackage{mathtools}
 editor_options: 
@@
 library(MetBrewer)
+library(grateful)
 library(permuco)
 library(scales)
+library(pakret)
 library(scico)
@@
 # setting a default ggplot2 theme
-# theme_set(theme_light(base_size = 10, base_family = "CMU Sans Serif") )
 theme_set(theme_light(base_size = 10, base_family = "Open Sans") )
+
+# template for citing R packages
+pkrt_set(pkg = "the `R` package `:pkg` v :ver [:ref]")
 ```
@@
 y_{i} &\sim \mathrm{EF}\left(\mu_{i}, \phi\right)\\
-g\left(\mu_i\right) &= \eta_{i} = A_{i} + \mathbf{X}_{i} \gamma + \sum_{j=1}^{J} f_{j}\left(x_{ij}\right)
+g\left(\mu_i\right) &= A_{i} + \mathbf{X}_{i} \gamma + \sum_{j=1}^{J} f_{j}\left(x_{ij}\right)
 \end{aligned}
@@
 
-where $g(\cdot)$ is the link function, $A$ is an offset, $\mathbf{X}_{i}$ is the $i$ th row of a parametric model matrix, $\gamma$ is a vector of parameters for the parametric terms, $f_{j}$ is a smooth function of covariate $x_{j}$, $y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)$ denotes that the observations $y_{i}$ are distributed as some member of the exponential family of distributions (e.g., Gaussian, Gamma, Beta, Poisson) with mean $\mu_{i}$ and scale parameter $\phi$.
-
+where $y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)$ denotes that the observations $y_{i}$ are distributed as some member of the exponential family of distributions (e.g., Gaussian, Gamma, Beta, Poisson) with mean $\mu_{i}$ and scale parameter $\phi$; $g(\cdot)$ is the link function, $A$ is an offset, $\mathbf{X}_{i}$ is the $i$th row of a parametric model matrix, $\gamma$ is a vector of parameters for the parametric terms, $f_{j}$ is a smooth function of covariate $x_{j}$.
 The smooth functions $f_{j}$ are represented in the model via penalised splines basis expansions of the covariates, that are a weighted sum of basis functions:
@@
 
-where $\beta_{jk}$ is the weight (coefficient) associated with the $k$ th basis function $b_{j k}()$ evaluated at the covariate value $x_{ij}$ for the $j$ th smooth function $f_{j}$...
+where $\beta_{jk}$ is the weight (coefficient) associated with the $k$th basis function $b_{jk}()$ evaluated at the covariate value $x_{ij}$ for the $j$th smooth function $f_{j}$.
+Splines' coefficients are penalised...
 
@@
 A Gaussian process (GP) is a stochastic process that defines the distribution over a collection of random variables indexed by a continuous variable, that is $\{f(t): t \in \mathcal{T}\}$ for some index set $\mathcal{T}$ [@riutort-mayol_practical_2023; @rasmussen2005].
-
-Note that nonparametric does not mean there aren't parameters, it means that there are infinitely many parameters.
-
-From [brms documentation](https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp): A GP is a stochastic process, which describes the relation between one or more predictors $x=\left(x_1, \ldots, x_d\right)$ and a response $f(x)$, where $d$ is the number of predictors.
+Whereas Bayesian linear regression outputs a distribution over the parameters of some predefind parametric model, the GP approach, in contrast, is a non-parametric approach, in that it finds a distribution over the possible functions that are consistent with the observed data.
+However, note that nonparametric does not mean there aren't parameters, it means that there are infinitely many parameters.
+
+From [brms documentation](https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp): A GP is a stochastic process, which describes the relation between one or more predictors $x=\left(x_{1}, \ldots, x_{d}\right)$ and a response $f(x)$, where $d$ is the number of predictors.
 A GP is the generalization of the multivariate normal distribution to an infinite number of dimensions.
@@
 $$
-\left(f\left(x_1\right), \ldots f\left(x_n\right) \sim \operatorname{MVN}\left(0,\left(k_p\left(x_i, x_j\right)\right)_{i, j=1}^n\right)\right.
-$$
-
-The smoothness and general behaviour of the function $f$ depends only on the choice of covariance kernel, see <https://mc-stan.org/docs/functions-reference/matrix_operations.html#gaussian-process-covariance-functions>.
-It ensures that values that are close together in the input space will be mapped to similar output values...
+\left(f\left(x_{1}\right), \ldots f\left(x_{n}\right) \sim \operatorname{MVN}\left(0, \left(k_p\left(x_{i}, x_{j}\right)\right)_{i, j=1}^{n}\right)\right.
+$$
+
+The smoothness and general behaviour of the function $f$ depends only on the choice of covariance kernel, which ensures that values that are close together in the input space will be mapped to similar output values...
+
 From this perspective, $f$ is a realisation of an infinite dimensional normal distribution:
@@
 In this article, we consider the squared-exponential (a.k.a.
-Gaussian or radial basis function) kernel, which computes the squared distance between points and converts it into a measure of similarity.
+radial basis function) kernel, which computes the squared distance between points and converts it into a measure of similarity.
 It is defined as:
@@
 $$
-C(\lambda):=C\left(t_1, t_2, \sigma, \gamma\right):=\sigma^2 \exp \left(-\frac{\left(t_1-t_2\right)^2}{2 \gamma^2}\right)
-$$
-
-with hyperparameters $\lambda=(\sigma, \gamma)$, expressing the overall scale of GP and the length-scale, respectively [@rasmussen2005].
+C(\lambda) := C\left(t_1, t_2, \sigma, \gamma\right) := \sigma^2 \exp \left(-\frac{||t_1-t_2||^{2}}{2 \gamma^2}\right)
+$$
+
+with hyperparameters $\lambda = (\sigma, \gamma)$, expressing the overall scale of GP and the length-scale, respectively [@rasmussen2005].
 The advantages of this kernel are that it is computationally efficient and (infinitely) smooth making it a reasonable choice for the purposes of the present article.
@@
 
+Taken from <https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10c-Gaussian-processes.html>: For a given vector $\mathbf{x}$, we can use the kernel to construct finite multi-variate normal distribution associated with it like so:
+
+$$
+\mathbf{x} \mapsto_{G P} \operatorname{MVNormal}(m(\mathbf{x}), k(\mathbf{x}, \mathbf{x}))
+$$
+
+where $m$ is a function that specifies the mean for the distribution associated with $\mathbf{x}$.
+This mapping is essentially the Gaussian process: a systematic association of vectors of arbitrary length with a suitable multi-variate normal distribution.
+
 Low-rank approximate Gaussian processes are of main interest in machine learning and statistics due to the high computational demands of exact Gaussian process models [@riutort-mayol_practical_2023]...
@@
 
+<!--
+
+For simulating EEG data, see <https://github.com/GRousselet/onsetsim> and <https://github.com/unfoldtoolbox/UnfoldSim.jl>...
+
+-->
+
 Following the approach used by @sassenhagen2019 and @rousselet_using_2025, we simulated EEG data stemming from two conditions, one with noise only, and the other with noise + signal.
-As in previous studies, the noise was generated by superimposing 50 sinusoids at different frequencies, following an EEG-like spectrum [see details and code in @yeung2004]...
-
-For simulating EEG data, see <https://github.com/GRousselet/onsetsim> and <https://github.com/unfoldtoolbox/UnfoldSim.jl>...
-
-Copy-pasted from @rousselet_using_2025: "The signal was a truncated Gaussian defining an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
-It was created by getting the probability density function of a standard normal with 93 points from x= 1.5 to 1.5, rescaled in amplitude between 0 and 1 and padded with 79 zeros to the left and to the right, to form a time series of 251 points, from 0 to 500 ms, in steps of 2 ms (500 Hz sampling rate)."
+As in previous studies, the noise was generated by superimposing 50 sinusoids at different frequencies, following an EEG-like spectrum [see details and code in @yeung2004].
+As in @rousselet_using_2025, the signal was generated from truncated Gaussian with an objective onset at 160 ms, a peak at 250 ms, and an offset at 342 ms.
+We simulated this signal for 250 timesteps between 0 and 0.5s, akin to & 500 Hz sampling rate.
+We simulated such data for a group of 20 participants with 50 trials per participant and condition.
 
@@
 source("code/erp_template.R")
+
+# importing helper functions
+source("code/functions.R")
 
@@
 ```{r, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Slope for the difference between conditions according to the GAM (left) or the GP (right)."}
+# check posterior_smooths()?
+# https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/posterior_smooths.brmsfit
 # defining a function to plot posterior slope
@@
         x = "Time (ms)",
-        y = expression(Pr(beta>0~"|"~data, model) / (1 - Pr(beta>0~"|"~data, model) ) )
+        y = expression(Pr(beta>0~"|"~data, model) / (1 - Pr(beta>0~"|"~data) ) )
         )
@@
     ylim(0, 1) +
+    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
+```
+
+```{r mne-decoding, out.width = "75%", eval = FALSE}
+# loading the reticulate package
+library(reticulate)
+
+# importing the numpy and mne python modules
+np <- import("numpy")
+mne <- import("mne")
+sklearn <- import("sklearn")
+
+# defining the function in R (it will be executed in Python)
+mne_decoding <- function (X, labels, ncores = 1) {
+
+    # converting R dataframe to NumPy array (reshaping if needed)
+    # X should be a matrix before conversion
+    X_np <- np$array(X)
+
+    if (length(dim(X_np) ) == 2) {
+
+        # adding a second dimension (channels) if missing
+        X_np <- np$expand_dims(X_np, axis = as.integer(1) )
+
+    }
+
+    # defining the classifier
+    clf <- sklearn$linear_model$LogisticRegression(solver = "liblinear")
+
+    # sliding the estimator on all time frames
+    time_decod <- mne$decoding$SlidingEstimator(
+        clf, n_jobs = as.integer(ncores),
+        scoring = "roc_auc", verbose = TRUE
+        )
+
+    # or using N-fold cross-validation
+    scores <- mne$decoding$cross_val_multiscore(
+        time_decod,
+        X_np,
+        labels,
+        cv = as.integer(4),
+        n_jobs = as.integer(ncores),
+        verbose = TRUE
+        )
+
+    # returning the scores (averaged over CV folds)
+    return (scores)
+
+}
+
+# listing all participants
+participants <- unique(raw_df$participant)
+
+# initialising empty decoding results
+group_decoding_scores <- data.frame()
+
+# running decoding for each participant
+for (ppt in participants) {
+
+    # printing progress
+    print(ppt)
+
+    # retrieve data from one participant
+    ppt_data <- raw_df %>%
+        filter(participant == ppt) %>%
+        select(-participant) %>%
+        pivot_wider(names_from = x, values_from = y) %>%
+        select(-condition, -trial)
+
+    # extracting the labels
+    labels <- raw_df %>%
+        filter(participant == ppt) %>%
+        select(-participant) %>%
+        pivot_wider(names_from = x, values_from = y) %>%
+        pull(condition) %>%
+        as.numeric()
+
+    # extracting the timesteps
+    timesteps <- raw_df %>%
+        filter(participant == ppt) %>%
+        select(-participant) %>%
+        pull(x) %>%
+        unique()
+
+    # running the decoding
+    decoding_scores <- data.frame(
+        mne_decoding(X = ppt_data, labels = labels-1)
+        ) %>%
+        # computing the average over CV folds
+        summarise(across(where(is.numeric), mean) )
+
+    # appending to previous results
+    group_decoding_scores <- bind_rows(group_decoding_scores, decoding_scores)
+
+}
+
+# saving the scores
+saveRDS(object = group_decoding_scores, file = "results/decoding_scores.rds")
+```
+
+```{r, out.width = "75%", fig.cap = "Exemplary group-level average timecourse of binary decoding accuracy (ROC AUC). Decoding was performed using MNE-Python."}
+# importing the decoding scores
+group_decoding_scores <- readRDS(file = "results/decoding_scores.rds")
+
+# extracting the timesteps
+timesteps <- raw_df %>%
+    # filter(participant == ppt) %>%
+    # select(-participant) %>%
+    pull(x) %>%
+    unique()
+
+# plotting it
+group_decoding_scores %>%
+    t() %>%
+    data.frame() %>%
+    mutate(
+        mean = rowMeans(across(1:20) ),
+        se = apply(across(1:20), 1, function (x) sd(x) / sqrt(length(x) ) )
+        ) %>%
+    mutate(lower = mean - se, upper = mean + se) %>%
+    mutate(time = timesteps) %>%
+    ggplot(aes(x = time, y = mean) ) +
+    geom_hline(yintercept = 0.5, linetype = "dashed") +
+    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "steelblue", alpha = 0.2) +
+    geom_line(color = "steelblue", size = 0.8) +
     labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
@@
             x = "Time (s)",
-            y = expression(Pr(AUC>chance~"|"~data, model) / (1 - Pr(AUC>chance~"|"~data, model) ) )
+            y = expression(
+                Pr(AUC>chance~"|"~data, model) / (1 - Pr(AUC>chance~"|"~data, model) )
+                )
             )
@@
 ```{r fig-sim-timegen, echo = FALSE, dpi = 300, fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
+# number of participants
+n_participants <- 10
+
 # number of timepoints
-n_participants <- 20
 # n_timepoints <- 240 # (1.2s at 200Hz)
@@
     summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>%
-    # pull(auc) %>% hist()
     ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
@@
     scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
-    # theme_light() +
     labs(x = "Testing time (s)", y = "Training time (s)")
@@
 
-Now, we want to test whether and when decoding performance is above chance level (0.5 for a binary decoding task). This one is computationally very intense...
-
-```{r gam-timegen, message = FALSE}
-# fitting a GAM with a 2D smooth function
+Now, we want to test whether and when decoding performance is above chance level (0.5 for a binary decoding task).
+These two models are computationally heavier to fit (more observations and 2D smooth functions)...
+
+```{r gam-gp-timegen, eval = FALSE, message = FALSE}
+# fitting a GAM with two temporal dimensions
 timegen_gam <- brm(
     # 2D thin-plate spline (tp) to model smooth interactions between training and testing time
-    auc ~ s(train_time, test_time, bs = "tp", k = 10),
+    # auc ~ s(train_time, test_time, bs = "tp", k = 10),
+    auc ~ t2(train_time, test_time, bs = "tp", k = 10),
     data = timegen_data,
     family = Beta(),
-    iter = 5000,
+    iter = 2000,
     chains = 4,
     cores = 4,
-    file = "models/timegen_gam.rds"
+    file = "models/timegen_gam_t2.rds"
+    )
+
+# fitting a GP with two temporal dimensions
+timegen_gp <- brm(
+    auc ~ gp(train_time, test_time, k = 20),
+    # data = timegen_data %>% mutate(auc_c = scale(x = auc, center = TRUE, scale = FALSE) ),
+    data = timegen_data,
+    family = Beta(),
+    control = list(adapt_delta = 0.95, max_treedepth = 20),
+    iter = 2000,
+    chains = 4,
+    cores = 4,
+    file = "models/timegen_gp.rds"
     )
@@
 
-```{r gp-timegen, eval = FALSE, message = FALSE}
-# fitting a GP with two temporal dimensions
-timegen_gp <- brm(
-    auc ~ gp(train_time, test_time, k = 20),
-    data = timegen_data,
-    family = Beta(),
-    control = list(adapt_delta = 0.95, max_treedepth = 20),
-    iter = 2000,
-    chains = 4,
-    cores = 4,
-    file = "models/timegen_gp.rds"
-    )
-```
-
 ```{r gp-timegen-preds, eval = FALSE, echo = FALSE, fig.cap = "Posterior predictions of the GP fitted on decoding accuracy over time."}
@@
     geom_line() +
-    # theme_light() +
     labs(x = "Time (s)", y = "AUC")
@@
 
-## Comparing the identified onsets to other approaches
-
-Comparing performance with uncorrected massive test, classical correction procedures (e.g., Bonferroni, FDR, FWER), cluster-based permutation test, cluster-depth, change point, [TFCE](https://jaromilfrossard.github.io/permuco4brain/articles/tfce.html)...
-We used the `permuco` package [@frossard2021]...
-
-```{r permutations-1d, echo = FALSE, fig.cap = "T-values timecourse with onset identified using the permuco package (clustermass method)."}
+## Comparing the identified onsets/offsets to other approaches
+
+From Rousselet (2024): Five methods for multiple comparison correction were considered: two FDR methods; two cluster-based methods; and the maximum statistics.
+The two FDR methods were BH95 (Benjamini & Hochberg, 1995) and BY01 (Benjamini & Yekutieli, 2001), which were applied to the permutation p-values using the `p.adjust` function in `R`.
+The first cluster-based inference was implemented using a cluster-sum statistic of squared t-values...
+
+Comparing performance of the GAM and GP models with:
+
+- uncorrected univariate tests
+- univariate tests + Bonferroni
+- univariate tests + BH95 (Benjamini-Hochberg, 1995) aka FDR
+- univariate tests + BY01 (Benjamini-Yekutieli, 2001)
+- cluster-based permutation test (cluster sum)
+- cluster-based permutation test (cluster depth)
+- threshold-free cluster enhancement
+- change point
+
+We used `r pkrt("permuco")` and `r pkrt("changepoint")`...
+
+```{r permutations-1d, echo = FALSE, eval = FALSE, fig.cap = "T-values timecourse with onset identified using the permuco package (clustermass method)."}
 # centering AUC relative to chance level (0.5)
@@
 
-```{r changepoint, echo = FALSE, fig.cap = "T-values timecourse with true onset and onset identified using the changepoint package (binary segmentation method)."}
-# computing the timecourse of t-values^2
-t_values <- raw_df %>%
-    summarise(y = mean(y), .by = c(x, condition, participant) ) %>%
+```{r test-1d, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Timecourse of squared t-values and s-values (continuous measure of evidence given by -log2(p-value)) with true onset and onset identified using the raw (uncorrected) p-values or the corrected p-values (BH, BY, Bonferroni, or Holm)."}
+# https://github.com/GRousselet/onsetsim/blob/main/docs/weakstrongfwer_demo.md
+# BH95: p.adjust(pvals, method = "BH")
+# BY01: p.adjust(perm.pvals, method = "BY")
+
+# defining an arbitrary alpha threshold
+aath <- 0.05
+
+# summarising raw_data per participant
+ppt_data <- raw_df %>%
+    summarise(y = mean(y), .by = c(participant, condition, x) ) %>%
     pivot_wider(names_from = condition, values_from = y) %>%
-    # head()
+    mutate(y_diff = cond2 - cond1)
+
+# massive univariate t-tests
+tests_results <- ppt_data %>%
     group_by(x) %>%
-    summarise(tvalues = t.test(cond1, cond2)$statistic^2 ) %>%
+    summarise(
+        # tval = t.test(x = auc_centered, mu = 0)$statistic^2,
+        # pval = t.test(x = auc_centered, mu = 0)$p.value
+        tval = t.test(x = y_diff, mu = 0)$statistic^2,
+        pval = t.test(x = y_diff, mu = 0)$p.value
+        ) %>%
+    mutate(
+        pval_bh = p.adjust(p = pval, method = "BH"),
+        pval_by = p.adjust(p = pval, method = "BY"),
+        pval_bonf = p.adjust(p = pval, method = "bonferroni"),
+        pval_holm = p.adjust(p = pval, method = "holm")
+        ) %>%
     ungroup()
 
+# finding onsets
+from_timestep <- 20
+onset_p <- find_onset(
+    tests_results$pval[from_timestep:nrow(tests_results)] <= aath,
+    tests_results$x[from_timestep:nrow(tests_results)]
+    )
+onset_bh <- find_onset(
+    tests_results$pval_bh[from_timestep:nrow(tests_results)] <= aath,
+    tests_results$x[from_timestep:nrow(tests_results)]
+    )
+onset_by <- find_onset(
+    tests_results$pval_by[from_timestep:nrow(tests_results)] <= aath,
+    tests_results$x[from_timestep:nrow(tests_results)]
+    )
+onset_bonf <- find_onset(
+    tests_results$pval_bonf[from_timestep:nrow(tests_results)] <= aath,
+    tests_results$x[from_timestep:nrow(tests_results)]
+    )
+onset_holm <- find_onset(
+    tests_results$pval_holm[from_timestep:nrow(tests_results)] <= aath,
+    tests_results$x[from_timestep:nrow(tests_results)]
+    )
+
+# plotting the results
+default_colours <- hue_pal()(5)
+true_onset <- 160
+true_offset <- 342
+tvals_plot <- tests_results %>%
+    ggplot(aes(x = x, y = tval) ) +
+    geom_area(position = "identity") +
+    geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = onset_p, linetype = 2, color = default_colours[1]) +
+    geom_vline(xintercept = onset_bh, linetype = 2, color = default_colours[2]) +
+    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[4]) +
+    geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
+    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[5]) +
+    labs(x = "Time (ms)", y = bquote(t^2) )
+
+plot_title <- paste0(
+    "True onset: ", true_onset, "ms, ",
+    "Uncorrected onset: ", round(onset_p, 3), "ms, ",
+    "BH onset: ", round(onset_bh, 3), "ms, ",
+    "BY onset: ", round(onset_by, 3), "ms,\n",
+    "Bonf onset: ", round(onset_bonf, 3), "ms, ",
+    "Holm onset: ", round(onset_holm, 3), "ms"
+    )
+
+# plotting -log2(p) (surprisal under H0) with thresholds (on p-values)
+# https://lesslikely.com/statistics/s-values/
+# a s-value of 4 (bits of information) is no more surprising than
+# getting all heads on 4 fair coin tosses
+pvals_plot <- tests_results %>%
+    pivot_longer(cols = pval:pval_holm) %>%
+    mutate(sval = -log2(value) ) %>%
+    ggplot(aes(x = x, y = sval, colour = name) ) +
+    geom_line(linewidth = 0.5) +
+    geom_hline(yintercept = -log2(aath), linetype = 2) +
+    geom_vline(xintercept = onset_p, linetype = 2, color = default_colours[1]) +
+    geom_vline(xintercept = onset_bh, linetype = 2, color = default_colours[2]) +
+    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[4]) +
+    geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
+    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[5]) +
+    labs(
+        x = "Time (ms)",
+        y = "-log2(p-value)",
+        colour = "correction"
+        )
+
+# combining the two plots
+tvals_plot + pvals_plot +
+    plot_annotation(title = plot_title) & theme(plot.title = element_text(hjust = 0.5) )
+```
+
+```{r changepoint, echo = FALSE, out.width = "75%", fig.cap = "T-values timecourse with true onset/offset and onset/offset identified using the changepoint package (binary segmentation method, in green)."}
 # using the changepoint package to identify onsets and offsets
-res <- cpt.meanvar(data = t_values$tvalues, method = "BinSeg", Q = 2)
+res <- cpt.meanvar(data = tests_results$tval, method = "BinSeg", Q = 2)
 
 # plotting the results
-true_onset <- 160
-t_values %>%
-    ggplot(aes(x = x, y = tvalues) ) +
+tests_results %>%
+    ggplot(aes(x = time, y = tval) ) +
     geom_line(linewidth = 1) +
-    geom_vline(xintercept = true_onset) +
-    geom_vline(xintercept = Xf[res@cpts[1]], linetype = "dotted") +
-    labs(x = "Time (ms)", y = bquote(t^2) ) +
-    # theme_light() +
-    ggtitle(paste0("Change point onset: ", Xf[res@cpts[1]], "ms, true onset: ", true_onset, "ms") )
-```
-
-Now cluster-based permutations in 2D...
-
-```{r permutations-2d, fig.cap = "Blah blah..."}
-# See https://github.com/jaromilfrossard/permuco4brain
-# devtools::install_github("jaromilfrossard/permuco4brain", build_vignettes = TRUE)
-library(permuco4brain)
-```
-
-```{python, echo = FALSE, eval = FALSE}
-# or using the MNE package in Python
-import mne
-```
+    geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = true_offset, linetype = 2) +
+    geom_vline(
+        xintercept = tests_results$time[res@cpts[1]],
+        linetype = 3, color = "darkgreen"
+        ) +
+    geom_vline(
+        xintercept = tests_results$time[res@cpts[2]],
+        linetype = 3, color = "darkgreen"
+        ) +
+    labs(x = "Time (s)", y = bquote(t^2) ) +
+    ggtitle(
+        paste0(
+            "True onset: ", true_onset, "s, ", "Change point onset: ",
+            round(tests_results$time[res@cpts[1]], 3), "s"
+            )
+        )
+```
+
+Now cluster-based permutations using `MNE-Python` [@gramfort2013] via `r pkrt("reticulate")`...
+
+```{r mne-cluster, echo = FALSE, eval = FALSE, fig.cap = "Cluster-based permutation tests via MNE-Python."}
+# loading the reticulate package
+library(reticulate)
+
+# importing the numpy and mne python modules
+np <- import("numpy")
+mne <- import("mne")
+
+# defining the function in R (it will be executed in Python)
+freq_stats_gat_matrix <- function (X) {
+
+    # converting R matrix to NumPy array
+    X_np <- np$array(X, dtype = "float64")
+
+    # running the statistical test
+    results <- mne$stats$spatio_temporal_cluster_1samp_test(
+        X_np,
+        out_type = "mask",
+        n_permutations = as.integer(1000),
+        n_jobs = as.integer(4),
+        verbose = TRUE
+        )
+
+    # extracting results
+    T_obs_ <- results[[1]]
+    clusters <- results[[2]]
+    p_values <- results[[3]]
+
+    # using the first slice of the 3D array
+    p_values_ <- np$transpose(np$ones_like(X_np[1, , drop = FALSE]) )
+
+    # assigning p-values to clusters
+    for (i in seq_along(clusters) ) {
+
+        # retrieving the current cluster
+        cluster_mask <- clusters[[i]][[1]]
+        idx_start <- cluster_mask$start+1
+        idx_stop <- cluster_mask$stop
+
+        # retrieving the p-value for this cluster
+        pval <- p_values[[i]]
+
+        # assigning this p-value to timesteps belonging
+        # to the current cluster
+        p_values_[idx_start:idx_stop] <- pval
+
+    }
+
+    # converting result back to R format
+    return (as.matrix(np$squeeze(np$transpose(p_values_) ) ) )
+
+}
+
+# converting decoding_data to a matrix
+data_matrix <- matrix(
+    decoding_data$auc,
+    ncol = length(unique(decoding_data$time) ),
+    byrow = TRUE
+    )
+
+# running the function
+chance_level <- 0.5
+p_values_results <- freq_stats_gat_matrix(data_matrix-chance_level)
+
+# converting back to a dataframe if needed
+p_values_df <- data.frame(
+    pval = p_values_results,
+    time = decoding_data$time
+    )
+
+# saving the scores
+saveRDS(object = p_values_df, file = "results/mne_permutation_decoding_scores.rds")
+```
+
+```{r fig-mne-cluster, out.width = "75%", fig.cap = "Cluster-based permutation tests via MNE-Python."}
+# importing the decoding scores
+p_values_df <- readRDS(file = "results/mne_permutation_decoding_scores.rds")
+
+# plotting the s-values (-log2(p-values))
+p_values_df %>%
+    mutate(sval = -log2(pval) ) %>%
+    ggplot(aes(x = time, y = sval) ) +
+    geom_line(linewidth = 0.5) +
+    geom_hline(yintercept = -log2(0.05), linetype = 2) +
+    # geom_area(position = "identity") +
+    labs(
+        x = "Time (ms)",
+        y = "-log2(p-value)"
+        )
+```
+
+\newpage
 
@@
 
+\newpage
+
 # Discussion
@@
 
-TO-DO: Prepare a wrapper `R` package and show how to call it in `Python` and integrate it with `MNE-Python` [@gramfort2013] pipelines...
+Prepare a wrapper `R` package and show how to call it in `Python` and integrate it with `MNE-Python` [@gramfort2013] pipelines...
 
@@
 ...
+
+\newpage
+
+# Packages
+
+```{r, echo = FALSE}
+# citing the R packages we have used
+cite_packages(output = "paragraph", omit = NULL, out.dir = ".")
+```
 
@@
 $$
-\text{TFCE} = \int_{h} e(h)^{E} h^{H} dh
-$$
-
-Where... Then, we can compute p-values for each timestep.... we permute conditions, compute the TFCE score for the permuted set, and take the max(TFCE)... But see @sassenhagen2019...
+\text{TFCE} = \int_{h} e(h)^{E} h^{H} \mathrm{d}h
+$$
+
+Where...
+the parameters $E$ and $H$ are set a priori and control the influence of the extend and height on the TFCE.
+Then, p-value for timestep/voxel $i$ is computed by comparing it TFCE with the null distribution of TFCE values.
+For each permuted signal, we keep the maximal value over the whole signal for the null distribution of the TFCE....
+But see @sassenhagen2019...
+
+\newpage
+
+# Using the `R` package and integration with `MNE-Python`
+
+Explain how to use the `R` package and to integrate it with `MNE` epochs...
+
+```{r r-python, eval = FALSE}
+# to-do adding some code here...
+```

### 2025-02-06 brms_meeg.tex
now modelling actual decoding results (instead of simulated ones)

@@
 
+\usepackage{mathtools}
 \makeatletter
@@
 \citeproc{ref-ehinger_unfold_2019}{Ehinger \& Dimigen, 2019};
-\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021},
+\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021a},
 \citeproc{ref-frossard2022}{2022}; \citeproc{ref-gramfort2013}{Gramfort,
@@
 \citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
-\citeproc{ref-pedersen_hierarchical_2019}{Pedersen et al., 2019};
+\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al., 2019};
 \citeproc{ref-pernet2015}{Pernet et al., 2015};
@@
 
-where \(g(\cdot)\) is the link function, \(A\) is an offset,
-\(\mathbf{X}_{i}\) is the \(i\) th row of a parametric model matrix,
+where \(y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)\) denotes that
+the observations \(y_{i}\) are distributed as some member of the
+exponential family of distributions (e.g., Gaussian, Gamma, Beta,
+Poisson) with mean \(\mu_{i}\) and scale parameter \(\phi\);
+\(g(\cdot)\) is the link function, \(A\) is an offset,
+\(\mathbf{X}_{i}\) is the \(i\)th row of a parametric model matrix,
 \(\gamma\) is a vector of parameters for the parametric terms, \(f_{j}\)
-is a smooth function of covariate \(x_{j}\),
-\(y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)\) denotes that the
-observations \(y_{i}\) are distributed as some member of the exponential
-family of distributions (e.g., Gaussian, Gamma, Beta, Poisson) with mean
-\(\mu_{i}\) and scale parameter \(\phi\).
-
-The smooth functions \(f_{j}\) are represented in the model via
-penalised splines basis expansions of the covariates, that are a
-weighted sum of basis functions:
+is a smooth function of covariate \(x_{j}\). The smooth functions
+\(f_{j}\) are represented in the model via penalised splines basis
+expansions of the covariates, that are a weighted sum of basis
+functions:
 
@@
 where \(\beta_{jk}\) is the weight (coefficient) associated with the
-\(k\) th basis function \(b_{j k}()\) evaluated at the covariate value
-\(x_{ij}\) for the \(j\) th smooth function \(f_{j}\)\ldots{}
+\(k\)th basis function \(b_{jk}()\) evaluated at the covariate value
+\(x_{ij}\) for the \(j\)th smooth function \(f_{j}\). Splines'
+coefficients are penalised\ldots{}
 
@@
 \citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
-2023}).
-
-Note that nonparametric does not mean there aren't parameters, it means
-that there are infinitely many parameters.
+2023}). Whereas Bayesian linear regression outputs a distribution over
+the parameters of some predefind parametric model, the GP approach, in
+contrast, is a non-parametric approach, in that it finds a distribution
+over the possible functions that are consistent with the observed data.
+However, note that nonparametric does not mean there aren't parameters,
+it means that there are infinitely many parameters.
 
@@
 relation between one or more predictors
-\(x=\left(x_1, \ldots, x_d\right)\) and a response \(f(x)\), where \(d\)
-is the number of predictors. A GP is the generalization of the
+\(x=\left(x_{1}, \ldots, x_{d}\right)\) and a response \(f(x)\), where
+\(d\) is the number of predictors. A GP is the generalization of the
 multivariate normal distribution to an infinite number of dimensions.
@@
 \[
-\left(f\left(x_1\right), \ldots f\left(x_n\right) \sim \operatorname{MVN}\left(0,\left(k_p\left(x_i, x_j\right)\right)_{i, j=1}^n\right)\right.
+\left(f\left(x_{1}\right), \ldots f\left(x_{n}\right) \sim \operatorname{MVN}\left(0, \left(k_p\left(x_{i}, x_{j}\right)\right)_{i, j=1}^{n}\right)\right.
 \]
@@
 The smoothness and general behaviour of the function \(f\) depends only
-on the choice of covariance kernel, see
-\url{https://mc-stan.org/docs/functions-reference/matrix_operations.html\#gaussian-process-covariance-functions}.
-It ensures that values that are close together in the input space will
-be mapped to similar output values\ldots{} From this perspective, \(f\)
-is a realisation of an infinite dimensional normal distribution:
+on the choice of covariance kernel, which ensures that values that are
+close together in the input space will be mapped to similar output
+values\ldots{}
+
+From this perspective, \(f\) is a realisation of an infinite dimensional
+normal distribution:
 
@@
 choices of the covariance kernel lead to different GPs. In this article,
-we consider the squared-exponential (a.k.a. Gaussian or radial basis
-function) kernel, which computes the squared distance between points and
-converts it into a measure of similarity. It is defined as:
+we consider the squared-exponential (a.k.a. radial basis function)
+kernel, which computes the squared distance between points and converts
+it into a measure of similarity. It is defined as:
 
 \[
-C(\lambda):=C\left(t_1, t_2, \sigma, \gamma\right):=\sigma^2 \exp \left(-\frac{\left(t_1-t_2\right)^2}{2 \gamma^2}\right)
+C(\lambda) := C\left(t_1, t_2, \sigma, \gamma\right) := \sigma^2 \exp \left(-\frac{||t_1-t_2||^{2}}{2 \gamma^2}\right)
 \]
 
-with hyperparameters \(\lambda=(\sigma, \gamma)\), expressing the
+with hyperparameters \(\lambda = (\sigma, \gamma)\), expressing the
 overall scale of GP and the length-scale, respectively
@@
 estimated from the data, along with all other model parameters.
+
+Taken from
+\url{https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10c-Gaussian-processes.html}:
+For a given vector \(\mathbf{x}\), we can use the kernel to construct
+finite multi-variate normal distribution associated with it like so:
+
+\[
+\mathbf{x} \mapsto_{G P} \operatorname{MVNormal}(m(\mathbf{x}), k(\mathbf{x}, \mathbf{x}))
+\]
+
+where \(m\) is a function that specifies the mean for the distribution
+associated with \(\mathbf{x}\). This mapping is essentially the Gaussian
+process: a systematic association of vectors of arbitrary length with a
+suitable multi-variate normal distribution.
 
@@
 EEG-like spectrum (see details and code in
-\citeproc{ref-yeung2004}{Yeung et al., 2004})\ldots{}
-
-For simulating EEG data, see
-\url{https://github.com/GRousselet/onsetsim} and
-\url{https://github.com/unfoldtoolbox/UnfoldSim.jl}\ldots{}
-
-Copy-pasted from Rousselet (\citeproc{ref-rousselet_using_2025}{2025}):
-``The signal was a truncated Gaussian defining an objective onset at 160
-ms, a maximum at 250 ms, and an offset at 342 ms. It was created by
-getting the probability density function of a standard normal with 93
-points from x= 1.5 to 1.5, rescaled in amplitude between 0 and 1 and
-padded with 79 zeros to the left and to the right, to form a time series
-of 251 points, from 0 to 500 ms, in steps of 2 ms (500 Hz sampling
-rate).''
+\citeproc{ref-yeung2004}{Yeung et al., 2004}). As in Rousselet
+(\citeproc{ref-rousselet_using_2025}{2025}), the signal was generated
+from truncated Gaussian with an objective onset at 160 ms, a peak at 250
+ms, and an offset at 342 ms. We simulated this signal for 250 timesteps
+between 0 and 0.5s, akin to \& 500 Hz sampling rate. We simulated such
+data for a group of 20 participants with 50 trials per participant and
+condition.
 
@@
 Models were fitted using the \texttt{brms} package
-(\citeproc{ref-burkner_brms_2017}{Bürkner, 2017},
-\citeproc{ref-burkner_advanced_2018}{2018};
+(\citeproc{ref-burkner_brms_2017}{Bürkner, 2017a},
+\citeproc{ref-burkner_advanced_2018}{2018a};
 \citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019})\ldots{}
@@
     \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
-    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{20}\NormalTok{),}
+    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{20}\NormalTok{),}
     \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
@@
 \includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-decoding-1.pdf}
+
+}
+
+\end{figure}%
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# loading the reticulate package}
+\FunctionTok{library}\NormalTok{(reticulate)}
+
+\CommentTok{\# importing the numpy and mne python modules}
+\NormalTok{np }\OtherTok{\textless{}{-}} \FunctionTok{import}\NormalTok{(}\StringTok{"numpy"}\NormalTok{)}
+\NormalTok{mne }\OtherTok{\textless{}{-}} \FunctionTok{import}\NormalTok{(}\StringTok{"mne"}\NormalTok{)}
+\NormalTok{sklearn }\OtherTok{\textless{}{-}} \FunctionTok{import}\NormalTok{(}\StringTok{"sklearn"}\NormalTok{)}
+
+\CommentTok{\# defining the function in R (it will be executed in Python)}
+\NormalTok{mne\_decoding }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{ (X, labels, }\AttributeTok{ncores =} \DecValTok{1}\NormalTok{) \{}
+
+    \CommentTok{\# converting R dataframe to NumPy array (reshaping if needed)}
+    \CommentTok{\# X should be a matrix before conversion}
+\NormalTok{    X\_np }\OtherTok{\textless{}{-}}\NormalTok{ np}\SpecialCharTok{$}\FunctionTok{array}\NormalTok{(X)}
+
+    \ControlFlowTok{if}\NormalTok{ (}\FunctionTok{length}\NormalTok{(}\FunctionTok{dim}\NormalTok{(X\_np) ) }\SpecialCharTok{==} \DecValTok{2}\NormalTok{) \{}
+
+        \CommentTok{\# adding a second dimension (channels) if missing}
+\NormalTok{        X\_np }\OtherTok{\textless{}{-}}\NormalTok{ np}\SpecialCharTok{$}\FunctionTok{expand\_dims}\NormalTok{(X\_np, }\AttributeTok{axis =} \FunctionTok{as.integer}\NormalTok{(}\DecValTok{1}\NormalTok{) )}
+
+\NormalTok{    \}}
+
+    \CommentTok{\# defining the classifier}
+\NormalTok{    clf }\OtherTok{\textless{}{-}}\NormalTok{ sklearn}\SpecialCharTok{$}\NormalTok{linear\_model}\SpecialCharTok{$}\FunctionTok{LogisticRegression}\NormalTok{(}\AttributeTok{solver =} \StringTok{"liblinear"}\NormalTok{)}
+
+    \CommentTok{\# sliding the estimator on all time frames}
+\NormalTok{    time\_decod }\OtherTok{\textless{}{-}}\NormalTok{ mne}\SpecialCharTok{$}\NormalTok{decoding}\SpecialCharTok{$}\FunctionTok{SlidingEstimator}\NormalTok{(}
+\NormalTok{        clf, }\AttributeTok{n\_jobs =} \FunctionTok{as.integer}\NormalTok{(ncores),}
+        \AttributeTok{scoring =} \StringTok{"roc\_auc"}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{TRUE}
+\NormalTok{        )}
+
+    \CommentTok{\# or using N{-}fold cross{-}validation}
+\NormalTok{    scores }\OtherTok{\textless{}{-}}\NormalTok{ mne}\SpecialCharTok{$}\NormalTok{decoding}\SpecialCharTok{$}\FunctionTok{cross\_val\_multiscore}\NormalTok{(}
+\NormalTok{        time\_decod,}
+\NormalTok{        X\_np,}
+\NormalTok{        labels,}
+        \AttributeTok{cv =} \FunctionTok{as.integer}\NormalTok{(}\DecValTok{4}\NormalTok{),}
+        \AttributeTok{n\_jobs =} \FunctionTok{as.integer}\NormalTok{(ncores),}
+        \AttributeTok{verbose =} \ConstantTok{TRUE}
+\NormalTok{        )}
+
+    \CommentTok{\# returning the scores (averaged over CV folds)}
+    \FunctionTok{return}\NormalTok{ (scores)}
+
+\NormalTok{\}}
+
+\CommentTok{\# listing all participants}
+\NormalTok{participants }\OtherTok{\textless{}{-}} \FunctionTok{unique}\NormalTok{(raw\_df}\SpecialCharTok{$}\NormalTok{participant)}
+
+\CommentTok{\# initialising empty decoding results}
+\NormalTok{group\_decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{()}
+
+\CommentTok{\# running decoding for each participant}
+\ControlFlowTok{for}\NormalTok{ (ppt }\ControlFlowTok{in}\NormalTok{ participants) \{}
+
+    \CommentTok{\# printing progress}
+    \FunctionTok{print}\NormalTok{(ppt)}
+
+    \CommentTok{\# retrieve data from one participant}
+\NormalTok{    ppt\_data }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{filter}\NormalTok{(participant }\SpecialCharTok{==}\NormalTok{ ppt) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{participant) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ x, }\AttributeTok{values\_from =}\NormalTok{ y) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{condition, }\SpecialCharTok{{-}}\NormalTok{trial)}
+
+    \CommentTok{\# extracting the labels}
+\NormalTok{    labels }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{filter}\NormalTok{(participant }\SpecialCharTok{==}\NormalTok{ ppt) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{participant) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ x, }\AttributeTok{values\_from =}\NormalTok{ y) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{pull}\NormalTok{(condition) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{as.numeric}\NormalTok{()}
+
+    \CommentTok{\# extracting the timesteps}
+\NormalTok{    timesteps }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{filter}\NormalTok{(participant }\SpecialCharTok{==}\NormalTok{ ppt) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{participant) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{pull}\NormalTok{(x) }\SpecialCharTok{\%\textgreater{}\%}
+        \FunctionTok{unique}\NormalTok{()}
+
+    \CommentTok{\# running the decoding}
+\NormalTok{    decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
+        \FunctionTok{mne\_decoding}\NormalTok{(}\AttributeTok{X =}\NormalTok{ ppt\_data, }\AttributeTok{labels =}\NormalTok{ labels}\DecValTok{{-}1}\NormalTok{)}
+\NormalTok{        ) }\SpecialCharTok{\%\textgreater{}\%}
+        \CommentTok{\# computing the average over CV folds}
+        \FunctionTok{summarise}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{where}\NormalTok{(is.numeric), mean) )}
+
+    \CommentTok{\# appending to previous results}
+\NormalTok{    group\_decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(group\_decoding\_scores, decoding\_scores)}
+
+\NormalTok{\}}
+
+\CommentTok{\# saving the scores}
+\FunctionTok{saveRDS}\NormalTok{(}\AttributeTok{object =}\NormalTok{ group\_decoding\_scores, }\AttributeTok{file =} \StringTok{"results/decoding\_scores.rds"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# importing the decoding scores}
+\NormalTok{group\_decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\AttributeTok{file =} \StringTok{"results/decoding\_scores.rds"}\NormalTok{)}
+
+\CommentTok{\# extracting the timesteps}
+\NormalTok{timesteps }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
+    \CommentTok{\# filter(participant == ppt) \%\textgreater{}\%}
+    \CommentTok{\# select({-}participant) \%\textgreater{}\%}
+    \FunctionTok{pull}\NormalTok{(x) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{unique}\NormalTok{()}
+
+\CommentTok{\# plotting it}
+\NormalTok{group\_decoding\_scores }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{t}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{data.frame}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}
+        \AttributeTok{mean =} \FunctionTok{rowMeans}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{20}\NormalTok{) ),}
+        \AttributeTok{se =} \FunctionTok{apply}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{20}\NormalTok{), }\DecValTok{1}\NormalTok{, }\ControlFlowTok{function}\NormalTok{ (x) }\FunctionTok{sd}\NormalTok{(x) }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(}\FunctionTok{length}\NormalTok{(x) ) )}
+\NormalTok{        ) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lower =}\NormalTok{ mean }\SpecialCharTok{{-}}\NormalTok{ se, }\AttributeTok{upper =}\NormalTok{ mean }\SpecialCharTok{+}\NormalTok{ se) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{time =}\NormalTok{ timesteps) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time, }\AttributeTok{y =}\NormalTok{ mean) ) }\SpecialCharTok{+}
+    \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+    \FunctionTok{geom\_ribbon}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{ymin =}\NormalTok{ lower, }\AttributeTok{ymax =}\NormalTok{ upper), }\AttributeTok{fill =} \StringTok{"steelblue"}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\SpecialCharTok{+}
+    \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{color =} \StringTok{"steelblue"}\NormalTok{, }\AttributeTok{size =} \FloatTok{0.8}\NormalTok{) }\SpecialCharTok{+}
+    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (s)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Decoding accuracy (ROC AUC)"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[H]
+
+\caption{Exemplary group-level average timecourse of binary decoding
+accuracy (ROC AUC). Decoding was performed using MNE-Python.}
+
+{\centering \includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/unnamed-chunk-2-1.pdf}
 
@@
 Now, we want to test whether and when decoding performance is above
-chance level (0.5 for a binary decoding task). This one is
-computationally very intense\ldots{}
+chance level (0.5 for a binary decoding task). These two models are
+computationally heavier to fit (more observations and 2D smooth
+functions)\ldots{}
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# fitting a GAM with a 2D smooth function}
+\CommentTok{\# fitting a GAM with two temporal dimensions}
 \NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
     \CommentTok{\# 2D thin{-}plate spline (tp) to model smooth interactions between training and testing time}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{s}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
+    \CommentTok{\# auc \textasciitilde{} s(train\_time, test\_time, bs = "tp", k = 10),}
+\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
     \AttributeTok{data =}\NormalTok{ timegen\_data,}
     \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
+    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
     \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
     \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/timegen\_gam.rds"}
+    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"}
 \NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-\caption{Posterior probability of decoding accuracy being above chance
-level (2D GAM).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-timegen-post-preds-1.pdf}
-
-}
-
-\end{figure}%
-
-\begin{Shaded}
-\begin{Highlighting}[]
+
 \CommentTok{\# fitting a GP with two temporal dimensions}
@@
 \NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gp}\NormalTok{(train\_time, test\_time, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
+    \CommentTok{\# data = timegen\_data \%\textgreater{}\% mutate(auc\_c = scale(x = auc, center = TRUE, scale = FALSE) ),}
     \AttributeTok{data =}\NormalTok{ timegen\_data,}
     \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
+    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{20}\NormalTok{),}
+    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
     \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
@@
 
+\begin{figure}[H]
+
+\caption{Posterior probability of decoding accuracy being above chance
+level (2D GAM).}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-timegen-post-preds-1.pdf}
+
+}
+
+\end{figure}%
+
 \subsection{Application to actual M/EEG
@@
 
-\subsection{Comparing the identified onsets to other
-approaches}\label{comparing-the-identified-onsets-to-other-approaches}
-
-Comparing performance with uncorrected massive test, classical
-correction procedures (e.g., Bonferroni, FDR, FWER), cluster-based
-permutation test, cluster-depth, change point,
-\href{https://jaromilfrossard.github.io/permuco4brain/articles/tfce.html}{TFCE}\ldots{}
-We used the \texttt{permuco} package
-(\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021})\ldots{}
-
-\begin{verbatim}
-Effect: intercept.
-Alternative Hypothesis: two.sided.
-Statistic: fisher(1, 19).
-Resampling Method: freedman_lane.
-Type of Resampling: signflip.
-Number of Dependant Variables: 240.
-Number of Resamples: 5000.
-Multiple Comparisons Procedure: clustermass.
-Threshold: 4.38075.
-Mass Function: the sum.
-Table of clusters.
-
-  start end cluster mass P(>mass)
-1    25  25     4.883149   1.0000
-2   127 127     6.183416   0.4212
-3   218 218     5.080433   1.0000
-\end{verbatim}
+\subsection{Comparing the identified onsets/offsets to other
+approaches}\label{comparing-the-identified-onsetsoffsets-to-other-approaches}
+
+From Rousselet (2024): Five methods for multiple comparison correction
+were considered: two FDR methods; two cluster-based methods; and the
+maximum statistics. The two FDR methods were BH95 (Benjamini \&
+Hochberg, 1995) and BY01 (Benjamini \& Yekutieli, 2001), which were
+applied to the permutation p-values using the \texttt{p.adjust} function
+in \texttt{R}. The first cluster-based inference was implemented using a
+cluster-sum statistic of squared t-values\ldots{}
+
+Comparing performance of the GAM and GP models with:
+
+\begin{itemize}
+\tightlist
+\item
+  uncorrected univariate tests
+\item
+  univariate tests + Bonferroni
+\item
+  univariate tests + BH95 (Benjamini-Hochberg, 1995) aka FDR
+\item
+  univariate tests + BY01 (Benjamini-Yekutieli, 2001)
+\item
+  cluster-based permutation test (cluster sum)
+\item
+  cluster-based permutation test (cluster depth)
+\item
+  threshold-free cluster enhancement
+\item
+  change point
+\end{itemize}
+
+We used the \texttt{R} package \texttt{permuco} v 1.1.3
+(\citeproc{ref-permuco}{Frossard \& Renaud, 2021b}) and the \texttt{R}
+package \texttt{changepoint} v 2.2.4 (\citeproc{ref-changepoint}{Killick
+et al., 2022a})\ldots{}
 
@@
 
-\caption{T-values timecourse with onset identified using the permuco
-package (clustermass method).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/permutations-1d-1.pdf}
+\caption{Timecourse of squared t-values and s-values (continuous measure
+of evidence given by -log2(p-value)) with true onset and onset
+identified using the raw (uncorrected) p-values or the corrected
+p-values (BH, BY, Bonferroni, or Holm).}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/test-1d-1.pdf}
 
@@
 
-\caption{T-values timecourse with true onset and onset identified using
-the changepoint package (binary segmentation method).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/changepoint-1.pdf}
+\caption{T-values timecourse with true onset/offset and onset/offset
+identified using the changepoint package (binary segmentation method, in
+green).}
+
+{\centering \includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/changepoint-1.pdf}
 
@@
 
-Now cluster-based permutations in 2D\ldots{}
+Now cluster-based permutations using \texttt{MNE-Python}
+(\citeproc{ref-gramfort2013}{Gramfort, 2013}) via the \texttt{R} package
+\texttt{reticulate} v 1.35.0 (\citeproc{ref-reticulate}{Ushey et al.,
+2024})\ldots{}
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# See https://github.com/jaromilfrossard/permuco4brain}
-\CommentTok{\# devtools::install\_github("jaromilfrossard/permuco4brain", build\_vignettes = TRUE)}
-\FunctionTok{library}\NormalTok{(permuco4brain)}
+\CommentTok{\# importing the decoding scores}
+\NormalTok{p\_values\_df }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\AttributeTok{file =} \StringTok{"results/mne\_permutation\_decoding\_scores.rds"}\NormalTok{)}
+
+\CommentTok{\# plotting the s{-}values ({-}log2(p{-}values))}
+\NormalTok{p\_values\_df }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{sval =} \SpecialCharTok{{-}}\FunctionTok{log2}\NormalTok{(pval) ) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time, }\AttributeTok{y =}\NormalTok{ sval) ) }\SpecialCharTok{+}
+    \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{linewidth =} \FloatTok{0.5}\NormalTok{) }\SpecialCharTok{+}
+    \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \SpecialCharTok{{-}}\FunctionTok{log2}\NormalTok{(}\FloatTok{0.05}\NormalTok{), }\AttributeTok{linetype =} \DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
+    \CommentTok{\# geom\_area(position = "identity") +}
+    \FunctionTok{labs}\NormalTok{(}
+        \AttributeTok{x =} \StringTok{"Time (ms)"}\NormalTok{,}
+        \AttributeTok{y =} \StringTok{"{-}log2(p{-}value)"}
+\NormalTok{        )}
 \end{Highlighting}
@@
 
+\begin{figure}[!htb]
+
+\caption{\label{fig-mne-cluster}Cluster-based permutation tests via
+MNE-Python.}
+
+\centering{
+
+\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-mne-cluster-1.pdf}
+
+}
+
+\end{figure}%
+
+\newpage
+
 \section{Results}\label{results}
@@
 \ldots{}
+
+\newpage
 
@@
 
-TO-DO: Prepare a wrapper \texttt{R} package and show how to call it in
+Prepare a wrapper \texttt{R} package and show how to call it in
 \texttt{Python} and integrate it with \texttt{MNE-Python}
@@
 
+\section{Packages}\label{packages}
+
+We used R version 4.2.3 (\citeproc{ref-base}{R Core Team, 2023}) and the
+following R packages: brms v. 2.22.0 (\citeproc{ref-brms2017}{Bürkner,
+2017b}, \citeproc{ref-brms2018}{2018b}, \citeproc{ref-brms2021}{2021}),
+changepoint v. 2.2.4 (\citeproc{ref-changepoint2022}{Killick et al.,
+2022b}; \citeproc{ref-changepoint2014}{Killick \& Eckley, 2014}),
+grateful v. 0.2.10 (\citeproc{ref-grateful}{Rodriguez-Sanchez \&
+Jackson, 2023}), knitr v. 1.45 (\citeproc{ref-knitr2014}{Xie, 2014},
+\citeproc{ref-knitr2015}{2015}, \citeproc{ref-knitr2023}{2023}),
+MetBrewer v. 0.2.0 (\citeproc{ref-MetBrewer}{Mills, 2022}), pakret v.
+0.2.2 (\citeproc{ref-pakret}{Gallou, 2024}), patchwork v. 1.2.0
+(\citeproc{ref-patchwork}{T. L. Pedersen, 2024}), permuco v. 1.1.3
+(\citeproc{ref-permuco}{Frossard \& Renaud, 2021b}), rmarkdown v. 2.29
+(\citeproc{ref-rmarkdown2024}{Allaire et al., 2024};
+\citeproc{ref-rmarkdown2018}{Xie et al., 2018},
+\citeproc{ref-rmarkdown2020}{2020}), scales v. 1.3.0
+(\citeproc{ref-scales}{Wickham et al., 2023}), scico v. 1.5.0
+(\citeproc{ref-scico}{T. L. Pedersen \& Crameri, 2023}), tidybayes v.
+3.0.6 (\citeproc{ref-tidybayes}{Kay, 2023}), tidyverse v. 2.0.0
+(\citeproc{ref-tidyverse}{Wickham et al., 2019}).
+
+\newpage
+
 \section{References}\label{references}
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-rmarkdown2024}
+Allaire, J., Xie, Y., Dervieux, C., McPherson, J., Luraschi, J., Ushey,
+K., Atkins, A., Wickham, H., Cheng, J., Chang, W., \& Iannone, R.
+(2024). \emph{{rmarkdown}: Dynamic documents for r}.
+\url{https://github.com/rstudio/rmarkdown}
+
 \bibitem[\citeproctext]{ref-burkner_brms_2017}
-Bürkner, P.-C. (2017). \textbf{Brms} : {An} {{\emph{R}}} {Package} for
+Bürkner, P.-C. (2017a). \textbf{Brms} : {An} {{\emph{R}}} {Package} for
 {Bayesian Multilevel Models Using} {{\emph{Stan}}}. \emph{Journal of
@@
 
+\bibitem[\citeproctext]{ref-brms2017}
+Bürkner, P.-C. (2017b). {brms}: An {R} package for {Bayesian} multilevel
+models using {Stan}. \emph{Journal of Statistical Software},
+\emph{80}(1), 1--28. \url{https://doi.org/10.18637/jss.v080.i01}
+
 \bibitem[\citeproctext]{ref-burkner_advanced_2018}
-Bürkner, P.-C. (2018). Advanced {Bayesian Multilevel Modeling} with the
+Bürkner, P.-C. (2018a). Advanced {Bayesian Multilevel Modeling} with the
 {R Package} brms. \emph{The R Journal}, \emph{10}(1), 395.
 \url{https://doi.org/10.32614/RJ-2018-017}
+
+\bibitem[\citeproctext]{ref-brms2018}
+Bürkner, P.-C. (2018b). Advanced {Bayesian} multilevel modeling with the
+{R} package {brms}. \emph{The R Journal}, \emph{10}(1), 395--411.
+\url{https://doi.org/10.32614/RJ-2018-017}
+
+\bibitem[\citeproctext]{ref-brms2021}
+Bürkner, P.-C. (2021). Bayesian item response modeling in {R} with
+{brms} and {Stan}. \emph{Journal of Statistical Software},
+\emph{100}(5), 1--54. \url{https://doi.org/10.18637/jss.v100.i05}
 
@@
 \bibitem[\citeproctext]{ref-frossard2021}
-Frossard, J., \& Renaud, O. (2021). Permutation Tests for Regression,
+Frossard, J., \& Renaud, O. (2021a). Permutation Tests for Regression,
 ANOVA, and Comparison of Signals: The {\textbf{permuco}} Package.
 \emph{Journal of Statistical Software}, \emph{99}(15).
+\url{https://doi.org/10.18637/jss.v099.i15}
+
+\bibitem[\citeproctext]{ref-permuco}
+Frossard, J., \& Renaud, O. (2021b). Permutation tests for regression,
+{ANOVA}, and comparison of signals: The {permuco} package. \emph{Journal
+of Statistical Software}, \emph{99}(15), 1--32.
 \url{https://doi.org/10.18637/jss.v099.i15}
@@
 
+\bibitem[\citeproctext]{ref-pakret}
+Gallou, A. (2024). \emph{{pakret}: Cite {``{R}''} packages on the fly in
+{``{R Markdown}''} and {``{Quarto}''}}.
+\url{https://CRAN.R-project.org/package=pakret}
+
 \bibitem[\citeproctext]{ref-gramfort2013}
@@
 \url{https://doi.org/10.1016/j.neuroimage.2003.08.003}
+
+\bibitem[\citeproctext]{ref-tidybayes}
+Kay, M. (2023). \emph{{tidybayes}: Tidy data and geoms for {Bayesian}
+models}. \url{https://doi.org/10.5281/zenodo.1308151}
+
+\bibitem[\citeproctext]{ref-changepoint2014}
+Killick, R., \& Eckley, I. A. (2014). {changepoint}: An {R} package for
+changepoint analysis. \emph{Journal of Statistical Software},
+\emph{58}(3), 1--19.
+\url{https://www.jstatsoft.org/article/view/v058i03}
+
+\bibitem[\citeproctext]{ref-changepoint}
+Killick, R., Haynes, K., \& Eckley, I. A. (2022a). \emph{{changepoint}:
+An {R} package for changepoint analysis}.
+\url{https://CRAN.R-project.org/package=changepoint}
+
+\bibitem[\citeproctext]{ref-changepoint2022}
+Killick, R., Haynes, K., \& Eckley, I. A. (2022b). \emph{{changepoint}:
+An {R} package for changepoint analysis}.
+\url{https://CRAN.R-project.org/package=changepoint}
 
@@
 \url{https://doi.org/10.1016/j.jneumeth.2007.03.024}
+
+\bibitem[\citeproctext]{ref-MetBrewer}
+Mills, B. R. (2022). \emph{{MetBrewer}: Color palettes inspired by works
+at the metropolitan museum of art}.
+\url{https://CRAN.R-project.org/package=MetBrewer}
 
@@
 
+\bibitem[\citeproctext]{ref-patchwork}
+Pedersen, T. L. (2024). \emph{{patchwork}: The composer of plots}.
+\url{https://CRAN.R-project.org/package=patchwork}
+
+\bibitem[\citeproctext]{ref-scico}
+Pedersen, T. L., \& Crameri, F. (2023). \emph{{scico}: Colour palettes
+based on the scientific colour-maps}.
+\url{https://CRAN.R-project.org/package=scico}
+
 \bibitem[\citeproctext]{ref-pernet2015}
@@
 
+\bibitem[\citeproctext]{ref-base}
+R Core Team. (2023). \emph{{R}: A language and environment for
+statistical computing}. R Foundation for Statistical Computing.
+\url{https://www.R-project.org/}
+
 \bibitem[\citeproctext]{ref-rasmussen2005}
@@
 
+\bibitem[\citeproctext]{ref-grateful}
+Rodriguez-Sanchez, F., \& Jackson, C. P. (2023). \emph{{grateful}:
+Facilitate citation of r packages}.
+\url{https://pakillo.github.io/grateful/}
+
 \bibitem[\citeproctext]{ref-rosenblatt2018}
@@
 
+\bibitem[\citeproctext]{ref-reticulate}
+Ushey, K., Allaire, J., \& Tang, Y. (2024). \emph{Reticulate: Interface
+to 'python'}. \url{https://CRAN.R-project.org/package=reticulate}
+
+\bibitem[\citeproctext]{ref-tidyverse}
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
+François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
+Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
+Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
+Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
+\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
+
+\bibitem[\citeproctext]{ref-scales}
+Wickham, H., Pedersen, T. L., \& Seidel, D. (2023). \emph{{scales}:
+Scale functions for visualization}.
+\url{https://CRAN.R-project.org/package=scales}
+
 \bibitem[\citeproctext]{ref-wood2003}
@@
 \url{https://doi.org/10.1198/016214504000000980}
+
+\bibitem[\citeproctext]{ref-knitr2014}
+Xie, Y. (2014). {knitr}: A comprehensive tool for reproducible research
+in {R}. In V. Stodden, F. Leisch, \& R. D. Peng (Eds.),
+\emph{Implementing reproducible computational research}. Chapman;
+Hall/CRC.
+
+\bibitem[\citeproctext]{ref-knitr2015}
+Xie, Y. (2015). \emph{Dynamic documents with {R} and knitr} (2nd ed.).
+Chapman; Hall/CRC. \url{https://yihui.org/knitr/}
+
+\bibitem[\citeproctext]{ref-knitr2023}
+Xie, Y. (2023). \emph{{knitr}: A general-purpose package for dynamic
+report generation in r}. \url{https://yihui.org/knitr/}
+
+\bibitem[\citeproctext]{ref-rmarkdown2018}
+Xie, Y., Allaire, J. J., \& Grolemund, G. (2018). \emph{R markdown: The
+definitive guide}. Chapman; Hall/CRC.
+\url{https://bookdown.org/yihui/rmarkdown}
+
+\bibitem[\citeproctext]{ref-rmarkdown2020}
+Xie, Y., Dervieux, C., \& Riederer, E. (2020). \emph{R markdown
+cookbook}. Chapman; Hall/CRC.
+\url{https://bookdown.org/yihui/rmarkdown-cookbook}
 
@@
 \[
-\text{TFCE} = \int_{h} e(h)^{E} h^{H} dh
+\text{TFCE} = \int_{h} e(h)^{E} h^{H} \mathrm{d}h
 \]
 
-Where\ldots{} Then, we can compute p-values for each timestep\ldots. we
-permute conditions, compute the TFCE score for the permuted set, and
-take the max(TFCE)\ldots{} But see Sassenhagen \& Draschkow
+Where\ldots{} the parameters \(E\) and \(H\) are set a priori and
+control the influence of the extend and height on the TFCE. Then,
+p-value for timestep/voxel \(i\) is computed by comparing it TFCE with
+the null distribution of TFCE values. For each permuted signal, we keep
+the maximal value over the whole signal for the null distribution of the
+TFCE\ldots. But see Sassenhagen \& Draschkow
 (\citeproc{ref-sassenhagen2019}{2019})\ldots{}
 
+\newpage
+
+\section{\texorpdfstring{Using the \texttt{R} package and integration
+with
+\texttt{MNE-Python}}{Using the R package and integration with MNE-Python}}\label{using-the-r-package-and-integration-with-mne-python}
+
+Explain how to use the \texttt{R} package and to integrate it with
+\texttt{MNE} epochs\ldots{}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# to{-}do adding some code here...}
+\end{Highlighting}
+\end{Shaded}
+
 \end{document}

### 2025-02-06 code/functions.R

@@ -1,7 +1,9 @@
-##############################################################################
-# # Functions from https://github.com/GRousselet/onsetsim/blob/main/code/functions.R
-# # Generate permutation distributions of t values for independent groups
-# ## cond1 and cond2 are Nt trials x Nf time frames matrices
+######################################################################################
+# Functions from https://github.com/GRousselet/onsetsim/blob/main/code/functions.R
+# Generate permutation distributions of t values for independent groups
+# cond1 and cond2 are Nt trials x Nf time frames matrices
+######################################################################################
+
 # permtdist <- function(cond1, cond2, Nt, Nf, nboot = 2000){
 #   
 #   st2 <- Nt+1 # index to start condition 2 
@@ -47,28 +49,32 @@
 #   }
 #   perm.tvals
 # }
-# 
-# # Cluster correction
-# 
-# ## Form clusters using binary vector: pvals < alpha
-# ## clusters must be at least 2 time frames
-# cluster.make <- function(x){
-#   y <- rle(x)
-#   cmap <- vector(mode = "numeric", length = 0)
-#   nC <- length(y$values) # number of clusters
-#   indx <- 0 # cluster counter
-#   for(CL in 1:nC){
-#     if(y$values[CL] == 0 || y$lengths[CL] == 1){
-#       val <- 0
-#     } else {
-#       indx <- indx + 1
-#       val <- indx
-#     }
-#     cmap <- c(cmap, rep(val, y$lengths[CL]))
-#   }
-#   cmap
-# }
-# 
+
+# Cluster correction
+# Form clusters using binary vector: pvals < alpha
+# clusters must be at least 2 time frames
+cluster.make <- function (x) {
+    
+    y <- rle(x)
+    cmap <- vector(mode = "numeric", length = 0)
+    nC <- length(y$values) # number of clusters
+    indx <- 0 # cluster counter
+    
+    for (CL in 1:nC) {
+        
+    if(y$values[CL] == 0 || y$lengths[CL] == 1){
+      val <- 0
+    } else {
+      indx <- indx + 1
+      val <- indx
+    }
+    cmap <- c(cmap, rep(val, y$lengths[CL]))
+    }
+    
+    cmap
+    
+}
+
 # ## Save sum for each cluster
 # # values = statistics (t values)
 # # cmap = output from cluster.make
[…]

### 2025-02-06 brms_meeg.qmd
minor cosmetic updates

@@
 
-Models were fitted using the `brms` package [@burkner_brms_2017; @burkner_advanced_2018; @nalborczyk2019]...
+Models were fitted using the `brms` package [@brms2017; @brms2018; @nalborczyk2019]...
 
@@
 Now, we want to *test* whether the group-level average decoding accuracy is above chance (i.e., 0.5) at each timestep.
-We use a similar GAM/GP as previously, but we replace the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one.
+We use a similar GAM/GP as previously, but we replace the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one to account for the bounded nature of AUC values (between 0 and 1).
 
@@
 # citing the R packages we have used
-cite_packages(output = "paragraph", omit = NULL, out.dir = ".")
+cite_packages(
+    output = "paragraph",
+    # omit = c("permuco", "brms", "reticulate"),
+    omit = c("permuco", "reticulate"),
+    out.dir = "."
+    )
 ```

### 2025-02-06 brms_meeg.tex
minor cosmetic updates

@@
 Models were fitted using the \texttt{brms} package
-(\citeproc{ref-burkner_brms_2017}{Bürkner, 2017a},
-\citeproc{ref-burkner_advanced_2018}{2018a};
+(\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018};
 \citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019})\ldots{}
@@
 GAM/GP as previously, but we replace the \(\mathrm{Normal}\) likelihood
-function by a \(\mathrm{Beta}\) one.
+function by a \(\mathrm{Beta}\) one to account for the bounded nature of
+AUC values (between 0 and 1).
 
@@
 following R packages: brms v. 2.22.0 (\citeproc{ref-brms2017}{Bürkner,
-2017b}, \citeproc{ref-brms2018}{2018b}, \citeproc{ref-brms2021}{2021}),
+2017}, \citeproc{ref-brms2018}{2018}, \citeproc{ref-brms2021}{2021}),
 changepoint v. 2.2.4 (\citeproc{ref-changepoint2022}{Killick et al.,
@@
 0.2.2 (\citeproc{ref-pakret}{Gallou, 2024}), patchwork v. 1.2.0
-(\citeproc{ref-patchwork}{T. L. Pedersen, 2024}), permuco v. 1.1.3
-(\citeproc{ref-permuco}{Frossard \& Renaud, 2021b}), rmarkdown v. 2.29
+(\citeproc{ref-patchwork}{T. L. Pedersen, 2024}), rmarkdown v. 2.29
 (\citeproc{ref-rmarkdown2024}{Allaire et al., 2024};
@@
 
-\bibitem[\citeproctext]{ref-burkner_brms_2017}
-Bürkner, P.-C. (2017a). \textbf{Brms} : {An} {{\emph{R}}} {Package} for
-{Bayesian Multilevel Models Using} {{\emph{Stan}}}. \emph{Journal of
-Statistical Software}, \emph{80}(1).
-\url{https://doi.org/10.18637/jss.v080.i01}
-
 \bibitem[\citeproctext]{ref-brms2017}
-Bürkner, P.-C. (2017b). {brms}: An {R} package for {Bayesian} multilevel
+Bürkner, P.-C. (2017). {brms}: An {R} package for {Bayesian} multilevel
 models using {Stan}. \emph{Journal of Statistical Software},
@@
 
-\bibitem[\citeproctext]{ref-burkner_advanced_2018}
-Bürkner, P.-C. (2018a). Advanced {Bayesian Multilevel Modeling} with the
-{R Package} brms. \emph{The R Journal}, \emph{10}(1), 395.
-\url{https://doi.org/10.32614/RJ-2018-017}
-
 \bibitem[\citeproctext]{ref-brms2018}
-Bürkner, P.-C. (2018b). Advanced {Bayesian} multilevel modeling with the
+Bürkner, P.-C. (2018). Advanced {Bayesian} multilevel modeling with the
 {R} package {brms}. \emph{The R Journal}, \emph{10}(1), 395--411.

### 2025-02-09 brms_meeg.qmd
including some more refs and using t2 splines for timegen data

@@
     # financial-support: "This study was supported by Grant [Grant Number] from [Funding Source]."
-abstract: "Time-resolved electrophysoiological measurements such as those offered by magneto- or electro-encephalography (M/EEG) provide a unique window onto neural activity underlying cognitive process and how they unfold over time.
+abstract: "Time-resolved electrophysiological measurements such as those offered by magneto- or electro-encephalography (M/EEG) provide a unique window onto neural activity underlying cognitive process and how they unfold over time.
 Typically, we are interested in testing whether such measures differ across conditions and/or groups.
@@
           \usepackage{mathtools}
-editor_options: 
-  chunk_output_type: console
 ---
@@
 library(grateful)
-library(permuco)
 library(scales)
@@
 
-Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]... Using Bayes factors [@teichmann2022]...
+Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]... Using Bayes factors [@teichmann2022]... Weighting single trials [@pernet2022]...
+
+See <https://www.let.rug.nl/nerbonne/teach/rema-stats-meth-seminar/presentations/Wieling-2014-GAMs.pdf> and <https://www.let.rug.nl/wieling/Statistics/GAM-EEG/lab/>, <https://www.let.rug.nl/wieling/Statistics/GAM-EEG/GAM-EEG.pdf> [@abugaber2023; @meulman2015]...
+
+Recent example of GLM for EEG [@wüllhorst2025; @fischer2013]...
+"Similar approaches have been successfully applied to EEG time- (Rousselet et al., 2008) and frequency-domain (Cohen and Cavanagh, 2011) data and allow the simultaneous investigation of multiple independent variables while preserving the high temporal resolution of the EEG."...
+See also [@rousselet2008; @hauk2006]...
+Example of two-stage regression analysis [individual-level then group-level, 10.1101/2024.04.10.588861, @dunagan2024]...
 
@@
 
+See also <https://jtimonen.github.io/lgpr-usage/articles/math.html>...
+
 Low-rank approximate Gaussian processes are of main interest in machine learning and statistics due to the high computational demands of exact Gaussian process models [@riutort-mayol_practical_2023]...
@@
 
+```{=html}
 <!--
@@
 -->
+```
 
@@
 As in @rousselet_using_2025, the signal was generated from truncated Gaussian with an objective onset at 160 ms, a peak at 250 ms, and an offset at 342 ms.
-We simulated this signal for 250 timesteps between 0 and 0.5s, akin to & 500 Hz sampling rate.
-We simulated such data for a group of 20 participants with 50 trials per participant and condition.
-
-```{r eeg-group, echo = FALSE, fig.width = 9, fig.cap = "Some ERPs in two conditions with 50 trials each, for a group of 20 participants."}
+We simulated this signal for 250 timesteps between 0 and 0.5s, akin to a 500 Hz sampling rate.
+We simulated such data for a group of 20 participants with 50 trials per participant and condition (@fig-eeg).
+
+```{r fig-eeg, echo = FALSE, fig.width = 9, fig.cap = "Some ERPs in two conditions with 50 trials each, for a group of 20 participants."}
 # importing R version of Matlab code from Yeung et al. (2004)
@@
 
+# removing raw_df
+if (exists("raw_df") ) rm(raw_df)
+
 for (P in 1:n_ppt) { # for each participant
@@
         trial = c(rep(1:n_trials, each = length(Xf) ), rep(1:n_trials, each = length(Xf) ) ),
-        condition = factor(rep(c("cond1", "cond2"), each = Nf*n_trials) ),
+        condition = factor(rep(c("cond1", "cond2"), each = Nf * n_trials) ),
         participant = paste0("participant_", sprintf("%02d", P) )
@@
 
+# converting time from ms to seconds
+raw_df <- raw_df %>%
+    # converting time from ms to seconds
+    mutate(x = x / 1000) %>%
+    # renaming columns
+    rename(time = x, eeg = y)
+
+# defining the true onset and offset in seconds
+true_onset <- 0.160
+true_offset <- 0.342
+
 # plotting the data
 raw_df %>%
-    summarise(y = mean(y), .by = c(participant, condition, x) ) %>%
-    ggplot(aes(x, y, colour = condition) ) +
+    summarise(eeg = mean(eeg), .by = c(participant, condition, time) ) %>%
+    ggplot(aes(x = time, y = eeg, colour = condition) ) +
     geom_hline(yintercept = 0, linetype = 2) +
@@
     facet_wrap(~participant) +
-    labs(x = "Time (ms)", y = "EEG signal (a.u.)")
-```
-
-```{r erp-diff, out.width = "75%", fig.cap = "Group-level average difference between conditions (mean +/- standard error of the mean).
+    labs(x = "Time (s)", y = "EEG signal (a.u.)")
+```
+
+We computed the average of the ERP difference (@fig-erp)...
+
+```{r fig-erp, echo = FALSE, out.width = "75%", fig.cap = "Group-level average difference between conditions (mean +/- standard error of the mean).
 The 'true' (population value of the) onset is indicated by the vertical dashed line."}
 # averaging across participants
-group_df <- raw_df %>%
-    pivot_wider(names_from = condition, values_from = y, values_fn = mean) %>%
-    mutate(y_diff = cond2 - cond1) %>%
+raw_df %>%
+    pivot_wider(names_from = condition, values_from = eeg, values_fn = mean) %>%
+    mutate(eeg_diff = cond2 - cond1) %>%
     summarise(
-        y_mean = mean(y_diff),
-        y_sem = sd(y_diff) / sqrt(n() ),
-        .by = x
-        ) %>%
-    mutate(lower = y_mean - y_sem, upper = y_mean + y_sem)
-
+        eeg_mean = mean(eeg_diff),
+        eeg_sem = sd(eeg_diff) / sqrt(n() ),
+        .by = time
+        ) %>%
+    mutate(lower = eeg_mean - eeg_sem, upper = eeg_mean + eeg_sem) %>%
     # plotting the data
-group_df %>%
-    ggplot(aes(x = x, y = y_mean) ) +
+    ggplot(aes(x = time, y = eeg_mean) ) +
     geom_hline(yintercept = 0, linetype = 2) +
@@
     geom_line(show.legend = FALSE) +
-    labs(x = "Time (ms)", y = "EEG difference (a.u.)")
+    labs(x = "Time (s)", y = "EEG difference (a.u.)")
 ```
@@
 
-Models were fitted using the `brms` package [@brms2017; @brms2018; @nalborczyk2019]...
+We then fitted a GAM and GP regression model using the `brms` package [@brms2017; @brms2018; @nalborczyk2019]...
 
 ```{r gam, message = FALSE}
+# averaging across participants
+ppt_df <- raw_df %>%
+    group_by(participant, condition, time) %>%
+    summarise(eeg = mean(eeg) ) %>%
+    # mutate(eeg = as.numeric(scale(x = eeg, scale = FALSE) ) ) %>%
+    ungroup()
+
 # defining a contrast for condition
-contrasts(group_df$condition) <- c(-0.5, 0.5)
+contrasts(ppt_df$condition) <- c(-0.5, 0.5)
 
@@
     # cubic regression splines with k-1 basis functions
-    # then we should try s(x, participant, bs = "fs")
-    # that is, random factor smooth interactions...
-    y ~ condition + s(x, bs = "cr", k = 20, by = condition),
-    data = group_df,
+    eeg ~ condition + s(time, bs = "cr", k = 10, by = condition),
+    data = ppt_df,
     family = gaussian(),
@@
 ```{r gp, message = FALSE}
+# computing eeg_diff and averaging across participants
+# group_df <- raw_df %>%
+#     group_by(participant, time) %>%
+#     pivot_wider(names_from = condition, values_from = eeg, values_fn = mean) %>%
+#     mutate(eeg_diff = cond2 - cond1) %>%
+#     summarise(eeg_diff = mean(eeg_diff) ) %>%
+#     ungroup()
+
+# see https://betanalpha.github.io/assets/case_studies/gp_part3/part3.html
+# and https://betanalpha.github.io/assets/case_studies/gaussian_processes.html
+gp_priors <- c(
+    set_prior("normal(0, 1)", class = "Intercept"),
+    set_prior("normal(0, 1)", class = "b"),
+    set_prior("exponential(1)", class = "sigma"),
+    set_prior("exponential(1)", class = "sdgp")
+    )
+
 gp_model <- brm(
@@
     # computing Hilbert-space approximate GPs
-    y ~ gp(x, k = 20, by = condition),
     # if k = NA (default), exact GPs are computed
-    # y ~ gp(x, by = condition),
-    data = group_df,
+    # eeg ~ gp(time, by = condition),
+    eeg ~ condition + gp(time, by = condition, k = 20, cov = "exp_quad"),
+    data = ppt_df,
     family = gaussian(),
-    control = list(adapt_delta = 0.99),
-    iter = 5000,
+    # prior = gp_priors,
+    control = list(adapt_delta = 0.99, max_treedepth = 15),
+    backend = "cmdstanr",
+    iter = 2000,
     chains = 4,
@@
 # defining a function to plot posterior predictions
-plot_post_preds <- function (model, data) {
+plot_post_preds <- function (model, data = NULL) {
+
+    if (is.null(data) ) data <- model$data
 
@@
         x = model,
-        effect = "x:condition",
+        effect = "time:condition",
         method = "posterior_epred",
@@
     post_preds %>%
-        ggplot(aes(x = x, y = y, colour = condition, fill = condition) ) +
+        ggplot(aes(x = time, y = eeg, colour = condition, fill = condition) ) +
+        geom_vline(xintercept = true_onset, linetype = 2) +
         geom_hline(yintercept = 0, linetype = 2) +
         geom_point(
-            data = data %>% summarise(y = mean(y), .by = c(x, condition) ),
+            data = data %>% summarise(eeg = mean(eeg), .by = c(time, condition) ),
             aes(fill = NULL),
@@
             ) +
-        labs(x = "Time (ms)", y = "EEG signal (a.u.)")
+        labs(x = "Time (s)", y = "EEG signal (a.u.)")
 
@@
 # plotting the posterior predictions
-plot_post_preds(model = gam, data = group_df) +
-    plot_post_preds(model = gp_model, data = group_df)
+plot_post_preds(model = gam) + plot_post_preds(model = gp_model)
 ```
@@
 # defining a function to plot posterior slope
-plot_post_slope <- function (model, data, eps = 0.1) {
-
-  # defining a sequence of x values to make predictions
-  x_seq <- crossing(
-      x = seq(min(data$x), max(data$x), length.out = 100),
+plot_post_slope <- function (model, data = NULL, eps = 0.1) {
+
+    # if no data is specified, use model$data
+    if (is.null(data) ) data <- model$data
+
+    # defining a sequence of time values to make predictions
+    time_seq <- crossing(
+        time = seq(min(data$time), max(data$time), length.out = 100),
         condition = c("cond1", "cond2")
@@
     # retrieving posterior samples
-  posterior_samples <- posterior_epred(object = model, newdata = x_seq)
+    posterior_samples <- posterior_epred(object = model, newdata = time_seq)
 
@@
     prob_y_above_0 <- data.frame(
-      x = seq(min(data$x), max(data$x), length.out = 100)
+        time = seq(min(data$time), max(data$time), length.out = 100)
         ) %>%
@@
     prob_y_above_0 %>%
-        ggplot(aes(x = x, y = m) ) +
+        ggplot(aes(x = time, y = m) ) +
         geom_hline(yintercept = 0, linetype = 2) +
+        geom_vline(xintercept = true_onset, linetype = 2) +
         geom_line() +
         geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
-        # theme_light(base_size = 12) +
-        labs(x = "Time (ms)", y = expression(Pr(beta~"|"~data, model) ) )
+        labs(x = "Time (s)", y = expression(Pr(beta~"|"~data, model) ) )
 
@@
 # using the function with model
-plot_post_slope(model = gam, data = group_df, eps = 0.1) +
-    plot_post_slope(model = gp_model, data = group_df, eps = 0.1)
+plot_post_slope(model = gam, eps = 0.1) +
+    plot_post_slope(model = gp_model, eps = 0.1)
 ```
@@
 # defining a function to plot posterior prob of slope above 0
-plot_post_test <- function (model, data, eps = 0.1) {
-
-    # defining a sequence of x values to make predictions
-    x_seq <- crossing(
-        x = seq(min(data$x), max(data$x), length.out = 100),
+plot_post_test <- function (model, data = NULL, eps = 0.1) {
+
+    # if no data is specified, use model$data
+    if (is.null(data) ) data <- model$data
+
+    # defining a sequence of time values to make predictions
+    time_seq <- crossing(
+        time = seq(min(data$time), max(data$time), length.out = 100),
         condition = c("cond1", "cond2")
@@
     # retrieving posterior samples
-    posterior_samples <- posterior_epred(object = model, newdata = x_seq)
+    posterior_samples <- posterior_epred(object = model, newdata = time_seq)
 
@@
     prob_y_above_0 <- data.frame(
-        x = x_seq$x,
+        time = time_seq$time,
         m = colMeans(posterior_samples > (0 + eps) ),
@@
     prob_y_above_0 %>%
-    ggplot(aes(x = x, y = m) ) +
+    ggplot(aes(x = time, y = m) ) +
     geom_hline(yintercept = 0.5, linetype = 2) +
+    geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = true_offset, linetype = 2) +
     geom_line() +
-    # geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
-    # theme_light(base_size = 12) +
-    # labs(x = "Time (ms)", y = "Pr(EEG>0)|data)")
-    labs(x = "Time (ms)", y = expression(Pr(beta>0~"|"~data, model) ) )
+    labs(x = "Time (s)", y = expression(Pr(beta>0~"|"~data, model) ) )
 
@@
 # using the function with the two models
-plot_post_test(model = gam, data = group_df, eps = 0.1) +
-    plot_post_test(model = gp_model, data = group_df, eps = 0.1)
+plot_post_test(model = gam, eps = 0.1) +
+    plot_post_test(model = gp_model, eps = 0.1)
 ```
@@
 We can also express this as the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional BF thresholds (@fig-post-prob-ratio).
+Also check this paper (<https://link.springer.com/article/10.3758/s13428-021-01613-6?fromPaywallRec=false>)?
 
@@
 n_approx <- 1000
-x_seq <- crossing(
-    x = seq(min(group_df$x), max(group_df$x), length.out = n_approx),
+time_seq <- crossing(
+    time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx),
     condition = c("cond1", "cond2")
@@
 # retrieving posterior samples
-posterior_samples <- posterior_epred(object = gam, newdata = x_seq)
+posterior_samples <- posterior_epred(object = gam, newdata = time_seq)
 
@@
 
-# sanity checks
-# ts.plot(colMeans(posterior_samples) )
-# ts.plot(apply(X = posterior_samples, MARGIN = 2, FUN = mean) )
-# ts.plot(apply(X = posterior_samples, MARGIN = 2, FUN = sd) )
-# mean(apply(X = posterior_samples, MARGIN = 2, FUN = sd) )
-
 # computing the probability that y is above 0
@@
 prob_y_above_0 <- data.frame(
-    x = seq(min(group_df$x), max(group_df$x), length.out = n_approx)
+    time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx)
     ) %>%
@@
     dplyr::filter(prob_ratio > threshold) %>%
-    summarise(cluster_onset = min(x), cluster_offset = max(x) ) %>%
+    summarise(cluster_onset = min(time), cluster_offset = max(time) ) %>%
     mutate(
@@
             dplyr::filter(post_prob == max(post_prob) ) %>%
-            pull(x)
-        ) %>%
-    mutate(true_onset = 160, true_peak = 250, true_offset = 342) %>%
+            pull(time)
+        ) %>%
+    mutate(true_onset = 0.160, true_peak = 0.250, true_offset = 0.342) %>%
     mutate(model = "GAM")
@@
     mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
-    ggplot(aes(x = x, y = prob_ratio) ) +
+    ggplot(aes(x = time, y = prob_ratio) ) +
     geom_hline(yintercept = 1, linetype = "dashed") +
@@
     geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
+    geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = true_offset, linetype = 2) +
     geom_line() +
@@
     labs(
-        x = "Time (ms)",
-        y = expression(Pr(beta>0~"|"~data, model) / (1 - Pr(beta>0~"|"~data) ) )
+        x = "Time (s)",
+        y = expression(Pr(beta>0~"|"~data) / (1 - Pr(beta>0~"|"~data) ) )
         )
@@
 # retrieving posterior samples
-posterior_samples <- posterior_epred(object = gp_model, newdata = x_seq)
+posterior_samples <- posterior_epred(object = gp_model, newdata = time_seq)
 
@@
 # computing the probability that y is above 0
-eps <- 0.05
 prob_y_above_0 <- data.frame(
-    x = seq(min(group_df$x), max(group_df$x), length.out = n_approx)
+    time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx)
     ) %>%
@@
 # the signal was a truncated Gaussian defining an objective onset at 160 ms,
-# a maximum at 250 ms, and an offset at 342 ms.
-threshold <- 10
+# a maximum at 250 ms, and an offset at 342 ms
 exceeding_times_gp <- prob_y_above_0 %>%
     dplyr::filter(prob_ratio > threshold) %>%
-    summarise(cluster_onset = min(x), cluster_offset = max(x) ) %>%
+    summarise(cluster_onset = min(time), cluster_offset = max(time) ) %>%
     mutate(
@@
             dplyr::filter(post_prob == max(post_prob) ) %>%
-            pull(x)
-        ) %>%
-    mutate(true_onset = 160, true_peak = 250, true_offset = 342) %>%
+            pull(time)
+        ) %>%
+    mutate(true_onset = 0.160, true_peak = 0.250, true_offset = 0.342) %>%
     mutate(model = "GP")
@@
 
-Below we are plotting the error function (where the error is computed as $|\hat{\theta}-\theta|$) for both the onset and offset values of the ERP difference, according to various `eps` and `threshold` values.
+Below we are plotting the error function (where the error is computed as $\text{error}:=|\hat{\theta}-\theta|$) for both the onset and offset values of the ERP difference, according to various `eps` and `threshold` values.
 Remember that the signal is generated from a truncated Gaussian defining an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
@@
 threshold_values <- seq(from = 1, to = 50, by = 1)
-n_approx <- n_distinct(group_df$x)
+
+# using as many values of time as in the original data
+n_approx <- n_distinct(ppt_df$time)
 
@@
 
-# generating a grid of values to compute predictions for
-x_seq <- crossing(
-    x = seq(min(group_df$x), max(group_df$x), length.out = n_approx*2-1),
-    condition = c("cond1", "cond2")
+# generating a grid of time values to compute predictions for
+time_seq <- crossing(
+    # participant = unique(ppt_df$participant),
+    condition = c("cond1", "cond2"),
+    time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx*2-1),
     ) %>%
@@
 # retrieving posterior samples
-posterior_samples <- posterior_epred(object = gp_model, newdata = x_seq)
+# posterior_samples <- posterior_epred(object = gp_model, newdata = time_seq)
+posterior_samples <- posterior_epred(object = gam, newdata = time_seq)
 
@@
     prob_y_above_0 <- data.frame(
-        x = seq(min(group_df$x), max(group_df$x), length.out = n_approx*2-1)
+        time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx*2-1)
         ) %>%
@@
         summarise(
-            cluster_onset = min(x, na.rm = TRUE),
-            cluster_offset = max(x, na.rm = TRUE)
+            cluster_onset = min(time, na.rm = TRUE),
+            cluster_offset = max(time, na.rm = TRUE)
             ) %>%
@@
             dplyr::filter(post_prob == max(post_prob, na.rm = TRUE) ) %>%
-            pull(x)
-            ) %>%
-        mutate(true_onset = 160, true_peak = 250, true_offset = 342)
+            pull(time)
+            ) %>%
+        mutate(true_onset = 0.160, true_peak = 0.250, true_offset = 0.342)
 
@@
     scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-    # scale_fill_viridis_c() +
-    # scale_y_log10(labels = label_log(digits = 2) ) +
-    # theme_light(base_size = 10) +
     labs(
@@
     scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-    # theme_light(base_size = 10) +
     labs(
@@
 # displaying best parameters values (according to onset error)
-# library(flextable)
 results %>%
@@
 # displaying best parameters values (according to offset error)
-# library(flextable)
 results %>%
@@
     head()
+
+# displaying best parameters values (according to both onset/offset error)
+results %>%
+    filter(error_onset == min(error_onset) & error_offset == min(error_offset) ) %>%
+    data.frame()
 ```
@@
     summarise(
-        eeg = mean(y),
-        eeg_sd = sd(y),
-        .by = c(participant, condition, x)
+        eeg_mean = mean(eeg),
+        eeg_sd = sd(eeg),
+        .by = c(participant, condition, time)
         )
@@
     # using by-participant SD of ERPs across trials
-    eeg | se(eeg_sd) ~
-        condition + s(x, bs = "cr", k = 10, by = condition) +
+    eeg_mean | se(eeg_sd) ~
+        condition + s(time, bs = "cr", k = 10, by = condition) +
         (1 | participant),
@@
 
-```{r meta-gam-preds, fig.cap = "Hierarchical GAM posterior predictions."}
-# plotting the posterior predictions
-plot(
-    conditional_effects(x = meta_gam, effect = "x:condition"),
-    points = FALSE, theme = theme_light(), plot = FALSE
-    )[[1]] +
-    labs(x = "Time (ms)", y = "EEG signal (a.u.)")
-```
-
-```{r meta-gp, eval = FALSE, message = FALSE}
+```{r meta-gp, message = FALSE}
 # fitting the GP
@@
     # using by-participant SD of ERPs across trials
-    eeg | se(eeg_sd) ~
-        condition + gp(x, k = 20, by = condition) +
+    eeg_mean | se(eeg_sd) ~
+        condition + gp(time, k = 20, by = condition) +
         (1 | participant),
@@
     family = gaussian(),
-    control = list(adapt_delta = 0.95, max_treedepth = 20),
+    control = list(adapt_delta = 0.99),
     iter = 5000,
@@
 
-```{r meta-gp-preds, eval = FALSE, fig.cap = "Hierarchical GP posterior predictions."}
+```{r meta-gp-preds, fig.width = 9, fig.asp = 0.5, fig.cap = "Posterior predictions for the hierarchical GAM and GP."}
 # plotting the posterior predictions
-plot(
-    conditional_effects(x = meta_gp, effect = "x:condition"),
+meta_gam_preds <- plot(
+    conditional_effects(x = meta_gam, effect = "time:condition"),
     points = FALSE, theme = theme_light(), plot = FALSE
     )[[1]] +
-    labs(x = "Time (ms)", y = "EEG signal (a.u.)")
-```
-
-## Application to 1D decoding results (accuracy over time)
-
-Assume we have M/EEG data and we conducted time-resolved multivariate pattern analysis (MVPA), also known as decoding.
-As a result, we have a timecourse of decoding accuracies (e.g., ROC AUC), bounded between 0 and 1, per participant (@fig-sim-decoding).
-
-```{r fig-sim-decoding, echo = FALSE, out.width = "75%", fig.cap = "Exemplary (simulated) group-level average timecourse of binary decoding accuracy (ROC AUC)."}
+    labs(x = "Time (s)", y = "EEG signal (a.u.)")
+meta_gp_preds <- plot(
+    conditional_effects(x = meta_gp, effect = "time:condition"),
+    points = FALSE, theme = theme_light(), plot = FALSE
+    )[[1]] +
+    labs(x = "Time (s)", y = "EEG signal (a.u.)")
+
+# combining the two plots
+meta_gam_preds + meta_gp_preds
+```
+
+## Application to time-resolved decoding results (accuracy over time)
+
+We conducted time-resolved multivariate pattern analysis (MVPA), also known as decoding.
+As a result, we have a timecourse of decoding accuracies (e.g., ROC AUC), bounded between 0 and 1, per participant (@fig-mne-decoding)...
+
+```{r fig-sim-decoding, echo = FALSE, eval = TRUE, out.width = "75%", fig.cap = "Exemplary (simulated) group-level average timecourse of binary decoding accuracy (ROC AUC)."}
 # simulating decoding accuracies (ROC AUC)
@@
 # plotting it
-group_data %>%
-    ggplot(aes(x = time, y = mean_auc) ) +
-    geom_hline(yintercept = 0.5, linetype = "dashed") +
-    geom_vline(xintercept = 0.0, linetype = "dashed") +
-    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "steelblue", alpha = 0.2) +
-    geom_line(color = "steelblue", size = 1) +
-    ylim(0, 1) +
-    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
-```
-
-```{r mne-decoding, out.width = "75%", eval = FALSE}
+# group_data %>%
+#     ggplot(aes(x = time, y = mean_auc) ) +
+#     geom_hline(yintercept = 0.5, linetype = "dashed") +
+#     geom_vline(xintercept = 0.0, linetype = "dashed") +
+#     geom_ribbon(aes(ymin = lower, ymax = upper), fill = "steelblue", alpha = 0.2) +
+#     geom_line(color = "steelblue", size = 1) +
+#     ylim(0, 1) +
+#     labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
+```
+
+```{r mne-decoding, eval = FALSE, echo = FALSE}
 # loading the reticulate package
@@
 # defining the function in R (it will be executed in Python)
-mne_decoding <- function (X, labels, ncores = 1) {
+mne_decoding <- function (X, labels, ncores = 8) {
 
@@
         select(-participant) %>%
-        pivot_wider(names_from = x, values_from = y) %>%
+        pivot_wider(names_from = time, values_from = eeg) %>%
         select(-condition, -trial)
@@
         select(-participant) %>%
-        pivot_wider(names_from = x, values_from = y) %>%
+        pivot_wider(names_from = time, values_from = eeg) %>%
         pull(condition) %>%
@@
         select(-participant) %>%
-        pull(x) %>%
+        pull(time) %>%
         unique()
@@
 
-```{r, out.width = "75%", fig.cap = "Exemplary group-level average timecourse of binary decoding accuracy (ROC AUC). Decoding was performed using MNE-Python."}
+```{r fig-mne-decoding, echo = FALSE, fig.width = 9, fig.cap = "Exemplary average timecourse of binary decoding accuracy (ROC AUC) for each participant.
+Group-level average decoding accuracy is depicted as a grey background density in each panel."}
 # importing the decoding scores
@@
 timesteps <- raw_df %>%
-    # filter(participant == ppt) %>%
-    # select(-participant) %>%
-    pull(x) %>%
+    pull(time) %>%
     unique()
@@
 # plotting it
-group_decoding_scores %>%
+decoding_summary <- group_decoding_scores %>%
     t() %>%
@@
     mutate(time = timesteps) %>%
-    ggplot(aes(x = time, y = mean) ) +
+    mutate(
+        ymin = pmin(mean, 0.5),
+        ymax = pmax(mean, 0.5)
+        )
+
+# reshaping the decoding data
+decoding_data <- group_decoding_scores %>%
+    t() %>%
+    data.frame() %>%
+    mutate(time = timesteps) %>%
+    pivot_longer(cols = 1:20) %>%
+    mutate(
+        participant = rep(
+            unique(raw_df$participant),
+            n_distinct(raw_df$time)
+            )
+        ) %>%
+    select(participant, time, auc = value)
+
+# plotting it
+# group_decoding_scores %>%
+#     t() %>%
+#     data.frame() %>%
+#     mutate(time = timesteps) %>%
+#     pivot_longer(cols = 1:20) %>%
+#     mutate(
+#         participant = rep(
+#             unique(raw_df$participant),
+#             n_distinct(raw_df$time)
+#             )
+#         ) %>%
+decoding_data %>%
+    ggplot(aes(x = time) ) +
     geom_hline(yintercept = 0.5, linetype = "dashed") +
-    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "steelblue", alpha = 0.2) +
-    geom_line(color = "steelblue", size = 0.8) +
+    geom_vline(xintercept = true_onset, linetype = "dashed") +
+    geom_line(aes(y = auc), linewidth = 0.5) +
+    facet_wrap(~participant) +
+    geom_ribbon(
+        data = decoding_summary,
+        aes(x = time, ymin = ymin, ymax = ymax),
+        alpha = 0.5
+        ) +
     labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
@@
     family = Beta(),
-    control = list(adapt_delta = 0.9),
+    control = list(adapt_delta = 0.99),
     iter = 5000,
@@
     geom_hline(yintercept = 0.5, linetype = 2) +
-    geom_vline(xintercept = 0.0, linetype = 2) +
+    geom_vline(xintercept = true_onset, linetype = 2) +
     ylim(0, 1) +
@@
 
-    # defining a sequence of x values to make predictions
-    x_seq <- crossing(time = seq(min(data$time), max(data$time), length.out = 100) )
+    # defining a sequence of time values to make predictions
+    time_seq <- crossing(time = seq(min(data$time), max(data$time), length.out = 100) )
 
     # retrieving posterior samples
-    posterior_samples <- posterior_epred(object = model, newdata = x_seq)
+    posterior_samples <- posterior_epred(object = model, newdata = time_seq)
 
@@
     prob_y_above_0 <- data.frame(
-        time = x_seq$time,
+        time = time_seq$time,
         m = colMeans(posterior_samples > (chance_level + eps) ),
@@
     geom_hline(yintercept = chance_level, linetype = 2) +
-    geom_vline(xintercept = 0, linetype = 2) +
+    geom_vline(xintercept = true_onset, linetype = 2) +
     geom_line() +
-    # geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
     labs(x = "Time (s)", y = expression(Pr(AUC>chance~"|"~data, model) ) )
@@
 
-    # defining a sequence of x values to make predictions
-    # x_seq <- crossing(time = seq(min(data$time), max(data$time), length.out = 100) )
-    x_seq <- data.frame(time = data$time)
+    # defining a sequence of time values to make predictions
+    time_seq <- data.frame(time = data$time)
 
     # retrieving posterior samples
-    posterior_samples <- posterior_epred(object = model, newdata = x_seq)
+    posterior_samples <- posterior_epred(object = model, newdata = time_seq)
 
@@
         summarise(cluster_onset = min(time), cluster_offset = max(time) )
-    print(exceeding_times)
+
+    # printing the identified clusters
+    # print(exceeding_times)
 
@@
         geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
-        geom_vline(xintercept = 0, linetype = 2) +
+        geom_vline(xintercept = true_onset, linetype = 2) +
         geom_line() +
@@
             y = expression(
-                Pr(AUC>chance~"|"~data, model) / (1 - Pr(AUC>chance~"|"~data, model) )
+                Pr(AUC>chance~"|"~data) / (1 - Pr(AUC>chance~"|"~data) )
                 )
@@
 
-\newpage
-
-## Application to 2D decoding results (cross-temporal generalisation)
-
-Assume we have M/EEG data and we have conducted cross-temporal generalisation analyses [@king2014].
-As a result, we have a 2D matrix where each element contains the decoding accuracy (e.g., ROC AUC) of a classifier trained at timestep $\text{training}_{i}$ and tested at timestep $\text{testing}_{j}$ (@fig-sim-timegen).
-
-```{r fig-sim-timegen, echo = FALSE, dpi = 300, fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
-# number of participants
-n_participants <- 10
-
-# number of timepoints
-# n_timepoints <- 240 # (1.2s at 200Hz)
-n_timepoints <- 60 # (1.2s at 50Hz)
-timepoints <- seq(from = -0.2, to = 1, length.out = n_timepoints)
-
-# smooth onset and offset parameters
-decode_onset <- 0.2
-decode_offset <- 0.3
-
-# smoothness of onset/offset
-sigma_smooth <- 0.05
-
-# generating participant-specific cross-temporal generalisation matrices
-timegen_data <- tibble()
-
-# for each participant
-for (p in 1:n_participants) {
-
-    # adding variability in peak amplitude and width per participant
-    peak_variability <- runif(1, 30, 50) # random peak strength (higher = stronger decoding)
-    smooth_variability <- runif(1, 0.04, 0.06) # random smoothness variation
-    onset_variability <- runif(1, -0.05, 0.05) # small jitter in onset time
-    offset_variability <- runif(1, -0.05, 0.05) # small jitter in offset time
-
-    # computing smooth transition function (soft onset and offset with participant variation)
-    decode_window <- outer(timepoints, timepoints, function (t1, t2) {
-
-        # smooth onset
-        onset_factor <- 1 /
-            (1 + exp(-(t1 - (decode_onset + onset_variability) ) / sigma_smooth) )
-
-        # smooth offset
-        offset_factor <- 1 /
-            (1 + exp((t1 - (decode_offset + offset_variability)) / sigma_smooth) )
-
-        # diagonal Gaussian
-        onset_factor * offset_factor * exp(-((t1 - t2)^2) / (2 * smooth_variability^2) )
-
-    })
-
-    # defining smooth alpha and beta parameters for Beta distribution
-    base_alpha <- 10
-    peak_alpha <- peak_variability
-    alpha_matrix <- base_alpha + peak_alpha * decode_window
-    beta_matrix <- base_alpha
-
-    # generating AUC values using the Beta distribution
-    auc_matrix_beta <- matrix(
-        rbeta(n_timepoints^2, shape1 = alpha_matrix, shape2 = beta_matrix),
-        nrow = n_timepoints, ncol = n_timepoints
-        )
-
-    # converting to long format for plotting
-    temp_data <- expand.grid(
-        train_time = timepoints, test_time = timepoints
-        ) %>%
-        mutate(auc = as.vector(auc_matrix_beta), participant = p)
-
-    # storing results
-    timegen_data <- bind_rows(timegen_data, temp_data)
-
-}
-
-# plotting cross-temporal generalization matrix
-timegen_data %>%
-    summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>%
-    ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
-    geom_tile() +
-    geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
-    geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
-    geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
-    scale_x_continuous(expand = c(0, 0) ) +
-    scale_y_continuous(expand = c(0, 0) ) + 
-    coord_fixed() +
-    scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
-    labs(x = "Testing time (s)", y = "Training time (s)")
-```
-
-Now, we want to test whether and when decoding performance is above chance level (0.5 for a binary decoding task).
-These two models are computationally heavier to fit (more observations and 2D smooth functions)...
-
-```{r gam-gp-timegen, eval = FALSE, message = FALSE}
-# fitting a GAM with two temporal dimensions
-timegen_gam <- brm(
-    # 2D thin-plate spline (tp) to model smooth interactions between training and testing time
-    # auc ~ s(train_time, test_time, bs = "tp", k = 10),
-    auc ~ t2(train_time, test_time, bs = "tp", k = 10),
-    data = timegen_data,
-    family = Beta(),
-    iter = 2000,
-    chains = 4,
-    cores = 4,
-    file = "models/timegen_gam_t2.rds"
-    )
-
-# fitting a GP with two temporal dimensions
-timegen_gp <- brm(
-    auc ~ gp(train_time, test_time, k = 20),
-    # data = timegen_data %>% mutate(auc_c = scale(x = auc, center = TRUE, scale = FALSE) ),
-    data = timegen_data,
-    family = Beta(),
-    control = list(adapt_delta = 0.95, max_treedepth = 20),
-    iter = 2000,
-    chains = 4,
-    cores = 4,
-    file = "models/timegen_gp.rds"
-    )
-```
-
-```{r gam-timegen-preds, eval = FALSE, echo = FALSE, fig.cap = "Posterior predictions of the GAM fitted on decoding accuracy over time."}
-# plotting the posterior predictions
-plot(
-    conditional_effects(x = timegen_gam, prob = 0.95),
-    points = FALSE,
-    # point_args = list(size = 1, pch = 21, colour = "grey"),
-    # line_args = list(colour = "black"),
-    method = "posterior_epred",
-    surface = TRUE,
-    theme = theme_light(),
-    plot = TRUE
-    )#[[1]] +
-    # geom_hline(yintercept = 0.5, linetype = 2) +
-    # geom_vline(xintercept = 0.0, linetype = 2) +
-    # ylim(0, 1) +
-    # labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
-
-# posterior predictive checks
-pp_check(timegen_gam)
-```
-
-```{r gam-timegen-post-preds, echo = FALSE, fig.width = 8, fig.asp = 0.6, fig.cap = "Posterior probability of decoding accuracy being above chance level (2D GAM)."}
-# defining a function to plot posterior prob of slope above 0
-plot_post_timegen <- function (model, data, chance_level = 0.5, eps = 0.1) {
-
-    # defining a sequence of x values to make predictions
-    preds_conds <- crossing(
-        train_time = unique(data$train_time)[c(TRUE, FALSE)],
-        test_time = unique(data$test_time)[c(TRUE, FALSE)]
-        )
-
-    # retrieving posterior predictions
-    post_preds <- conditional_effects(
-        x = model,
-        conditions = preds_conds,
-        method = "posterior_epred",
-        prob = 0.95,
-        ndraws = 100
-        )[[1]]
-
-    # plotting cross-temporal generalization matrix
-    post_preds_plot <- post_preds %>%
-        summarise(auc = mean(estimate__), .by = c(train_time, test_time) ) %>%
-        ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
-        geom_tile() +
-        geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
-        geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
-        geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
-        scale_x_continuous(expand = c(0, 0) ) +
-        scale_y_continuous(expand = c(0, 0) ) + 
-        coord_fixed() +
-        scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
-        # theme_light() +
-        labs(x = "Testing time (s)", y = "Training time (s)")
-
-    # retrieving posterior samples
-    posterior_samples <- posterior_epred(object = model, newdata = preds_conds)
-
-    # computing the probability that y is above 0
-    prob_y_above_0 <- preds_conds %>%
-        mutate(post_prob = colMeans(posterior_samples) ) %>%
-        mutate(m = colMeans(posterior_samples > (chance_level + eps) ) ) %>%
-        mutate(prob_ratio = m / (1 - m) )
-
-    # plotting it
-    # prob_y_above_0 %>% filter(prob_ratio != Inf) %>% pull(prob_ratio) %>% range()
-    prob_above_chance_plot <- prob_y_above_0 %>%
-        mutate(prob_ratio = ifelse(prob_ratio == Inf, 2000, prob_ratio) ) %>%
-        mutate(prob_ratio = ifelse(prob_ratio == 0, 0.0005, prob_ratio) ) %>%
-        ggplot(aes(x = train_time, y = test_time, fill = prob_ratio) ) +
-        geom_tile() +
-        geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
-        geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
-        geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
-        scale_x_continuous(expand = c(0, 0) ) +
-        scale_y_continuous(expand = c(0, 0) ) + 
-        coord_fixed() +
-        scale_fill_scico(palette = "vik", midpoint = 0, name = "p/(1-p)", trans = "log10") +
-        # scale_fill_scico(palette = "vik", midpoint = 1, name = "p/(1-p)") +
-        # theme_light() +
-        labs(x = "Testing time (s)", y = "Training time (s)")
-
-    # combining the two plots
-    post_preds_plot + prob_above_chance_plot
-
-    # some tests
-    # post_preds_plot +
-    #     geom_contour(
-    #         data = prob_y_above_0,
-    #         aes(x = train_time, y = test_time, z = prob_ratio),
-    #         breaks = 10,
-    #         color = "red",
-    #         linewidth = 1,
-    #         inherit.aes = FALSE
-    #         )
-
-}
-
-# using the function with the decoding GAM
-plot_post_timegen(model = timegen_gam, data = timegen_data, chance_level = 0.5, eps = 0.01)
-```
-
-```{r gp-timegen-preds, eval = FALSE, echo = FALSE, fig.cap = "Posterior predictions of the GP fitted on decoding accuracy over time."}
-# plotting the posterior predictions
-plot(
-    conditional_effects(x = timegen_gp, prob = 0.95),
-    points = FALSE,
-    method = "posterior_epred",
-    surface = TRUE,
-    theme = theme_light(),
-    plot = TRUE
-    )
-
-# posterior predictive checks
-pp_check(timegen_gp)
-
-# checking timecourse of timegen_data's diagonal
-timegen_data %>%
-    filter(train_time == test_time) %>%
-    summarise(auc = mean(auc), .by = train_time) %>%
-    ggplot(aes(x = train_time, y = auc) ) +
-    geom_hline(yintercept = 0.5, linetype = 2) +
-    geom_line() +
-    labs(x = "Time (s)", y = "AUC")
-```
-
 ## Application to actual M/EEG data
@@
 -   uncorrected univariate tests
-- univariate tests + Bonferroni
+-   univariate tests + Bonferroni/Holm
 -   univariate tests + BH95 (Benjamini-Hochberg, 1995) aka FDR
@@
 
-We used `r pkrt("permuco")` and `r pkrt("changepoint")`...
+We used `r pkrt("changepoint")`...
 
@@
 
-```{r test-1d, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Timecourse of squared t-values and s-values (continuous measure of evidence given by -log2(p-value)) with true onset and onset identified using the raw (uncorrected) p-values or the corrected p-values (BH, BY, Bonferroni, or Holm)."}
+S-values [@greenland2019]...
+
+```{r test-1d, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Timecourse of squared t-values and s-values (continuous measure of evidence given by -log2(p-value)) with true onset and onset identified using the raw (uncorrected) p-values or the corrected p-values (BH, BY, Holm)."}
 # https://github.com/GRousselet/onsetsim/blob/main/docs/weakstrongfwer_demo.md
@@
 ppt_data <- raw_df %>%
-    summarise(y = mean(y), .by = c(participant, condition, x) ) %>%
-    pivot_wider(names_from = condition, values_from = y) %>%
-    mutate(y_diff = cond2 - cond1)
+    summarise(eeg = mean(eeg), .by = c(participant, condition, time) ) %>%
+    pivot_wider(names_from = condition, values_from = eeg) %>%
+    mutate(eeg_diff = cond2 - cond1)
 
@@
 tests_results <- ppt_data %>%
-    group_by(x) %>%
+    group_by(time) %>%
     summarise(
-        # tval = t.test(x = auc_centered, mu = 0)$statistic^2,
-        # pval = t.test(x = auc_centered, mu = 0)$p.value
-        tval = t.test(x = y_diff, mu = 0)$statistic^2,
-        pval = t.test(x = y_diff, mu = 0)$p.value
+        tval = t.test(x = eeg_diff, mu = 0)$statistic^2,
+        pval = t.test(x = eeg_diff, mu = 0)$p.value
         ) %>%
@@
         pval_by = p.adjust(p = pval, method = "BY"),
-        pval_bonf = p.adjust(p = pval, method = "bonferroni"),
+        # pval_bonf = p.adjust(p = pval, method = "bonferroni"),
         pval_holm = p.adjust(p = pval, method = "holm")
@@
 # finding onsets
-from_timestep <- 20
+from_timestep <- 1
 onset_p <- find_onset(
     tests_results$pval[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$x[from_timestep:nrow(tests_results)]
+    tests_results$time[from_timestep:nrow(tests_results)]
     )
@@
     tests_results$pval_bh[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$x[from_timestep:nrow(tests_results)]
+    tests_results$time[from_timestep:nrow(tests_results)]
     )
@@
     tests_results$pval_by[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$x[from_timestep:nrow(tests_results)]
-    )
-onset_bonf <- find_onset(
-    tests_results$pval_bonf[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$x[from_timestep:nrow(tests_results)]
-    )
+    tests_results$time[from_timestep:nrow(tests_results)]
+    )
+# onset_bonf <- find_onset(
+#     tests_results$pval_bonf[from_timestep:nrow(tests_results)] <= aath,
+#     tests_results$time[from_timestep:nrow(tests_results)]
+#     )
 onset_holm <- find_onset(
     tests_results$pval_holm[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$x[from_timestep:nrow(tests_results)]
+    tests_results$time[from_timestep:nrow(tests_results)]
     )
@@
 # plotting the results
-default_colours <- hue_pal()(5)
-true_onset <- 160
-true_offset <- 342
+default_colours <- hue_pal()(10)
 tvals_plot <- tests_results %>%
-    ggplot(aes(x = x, y = tval) ) +
+    ggplot(aes(x = time, y = tval) ) +
     geom_area(position = "identity") +
@@
     geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[4]) +
-    geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
+    # geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
     geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[5]) +
-    labs(x = "Time (ms)", y = bquote(t^2) )
+    labs(x = "Time (s)", y = bquote(t^2) )
 
 plot_title <- paste0(
-    "True onset: ", true_onset, "ms, ",
-    "Uncorrected onset: ", round(onset_p, 3), "ms, ",
-    "BH onset: ", round(onset_bh, 3), "ms, ",
-    "BY onset: ", round(onset_by, 3), "ms,\n",
-    "Bonf onset: ", round(onset_bonf, 3), "ms, ",
-    "Holm onset: ", round(onset_holm, 3), "ms"
+    "True onset: ", true_onset, "s, ",
+    "Uncorrected onset: ", round(onset_p, 3), "s, ",
+    "BH onset: ", round(onset_bh, 3), "s, ",
+    "BY onset: ", round(onset_by, 3), "s, ",
+    # "Bonf onset: ", round(onset_bonf, 3), "s, ",
+    "Holm onset: ", round(onset_holm, 3),  "s"
     )
@@
     mutate(sval = -log2(value) ) %>%
-    ggplot(aes(x = x, y = sval, colour = name) ) +
+    ggplot(aes(x = time, y = sval, colour = name) ) +
     geom_line(linewidth = 0.5) +
     geom_hline(yintercept = -log2(aath), linetype = 2) +
+    geom_vline(xintercept = true_onset, linetype = 2) +
     geom_vline(xintercept = onset_p, linetype = 2, color = default_colours[1]) +
@@
     geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[4]) +
-    geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
+    # geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
     geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[5]) +
     labs(
-        x = "Time (ms)",
+        x = "Time (s)",
         y = "-log2(p-value)",
@@
             "True onset: ", true_onset, "s, ", "Change point onset: ",
-            round(tests_results$time[res@cpts[1]], 3), "s"
+            round(tests_results$time[res@cpts[1]], 3), "s, ",
+            "True offset: ", true_offset, "s, ", "Change point offset: ",
+            round(tests_results$time[res@cpts[2]], 3), "s"
             )
@@
     geom_hline(yintercept = -log2(0.05), linetype = 2) +
-    # geom_area(position = "identity") +
     labs(
-        x = "Time (ms)",
+        x = "Time (s)",
         y = "-log2(p-value)"
@@
 
-...
+Extending the approach to spatiotemporal data (i.e., time + sensors)...
 
@@
 
-````{=html}
-<!--
-
-# Generalised additive models
-
-In brief, a GAM is a GLM that includes a sum of smooth functions of predictor(s)... The general formula for a GAM is:
-
-$$
-g(\mathbb{E}[Y]) = \alpha + \sum_{j=1}^p s_j\left(X_j\right)
-$$
-
-where:
-
-- $Y$ is the response variable,
-- $g()$ is the link function that relates the expected value of $Y(E(Y)$ to the linear predictor.
-- $\alpha$ is the intercept term.
-- $s_j\left(X_j\right)$ are smooth functions, such as splines, applied to the predictor variable $X_j$.
-- $p$ is the number of predictor variables.
-
-Explain GAMs with code and figures here...
-
-```{r gam-tutorial, echo = TRUE, eval = FALSE}
-# see https://gavinsimpson.github.io/gratia/articles/posterior-simulation.html
-# see https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10b-GAMs.html
-# and https://m-clark.github.io/generalized-additive-models/technical.html
-# create a set of /k/ splines for vector /x/
-get_basis <- function(x, k = 20){
-  # fit a dummy GAM
-  fit_gam <- mgcv::gam(
-    formula = y ~ s(times, bs = "cr", k = 20),
-    data    = tibble(y = 1, times = x))
-  # extract model matrix
-  model_matrix <- model.matrix(fit_gam)
-  # wrangle to long tibble
-  some_curves <- model_matrix |> 
-    as_tibble() |> 
-    mutate(x = x) |> 
-    pivot_longer(cols = -x) |> 
-    mutate(
-      name = str_replace(name, "s\\(times\\).", "curve_"))
-  return(some_curves)
+# Application to 2D time-resolved decoding results (cross-temporal generalisation)
+
+Assume we have M/EEG data and we have conducted cross-temporal generalisation analyses [@king2014].
+As a result, we have a 2D matrix where each element contains the decoding accuracy (e.g., ROC AUC) of a classifier trained at timestep $\text{training}_{i}$ and tested at timestep $\text{testing}_{j}$ (@fig-sim-timegen).
+
+```{r fig-sim-timegen, echo = FALSE, dpi = 300, fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
+# number of participants
+n_participants <- 10
+
+# number of timepoints
+# n_timepoints <- 240 # (1.2s at 200Hz)
+n_timepoints <- 60 # (1.2s at 50Hz)
+timepoints <- seq(from = -0.2, to = 1, length.out = n_timepoints)
+
+# smooth onset and offset parameters
+decode_onset <- 0.2
+decode_offset <- 0.3
+
+# smoothness of onset/offset
+sigma_smooth <- 0.05
+
+# generating participant-specific cross-temporal generalisation matrices
+timegen_data <- tibble()
+
+# for each participant
+for (p in 1:n_participants) {
+
+    # adding variability in peak amplitude and width per participant
+    peak_variability <- runif(1, 30, 50) # random peak strength (higher = stronger decoding)
+    smooth_variability <- runif(1, 0.04, 0.06) # random smoothness variation
+    onset_variability <- runif(1, -0.05, 0.05) # small jitter in onset time
+    offset_variability <- runif(1, -0.05, 0.05) # small jitter in offset time
+
+    # computing smooth transition function (soft onset and offset with participant variation)
+    decode_window <- outer(timepoints, timepoints, function (t1, t2) {
+
+        # smooth onset
+        onset_factor <- 1 /
+            (1 + exp(-(t1 - (decode_onset + onset_variability) ) / sigma_smooth) )
+
+        # smooth offset
+        offset_factor <- 1 /
+            (1 + exp((t1 - (decode_offset + offset_variability)) / sigma_smooth) )
+
+        # diagonal Gaussian
+        onset_factor * offset_factor * exp(-((t1 - t2)^2) / (2 * smooth_variability^2) )
+
+    })
+
+    # defining smooth alpha and beta parameters for Beta distribution
+    base_alpha <- 10
+    peak_alpha <- peak_variability
+    alpha_matrix <- base_alpha + peak_alpha * decode_window
+    beta_matrix <- base_alpha
+
+    # generating AUC values using the Beta distribution
+    auc_matrix_beta <- matrix(
+        rbeta(n_timepoints^2, shape1 = alpha_matrix, shape2 = beta_matrix),
+        nrow = n_timepoints, ncol = n_timepoints
+        )
+
+    # converting to long format for plotting
+    temp_data <- expand.grid(
+        train_time = timepoints, test_time = timepoints
+        ) %>%
+        mutate(auc = as.vector(auc_matrix_beta), participant = p)
+
+    # storing results
+    timegen_data <- bind_rows(timegen_data, temp_data)
+
 }
 
-make_a_wiggle <- function(some_curves, weights, data = NULL) {
-
-  your_wiggly_line <- some_curves |> 
-    mutate(
-      weight = rep(weights, n_x),
-      weighted_value = value * weight
-      ) |> 
-    group_by(x) |> 
-    summarize(wiggly_line = sum(weighted_value)) |> 
-    ungroup()
-
-  your_wiggle_plot <- your_wiggly_line |> 
-    ggplot(aes(x = x, y = wiggly_line)) + 
-    geom_line(color = project_colors[1], size=1.5) + 
-    ggtitle("your wiggly line") +
-    xlab("x") + ylab("")
-
-  your_wiggle_plot
+# plotting cross-temporal generalization matrix
+timegen_data %>%
+    summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>%
+    ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
+    geom_tile() +
+    geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
+    geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
+    geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
+    scale_x_continuous(expand = c(0, 0) ) +
+    scale_y_continuous(expand = c(0, 0) ) + 
+    coord_fixed() +
+    scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
+    labs(x = "Testing time (s)", y = "Training time (s)")
+```
+
+```{r mne-timegen, eval = FALSE, echo = FALSE}
+# defining the function in R (it will be executed in Python)
+mne_timegen <- function (X, labels, ncores = 8) {
+
+    # converting R dataframe to NumPy array (reshaping if needed)
+    # X should be a matrix before conversion
+    X_np <- np$array(X)
+
+    if (length(dim(X_np) ) == 2) {
+
+        # adding a second dimension (channels) if missing
+        X_np <- np$expand_dims(X_np, axis = as.integer(1) )
+
     }
 
-weights <- c(0, 1, 10, rep(0, times = 17) )
-make_a_wiggle(some_curves, weights)
-```
-
-# Gaussian processes
-
-Explain GPs with code and figures here...
-
-```{r gp-tutorial, echo = TRUE, eval = FALSE}
-# see https://m-clark.github.io/models-by-example/gaussian-process.html
-# see https://distill.pub/2019/visual-exploration-gaussian-processes/
-# and http://www.infinitecuriosity.org/vizgp/
-# defining the mean-centered function
-gp_mu <- function (x) {
-
-    map_dbl(x, function (x) x = 0)
+    # defining the classifier
+    clf <- sklearn$linear_model$LogisticRegression(solver = "liblinear")
+
+    # sliding the estimator on all time frames
+    time_gen <- mne$decoding$GeneralizingEstimator(
+        clf,
+        n_jobs = as.integer(ncores),
+        scoring = "roc_auc",
+        verbose = FALSE
+        )
+
+    # or using N-fold cross-validation
+    scores <- mne$decoding$cross_val_multiscore(
+        time_gen,
+        X_np,
+        labels,
+        cv = as.integer(4),
+        n_jobs = as.integer(ncores),
+        verbose = FALSE
+        )
+
+    # returning the scores
+    return (scores)
 
@@
 
-# defining the covariance function (squared exponential kernel)
-gp_K <- function (x, l = 1, sigma_f = 1) {
-
-    sigma_f * exp( -(1/(2 * l^2)) * as.matrix(dist(x, upper = TRUE, diag = TRUE) ^ 2) )
+# listing all participants
+participants <- unique(raw_df$participant)
+
+# initialising empty decoding results
+group_timegen_scores <- data.frame()
+
+# running decoding for each participant
+for (ppt in participants) {
+
+    # printing progress
+    print(ppt)
+
+    # retrieve data from one participant
+    ppt_data <- raw_df %>%
+        filter(participant == ppt) %>%
+        select(-participant) %>%
+        pivot_wider(names_from = time, values_from = eeg) %>%
+        select(-condition, -trial)
+
+    # extracting the labels
+    labels <- raw_df %>%
+        filter(participant == ppt) %>%
+        select(-participant) %>%
+        pivot_wider(names_from = time, values_from = eeg) %>%
+        pull(condition) %>%
+        as.numeric()
+
+    # extracting the timesteps
+    timesteps <- raw_df %>%
+        filter(participant == ppt) %>%
+        select(-participant) %>%
+        pull(time) %>%
+        unique()
+
+    # running the decoding
+    timegen_scores <- mne_timegen(
+        X = ppt_data,
+        labels = labels-1
+        )
+
+    # computing the average over CV folds
+    # a <- timegen_scores %>%
+    #     apply(MARGIN = c(2, 3), FUN = mean) %>%
+    #     data.frame()
+    timegen_scores <- data.frame(np$mean(timegen_scores, axis = as.integer(0) ) )
+
+    # sanity check
+    # image(as.matrix(timegen_scores) )
+
+    # appending to previous results
+    group_timegen_scores <- bind_rows(group_timegen_scores, timegen_scores)
 
@@
 
-# uniform prior
-x_prior <- seq(-5, 5, 0.2)
-
-# multivariate draws
-y_prior = MASS::mvrnorm(
-    n  = n_prior,
-    mu = gp_mu(x_prior),
-    Sigma = gp_K(x_prior, l = l, sigma_f = sigma_f)
-    )
+# saving the scores
+saveRDS(object = group_timegen_scores, file = "results/timegen_scores.rds")
+
+# reshaping the data
+colnames(group_timegen_scores) <- unique(raw_df$time)
+timegen_data <- group_timegen_scores %>%
+    mutate(participant = rep(x = participants, each = n_distinct(raw_df$time) ) ) %>%
+    mutate(train_time = rep(x = unique(raw_df$time), times = length(participants) ) ) %>%
+    pivot_longer(
+        cols = -c(participant, train_time),
+        names_to = "test_time",
+        values_to = "auc"
+        ) %>%
+    mutate(test_time = as.numeric(test_time) )
+```
+
+Now, we want to test whether and when decoding performance is above chance level (0.5 for a binary decoding task).
+These two models are computationally heavier to fit (more observations and 2D smooth functions)...
+
+```{r gam-gp-timegen, message = FALSE}
+# fitting a GAM with two temporal dimensions
+timegen_gam <- brm(
+    # 2D thin-plate spline (tp)
+    # auc ~ s(train_time, test_time, bs = "tp", k = 10),
+    auc ~ t2(train_time, test_time, bs = "tp", k = 10),
+    data = timegen_data,
+    family = Beta(),
+    iter = 2000,
+    chains = 4,
+    cores = 4,
+    file = "models/timegen_gam_t2.rds"
+    )
+
+# fitting a GP with two temporal dimensions
+# timegen_gp <- brm(
+#     auc ~ gp(train_time, test_time, k = 20),
+#     data = timegen_data,
+#     family = Beta(),
+#     control = list(adapt_delta = 0.95),
+#     iter = 2000,
+#     chains = 4,
+#     cores = 4,
+#     file = "models/timegen_gp.rds"
+#     )
+```
+
+```{r gam-timegen-preds, eval = FALSE, echo = FALSE, fig.cap = "Posterior predictions of the GAM fitted on decoding accuracy over time."}
+# plotting the posterior predictions
+plot(
+    conditional_effects(x = timegen_gam, prob = 0.95),
+    points = FALSE,
+    # point_args = list(size = 1, pch = 21, colour = "grey"),
+    # line_args = list(colour = "black"),
+    method = "posterior_epred",
+    surface = TRUE,
+    theme = theme_light(),
+    plot = TRUE
+    )#[[1]] +
+    # geom_hline(yintercept = 0.5, linetype = 2) +
+    # geom_vline(xintercept = 0.0, linetype = 2) +
+    # ylim(0, 1) +
+    # labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
+
+# posterior predictive checks
+pp_check(timegen_gam)
+```
+
+```{r gam-timegen-post-preds, echo = FALSE, fig.width = 8, fig.asp = 0.6, fig.cap = "Posterior probability of decoding accuracy being above chance level (2D GAM)."}
+# defining a function to plot posterior prob of slope above 0
+plot_post_timegen <- function (model, data, chance_level = 0.5, eps = 0.1) {
+
+    # defining a sequence of x values to make predictions
+    preds_conds <- crossing(
+        train_time = unique(data$train_time)[c(TRUE, FALSE)],
+        test_time = unique(data$test_time)[c(TRUE, FALSE)]
+        )
+
+    # retrieving posterior predictions
+    post_preds <- conditional_effects(
+        x = model,
+        conditions = preds_conds,
+        method = "posterior_epred",
+        prob = 0.95,
+        ndraws = 100
+        )[[1]]
+
+    # plotting cross-temporal generalization matrix
+    post_preds_plot <- post_preds %>%
+        summarise(auc = mean(estimate__), .by = c(train_time, test_time) ) %>%
+        ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
+        geom_tile() +
+        geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
+        geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
+        geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
+        scale_x_continuous(expand = c(0, 0) ) +
+        scale_y_continuous(expand = c(0, 0) ) + 
+        coord_fixed() +
+        scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
+        labs(x = "Testing time (s)", y = "Training time (s)")
+
+    # retrieving posterior samples
+    posterior_samples <- posterior_epred(object = model, newdata = preds_conds)
+
+    # computing the probability that y is above 0
+    prob_y_above_0 <- preds_conds %>%
+        mutate(post_prob = colMeans(posterior_samples) ) %>%
+        mutate(m = colMeans(posterior_samples > (chance_level + eps) ) ) %>%
+        mutate(prob_ratio = m / (1 - m) )
 
     # plotting it
-prior_data = data.frame(
-    x  = x_prior,
-    y  = t(y_prior),
-    sd = apply(y_prior, 2, sd)
-    ) %>% 
-    pivot_longer(-c(x, sd), names_to = "variable")
-
-prior_data %>%
-    ggplot(aes(x = x, y = value) ) +
-    geom_line(aes(group = variable), color = "#FF550080", alpha = 0.5) +
-    labs(title = "Prior")
-```
-
--->
-````
+    # prob_y_above_0 %>% filter(prob_ratio != Inf) %>% pull(prob_ratio) %>% range()
+    prob_above_chance_plot <- prob_y_above_0 %>%
+        mutate(prob_ratio = ifelse(prob_ratio == Inf, 2000, prob_ratio) ) %>%
+        mutate(prob_ratio = ifelse(prob_ratio == 0, 0.0005, prob_ratio) ) %>%
+        ggplot(aes(x = train_time, y = test_time, fill = prob_ratio) ) +
+        geom_tile() +
+        geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
+        geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
+        geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
+        scale_x_continuous(expand = c(0, 0) ) +
+        scale_y_continuous(expand = c(0, 0) ) + 
+        coord_fixed() +
+        scale_fill_scico(palette = "vik", midpoint = 0, name = "p/(1-p)", trans = "log10") +
+        # scale_fill_scico(palette = "vik", midpoint = 1, name = "p/(1-p)") +
+        # theme_light() +
+        labs(x = "Testing time (s)", y = "Training time (s)")
+
+    # combining the two plots
+    post_preds_plot + prob_above_chance_plot
+
+    # some tests
+    # post_preds_plot +
+    #     geom_contour(
+    #         data = prob_y_above_0,
+    #         aes(x = train_time, y = test_time, z = prob_ratio),
+    #         breaks = 10,
+    #         color = "red",
+    #         linewidth = 1,
+    #         inherit.aes = FALSE
+    #         )
+
+}
+
+# using the function with the decoding GAM
+plot_post_timegen(model = timegen_gam, data = timegen_data, chance_level = 0.5, eps = 0.01)
+```
+
+```{r gp-timegen-preds, eval = FALSE, echo = FALSE, fig.cap = "Posterior predictions of the GP fitted on decoding accuracy over time."}
+# plotting the posterior predictions
+plot(
+    conditional_effects(x = timegen_gp, prob = 0.95),
+    points = FALSE,
+    method = "posterior_epred",
+    surface = TRUE,
+    theme = theme_light(),
+    plot = TRUE
+    )
+
+# posterior predictive checks
+pp_check(timegen_gp)
+
+# checking timecourse of timegen_data's diagonal
+timegen_data %>%
+    filter(train_time == test_time) %>%
+    summarise(auc = mean(auc), .by = train_time) %>%
+    ggplot(aes(x = train_time, y = auc) ) +
+    geom_hline(yintercept = 0.5, linetype = 2) +
+    geom_line() +
+    labs(x = "Time (s)", y = "AUC")
+```
+
+\newpage
 

### 2025-02-09 brms_meeg.tex
including some more refs and using t2 splines for timegen data

@@
 \citeproc{ref-ehinger_unfold_2019}{Ehinger \& Dimigen, 2019};
-\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021a},
+\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021},
 \citeproc{ref-frossard2022}{2022}; \citeproc{ref-gramfort2013}{Gramfort,
@@
 \citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al., 2019};
-\citeproc{ref-pernet2015}{Pernet et al., 2015};
+\citeproc{ref-pernet2015}{C. R. Pernet et al., 2015};
 \citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al., 2023};
@@
 Using Bayes factors (\citeproc{ref-teichmann2022}{Teichmann,
-2022})\ldots{}
+2022})\ldots{} Weighting single trials (\citeproc{ref-pernet2022}{C.
+Pernet, 2022})\ldots{}
+
+See
+\url{https://www.let.rug.nl/nerbonne/teach/rema-stats-meth-seminar/presentations/Wieling-2014-GAMs.pdf}
+and \url{https://www.let.rug.nl/wieling/Statistics/GAM-EEG/lab/},
+\url{https://www.let.rug.nl/wieling/Statistics/GAM-EEG/GAM-EEG.pdf}
+(\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
+\citeproc{ref-meulman2015}{Meulman et al., 2015})\ldots{}
+
+Recent example of GLM for EEG (\citeproc{ref-fischer2013}{Fischer \&
+Ullsperger, 2013}; \citeproc{ref-wuxfcllhorst2025}{Wüllhorst et al.,
+2025})\ldots{} ``Similar approaches have been successfully applied to
+EEG time- (Rousselet et al., 2008) and frequency-domain (Cohen and
+Cavanagh, 2011) data and allow the simultaneous investigation of
+multiple independent variables while preserving the high temporal
+resolution of the EEG.''\ldots{} See also (\citeproc{ref-hauk2006}{Hauk
+et al., 2006}; \citeproc{ref-rousselet2008}{Rousselet et al.,
+2008})\ldots{} Example of two-stage regression analysis
+(individual-level then group-level, 10.1101/2024.04.10.588861,
+\citeproc{ref-dunagan2024}{Dunagan et al., 2024})\ldots{}
 
@@
 
+See also
+\url{https://jtimonen.github.io/lgpr-usage/articles/math.html}\ldots{}
+
 Low-rank approximate Gaussian processes are of main interest in machine
@@
 ms, and an offset at 342 ms. We simulated this signal for 250 timesteps
-between 0 and 0.5s, akin to \& 500 Hz sampling rate. We simulated such
+between 0 and 0.5s, akin to a 500 Hz sampling rate. We simulated such
 data for a group of 20 participants with 50 trials per participant and
-condition.
-
-\begin{figure}[H]
-
-\caption{Some ERPs in two conditions with 50 trials each, for a group of
-20 participants.}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/eeg-group-1.pdf}
+condition (Figure~\ref{fig-eeg}).
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-eeg}Some ERPs in two conditions with 50 trials each,
+for a group of 20 participants.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-eeg-1.pdf}
 
@@
 \end{figure}%
+
+We computed the average of the ERP difference
+(Figure~\ref{fig-erp})\ldots{}
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-erp}Group-level average difference between
+conditions (mean +/- standard error of the mean). The `true' (population
+value of the) onset is indicated by the vertical dashed line.}
+
+\centering{
+
+\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-erp-1.pdf}
+
+}
+
+\end{figure}%
+
+\subsection{Model fitting}\label{model-fitting}
+
+We then fitted a GAM and GP regression model using the \texttt{brms}
+package (\citeproc{ref-brms2017}{Bürkner, 2017},
+\citeproc{ref-brms2018}{2018}; \citeproc{ref-nalborczyk2019}{Nalborczyk
+et al., 2019})\ldots{}
 
@@
 \CommentTok{\# averaging across participants}
-\NormalTok{group\_df }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ condition, }\AttributeTok{values\_from =}\NormalTok{ y, }\AttributeTok{values\_fn =}\NormalTok{ mean) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{y\_diff =}\NormalTok{ cond2 }\SpecialCharTok{{-}}\NormalTok{ cond1) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{summarise}\NormalTok{(}
-        \AttributeTok{y\_mean =} \FunctionTok{mean}\NormalTok{(y\_diff),}
-        \AttributeTok{y\_sem =} \FunctionTok{sd}\NormalTok{(y\_diff) }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(}\FunctionTok{n}\NormalTok{() ),}
-        \AttributeTok{.by =}\NormalTok{ x}
-\NormalTok{        ) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lower =}\NormalTok{ y\_mean }\SpecialCharTok{{-}}\NormalTok{ y\_sem, }\AttributeTok{upper =}\NormalTok{ y\_mean }\SpecialCharTok{+}\NormalTok{ y\_sem)}
-
-\CommentTok{\# plotting the data}
-\NormalTok{group\_df }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y\_mean) ) }\SpecialCharTok{+}
-    \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =}\NormalTok{ true\_onset, }\AttributeTok{linetype =} \DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{geom\_ribbon}\NormalTok{(}
-        \FunctionTok{aes}\NormalTok{(}\AttributeTok{ymin =}\NormalTok{ lower, }\AttributeTok{ymax =}\NormalTok{ upper, }\AttributeTok{colour =} \ConstantTok{NULL}\NormalTok{),}
-        \AttributeTok{alpha =} \FloatTok{0.25}\NormalTok{, }\AttributeTok{show.legend =} \ConstantTok{FALSE}
-\NormalTok{        ) }\SpecialCharTok{+}
-    \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{show.legend =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (ms)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG difference (a.u.)"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-\caption{Group-level average difference between conditions (mean +/-
-standard error of the mean). The `true' (population value of the) onset
-is indicated by the vertical dashed line.}
-
-{\centering \includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/erp-diff-1.pdf}
-
-}
-
-\end{figure}%
-
-\subsection{Model fitting}\label{model-fitting}
-
-Models were fitted using the \texttt{brms} package
-(\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018};
-\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019})\ldots{}
-
-\begin{Shaded}
-\begin{Highlighting}[]
+\NormalTok{ppt\_df }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{group\_by}\NormalTok{(participant, condition, time) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{summarise}\NormalTok{(}\AttributeTok{eeg =} \FunctionTok{mean}\NormalTok{(eeg) ) }\SpecialCharTok{\%\textgreater{}\%}
+    \CommentTok{\# mutate(eeg = as.numeric(scale(x = eeg, scale = FALSE) ) ) \%\textgreater{}\%}
+    \FunctionTok{ungroup}\NormalTok{()}
+
 \CommentTok{\# defining a contrast for condition}
-\FunctionTok{contrasts}\NormalTok{(group\_df}\SpecialCharTok{$}\NormalTok{condition) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{)}
+\FunctionTok{contrasts}\NormalTok{(ppt\_df}\SpecialCharTok{$}\NormalTok{condition) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{)}
 
@@
     \CommentTok{\# cubic regression splines with k{-}1 basis functions}
-    \CommentTok{\# then we should try s(x, participant, bs = "fs")}
-    \CommentTok{\# that is, random factor smooth interactions...}
-\NormalTok{    y }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(x, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
-    \AttributeTok{data =}\NormalTok{ group\_df,}
+\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
+    \AttributeTok{data =}\NormalTok{ ppt\_df,}
     \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
@@
 \begin{Highlighting}[]
+\CommentTok{\# computing eeg\_diff and averaging across participants}
+\CommentTok{\# group\_df \textless{}{-} raw\_df \%\textgreater{}\%}
+\CommentTok{\#     group\_by(participant, time) \%\textgreater{}\%}
+\CommentTok{\#     pivot\_wider(names\_from = condition, values\_from = eeg, values\_fn = mean) \%\textgreater{}\%}
+\CommentTok{\#     mutate(eeg\_diff = cond2 {-} cond1) \%\textgreater{}\%}
+\CommentTok{\#     summarise(eeg\_diff = mean(eeg\_diff) ) \%\textgreater{}\%}
+\CommentTok{\#     ungroup()}
+
+\CommentTok{\# see https://betanalpha.github.io/assets/case\_studies/gp\_part3/part3.html}
+\CommentTok{\# and https://betanalpha.github.io/assets/case\_studies/gaussian\_processes.html}
+\NormalTok{gp\_priors }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
+    \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"Intercept"}\NormalTok{),}
+    \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{),}
+    \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"exponential(1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sigma"}\NormalTok{),}
+    \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"exponential(1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sdgp"}\NormalTok{)}
+\NormalTok{    )}
+
 \NormalTok{gp\_model }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
@@
     \CommentTok{\# computing Hilbert{-}space approximate GPs}
-\NormalTok{    y }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gp}\NormalTok{(x, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
     \CommentTok{\# if k = NA (default), exact GPs are computed}
-    \CommentTok{\# y \textasciitilde{} gp(x, by = condition),}
-    \AttributeTok{data =}\NormalTok{ group\_df,}
+    \CommentTok{\# eeg \textasciitilde{} gp(time, by = condition),}
+\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{gp}\NormalTok{(time, }\AttributeTok{by =}\NormalTok{ condition, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{cov =} \StringTok{"exp\_quad"}\NormalTok{),}
+    \AttributeTok{data =}\NormalTok{ ppt\_df,}
     \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
-    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{),}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
+    \CommentTok{\# prior = gp\_priors,}
+    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{15}\NormalTok{),}
+    \AttributeTok{backend =} \StringTok{"cmdstanr"}\NormalTok{,}
+    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
     \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
@@
 \CommentTok{\# plotting the posterior predictions}
-\FunctionTok{plot\_post\_preds}\NormalTok{(}\AttributeTok{model =}\NormalTok{ gam, }\AttributeTok{data =}\NormalTok{ group\_df) }\SpecialCharTok{+}
-    \FunctionTok{plot\_post\_preds}\NormalTok{(}\AttributeTok{model =}\NormalTok{ gp\_model, }\AttributeTok{data =}\NormalTok{ group\_df)}
+\FunctionTok{plot\_post\_preds}\NormalTok{(}\AttributeTok{model =}\NormalTok{ gam) }\SpecialCharTok{+} \FunctionTok{plot\_post\_preds}\NormalTok{(}\AttributeTok{model =}\NormalTok{ gp\_model)}
 \end{Highlighting}
@@
 with the conventional BF thresholds (Figure~\ref{fig-post-prob-ratio}).
+Also check this paper
+(\url{https://link.springer.com/article/10.3758/s13428-021-01613-6?fromPaywallRec=false})?
 
@@
 Below we are plotting the error function (where the error is computed as
-\(|\hat{\theta}-\theta|\)) for both the onset and offset values of the
-ERP difference, according to various \texttt{eps} and \texttt{threshold}
-values. Remember that the signal is generated from a truncated Gaussian
-defining an objective onset at 160 ms, a maximum at 250 ms, and an
-offset at 342 ms. Figure~\ref{fig-onset-error} shows that the GP model
-can \emph{exactly} recover the true onset and offset values, given some
-reasonable choice of \texttt{eps} and \texttt{threshold} values.
+\(\text{error}:=|\hat{\theta}-\theta|\)) for both the onset and offset
+values of the ERP difference, according to various \texttt{eps} and
+\texttt{threshold} values. Remember that the signal is generated from a
+truncated Gaussian defining an objective onset at 160 ms, a maximum at
+250 ms, and an offset at 342 ms. Figure~\ref{fig-onset-error} shows that
+the GP model can \emph{exactly} recover the true onset and offset
+values, given some reasonable choice of \texttt{eps} and
+\texttt{threshold} values.
 
@@
    eps threshold estimated_onset estimated_offset error_onset error_offset
-1 0.06        15             160              342           0            0
-2 0.06        16             160              342           0            0
-3 0.06        17             160              342           0            0
-4 0.06        18             160              342           0            0
-5 0.06        19             160              342           0            0
-6 0.06        20             160              342           0            0
+1 0.03         6            0.16            0.351           0        0.009
+2 0.03         7            0.16            0.351           0        0.009
+3 0.03         8            0.16            0.351           0        0.009
+4 0.03         9            0.16            0.351           0        0.009
+5 0.03        10            0.16            0.351           0        0.009
+6 0.03        11            0.16            0.351           0        0.009
 \end{verbatim}
@@
    eps threshold estimated_onset estimated_offset error_onset error_offset
-1 0.06         3             159              342           1            0
-2 0.06         4             159              342           1            0
-3 0.06         5             159              342           1            0
-4 0.06         6             159              342           1            0
-5 0.06         7             159              342           1            0
-6 0.06         8             159              342           1            0
+1 0.07         8           0.166            0.342       0.006            0
+2 0.07         9           0.166            0.342       0.006            0
+3 0.07        10           0.166            0.342       0.006            0
+4 0.07        11           0.166            0.342       0.006            0
+5 0.07        12           0.166            0.342       0.006            0
+6 0.07        13           0.166            0.342       0.006            0
+\end{verbatim}
+
+\begin{verbatim}
+[1] eps              threshold        estimated_onset  estimated_peak  
+[5] estimated_offset error_onset      error_peak       error_offset    
+<0 rows> (or 0-length row.names)
 \end{verbatim}
@@
     \FunctionTok{summarise}\NormalTok{(}
-        \AttributeTok{eeg =} \FunctionTok{mean}\NormalTok{(y),}
-        \AttributeTok{eeg\_sd =} \FunctionTok{sd}\NormalTok{(y),}
-        \AttributeTok{.by =} \FunctionTok{c}\NormalTok{(participant, condition, x)}
+        \AttributeTok{eeg\_mean =} \FunctionTok{mean}\NormalTok{(eeg),}
+        \AttributeTok{eeg\_sd =} \FunctionTok{sd}\NormalTok{(eeg),}
+        \AttributeTok{.by =} \FunctionTok{c}\NormalTok{(participant, condition, time)}
 \NormalTok{        )}
@@
     \CommentTok{\# using by{-}participant SD of ERPs across trials}
-\NormalTok{    eeg }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
-\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(x, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
+\NormalTok{    eeg\_mean }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
+\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
 \NormalTok{        (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ participant),}
@@
 \begin{Highlighting}[]
-\CommentTok{\# plotting the posterior predictions}
-\FunctionTok{plot}\NormalTok{(}
-    \FunctionTok{conditional\_effects}\NormalTok{(}\AttributeTok{x =}\NormalTok{ meta\_gam, }\AttributeTok{effect =} \StringTok{"x:condition"}\NormalTok{),}
-    \AttributeTok{points =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{theme =} \FunctionTok{theme\_light}\NormalTok{(), }\AttributeTok{plot =} \ConstantTok{FALSE}
-\NormalTok{    )[[}\DecValTok{1}\NormalTok{]] }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (ms)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG signal (a.u.)"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-\caption{Hierarchical GAM posterior predictions.}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/meta-gam-preds-1.pdf}
-
-}
-
-\end{figure}%
-
-\begin{Shaded}
-\begin{Highlighting}[]
 \CommentTok{\# fitting the GP}
@@
     \CommentTok{\# using by{-}participant SD of ERPs across trials}
-\NormalTok{    eeg }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
-\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{gp}\NormalTok{(x, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
+\NormalTok{    eeg\_mean }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
+\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{gp}\NormalTok{(time, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
 \NormalTok{        (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ participant),}
@@
     \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
-    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{20}\NormalTok{),}
+    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{),}
     \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
@@
 \CommentTok{\# plotting the posterior predictions}
-\FunctionTok{plot}\NormalTok{(}
-    \FunctionTok{conditional\_effects}\NormalTok{(}\AttributeTok{x =}\NormalTok{ meta\_gp, }\AttributeTok{effect =} \StringTok{"x:condition"}\NormalTok{),}
+\NormalTok{meta\_gam\_preds }\OtherTok{\textless{}{-}} \FunctionTok{plot}\NormalTok{(}
+    \FunctionTok{conditional\_effects}\NormalTok{(}\AttributeTok{x =}\NormalTok{ meta\_gam, }\AttributeTok{effect =} \StringTok{"time:condition"}\NormalTok{),}
     \AttributeTok{points =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{theme =} \FunctionTok{theme\_light}\NormalTok{(), }\AttributeTok{plot =} \ConstantTok{FALSE}
 \NormalTok{    )[[}\DecValTok{1}\NormalTok{]] }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (ms)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG signal (a.u.)"}\NormalTok{)}
+    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (s)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG signal (a.u.)"}\NormalTok{)}
+\NormalTok{meta\_gp\_preds }\OtherTok{\textless{}{-}} \FunctionTok{plot}\NormalTok{(}
+    \FunctionTok{conditional\_effects}\NormalTok{(}\AttributeTok{x =}\NormalTok{ meta\_gp, }\AttributeTok{effect =} \StringTok{"time:condition"}\NormalTok{),}
+    \AttributeTok{points =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{theme =} \FunctionTok{theme\_light}\NormalTok{(), }\AttributeTok{plot =} \ConstantTok{FALSE}
+\NormalTok{    )[[}\DecValTok{1}\NormalTok{]] }\SpecialCharTok{+}
+    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (s)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG signal (a.u.)"}\NormalTok{)}
+
+\CommentTok{\# combining the two plots}
+\NormalTok{meta\_gam\_preds }\SpecialCharTok{+}\NormalTok{ meta\_gp\_preds}
 \end{Highlighting}
@@
 
-\subsection{Application to 1D decoding results (accuracy over
-time)}\label{application-to-1d-decoding-results-accuracy-over-time}
-
-Assume we have M/EEG data and we conducted time-resolved multivariate
-pattern analysis (MVPA), also known as decoding. As a result, we have a
-timecourse of decoding accuracies (e.g., ROC AUC), bounded between 0 and
-1, per participant (Figure~\ref{fig-sim-decoding}).
+\begin{figure}[H]
+
+\caption{Posterior predictions for the hierarchical GAM and GP.}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/meta-gp-preds-1.pdf}
+
+}
+
+\end{figure}%
+
+\subsection{Application to time-resolved decoding results (accuracy over
+time)}\label{application-to-time-resolved-decoding-results-accuracy-over-time}
+
+We conducted time-resolved multivariate pattern analysis (MVPA), also
+known as decoding. As a result, we have a timecourse of decoding
+accuracies (e.g., ROC AUC), bounded between 0 and 1, per participant
+(Figure~\ref{fig-mne-decoding})\ldots{}
 
@@
 
-\caption{\label{fig-sim-decoding}Exemplary (simulated) group-level
-average timecourse of binary decoding accuracy (ROC AUC).}
+\caption{\label{fig-mne-decoding}Exemplary average timecourse of binary
+decoding accuracy (ROC AUC) for each participant. Group-level average
+decoding accuracy is depicted as a grey background density in each
+panel.}
 
@@
 
-\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-decoding-1.pdf}
-
-}
-
-\end{figure}%
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# loading the reticulate package}
-\FunctionTok{library}\NormalTok{(reticulate)}
-
-\CommentTok{\# importing the numpy and mne python modules}
-\NormalTok{np }\OtherTok{\textless{}{-}} \FunctionTok{import}\NormalTok{(}\StringTok{"numpy"}\NormalTok{)}
-\NormalTok{mne }\OtherTok{\textless{}{-}} \FunctionTok{import}\NormalTok{(}\StringTok{"mne"}\NormalTok{)}
-\NormalTok{sklearn }\OtherTok{\textless{}{-}} \FunctionTok{import}\NormalTok{(}\StringTok{"sklearn"}\NormalTok{)}
-
-\CommentTok{\# defining the function in R (it will be executed in Python)}
-\NormalTok{mne\_decoding }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{ (X, labels, }\AttributeTok{ncores =} \DecValTok{1}\NormalTok{) \{}
-
-    \CommentTok{\# converting R dataframe to NumPy array (reshaping if needed)}
-    \CommentTok{\# X should be a matrix before conversion}
-\NormalTok{    X\_np }\OtherTok{\textless{}{-}}\NormalTok{ np}\SpecialCharTok{$}\FunctionTok{array}\NormalTok{(X)}
-
-    \ControlFlowTok{if}\NormalTok{ (}\FunctionTok{length}\NormalTok{(}\FunctionTok{dim}\NormalTok{(X\_np) ) }\SpecialCharTok{==} \DecValTok{2}\NormalTok{) \{}
-
-        \CommentTok{\# adding a second dimension (channels) if missing}
-\NormalTok{        X\_np }\OtherTok{\textless{}{-}}\NormalTok{ np}\SpecialCharTok{$}\FunctionTok{expand\_dims}\NormalTok{(X\_np, }\AttributeTok{axis =} \FunctionTok{as.integer}\NormalTok{(}\DecValTok{1}\NormalTok{) )}
-
-\NormalTok{    \}}
-
-    \CommentTok{\# defining the classifier}
-\NormalTok{    clf }\OtherTok{\textless{}{-}}\NormalTok{ sklearn}\SpecialCharTok{$}\NormalTok{linear\_model}\SpecialCharTok{$}\FunctionTok{LogisticRegression}\NormalTok{(}\AttributeTok{solver =} \StringTok{"liblinear"}\NormalTok{)}
-
-    \CommentTok{\# sliding the estimator on all time frames}
-\NormalTok{    time\_decod }\OtherTok{\textless{}{-}}\NormalTok{ mne}\SpecialCharTok{$}\NormalTok{decoding}\SpecialCharTok{$}\FunctionTok{SlidingEstimator}\NormalTok{(}
-\NormalTok{        clf, }\AttributeTok{n\_jobs =} \FunctionTok{as.integer}\NormalTok{(ncores),}
-        \AttributeTok{scoring =} \StringTok{"roc\_auc"}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{TRUE}
-\NormalTok{        )}
-
-    \CommentTok{\# or using N{-}fold cross{-}validation}
-\NormalTok{    scores }\OtherTok{\textless{}{-}}\NormalTok{ mne}\SpecialCharTok{$}\NormalTok{decoding}\SpecialCharTok{$}\FunctionTok{cross\_val\_multiscore}\NormalTok{(}
-\NormalTok{        time\_decod,}
-\NormalTok{        X\_np,}
-\NormalTok{        labels,}
-        \AttributeTok{cv =} \FunctionTok{as.integer}\NormalTok{(}\DecValTok{4}\NormalTok{),}
-        \AttributeTok{n\_jobs =} \FunctionTok{as.integer}\NormalTok{(ncores),}
-        \AttributeTok{verbose =} \ConstantTok{TRUE}
-\NormalTok{        )}
-
-    \CommentTok{\# returning the scores (averaged over CV folds)}
-    \FunctionTok{return}\NormalTok{ (scores)}
-
-\NormalTok{\}}
-
-\CommentTok{\# listing all participants}
-\NormalTok{participants }\OtherTok{\textless{}{-}} \FunctionTok{unique}\NormalTok{(raw\_df}\SpecialCharTok{$}\NormalTok{participant)}
-
-\CommentTok{\# initialising empty decoding results}
-\NormalTok{group\_decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{()}
-
-\CommentTok{\# running decoding for each participant}
-\ControlFlowTok{for}\NormalTok{ (ppt }\ControlFlowTok{in}\NormalTok{ participants) \{}
-
-    \CommentTok{\# printing progress}
-    \FunctionTok{print}\NormalTok{(ppt)}
-
-    \CommentTok{\# retrieve data from one participant}
-\NormalTok{    ppt\_data }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{filter}\NormalTok{(participant }\SpecialCharTok{==}\NormalTok{ ppt) }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{participant) }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ x, }\AttributeTok{values\_from =}\NormalTok{ y) }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{condition, }\SpecialCharTok{{-}}\NormalTok{trial)}
-
-    \CommentTok{\# extracting the labels}
-\NormalTok{    labels }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{filter}\NormalTok{(participant }\SpecialCharTok{==}\NormalTok{ ppt) }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{participant) }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ x, }\AttributeTok{values\_from =}\NormalTok{ y) }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{pull}\NormalTok{(condition) }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{as.numeric}\NormalTok{()}
-
-    \CommentTok{\# extracting the timesteps}
-\NormalTok{    timesteps }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{filter}\NormalTok{(participant }\SpecialCharTok{==}\NormalTok{ ppt) }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{participant) }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{pull}\NormalTok{(x) }\SpecialCharTok{\%\textgreater{}\%}
-        \FunctionTok{unique}\NormalTok{()}
-
-    \CommentTok{\# running the decoding}
-\NormalTok{    decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(}
-        \FunctionTok{mne\_decoding}\NormalTok{(}\AttributeTok{X =}\NormalTok{ ppt\_data, }\AttributeTok{labels =}\NormalTok{ labels}\DecValTok{{-}1}\NormalTok{)}
-\NormalTok{        ) }\SpecialCharTok{\%\textgreater{}\%}
-        \CommentTok{\# computing the average over CV folds}
-        \FunctionTok{summarise}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{where}\NormalTok{(is.numeric), mean) )}
-
-    \CommentTok{\# appending to previous results}
-\NormalTok{    group\_decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{bind\_rows}\NormalTok{(group\_decoding\_scores, decoding\_scores)}
-
-\NormalTok{\}}
-
-\CommentTok{\# saving the scores}
-\FunctionTok{saveRDS}\NormalTok{(}\AttributeTok{object =}\NormalTok{ group\_decoding\_scores, }\AttributeTok{file =} \StringTok{"results/decoding\_scores.rds"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# importing the decoding scores}
-\NormalTok{group\_decoding\_scores }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\AttributeTok{file =} \StringTok{"results/decoding\_scores.rds"}\NormalTok{)}
-
-\CommentTok{\# extracting the timesteps}
-\NormalTok{timesteps }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
-    \CommentTok{\# filter(participant == ppt) \%\textgreater{}\%}
-    \CommentTok{\# select({-}participant) \%\textgreater{}\%}
-    \FunctionTok{pull}\NormalTok{(x) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{unique}\NormalTok{()}
-
-\CommentTok{\# plotting it}
-\NormalTok{group\_decoding\_scores }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{t}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{data.frame}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}
-        \AttributeTok{mean =} \FunctionTok{rowMeans}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{20}\NormalTok{) ),}
-        \AttributeTok{se =} \FunctionTok{apply}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{20}\NormalTok{), }\DecValTok{1}\NormalTok{, }\ControlFlowTok{function}\NormalTok{ (x) }\FunctionTok{sd}\NormalTok{(x) }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(}\FunctionTok{length}\NormalTok{(x) ) )}
-\NormalTok{        ) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{lower =}\NormalTok{ mean }\SpecialCharTok{{-}}\NormalTok{ se, }\AttributeTok{upper =}\NormalTok{ mean }\SpecialCharTok{+}\NormalTok{ se) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{time =}\NormalTok{ timesteps) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time, }\AttributeTok{y =}\NormalTok{ mean) ) }\SpecialCharTok{+}
-    \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{geom\_ribbon}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{ymin =}\NormalTok{ lower, }\AttributeTok{ymax =}\NormalTok{ upper), }\AttributeTok{fill =} \StringTok{"steelblue"}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.2}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{color =} \StringTok{"steelblue"}\NormalTok{, }\AttributeTok{size =} \FloatTok{0.8}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (s)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Decoding accuracy (ROC AUC)"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-\caption{Exemplary group-level average timecourse of binary decoding
-accuracy (ROC AUC). Decoding was performed using MNE-Python.}
-
-{\centering \includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/unnamed-chunk-2-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-mne-decoding-1.pdf}
 
@@
     \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.9}\NormalTok{),}
+    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{),}
     \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
@@
 
-\begin{verbatim}
-  cluster_onset cluster_offset
-1     0.1263598      0.6937238
-  cluster_onset cluster_offset
-1     0.1615063      0.6786611
-\end{verbatim}
-
 \begin{figure}[!htb]
@@
 \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-ratio-1.pdf}
-
-}
-
-\end{figure}%
-
-\newpage
-
-\subsection{Application to 2D decoding results (cross-temporal
-generalisation)}\label{application-to-2d-decoding-results-cross-temporal-generalisation}
-
-Assume we have M/EEG data and we have conducted cross-temporal
-generalisation analyses (\citeproc{ref-king2014}{King \& Dehaene,
-2014}). As a result, we have a 2D matrix where each element contains the
-decoding accuracy (e.g., ROC AUC) of a classifier trained at timestep
-\(\text{training}_{i}\) and tested at timestep \(\text{testing}_{j}\)
-(Figure~\ref{fig-sim-timegen}).
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-sim-timegen}Exemplary (simulated) group-level
-average cross-temporal generalisation matrix of decoding performance
-(ROC AUC).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-timegen-1.pdf}
-
-}
-
-\end{figure}%
-
-Now, we want to test whether and when decoding performance is above
-chance level (0.5 for a binary decoding task). These two models are
-computationally heavier to fit (more observations and 2D smooth
-functions)\ldots{}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# fitting a GAM with two temporal dimensions}
-\NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-    \CommentTok{\# 2D thin{-}plate spline (tp) to model smooth interactions between training and testing time}
-    \CommentTok{\# auc \textasciitilde{} s(train\_time, test\_time, bs = "tp", k = 10),}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
-    \AttributeTok{data =}\NormalTok{ timegen\_data,}
-    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"}
-\NormalTok{    )}
-
-\CommentTok{\# fitting a GP with two temporal dimensions}
-\NormalTok{timegen\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gp}\NormalTok{(train\_time, test\_time, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
-    \CommentTok{\# data = timegen\_data \%\textgreater{}\% mutate(auc\_c = scale(x = auc, center = TRUE, scale = FALSE) ),}
-    \AttributeTok{data =}\NormalTok{ timegen\_data,}
-    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{20}\NormalTok{),}
-    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/timegen\_gp.rds"}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-\caption{Posterior probability of decoding accuracy being above chance
-level (2D GAM).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-timegen-post-preds-1.pdf}
 
@@
 \item
-  univariate tests + Bonferroni
+  univariate tests + Bonferroni/Holm
 \item
@@
 
-We used the \texttt{R} package \texttt{permuco} v 1.1.3
-(\citeproc{ref-permuco}{Frossard \& Renaud, 2021b}) and the \texttt{R}
-package \texttt{changepoint} v 2.2.4 (\citeproc{ref-changepoint}{Killick
-et al., 2022a})\ldots{}
+We used the \texttt{R} package \texttt{changepoint} v 2.2.4
+(\citeproc{ref-changepoint}{Killick et al., 2022a})\ldots{}
+
+S-values (\citeproc{ref-greenland2019}{Greenland, 2019})\ldots{}
 
@@
 identified using the raw (uncorrected) p-values or the corrected
-p-values (BH, BY, Bonferroni, or Holm).}
+p-values (BH, BY, Holm).}
 
@@
     \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \SpecialCharTok{{-}}\FunctionTok{log2}\NormalTok{(}\FloatTok{0.05}\NormalTok{), }\AttributeTok{linetype =} \DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-    \CommentTok{\# geom\_area(position = "identity") +}
     \FunctionTok{labs}\NormalTok{(}
-        \AttributeTok{x =} \StringTok{"Time (ms)"}\NormalTok{,}
+        \AttributeTok{x =} \StringTok{"Time (s)"}\NormalTok{,}
         \AttributeTok{y =} \StringTok{"{-}log2(p{-}value)"}
@@
 
-\ldots{}
+Extending the approach to spatiotemporal data (i.e., time +
+sensors)\ldots{}
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-abugaber2023}
+Abugaber, D., Finestrat, I., Luque, A., \& Morgan-Short, K. (2023).
+Generalized additive mixed modeling of EEG supports dual-route accounts
+of morphosyntax in suggesting no word frequency effects on processing of
+regular grammatical forms. \emph{Journal of Neurolinguistics},
+\emph{67}, 101137.
+\url{https://doi.org/10.1016/j.jneuroling.2023.101137}
+
 \bibitem[\citeproctext]{ref-rmarkdown2024}
@@
 
+\bibitem[\citeproctext]{ref-dunagan2024}
+Dunagan, D., Jordan, T., Hale, J. T., Pylkkänen, L., \& Chacón, D. A.
+(2024). \emph{Evaluating the timecourses of morpho-orthographic,
+lexical, and grammatical processing following rapid parallel visual
+presentation: An EEG investigation in english}.
+\url{http://dx.doi.org/10.1101/2024.04.10.588861}
+
 \bibitem[\citeproctext]{ref-ehinger_unfold_2019}
@@
 
+\bibitem[\citeproctext]{ref-fischer2013}
+Fischer, Adrian~G., \& Ullsperger, M. (2013). Real and Fictive Outcomes
+Are Processed Differently but Converge on a Common Adaptive Mechanism.
+\emph{Neuron}, \emph{79}(6), 1243--1255.
+\url{https://doi.org/10.1016/j.neuron.2013.07.006}
+
 \bibitem[\citeproctext]{ref-frossard2021}
-Frossard, J., \& Renaud, O. (2021a). Permutation Tests for Regression,
+Frossard, J., \& Renaud, O. (2021). Permutation Tests for Regression,
 ANOVA, and Comparison of Signals: The {\textbf{permuco}} Package.
 \emph{Journal of Statistical Software}, \emph{99}(15).
-\url{https://doi.org/10.18637/jss.v099.i15}
-
-\bibitem[\citeproctext]{ref-permuco}
-Frossard, J., \& Renaud, O. (2021b). Permutation tests for regression,
-{ANOVA}, and comparison of signals: The {permuco} package. \emph{Journal
-of Statistical Software}, \emph{99}(15), 1--32.
 \url{https://doi.org/10.18637/jss.v099.i15}
@@
 
+\bibitem[\citeproctext]{ref-greenland2019}
+Greenland, S. (2019). Valid {\emph{P}} -Values Behave Exactly as They
+Should: Some Misleading Criticisms of {\emph{P}} -Values and Their
+Resolution With {\emph{S}} -Values. \emph{The American Statistician},
+\emph{73}(sup1), 106--114.
+\url{https://doi.org/10.1080/00031305.2018.1529625}
+
+\bibitem[\citeproctext]{ref-hauk2006}
+Hauk, O., Davis, M. H., Ford, M., Pulvermüller, F., \& Marslen-Wilson,
+W. D. (2006). The time course of visual word recognition as revealed by
+linear regression analysis of ERP data. \emph{NeuroImage}, \emph{30}(4),
+1383--1400. \url{https://doi.org/10.1016/j.neuroimage.2005.11.048}
+
 \bibitem[\citeproctext]{ref-hayasaka_validating_2003}
@@
 \url{https://doi.org/10.1016/j.jneumeth.2007.03.024}
+
+\bibitem[\citeproctext]{ref-meulman2015}
+Meulman, N., Wieling, M., Sprenger, S. A., Stowe, L. A., \& Schmid, M.
+S. (2015). Age Effects in L2 Grammar Processing as Revealed by ERPs and
+How (Not) to Study Them. \emph{PLOS ONE}, \emph{10}(12), e0143328.
+\url{https://doi.org/10.1371/journal.pone.0143328}
 
@@
 
+\bibitem[\citeproctext]{ref-pernet2022}
+Pernet, C. (2022). Electroencephalography robust statistical linear
+modelling using a single weight per trial. \emph{Aperture Neuro},
+\emph{2}, 1--22.
+\url{https://doi.org/10.52294/apertureneuro.2022.2.seoo9435}
+
 \bibitem[\citeproctext]{ref-pernet2015}
@@
 \url{https://doi.org/10.1111/ejn.16618}
+
+\bibitem[\citeproctext]{ref-rousselet2008}
+Rousselet, G. A., Pernet, C. R., Bennett, P. J., \& Sekuler, A. B.
+(2008). Parametric study of EEG sensitivity to phase noise during face
+processing. \emph{BMC Neuroscience}, \emph{9}(1).
+\url{https://doi.org/10.1186/1471-2202-9-98}
 
@@
 
+\bibitem[\citeproctext]{ref-wuxfcllhorst2025}
+Wüllhorst, V., Wüllhorst, R., Overmeyer, R., \& Endrass, T. (2025).
+Comprehensive Analysis of Event{-}Related Potentials of Response
+Inhibition: The Role of Negative Urgency and Compulsivity.
+\emph{Psychophysiology}, \emph{62}(2).
+\url{https://doi.org/10.1111/psyp.70000}
+
 \bibitem[\citeproctext]{ref-knitr2014}
@@
 \appendix
+
+\section{Application to 2D time-resolved decoding results
+(cross-temporal
+generalisation)}\label{application-to-2d-time-resolved-decoding-results-cross-temporal-generalisation}
+
+Assume we have M/EEG data and we have conducted cross-temporal
+generalisation analyses (\citeproc{ref-king2014}{King \& Dehaene,
+2014}). As a result, we have a 2D matrix where each element contains the
+decoding accuracy (e.g., ROC AUC) of a classifier trained at timestep
+\(\text{training}_{i}\) and tested at timestep \(\text{testing}_{j}\)
+(Figure~\ref{fig-sim-timegen}).
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-sim-timegen}Exemplary (simulated) group-level
+average cross-temporal generalisation matrix of decoding performance
+(ROC AUC).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-timegen-1.pdf}
+
+}
+
+\end{figure}%
+
+Now, we want to test whether and when decoding performance is above
+chance level (0.5 for a binary decoding task). These two models are
+computationally heavier to fit (more observations and 2D smooth
+functions)\ldots{}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# fitting a GAM with two temporal dimensions}
+\NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
+    \CommentTok{\# 2D thin{-}plate spline (tp)}
+    \CommentTok{\# auc \textasciitilde{} s(train\_time, test\_time, bs = "tp", k = 10),}
+\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
+    \AttributeTok{data =}\NormalTok{ timegen\_data,}
+    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
+    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
+    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"}
+\NormalTok{    )}
+
+\CommentTok{\# fitting a GP with two temporal dimensions}
+\CommentTok{\# timegen\_gp \textless{}{-} brm(}
+\CommentTok{\#     auc \textasciitilde{} gp(train\_time, test\_time, k = 20),}
+\CommentTok{\#     data = timegen\_data,}
+\CommentTok{\#     family = Beta(),}
+\CommentTok{\#     control = list(adapt\_delta = 0.95),}
+\CommentTok{\#     iter = 2000,}
+\CommentTok{\#     chains = 4,}
+\CommentTok{\#     cores = 4,}
+\CommentTok{\#     file = "models/timegen\_gp.rds"}
+\CommentTok{\#     )}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[H]
+
+\caption{Posterior probability of decoding accuracy being above chance
+level (2D GAM).}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-timegen-post-preds-1.pdf}
+
+}
+
+\end{figure}%
+
+\newpage
 

### 2025-02-09 code/errors_simulation.R

@@ -1,0 +1,1258 @@
+###########################################################
+# Monte-Carlo simulation of onset/offset error properties #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on February 9, 2025                        #
+###########################################################
+
+library(changepoint)
+library(tidyverse)
+library(brms)
+
+# retrieving the code from the qdm document
+# knitr::purl(input = "brms_meeg.qmd", output = "brms_meeg.R", documentation = 0)
+
+# importing R version of Matlab code from Yeung et al. (2004)
+source("eeg_noise.R")
+
+# importing the ERP template with true onset = 160 ms, F=81, and max at F=126
+source("erp_template.R")
+
+# importing helper functions
+source("functions.R")
+
+# to use with the eeg_noise function
+meanpower <- unlist(read.table("meanpower.txt") )
+
+# defining simulation parameters
+n_trials <- 50 # number of trials
+n_ppt <- 20 # number of participants
+outvar <- 1 # noise variance
+srate <- 500 # sampling rate in Hz
+ronset <- seq(150, 170, 2) # random onset for each participant
+
+# removing raw_df
+if (exists("raw_df") ) rm(raw_df)
+
+for (P in 1:n_ppt) { # for each participant
+    
+    # get random onset
+    ponset <- sample(ronset, 1)
+    
+    # find starting point
+    st <- which(Xf == ponset)
+    
+    # pad vector
+    temp2 <- c(rep(0, st - 2), erp, rep(0, Nf - st - length(erp) + 2) )
+    
+    # initialising empty conditions
+    cond1 <- matrix(0, nrow = n_trials, ncol = Nf)
+    cond2 <- matrix(0, nrow = n_trials, ncol = Nf)
+    
+    for (T in 1:n_trials) { # for each trial
+        
+        cond1[T, ] <- temp1 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
+        cond2[T, ] <- temp2 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
+        
+    }
+    
+    # converting results to dataframe
+    temp_df <- data.frame(
+        x = rep(Xf, 2*nrow(cond1) ),
+        y = c(c(t(cond1) ), c(t(cond2) ) ),
[…]

### 2025-02-10 brms_meeg.qmd
reshaping the document and drafting a script to conduct proper simulations of onset/offset recovery

@@
 ---
-title: "Modelling M/EEG data with Bayesian nonparametric multilevel models"
+title: "Modelling M/EEG data with Bayesian multilevel generalised additive models"
 shorttitle: "Bayesian M/EEG modelling"
@@
 Typically, we are interested in testing whether such measures differ across conditions and/or groups.
-The conventional approach consists in conducting mass-univariate statistics followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster-based inference.
+The conventional approach consists in conducting mass-univariate statistics through followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster-based inference.
 However, these cluster-based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the onset and offset of effects (e.g., differences across conditions).
-Here, we introduce a novel *model-based approch* for analysing one-dimensional M/EEG timeseries such as ERPs or decoding timecourses and their differences across conditions.
-This approach relies on Bayesian nonparametric multilevel modelling (multilevel generalised additive models or Gaussian processes), which outputs posterior probabililities of the effect being above chance at every timestep/voxel, while naturally taking into account the temporal dependencies and between-subject variability present in such data."
+Here, we introduce a novel *model-based approch* for analysing one-dimensional M/EEG timeseries such as ERPs or decoding timecourses and their differences across conditions or group.
+This approach relies on Bayesian nonparametric multilevel modelling (multilevel generalised additive models), which outputs the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between-subject variability present in such data."
 keywords: [EEG, MEG, generalised additive models, gaussian processes, mixed-effects models, Bayesian statistics, brms]
@@
   apaquarto-pdf:
+    latex-engine: xelatex
     documentmode: doc # man, jou, doc, stu
@@
 
+From @dimigen2021: "Recently, spline regression has been applied to ERPs (Hendrix, Baayen, & Bolger, 2017; Kryuchkova,Tucker,Wurm,&Baayen,2012;Tremblay& Baayen, 2010; Tremblay & Newman, 2015)"...
+
 Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]... Using Bayes factors [@teichmann2022]... Weighting single trials [@pernet2022]...
@@
 See also [@rousselet2008; @hauk2006]...
-Example of two-stage regression analysis [individual-level then group-level, 10.1101/2024.04.10.588861, @dunagan2024]...
+Example of two-stage regression analysis [i.e., individual-level then group-level, @dunagan2024]...
 
@@
 
+```{=html}
+<!--
+
 ## Gaussian process regression
@@
 
+-->
+```
+
 ## Objectives
@@
 srate <- 500 # sampling rate in Hz
-ronset <- seq(150, 170, 2) # random onset for each participant
+ronset <- seq(from = 150, to = 170, by = 2) # random onset for each participant
 
@@
 true_onset <- 0.160
+true_peak <- 0.250
 true_offset <- 0.342
@@
     geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = true_offset, linetype = 2) +
     geom_line(show.legend = FALSE) +
     facet_wrap(~participant) +
+    scale_colour_met_d(name = "Johnson") +
     labs(x = "Time (s)", y = "EEG signal (a.u.)")
@@
 ```{r fig-erp, echo = FALSE, out.width = "75%", fig.cap = "Group-level average difference between conditions (mean +/- standard error of the mean).
-The 'true' (population value of the) onset is indicated by the vertical dashed line."}
+The 'true' onset and offset are indicated by the vertical dashed lines."}
 # averaging across participants
@@
     geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = true_offset, linetype = 2) +
     geom_ribbon(
@@
 
-We then fitted a GAM and GP regression model using the `brms` package [@brms2017; @brms2018; @nalborczyk2019]...
+We then fitted a GAM and GP regression model using the `brms` package [@brms2017; @brms2018; @nalborczyk2019].
+We used the default priors in `brms`, that is, weakly informative priors.
+Four Markov Chain Monte-Carlo (MCMC) were ran for each model to approximate the posterior distribution, including each 5000 iterations and a warmup of 2000 iterations.
+Posterior convergence was assessed examining trace plots as well as the Gelman–Rubin statisti $\hat{R}$.
 
@@
     family = gaussian(),
+    warmup = 2000,
     iter = 5000,
@@
 
-Now we fit the GP...
+Now we fit the GP regression model...
 
@@
         geom_vline(xintercept = true_onset, linetype = 2) +
+        geom_vline(xintercept = true_offset, linetype = 2) +
         geom_hline(yintercept = 0, linetype = 2) +
@@
             ) +
+        scale_colour_met_d(name = "Johnson") +
+        scale_fill_met_d(name = "Johnson") +
         labs(x = "Time (s)", y = "EEG signal (a.u.)")
@@
 }
+
 ```
@@
 plot_post_preds(model = gam) + plot_post_preds(model = gp_model)
+
 ```
@@
 
-We can retrieve the posterior probability of the slope for `condition` at each timestep.
-
-```{r, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Slope for the difference between conditions according to the GAM (left) or the GP (right)."}
-# check posterior_smooths()?
-# https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/posterior_smooths.brmsfit
+We can retrieve the posterior probability of the slope for `condition` at each timestep (@fig-plot-post-slope).
+
+```{r fig-plot-post-slope, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Slope for the difference between conditions according to the GAM (left) or the GP (right)."}
 # defining a function to plot posterior slope
@@
         geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
+        scale_colour_met_d(name = "Johnson") +
+        scale_fill_met_d(name = "Johnson") +
         labs(x = "Time (s)", y = expression(Pr(beta~"|"~data, model) ) )
@@
 # using the function with model
-plot_post_slope(model = gam, eps = 0.1) +
-    plot_post_slope(model = gp_model, eps = 0.1)
-```
-
-We can also compute the posterior probability of the slope for `condition` being above 0 (@fig-post-prob-test).
-
-```{r fig-post-prob-test, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior probability of the EEG signal being above 0+eps according to the GAM (left) and the GP (right)."}
+plot_post_slope(model = gam, eps = 0.1) + plot_post_slope(model = gp_model, eps = 0.1)
+```
+
+We can also compute the posterior probability of the slope for `condition` being above $0+\epsilon$ (@fig-post-prob-test), with $\epsilon$ defined as 10% of the standard deviation of the raw EEG signal.
+
+```{r fig-post-prob-test, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior probability of the ERP difference (slope) being above 0 according to the GAM (left) and the GP (right)."}
 # defining a function to plot posterior prob of slope above 0
-plot_post_test <- function (model, data = NULL, eps = 0.1) {
+plot_post_test <- function (model, data = NULL, eps = 0.1, n_approx = 1e3) {
 
@@
     time_seq <- crossing(
-        time = seq(min(data$time), max(data$time), length.out = 100),
+        time = seq(min(data$time), max(data$time), length.out = n_approx),
         condition = c("cond1", "cond2")
@@
     # computing the difference between cond2 and cond1
-    posterior_samples <- posterior_samples[, 101:200] - posterior_samples[, 1:100]
+    posterior_samples <- posterior_samples[, (nrow(time_seq)/2+1):nrow(time_seq)] -
+        posterior_samples[, 1:(nrow(time_seq)/2)]
 
@@
         time = time_seq$time,
-        m = colMeans(posterior_samples > (0 + eps) ),
-        lower = apply(
-            X = posterior_samples > (0 + eps),
-            MARGIN = 2, FUN = quantile, probs = 0.025
-            ),
-        upper = apply(
-            X = posterior_samples > (0 + eps),
-            MARGIN = 2, FUN = quantile, probs = 0.975
-            )
+        m = colMeans(posterior_samples > (0 + eps) )
+        # lower = apply(
+        #     X = posterior_samples > (0 + eps),
+        #     MARGIN = 2, FUN = quantile, probs = 0.025
+        #     ),
+        # upper = apply(
+        #     X = posterior_samples > (0 + eps),
+        #     MARGIN = 2, FUN = quantile, probs = 0.975
+        #     )
+        # )
         )
@@
         geom_line() +
+        ylim(c(0, 1) ) +
         labs(x = "Time (s)", y = expression(Pr(beta>0~"|"~data, model) ) )
@@
 
-We can also express this as the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional BF thresholds (@fig-post-prob-ratio).
-Also check this paper (<https://link.springer.com/article/10.3758/s13428-021-01613-6?fromPaywallRec=false>)?
-
-```{r fig-post-prob-ratio, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Ratio of posterior probability according to the GAM (left) and the GP (right).
-Timesteps above threshold (>10) are highlighted in green."}
+We can also express this as the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio).
+
+```{r fig-post-prob-ratio, echo = FALSE, fig.width = 8, fig.asp = 0.75, fig.cap = "Ratio of posterior probability according to the GAM. Timesteps above threshold (10) are highlighted in green."}
+# defining some parameters
+n_approx <- 1e3
+eps <- 0.1 * sd(ppt_df$eeg)
+# eps <- 0.1
+threshold <- 10
+true_peak <- 0.250
+
 # defining a sequence of x values to make predictions
-n_approx <- 1000
 time_seq <- crossing(
@@
 # computing the probability that y is above 0
-eps <- 0.05
 prob_y_above_0 <- data.frame(
@@
     mutate(m = colMeans(posterior_samples > eps) ) %>%
-    mutate(prob_ratio = m / (1 - m) )
+    mutate(prob_ratio = m / (1 - m) ) %>%
+    mutate(prob_ratio = ifelse(is.infinite(prob_ratio), 1e4, prob_ratio) ) %>%
+    mutate(prob_ratio = ifelse(prob_ratio == 0, 0.0001, prob_ratio) )
 
@@
 # a maximum at 250 ms, and an offset at 342 ms.
-threshold <- 10
 exceeding_times_gam <- prob_y_above_0 %>%
@@
         ) %>%
-    mutate(true_onset = 0.160, true_peak = 0.250, true_offset = 0.342) %>%
+    mutate(true_onset = true_onset, true_peak = true_peak, true_offset = true_offset) %>%
     mutate(model = "GAM")
+
+# creates grid for background gradient
+# Create a grid covering the range of time and log-transformed y-values
+bg_data <- expand.grid(
+    time = seq(min(prob_y_above_0$time), max(prob_y_above_0$time), length.out = 100),
+    # prob_ratio = 10^seq(log10(0.0001), log10(1e4), length.out = 100)
+    prob_ratio = 10^seq(log10(min(prob_y_above_0$prob_ratio) ), log10(max(prob_y_above_0$prob_ratio) ), length.out = 100)
+    )
+
+# assigning a fill value, for example, using prob_ratio or another variable
+bg_data$fill_value <- log10(bg_data$prob_ratio)
 
@@
 gam_ratio_plot <- prob_y_above_0 %>%
+# prob_y_above_0 %>%
     mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
     ggplot(aes(x = time, y = prob_ratio) ) +
+    geom_tile(
+        data = bg_data,
+        aes(x = time, y = prob_ratio, fill = fill_value),
+        alpha = 0.4,
+        inherit.aes = FALSE,
+        show.legend = FALSE
+        ) +
+    scale_fill_scico(palette = "vik", midpoint = log10(1) ) +
     geom_hline(yintercept = 1, linetype = "dashed") +
@@
     geom_vline(xintercept = true_offset, linetype = 2) +
-    geom_line() +
+    geom_line(linewidth = 0.8) +
     geom_point(
@@
         colour = "darkgreen",
-        shape = 15,
+        shape = 16,
         size = 3,
@@
         ) +
-    scale_y_log10(labels = label_log(digits = 2), limits = c(NA, 1e4) ) +
+    # scale_y_log10(labels = label_log(digits = 2), limits = c(0.0001, 1e4) ) +
+    scale_y_log10(
+        labels = label_log(digits = 2),
+        breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4),
+        limits = c(0.0001, 1e4)
+        ) +
+    coord_cartesian(expand = FALSE) +
     labs(
@@
 # retrieving posterior samples
-posterior_samples <- posterior_epred(object = gp_model, newdata = time_seq)
+# posterior_samples <- posterior_epred(object = gp_model, newdata = time_seq)
 
 # computing the difference between cond2 and cond1
-posterior_samples <- posterior_samples[, (ncol(posterior_samples)/2+1):ncol(posterior_samples)] -
-    posterior_samples[, 1:(ncol(posterior_samples)/2)]
+# posterior_samples <- posterior_samples[, (ncol(posterior_samples)/2+1):ncol(posterior_samples)] -
+#     posterior_samples[, 1:(ncol(posterior_samples)/2)]
 
 # computing the probability that y is above 0
-prob_y_above_0 <- data.frame(
-    time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx)
-    ) %>%
-    mutate(post_prob = colMeans(posterior_samples) ) %>%
-    mutate(m = colMeans(posterior_samples > eps) ) %>%
-    mutate(prob_ratio = m / (1 - m) )
+# prob_y_above_0 <- data.frame(
+#     time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx)
+#     ) %>%
+#     mutate(post_prob = colMeans(posterior_samples) ) %>%
+#     mutate(m = colMeans(posterior_samples > eps) ) %>%
+#     mutate(prob_ratio = m / (1 - m) )
 
@@
 # a maximum at 250 ms, and an offset at 342 ms
-exceeding_times_gp <- prob_y_above_0 %>%
-    dplyr::filter(prob_ratio > threshold) %>%
-    summarise(cluster_onset = min(time), cluster_offset = max(time) ) %>%
-    mutate(
-        cluster_peak = prob_y_above_0 %>%
-            dplyr::filter(post_prob == max(post_prob) ) %>%
-            pull(time)
-        ) %>%
-    mutate(true_onset = 0.160, true_peak = 0.250, true_offset = 0.342) %>%
-    mutate(model = "GP")
+# exceeding_times_gp <- prob_y_above_0 %>%
+#     dplyr::filter(prob_ratio > threshold) %>%
+#     summarise(cluster_onset = min(time), cluster_offset = max(time) ) %>%
+#     mutate(
+#         cluster_peak = prob_y_above_0 %>%
+#             dplyr::filter(post_prob == max(post_prob) ) %>%
+#             pull(time)
+#         ) %>%
+#     mutate(true_onset = true_onset, true_peak = true_peak, true_offset = true_offset) %>%
+#     mutate(model = "GP")
 
 # plotting it
-gp_ratio_plot <- prob_y_above_0 %>%
-    mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
-    ggplot(aes(x = time, y = prob_ratio) ) +
-    geom_hline(yintercept = 1, linetype = "dashed") +
-    geom_hline(yintercept = 3, linetype = "dashed", color = "darkred") +
-    geom_hline(yintercept = 1/3, linetype = "dashed", color = "darkred") +
-    geom_hline(yintercept = 10, linetype = "dashed", color = "red") +
-    geom_hline(yintercept = 1/10, linetype = "dashed", color = "red") +
-    geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
-    geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = true_offset, linetype = 2) +
-    geom_line() +
-    geom_point(
-        data = . %>% dplyr::filter(!is.na(above_thres) ),
-        aes(y = threshold),
-        colour = "darkgreen",
-        shape = 15,
-        size = 3,
-        na.rm = TRUE
-        ) +
-    scale_y_log10(labels = label_log(digits = 2), limits = c(NA, 1e4) ) +
-    labs(
-        x = "Time (s)",
-        y = expression(Pr(beta>0~"|"~data) / (1 - Pr(beta>0~"|"~data) ) )
-        )
+# gp_ratio_plot <- prob_y_above_0 %>%
+#     mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
+#     ggplot(aes(x = time, y = prob_ratio) ) +
+#     geom_hline(yintercept = 1, linetype = "dashed") +
+#     geom_hline(yintercept = 3, linetype = "dashed", color = "darkred") +
+#     geom_hline(yintercept = 1/3, linetype = "dashed", color = "darkred") +
+#     geom_hline(yintercept = 10, linetype = "dashed", color = "red") +
+#     geom_hline(yintercept = 1/10, linetype = "dashed", color = "red") +
+#     geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
+#     geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
+#     geom_vline(xintercept = true_onset, linetype = 2) +
+#     geom_vline(xintercept = true_offset, linetype = 2) +
+#     geom_line() +
+#     geom_point(
+#         data = . %>% dplyr::filter(!is.na(above_thres) ),
+#         aes(y = threshold),
+#         colour = "darkgreen",
+#         shape = 15,
+#         size = 3,
+#         na.rm = TRUE
+#         ) +
+#     scale_y_log10(labels = label_log(digits = 2), limits = c(NA, 1e4) ) +
+#     labs(
+#         x = "Time (s)",
+#         y = expression(Pr(beta>0~"|"~data) / (1 - Pr(beta>0~"|"~data) ) )
+#         )
 
 # plotting results for the two models
-gam_ratio_plot + gp_ratio_plot
+# gam_ratio_plot + gp_ratio_plot
+gam_ratio_plot
 
@@
 
+## Multilevel modelling using ERP summary statistics
+
+Next we fit a hierarchical/multilevel GAM using summary statistics of ERPs (mean and SD) at the participant level (similar to what is done in meta-analysis).
+
+```{r meta-gam, message = FALSE}
+# averaging across participants
+summary_df <- raw_df %>%
+    summarise(
+        eeg_mean = mean(eeg),
+        eeg_sd = sd(eeg),
+        .by = c(participant, condition, time)
+        )
+
+# defining a contrast for condition
+contrasts(summary_df$condition) <- c(-0.5, 0.5)
+
+# fitting the GAM
+meta_gam <- brm(
+    # using by-participant SD of ERPs across trials
+    eeg_mean | se(eeg_sd) ~
+        condition + s(time, bs = "cr", k = 10, by = condition) +
+        (1 | participant),
+    data = summary_df,
+    family = gaussian(),
+    warmup = 2000,
+    iter = 5000,
+    chains = 4,
+    cores = 4,
+    file = "models/meta_gam.rds"
+    )
+```
+
+```{r meta-gp, message = FALSE}
+# fitting the GP
+meta_gp <- brm(
+    # using by-participant SD of ERPs across trials
+    eeg_mean | se(eeg_sd) ~
+        condition + gp(time, k = 20, by = condition) +
+        (1 | participant),
+    data = summary_df,
+    family = gaussian(),
+    control = list(adapt_delta = 0.99),
+    iter = 5000,
+    chains = 4,
+    cores = 4,
+    file = "models/meta_gp.rds"
+    )
+```
+
+```{r meta-gp-preds, fig.width = 9, fig.asp = 0.5, fig.cap = "Posterior predictions for the hierarchical GAM and GP."}
+# plotting the posterior predictions
+meta_gam_preds <- plot(
+    conditional_effects(x = meta_gam, effect = "time:condition"),
+    points = FALSE, theme = theme_light(), plot = FALSE
+    )[[1]] +
+    scale_colour_met_d(name = "Johnson") +
+    scale_fill_met_d(name = "Johnson") +
+    labs(x = "Time (s)", y = "EEG signal (a.u.)")
+meta_gp_preds <- plot(
+    conditional_effects(x = meta_gp, effect = "time:condition"),
+    points = FALSE, theme = theme_light(), plot = FALSE
+    )[[1]] +
+    scale_colour_met_d(name = "Johnson") +
+    scale_fill_met_d(name = "Johnson") +
+    labs(x = "Time (s)", y = "EEG signal (a.u.)")
+
+# combining the two plots
+meta_gam_preds + meta_gp_preds
+```
+
 ## Error properties of the proposed approach
 
-Below we are plotting the error function (where the error is computed as $\text{error}:=|\hat{\theta}-\theta|$) for both the onset and offset values of the ERP difference, according to various `eps` and `threshold` values.
-Remember that the signal is generated from a truncated Gaussian defining an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
-@fig-onset-error shows that the GP model can *exactly* recover the true onset and offset values, given some reasonable choice of `eps` and `threshold` values.
-
-```{r fig-onset-error, echo = FALSE, dpi = 300, fig.width = 8, fig.asp = 0.5, fig.cap = "Error function of onset (left) and offset (right) estimation according to various eps and threshold values (according to the GP model).
+We then computed the difference between the true and estimated onset/offset of the ERP difference ($\text{error}:=|\hat{\theta}-\theta|$), according to various `eps` and `threshold` values.
+Remember that the signal is generated from a truncated Gaussian with an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
+@fig-onset-error shows that the hierarchical GAM can *exactly* recover the true onset and offset values, given some reasonable choice of `eps` and `threshold` values (e.g., a threshold of 20).
+
+```{r fig-onset-error, echo = FALSE, dpi = 300, fig.width = 8, fig.asp = 0.5, fig.cap = "Error function of onset (left) and offset (right) estimation according to various eps and threshold values (according to the hierarchical GAM).
 Minimum error values are indicated by red crosses."}
 # defining values of eps and threshold to test
-eps_values <- seq(from = 0, to = 0.2, by = 0.01)
-threshold_values <- seq(from = 1, to = 50, by = 1)
+eps_values <- seq(from = 0, to = 0.1, by = 0.01)
+threshold_values <- seq(from = 1, to = 30, by = 1)
 
 # using as many values of time as in the original data
-n_approx <- n_distinct(ppt_df$time)
+# n_approx <- n_distinct(ppt_df$time)
+
+# number of posterior samples to use
+n_post_samples <- 1e3
 
@@
 
+# which model should we use?
+error_model <- meta_gam
+
 # generating a grid of time values to compute predictions for
-time_seq <- crossing(
-    # participant = unique(ppt_df$participant),
-    condition = c("cond1", "cond2"),
-    time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx*2-1),
-    ) %>%
-    arrange(condition)
+# time_seq <- crossing(
+#     condition = c("cond1", "cond2"),
+#     time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx*2-1),
+#     ) %>%
+#     arrange(condition)
+
+# eps <- 0.05
+# preds <- meta_gam$data %>%
+#     add_epred_draws(meta_gam, ndraws = 1e3) %>%
+#     data.frame() %>%
+#     dplyr::select(participant, time, condition, .epred, .draw) %>%
+#     group_by(participant, time) %>%
+#     pivot_wider(names_from = condition, values_from = .epred) %>%
+#     mutate(epred_diff = cond2 - cond1) %>%
+#     summarise(m = mean(epred_diff > eps) ) %>%
+#     mutate(prob_ratio = m / (1 - m) ) %>%
+#     ungroup()
+
+# plotting these predictions
+# preds %>%
+#     ggplot(aes(x = time, y = m) ) +
+#     geom_line() +
+#     facet_wrap(~participant)
 
@@
 # posterior_samples <- posterior_epred(object = gp_model, newdata = time_seq)
-posterior_samples <- posterior_epred(object = gam, newdata = time_seq)
+# posterior_samples <- posterior_epred(object = gam, newdata = time_seq)
+# library(modelr)
+# meta_gam$data %>%
+#     add_epred_draws(meta_gam, ndraws = 20, re_formula = NA) %>%
+#     ggplot(
+#         aes(
+#             x = time, y = .epred,
+#             colour = condition,
+#             group = interaction(participant, condition, .draw)
+#             )
+#         ) +
+#     geom_hline(yintercept = 0, linetype = 2) +
+#     geom_vline(xintercept = true_onset, linetype = 2) +
+#     geom_vline(xintercept = true_offset, linetype = 2) +
+#     geom_line(show.legend = TRUE, alpha = 0.5, linewidth = 0.5) +
+#     facet_wrap(~participant) +
+#     labs(x = "Time (s)", y = "EEG signal (a.u.)")
 
 # computing the difference between cond2 and cond1
-posterior_samples <- posterior_samples[, (ncol(posterior_samples)/2+1):ncol(posterior_samples)] -
-    posterior_samples[, 1:(ncol(posterior_samples)/2)]
+# posterior_samples <- posterior_samples[, (ncol(posterior_samples)/2):ncol(posterior_samples)] -
+#     posterior_samples[, 1:(ncol(posterior_samples)/2+1)]
 
@@
 for (i in seq_len(nrow(results) ) ) {
+
+    # printing progress
+    # cat("Assessing combination:", i, "out of", nrow(results), "combinations...")
 
@@
     # computing probability metrics with the current eps value
-    prob_y_above_0 <- data.frame(
-        time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx*2-1)
-        ) %>%
-    mutate(post_prob = colMeans(posterior_samples) ) %>%
-    mutate(m = colMeans(posterior_samples > eps) ) %>%
-    mutate(prob_ratio = m / (1 - m) )
+    # prob_y_above_0 <- data.frame(
+    #     time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx*2-1)
+    #     ) %>%
+    # mutate(post_prob = colMeans(posterior_samples) ) %>%
+    # mutate(m = colMeans(posterior_samples > eps) ) %>%
+    # mutate(prob_ratio = m / (1 - m) )
+
+    prob_y_above_0 <- error_model$data %>%
+        add_epred_draws(object = error_model, ndraws = n_post_samples) %>%
+        data.frame() %>%
+        dplyr::select(participant, time, condition, .epred, .draw) %>%
+        # dplyr::select(time, condition, .epred, .draw) %>%
+        pivot_wider(names_from = condition, values_from = .epred) %>%
+        mutate(epred_diff = cond2 - cond1) %>%
+        # computing mean posterior prob at the participant level
+        group_by(participant, time) %>%
+        summarise(m = mean(epred_diff > eps) ) %>%
+        ungroup() %>%
+        # computing mean posterior prob at the group level
+        group_by(time) %>%
+        summarise(m = mean(m) ) %>%
+        mutate(prob_ratio = m / (1 - m) ) %>%
+        ungroup()
 
@@
             ) %>%
-        mutate(
-            cluster_peak = prob_y_above_0 %>%
-            dplyr::filter(post_prob == max(post_prob, na.rm = TRUE) ) %>%
-            pull(time)
-            ) %>%
-        mutate(true_onset = 0.160, true_peak = 0.250, true_offset = 0.342)
+        # mutate(
+        #     cluster_peak = prob_y_above_0 %>%
+        #     dplyr::filter(post_prob == max(post_prob, na.rm = TRUE) ) %>%
+        #     pull(time)
+        #     ) %>%
+        # mutate(true_onset = true_onset, true_peak = true_peak, true_offset = true_offset)
+        mutate(true_onset = true_onset, true_offset = true_offset)
 
@@
         results$estimated_onset[i] <- exceeding_times$cluster_onset
-        results$estimated_peak[i] <- exceeding_times$cluster_peak
+        # results$estimated_peak[i] <- exceeding_times$cluster_peak
         results$estimated_offset[i] <- exceeding_times$cluster_offset
@@
             )
-        results$error_peak[i] <- abs(
-            exceeding_times$cluster_peak - exceeding_times$true_peak
-            )
+        # results$error_peak[i] <- abs(
+        #     exceeding_times$cluster_peak - exceeding_times$true_peak
+        #     )
         results$error_offset[i] <- abs(
@@
 # displaying best parameters values (according to onset error)
+# results %>%
+#     dplyr::select(-estimated_peak, -error_peak) %>%
+#     arrange(error_onset) %>%
+#     data.frame() %>%
+#     head()
+
+# displaying best parameters values (according to offset error)
+# results %>%
+#     dplyr::select(-estimated_peak, -error_peak) %>%
+#     arrange(error_offset) %>%
+#     data.frame() %>%
+#     head()
+
+# displaying best parameters values (according to both onset and offset error)
 results %>%
     dplyr::select(-estimated_peak, -error_peak) %>%
-    arrange(error_onset) %>%
+    arrange(error_onset, error_offset) %>%
     data.frame() %>%
     head()
-
-# displaying best parameters values (according to offset error)
-results %>%
-    dplyr::select(-estimated_peak, -error_peak) %>%
-    arrange(error_offset) %>%
+```
+
+\newpage
+
+## Application to actual M/EEG data
+
+Assessing the reliability of the proposed approach (in comparison to other methods) using some sort of split-half reliability [@rosenblatt2018]?
+
+## Comparing the identified onsets/offsets to other approaches
+
+We considered five methods for multiplicity corrections: two FDR methods; two cluster-based methods; and the maximum statistics.
+The two FDR methods were BH95 (Benjamini & Hochberg, 1995; 10.1111/j.2517-6161.1995.tb02031.x) and BY01 (Benjamini & Yekutieli, 2001; 10.1214/aos/1013699998), which were applied to the permutation p-values using the `p.adjust` function in `R`.
+The first cluster-based inference was implemented using a cluster-sum statistic of squared t-values...
+
+We also compared the performance of the GAMM (in correctly identifying/estimating the onset and offset of ERP difference across conditions) to the performance of the binary segmentation method as implemented in `r pkrt("changepoint")`, threshold-free cluster enhancement [TFCE, @smith2009], as implemented in `MNE-Python` [@gramfort2013], and using the raw t-, p-, or s-values timecourse.
+S-values [@greenland2019]...
+
+<!--
+
+Comparing performance of the GAM and GP models with:
+
+-   uncorrected univariate tests
+-   univariate tests + Bonferroni/Holm
+-   univariate tests + BH95 (Benjamini-Hochberg, 1995) aka FDR
+-   univariate tests + BY01 (Benjamini-Yekutieli, 2001)
+-   cluster-based permutation test (cluster sum)
+-   cluster-based permutation test (cluster depth)
+-   threshold-free cluster enhancement
+-   change point
+
+We used `r pkrt("changepoint")`...
+
+-->
+
+```{r permutations-1d, echo = FALSE, eval = FALSE, fig.cap = "T-values timecourse with onset identified using the permuco package (clustermass method)."}
+# centering AUC relative to chance level (0.5)
+decoding_data_perm <- decoding_data %>%
+    mutate(auc_centered = auc - 0.5) %>%
+    dplyr::select(-auc)
+
+# formatting the AUC data to be used with permuco
+dv <- decoding_data_perm %>%
+    pivot_wider(names_from = time, values_from = auc_centered) %>%
+    dplyr::select(-participant) %>%
+    data.frame()
+
+# formatting the intercept to be used with permuco
+iv <- decoding_data %>%
+    dplyr::select(participant) %>%
+    distinct() %>%
+    mutate(participant = as.factor(participant) ) %>%
     data.frame() %>%
-    head()
-
-# displaying best parameters values (according to both onset/offset error)
-results %>%
-    filter(error_onset == min(error_onset) & error_offset == min(error_offset) ) %>%
-    data.frame()
-```
+    mutate(intercept = 1)
+
+# performing cluster-based permutation test
+cluster_results <- clusterlm(
+    formula = dv ~ intercept,
+    data = iv,
+    # data = decoding_data_perm,
+    # formula = auc_centered ~ participant,
+    # number of permutations (increase for accuracy)
+    # np = 2000,
+    # type of transformation
+    type = "signflip"
+    # using a t-test
+    # test = "t"
+    # multiple comparison correction
+    # multcomp = "clustermass"
+    # multcomp = "clusterdepth_head",
+    # multcomp = "tfce"
+    )
+
+# printing results
+print(cluster_results)
+
+# plotting significant clusters
+plot(cluster_results)
+
+# retrieving t and p-values
+f_values <- cluster_results$multiple_comparison$intercept$clustermass$main[, 1]
+p_values <- cluster_results$multiple_comparison$intercept$clustermass$main[, 2]
+```
+
+```{r test-1d, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Timecourse of squared t-values and s-values (continuous measure of evidence given by -log2(p-value)) with true onset and onset identified using the raw (uncorrected) p-values or the corrected p-values (BH, BY, Holm)."}
+# https://github.com/GRousselet/onsetsim/blob/main/docs/weakstrongfwer_demo.md
+# BH95: p.adjust(pvals, method = "BH")
+# BY01: p.adjust(perm.pvals, method = "BY")
+
+# defining an arbitrary alpha threshold
+aath <- 0.05
+
+# summarising raw_data per participant
+ppt_data <- raw_df %>%
+    summarise(eeg = mean(eeg), .by = c(participant, condition, time) ) %>%
+    pivot_wider(names_from = condition, values_from = eeg) %>%
+    mutate(eeg_diff = cond2 - cond1)
+
+# massive univariate t-tests
+tests_results <- ppt_data %>%
+    group_by(time) %>%
+    summarise(
+        tval = t.test(x = eeg_diff, mu = 0)$statistic^2,
+        pval = t.test(x = eeg_diff, mu = 0)$p.value
+        ) %>%
+    mutate(
+        pval_bh = p.adjust(p = pval, method = "BH"),
+        pval_by = p.adjust(p = pval, method = "BY"),
+        # pval_bonf = p.adjust(p = pval, method = "bonferroni"),
+        pval_holm = p.adjust(p = pval, method = "holm")
+        ) %>%
+    ungroup()
+
+# finding onsets
+from_timestep <- 1
+onset_p <- find_onset(
+    tests_results$pval[from_timestep:nrow(tests_results)] <= aath,
+    tests_results$time[from_timestep:nrow(tests_results)]
+    )
+onset_bh <- find_onset(
+    tests_results$pval_bh[from_timestep:nrow(tests_results)] <= aath,
+    tests_results$time[from_timestep:nrow(tests_results)]
+    )
+onset_by <- find_onset(
+    tests_results$pval_by[from_timestep:nrow(tests_results)] <= aath,
+    tests_results$time[from_timestep:nrow(tests_results)]
+    )
+# onset_bonf <- find_onset(
+#     tests_results$pval_bonf[from_timestep:nrow(tests_results)] <= aath,
+#     tests_results$time[from_timestep:nrow(tests_results)]
+#     )
+onset_holm <- find_onset(
+    tests_results$pval_holm[from_timestep:nrow(tests_results)] <= aath,
+    tests_results$time[from_timestep:nrow(tests_results)]
+    )
+
+# plotting the results
+default_colours <- hue_pal()(10)
+tvals_plot <- tests_results %>%
+    ggplot(aes(x = time, y = tval) ) +
+    geom_area(position = "identity") +
+    geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = onset_p, linetype = 2, color = default_colours[1]) +
+    geom_vline(xintercept = onset_bh, linetype = 2, color = default_colours[2]) +
+    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[4]) +
+    # geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
+    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[5]) +
+    labs(x = "Time (s)", y = bquote(t^2) )
+
+plot_title <- paste0(
+    "True onset: ", true_onset, "s, ",
+    "Uncorrected onset: ", round(onset_p, 3), "s, ",
+    "BH onset: ", round(onset_bh, 3), "s, ",
+    "BY onset: ", round(onset_by, 3), "s, ",
+    # "Bonf onset: ", round(onset_bonf, 3), "s, ",
+    "Holm onset: ", round(onset_holm, 3),  "s"
+    )
+
+# plotting -log2(p) (surprisal under H0) with thresholds (on p-values)
+# https://lesslikely.com/statistics/s-values/
+# a s-value of 4 (bits of information) is no more surprising than
+# getting all heads on 4 fair coin tosses
+pvals_plot <- tests_results %>%
+    pivot_longer(cols = pval:pval_holm) %>%
+    mutate(sval = -log2(value) ) %>%
+    ggplot(aes(x = time, y = sval, colour = name) ) +
+    geom_line(linewidth = 0.5) +
+    geom_hline(yintercept = -log2(aath), linetype = 2) +
+    geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = onset_p, linetype = 2, color = default_colours[1]) +
+    geom_vline(xintercept = onset_bh, linetype = 2, color = default_colours[2]) +
+    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[4]) +
+    # geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
+    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[5]) +
+    labs(
+        x = "Time (s)",
+        y = "-log2(p-value)",
+        colour = "correction"
+        )
+
+# combining the two plots
+tvals_plot + pvals_plot +
+    plot_annotation(title = plot_title) & theme(plot.title = element_text(hjust = 0.5) )
+```
+
+```{r changepoint, echo = FALSE, out.width = "75%", fig.cap = "T-values timecourse with true onset/offset and onset/offset identified using the changepoint package (binary segmentation method, in green)."}
+# using the changepoint package to identify onsets and offsets
+res <- cpt.meanvar(data = tests_results$tval, method = "BinSeg", Q = 2)
+
+# plotting the results
+tests_results %>%
+    ggplot(aes(x = time, y = tval) ) +
+    geom_line(linewidth = 1) +
+    geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = true_offset, linetype = 2) +
+    geom_vline(
+        xintercept = tests_results$time[res@cpts[1]],
+        linetype = 3, color = "darkgreen"
+        ) +
+    geom_vline(
+        xintercept = tests_results$time[res@cpts[2]],
+        linetype = 3, color = "darkgreen"
+        ) +
+    labs(x = "Time (s)", y = bquote(t^2) ) +
+    ggtitle(
+        paste0(
+            "True onset: ", true_onset, "s, ", "Change point onset: ",
+            round(tests_results$time[res@cpts[1]], 3), "s, ",
+            "True offset: ", true_offset, "s, ", "Change point offset: ",
+            round(tests_results$time[res@cpts[2]], 3), "s"
+            )
+        )
+```
+
+Now cluster-based permutations using `MNE-Python` [@gramfort2013] via `r pkrt("reticulate")`...
+
+```{r mne-cluster, echo = FALSE, eval = FALSE, fig.cap = "Cluster-based permutation tests via MNE-Python."}
+# loading the reticulate package
+library(reticulate)
+
+# importing the numpy and mne python modules
+np <- import("numpy")
+mne <- import("mne")
+
+# defining the function in R (it will be executed in Python)
+freq_stats_gat_matrix <- function (X) {
+
+    # converting R matrix to NumPy array
+    X_np <- np$array(X, dtype = "float64")
+
+    # running the statistical test
+    results <- mne$stats$spatio_temporal_cluster_1samp_test(
+        X_np,
+        out_type = "mask",
+        n_permutations = as.integer(1000),
+        n_jobs = as.integer(4),
+        verbose = TRUE
+        )
+
+    # extracting results
+    T_obs_ <- results[[1]]
+    clusters <- results[[2]]
+    p_values <- results[[3]]
+
+    # using the first slice of the 3D array
+    p_values_ <- np$transpose(np$ones_like(X_np[1, , drop = FALSE]) )
+
+    # assigning p-values to clusters
+    for (i in seq_along(clusters) ) {
+
+        # retrieving the current cluster
+        cluster_mask <- clusters[[i]][[1]]
+        idx_start <- cluster_mask$start+1
+        idx_stop <- cluster_mask$stop
+
+        # retrieving the p-value for this cluster
+        pval <- p_values[[i]]
+
+        # assigning this p-value to timesteps belonging
+        # to the current cluster
+        p_values_[idx_start:idx_stop] <- pval
+
+    }
+
+    # converting result back to R format
+    return (as.matrix(np$squeeze(np$transpose(p_values_) ) ) )
+
+}
+
+# converting decoding_data to a matrix
+data_matrix <- matrix(
+    decoding_data$auc,
+    ncol = length(unique(decoding_data$time) ),
+    byrow = TRUE
+    )
+
+# running the function
+chance_level <- 0.5
+p_values_results <- freq_stats_gat_matrix(data_matrix-chance_level)
+
+# converting back to a dataframe if needed
+p_values_df <- data.frame(
+    pval = p_values_results,
+    time = decoding_data$time
+    )
+
+# saving the scores
+saveRDS(object = p_values_df, file = "results/mne_permutation_decoding_scores.rds")
+```
+
+```{r fig-mne-cluster, out.width = "75%", fig.cap = "Cluster-based permutation tests via MNE-Python."}
+# importing the decoding scores
+p_values_df <- readRDS(file = "results/mne_permutation_decoding_scores.rds")
+
+# plotting the s-values (-log2(p-values))
+p_values_df %>%
+    mutate(sval = -log2(pval) ) %>%
+    ggplot(aes(x = time, y = sval) ) +
+    geom_line(linewidth = 0.5) +
+    geom_hline(yintercept = -log2(0.05), linetype = 2) +
+    labs(
+        x = "Time (s)",
+        y = "-log2(p-value)"
+        )
+```
+
+## Simulation study
+
+Onset/offset estimation methods were assessed using these summary statistics of the Monte Carlo sampling distributions of onsets: bias (median-based), mean absolute error (MAE) and variance, with 10.000 iterations (i.e., 10.000 simulated datasets)...
 
@@
 
-## Using summary statistics of ERPs
-
-Next we fit a hierarchical GAM using summary statistics of ERPs (mean and SD) at the participant level (similar to what is done in meta-analysis).
-
-```{r meta-gam, message = FALSE}
-# averaging across participants
-summary_df <- raw_df %>%
-    summarise(
-        eeg_mean = mean(eeg),
-        eeg_sd = sd(eeg),
-        .by = c(participant, condition, time)
-        )
-
-# defining a contrast for condition
-contrasts(summary_df$condition) <- c(-0.5, 0.5)
-
-# fitting the GAM
-meta_gam <- brm(
-    # using by-participant SD of ERPs across trials
-    eeg_mean | se(eeg_sd) ~
-        condition + s(time, bs = "cr", k = 10, by = condition) +
-        (1 | participant),
-    data = summary_df,
-    family = gaussian(),
-    iter = 5000,
-    chains = 4,
-    cores = 4,
-    file = "models/meta_gam.rds"
-    )
-```
-
-```{r meta-gp, message = FALSE}
-# fitting the GP
-meta_gp <- brm(
-    # using by-participant SD of ERPs across trials
-    eeg_mean | se(eeg_sd) ~
-        condition + gp(time, k = 20, by = condition) +
-        (1 | participant),
-    data = summary_df,
-    family = gaussian(),
-    control = list(adapt_delta = 0.99),
-    iter = 5000,
-    chains = 4,
-    cores = 4,
-    file = "models/meta_gp.rds"
-    )
-```
-
-```{r meta-gp-preds, fig.width = 9, fig.asp = 0.5, fig.cap = "Posterior predictions for the hierarchical GAM and GP."}
-# plotting the posterior predictions
-meta_gam_preds <- plot(
-    conditional_effects(x = meta_gam, effect = "time:condition"),
-    points = FALSE, theme = theme_light(), plot = FALSE
-    )[[1]] +
-    labs(x = "Time (s)", y = "EEG signal (a.u.)")
-meta_gp_preds <- plot(
-    conditional_effects(x = meta_gp, effect = "time:condition"),
-    points = FALSE, theme = theme_light(), plot = FALSE
-    )[[1]] +
-    labs(x = "Time (s)", y = "EEG signal (a.u.)")
-
-# combining the two plots
-meta_gam_preds + meta_gp_preds
-```
-
-## Application to time-resolved decoding results (accuracy over time)
+# Results
+
+...
+
+\newpage
+
+# Discussion
+
+...
+
+## Summary of the proposed approach
+
+...
+
+## Increasing potential usage
+
+Prepare a wrapper `R` package and show how to call it in `Python` and integrate it with `MNE-Python` [@gramfort2013] pipelines...
+
+## Limitations and future directions
+
+Can be applied to any 1D timeseries (e.g., pupillometry, electromyography)... Extending the approach to spatiotemporal data (i.e., time + sensors)...
+
+## Conclusions
+
+...
+
+\newpage
+
+# Packages
+
+```{r, echo = FALSE}
+# citing the R packages we have used
+cite_packages(
+    output = "paragraph",
+    omit = c("permuco", "reticulate"),
+    out.dir = "."
+    )
+```
+
+\newpage
+
+# References
+
+::: {#refs}
+:::
+
+\newpage
+
+# Appendix
+
+# Application to time-resolved decoding results (accuracy over time)
 
@@
 
-## Application to actual M/EEG data
-
-Assessing the reliability of the proposed approach (in comparison to other methods) using some sort of split-half reliability [@rosenblatt2018]?
-
-## Comparing the identified onsets/offsets to other approaches
-
-From Rousselet (2024): Five methods for multiple comparison correction were considered: two FDR methods; two cluster-based methods; and the maximum statistics.
-The two FDR methods were BH95 (Benjamini & Hochberg, 1995) and BY01 (Benjamini & Yekutieli, 2001), which were applied to the permutation p-values using the `p.adjust` function in `R`.
-The first cluster-based inference was implemented using a cluster-sum statistic of squared t-values...
-
-Comparing performance of the GAM and GP models with:
-
--   uncorrected univariate tests
--   univariate tests + Bonferroni/Holm
--   univariate tests + BH95 (Benjamini-Hochberg, 1995) aka FDR
--   univariate tests + BY01 (Benjamini-Yekutieli, 2001)
--   cluster-based permutation test (cluster sum)
--   cluster-based permutation test (cluster depth)
--   threshold-free cluster enhancement
--   change point
-
-We used `r pkrt("changepoint")`...
-
-```{r permutations-1d, echo = FALSE, eval = FALSE, fig.cap = "T-values timecourse with onset identified using the permuco package (clustermass method)."}
-# centering AUC relative to chance level (0.5)
-decoding_data_perm <- decoding_data %>%
-    mutate(auc_centered = auc - 0.5) %>%
-    dplyr::select(-auc)
-
-# formatting the AUC data to be used with permuco
-dv <- decoding_data_perm %>%
-    pivot_wider(names_from = time, values_from = auc_centered) %>%
-    dplyr::select(-participant) %>%
-    data.frame()
-
-# formatting the intercept to be used with permuco
-iv <- decoding_data %>%
-    dplyr::select(participant) %>%
-    distinct() %>%
-    mutate(participant = as.factor(participant) ) %>%
-    data.frame() %>%
-    mutate(intercept = 1)
-
-# performing cluster-based permutation test
-cluster_results <- clusterlm(
-    formula = dv ~ intercept,
-    data = iv,
-    # data = decoding_data_perm,
-    # formula = auc_centered ~ participant,
-    # number of permutations (increase for accuracy)
-    # np = 2000,
-    # type of transformation
-    type = "signflip"
-    # using a t-test
-    # test = "t"
-    # multiple comparison correction
-    # multcomp = "clustermass"
-    # multcomp = "clusterdepth_head",
-    # multcomp = "tfce"
-    )
-
-# printing results
-print(cluster_results)
-
-# plotting significant clusters
-plot(cluster_results)
-
-# retrieving t and p-values
-f_values <- cluster_results$multiple_comparison$intercept$clustermass$main[, 1]
-p_values <- cluster_results$multiple_comparison$intercept$clustermass$main[, 2]
-```
-
-S-values [@greenland2019]...
-
-```{r test-1d, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Timecourse of squared t-values and s-values (continuous measure of evidence given by -log2(p-value)) with true onset and onset identified using the raw (uncorrected) p-values or the corrected p-values (BH, BY, Holm)."}
-# https://github.com/GRousselet/onsetsim/blob/main/docs/weakstrongfwer_demo.md
-# BH95: p.adjust(pvals, method = "BH")
-# BY01: p.adjust(perm.pvals, method = "BY")
-
-# defining an arbitrary alpha threshold
-aath <- 0.05
-
-# summarising raw_data per participant
-ppt_data <- raw_df %>%
-    summarise(eeg = mean(eeg), .by = c(participant, condition, time) ) %>%
-    pivot_wider(names_from = condition, values_from = eeg) %>%
-    mutate(eeg_diff = cond2 - cond1)
-
-# massive univariate t-tests
-tests_results <- ppt_data %>%
-    group_by(time) %>%
-    summarise(
-        tval = t.test(x = eeg_diff, mu = 0)$statistic^2,
-        pval = t.test(x = eeg_diff, mu = 0)$p.value
-        ) %>%
-    mutate(
-        pval_bh = p.adjust(p = pval, method = "BH"),
-        pval_by = p.adjust(p = pval, method = "BY"),
-        # pval_bonf = p.adjust(p = pval, method = "bonferroni"),
-        pval_holm = p.adjust(p = pval, method = "holm")
-        ) %>%
-    ungroup()
-
-# finding onsets
-from_timestep <- 1
-onset_p <- find_onset(
-    tests_results$pval[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$time[from_timestep:nrow(tests_results)]
-    )
-onset_bh <- find_onset(
-    tests_results$pval_bh[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$time[from_timestep:nrow(tests_results)]
-    )
-onset_by <- find_onset(
-    tests_results$pval_by[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$time[from_timestep:nrow(tests_results)]
-    )
-# onset_bonf <- find_onset(
-#     tests_results$pval_bonf[from_timestep:nrow(tests_results)] <= aath,
-#     tests_results$time[from_timestep:nrow(tests_results)]
-#     )
-onset_holm <- find_onset(
-    tests_results$pval_holm[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$time[from_timestep:nrow(tests_results)]
-    )
-
-# plotting the results
-default_colours <- hue_pal()(10)
-tvals_plot <- tests_results %>%
-    ggplot(aes(x = time, y = tval) ) +
-    geom_area(position = "identity") +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = onset_p, linetype = 2, color = default_colours[1]) +
-    geom_vline(xintercept = onset_bh, linetype = 2, color = default_colours[2]) +
-    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[4]) +
-    # geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
-    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[5]) +
-    labs(x = "Time (s)", y = bquote(t^2) )
-
-plot_title <- paste0(
-    "True onset: ", true_onset, "s, ",
-    "Uncorrected onset: ", round(onset_p, 3), "s, ",
-    "BH onset: ", round(onset_bh, 3), "s, ",
-    "BY onset: ", round(onset_by, 3), "s, ",
-    # "Bonf onset: ", round(onset_bonf, 3), "s, ",
-    "Holm onset: ", round(onset_holm, 3),  "s"
-    )
-
-# plotting -log2(p) (surprisal under H0) with thresholds (on p-values)
-# https://lesslikely.com/statistics/s-values/
-# a s-value of 4 (bits of information) is no more surprising than
-# getting all heads on 4 fair coin tosses
-pvals_plot <- tests_results %>%
-    pivot_longer(cols = pval:pval_holm) %>%
-    mutate(sval = -log2(value) ) %>%
-    ggplot(aes(x = time, y = sval, colour = name) ) +
-    geom_line(linewidth = 0.5) +
-    geom_hline(yintercept = -log2(aath), linetype = 2) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = onset_p, linetype = 2, color = default_colours[1]) +
-    geom_vline(xintercept = onset_bh, linetype = 2, color = default_colours[2]) +
-    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[4]) +
-    # geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
-    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[5]) +
-    labs(
-        x = "Time (s)",
-        y = "-log2(p-value)",
-        colour = "correction"
-        )
-
-# combining the two plots
-tvals_plot + pvals_plot +
-    plot_annotation(title = plot_title) & theme(plot.title = element_text(hjust = 0.5) )
-```
-
-```{r changepoint, echo = FALSE, out.width = "75%", fig.cap = "T-values timecourse with true onset/offset and onset/offset identified using the changepoint package (binary segmentation method, in green)."}
-# using the changepoint package to identify onsets and offsets
-res <- cpt.meanvar(data = tests_results$tval, method = "BinSeg", Q = 2)
-
-# plotting the results
-tests_results %>%
-    ggplot(aes(x = time, y = tval) ) +
-    geom_line(linewidth = 1) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = true_offset, linetype = 2) +
-    geom_vline(
-        xintercept = tests_results$time[res@cpts[1]],
-        linetype = 3, color = "darkgreen"
-        ) +
-    geom_vline(
-        xintercept = tests_results$time[res@cpts[2]],
-        linetype = 3, color = "darkgreen"
-        ) +
-    labs(x = "Time (s)", y = bquote(t^2) ) +
-    ggtitle(
-        paste0(
-            "True onset: ", true_onset, "s, ", "Change point onset: ",
-            round(tests_results$time[res@cpts[1]], 3), "s, ",
-            "True offset: ", true_offset, "s, ", "Change point offset: ",
-            round(tests_results$time[res@cpts[2]], 3), "s"
-            )
-        )
-```
-
-Now cluster-based permutations using `MNE-Python` [@gramfort2013] via `r pkrt("reticulate")`...
-
-```{r mne-cluster, echo = FALSE, eval = FALSE, fig.cap = "Cluster-based permutation tests via MNE-Python."}
-# loading the reticulate package
-library(reticulate)
-
-# importing the numpy and mne python modules
-np <- import("numpy")
-mne <- import("mne")
-
-# defining the function in R (it will be executed in Python)
-freq_stats_gat_matrix <- function (X) {
-
-    # converting R matrix to NumPy array
-    X_np <- np$array(X, dtype = "float64")
-
-    # running the statistical test
-    results <- mne$stats$spatio_temporal_cluster_1samp_test(
-        X_np,
-        out_type = "mask",
-        n_permutations = as.integer(1000),
-        n_jobs = as.integer(4),
-        verbose = TRUE
-        )
-
-    # extracting results
-    T_obs_ <- results[[1]]
-    clusters <- results[[2]]
-    p_values <- results[[3]]
-
-    # using the first slice of the 3D array
-    p_values_ <- np$transpose(np$ones_like(X_np[1, , drop = FALSE]) )
-
-    # assigning p-values to clusters
-    for (i in seq_along(clusters) ) {
-
-        # retrieving the current cluster
-        cluster_mask <- clusters[[i]][[1]]
-        idx_start <- cluster_mask$start+1
-        idx_stop <- cluster_mask$stop
-
-        # retrieving the p-value for this cluster
-        pval <- p_values[[i]]
-
-        # assigning this p-value to timesteps belonging
-        # to the current cluster
-        p_values_[idx_start:idx_stop] <- pval
-
-    }
-
-    # converting result back to R format
-    return (as.matrix(np$squeeze(np$transpose(p_values_) ) ) )
-
-}
-
-# converting decoding_data to a matrix
-data_matrix <- matrix(
-    decoding_data$auc,
-    ncol = length(unique(decoding_data$time) ),
-    byrow = TRUE
-    )
-
-# running the function
-chance_level <- 0.5
-p_values_results <- freq_stats_gat_matrix(data_matrix-chance_level)
-
-# converting back to a dataframe if needed
-p_values_df <- data.frame(
-    pval = p_values_results,
-    time = decoding_data$time
-    )
-
-# saving the scores
-saveRDS(object = p_values_df, file = "results/mne_permutation_decoding_scores.rds")
-```
-
-```{r fig-mne-cluster, out.width = "75%", fig.cap = "Cluster-based permutation tests via MNE-Python."}
-# importing the decoding scores
-p_values_df <- readRDS(file = "results/mne_permutation_decoding_scores.rds")
-
-# plotting the s-values (-log2(p-values))
-p_values_df %>%
-    mutate(sval = -log2(pval) ) %>%
-    ggplot(aes(x = time, y = sval) ) +
-    geom_line(linewidth = 0.5) +
-    geom_hline(yintercept = -log2(0.05), linetype = 2) +
-    labs(
-        x = "Time (s)",
-        y = "-log2(p-value)"
-        )
-```
-
 \newpage
-
-# Results
-
-...
-
-\newpage
-
-# Discussion
-
-...
-
-## Summary of the proposed approach
-
-...
-
-## Increasing potential usage
-
-Prepare a wrapper `R` package and show how to call it in `Python` and integrate it with `MNE-Python` [@gramfort2013] pipelines...
-
-## Limitations and future directions
-
-Extending the approach to spatiotemporal data (i.e., time + sensors)...
-
-## Conclusions
-
-...
-
-\newpage
-
-# Packages
-
-```{r, echo = FALSE}
-# citing the R packages we have used
-cite_packages(
-    output = "paragraph",
-    # omit = c("permuco", "brms", "reticulate"),
-    omit = c("permuco", "reticulate"),
-    out.dir = "."
-    )
-```
-
-\newpage
-
-# References
-
-::: {#refs}
-:::
-
-\newpage
-
-# Appendix
 
@@
     family = Beta(),
-    iter = 2000,
+    iter = 5000,
     chains = 4,

### 2025-02-10 brms_meeg.tex
reshaping the document and drafting a script to conduct proper simulations of onset/offset recovery

@@
 
-\title{Modelling M/EEG data with Bayesian nonparametric multilevel
-models}
+\title{Modelling M/EEG data with Bayesian multilevel generalised
+additive models}
 
@@
 
-\abstract{Time-resolved electrophysoiological measurements such as those
+\abstract{Time-resolved electrophysiological measurements such as those
 offered by magneto- or electro-encephalography (M/EEG) provide a unique
@@
 measures differ across conditions and/or groups. The conventional
-approach consists in conducting mass-univariate statistics followed by
-some form of multiplicity correction (e.g., FDR, FWER) or cluster-based
-inference. However, these cluster-based methods have an important
-downside: they shift the focus of inference from the timepoint to the
-cluster level, thus preventing any conclusion to be made about the onset
-and offset of effects (e.g., differences across conditions). Here, we
-introduce a novel \emph{model-based approch} for analysing
+approach consists in conducting mass-univariate statistics through
+followed by some form of multiplicity correction (e.g., FDR, FWER) or
+cluster-based inference. However, these cluster-based methods have an
+important downside: they shift the focus of inference from the timepoint
+to the cluster level, thus preventing any conclusion to be made about
+the onset and offset of effects (e.g., differences across conditions).
+Here, we introduce a novel \emph{model-based approch} for analysing
 one-dimensional M/EEG timeseries such as ERPs or decoding timecourses
-and their differences across conditions. This approach relies on
-Bayesian nonparametric multilevel modelling (multilevel generalised
-additive models or Gaussian processes), which outputs posterior
-probabililities of the effect being above chance at every
-timestep/voxel, while naturally taking into account the temporal
+and their differences across conditions or group. This approach relies
+on Bayesian nonparametric multilevel modelling (multilevel generalised
+additive models or Gaussian processes), which outputs the posterior
+probabilility of the effect being different than 0 (or above chance) at
+every timestep, while naturally taking into account the temporal
 dependencies and between-subject variability present in such data.}
@@
 \newpage
-\section[Introduction]{Modelling M/EEG data with Bayesian nonparametric
-multilevel models}
+\section[Introduction]{Modelling M/EEG data with Bayesian multilevel
+generalised additive models}
 
@@
 
+From Dimigen \& Ehinger (\citeproc{ref-dimigen2021}{2021}): ``Recently,
+spline regression has been applied to ERPs (Hendrix, Baayen, \& Bolger,
+2017; Kryuchkova,Tucker,Wurm,\&Baayen,2012;Tremblay\& Baayen, 2010;
+Tremblay \& Newman, 2015)''\ldots{}
+
 Disentangling overlapping processes
@@
 et al., 2006}; \citeproc{ref-rousselet2008}{Rousselet et al.,
-2008})\ldots{} Example of two-stage regression analysis
-(individual-level then group-level, 10.1101/2024.04.10.588861,
-\citeproc{ref-dunagan2024}{Dunagan et al., 2024})\ldots{}
+2008})\ldots{} Example of two-stage regression analysis (i.e.,
+individual-level then group-level, \citeproc{ref-dunagan2024}{Dunagan et
+al., 2024})\ldots{}
 
@@
 coefficients are penalised\ldots{}
-
-\subsection{Gaussian process
-regression}\label{gaussian-process-regression}
-
-A Gaussian process (GP) is a stochastic process that defines the
-distribution over a collection of random variables indexed by a
-continuous variable, that is \(\{f(t): t \in \mathcal{T}\}\) for some
-index set \(\mathcal{T}\) (\citeproc{ref-rasmussen2005}{Rasmussen \&
-Williams, 2005};
-\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
-2023}). Whereas Bayesian linear regression outputs a distribution over
-the parameters of some predefind parametric model, the GP approach, in
-contrast, is a non-parametric approach, in that it finds a distribution
-over the possible functions that are consistent with the observed data.
-However, note that nonparametric does not mean there aren't parameters,
-it means that there are infinitely many parameters.
-
-From
-\href{https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp}{brms
-documentation}: A GP is a stochastic process, which describes the
-relation between one or more predictors
-\(x=\left(x_{1}, \ldots, x_{d}\right)\) and a response \(f(x)\), where
-\(d\) is the number of predictors. A GP is the generalization of the
-multivariate normal distribution to an infinite number of dimensions.
-Thus, it can be interpreted as a prior over functions. The values of
-\(f()\) at any finite set of locations are jointly multivariate normal,
-with a covariance matrix defined by the covariance kernel
-\(k_p\left(x_i, x_j\right)\), where \(p\) is the vector of parameters of
-the GP:
-
-\[
-\left(f\left(x_{1}\right), \ldots f\left(x_{n}\right) \sim \operatorname{MVN}\left(0, \left(k_p\left(x_{i}, x_{j}\right)\right)_{i, j=1}^{n}\right)\right.
-\]
-
-The smoothness and general behaviour of the function \(f\) depends only
-on the choice of covariance kernel, which ensures that values that are
-close together in the input space will be mapped to similar output
-values\ldots{}
-
-From this perspective, \(f\) is a realisation of an infinite dimensional
-normal distribution:
-
-\[
-f \sim \mathrm{Normal}(0, C(\lambda))
-\]
-
-where \(C\) is a covariance kernel with hyperparameters \(\lambda\) that
-defines the covariance between two function values \(f\left(t_1\right)\)
-and \(f\left(t_2\right)\) for two time points \(t_1\) and \(t_2\)
-(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). Similar to
-the different choices of the basis function for splines, different
-choices of the covariance kernel lead to different GPs. In this article,
-we consider the squared-exponential (a.k.a. radial basis function)
-kernel, which computes the squared distance between points and converts
-it into a measure of similarity. It is defined as:
-
-\[
-C(\lambda) := C\left(t_1, t_2, \sigma, \gamma\right) := \sigma^2 \exp \left(-\frac{||t_1-t_2||^{2}}{2 \gamma^2}\right)
-\]
-
-with hyperparameters \(\lambda = (\sigma, \gamma)\), expressing the
-overall scale of GP and the length-scale, respectively
-(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). The
-advantages of this kernel are that it is computationally efficient and
-(infinitely) smooth making it a reasonable choice for the purposes of
-the present article. Here again, \(\lambda\) hyperparameters are
-estimated from the data, along with all other model parameters.
-
-Taken from
-\url{https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10c-Gaussian-processes.html}:
-For a given vector \(\mathbf{x}\), we can use the kernel to construct
-finite multi-variate normal distribution associated with it like so:
-
-\[
-\mathbf{x} \mapsto_{G P} \operatorname{MVNormal}(m(\mathbf{x}), k(\mathbf{x}, \mathbf{x}))
-\]
-
-where \(m\) is a function that specifies the mean for the distribution
-associated with \(\mathbf{x}\). This mapping is essentially the Gaussian
-process: a systematic association of vectors of arbitrary length with a
-suitable multi-variate normal distribution.
-
-See also
-\url{https://jtimonen.github.io/lgpr-usage/articles/math.html}\ldots{}
-
-Low-rank approximate Gaussian processes are of main interest in machine
-learning and statistics due to the high computational demands of exact
-Gaussian process models
-(\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
-2023})\ldots{}
 
@@
 \caption{\label{fig-erp}Group-level average difference between
-conditions (mean +/- standard error of the mean). The `true' (population
-value of the) onset is indicated by the vertical dashed line.}
+conditions (mean +/- standard error of the mean). The `true' onset and
+offset are indicated by the vertical dashed lines.}
 
@@
 \citeproc{ref-brms2018}{2018}; \citeproc{ref-nalborczyk2019}{Nalborczyk
-et al., 2019})\ldots{}
+et al., 2019}). We used the default priors in \texttt{brms}, that is,
+weakly informative priors. Four Markov Chain Monte-Carlo (MCMC) were ran
+for each model to approximate the posterior distribution, including each
+5000 iterations and a warmup of 2000 iterations. Posterior convergence
+was assessed examining trace plots as well as the Gelman--Rubin statisti
+\(\hat{R}\).
 
@@
     \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
+    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
     \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
@@
 
-Now we fit the GP\ldots{}
+Now we fit the GP regression model\ldots{}
 
@@
 We can retrieve the posterior probability of the slope for
-\texttt{condition} at each timestep.
-
-\begin{figure}[H]
-
-\caption{Slope for the difference between conditions according to the
-GAM (left) or the GP (right).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/unnamed-chunk-1-1.pdf}
+\texttt{condition} at each timestep (Figure~\ref{fig-plot-post-slope}).
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-plot-post-slope}Slope for the difference between
+conditions according to the GAM (left) or the GP (right).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-plot-post-slope-1.pdf}
 
@@
 We can also compute the posterior probability of the slope for
-\texttt{condition} being above 0 (Figure~\ref{fig-post-prob-test}).
+\texttt{condition} being above \(0+\epsilon\)
+(Figure~\ref{fig-post-prob-test}), with \(\epsilon\) defined as 10\% of
+the standard deviation of the raw EEG signal.
 
@@
 
-\caption{\label{fig-post-prob-test}Posterior probability of the EEG
-signal being above 0+eps according to the GAM (left) and the GP
+\caption{\label{fig-post-prob-test}Posterior probability of the ERP
+difference (slope) being above 0 according to the GAM (left) and the GP
 (right).}
@@
 \(p/(1-p)\)) and visualise the timecourse of this ratio superimposed
-with the conventional BF thresholds (Figure~\ref{fig-post-prob-ratio}).
-Also check this paper
-(\url{https://link.springer.com/article/10.3758/s13428-021-01613-6?fromPaywallRec=false})?
+with the conventional thresholds on evidence ratios
+(Figure~\ref{fig-post-prob-ratio}).
 
@@
 \caption{\label{fig-post-prob-ratio}Ratio of posterior probability
-according to the GAM (left) and the GP (right). Timesteps above
-threshold (\textgreater10) are highlighted in green.}
+according to the GAM. Timesteps above threshold (10) are highlighted in
+green.}
 
@@
 
-\subsection{Error properties of the proposed
-approach}\label{error-properties-of-the-proposed-approach}
-
-Below we are plotting the error function (where the error is computed as
-\(\text{error}:=|\hat{\theta}-\theta|\)) for both the onset and offset
-values of the ERP difference, according to various \texttt{eps} and
-\texttt{threshold} values. Remember that the signal is generated from a
-truncated Gaussian defining an objective onset at 160 ms, a maximum at
-250 ms, and an offset at 342 ms. Figure~\ref{fig-onset-error} shows that
-the GP model can \emph{exactly} recover the true onset and offset
-values, given some reasonable choice of \texttt{eps} and
-\texttt{threshold} values.
-
-\begin{verbatim}
-   eps threshold estimated_onset estimated_offset error_onset error_offset
-1 0.03         6            0.16            0.351           0        0.009
-2 0.03         7            0.16            0.351           0        0.009
-3 0.03         8            0.16            0.351           0        0.009
-4 0.03         9            0.16            0.351           0        0.009
-5 0.03        10            0.16            0.351           0        0.009
-6 0.03        11            0.16            0.351           0        0.009
-\end{verbatim}
-
-\begin{verbatim}
-   eps threshold estimated_onset estimated_offset error_onset error_offset
-1 0.07         8           0.166            0.342       0.006            0
-2 0.07         9           0.166            0.342       0.006            0
-3 0.07        10           0.166            0.342       0.006            0
-4 0.07        11           0.166            0.342       0.006            0
-5 0.07        12           0.166            0.342       0.006            0
-6 0.07        13           0.166            0.342       0.006            0
-\end{verbatim}
-
-\begin{verbatim}
-[1] eps              threshold        estimated_onset  estimated_peak  
-[5] estimated_offset error_onset      error_peak       error_offset    
-<0 rows> (or 0-length row.names)
-\end{verbatim}
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-onset-error}Error function of onset (left) and
-offset (right) estimation according to various eps and threshold values
-(according to the GP model). Minimum error values are indicated by red
-crosses.}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-error-1.pdf}
-
-}
-
-\end{figure}%
-
-\newpage
-
-\subsection{Using summary statistics of
-ERPs}\label{using-summary-statistics-of-erps}
-
-Next we fit a hierarchical GAM using summary statistics of ERPs (mean
-and SD) at the participant level (similar to what is done in
+\subsection{Multilevel modelling using ERP summary
+statistics}\label{multilevel-modelling-using-erp-summary-statistics}
+
+Next we fit a hierarchical/multilevel GAM using summary statistics of
+ERPs (mean and SD) at the participant level (similar to what is done in
 meta-analysis).
@@
 \NormalTok{    )[[}\DecValTok{1}\NormalTok{]] }\SpecialCharTok{+}
+    \FunctionTok{scale\_colour\_met\_d}\NormalTok{(}\AttributeTok{name =} \StringTok{"Johnson"}\NormalTok{) }\SpecialCharTok{+}
+    \FunctionTok{scale\_fill\_met\_d}\NormalTok{(}\AttributeTok{name =} \StringTok{"Johnson"}\NormalTok{) }\SpecialCharTok{+}
     \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (s)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG signal (a.u.)"}\NormalTok{)}
@@
 
-\subsection{Application to time-resolved decoding results (accuracy over
-time)}\label{application-to-time-resolved-decoding-results-accuracy-over-time}
-
-We conducted time-resolved multivariate pattern analysis (MVPA), also
-known as decoding. As a result, we have a timecourse of decoding
-accuracies (e.g., ROC AUC), bounded between 0 and 1, per participant
-(Figure~\ref{fig-mne-decoding})\ldots{}
+\newpage
+
+\subsection{Error properties of the proposed
+approach}\label{error-properties-of-the-proposed-approach}
+
+We then computed the difference between the true and estimated
+onset/offset of the ERP difference
+(\(\text{error}:=|\hat{\theta}-\theta|\)), according to various
+\texttt{eps} and \texttt{threshold} values. Remember that the signal is
+generated from a truncated Gaussian with an objective onset at 160 ms, a
+maximum at 250 ms, and an offset at 342 ms. Figure~\ref{fig-onset-error}
+shows that the hierarchical GAM can \emph{exactly} recover the true
+onset and offset values, given some reasonable choice of \texttt{eps}
+and \texttt{threshold} values (e.g., a threshold of 20).
+
+\begin{verbatim}
+   eps threshold estimated_onset estimated_offset error_onset error_offset
+1 0.00        13            0.16             0.34           0        0.002
+2 0.00        14            0.16             0.34           0        0.002
+3 0.00        15            0.16             0.34           0        0.002
+4 0.00        16            0.16             0.34           0        0.002
+5 0.00        17            0.16             0.34           0        0.002
+6 0.01        10            0.16             0.34           0        0.002
+\end{verbatim}
 
@@
 
-\caption{\label{fig-mne-decoding}Exemplary average timecourse of binary
-decoding accuracy (ROC AUC) for each participant. Group-level average
-decoding accuracy is depicted as a grey background density in each
-panel.}
+\caption{\label{fig-onset-error}Error function of onset (left) and
+offset (right) estimation according to various eps and threshold values
+(according to the hierarchical GAM). Minimum error values are indicated
+by red crosses.}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-mne-decoding-1.pdf}
-
-}
-
-\end{figure}%
-
-Now, we want to \emph{test} whether the group-level average decoding
-accuracy is above chance (i.e., 0.5) at each timestep. We use a similar
-GAM/GP as previously, but we replace the \(\mathrm{Normal}\) likelihood
-function by a \(\mathrm{Beta}\) one to account for the bounded nature of
-AUC values (between 0 and 1).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# fitting the GAM}
-\NormalTok{decoding\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
-    \AttributeTok{data =}\NormalTok{ decoding\_data,}
-    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/decoding\_gam.rds"}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# fitting the GP}
-\NormalTok{decoding\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gp}\NormalTok{(time, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
-    \AttributeTok{data =}\NormalTok{ decoding\_data,}
-    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{),}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/decoding\_gp.rds"}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-\caption{Posterior predictions of the GAM (left) and GP (right) fitted
-on decoding accuracy over time.}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/decoding-preds-1.pdf}
-
-}
-
-\end{figure}%
-
-Next, we plot the posterior probability of decoding accuracy being above
-chance level (plus some epsilon) (Figure~\ref{fig-decoding-post}).
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-decoding-post}Posterior probability of decoding
-accuracy being above chance level according to the GAM (left) or the GP
-(right).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-post-1.pdf}
-
-}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-decoding-ratio}Ratio of posterior probabilities of
-decoding accuracy being above chance level according to the GAM (left)
-or the GP (right).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-ratio-1.pdf}
-
-}
-
-\end{figure}%
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-error-1.pdf}
+
+}
+
+\end{figure}%
+
+\newpage
 
@@
 
-From Rousselet (2024): Five methods for multiple comparison correction
-were considered: two FDR methods; two cluster-based methods; and the
-maximum statistics. The two FDR methods were BH95 (Benjamini \&
-Hochberg, 1995) and BY01 (Benjamini \& Yekutieli, 2001), which were
-applied to the permutation p-values using the \texttt{p.adjust} function
-in \texttt{R}. The first cluster-based inference was implemented using a
-cluster-sum statistic of squared t-values\ldots{}
-
-Comparing performance of the GAM and GP models with:
-
-\begin{itemize}
-\tightlist
-\item
-  uncorrected univariate tests
-\item
-  univariate tests + Bonferroni/Holm
-\item
-  univariate tests + BH95 (Benjamini-Hochberg, 1995) aka FDR
-\item
-  univariate tests + BY01 (Benjamini-Yekutieli, 2001)
-\item
-  cluster-based permutation test (cluster sum)
-\item
-  cluster-based permutation test (cluster depth)
-\item
-  threshold-free cluster enhancement
-\item
-  change point
-\end{itemize}
-
-We used the \texttt{R} package \texttt{changepoint} v 2.2.4
-(\citeproc{ref-changepoint}{Killick et al., 2022a})\ldots{}
-
-S-values (\citeproc{ref-greenland2019}{Greenland, 2019})\ldots{}
+We considered five methods for multiplicity corrections: two FDR
+methods; two cluster-based methods; and the maximum statistics. The two
+FDR methods were BH95 (Benjamini \& Hochberg, 1995;
+10.1111/j.2517-6161.1995.tb02031.x) and BY01 (Benjamini \& Yekutieli,
+2001; 10.1214/aos/1013699998), which were applied to the permutation
+p-values using the \texttt{p.adjust} function in \texttt{R}. The first
+cluster-based inference was implemented using a cluster-sum statistic of
+squared t-values\ldots{}
+
+We also compared the performance of the GAMM (in correctly
+identifying/estimating the onset and offset of ERP difference across
+conditions) to the performance of the binary segmentation method as
+implemented in the \texttt{R} package \texttt{changepoint} v 2.2.4
+(\citeproc{ref-changepoint}{Killick et al., 2022a}), threshold-free
+cluster enhancement (TFCE, \citeproc{ref-smith2009}{Smith \& Nichols,
+2009}), as implemented in \texttt{MNE-Python}
+(\citeproc{ref-gramfort2013}{Gramfort, 2013}), and using the raw t-, p-,
+or s-values timecourse. S-values
+(\citeproc{ref-greenland2019}{Greenland, 2019})\ldots{}
 
@@
 
+\subsection{Simulation study}\label{simulation-study}
+
+Onset/offset estimation methods were assessed using these summary
+statistics of the Monte Carlo sampling distributions of onsets: bias
+(median-based), mean absolute error (MAE) and variance, with 10.000
+iterations (i.e., 10.000 simulated datasets)\ldots{}
+
 \newpage
@@
 
-Extending the approach to spatiotemporal data (i.e., time +
-sensors)\ldots{}
+Can be applied to any 1D timeseries (e.g., pupillometry,
+electromyography)\ldots{} Extending the approach to spatiotemporal data
+(i.e., time + sensors)\ldots{}
 
@@
 
+\bibitem[\citeproctext]{ref-dimigen2021}
+Dimigen, O., \& Ehinger, B. V. (2021). Regression-based analysis of
+combined EEG and eye-tracking data: Theory and applications.
+\emph{Journal of Vision}, \emph{21}(1), 3.
+\url{https://doi.org/10.1167/jov.21.1.3}
+
 \bibitem[\citeproctext]{ref-dunagan2024}
@@
 \appendix
+
+\section{Application to time-resolved decoding results (accuracy over
+time)}\label{application-to-time-resolved-decoding-results-accuracy-over-time}
+
+We conducted time-resolved multivariate pattern analysis (MVPA), also
+known as decoding. As a result, we have a timecourse of decoding
+accuracies (e.g., ROC AUC), bounded between 0 and 1, per participant
+(Figure~\ref{fig-mne-decoding})\ldots{}
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-mne-decoding}Exemplary average timecourse of binary
+decoding accuracy (ROC AUC) for each participant. Group-level average
+decoding accuracy is depicted as a grey background density in each
+panel.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-mne-decoding-1.pdf}
+
+}
+
+\end{figure}%
+
+Now, we want to \emph{test} whether the group-level average decoding
+accuracy is above chance (i.e., 0.5) at each timestep. We use a similar
+GAM/GP as previously, but we replace the \(\mathrm{Normal}\) likelihood
+function by a \(\mathrm{Beta}\) one to account for the bounded nature of
+AUC values (between 0 and 1).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# fitting the GAM}
+\NormalTok{decoding\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
+\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
+    \AttributeTok{data =}\NormalTok{ decoding\_data,}
+    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
+    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
+    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{file =} \StringTok{"models/decoding\_gam.rds"}
+\NormalTok{    )}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# fitting the GP}
+\NormalTok{decoding\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
+\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gp}\NormalTok{(time, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
+    \AttributeTok{data =}\NormalTok{ decoding\_data,}
+    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
+    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{),}
+    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
+    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{file =} \StringTok{"models/decoding\_gp.rds"}
+\NormalTok{    )}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[H]
+
+\caption{Posterior predictions of the GAM (left) and GP (right) fitted
+on decoding accuracy over time.}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/decoding-preds-1.pdf}
+
+}
+
+\end{figure}%
+
+Next, we plot the posterior probability of decoding accuracy being above
+chance level (plus some epsilon) (Figure~\ref{fig-decoding-post}).
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-decoding-post}Posterior probability of decoding
+accuracy being above chance level according to the GAM (left) or the GP
+(right).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-post-1.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-decoding-ratio}Ratio of posterior probabilities of
+decoding accuracy being above chance level according to the GAM (left)
+or the GP (right).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-ratio-1.pdf}
+
+}
+
+\end{figure}%
+
+\newpage
 
@@
     \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
+    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
     \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}

### 2025-02-10 code/errors_simulation.R

@@ -2,14 +2,15 @@
 # Monte-Carlo simulation of onset/offset error properties #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on February 9, 2025                        #
+# Last updated on February 10, 2025                       #
 ###########################################################
 
 library(changepoint)
+library(reticulate)
 library(tidyverse)
 library(brms)
 
-# retrieving the code from the qdm document
+# retrieving the code from the Quarto document
 # knitr::purl(input = "brms_meeg.qmd", output = "brms_meeg.R", documentation = 0)
 
 # importing R version of Matlab code from Yeung et al. (2004)
@@ -24,1235 +25,296 @@ source("functions.R")
 # to use with the eeg_noise function
 meanpower <- unlist(read.table("meanpower.txt") )
 
+# defining the number of simulations to be performed
+nsims <- 10
+
 # defining simulation parameters
 n_trials <- 50 # number of trials
 n_ppt <- 20 # number of participants
 outvar <- 1 # noise variance
 srate <- 500 # sampling rate in Hz
-ronset <- seq(150, 170, 2) # random onset for each participant
-
-# removing raw_df
-if (exists("raw_df") ) rm(raw_df)
-
-for (P in 1:n_ppt) { # for each participant
-    
-    # get random onset
-    ponset <- sample(ronset, 1)
-    
-    # find starting point
-    st <- which(Xf == ponset)
-    
-    # pad vector
-    temp2 <- c(rep(0, st - 2), erp, rep(0, Nf - st - length(erp) + 2) )
-    
-    # initialising empty conditions
-    cond1 <- matrix(0, nrow = n_trials, ncol = Nf)
-    cond2 <- matrix(0, nrow = n_trials, ncol = Nf)
-    
-    for (T in 1:n_trials) { # for each trial
-        
-        cond1[T, ] <- temp1 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
-        cond2[T, ] <- temp2 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
-        
-    }
-    
-    # converting results to dataframe
-    temp_df <- data.frame(
-        x = rep(Xf, 2*nrow(cond1) ),
-        y = c(c(t(cond1) ), c(t(cond2) ) ),
[…]

### 2025-02-13 brms_meeg.qmd
pushing some preliminary simulation results

@@
 ---
-title: "Modelling M/EEG data with Bayesian multilevel generalised additive models"
+title: "Modelling M/EEG data with Bayesian generalised additive multilevel models"
 shorttitle: "Bayesian M/EEG modelling"
@@
 Here, we introduce a novel *model-based approch* for analysing one-dimensional M/EEG timeseries such as ERPs or decoding timecourses and their differences across conditions or group.
-This approach relies on Bayesian nonparametric multilevel modelling (multilevel generalised additive models), which outputs the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between-subject variability present in such data."
-keywords: [EEG, MEG, generalised additive models, gaussian processes, mixed-effects models, Bayesian statistics, brms]
+This approach relies on Bayesian generalised additive multilevel models, which output the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between-subject variability present in such data."
+keywords: [EEG, MEG, generalised additive models, mixed-effects models, multilevel models, Bayesian statistics, brms]
 floatsintext: true
@@
 library(grateful)
+library(ggrepel)
 library(scales)
@@
 
-# Introduction (in progress)
-
-Here are some useful references to be discussed [@combrisson_exceeding_2015; @ehinger_unfold_2019; @hayasaka_validating_2003; @luck_how_2017; @pedersen_hierarchical_2019; @riutort-mayol_practical_2023; @rousselet_using_2025; @frossard2021; @frossard2022; @gramfort2013; @maris2007; @pernet2015]...
+# Introduction
+
+Here are some useful references to be discussed [@combrisson_exceeding_2015; @ehinger_unfold_2019; @hayasaka_validating_2003; @luck_how_2017; @pedersen_hierarchical_2019; @frossard2021; @frossard2022; @gramfort2013; @maris2007; @pernet2015]...
 See also [@maris2011]...
@@
 
-In the following, we consider two approaches to modelling the non-linear timecourse of M/EEG ERPs or decoding performance: i) generalised additive models (GAMs) with thin-plate smoothing splines [@wood2003; @wood2004] and ii) Gaussian processes (GPs) with a smooth covariance kernel [@rasmussen2005] and low-rank approximation [@riutort-mayol_practical_2023].
+```{=html}
+<!--
+
+In the following, we consider two approaches to modelling the non-linear timecourse of M/EEG ERPs (or decoding performance): i) generalised additive models (GAMs) with smoothing splines [@wood2003; @wood2004]...
+
+and ii) Gaussian processes (GPs) with a smooth covariance kernel [@rasmussen2005] and low-rank approximation [@riutort-mayol_practical_2023].
+
+-->
+```
 
 ## Previous modelling work
-
-From @dimigen2021: "Recently, spline regression has been applied to ERPs (Hendrix, Baayen, & Bolger, 2017; Kryuchkova,Tucker,Wurm,&Baayen,2012;Tremblay& Baayen, 2010; Tremblay & Newman, 2015)"...
-
-Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]... Using Bayes factors [@teichmann2022]... Weighting single trials [@pernet2022]...
-
-See <https://www.let.rug.nl/nerbonne/teach/rema-stats-meth-seminar/presentations/Wieling-2014-GAMs.pdf> and <https://www.let.rug.nl/wieling/Statistics/GAM-EEG/lab/>, <https://www.let.rug.nl/wieling/Statistics/GAM-EEG/GAM-EEG.pdf> [@abugaber2023; @meulman2015]...
 
@@
 
+From @dimigen2021: "Recently, spline regression has been applied to ERPs (Hendrix, Baayen, & Bolger, 2017; Kryuchkova,Tucker, Wurm, & Baayen,2012; Tremblay & Baayen, 2010; Tremblay & Newman, 2015)"...
+GAMMs for EEG data [@abugaber2023; @meulman2015]...
+
+Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]...
+Weighting single trials [@pernet2022]...
+The LIMO toolbox (https://onlinelibrary.wiley.com/doi/10.1155/2011/831409)...
+Using Bayes factors [@teichmann2022]...
+
 ## Generalised additive models
+
+See for instance these tutorials [@sóskuthy2017; @winter2016] or application to phonetic data [@wieling2018] or this introduction [@baayen2020] or these reference books [@hastie2017; @wood2017]...
 
@@
 
-...
+Focusing on identifying onset and offset of effects (as assessed by ERP differences or decoding performance)...
+Assessing the performance of a model-based approach (i.e., Bayesian GAMMs) to conventional methods (multiplicity corrections or cluster-based permutation)...
 
@@
 ## M/EEG data simulation
-
-```{=html}
-<!--
-
-For simulating EEG data, see <https://github.com/GRousselet/onsetsim> and <https://github.com/unfoldtoolbox/UnfoldSim.jl>...
-
--->
-```
 
@@
 
-We then fitted a GAM and GP regression model using the `brms` package [@brms2017; @brms2018; @nalborczyk2019].
+We then fitted a Bayesian GAM using the `brms` package [@brms2017; @brms2018; @nalborczyk2019].
 We used the default priors in `brms`, that is, weakly informative priors.
-Four Markov Chain Monte-Carlo (MCMC) were ran for each model to approximate the posterior distribution, including each 5000 iterations and a warmup of 2000 iterations.
-Posterior convergence was assessed examining trace plots as well as the Gelman–Rubin statisti $\hat{R}$.
+We ran eight Markov Chain Monte-Carlo (MCMC) to approximate the posterior distribution, including each 5000 iterations and a warmup of 2000 iterations, yielding a total of $8 \times (5000-2000) = 24000$ posterior samples to use for inference.
+Posterior convergence was assessed examining trace plots as well as the Gelman–Rubin statistic $\hat{R}$.
 
@@
     summarise(eeg = mean(eeg) ) %>%
-    # mutate(eeg = as.numeric(scale(x = eeg, scale = FALSE) ) ) %>%
     ungroup()
@@
     iter = 5000,
-    chains = 4,
-    cores = 4,
+    chains = 8,
+    cores = 8,
     file = "models/gam.rds"
@@
 
-Now we fit the GP regression model...
-
-```{r gp, message = FALSE}
+```{r gp, echo = FALSE, eval = FALSE, message = FALSE}
+# Now we fit the GP regression model...
 # computing eeg_diff and averaging across participants
@@
 
-And we plot the posterior predictions...
-
-```{r gam-preds, echo = FALSE}
+```{r model-preds, echo = FALSE}
 # defining a function to plot posterior predictions
@@
 }
-
-```
-
-```{r gam-gp-preds, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior predictions of the GAM (left) and GP (right) models."}
+```
+
+```{r gam-preds, echo = FALSE, eval = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior predictions of the GAM (left) and GP (right) models."}
 # plotting the posterior predictions
 plot_post_preds(model = gam) + plot_post_preds(model = gp_model)
-
 ```
@@
 
-We can retrieve the posterior probability of the slope for `condition` at each timestep (@fig-plot-post-slope).
-
-```{r fig-plot-post-slope, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Slope for the difference between conditions according to the GAM (left) or the GP (right)."}
+We then plot the posterior predictions together with the posterior estimate of the slope for `condition` at each timestep (@fig-plot-post-slope).
+
+```{r fig-plot-post-slope, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior estimate of the ERP in each condition (left) or directly for the difference of ERPs (right) according to the GAM."}
 # defining a function to plot posterior slope
-plot_post_slope <- function (model, data = NULL, eps = 0.1) {
+plot_post_slope <- function (model, data = NULL) {
 
@@
         scale_fill_met_d(name = "Johnson") +
-        labs(x = "Time (s)", y = expression(Pr(beta~"|"~data, model) ) )
+        labs(
+            x = "Time (s)",
+            y = expression(p(beta~"|"~data, model) )
+            )
 
@@
 
-# using the function with model
-plot_post_slope(model = gam, eps = 0.1) + plot_post_slope(model = gp_model, eps = 0.1)
-```
-
-We can also compute the posterior probability of the slope for `condition` being above $0+\epsilon$ (@fig-post-prob-test), with $\epsilon$ defined as 10% of the standard deviation of the raw EEG signal.
-
-```{r fig-post-prob-test, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior probability of the ERP difference (slope) being above 0 according to the GAM (left) and the GP (right)."}
+# plotting it
+# plot_post_slope(model = gam, eps = 0.1) + plot_post_slope(model = gp_model, eps = 0.1)
+plot_post_preds(model = gam) + plot_post_slope(model = gam)
+```
+
+We then compute the posterior probability of the slope for `condition` being above $0+\epsilon$ (@fig-post-prob-test), with $\epsilon := 0.05$, which can be interpreted as the smallest effect size of interest.
+
+```{r fig-post-prob-test, echo = FALSE, out.width = "75%", fig.asp = 0.5, fig.cap = "Posterior probability of the ERP difference (slope) being above 0 according to the GAM."}
 # defining a function to plot posterior prob of slope above 0
-plot_post_test <- function (model, data = NULL, eps = 0.1, n_approx = 1e3) {
+plot_post_test <- function (model, data = NULL, eps = 0.1, n_approx = 1e3, n_draws = 1e3) {
 
@@
         time = seq(min(data$time), max(data$time), length.out = n_approx),
+        # time = unique(data$time),
         condition = c("cond1", "cond2")
@@
     # retrieving posterior samples
-    posterior_samples <- posterior_epred(object = model, newdata = time_seq)
+    # posterior_samples <- posterior_epred(object = model, newdata = time_seq)
 
     # computing the difference between cond2 and cond1
-    posterior_samples <- posterior_samples[, (nrow(time_seq)/2+1):nrow(time_seq)] -
-        posterior_samples[, 1:(nrow(time_seq)/2)]
+    # posterior_samples <- posterior_samples[, (nrow(time_seq)/2+1):nrow(time_seq)] -
+    #     posterior_samples[, 1:(nrow(time_seq)/2)]
 
     # computing the probability that y is above 0
-    prob_y_above_0 <- data.frame(
-        time = time_seq$time,
-        m = colMeans(posterior_samples > (0 + eps) )
-        # lower = apply(
-        #     X = posterior_samples > (0 + eps),
-        #     MARGIN = 2, FUN = quantile, probs = 0.025
-        #     ),
-        # upper = apply(
-        #     X = posterior_samples > (0 + eps),
-        #     MARGIN = 2, FUN = quantile, probs = 0.975
+    # prob_y_above_0 <- data.frame(
+    #     time = time_seq$time,
+    #     m = colMeans(posterior_samples > (0 + eps) )
     #     )
-        # )
-        )
+
+    # computing probability metrics with the current eps value
+    prob_y_above_0 <- time_seq %>%
+        add_epred_draws(object = model, ndraws = n_draws) %>%
+        data.frame() %>%
+        dplyr::select(time, condition, .epred, .draw) %>%
+        pivot_wider(names_from = condition, values_from = .epred) %>%
+        mutate(epred_diff = cond2 - cond1) %>%
+        group_by(time) %>%
+        summarise(m = mean(epred_diff > 0 + eps) ) %>%
+        ungroup()
 
@@
 # using the function with the two models
-plot_post_test(model = gam, eps = 0.1) +
-    plot_post_test(model = gp_model, eps = 0.1)
+plot_post_test(model = gam, eps = 0.1)
 ```
@@
 n_approx <- 1e3
-eps <- 0.1 * sd(ppt_df$eeg)
-# eps <- 0.1
+eps <- 0.05
 threshold <- 10
-true_peak <- 0.250
 
@@
     time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx),
+    # time = unique(ppt_df$time),
     condition = c("cond1", "cond2")
@@
 
-# retrieving posterior samples
-posterior_samples <- posterior_epred(object = gam, newdata = time_seq)
-
-# computing the difference between cond2 and cond1
-posterior_samples <- posterior_samples[, (ncol(posterior_samples)/2+1):ncol(posterior_samples)] -
-    posterior_samples[, 1:(ncol(posterior_samples)/2)]
-
-# computing the probability that y is above 0
-prob_y_above_0 <- data.frame(
-    time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx)
-    ) %>%
-    mutate(post_prob = colMeans(posterior_samples) ) %>%
-    mutate(m = colMeans(posterior_samples > eps) ) %>%
+# computing the posterior probability that y is above 0
+n_posterior_samples <- 24000
+prob_y_above_0 <- time_seq %>%
+    add_epred_draws(object = gam) %>%
+    data.frame() %>%
+    dplyr::select(time, condition, .epred, .draw) %>%
+    pivot_wider(names_from = condition, values_from = .epred) %>%
+    mutate(epred_diff = cond2 - cond1) %>%
+    group_by(time) %>%
+    summarise(m = mean(epred_diff > 0 + eps) ) %>%
+    ungroup() %>%
     mutate(prob_ratio = m / (1 - m) ) %>%
-    mutate(prob_ratio = ifelse(is.infinite(prob_ratio), 1e4, prob_ratio) ) %>%
-    mutate(prob_ratio = ifelse(prob_ratio == 0, 0.0001, prob_ratio) )
+    # replacing infinite values by maximum possible values (number of posterior samples)
+    mutate(prob_ratio = ifelse(is.infinite(prob_ratio), n_posterior_samples, prob_ratio) ) %>%
+    mutate(prob_ratio = ifelse(prob_ratio == 0, 1/n_posterior_samples, prob_ratio) )
 
@@
 # a maximum at 250 ms, and an offset at 342 ms.
-exceeding_times_gam <- prob_y_above_0 %>%
-    dplyr::filter(prob_ratio > threshold) %>%
-    summarise(cluster_onset = min(time), cluster_offset = max(time) ) %>%
-    mutate(
-        cluster_peak = prob_y_above_0 %>%
-            dplyr::filter(post_prob == max(post_prob) ) %>%
-            pull(time)
-        ) %>%
-    mutate(true_onset = true_onset, true_peak = true_peak, true_offset = true_offset) %>%
-    mutate(model = "GAM")
-
-# creates grid for background gradient
-# Create a grid covering the range of time and log-transformed y-values
+# exceeding_times_gam <- prob_y_above_0 %>%
+#     dplyr::filter(prob_ratio > threshold) %>%
+#     summarise(cluster_onset = min(time), cluster_offset = max(time) ) %>%
+#     # mutate(
+#     #     cluster_peak = prob_y_above_0 %>%
+#     #         dplyr::filter(post_prob == max(post_prob) ) %>%
+#     #         pull(time)
+#     #     ) %>%
+#     # mutate(true_onset = true_onset, true_peak = true_peak, true_offset = true_offset) %>%
+#     mutate(true_onset = true_onset, true_offset = true_offset) %>%
+#     mutate(model = "GAM")
+
+# creating a grid covering the range of time and log-transformed y-values
 bg_data <- expand.grid(
@@
 gam_ratio_plot <- prob_y_above_0 %>%
-# prob_y_above_0 %>%
     mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
     ggplot(aes(x = time, y = prob_ratio) ) +
-    geom_tile(
+    geom_raster(
         data = bg_data,
         aes(x = time, y = prob_ratio, fill = fill_value),
-        alpha = 0.4,
+        alpha = 0.5,
+        interpolate = FALSE,
         inherit.aes = FALSE,
@@
         ) +
-    # scale_y_log10(labels = label_log(digits = 2), limits = c(0.0001, 1e4) ) +
     scale_y_log10(
         labels = label_log(digits = 2),
-        breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4),
-        limits = c(0.0001, 1e4)
+        breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
+        # limits = c(0.0001, 1e4)
         ) +
@@
 # gam_ratio_plot + gp_ratio_plot
+
+# or plotting only the GAM
 gam_ratio_plot
@@
     iter = 5000,
-    chains = 4,
-    cores = 4,
+    chains = 8,
+    cores = 8,
     file = "models/meta_gam.rds"
@@
 
-```{r meta-gp, message = FALSE}
+```{r meta-gp, echo = FALSE, eval = FALSE, message = FALSE}
 # fitting the GP
@@
 
-```{r meta-gp-preds, fig.width = 9, fig.asp = 0.5, fig.cap = "Posterior predictions for the hierarchical GAM and GP."}
+```{r meta-preds, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Posterior predictions for the hierarchical GAM."}
 # plotting the posterior predictions
@@
     labs(x = "Time (s)", y = "EEG signal (a.u.)")
-meta_gp_preds <- plot(
-    conditional_effects(x = meta_gp, effect = "time:condition"),
-    points = FALSE, theme = theme_light(), plot = FALSE
-    )[[1]] +
-    scale_colour_met_d(name = "Johnson") +
-    scale_fill_met_d(name = "Johnson") +
-    labs(x = "Time (s)", y = "EEG signal (a.u.)")
+# meta_gp_preds <- plot(
+#     conditional_effects(x = meta_gp, effect = "time:condition"),
+#     points = FALSE, theme = theme_light(), plot = FALSE
+#     )[[1]] +
+#     scale_colour_met_d(name = "Johnson") +
+#     scale_fill_met_d(name = "Johnson") +
+#     labs(x = "Time (s)", y = "EEG signal (a.u.)")
 
 # combining the two plots
-meta_gam_preds + meta_gp_preds
+# meta_gam_preds + meta_gp_preds
 ```
@@
 
-## Application to actual M/EEG data
-
-Assessing the reliability of the proposed approach (in comparison to other methods) using some sort of split-half reliability [@rosenblatt2018]?
-
 ## Comparing the identified onsets/offsets to other approaches
 
-We considered five methods for multiplicity corrections: two FDR methods; two cluster-based methods; and the maximum statistics.
-The two FDR methods were BH95 (Benjamini & Hochberg, 1995; 10.1111/j.2517-6161.1995.tb02031.x) and BY01 (Benjamini & Yekutieli, 2001; 10.1214/aos/1013699998), which were applied to the permutation p-values using the `p.adjust` function in `R`.
-The first cluster-based inference was implemented using a cluster-sum statistic of squared t-values...
-
-We also compared the performance of the GAMM (in correctly identifying/estimating the onset and offset of ERP difference across conditions) to the performance of the binary segmentation method as implemented in `r pkrt("changepoint")`, threshold-free cluster enhancement [TFCE, @smith2009], as implemented in `MNE-Python` [@gramfort2013], and using the raw t-, p-, or s-values timecourse.
-S-values [@greenland2019]...
-
-<!--
-
-Comparing performance of the GAM and GP models with:
-
--   uncorrected univariate tests
--   univariate tests + Bonferroni/Holm
--   univariate tests + BH95 (Benjamini-Hochberg, 1995) aka FDR
--   univariate tests + BY01 (Benjamini-Yekutieli, 2001)
--   cluster-based permutation test (cluster sum)
--   cluster-based permutation test (cluster depth)
--   threshold-free cluster enhancement
--   change point
-
-We used `r pkrt("changepoint")`...
-
--->
+We compared the ability of the GAMM to correctly estimate the onset and offset of the ERP difference to widely used methods.
+First, we conducted mass-univariate t-tests (thus treating each timestep independently) and identified the onset and offset of the ERP difference as the first and last values crossing an arbitrary significance threshold ($\alpha = 0.01$).
+We then followed the same approach but after applying different forms of multiplicity correction the the p-values.
+We compared two methods that control the false discovery rate (FDR) [i.e., BH95, @benjamini1995; and BY01, @benjamini2001], one method that controls the familywise error rate (FWER) (i.e., Holm–Bonferroni method, https://www.jstor.org/stable/4615733), and two cluster-based methods [permutation with a single threshold and TFCE, @smith2009].
+The BH95, BY01, and Holm corrections were applied to the p-values using the `p.adjust()` function in `R`.
+The cluster-based inference was implemented using a cluster-sum statistic of squared t-values, as implemented in `MNE-Python` [@gramfort2013], called via `r pkrt("reticulate")`.
+We also compared these estimates to the onset and offset as estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared t-values [as in @rousselet_using_2025].
+For visualisation and interpretability purposes, we converted p-values to s-values, which can be interpreted as bits of surprising information, assuming a null effect [@greenland2019] (@fig-corrections).
 
@@
 
-```{r test-1d, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Timecourse of squared t-values and s-values (continuous measure of evidence given by -log2(p-value)) with true onset and onset identified using the raw (uncorrected) p-values or the corrected p-values (BH, BY, Holm)."}
-# https://github.com/GRousselet/onsetsim/blob/main/docs/weakstrongfwer_demo.md
-# BH95: p.adjust(pvals, method = "BH")
-# BY01: p.adjust(perm.pvals, method = "BY")
-
+```{r fig-corrections, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Timecourse of squared t-values and s-values, with true onset (black dashed line) and onsets identified using the raw (uncorrected) p-values or the corrected p-values (BH, BY, Holm)."}
 # defining an arbitrary alpha threshold
-aath <- 0.05
+aath <- 0.01
 
@@
         pval_by = p.adjust(p = pval, method = "BY"),
-        # pval_bonf = p.adjust(p = pval, method = "bonferroni"),
         pval_holm = p.adjust(p = pval, method = "holm")
@@
     )
-# onset_bonf <- find_onset(
-#     tests_results$pval_bonf[from_timestep:nrow(tests_results)] <= aath,
-#     tests_results$time[from_timestep:nrow(tests_results)]
-#     )
 onset_holm <- find_onset(
@@
 # plotting the results
-default_colours <- hue_pal()(10)
+default_colours <- met.brewer(name = "Johnson", n = 4)
 tvals_plot <- tests_results %>%
@@
     geom_vline(xintercept = onset_bh, linetype = 2, color = default_colours[2]) +
-    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[4]) +
-    # geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
-    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[5]) +
+    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[3]) +
+    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[4]) +
     labs(x = "Time (s)", y = bquote(t^2) )
@@
     "BY onset: ", round(onset_by, 3), "s, ",
-    # "Bonf onset: ", round(onset_bonf, 3), "s, ",
     "Holm onset: ", round(onset_holm, 3),  "s"
@@
     geom_vline(xintercept = onset_bh, linetype = 2, color = default_colours[2]) +
-    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[4]) +
-    # geom_vline(xintercept = onset_bonf, linetype = 2, color = default_colours[3]) +
-    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[5]) +
+    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[3]) +
+    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[4]) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 4) ) +
     labs(
@@
 
-```{r changepoint, echo = FALSE, out.width = "75%", fig.cap = "T-values timecourse with true onset/offset and onset/offset identified using the changepoint package (binary segmentation method, in green)."}
+```{r changepoint, echo = FALSE, eval = FALSE, out.width = "75%", fig.cap = "T-values timecourse with true onset/offset and onset/offset identified using the changepoint package (binary segmentation method, in green)."}
 # using the changepoint package to identify onsets and offsets
@@
 
-Now cluster-based permutations using `MNE-Python` [@gramfort2013] via `r pkrt("reticulate")`...
-
 ```{r mne-cluster, echo = FALSE, eval = FALSE, fig.cap = "Cluster-based permutation tests via MNE-Python."}
@@
 
-```{r fig-mne-cluster, out.width = "75%", fig.cap = "Cluster-based permutation tests via MNE-Python."}
+```{r fig-mne-cluster, echo = FALSE, eval = FALSE, out.width = "75%", fig.cap = "Cluster-based permutation tests via MNE-Python."}
 # importing the decoding scores
@@
 
+\newpage
+
 ## Simulation study
 
-Onset/offset estimation methods were assessed using these summary statistics of the Monte Carlo sampling distributions of onsets: bias (median-based), mean absolute error (MAE) and variance, with 10.000 iterations (i.e., 10.000 simulated datasets)...
+Onset/offset estimation methods were assessed using these summary statistics of the Monte Carlo sampling distributions of onsets: median absolute error (MAE) and variance, with 10.000 iterations (i.e., 10.000 simulated datasets)...
+
+## Application to actual MEG data
+
+Assessing the reliability of the proposed approach using some sort of split-half reliability [e.g., @rosenblatt2018]?
+Using the MEG decoding results of Nalborczyk et al.
+(in preparation) in 33 participants:
+
+-   Create many (e.g., 10 or 20) half train/test splits of the data
+-   For each fold, estimate the onset/offset on both splits using all methods
+-   Then summarise the distribution of between-split difference across folds
+
+This will allow checking that the proposed approach produces more reliable onset/offset estimates.
+
+# Results
+
+...
+
+## Simulation results (bias and variance)
+
+@fig-simulation-mae-variance shows a summary of the simulation results, revealing that the proposed approach (`brms`) has the lowest MAE and variance for both the onset and offset estimates...
+
+```{r fig-simulation-mae-variance, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Median absolute error and variance of onset and offset estimates for each method."}
+# loading the simulation results
+sim_results <- readRDS(file = "results/sim_results.rds")
+
+# identifying the best brms results
+# eps=0 and threshold = 10 or 20 seems pretty good...
+# sim_results %>%
+#     select(simulation_id:offset_brms) %>%
+#     summarise(
+#         MAE_onset = median(abs(onset_brms - true_onset) ),
+#         variance_onset = var(onset_brms),
+#         MAE_offset = median(abs(offset_brms - true_offset) ),
+#         variance_offset = var(offset_brms),
+#         .by = c(eps, threshold)
+#         ) %>%
+#     arrange(MAE_onset, MAE_offset) %>%
+#     head(10)
+
+# computing and plotting bias, MAE, and variance
+# bias := true_onset - median(sim_onset_distrib)
+# MAE :=  mean(abs(sim_onset-true_onset))
+# variance := variance(sim_onset_distrib)
+sim_results %>%
+    # keeping only the best brms results
+    filter(eps == 0 & threshold == 20) %>%
+    select(-eps,-threshold) %>%
+    pivot_longer(cols = -simulation_id, names_to = "measure", values_to = "value") %>%
+    data.frame() %>%
+    mutate(
+        error = case_when(
+            grepl("onset", measure) ~ abs(value - true_onset),
+            grepl("offset", measure) ~ abs(value - true_offset)
+            )
+        ) %>%
+    separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
+    group_by(type, method) %>%
+    summarise(
+        MAE = median(error, na.rm = TRUE),
+        variance = var(error, na.rm = TRUE),
+        .groups = "drop"
+        ) %>%
+    ungroup() %>%
+    mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c(
+                "p", "bh", "by", "holm", "cluster_mass",
+                "cluster_tfce", "cpt", "brms"
+                ),
+            labels = c(
+                "p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
+                "TFCE", "Change point", "brms"
+                ),
+            )
+        ) %>%
+    ggplot(
+        aes(
+            x = MAE,
+            y = variance,
+            colour = method,
+            fill = method
+            )
+        ) +
+    geom_point(
+        size = 2,
+        pch = 21,
+        colour = "white",
+        show.legend = FALSE
+        ) +
+    geom_label_repel(
+        aes(label = method),
+        colour = "white",
+        size = 3,
+        segment.color = NA,
+        show.legend = FALSE
+        ) +
+    scale_y_log10() +
+    facet_wrap(~type) +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    labs(x = "Median absolute error (s)", y = "Variance (log-scale)")
+```
+
+```{r fig-simulation-distribution, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Distributions of onset and offset estimates for each method."}
+sim_results %>%
+    # keeping only the best brms results
+    filter(eps == 0 & threshold == 20) %>%
+    select(-eps,-threshold) %>%
+    pivot_longer(
+        cols = -simulation_id, 
+        names_to = "measure",
+        values_to = "value"
+        ) %>%
+    separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
+    data.frame() %>%
+    mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c(
+                "p", "bh", "by", "holm", "cluster_mass",
+                "cluster_tfce", "cpt", "brms"
+                ),
+            labels = c(
+                "p-value", "FDR BH95", "FDR BY01", "Holm",
+                "Cluster-mass", "TFCE", "Change point", "brms"
+                ),
+            )
+        ) %>%
+    mutate(
+        plot_row = ifelse(
+            method %in% c("p-value", "FDR BH95", "FDR BY01", "Holm"), 0, 1
+            )
+        ) %>%
+    ggplot(
+        aes(
+            x = value,
+            colour = method,
+            fill = method
+            )
+        ) +
+    geom_vline(xintercept = true_onset) +
+    geom_vline(xintercept = true_offset) +
+    geom_density(aes(fill = NULL) ) +
+    # facet_wrap(~type, scales = "free_x") +
+    # facet_wrap(~type) +
+    facet_grid(plot_row~type, scales = "free_x") +
+    theme_light(base_size = 12, base_family = "Open Sans") +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    labs(x = "Onset/Offset (s)", y = "Density")
+```
+
+## Application to actual EEG data (reliability)
+
+...
 
@@
 
-# Results
+# Discussion
 
@@
 
+## Summary of the proposed approach
+
+...
+
+## Increasing potential usage
+
+Prepare a wrapper `R` package and show how to call it in `Python` and integrate it with `MNE-Python` [@gramfort2013] pipelines...
+
+## Limitations and future directions
+
+Can be applied to any 1D timeseries (e.g., pupillometry, electromyography)... Extending the approach to spatiotemporal data (i.e., time + sensors)...
+
+The error properties depend on the threshold parameter, a value of 10 or 20 seems to be a reasonable default, but the optimal threshold parameter can be adjusted using split-half reliability assessment...
+
+## Conclusions
+
+...
+
 \newpage
 
-# Discussion
-
-...
-
-## Summary of the proposed approach
-
-...
-
-## Increasing potential usage
-
-Prepare a wrapper `R` package and show how to call it in `Python` and integrate it with `MNE-Python` [@gramfort2013] pipelines...
-
-## Limitations and future directions
-
-Can be applied to any 1D timeseries (e.g., pupillometry, electromyography)... Extending the approach to spatiotemporal data (i.e., time + sensors)...
-
-## Conclusions
-
-...
-
-\newpage
+# Data and code availability
+
+The simulation results as well as the `R` code to reproduce the simulations are available on GitHub: <https://github.com/lnalborczyk/brms_meeg>.
 
@@
 Note that $b_{m}(,)$ is a basis function that maps $R \times R \rightarrow R$.
-A popular bivariate basis function uses *thin-plate splines*, which extend to $\mathbf{s}_{i} \in \mathbb{R}^{d}$ and $\partial l_{g}$ penalties.
+A popular bivariate basis function uses *thin-plate splines* [@wood2003], which extend to $\mathbf{s}_{i} \in \mathbb{R}^{d}$ and $\partial l_{g}$ penalties.
 These splines are designed to interpolate and approximate smooth surfaces over two dimensions (hence the "bivariate" term).

### 2025-02-13 brms_meeg.tex
pushing some preliminary simulation results

@@
 
-\title{Modelling M/EEG data with Bayesian multilevel generalised
-additive models}
+\title{Modelling M/EEG data with Bayesian generalised additive
+multilevel models}
 
@@
 and their differences across conditions or group. This approach relies
-on Bayesian nonparametric multilevel modelling (multilevel generalised
-additive models or Gaussian processes), which outputs the posterior
-probabilility of the effect being different than 0 (or above chance) at
+on Bayesian generalised additive multilevel models, which output the
+posterior probabilility of the effect being above 0 (or above chance) at
 every timestep, while naturally taking into account the temporal
@@
 
-\keywords{EEG, MEG, generalised additive models, gaussian
-processes, mixed-effects models, Bayesian statistics, brms}
+\keywords{EEG, MEG, generalised additive models, mixed-effects
+models, multilevel models, Bayesian statistics, brms}
 
@@
 \newpage
-\section[Introduction]{Modelling M/EEG data with Bayesian multilevel
-generalised additive models}
+\section[Introduction]{Modelling M/EEG data with Bayesian generalised
+additive multilevel models}
 
@@
 
-\section{Introduction (in progress)}\label{introduction-in-progress}
+\section{Introduction}\label{introduction}
 
@@
 \citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al., 2019};
-\citeproc{ref-pernet2015}{C. R. Pernet et al., 2015};
-\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al., 2023};
-\citeproc{ref-rousselet_using_2025}{Rousselet, 2025})\ldots{} See also
+\citeproc{ref-pernet2015}{C. R. Pernet et al., 2015})\ldots{} See also
 (\citeproc{ref-maris2011}{Maris, 2011})\ldots{} and
@@
 
-In the following, we consider two approaches to modelling the non-linear
-timecourse of M/EEG ERPs or decoding performance: i) generalised
-additive models (GAMs) with thin-plate smoothing splines
-(\citeproc{ref-wood2003}{Wood, 2003}; \citeproc{ref-wood2004}{Wood,
-2004}) and ii) Gaussian processes (GPs) with a smooth covariance kernel
-(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}) and low-rank
-approximation (\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol
-et al., 2023}).
-
 \subsection{Previous modelling work}\label{previous-modelling-work}
-
-From Dimigen \& Ehinger (\citeproc{ref-dimigen2021}{2021}): ``Recently,
-spline regression has been applied to ERPs (Hendrix, Baayen, \& Bolger,
-2017; Kryuchkova,Tucker,Wurm,\&Baayen,2012;Tremblay\& Baayen, 2010;
-Tremblay \& Newman, 2015)''\ldots{}
-
-Disentangling overlapping processes
-(\citeproc{ref-skukies_brain_2024}{Skukies et al., 2024};
-\citeproc{ref-skukies_modelling_2021}{Skukies \& Ehinger, 2021})\ldots{}
-Using Bayes factors (\citeproc{ref-teichmann2022}{Teichmann,
-2022})\ldots{} Weighting single trials (\citeproc{ref-pernet2022}{C.
-Pernet, 2022})\ldots{}
-
-See
-\url{https://www.let.rug.nl/nerbonne/teach/rema-stats-meth-seminar/presentations/Wieling-2014-GAMs.pdf}
-and \url{https://www.let.rug.nl/wieling/Statistics/GAM-EEG/lab/},
-\url{https://www.let.rug.nl/wieling/Statistics/GAM-EEG/GAM-EEG.pdf}
-(\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
-\citeproc{ref-meulman2015}{Meulman et al., 2015})\ldots{}
 
@@
 
+From Dimigen \& Ehinger (\citeproc{ref-dimigen2021}{2021}): ``Recently,
+spline regression has been applied to ERPs (Hendrix, Baayen, \& Bolger,
+2017; Kryuchkova,Tucker, Wurm, \& Baayen,2012; Tremblay \& Baayen, 2010;
+Tremblay \& Newman, 2015)''\ldots{} GAMMs for EEG data
+(\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
+\citeproc{ref-meulman2015}{Meulman et al., 2015})\ldots{}
+
+Disentangling overlapping processes
+(\citeproc{ref-skukies_brain_2024}{Skukies et al., 2024};
+\citeproc{ref-skukies_modelling_2021}{Skukies \& Ehinger, 2021})\ldots{}
+Weighting single trials (\citeproc{ref-pernet2022}{C. Pernet,
+2022})\ldots{} The LIMO toolbox
+(https://onlinelibrary.wiley.com/doi/10.1155/2011/831409)\ldots{} Using
+Bayes factors (\citeproc{ref-teichmann2022}{Teichmann, 2022})\ldots{}
+
 \subsection{Generalised additive
 models}\label{generalised-additive-models}
+
+See for instance these tutorials
+(\citeproc{ref-suxf3skuthy2017}{Sóskuthy, 2017};
+\citeproc{ref-winter2016}{Winter \& Wieling, 2016}) or application to
+phonetic data (\citeproc{ref-wieling2018}{Wieling, 2018}) or this
+introduction (\citeproc{ref-baayen2020}{Baayen \& Linke, 2020}) or these
+reference books (\citeproc{ref-hastie2017}{Hastie \& Tibshirani, 2017};
+\citeproc{ref-wood2017}{Wood, 2017})\ldots{}
 
@@
 
-\ldots{}
+Focusing on identifying onset and offset of effects (as assessed by ERP
+differences or decoding performance)\ldots{} Assessing the performance
+of a model-based approach (i.e., Bayesian GAMMs) to conventional methods
+(multiplicity corrections or cluster-based permutation)\ldots{}
 
@@
 
-We then fitted a GAM and GP regression model using the \texttt{brms}
-package (\citeproc{ref-brms2017}{Bürkner, 2017},
-\citeproc{ref-brms2018}{2018}; \citeproc{ref-nalborczyk2019}{Nalborczyk
-et al., 2019}). We used the default priors in \texttt{brms}, that is,
-weakly informative priors. Four Markov Chain Monte-Carlo (MCMC) were ran
-for each model to approximate the posterior distribution, including each
-5000 iterations and a warmup of 2000 iterations. Posterior convergence
-was assessed examining trace plots as well as the Gelman--Rubin statisti
+We then fitted a Bayesian GAM using the \texttt{brms} package
+(\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018};
+\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}). We used the
+default priors in \texttt{brms}, that is, weakly informative priors. We
+ran eight Markov Chain Monte-Carlo (MCMC) to approximate the posterior
+distribution, including each 5000 iterations and a warmup of 2000
+iterations, yielding a total of \(8 \times (5000-2000) = 24000\)
+posterior samples to use for inference. Posterior convergence was
+assessed examining trace plots as well as the Gelman--Rubin statistic
 \(\hat{R}\).
@@
     \FunctionTok{summarise}\NormalTok{(}\AttributeTok{eeg =} \FunctionTok{mean}\NormalTok{(eeg) ) }\SpecialCharTok{\%\textgreater{}\%}
-    \CommentTok{\# mutate(eeg = as.numeric(scale(x = eeg, scale = FALSE) ) ) \%\textgreater{}\%}
     \FunctionTok{ungroup}\NormalTok{()}
@@
     \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
+    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
     \AttributeTok{file =} \StringTok{"models/gam.rds"}
@@
 
-Now we fit the GP regression model\ldots{}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# computing eeg\_diff and averaging across participants}
-\CommentTok{\# group\_df \textless{}{-} raw\_df \%\textgreater{}\%}
-\CommentTok{\#     group\_by(participant, time) \%\textgreater{}\%}
-\CommentTok{\#     pivot\_wider(names\_from = condition, values\_from = eeg, values\_fn = mean) \%\textgreater{}\%}
-\CommentTok{\#     mutate(eeg\_diff = cond2 {-} cond1) \%\textgreater{}\%}
-\CommentTok{\#     summarise(eeg\_diff = mean(eeg\_diff) ) \%\textgreater{}\%}
-\CommentTok{\#     ungroup()}
-
-\CommentTok{\# see https://betanalpha.github.io/assets/case\_studies/gp\_part3/part3.html}
-\CommentTok{\# and https://betanalpha.github.io/assets/case\_studies/gaussian\_processes.html}
-\NormalTok{gp\_priors }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
-    \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"Intercept"}\NormalTok{),}
-    \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"normal(0, 1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"b"}\NormalTok{),}
-    \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"exponential(1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sigma"}\NormalTok{),}
-    \FunctionTok{set\_prior}\NormalTok{(}\StringTok{"exponential(1)"}\NormalTok{, }\AttributeTok{class =} \StringTok{"sdgp"}\NormalTok{)}
-\NormalTok{    )}
-
-\NormalTok{gp\_model }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-    \CommentTok{\# k refers to the number of basis functions for}
-    \CommentTok{\# computing Hilbert{-}space approximate GPs}
-    \CommentTok{\# if k = NA (default), exact GPs are computed}
-    \CommentTok{\# eeg \textasciitilde{} gp(time, by = condition),}
-\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{gp}\NormalTok{(time, }\AttributeTok{by =}\NormalTok{ condition, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{cov =} \StringTok{"exp\_quad"}\NormalTok{),}
-    \AttributeTok{data =}\NormalTok{ ppt\_df,}
-    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
-    \CommentTok{\# prior = gp\_priors,}
-    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{15}\NormalTok{),}
-    \AttributeTok{backend =} \StringTok{"cmdstanr"}\NormalTok{,}
-    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/gp.rds"}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
-And we plot the posterior predictions\ldots{}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# plotting the posterior predictions}
-\FunctionTok{plot\_post\_preds}\NormalTok{(}\AttributeTok{model =}\NormalTok{ gam) }\SpecialCharTok{+} \FunctionTok{plot\_post\_preds}\NormalTok{(}\AttributeTok{model =}\NormalTok{ gp\_model)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-\caption{Posterior predictions of the GAM (left) and GP (right) models.}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-gp-preds-1.pdf}
-
-}
-
-\end{figure}%
-
 \subsection{Posterior probability of difference above
@@
 
-We can retrieve the posterior probability of the slope for
-\texttt{condition} at each timestep (Figure~\ref{fig-plot-post-slope}).
+We then plot the posterior predictions together with the posterior
+estimate of the slope for \texttt{condition} at each timestep
+(Figure~\ref{fig-plot-post-slope}).
 
@@
 
-\caption{\label{fig-plot-post-slope}Slope for the difference between
-conditions according to the GAM (left) or the GP (right).}
+\caption{\label{fig-plot-post-slope}Posterior estimate of the ERP in
+each condition (left) or directly for the difference of ERPs (right)
+according to the GAM.}
 
@@
 
-We can also compute the posterior probability of the slope for
+We then compute the posterior probability of the slope for
 \texttt{condition} being above \(0+\epsilon\)
@@
 \caption{\label{fig-post-prob-test}Posterior probability of the ERP
-difference (slope) being above 0 according to the GAM (left) and the GP
-(right).}
+difference (slope) being above 0 according to the GAM.}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-test-1.pdf}
+\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-test-1.pdf}
 
@@
     \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
+    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
     \AttributeTok{file =} \StringTok{"models/meta\_gam.rds"}
@@
 \end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# fitting the GP}
-\NormalTok{meta\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-    \CommentTok{\# using by{-}participant SD of ERPs across trials}
-\NormalTok{    eeg\_mean }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
-\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{gp}\NormalTok{(time, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
-\NormalTok{        (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ participant),}
-    \AttributeTok{data =}\NormalTok{ summary\_df,}
-    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
-    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{),}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/meta\_gp.rds"}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# plotting the posterior predictions}
-\NormalTok{meta\_gam\_preds }\OtherTok{\textless{}{-}} \FunctionTok{plot}\NormalTok{(}
-    \FunctionTok{conditional\_effects}\NormalTok{(}\AttributeTok{x =}\NormalTok{ meta\_gam, }\AttributeTok{effect =} \StringTok{"time:condition"}\NormalTok{),}
-    \AttributeTok{points =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{theme =} \FunctionTok{theme\_light}\NormalTok{(), }\AttributeTok{plot =} \ConstantTok{FALSE}
-\NormalTok{    )[[}\DecValTok{1}\NormalTok{]] }\SpecialCharTok{+}
-    \FunctionTok{scale\_colour\_met\_d}\NormalTok{(}\AttributeTok{name =} \StringTok{"Johnson"}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{scale\_fill\_met\_d}\NormalTok{(}\AttributeTok{name =} \StringTok{"Johnson"}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (s)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG signal (a.u.)"}\NormalTok{)}
-\NormalTok{meta\_gp\_preds }\OtherTok{\textless{}{-}} \FunctionTok{plot}\NormalTok{(}
-    \FunctionTok{conditional\_effects}\NormalTok{(}\AttributeTok{x =}\NormalTok{ meta\_gp, }\AttributeTok{effect =} \StringTok{"time:condition"}\NormalTok{),}
-    \AttributeTok{points =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{theme =} \FunctionTok{theme\_light}\NormalTok{(), }\AttributeTok{plot =} \ConstantTok{FALSE}
-\NormalTok{    )[[}\DecValTok{1}\NormalTok{]] }\SpecialCharTok{+}
-    \FunctionTok{scale\_colour\_met\_d}\NormalTok{(}\AttributeTok{name =} \StringTok{"Johnson"}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{scale\_fill\_met\_d}\NormalTok{(}\AttributeTok{name =} \StringTok{"Johnson"}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Time (s)"}\NormalTok{, }\AttributeTok{y =} \StringTok{"EEG signal (a.u.)"}\NormalTok{)}
-
-\CommentTok{\# combining the two plots}
-\NormalTok{meta\_gam\_preds }\SpecialCharTok{+}\NormalTok{ meta\_gp\_preds}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-\caption{Posterior predictions for the hierarchical GAM and GP.}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/meta-gp-preds-1.pdf}
-
-}
-
-\end{figure}%
-
-\newpage
 
@@
 
-\subsection{Application to actual M/EEG
-data}\label{application-to-actual-meeg-data}
-
-Assessing the reliability of the proposed approach (in comparison to
-other methods) using some sort of split-half reliability
-(\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018})?
-
 \subsection{Comparing the identified onsets/offsets to other
@@
 
-We considered five methods for multiplicity corrections: two FDR
-methods; two cluster-based methods; and the maximum statistics. The two
-FDR methods were BH95 (Benjamini \& Hochberg, 1995;
-10.1111/j.2517-6161.1995.tb02031.x) and BY01 (Benjamini \& Yekutieli,
-2001; 10.1214/aos/1013699998), which were applied to the permutation
-p-values using the \texttt{p.adjust} function in \texttt{R}. The first
-cluster-based inference was implemented using a cluster-sum statistic of
-squared t-values\ldots{}
-
-We also compared the performance of the GAMM (in correctly
-identifying/estimating the onset and offset of ERP difference across
-conditions) to the performance of the binary segmentation method as
-implemented in the \texttt{R} package \texttt{changepoint} v 2.2.4
-(\citeproc{ref-changepoint}{Killick et al., 2022a}), threshold-free
-cluster enhancement (TFCE, \citeproc{ref-smith2009}{Smith \& Nichols,
-2009}), as implemented in \texttt{MNE-Python}
-(\citeproc{ref-gramfort2013}{Gramfort, 2013}), and using the raw t-, p-,
-or s-values timecourse. S-values
-(\citeproc{ref-greenland2019}{Greenland, 2019})\ldots{}
-
-\begin{figure}[H]
-
-\caption{Timecourse of squared t-values and s-values (continuous measure
-of evidence given by -log2(p-value)) with true onset and onset
-identified using the raw (uncorrected) p-values or the corrected
-p-values (BH, BY, Holm).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/test-1d-1.pdf}
+We compared the ability of the GAMM to correctly estimate the onset and
+offset of the ERP difference to widely used methods. First, we conducted
+mass-univariate t-tests (thus treating each timestep independently) and
+identified the onset and offset of the ERP difference as the first and
+last values crossing an arbitrary significance threshold
+(\(\alpha = 0.01\)). We then followed the same approach but after
+applying different forms of multiplicity correction the the p-values. We
+compared two methods that control the false discovery rate (FDR) (i.e.,
+BH95, \citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995}; and
+BY01, \citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001}), one
+method that controls the familywise error rate (FWER) (i.e.,
+Holm--Bonferroni method, https://www.jstor.org/stable/4615733), and two
+cluster-based methods (permutation with a single threshold and TFCE,
+\citeproc{ref-smith2009}{Smith \& Nichols, 2009}). The BH95, BY01, and
+Holm corrections were applied to the p-values using the
+\texttt{p.adjust()} function in \texttt{R}. The cluster-based inference
+was implemented using a cluster-sum statistic of squared t-values, as
+implemented in \texttt{MNE-Python}
+(\citeproc{ref-gramfort2013}{Gramfort, 2013}), called via the \texttt{R}
+package \texttt{reticulate} v 1.35.0 (\citeproc{ref-reticulate}{Ushey et
+al., 2024}). We also compared these estimates to the onset and offset as
+estimated using the binary segmentation algorithm, as implemented in the
+\texttt{R} package \texttt{changepoint} v 2.2.4
+(\citeproc{ref-changepoint}{Killick et al., 2022a}), and applied
+directly to the squared t-values (as in
+\citeproc{ref-rousselet_using_2025}{Rousselet, 2025}). For visualisation
+and interpretability purposes, we converted p-values to s-values, which
+can be interpreted as bits of surprising information, assuming a null
+effect (\citeproc{ref-greenland2019}{Greenland, 2019})
+(Figure~\ref{fig-corrections}).
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-corrections}Timecourse of squared t-values and
+s-values, with true onset (black dashed line) and onsets identified
+using the raw (uncorrected) p-values or the corrected p-values (BH, BY,
+Holm).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-corrections-1.pdf}
 
@@
 
-\begin{figure}[H]
-
-\caption{T-values timecourse with true onset/offset and onset/offset
-identified using the changepoint package (binary segmentation method, in
-green).}
-
-{\centering \includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/changepoint-1.pdf}
+\newpage
+
+\subsection{Simulation study}\label{simulation-study}
+
+Onset/offset estimation methods were assessed using these summary
+statistics of the Monte Carlo sampling distributions of onsets: median
+absolute error (MAE) and variance, with 10.000 iterations (i.e., 10.000
+simulated datasets)\ldots{}
+
+\subsection{Application to actual MEG
+data}\label{application-to-actual-meg-data}
+
+Assessing the reliability of the proposed approach using some sort of
+split-half reliability (e.g., \citeproc{ref-rosenblatt2018}{Rosenblatt
+et al., 2018})? Using the MEG decoding results of Nalborczyk et al.~(in
+preparation) in 33 participants:
+
+\begin{itemize}
+\tightlist
+\item
+  Create many (e.g., 10 or 20) half train/test splits of the data
+\item
+  For each fold, estimate the onset/offset on both splits using all
+  methods
+\item
+  Then summarise the distribution of between-split difference across
+  folds
+\end{itemize}
+
+This will allow checking that the proposed approach produces more
+reliable onset/offset estimates.
+
+\section{Results}\label{results}
+
+\ldots{}
+
+\subsection{Simulation results (bias and
+variance)}\label{simulation-results-bias-and-variance}
+
+Figure~\ref{fig-simulation-mae-variance} shows a summary of the
+simulation results, revealing that the proposed approach (\texttt{brms})
+has the lowest MAE and variance for both the onset and offset
+estimates\ldots{}
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-simulation-mae-variance}Median absolute error and
+variance of onset and offset estimates for each method.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-simulation-mae-variance-1.pdf}
 
@@
 
-Now cluster-based permutations using \texttt{MNE-Python}
-(\citeproc{ref-gramfort2013}{Gramfort, 2013}) via the \texttt{R} package
-\texttt{reticulate} v 1.35.0 (\citeproc{ref-reticulate}{Ushey et al.,
-2024})\ldots{}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# importing the decoding scores}
-\NormalTok{p\_values\_df }\OtherTok{\textless{}{-}} \FunctionTok{readRDS}\NormalTok{(}\AttributeTok{file =} \StringTok{"results/mne\_permutation\_decoding\_scores.rds"}\NormalTok{)}
-
-\CommentTok{\# plotting the s{-}values ({-}log2(p{-}values))}
-\NormalTok{p\_values\_df }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{sval =} \SpecialCharTok{{-}}\FunctionTok{log2}\NormalTok{(pval) ) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ time, }\AttributeTok{y =}\NormalTok{ sval) ) }\SpecialCharTok{+}
-    \FunctionTok{geom\_line}\NormalTok{(}\AttributeTok{linewidth =} \FloatTok{0.5}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{geom\_hline}\NormalTok{(}\AttributeTok{yintercept =} \SpecialCharTok{{-}}\FunctionTok{log2}\NormalTok{(}\FloatTok{0.05}\NormalTok{), }\AttributeTok{linetype =} \DecValTok{2}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}
-        \AttributeTok{x =} \StringTok{"Time (s)"}\NormalTok{,}
-        \AttributeTok{y =} \StringTok{"{-}log2(p{-}value)"}
-\NormalTok{        )}
-\end{Highlighting}
-\end{Shaded}
-
 \begin{figure}[!htb]
 
-\caption{\label{fig-mne-cluster}Cluster-based permutation tests via
-MNE-Python.}
+\caption{\label{fig-simulation-distribution}Distributions of onset and
+offset estimates for each method.}
 
@@
 
-\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-mne-cluster-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-simulation-distribution-1.pdf}
 
@@
 
-\subsection{Simulation study}\label{simulation-study}
-
-Onset/offset estimation methods were assessed using these summary
-statistics of the Monte Carlo sampling distributions of onsets: bias
-(median-based), mean absolute error (MAE) and variance, with 10.000
-iterations (i.e., 10.000 simulated datasets)\ldots{}
-
-\newpage
-
-\section{Results}\label{results}
+\subsection{Application to actual EEG data
+(reliability)}\label{application-to-actual-eeg-data-reliability}
 
@@
 
+The error properties depend on the threshold parameter, a value of 10 or
+20 seems to be a reasonable default, but the optimal threshold parameter
+can be adjusted using split-half reliability assessment\ldots{}
+
 \subsection{Conclusions}\label{conclusions}
@@
 \newpage
+
+\section{Data and code availability}\label{data-and-code-availability}
+
+The simulation results as well as the \texttt{R} code to reproduce the
+simulations are available on GitHub:
+\url{https://github.com/lnalborczyk/brms_meeg}.
 
@@
 
+\bibitem[\citeproctext]{ref-baayen2020}
+Baayen, R. H., \& Linke, M. (2020). \emph{Generalized Additive Mixed
+Models} (pp. 563--591). Springer International Publishing.
+\url{https://doi.org/10.1007/978-3-030-46216-1_23}
+
+\bibitem[\citeproctext]{ref-benjamini1995}
+Benjamini, Y., \& Hochberg, Y. (1995). Controlling the False Discovery
+Rate: A Practical and Powerful Approach to Multiple Testing.
+\emph{Journal of the Royal Statistical Society Series B: Statistical
+Methodology}, \emph{57}(1), 289--300.
+\url{https://doi.org/10.1111/j.2517-6161.1995.tb02031.x}
+
+\bibitem[\citeproctext]{ref-benjamini2001}
+Benjamini, Y., \& Yekutieli, D. (2001). The control of the false
+discovery rate in multiple testing under dependency. \emph{The Annals of
+Statistics}, \emph{29}(4). \url{https://doi.org/10.1214/aos/1013699998}
+
 \bibitem[\citeproctext]{ref-brms2017}
@@
 \url{https://doi.org/10.1080/00031305.2018.1529625}
+
+\bibitem[\citeproctext]{ref-hastie2017}
+Hastie, T. J., \& Tibshirani, R. J. (2017). \emph{Generalized Additive
+Models}. Routledge. \url{https://doi.org/10.1201/9780203753781}
 
@@
 
-\bibitem[\citeproctext]{ref-rasmussen2005}
-Rasmussen, C. E., \& Williams, C. K. I. (2005). \emph{Gaussian Processes
-for Machine Learning}.
-\url{https://doi.org/10.7551/mitpress/3206.001.0001}
-
-\bibitem[\citeproctext]{ref-riutort-mayol_practical_2023}
-Riutort-Mayol, G., Bürkner, P.-C., Andersen, M. R., Solin, A., \&
-Vehtari, A. (2023). Practical {Hilbert} space approximate {Bayesian
-Gaussian} processes for probabilistic programming. \emph{Statistics and
-Computing}, \emph{33}(1), 17.
-\url{https://doi.org/10.1007/s11222-022-10167-2}
-
 \bibitem[\citeproctext]{ref-grateful}
@@
 \url{https://doi.org/10.1016/j.neuroimage.2008.03.061}
+
+\bibitem[\citeproctext]{ref-suxf3skuthy2017}
+Sóskuthy, M. (2017). \emph{Generalised additive mixed models for dynamic
+analysis in linguistics: A practical introduction}.
+\url{https://doi.org/10.48550/ARXIV.1703.05339}
 
@@
 
+\bibitem[\citeproctext]{ref-wieling2018}
+Wieling, M. (2018). Analyzing dynamic phonetic data using generalized
+additive mixed modeling: A tutorial focusing on articulatory differences
+between L1 and L2 speakers of English. \emph{Journal of Phonetics},
+\emph{70}, 86--116. \url{https://doi.org/10.1016/j.wocn.2018.03.002}
+
+\bibitem[\citeproctext]{ref-winter2016}
+Winter, B., \& Wieling, M. (2016). How to analyze linguistic change
+using mixed models, Growth Curve Analysis and Generalized Additive
+Modeling. \emph{Journal of Language Evolution}, \emph{1}(1), 7--18.
+\url{https://doi.org/10.1093/jole/lzv003}
+
 \bibitem[\citeproctext]{ref-wood2003}
@@
 
-\bibitem[\citeproctext]{ref-wood2004}
-Wood, S. N. (2004). Stable and Efficient Multiple Smoothing Parameter
-Estimation for Generalized Additive Models. \emph{Journal of the
-American Statistical Association}, \emph{99}(467), 673--686.
-\url{https://doi.org/10.1198/016214504000000980}
+\bibitem[\citeproctext]{ref-wood2017}
+Wood, S. N. (2017). \emph{Generalized Additive Models}. Chapman;
+Hall/CRC. \url{https://doi.org/10.1201/9781315370279}
 
@@
 \(R \times R \rightarrow R\). A popular bivariate basis function uses
-\emph{thin-plate splines}, which extend to
-\(\mathbf{s}_{i} \in \mathbb{R}^{d}\) and \(\partial l_{g}\) penalties.
-These splines are designed to interpolate and approximate smooth
-surfaces over two dimensions (hence the ``bivariate'' term). For \(d=2\)
-dimensions and \(l=2\) (smoothness penalty involving second order
-derivative):
+\emph{thin-plate splines} (\citeproc{ref-wood2003}{Wood, 2003}), which
+extend to \(\mathbf{s}_{i} \in \mathbb{R}^{d}\) and \(\partial l_{g}\)
+penalties. These splines are designed to interpolate and approximate
+smooth surfaces over two dimensions (hence the ``bivariate'' term). For
+\(d=2\) dimensions and \(l=2\) (smoothness penalty involving second
+order derivative):
 

### 2025-02-13 code/analysing_sim_results.R

@@ -1,0 +1,183 @@
+###########################################################
+# Monte-Carlo simulation of onset/offset error properties #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on February 13, 2025                       #
+###########################################################
+
+library(patchwork)
+library(tidyverse)
+library(ggrepel)
+
+# loading simulation results
+sim_results <- readRDS(file = "results/sim_results.rds")
+
+# identifying the best brms results
+# eps=0 and threshold = 10 or 20 seems pretty good...
+sim_results %>%
+    select(simulation_id:offset_brms) %>%
+    summarise(
+        MAE_onset = median(abs(onset_brms - true_onset) ),
+        variance_onset = var(onset_brms),
+        MAE_offset = median(abs(offset_brms - true_offset) ),
+        variance_offset = var(offset_brms),
+        .by = c(eps, threshold)
+        ) %>%
+    arrange(MAE_onset, MAE_offset) %>%
+    head(10)
+
+# plotting some results
+sim_results %>%
+    summarise(error_brms = mean(abs(onset_brms-true_onset) ), .by = c(eps, threshold) ) %>%
+    ggplot(aes(x = eps, y = threshold, fill = error_brms) ) +
+    geom_tile(show.legend = FALSE) +
+    geom_point(
+        data = sim_results %>%
+            summarise(error_brms = mean(abs(onset_brms-true_onset) ), .by = c(eps, threshold) ) %>%
+            dplyr::filter(error_brms == min(error_brms, na.rm = TRUE) ),
+        aes(x = eps, y = threshold),
+        color = "orangered", size = 2, shape = 4,
+        show.legend = FALSE
+        ) +
+    scale_x_continuous(expand = c(0, 0) ) +
+    scale_y_continuous(expand = c(0, 0) ) +
+    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
+    labs(
+        title = "Onset error",
+        x = "eps",
+        y = "Threshold",
+        fill = "Error"
+        )
+
+sim_results %>%
+    summarise(error_brms = mean(abs(offset_brms-true_offset) ), .by = c(eps, threshold) ) %>%
+    # arrange(error_brms) %>% head()
+    ggplot(aes(x = eps, y = threshold, fill = error_brms) ) +
+    geom_tile(show.legend = FALSE) +
[…]

### 2025-02-13 code/errors_simulation.R

@@ -2,30 +2,34 @@
 # Monte-Carlo simulation of onset/offset error properties #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on February 10, 2025                       #
+# Last updated on February 13, 2025                       #
 ###########################################################
 
+# importing R packages
 library(changepoint)
 library(reticulate)
 library(tidyverse)
+library(tidybayes)
 library(brms)
 
-# retrieving the code from the Quarto document
-# knitr::purl(input = "brms_meeg.qmd", output = "brms_meeg.R", documentation = 0)
+# importing Python modules
+np <- import("numpy")
+mne <- import("mne")
 
 # importing R version of Matlab code from Yeung et al. (2004)
-source("eeg_noise.R")
+source("code/eeg_noise.R")
 
 # importing the ERP template with true onset = 160 ms, F=81, and max at F=126
-source("erp_template.R")
+source("code/erp_template.R")
 
 # importing helper functions
-source("functions.R")
+source("code/functions.R")
 
 # to use with the eeg_noise function
-meanpower <- unlist(read.table("meanpower.txt") )
+meanpower <- unlist(read.table("code/meanpower.txt") )
 
 # defining the number of simulations to be performed
+# nsims <- 1e4
 nsims <- 10
 
 # defining simulation parameters
@@ -60,220 +64,110 @@ generate_data <- function (n_trials, n_ppt, outvar, srate, ronset) {
         cond1 <- matrix(0, nrow = n_trials, ncol = Nf)
         cond2 <- matrix(0, nrow = n_trials, ncol = Nf)
         
-        for (T in 1:n_trials) { # for each trial
+        for (trial in 1:n_trials) { # for each trial
             
-            cond1[T, ] <- temp1 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
-            cond2[T, ] <- temp2 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
+            cond1[trial, ] <- temp1 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
+            cond2[trial, ] <- temp2 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
             
         }
         
         # converting results to dataframe
         temp_df <- data.frame(
[…]

### 2025-02-14 brms_meeg.qmd
pushin new simulation results and updated title and abstract

@@
 ---
-title: "Modelling M/EEG data with Bayesian generalised additive multilevel models"
-shorttitle: "Bayesian M/EEG modelling"
+title: "Precise temporal localisation of M/EEG effects with Bayesian generalised additive multilevel models"
+shorttitle: "Bayesian modelling of M/EEG data"
 author:
@@
     # financial-support: "This study was supported by Grant [Grant Number] from [Funding Source]."
-abstract: "Time-resolved electrophysiological measurements such as those offered by magneto- or electro-encephalography (M/EEG) provide a unique window onto neural activity underlying cognitive process and how they unfold over time.
-Typically, we are interested in testing whether such measures differ across conditions and/or groups.
-The conventional approach consists in conducting mass-univariate statistics through followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster-based inference.
-However, these cluster-based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the onset and offset of effects (e.g., differences across conditions).
-Here, we introduce a novel *model-based approch* for analysing one-dimensional M/EEG timeseries such as ERPs or decoding timecourses and their differences across conditions or group.
-This approach relies on Bayesian generalised additive multilevel models, which output the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between-subject variability present in such data."
+abstract: "Time-resolved electrophysiological measurements such as those offered by magneto- or electro-encephalography (M/EEG) provide a unique window onto neural activity underlying cognitive processes.
+Typically, researchers are interested in testing whether and when such measures differ across conditions and/or groups.
+The conventional approach consists in conducting mass-univariate statistics through time followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster-based inference.
+However, these cluster-based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the onset or offset of effects (e.g., differences across conditions or groups).
+Here, we introduce a *model-based* approch for analysing M/EEG timeseries such as ERPs or timecourses of decoding performance and their differences across conditions or groups.
+This approach relies on Bayesian generalised additive multilevel models, which output the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between-subject variability present in such data.
+Using both simulation and actual EEG data, we show that the proposed approach largely outperforms conventional methods in determining both the onset and offset of M/EEG effects (e.g., ERPs difference, decoding performance), producting more precise and more reliable estimates.
+We provide an R package implementing the approach and illustrate how to integrate it into M/EEG statistical pipelines in MNE-Python."
 keywords: [EEG, MEG, generalised additive models, mixed-effects models, multilevel models, Bayesian statistics, brms]
@@
 a4paper: true
+donotrepeattitle: true
 lang: en
@@
 # template for citing R packages
+# pkrt_set(pkg = "the `R` package `:pkg` [:ref]")
 pkrt_set(pkg = "the `R` package `:pkg` v :ver [:ref]")
@@
 Here are some useful references to be discussed [@combrisson_exceeding_2015; @ehinger_unfold_2019; @hayasaka_validating_2003; @luck_how_2017; @pedersen_hierarchical_2019; @frossard2021; @frossard2022; @gramfort2013; @maris2007; @pernet2015]...
-See also [@maris2011]...
-and [@rosenblatt2018] (history of cluster-based approaches and using a data split?)...
-Cluster failure [@eklund2016]...
-
-```{=html}
-<!--
-
-In the following, we consider two approaches to modelling the non-linear timecourse of M/EEG ERPs (or decoding performance): i) generalised additive models (GAMs) with smoothing splines [@wood2003; @wood2004]...
-
-and ii) Gaussian processes (GPs) with a smooth covariance kernel [@rasmussen2005] and low-rank approximation [@riutort-mayol_practical_2023].
-
--->
-```
-
-## Previous modelling work
+See also @maris2011 and @rosenblatt2018 (history of cluster-based approaches)...
+Clusters failures [@eklund2016]...
+
+## Previous work
 
 Recent example of GLM for EEG [@wüllhorst2025; @fischer2013]...
-"Similar approaches have been successfully applied to EEG time- (Rousselet et al., 2008) and frequency-domain (Cohen and Cavanagh, 2011) data and allow the simultaneous investigation of multiple independent variables while preserving the high temporal resolution of the EEG."...
 See also [@rousselet2008; @hauk2006]...
@@
 
-From @dimigen2021: "Recently, spline regression has been applied to ERPs (Hendrix, Baayen, & Bolger, 2017; Kryuchkova,Tucker, Wurm, & Baayen,2012; Tremblay & Baayen, 2010; Tremblay & Newman, 2015)"...
-GAMMs for EEG data [@abugaber2023; @meulman2015]...
-
-Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]...
-Weighting single trials [@pernet2022]...
-The LIMO toolbox (https://onlinelibrary.wiley.com/doi/10.1155/2011/831409)...
-Using Bayes factors [@teichmann2022]...
+From @dimigen2021: Recently, spline regression has been applied to ERPs [e.g., @hendrix2017; @kryuchkova2012]... GAMMs for EEG data [@abugaber2023; @meulman2015]...
+
+Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]... Weighting single trials [@pernet2022]... The LIMO toolbox [@pernet2011]...
+
+Using Bayes factors for decoding performance [@teichmann2022]...
+
+## Bayesian regression modelling
+
+Short intro/recap about Bayesian (linear and generalised) regression models...
 
@@
 See for instance these tutorials [@sóskuthy2017; @winter2016] or application to phonetic data [@wieling2018] or this introduction [@baayen2020] or these reference books [@hastie2017; @wood2017]...
+application to pupillometry [@vanrij2019]...
+GAMLSS for neuroimaging data [@dinga2021]...
 
@@
 y_{i} &\sim \mathrm{EF}\left(\mu_{i}, \phi\right)\\
-g\left(\mu_i\right) &= A_{i} + \mathbf{X}_{i} \gamma + \sum_{j=1}^{J} f_{j}\left(x_{ij}\right)
+g\left(\mu_i\right) &=  \underbrace{A_{i} + \mathbf{X}_{i} \gamma}_{\text{parametric part}} + \underbrace{\sum_{j=1}^{J} f_{j}\left(x_{ij}\right)}_{\text{non-parametric part}}
 \end{aligned}
@@
 where $\beta_{jk}$ is the weight (coefficient) associated with the $k$th basis function $b_{jk}()$ evaluated at the covariate value $x_{ij}$ for the $j$th smooth function $f_{j}$.
-Splines' coefficients are penalised...
+Splines coefficients are penalised (usually through the squared of the smooth functions' second derivative) in a way that can be interpreted as a prior on the "wiggliness" of the function...
+
+## Bayesian generalised additive multilevel models
+
+Now describe the Bayesian GAMM... Proper inclusion of varying/random effects in the model specification protects against overly wiggly curves [@baayen2020]...
 
@@
 
-```{r fig-eeg, echo = FALSE, fig.width = 9, fig.cap = "Some ERPs in two conditions with 50 trials each, for a group of 20 participants."}
+```{r fig-eeg, echo = FALSE, fig.width = 9, fig.cap = "Averaged (mean) simulated EEG activity in two conditions with 50 trials each, for a group of 20 participants.
+The error band represents the mean plus/minus 1 standard error of the mean."}
 # importing R version of Matlab code from Yeung et al. (2004)
@@
 raw_df %>%
-    summarise(eeg = mean(eeg), .by = c(participant, condition, time) ) %>%
-    ggplot(aes(x = time, y = eeg, colour = condition) ) +
+    summarise(
+        eeg_mean = mean(eeg),
+        eeg_se = sd(eeg) / sqrt(n() ),
+        .by = c(participant, condition, time),
+        ) %>%
+    ggplot(aes(x = time, y = eeg_mean, colour = condition, fill = condition) ) +
     geom_hline(yintercept = 0, linetype = 2) +
@@
     geom_vline(xintercept = true_offset, linetype = 2) +
+    geom_ribbon(
+        aes(ymin = eeg_mean - eeg_se, ymax = eeg_mean + eeg_se, colour = NULL),
+        alpha = 0.3, show.legend = FALSE
+        ) +
     geom_line(show.legend = FALSE) +
@@
         aes(ymin = lower, ymax = upper, colour = NULL),
-        alpha = 0.25, show.legend = FALSE
+        alpha = 0.3, show.legend = FALSE
         ) +
@@
 Posterior convergence was assessed examining trace plots as well as the Gelman–Rubin statistic $\hat{R}$.
+The `brms` package uses the same syntax as `r pkrt("mgcv")` for specifying smooth effects.
 
@@
     file = "models/gam.rds"
+    )
+```
+
+```{r gamm, echo = FALSE, eval = FALSE, message = FALSE}
+# defining a contrast for condition
+contrasts(raw_df$condition) <- c(-0.5, 0.5)
+
+# fitting the GAMM with factor smooth (per participant)
+gamm <- brm(
+    # cubic regression splines with k-1 basis functions
+    eeg ~ condition +
+        s(time, bs = "cr", k = 10, by = condition) +
+        s(time, participant, bs = "fs"),
+    data = raw_df,
+    family = gaussian(),
+    warmup = 2000,
+    iter = 5000,
+    chains = 4,
+    cores = 4,
+    file = "models/gamm.rds"
     )
@@
         geom_vline(xintercept = true_onset, linetype = 2) +
+        geom_vline(xintercept = true_offset, linetype = 2) +
         geom_line() +
@@
 
-```{r fig-post-prob-test, echo = FALSE, out.width = "75%", fig.asp = 0.5, fig.cap = "Posterior probability of the ERP difference (slope) being above 0 according to the GAM."}
+```{r fig-post-prob-test, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Posterior probability of the ERP difference (slope) being above 0 according to the GAM."}
 # defining a function to plot posterior prob of slope above 0
@@
 
-```{r fig-post-prob-ratio, echo = FALSE, fig.width = 8, fig.asp = 0.75, fig.cap = "Ratio of posterior probability according to the GAM. Timesteps above threshold (10) are highlighted in green."}
+```{r fig-post-prob-ratio, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Ratio of posterior probability according to the GAM. Timesteps above threshold (10) are highlighted in green."}
 # defining some parameters
@@
 We then followed the same approach but after applying different forms of multiplicity correction the the p-values.
-We compared two methods that control the false discovery rate (FDR) [i.e., BH95, @benjamini1995; and BY01, @benjamini2001], one method that controls the familywise error rate (FWER) (i.e., Holm–Bonferroni method, https://www.jstor.org/stable/4615733), and two cluster-based methods [permutation with a single threshold and TFCE, @smith2009].
+We compared two methods that control the false discovery rate (FDR) [i.e., BH95, @benjamini1995; and BY01, @benjamini2001], one method that controls the familywise error rate (FWER) [i.e., Holm–Bonferroni method, @holm1979], and two cluster-based methods [permutation with a single threshold and TFCE, @smith2009].
 The BH95, BY01, and Holm corrections were applied to the p-values using the `p.adjust()` function in `R`.
@@
 
-Onset/offset estimation methods were assessed using these summary statistics of the Monte Carlo sampling distributions of onsets: median absolute error (MAE) and variance, with 10.000 iterations (i.e., 10.000 simulated datasets)...
+Onset/offset estimation methods were assessed using the median absolute error (MAE) and variance of 10.000 simulated datasets...
 
@@
 -   For each fold, estimate the onset/offset on both splits using all methods
--   Then summarise the distribution of between-split difference across folds
-
-This will allow checking that the proposed approach produces more reliable onset/offset estimates.
+-   Then summarise the distribution of onset/offset with the median and variance across folds
+
+This will allow checking that the proposed approach produces reliable onset/offset estimates.
+
+```{r reliability, echo = FALSE, eval = FALSE}
+# library(caret)
+# see https://github.com/topepo/caret/
+# https://topepo.github.io/caret/data-splitting.html
+# and createDataPartition()
+# and createFolds()
+# trainIndex <- createDataPartition(
+#     y = summary_df$participant,
+#     p = 0.5,
+#     list = FALSE,
+#     times = 10
+#     )
+
+# subsetting the data
+# summary_df_train <- summary_df[trainIndex, ]
+# summary_df_test  <- summary_df[-trainIndex, ]
+
+# getting unique participants
+unique_participants <- unique(summary_df$participant)
+
+# getting the number of participants
+num_participants <- length(unique_participants)
+
+# ensuring participants are randomly shuffled
+shuffled_participants <- sample(unique_participants)
+
+# defining the number of splits
+num_splits <- 10
+
+# creating a dataframe where each participant appears in every split
+split_df <- crossing(participant = unique_participants, split_number = 1:10)
+
+# assigning alternating groups (A or B) within each split for each participant
+# split_df <- split_df %>% mutate(group = rep(c("A", "B"), length.out = n() ) )
+split_df <- split_df %>%
+    group_by(split_number) %>%
+    # shuffling groups for each split
+    mutate(group = sample(rep(c("A", "B"), length.out = n() ) ) ) %>%
+    ungroup()
+
+# sanity checks
+# table(split_df$group)
+# split_df %>%
+#     ggplot(aes(x = group) ) +
+#     geom_histogram(stat = "count") +
+#     facet_wrap(~participant, ncol = 5)
+
+# now for each split, assess the onset and offset
+participants_split1_train <- split_df %>%
+    filter(split_number == 1 & group == "A") %>%
+    pull(participant)
+participants_split1_test <- split_df %>%
+    filter(split_number == 1 & group == "B") %>%
+    pull(participant)
+
+# split1 data
+summary_df_split1_train <- summary_df %>%
+    filter(participant %in% participants_split1_train)
+
+# fitting the GAM
+meta_gam_split1_train <- brm(
+    eeg_mean | se(eeg_sd) ~
+        condition + s(time, bs = "cr", k = 10, by = condition) +
+        (1 | participant),
+    data = summary_df_split1_train,
+    family = gaussian(),
+    warmup = 2000,
+    iter = 5000,
+    chains = 4,
+    cores = 4
+    )
+```
 
@@
     select(-eps,-threshold) %>%
+    # filtering some alpha_level
+    filter(alpha_level == 0.05) %>%
+    select(-alpha_level) %>%
     pivot_longer(cols = -simulation_id, names_to = "measure", values_to = "value") %>%
@@
             labels = c(
-                "p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
                 "TFCE", "Change point", "brms"
@@
     filter(eps == 0 & threshold == 20) %>%
+    # filtering some alpha_level
+    filter(alpha_level == 0.05) %>%
+    select(-alpha_level) %>%
     select(-eps,-threshold) %>%
@@
             labels = c(
-                "p-value", "FDR BH95", "FDR BY01", "Holm",
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm",
                 "Cluster-mass", "TFCE", "Change point", "brms"
@@
     geom_density(aes(fill = NULL) ) +
-    # facet_wrap(~type, scales = "free_x") +
-    # facet_wrap(~type) +
     facet_grid(plot_row~type, scales = "free_x") +
@@
 Can be applied to any 1D timeseries (e.g., pupillometry, electromyography)... Extending the approach to spatiotemporal data (i.e., time + sensors)...
+
+We kept the exemplary models simple, but can be extended by adding varying/random effects (intercept and slope) for item (e.g., word)... but also continuous predictors at the trial level?
 
@@
 
-# Using the `R` package and integration with `MNE-Python`
-
-Explain how to use the `R` package and to integrate it with `MNE` epochs...
+# Integration with `MNE-Python`
+
+Explain how to use the `R` package with `MNE` epochs...
 
 ```{r r-python, eval = FALSE}
+# suggest a package name
+library(available)
+suggest(
+    text = "Time-resolved electrophysiological measurements such as those offered by magneto- or electro-encephalography (M/EEG) provide a unique window onto neural activity underlying cognitive processes.
+Typically, researchers are interested in testing whether and when such measures differ across conditions and/or groups.
+The conventional approach consists in conducting mass-univariate statistics through time followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster-based inference.
+However, these cluster-based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the onset or offset of effects (e.g., differences across conditions or groups).
+Here, we introduce a *model-based* approch for analysing M/EEG timeseries such as ERPs or timecourses of decoding performance and their differences across conditions or groups.
+This approach relies on Bayesian generalised additive multilevel models, which output the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between-subject variability present in such data.
+Using both simulation and actual EEG data, we show that the proposed approach largely outperforms conventional methods in determining both the onset and offset of M/EEG effects (e.g., ERPs difference, decoding performance), producting more precise and more reliable estimates.
+We provide an R package implementing the approach and illustrate how to integrate it into M/EEG statistical pipelines in MNE-Python."
+    )
+
+available:::namr(
+    title = "Precise temporal localisation of M/EEG effects with Bayesian generalised additive multilevel models",
+    verb = TRUE
+    )
+
 # to-do adding some code here...

### 2025-02-14 brms_meeg.tex
pushin new simulation results and updated title and abstract

@@
 
-\title{Modelling M/EEG data with Bayesian generalised additive
-multilevel models}
-
-\shorttitle{Bayesian M/EEG modelling}
+\title{Precise temporal localisation of M/EEG effects with Bayesian
+generalised additive multilevel models}
+
+\shorttitle{Bayesian modelling of M/EEG data}
 
@@
 offered by magneto- or electro-encephalography (M/EEG) provide a unique
-window onto neural activity underlying cognitive process and how they
-unfold over time. Typically, we are interested in testing whether such
-measures differ across conditions and/or groups. The conventional
-approach consists in conducting mass-univariate statistics through
-followed by some form of multiplicity correction (e.g., FDR, FWER) or
+window onto neural activity underlying cognitive processes. Typically,
+researchers are interested in testing whether and when such measures
+differ across conditions and/or groups. The conventional approach
+consists in conducting mass-univariate statistics through time followed
+by some form of multiplicity correction (e.g., FDR, FWER) or
 cluster-based inference. However, these cluster-based methods have an
@@
 to the cluster level, thus preventing any conclusion to be made about
-the onset and offset of effects (e.g., differences across conditions).
-Here, we introduce a novel \emph{model-based approch} for analysing
-one-dimensional M/EEG timeseries such as ERPs or decoding timecourses
-and their differences across conditions or group. This approach relies
-on Bayesian generalised additive multilevel models, which output the
+the onset or offset of effects (e.g., differences across conditions or
+groups). Here, we introduce a \emph{model-based} approch for analysing
+M/EEG timeseries such as ERPs or timecourses of decoding performance and
+their differences across conditions or groups. This approach relies on
+Bayesian generalised additive multilevel models, which output the
 posterior probabilility of the effect being above 0 (or above chance) at
 every timestep, while naturally taking into account the temporal
-dependencies and between-subject variability present in such data.}
+dependencies and between-subject variability present in such data. Using
+both simulation and actual EEG data, we show that the proposed approach
+largely outperforms conventional methods in determining both the onset
+and offset of M/EEG effects (e.g., ERPs difference, decoding
+performance), producting more precise and more reliable estimates. We
+provide an R package implementing the approach and illustrate how to
+integrate it into M/EEG statistical pipelines in MNE-Python.}
 
@@
 \newpage
-\section[Introduction]{Modelling M/EEG data with Bayesian generalised
-additive multilevel models}
+\section[Introduction]{Precise temporal localisation of M/EEG effects
+with Bayesian generalised additive multilevel models}
 
@@
 \citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al., 2019};
-\citeproc{ref-pernet2015}{C. R. Pernet et al., 2015})\ldots{} See also
-(\citeproc{ref-maris2011}{Maris, 2011})\ldots{} and
-(\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}) (history of
-cluster-based approaches and using a data split?)\ldots{} Cluster
-failure (\citeproc{ref-eklund2016}{Eklund et al., 2016})\ldots{}
-
-\subsection{Previous modelling work}\label{previous-modelling-work}
+\citeproc{ref-pernet2015}{Pernet et al., 2015})\ldots{} See also Maris
+(\citeproc{ref-maris2011}{2011}) and Rosenblatt et al.
+(\citeproc{ref-rosenblatt2018}{2018}) (history of cluster-based
+approaches)\ldots{} Clusters failures (\citeproc{ref-eklund2016}{Eklund
+et al., 2016})\ldots{}
+
+\subsection{Previous work}\label{previous-work}
 
@@
 Ullsperger, 2013}; \citeproc{ref-wuxfcllhorst2025}{Wüllhorst et al.,
-2025})\ldots{} ``Similar approaches have been successfully applied to
-EEG time- (Rousselet et al., 2008) and frequency-domain (Cohen and
-Cavanagh, 2011) data and allow the simultaneous investigation of
-multiple independent variables while preserving the high temporal
-resolution of the EEG.''\ldots{} See also (\citeproc{ref-hauk2006}{Hauk
-et al., 2006}; \citeproc{ref-rousselet2008}{Rousselet et al.,
-2008})\ldots{} Example of two-stage regression analysis (i.e.,
-individual-level then group-level, \citeproc{ref-dunagan2024}{Dunagan et
-al., 2024})\ldots{}
-
-From Dimigen \& Ehinger (\citeproc{ref-dimigen2021}{2021}): ``Recently,
-spline regression has been applied to ERPs (Hendrix, Baayen, \& Bolger,
-2017; Kryuchkova,Tucker, Wurm, \& Baayen,2012; Tremblay \& Baayen, 2010;
-Tremblay \& Newman, 2015)''\ldots{} GAMMs for EEG data
-(\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
+2025})\ldots{} See also (\citeproc{ref-hauk2006}{Hauk et al., 2006};
+\citeproc{ref-rousselet2008}{Rousselet et al., 2008})\ldots{} Example of
+two-stage regression analysis (i.e., individual-level then group-level,
+\citeproc{ref-dunagan2024}{Dunagan et al., 2024})\ldots{}
+
+From Dimigen \& Ehinger (\citeproc{ref-dimigen2021}{2021}): Recently,
+spline regression has been applied to ERPs (e.g.,
+\citeproc{ref-hendrix2017}{Hendrix et al., 2017};
+\citeproc{ref-kryuchkova2012}{Kryuchkova et al., 2012})\ldots{} GAMMs
+for EEG data (\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
 \citeproc{ref-meulman2015}{Meulman et al., 2015})\ldots{}
@@
 \citeproc{ref-skukies_modelling_2021}{Skukies \& Ehinger, 2021})\ldots{}
-Weighting single trials (\citeproc{ref-pernet2022}{C. Pernet,
-2022})\ldots{} The LIMO toolbox
-(https://onlinelibrary.wiley.com/doi/10.1155/2011/831409)\ldots{} Using
-Bayes factors (\citeproc{ref-teichmann2022}{Teichmann, 2022})\ldots{}
+Weighting single trials (\citeproc{ref-pernet2022}{Pernet,
+2022})\ldots{} The LIMO toolbox (\citeproc{ref-pernet2011}{Pernet et
+al., 2011})\ldots{}
+
+Using Bayes factors for decoding performance
+(\citeproc{ref-teichmann2022}{Teichmann, 2022})\ldots{}
+
+\subsection{Bayesian regression
+modelling}\label{bayesian-regression-modelling}
+
+Short intro/recap about Bayesian (linear and generalised) regression
+models\ldots{}
 
@@
 reference books (\citeproc{ref-hastie2017}{Hastie \& Tibshirani, 2017};
-\citeproc{ref-wood2017}{Wood, 2017})\ldots{}
+\citeproc{ref-wood2017}{Wood, 2017a})\ldots{} application to
+pupillometry (\citeproc{ref-vanrij2019}{Rij et al., 2019})\ldots{}
+GAMLSS for neuroimaging data (\citeproc{ref-dinga2021}{Dinga et al.,
+2021})\ldots{}
 
@@
 \(k\)th basis function \(b_{jk}()\) evaluated at the covariate value
-\(x_{ij}\) for the \(j\)th smooth function \(f_{j}\). Splines'
-coefficients are penalised\ldots{}
+\(x_{ij}\) for the \(j\)th smooth function \(f_{j}\). Splines
+coefficients are penalised (usually through the squared of the smooth
+functions' second derivative) in a way that can be interpreted as a
+prior on the ``wiggliness'' of the function\ldots{}
+
+\subsection{Bayesian generalised additive multilevel
+models}\label{bayesian-generalised-additive-multilevel-models}
+
+Now describe the Bayesian GAMM\ldots{} Proper inclusion of
+varying/random effects in the model specification protects against
+overly wiggly curves (\citeproc{ref-baayen2020}{Baayen \& Linke,
+2020})\ldots{}
 
@@
 
-\caption{\label{fig-eeg}Some ERPs in two conditions with 50 trials each,
-for a group of 20 participants.}
+\caption{\label{fig-eeg}Averaged (mean) simulated EEG activity in two
+conditions with 50 trials each, for a group of 20 participants. The
+error band represents the mean plus/minus 1 standard error of the mean.}
 
@@
 assessed examining trace plots as well as the Gelman--Rubin statistic
-\(\hat{R}\).
+\(\hat{R}\). The \texttt{brms} package uses the same syntax as the
+\texttt{R} package \texttt{mgcv} v 1.9-1 (\citeproc{ref-mgcv}{Wood,
+2017b}) for specifying smooth effects.
 
@@
 \texttt{condition} being above \(0+\epsilon\)
-(Figure~\ref{fig-post-prob-test}), with \(\epsilon\) defined as 10\% of
-the standard deviation of the raw EEG signal.
+(Figure~\ref{fig-post-prob-test}), with \(\epsilon := 0.05\), which can
+be interpreted as the smallest effect size of interest.
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.pdf}
+\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.pdf}
 
@@
 method that controls the familywise error rate (FWER) (i.e.,
-Holm--Bonferroni method, https://www.jstor.org/stable/4615733), and two
+Holm--Bonferroni method, \citeproc{ref-holm1979}{Holm, 1979}), and two
 cluster-based methods (permutation with a single threshold and TFCE,
@@
 
-Onset/offset estimation methods were assessed using these summary
-statistics of the Monte Carlo sampling distributions of onsets: median
-absolute error (MAE) and variance, with 10.000 iterations (i.e., 10.000
-simulated datasets)\ldots{}
+Onset/offset estimation methods were assessed using the median absolute
+error (MAE) and variance of 10.000 simulated datasets\ldots{}
 
@@
 \item
-  Then summarise the distribution of between-split difference across
-  folds
+  Then summarise the distribution of onset/offset with the median and
+  variance across folds
 \end{itemize}
 
-This will allow checking that the proposed approach produces more
-reliable onset/offset estimates.
+This will allow checking that the proposed approach produces reliable
+onset/offset estimates.
 
@@
 (i.e., time + sensors)\ldots{}
+
+We kept the exemplary models simple, but can be extended by adding
+varying/random effects (intercept and slope) for item (e.g.,
+word)\ldots{} but also continuous predictors at the trial level?
 
@@
 
+\bibitem[\citeproctext]{ref-dinga2021}
+Dinga, R., Fraza, C. J., Bayer, J. M. M., Kia, S. M., Beckmann, C. F.,
+\& Marquand, A. F. (2021). \emph{Normative modeling of neuroimaging data
+using generalized additive models of location scale and shape}.
+\url{http://dx.doi.org/10.1101/2021.06.14.448106}
+
 \bibitem[\citeproctext]{ref-dunagan2024}
@@
 
+\bibitem[\citeproctext]{ref-hendrix2017}
+Hendrix, P., Bolger, P., \& Baayen, H. (2017). Distinct ERP signatures
+of word frequency, phrase frequency, and prototypicality in speech
+production. \emph{Journal of Experimental Psychology: Learning, Memory,
+and Cognition}, \emph{43}(1), 128--149.
+\url{https://doi.org/10.1037/a0040332}
+
+\bibitem[\citeproctext]{ref-holm1979}
+Holm, S. (1979). A simple sequentially rejective multiple test
+procedure. \emph{Scandinavian Journal of Statistics}, \emph{6}(2),
+65--70. \url{http://www.jstor.org/stable/4615733}
+
 \bibitem[\citeproctext]{ref-tidybayes}
@@
 \url{https://doi.org/10.1016/j.tics.2014.01.002}
+
+\bibitem[\citeproctext]{ref-kryuchkova2012}
+Kryuchkova, T., Tucker, B. V., Wurm, L. H., \& Baayen, R. H. (2012).
+Danger and usefulness are detected early in auditory lexical processing:
+Evidence from electroencephalography. \emph{Brain and Language},
+\emph{122}(2), 81--91. \url{https://doi.org/10.1016/j.bandl.2012.05.005}
 
@@
 \bibitem[\citeproctext]{ref-pernet2022}
-Pernet, C. (2022). Electroencephalography robust statistical linear
+Pernet, C. R. (2022). Electroencephalography robust statistical linear
 modelling using a single weight per trial. \emph{Aperture Neuro},
@@
 \url{https://doi.org/10.52294/apertureneuro.2022.2.seoo9435}
+
+\bibitem[\citeproctext]{ref-pernet2011}
+Pernet, C. R., Chauveau, N., Gaspar, C., \& Rousselet, G. A. (2011).
+LIMO EEG: A Toolbox for Hierarchical LInear MOdeling of
+ElectroEncephaloGraphic Data. \emph{Computational Intelligence and
+Neuroscience}, \emph{2011}, 1--11.
+\url{https://doi.org/10.1155/2011/831409}
 
@@
 \url{https://www.R-project.org/}
+
+\bibitem[\citeproctext]{ref-vanrij2019}
+Rij, J. van, Hendriks, P., Rijn, H. van, Baayen, R. H., \& Wood, S. N.
+(2019). Analyzing the Time Course of Pupillometric Data. \emph{Trends in
+Hearing}, \emph{23}. \url{https://doi.org/10.1177/2331216519832483}
 
@@
 \bibitem[\citeproctext]{ref-wood2017}
-Wood, S. N. (2017). \emph{Generalized Additive Models}. Chapman;
+Wood, S. N. (2017a). \emph{Generalized Additive Models}. Chapman;
 Hall/CRC. \url{https://doi.org/10.1201/9781315370279}
+
+\bibitem[\citeproctext]{ref-mgcv}
+Wood, S. N. (2017b). \emph{Generalized additive models: An introduction
+with r} (2nd ed.). Chapman; Hall/CRC.
 
@@
 
-\section{\texorpdfstring{Using the \texttt{R} package and integration
-with
-\texttt{MNE-Python}}{Using the R package and integration with MNE-Python}}\label{using-the-r-package-and-integration-with-mne-python}
-
-Explain how to use the \texttt{R} package and to integrate it with
-\texttt{MNE} epochs\ldots{}
+\section{\texorpdfstring{Integration with
+\texttt{MNE-Python}}{Integration with MNE-Python}}\label{integration-with-mne-python}
+
+Explain how to use the \texttt{R} package with \texttt{MNE}
+epochs\ldots{}
 
@@
 \begin{Highlighting}[]
+\CommentTok{\# suggest a package name}
+\FunctionTok{library}\NormalTok{(available)}
+\FunctionTok{suggest}\NormalTok{(}
+    \AttributeTok{text =} \StringTok{"Time{-}resolved electrophysiological measurements such as those offered by magneto{-} or electro{-}encephalography (M/EEG) provide a unique window onto neural activity underlying cognitive processes.
+Typically, researchers are interested in testing whether and when such measures differ across conditions and/or groups.
+The conventional approach consists in conducting mass{-}univariate statistics through time followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster{-}based inference.
+However, these cluster{-}based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the onset or offset of effects (e.g., differences across conditions or groups).
+Here, we introduce a *model{-}based* approch for analysing M/EEG timeseries such as ERPs or timecourses of decoding performance and their differences across conditions or groups.
+This approach relies on Bayesian generalised additive multilevel models, which output the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between{-}subject variability present in such data.
+Using both simulation and actual EEG data, we show that the proposed approach largely outperforms conventional methods in determining both the onset and offset of M/EEG effects (e.g., ERPs difference, decoding performance), producting more precise and more reliable estimates.
+We provide an R package implementing the approach and illustrate how to integrate it into M/EEG statistical pipelines in MNE{-}Python."}
+\NormalTok{    )}
+
+\NormalTok{available}\SpecialCharTok{:::}\FunctionTok{namr}\NormalTok{(}
+    \AttributeTok{title =} \StringTok{"Precise temporal localisation of M/EEG effects with Bayesian generalised additive multilevel models"}\NormalTok{,}
+    \AttributeTok{verb =} \ConstantTok{TRUE}
+\NormalTok{    )}
+
 \CommentTok{\# to{-}do adding some code here...}

### 2025-02-14 code/errors_simulation.R

@@ -29,8 +29,7 @@ source("code/functions.R")
 meanpower <- unlist(read.table("code/meanpower.txt") )
 
 # defining the number of simulations to be performed
-# nsims <- 1e4
-nsims <- 10
+nsims <- 1e2
 
 # defining simulation parameters
 n_trials <- 50 # number of trials
@@ -43,8 +42,8 @@ ronset <- seq(from = 150, to = 170, by = 2) # random onset for each participant
 true_onset <- 0.160
 true_offset <- 0.342
 
-# defining an arbitrary alpha threshold
-aath <- 0.05
+# defining significance (alpha) thresholds
+significance_thresholds <- c(0.05, 0.01, 0.005, 0.001)
 
 # defining a function to generate a dataframe
 generate_data <- function (n_trials, n_ppt, outvar, srate, ronset) {
@@ -234,7 +233,7 @@ sim_results <- data.frame()
 for (i in seq(from = 1, to = nsims, by = 1) ) {
     
     # printing progress
-    cat("Simulation number:", i, "out of", nrow(nsims), "simulations...")
+    cat("Simulation number:", i, "out of", nsims, "simulations...\n")
     
     # simulating some EEG data
     raw_df <- generate_data(
@@ -341,49 +340,12 @@ for (i in seq(from = 1, to = nsims, by = 1) ) {
         
     }
     
-    # plotting the results
-    # onset_plot <- eps_threshold_results %>%
-    #     ggplot(aes(x = eps, y = threshold, fill = error_onset) ) +
-    #     geom_tile(show.legend = FALSE) +
-    #     geom_point(
-    #         data = eps_threshold_results %>%
-    #             dplyr::filter(error_onset == min(error_onset, na.rm = TRUE) ),
-    #         aes(x = eps, y = threshold),
-    #         color = "orangered", size = 2, shape = 4,
-    #         show.legend = FALSE
-    #         ) +
-    #     scale_x_continuous(expand = c(0, 0) ) +
-    #     scale_y_continuous(expand = c(0, 0) ) +
-    #     scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-    #     labs(
-    #         title = "Onset error",
-    #         x = "eps",
-    #         y = "Threshold",
-    #         fill = "Error"
-    #         )
-    # 
-    # offset_plot <- eps_threshold_results %>%
-    #     ggplot(aes(x = eps, y = threshold, fill = error_offset) ) +
[…]

### 2025-02-14 code/split_half_reliability.R

@@ -1,0 +1,216 @@
+###########################################################
+# Monte-Carlo simulation of onset/offset error properties #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on February 13, 2025                       #
+###########################################################
+
+# importing R packages
+library(changepoint)
+library(reticulate)
+library(tidyverse)
+library(tidybayes)
+library(brms)
+
+# importing Python modules
+np <- import("numpy")
+mne <- import("mne")
+
+# importing R version of Matlab code from Yeung et al. (2004)
+source("code/eeg_noise.R")
+
+# importing the ERP template with true onset = 160 ms, F=81, and max at F=126
+source("code/erp_template.R")
+
+# importing helper functions
+source("code/functions.R")
+
+# to use with the eeg_noise function
+meanpower <- unlist(read.table("code/meanpower.txt") )
+
+# defining the number of simulations to be performed
+# nsims <- 1e4
+nsims <- 20
+
+# defining simulation parameters
+n_trials <- 50 # number of trials
+n_ppt <- 20 # number of participants
+outvar <- 1 # noise variance
+srate <- 500 # sampling rate in Hz
+ronset <- seq(from = 150, to = 170, by = 2) # random onset for each participant
+
+# defining the true onset and offset in seconds
+true_onset <- 0.160
+true_offset <- 0.342
+
+# defining significance (alpha) thresholds
+significance_thresholds <- c(0.05, 0.01, 0.005, 0.001)
+
+# defining a function to generate a dataframe
+generate_data <- function (n_trials, n_ppt, outvar, srate, ronset) {
+    
+    for (P in 1:n_ppt) { # for each participant
+        
+        # get random onset
+        ponset <- sample(x = ronset, size = 1)
+        
+        # find starting point
+        st <- which(Xf == ponset)
+        
+        # pad vector
+        temp2 <- c(rep(0, st - 2), erp, rep(0, Nf - st - length(erp) + 2) )
+        
+        # initialising empty conditions
+        cond1 <- matrix(0, nrow = n_trials, ncol = Nf)
+        cond2 <- matrix(0, nrow = n_trials, ncol = Nf)
+        
+        for (trial in 1:n_trials) { # for each trial
+            
+            cond1[trial, ] <- temp1 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
[…]

### 2025-02-20 brms_meeg.qmd
adding preliminary results with actual MEG data

@@
 The conventional approach consists in conducting mass-univariate statistics through time followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster-based inference.
-However, these cluster-based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the onset or offset of effects (e.g., differences across conditions or groups).
-Here, we introduce a *model-based* approch for analysing M/EEG timeseries such as ERPs or timecourses of decoding performance and their differences across conditions or groups.
-This approach relies on Bayesian generalised additive multilevel models, which output the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between-subject variability present in such data.
-Using both simulation and actual EEG data, we show that the proposed approach largely outperforms conventional methods in determining both the onset and offset of M/EEG effects (e.g., ERPs difference, decoding performance), producting more precise and more reliable estimates.
+However, while allowing efficient error-rates control, cluster-based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the precise onset or offset of effects (e.g., differences across conditions or groups).
+Here, we introduce a *model-based* approch for analysing M/EEG timeseriesn such as ERPs or decoding performance through time, and their differences across conditions or groups.
+This approach relies on Bayesian generalised additive multilevel models and outputs the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between-subject variability present in such data.
+Using both simulated and actual M/EEG data, we show that the proposed approach largely outperforms conventional methods in estimating the onset and offset of M/EEG effects, producing more precise and more reliable estimates.
 We provide an R package implementing the approach and illustrate how to integrate it into M/EEG statistical pipelines in MNE-Python."
@@
 
+## Problem statement
+
+Description of cluster-based approaches [see @sassenhagen2019]...
+As pointed by the original authors themselves [@maris2007], "there There is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA \[false alarm\] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect" [@maris2007; @sassenhagen2019]...
+In contrast, the proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the probability of effects at the level of timesteps, voxels, sensors (etc), while naturally taking into account spatiotemporal dependencies.
+
 ## Previous work
@@
 
-From @dimigen2021: Recently, spline regression has been applied to ERPs [e.g., @hendrix2017; @kryuchkova2012]... GAMMs for EEG data [@abugaber2023; @meulman2015]...
+See also the rERP framework [@smith2014a; @smith2014b] and @tremblay2014...
+
+From @dimigen2021: Recently, spline regression has been applied to ERPs [e.g., @hendrix2017; @kryuchkova2012]... GAMMs for EEG data [@abugaber2023; @meulman2015; @meulman2023]...
 
@@
 
-Using Bayes factors for decoding performance [@teichmann2022]...
+Recently, @teichmann2022 provided a detailed tutorial on using Bayes factors (BFs) to analyse the 1D or 2D output from MVPA, that is, for testing, at every timestep, whether decoding performance is above chance level.
+However, this approach provides timeseries of BFs that ignores temporal dependencies..
 
@@
 
-See for instance these tutorials [@sóskuthy2017; @winter2016] or application to phonetic data [@wieling2018] or this introduction [@baayen2020] or these reference books [@hastie2017; @wood2017]...
+See for instance these tutorials [@sóskuthy2017; @winter2016] or application to phonetic data [@wieling2018; @sóskuthy2021] or this introduction [@baayen2020] or these reference books [@hastie2017; @wood2017]...
 application to pupillometry [@vanrij2019]...
 GAMLSS for neuroimaging data [@dinga2021]...
+Modelling auto-correlation in GAMMs + EEG example [@baayen2018]...
 
@@
 y_{i} &\sim \mathrm{EF}\left(\mu_{i}, \phi\right)\\
-g\left(\mu_i\right) &=  \underbrace{A_{i} + \mathbf{X}_{i} \gamma}_{\text{parametric part}} + \underbrace{\sum_{j=1}^{J} f_{j}\left(x_{ij}\right)}_{\text{non-parametric part}}
+g\left(\mu_i\right) &= \underbrace{\mathbf{A}_{i} \gamma}_{\text{parametric part}} + \underbrace{\sum_{j=1}^{J} f_{j}\left(x_{ij}\right)}_{\text{non-parametric part}}
 \end{aligned}
@@
 
-where $y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)$ denotes that the observations $y_{i}$ are distributed as some member of the exponential family of distributions (e.g., Gaussian, Gamma, Beta, Poisson) with mean $\mu_{i}$ and scale parameter $\phi$; $g(\cdot)$ is the link function, $A$ is an offset, $\mathbf{X}_{i}$ is the $i$th row of a parametric model matrix, $\gamma$ is a vector of parameters for the parametric terms, $f_{j}$ is a smooth function of covariate $x_{j}$.
+where $y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)$ denotes that the observations $y_{i}$ are distributed as some member of the exponential family of distributions (e.g., Gaussian, Gamma, Beta, Poisson) with mean $\mu_{i}$ and scale parameter $\phi$; $g(\cdot)$ is the link function, $\mathbf{A}_{i}$ is the $i$th row of a known parametric model matrix, $\gamma$ is a vector of parameters for the parametric terms (to be estimated), $f_{j}$ is a smooth function of covariate $x_{j}$ (to be estimated as well).
 The smooth functions $f_{j}$ are represented in the model via penalised splines basis expansions of the covariates, that are a weighted sum of basis functions:
@@
 
+Terminology: *splines* are functions composed of simpler functions.
+These simpler functions are basis functions (e.g., a cubic polynomial) and the set of basis functions is a *basis*.
+Each basis function gets its coefficient and the resultant spline is the sum of these weighted basis functions.
+
 ## Bayesian generalised additive multilevel models
 
-Now describe the Bayesian GAMM... Proper inclusion of varying/random effects in the model specification protects against overly wiggly curves [@baayen2020]...
+Now describe the Bayesian GAMM [@miller2025]...
+Proper inclusion of varying/random effects in the model specification protects against overly wiggly curves [@baayen2020]...
+Generalising to scale and shape or "distributional GAMs" [@rigby2005; @umlauf2018] and applied to neuroimaging data [@dinga2021]...
+
+Instead of averaring, obtain the smooth ERP signal from multilevel GAM... less susceptible to outliers (Meulman et al., 2023)...
 
@@
 
-Following the approach used by @sassenhagen2019 and @rousselet_using_2025, we simulated EEG data stemming from two conditions, one with noise only, and the other with noise + signal.
+Following the approach of @sassenhagen2019 and @rousselet_using_2025, we simulated EEG data stemming from two conditions, one with noise only, and the other with noise + signal.
 As in previous studies, the noise was generated by superimposing 50 sinusoids at different frequencies, following an EEG-like spectrum [see details and code in @yeung2004].
@@
 We then fitted a Bayesian GAM using the `brms` package [@brms2017; @brms2018; @nalborczyk2019].
-We used the default priors in `brms`, that is, weakly informative priors.
+We used the default priors in `brms` (i.e., weakly informative priors).
 We ran eight Markov Chain Monte-Carlo (MCMC) to approximate the posterior distribution, including each 5000 iterations and a warmup of 2000 iterations, yielding a total of $8 \times (5000-2000) = 24000$ posterior samples to use for inference.
@@
 The `brms` package uses the same syntax as `r pkrt("mgcv")` for specifying smooth effects.
+@fig-post-prob-test shows the predictions of this model together with the raw data.
 
@@
     # cubic regression splines with k-1 basis functions
-    eeg ~ condition + s(time, bs = "cr", k = 10, by = condition),
+    eeg ~ condition + s(time, bs = "cr", k = 20, by = condition),
     data = ppt_df,
@@
     eeg ~ condition +
-        s(time, bs = "cr", k = 10, by = condition) +
+        s(time, bs = "cr", k = 20, by = condition) +
         s(time, participant, bs = "fs"),
@@
 
-```{r fig-post-prob-ratio, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Ratio of posterior probability according to the GAM. Timesteps above threshold (10) are highlighted in green."}
+```{=html}
+<!--
+
+Maybe better to use the p-MAP (i.e., density ratio between the null and the MAP)?
+And/or look at the derivative (e.g., <https://fromthebottomoftheheap.net/2014/05/15/identifying-periods-of-change-with-gams/>)?
+
+-->
+```
+
+```{r fig-post-prob-ratio, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Ratio of posterior probability according to the GAM (on a log10 scale).
+Timesteps above threshold (10) are highlighted in green."}
 # defining some parameters
@@
 #     relocate(model, .before = cluster_onset)
+
 ```
@@
 Next we fit a hierarchical/multilevel GAM using summary statistics of ERPs (mean and SD) at the participant level (similar to what is done in meta-analysis).
+Although it is possible to fit a GAMM at the single-trial level, we present a resource-efficient version of the model that is fitted directly on the by-participant summary statistics (mean and SD)...
 
@@
     eeg_mean | se(eeg_sd) ~
-        condition + s(time, bs = "cr", k = 10, by = condition) +
+        condition + s(time, bs = "cr", k = 20, by = condition) +
         (1 | participant),
@@
 
-Assessing the reliability of the proposed approach using some sort of split-half reliability [e.g., @rosenblatt2018]?
-Using the MEG decoding results of Nalborczyk et al.
-(in preparation) in 33 participants:
-
--   Create many (e.g., 10 or 20) half train/test splits of the data
--   For each fold, estimate the onset/offset on both splits using all methods
--   Then summarise the distribution of onset/offset with the median and variance across folds
-
-This will allow checking that the proposed approach produces reliable onset/offset estimates.
-
-```{r reliability, echo = FALSE, eval = FALSE}
-# library(caret)
-# see https://github.com/topepo/caret/
-# https://topepo.github.io/caret/data-splitting.html
-# and createDataPartition()
-# and createFolds()
-# trainIndex <- createDataPartition(
-#     y = summary_df$participant,
-#     p = 0.5,
-#     list = FALSE,
-#     times = 10
-#     )
-
-# subsetting the data
-# summary_df_train <- summary_df[trainIndex, ]
-# summary_df_test  <- summary_df[-trainIndex, ]
-
-# getting unique participants
-unique_participants <- unique(summary_df$participant)
-
-# getting the number of participants
-num_participants <- length(unique_participants)
-
-# ensuring participants are randomly shuffled
-shuffled_participants <- sample(unique_participants)
-
-# defining the number of splits
-num_splits <- 10
-
-# creating a dataframe where each participant appears in every split
-split_df <- crossing(participant = unique_participants, split_number = 1:10)
-
-# assigning alternating groups (A or B) within each split for each participant
-# split_df <- split_df %>% mutate(group = rep(c("A", "B"), length.out = n() ) )
-split_df <- split_df %>%
-    group_by(split_number) %>%
-    # shuffling groups for each split
-    mutate(group = sample(rep(c("A", "B"), length.out = n() ) ) ) %>%
-    ungroup()
-
-# sanity checks
-# table(split_df$group)
-# split_df %>%
-#     ggplot(aes(x = group) ) +
-#     geom_histogram(stat = "count") +
-#     facet_wrap(~participant, ncol = 5)
-
-# now for each split, assess the onset and offset
-participants_split1_train <- split_df %>%
-    filter(split_number == 1 & group == "A") %>%
-    pull(participant)
-participants_split1_test <- split_df %>%
-    filter(split_number == 1 & group == "B") %>%
-    pull(participant)
-
-# split1 data
-summary_df_split1_train <- summary_df %>%
-    filter(participant %in% participants_split1_train)
-
-# fitting the GAM
-meta_gam_split1_train <- brm(
-    eeg_mean | se(eeg_sd) ~
-        condition + s(time, bs = "cr", k = 10, by = condition) +
-        (1 | participant),
-    data = summary_df_split1_train,
-    family = gaussian(),
+Next, we assessed the performance of the proposed approach on actual MEG data (decoding results of Nalborczyk et al., in preparation).
+We conducted time-resolved multivariate pattern analysis (MVPA), also known as decoding...
+As a result, we have a timecourse of decoding performance (ROC AUC), bounded between 0 and 1, for each participant (for a total of 32 participants).
+Now, we want to *test* whether the group-level average decoding accuracy is above chance (i.e., 0.5) at each timestep (@fig-decoding-data).
+We fitted a similar GAM as discussed previously, but we replaced the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one to account for the bounded nature of AUC values (between 0 and 1) [for a tutorial on Beta regression, see @coretta2025].
+Note that although we chose a basis dimension of $k=50$, which seems appropriate for the present data, this choice should be adapted according to the properties of the modelled data (e.g., signal-to-noise ratio, prior low-pass filtering, sampling rate, etc) and should be assessed by the usual model checking tools (e.g., posterior predictive checks).
+We also define a region of practical equivalence [ROPE, @kruschke2017], defined as the chance level plus the standard deviation of the (group-level average) decoding performance during the baseline period.
+This ensures that...[^1]
+
+[^1]: Could an alternative approach be to include a predictor for "baseline vs. after baseline"?
+
+```{r decoding-model, eval = FALSE}
+# fitting the Beta GAM
+meg_decoding_gam <- brm(
+    auc ~ s(time, bs = "cr", k = 50),
+    data = decoding_df,
+    family = Beta(),
     warmup = 2000,
@@
 
+We assessed the reliability of the proposed approach using a form of permutation-based split-half reliability [e.g., @rosenblatt2018], which consisted of the following steps:
+
+-   Create many (e.g., 10 or 20) half train/test splits of the data
+-   For each fold, estimate the onset/offset on both splits using all methods
+-   Then summarise the distribution of onset/offset with the median and variance across folds
+
+This will allow checking that the proposed approach produces reliable onset/offset estimates.
+
+```{r fig-decoding-data, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Group-level average decoding performance (N=32) superimposed with the GAM predictions (in blue) and the region of practical equivalence (ROPE, in orange) computed from the baseline period.
+The blue horizontal line indicates the timesteps at which the posterior probability ratio is equal to or greater than 20."}
+# importing the reshaped MEG decoding results
+decoding_df <- read.csv(file = "code/decoding_results_reshaped.csv") %>%
+    # removing some participants
+    mutate(participant_id = cur_group_id(), .by = participant) %>%
+    filter(participant_id < 33) %>%
+    select(-participant_id)
+
+# averaging across participants
+decoding_summary_df <- decoding_df %>%
+    summarise(
+        auc_mean = mean(auc),
+        auc_sd = sd(auc),
+        .by = c(participant, time)
+        )
+
+# defining the chance level
+chance_level <- 0.5
+
+# computing the smallest effect size of interest (SESOI)
+# as the SD of decoding performance during the baseline
+sesoi <- decoding_summary_df %>%
+    filter(time < 0) %>%
+    summarise(auc_mean = mean(auc_mean), .by = time) %>%
+    summarise(sesoi = sd(auc_mean) ) %>%
+    # summarise(sesoi = sd(auc_mean) / sqrt(n_distinct(decoding_summary_df$participant) ) ) %>%
+    pull(sesoi)
+
+# fitting the Beta GAM
+meg_decoding_gam <- brm(
+    auc_mean ~ s(time, bs = "cr", k = 50),
+    data = decoding_summary_df,
+    family = Beta(),
+    warmup = 2000,
+    iter = 5000,
+    chains = 4,
+    cores = 4,
+    file = "models/meg_decoding_gam.rds"
+    )
+
+# some posterior predictive checks
+# pp_check(gam_split, prefix = "ppd")
+# pp_check(gam_split)
+# pp_check(gam_split, type = "error_hist")
+# pp_check(gam_split, type = "stat_2d")
+# pp_check(gam_split, type = "ecdf_overlay")
+
+# computing the posterior probability
+prob_y_above_0 <- data.frame(time = unique(meg_decoding_gam$data$time) ) %>%
+    # retrieving the posterior samples
+    add_epred_draws(object = meg_decoding_gam) %>%
+    # converting to dataframe
+    data.frame() %>%
+    # computing mean posterior probability at the group level
+    group_by(time) %>%
+    summarise(m = mean(.epred > (0 + chance_level + sesoi) ) ) %>%
+    mutate(prob_ratio = m / (1 - m) ) %>%
+    # ensuring there is no 0 or +Inf values
+    mutate(prob_ratio = ifelse(is.infinite(prob_ratio), ndraws(meg_decoding_gam), prob_ratio) ) %>%
+    mutate(prob_ratio = ifelse(prob_ratio == 0, 1 / ndraws(meg_decoding_gam), prob_ratio) ) %>%
+    ungroup()
+
+# finding onset and offset for the current threshold
+# exceeding_times <- prob_y_above_0 %>%
+#     dplyr::filter(prob_ratio > post_prob_ratio_threshold) %>%
+#     summarise(
+#         cluster_onset = min(time, na.rm = TRUE),
+#         cluster_offset = max(time, na.rm = TRUE)
+#         )
+# or finding all clusters
+# defining a function to find onset and offset in timeseries
+# find_clusters <- function (mask, timeseries) {
+#     
+#     # initialising onset and offset values
+#     onset <- NA
+#     offset <- NA
+#     
+#     # identifying the onset and offset
+#     masked_timesteps <- try(timeseries[which(mask)], silent = TRUE)
+#     onset <- head(x = masked_timesteps, n = 1)
+#     offset <- tail(x = masked_timesteps, n = 1)
+#     
+#     # returning the onset and offset
+#     return (c(onset, offset) )
+#     
+# }
+# 
+# find_onset_offset(
+#         tests_results$pval <= alpha_level,
+#         tests_results$time
+#         )
+# prob_y_above_0 %>%
+#     dplyr::filter(prob_ratio >= post_prob_ratio_threshold) %>%
+#     dplyr::pull(time)
+find_clusters <- function (df, threshold = 20) {
+
+    df %>%
+        mutate(above_threshold = prob_ratio >= threshold) %>%
+        mutate(cluster_change = c(TRUE, diff(above_threshold) != 0) ) %>%
+        filter(above_threshold) %>%
+        mutate(cluster_id = cumsum(cluster_change)) %>%
+        group_by(cluster_id) %>%
+        summarise(cluster_onset = first(time), cluster_offset = last(time), .groups = "drop") %>%
+        select(cluster_onset, cluster_offset)
+
+}
+
+post_prob_ratio_threshold <- 20
+clusters <- find_clusters(prob_y_above_0, threshold = post_prob_ratio_threshold)
+
+# checking model's predictions against raw data
+plot(
+    conditional_effects(x = meg_decoding_gam),
+    line_args = list(colour = "steelblue", fill = "steelblue", alpha = 0.3),
+    points = FALSE, plot = FALSE
+    )[[1]] +
+    geom_hline(yintercept = 0.5, linetype = 2) +
+    geom_vline(xintercept = 0.0, linetype = 2) +
+    annotate(
+        geom = "rect",
+        xmin = -Inf, xmax = Inf,
+        ymin = chance_level - sesoi, ymax = chance_level + sesoi,
+        fill ="orangered",
+        alpha = 0.2
+        ) +
+    geom_segment(
+        data = clusters,
+        aes(
+            x = cluster_onset,
+            xend = cluster_offset,
+            y = 0.48, yend = 0.48
+            ),
+        colour = "steelblue",
+        inherit.aes = FALSE,
+        lineend = "round", linewidth = 2
+        ) +
+    geom_line(
+        data = decoding_summary_df %>% summarise(auc_mean = mean(auc_mean), .by = time),
+        aes(x = time, y = auc_mean), inherit.aes = FALSE
+        ) +
+    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
+```
+
+\newpage
+
 # Results
 
-...
+This section is divided in two parts.
+First, we present the results from the simulation study, assessing the bias and variance of each method when applied to simulated data in which the ground truth is known.
+Second, we present the results obtained when applying the different methods to actual MEG data (decoding performance through time), assessing the reliability of each method and the stability of its estimates.
 
@@
             grepl("offset", measure) ~ abs(value - true_offset)
+            # grepl("onset", measure) ~ value - true_onset,
+            # grepl("offset", measure) ~ value - true_offset
             )
@@
         ) +
+    # geom_vline(xintercept = 0, linetype = 2) +
     geom_point(
@@
     facet_wrap(~type) +
-    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
     labs(x = "Median absolute error (s)", y = "Variance (log-scale)")
@@
 
-```{r fig-simulation-distribution, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Distributions of onset and offset estimates for each method."}
+```{r fig-simulation-distribution, echo = FALSE, fig.width = 9, fig.cap = "Distributions of onset and offset estimates for each method."}
 sim_results %>%
@@
         ) %>%
+    # mutate(
+    #     plot_row = ifelse(
+    #         method %in% c("p-value", "FDR BH95", "FDR BY01", "Holm"), 0, 1
+    #         )
+    #     ) %>%
     mutate(
-        plot_row = ifelse(
-            method %in% c("p-value", "FDR BH95", "FDR BY01", "Holm"), 0, 1
+        plot_row = case_when(
+            method %in% c("Raw p-value", "Holm") ~ 1,
+            method %in% c("FDR BH95", "FDR BY01") ~ 2,
+            method %in% c("Cluster-mass", "TFCE") ~ 3,
+            method %in% c("Change point", "brms") ~ 4
             )
@@
     theme_light(base_size = 12, base_family = "Open Sans") +
-    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
     labs(x = "Onset/Offset (s)", y = "Density")
@@
 
-## Application to actual EEG data (reliability)
+\newpage
+
+## Application to actual MEG data (reliability)
+
+@fig-onset-offset shows the group-level average decoding performance through time with onset and offset estimates for each method.
+The `brms_full` method is similar to the `brms` method except that the ROPE is defined on the entire dataset rather than on the split dataset.
+Overall, this figure shows that both the `Raw p-value` and `FDR BH95` methods were extremely lenient, considering that the decoding performance was above chance before the onset of the stimulus (false positive) and until the end of the trial.
+The `Change point` and `Cluster mass` methods were the most conservative methods, identifying a time window from approximately +60ms to +500ms.
+The `Holm`, `TFCE`, `brms`, and `brms_full` methods produced somewhat similar estimates of onset and offset, from approximately +60ms to +650ms.^[It should be noted that although each method can produce several "clusters" of timesteps, we only considered the first (onset) and last (offset) timesteps identified by each method to compute the error (difference).]
+
+```{r fig-onset-offset, echo = FALSE, fig.width = 9, fig.asp = 0.7, fig.cap = "Group-level average decoding performance through time with onset and offset estimates for each method (data from Nalborczyk et al., in preparation)."}
+# loading the results
+reliability_results <- readRDS(file = "results/reliability_results.rds")
+
+# extract full data estimates
+full_reliability_results <- reliability_results %>%
+    tail(1) %>%
+    pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
+    data.frame() %>%
+    select(-split_id) %>%
+    separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
+    pivot_wider(names_from = type, values_from = value) %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c(
+                "p", "bh", "by", "holm", "cluster_mass",
+                "cluster_tfce", "cpt", "brms", "brms_full"
+                ),
+            labels = c(
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
+                "TFCE", "Change point", "brms", "brms_full"
+                )
+            )
+        ) %>%
+    # arrange(method) %>%
+    mutate(y_offset = 0.5 - as.numeric(factor(method) ) * 0.01 - 0.01)
+
+# checking model's predictions against raw data
+decoding_df %>%
+    summarise(
+        auc_mean = mean(auc),
+        auc_sd = sd(auc),
+        auc_se = sd(auc) / sqrt(n() ),
+        .by = time
+        ) %>%
+    # summarise(auc_mean = mean(auc_mean), .by = time) %>%
+    ggplot(aes(x = time, y = auc_mean) ) +
+    geom_hline(yintercept = 0.5, linetype = 2) +
+    geom_vline(xintercept = 0.0, linetype = 2) +
+    # annotate(
+    #     geom = "rect",
+    #     xmin = -Inf, xmax = Inf,
+    #     ymin = chance_level - sesoi, ymax = chance_level + sesoi,
+    #     fill ="orangered",
+    #     alpha = 0.2
+    #     ) +
+    geom_segment(
+        data = full_reliability_results,
+        aes(
+            x = onset, xend = offset,
+            y = y_offset, yend = y_offset,
+            colour = method,
+            ),
+        inherit.aes = FALSE,
+        lineend = "round", linewidth = 2
+        ) +
+    geom_ribbon(
+        aes(ymin = auc_mean - auc_se, ymax = auc_mean + auc_se),
+        alpha = 0.2,
+        ) +
+    geom_line(
+        data = decoding_summary_df %>%
+            summarise(auc_mean = mean(auc_mean), .by = time),
+        aes(x = time, y = auc_mean), inherit.aes = FALSE
+        ) +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)", colour = NULL)
+
+# saving the plot
+# ggsave(
+#     filename = "figures/meg_decoding_results_all_methods.png",
+#     width = 12, height = 6, dpi = 300,
+#     device = "png"
+#     )
+```
+
+@fig-reliability shows two properties of each method: i) the median difference between the onset and offset estimates from each data split and the onset and offset estimates from the full dataset (x-axis) and ii) and the variance of its onset and offset estimates across data splits (y-axis).
+This figure reveals that the `brms` *onset and offset* estimates on each split are the closest to the estimates from the full dataset on average (0ms difference for the onset estimate and 5ms difference for the offset estimate).
+The `Raw p-value` method has similar performance, but given the aberrant estimates it produces (cf.
+@fig-onset-offset), the result that it is consistent between data splits and the full dataset is not convincing on its own.
+Overall, the figure reveals that for all other methods, split datasets produce later onset estimates and earlier offset estimates (as compared to the estimates from the model fitted on the full dataset).
+
+```{r fig-reliability, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Median absolute error and variance of onset (left) and offset (right) estimates for each method."}
+# extract full data estimates
+full_reliability_results <- reliability_results %>%
+    tail(1) %>%
+    pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
+    data.frame() %>%
+    select(-split_id)
+
+# analysing the split results
+reliability_results %>%
+    slice(1:nrow(.)-1) %>%
+    pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
+    data.frame() %>%
+    left_join(
+        full_reliability_results,
+        by = "measure",
+        suffix = c("_split", "_full")
+        ) %>%
+    # mutate(error = abs(value_split - value_full) ) %>%
+    mutate(error = value_split - value_full) %>%
+    # mutate(error = (value_split - value_full)^2) %>%
+    separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
+    group_by(type, method) %>%
+    summarise(
+        MAE = median(error, na.rm = TRUE),
+        # variance = var(error, na.rm = TRUE),
+        variance = var(value_split, na.rm = TRUE)
+        # variance = sum(error, na.rm = TRUE)
+        # .groups = "drop"
+        ) %>%
+    ungroup() %>%
+    # now plotting
+    mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c(
+                "p", "bh", "by", "holm", "cluster_mass",
+                "cluster_tfce", "cpt", "brms", "brms_full"
+                ),
+            labels = c(
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
+                "TFCE", "Change point", "brms", "brms_full"
+                ),
+            )
+        ) %>%
+    ggplot(
+        aes(
+            x = MAE,
+            y = variance,
+            colour = method,
+            fill = method
+            )
+        ) +
+    geom_vline(xintercept = 0, linetype = 2) +
+    geom_point(
+        size = 2,
+        pch = 21,
+        colour = "white",
+        show.legend = FALSE
+        ) +
+    geom_label_repel(
+        aes(label = method),
+        colour = "white",
+        size = 3,
+        segment.color = NA,
+        max.overlaps = 20,
+        show.legend = FALSE
+        ) +
+    # scale_y_log10() +
+    facet_wrap(~type, scales = "free") +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    labs(
+        x = "Average (median) difference to full dataset (s)",
+        y = "Variance across splits"
+        )
+```
+
+```{r fig-reliability-distributions, echo = FALSE, eval = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Median absolute error and variance of onset (left) and offset (right) estimates for each method."}
+# plotting distributions
+reliability_results %>%
+    slice(1:nrow(.)-1) %>%
+    pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
+    data.frame() %>%
+    left_join(
+        full_reliability_results,
+        by = "measure",
+        suffix = c("_split", "_full")
+        ) %>%
+    # mutate(error = abs(value_split - value_full) ) %>%
+    mutate(error = value_split - value_full) %>%
+    # mutate(error = (value_split - value_full)^2) %>%
+    separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
+    # now plotting
+    mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c(
+                "p", "bh", "by", "holm", "cluster_mass",
+                "cluster_tfce", "cpt", "brms", "brms_full"
+                ),
+            labels = c(
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
+                "TFCE", "Change point", "brms", "brms_full"
+                ),
+            )
+        ) %>%
+    mutate(
+        plot_row = case_when(
+            method %in% c("Raw p-value", "Holm") ~ 1,
+            method %in% c("FDR BH95", "FDR BY01") ~ 2,
+            method %in% c("Cluster-mass", "TFCE") ~ 3,
+            method %in% c("Change point", "brms") ~ 4
+            )
+        ) %>%
+    ggplot(
+        aes(
+            x = error,
+            colour = method,
+            fill = method
+            )
+        ) +
+    geom_vline(xintercept = 0, linetype = 2) +
+    geom_density(aes(fill = NULL) ) +
+    facet_grid(plot_row~type, scales = "free_x") +
+    theme_light(base_size = 12, base_family = "Open Sans") +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    labs(x = "Onset/Offset (s)", y = "Density")
+```
+
+\newpage
+
+# Discussion
 
@@
 
-\newpage
-
-# Discussion
-
-...
-
 ## Summary of the proposed approach
 
-...
+Overall, before concluding on the onset/offset of effect based on the model, we need to ensure that the model provides a faithful description of the data-generating process (e.g., via posterior predictive checks etc)...
 
@@
 
+As in previous simulation work [e.g., @rousselet2008; @sassenhagen2019], the present simulation results depend on various choices such as the specific cluster-forming algorithm and threshold, signal-to-noise ratio, negative impact of preprocessing steps (e.g., low-pass filter) on temporal resolution...
+note however, that the same caveats apply to all methods...
+
+The error properties depend on the threshold parameter, a value of 10 or 20 seems to be a reasonable default, but the optimal threshold parameter can be adjusted using split-half reliability assessment...
+
 Can be applied to any 1D timeseries (e.g., pupillometry, electromyography)... Extending the approach to spatiotemporal data (i.e., time + sensors)...
@@
 We kept the exemplary models simple, but can be extended by adding varying/random effects (intercept and slope) for item (e.g., word)... but also continuous predictors at the trial level?
-
-The error properties depend on the threshold parameter, a value of 10 or 20 seems to be a reasonable default, but the optimal threshold parameter can be adjusted using split-half reliability assessment...
 
@@
 # Appendix
+
+````{=html}
+<!--
 
@@
 
+-->
+````
+
 # Application to 2D time-resolved decoding results (cross-temporal generalisation)
@@
     # 2D thin-plate spline (tp)
-    # auc ~ s(train_time, test_time, bs = "tp", k = 10),
-    auc ~ t2(train_time, test_time, bs = "tp", k = 10),
+    auc ~ t2(train_time, test_time, bs = "tp", k = 20),
     data = timegen_data,
@@
 
+Could be extended to spatial and temporal dimensions with formulas such as `te(x, y, Time, d = c(2, 1) )`...
+
 \newpage
@@
 
+```{=html}
+<!--
+
 # Threshold-free cluster enhancement
@@
 Cluster-based permutation approaches require defining a cluster-forming threshold (e.g., a t- or f-value) as the initial step of the algorithm.
-As different cluster-forming thresholds lead to clusters with different spatial or temporal extent, this threshold modulates the sensitivity of the subsequent permutation test.
-The threshold-free cluster enhancement method (TFCE) was introduced by @smith2009 to overcome this arbitrary threshold.
+As different cluster-forming thresholds lead to clusters with different spatial or temporal extent, this initial threshold modulates the sensitivity of the subsequent permutation test.
+The threshold-free cluster enhancement method (TFCE) was introduced by @smith2009 to overcome this choice of an arbitrary threshold.
 
 In brief, the TFCE method works as follows.
-Instead of picking an arbitrary cluster-forming threshold (e.g., $t=2$), we try all (or many) possible thresholds in a given range and check whether a given timestep/voxel belongs to a significant cluster under any of the set of thresholds...
-Then, instead of using cluster mass, we use a weighted average between the cluster extend ($e$, how broad is the cluster, that is, how many connected samples it contains) and the cluster height ($h$, how high is the cluster, that is, how large is the test statistic) according to the formula:
+Instead of picking an arbitrary cluster-forming threshold (e.g., $t=2$), we try all (or many) possible thresholds in a given range and check whether a given timestep/voxel belongs to a significant cluster under any of the set of thresholds.
+Then, instead of using cluster mass (e.g., the sum of squared t-values within the cluster), we use a weighted average between the cluster extend ($e$, how broad is the cluster, that is, how many connected samples it contains) and the cluster height ($h$, how high is the cluster, that is, how large is the test statistic).
+The TFCE score at each timestep/voxel $t$ is given by:
 
 $$
-\text{TFCE} = \int_{h} e(h)^{E} h^{H} \mathrm{d}h
+\text{TFCE(t)} = \int_{h=h_{0}}^{h_{t}} e(h)^{E} h^{H} \mathrm{d}h
 $$
 
-Where...
-the parameters $E$ and $H$ are set a priori and control the influence of the extend and height on the TFCE.
-Then, p-value for timestep/voxel $i$ is computed by comparing it TFCE with the null distribution of TFCE values.
+where $h_{0}$ is typically $0$ and parameters $E$ and $H$ are set a priori (typically to 0.5 and 2, respectively) and control the influence of the extend and height on the TFCE.
+Then, p-value for timestep/voxel $t$ is computed by comparing it TFCE with the null distribution of TFCE values.
 For each permuted signal, we keep the maximal value over the whole signal for the null distribution of the TFCE....
@@
 
+-->
+```
+
 # Integration with `MNE-Python`
@@
 ```{r r-python, eval = FALSE}
-# suggest a package name
-library(available)
-suggest(
-    text = "Time-resolved electrophysiological measurements such as those offered by magneto- or electro-encephalography (M/EEG) provide a unique window onto neural activity underlying cognitive processes.
-Typically, researchers are interested in testing whether and when such measures differ across conditions and/or groups.
-The conventional approach consists in conducting mass-univariate statistics through time followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster-based inference.
-However, these cluster-based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the onset or offset of effects (e.g., differences across conditions or groups).
-Here, we introduce a *model-based* approch for analysing M/EEG timeseries such as ERPs or timecourses of decoding performance and their differences across conditions or groups.
-This approach relies on Bayesian generalised additive multilevel models, which output the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between-subject variability present in such data.
-Using both simulation and actual EEG data, we show that the proposed approach largely outperforms conventional methods in determining both the onset and offset of M/EEG effects (e.g., ERPs difference, decoding performance), producting more precise and more reliable estimates.
-We provide an R package implementing the approach and illustrate how to integrate it into M/EEG statistical pipelines in MNE-Python."
-    )
-
-available:::namr(
-    title = "Precise temporal localisation of M/EEG effects with Bayesian generalised additive multilevel models",
-    verb = TRUE
-    )
-
-# to-do adding some code here...
-```
+# TO-DO: adding some code here...
+```

### 2025-02-20 brms_meeg.tex
adding preliminary results with actual MEG data

@@
 by some form of multiplicity correction (e.g., FDR, FWER) or
-cluster-based inference. However, these cluster-based methods have an
-important downside: they shift the focus of inference from the timepoint
-to the cluster level, thus preventing any conclusion to be made about
-the onset or offset of effects (e.g., differences across conditions or
-groups). Here, we introduce a \emph{model-based} approch for analysing
-M/EEG timeseries such as ERPs or timecourses of decoding performance and
-their differences across conditions or groups. This approach relies on
-Bayesian generalised additive multilevel models, which output the
-posterior probabilility of the effect being above 0 (or above chance) at
-every timestep, while naturally taking into account the temporal
-dependencies and between-subject variability present in such data. Using
-both simulation and actual EEG data, we show that the proposed approach
-largely outperforms conventional methods in determining both the onset
-and offset of M/EEG effects (e.g., ERPs difference, decoding
-performance), producting more precise and more reliable estimates. We
-provide an R package implementing the approach and illustrate how to
-integrate it into M/EEG statistical pipelines in MNE-Python.}
+cluster-based inference. However, while allowing efficient error-rates
+control, cluster-based methods have an important downside: they shift
+the focus of inference from the timepoint to the cluster level, thus
+preventing any conclusion to be made about the precise onset or offset
+of effects (e.g., differences across conditions or groups). Here, we
+introduce a \emph{model-based} approch for analysing M/EEG timeseriesn
+such as ERPs or decoding performance through time, and their differences
+across conditions or groups. This approach relies on Bayesian
+generalised additive multilevel models and outputs the posterior
+probabilility of the effect being above 0 (or above chance) at every
+timestep, while naturally taking into account the temporal dependencies
+and between-subject variability present in such data. Using both
+simulated and actual M/EEG data, we show that the proposed approach
+largely outperforms conventional methods in estimating the onset and
+offset of M/EEG effects, producing more precise and more reliable
+estimates. We provide an R package implementing the approach and
+illustrate how to integrate it into M/EEG statistical pipelines in
+MNE-Python.}
 
@@
 
+\subsection{Problem statement}\label{problem-statement}
+
+Description of cluster-based approaches (see
+\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019})\ldots{}
+As pointed by the original authors themselves
+(\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007}), ``there There is
+a conflict between this interest in localized effects and our choice for
+a global null hypothesis: by controlling the FA {[}false alarm{]} rate
+under this global null hypothesis, one cannot quantify the uncertainty
+in the spatiotemporal localization of the effect''
+(\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
+\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019})\ldots{}
+In contrast, the proposed approach, based on Bayesian generalised
+additive multilevel models, allows quantifying the probability of
+effects at the level of timesteps, voxels, sensors (etc), while
+naturally taking into account spatiotemporal dependencies.
+
 \subsection{Previous work}\label{previous-work}
@@
 
+See also the rERP framework (\citeproc{ref-smith2014a}{N. J. Smith \&
+Kutas, 2014a}, \citeproc{ref-smith2014b}{2014b}) and Tremblay \& Newman
+(\citeproc{ref-tremblay2014}{2014})\ldots{}
+
 From Dimigen \& Ehinger (\citeproc{ref-dimigen2021}{2021}): Recently,
@@
 for EEG data (\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
-\citeproc{ref-meulman2015}{Meulman et al., 2015})\ldots{}
+\citeproc{ref-meulman2015}{Meulman et al., 2015},
+\citeproc{ref-meulman2023}{2023})\ldots{}
 
@@
 
-Using Bayes factors for decoding performance
-(\citeproc{ref-teichmann2022}{Teichmann, 2022})\ldots{}
+Recently, Teichmann (\citeproc{ref-teichmann2022}{2022}) provided a
+detailed tutorial on using Bayes factors (BFs) to analyse the 1D or 2D
+output from MVPA, that is, for testing, at every timestep, whether
+decoding performance is above chance level. However, this approach
+provides timeseries of BFs that ignores temporal dependencies..
 
@@
 \citeproc{ref-winter2016}{Winter \& Wieling, 2016}) or application to
-phonetic data (\citeproc{ref-wieling2018}{Wieling, 2018}) or this
-introduction (\citeproc{ref-baayen2020}{Baayen \& Linke, 2020}) or these
-reference books (\citeproc{ref-hastie2017}{Hastie \& Tibshirani, 2017};
+phonetic data (\citeproc{ref-suxf3skuthy2021}{Sóskuthy, 2021};
+\citeproc{ref-wieling2018}{Wieling, 2018}) or this introduction
+(\citeproc{ref-baayen2020}{Baayen \& Linke, 2020}) or these reference
+books (\citeproc{ref-hastie2017}{Hastie \& Tibshirani, 2017};
 \citeproc{ref-wood2017}{Wood, 2017a})\ldots{} application to
@@
 GAMLSS for neuroimaging data (\citeproc{ref-dinga2021}{Dinga et al.,
-2021})\ldots{}
+2021})\ldots{} Modelling auto-correlation in GAMMs + EEG example
+(\citeproc{ref-baayen2018}{Baayen et al., 2018})\ldots{}
 
@@
 Poisson) with mean \(\mu_{i}\) and scale parameter \(\phi\);
-\(g(\cdot)\) is the link function, \(A\) is an offset,
-\(\mathbf{X}_{i}\) is the \(i\)th row of a parametric model matrix,
-\(\gamma\) is a vector of parameters for the parametric terms, \(f_{j}\)
-is a smooth function of covariate \(x_{j}\). The smooth functions
-\(f_{j}\) are represented in the model via penalised splines basis
-expansions of the covariates, that are a weighted sum of basis
+\(g(\cdot)\) is the link function, \(\mathbf{A}_{i}\) is the \(i\)th row
+of a known parametric model matrix, \(\gamma\) is a vector of parameters
+for the parametric terms (to be estimated), \(f_{j}\) is a smooth
+function of covariate \(x_{j}\) (to be estimated as well). The smooth
+functions \(f_{j}\) are represented in the model via penalised splines
+basis expansions of the covariates, that are a weighted sum of basis
 functions:
@@
 
+Terminology: \emph{splines} are functions composed of simpler functions.
+These simpler functions are basis functions (e.g., a cubic polynomial)
+and the set of basis functions is a \emph{basis}. Each basis function
+gets its coefficient and the resultant spline is the sum of these
+weighted basis functions.
+
 \subsection{Bayesian generalised additive multilevel
@@
 
-Now describe the Bayesian GAMM\ldots{} Proper inclusion of
-varying/random effects in the model specification protects against
-overly wiggly curves (\citeproc{ref-baayen2020}{Baayen \& Linke,
-2020})\ldots{}
+Now describe the Bayesian GAMM (\citeproc{ref-miller2025}{Miller,
+2025})\ldots{} Proper inclusion of varying/random effects in the model
+specification protects against overly wiggly curves
+(\citeproc{ref-baayen2020}{Baayen \& Linke, 2020})\ldots{} Generalising
+to scale and shape or ``distributional GAMs''
+(\citeproc{ref-rigby2005}{Rigby \& Stasinopoulos, 2005};
+\citeproc{ref-umlauf2018}{Umlauf et al., 2018}) and applied to
+neuroimaging data (\citeproc{ref-dinga2021}{Dinga et al., 2021})\ldots{}
+
+Instead of averaring, obtain the smooth ERP signal from multilevel
+GAM\ldots{} less susceptible to outliers (Meulman et al., 2023)\ldots{}
 
@@
 
-Following the approach used by Sassenhagen \& Draschkow
+Following the approach of Sassenhagen \& Draschkow
 (\citeproc{ref-sassenhagen2019}{2019}) and Rousselet
@@
 \citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}). We used the
-default priors in \texttt{brms}, that is, weakly informative priors. We
+default priors in \texttt{brms} (i.e., weakly informative priors). We
 ran eight Markov Chain Monte-Carlo (MCMC) to approximate the posterior
@@
 \texttt{R} package \texttt{mgcv} v 1.9-1 (\citeproc{ref-mgcv}{Wood,
-2017b}) for specifying smooth effects.
+2017b}) for specifying smooth effects. Figure~\ref{fig-post-prob-test}
+shows the predictions of this model together with the raw data.
 
@@
     \CommentTok{\# cubic regression splines with k{-}1 basis functions}
-\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
+\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
     \AttributeTok{data =}\NormalTok{ ppt\_df,}
@@
 \caption{\label{fig-post-prob-ratio}Ratio of posterior probability
-according to the GAM. Timesteps above threshold (10) are highlighted in
-green.}
+according to the GAM (on a log10 scale). Timesteps above threshold (10)
+are highlighted in green.}
 
@@
 ERPs (mean and SD) at the participant level (similar to what is done in
-meta-analysis).
+meta-analysis). Although it is possible to fit a GAMM at the
+single-trial level, we present a resource-efficient version of the model
+that is fitted directly on the by-participant summary statistics (mean
+and SD)\ldots{}
 
@@
 \NormalTok{    eeg\_mean }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
-\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
+\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
 \NormalTok{        (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ participant),}
@@
 cluster-based methods (permutation with a single threshold and TFCE,
-\citeproc{ref-smith2009}{Smith \& Nichols, 2009}). The BH95, BY01, and
-Holm corrections were applied to the p-values using the
+\citeproc{ref-smith2009}{S. Smith \& Nichols, 2009}). The BH95, BY01,
+and Holm corrections were applied to the p-values using the
 \texttt{p.adjust()} function in \texttt{R}. The cluster-based inference
@@
 
-Assessing the reliability of the proposed approach using some sort of
-split-half reliability (e.g., \citeproc{ref-rosenblatt2018}{Rosenblatt
-et al., 2018})? Using the MEG decoding results of Nalborczyk et al.~(in
-preparation) in 33 participants:
+Next, we assessed the performance of the proposed approach on actual MEG
+data (decoding results of Nalborczyk et al., in preparation). We
+conducted time-resolved multivariate pattern analysis (MVPA), also known
+as decoding\ldots{} As a result, we have a timecourse of decoding
+performance (ROC AUC), bounded between 0 and 1, for each participant
+(for a total of 32 participants). Now, we want to \emph{test} whether
+the group-level average decoding accuracy is above chance (i.e., 0.5) at
+each timestep (Figure~\ref{fig-decoding-data}). We fitted a similar GAM
+as discussed previously, but we replaced the \(\mathrm{Normal}\)
+likelihood function by a \(\mathrm{Beta}\) one to account for the
+bounded nature of AUC values (between 0 and 1) (for a tutorial on Beta
+regression, see \citeproc{ref-coretta2025}{Coretta \& Bürkner, 2025}).
+Note that although we chose a basis dimension of \(k=50\), which seems
+appropriate for the present data, this choice should be adapted
+according to the properties of the modelled data (e.g., signal-to-noise
+ratio, prior low-pass filtering, sampling rate, etc) and should be
+assessed by the usual model checking tools (e.g., posterior predictive
+checks). We also define a region of practical equivalence (ROPE,
+\citeproc{ref-kruschke2017}{Kruschke \& Liddell, 2017}), defined as the
+chance level plus the standard deviation of the (group-level average)
+decoding performance during the baseline period. This ensures
+that\ldots{}\footnote{Could an alternative approach be to include a
+  predictor for ``baseline vs.~after baseline''?}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# fitting the Beta GAM}
+\NormalTok{meg\_decoding\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
+\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{50}\NormalTok{),}
+    \AttributeTok{data =}\NormalTok{ decoding\_df,}
+    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
+    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
+    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
+    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{cores =} \DecValTok{4}
+\NormalTok{    )}
+\end{Highlighting}
+\end{Shaded}
+
+We assessed the reliability of the proposed approach using a form of
+permutation-based split-half reliability (e.g.,
+\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}), which consisted
+of the following steps:
 
@@
 
+\begin{figure}[!htb]
+
+\caption{\label{fig-decoding-data}Group-level average decoding
+performance (N=32) superimposed with the GAM predictions (in blue) and
+the region of practical equivalence (ROPE, in orange) computed from the
+baseline period. The blue horizontal line indicates the timesteps at
+which the posterior probability ratio is equal to or greater than 20.}
+
+\centering{
+
+\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-data-1.pdf}
+
+}
+
+\end{figure}%
+
+\newpage
+
 \section{Results}\label{results}
 
-\ldots{}
+This section is divided in two parts. First, we present the results from
+the simulation study, assessing the bias and variance of each method
+when applied to simulated data in which the ground truth is known.
+Second, we present the results obtained when applying the different
+methods to actual MEG data (decoding performance through time),
+assessing the reliability of each method and the stability of its
+estimates.
 
@@
 
-\subsection{Application to actual EEG data
-(reliability)}\label{application-to-actual-eeg-data-reliability}
-
-\ldots{}
+\newpage
+
+\subsection{Application to actual MEG data
+(reliability)}\label{application-to-actual-meg-data-reliability}
+
+Figure~\ref{fig-onset-offset} shows the group-level average decoding
+performance through time with onset and offset estimates for each
+method. The \texttt{brms\_full} method is similar to the \texttt{brms}
+method except that the ROPE is defined on the entire dataset rather than
+on the split dataset. Overall, this figure shows that both the
+\texttt{Raw\ p-value} and \texttt{FDR\ BH95} methods were extremely
+lenient, considering that the decoding performance was above chance
+before the onset of the stimulus (false positive) and until the end of
+the trial. The \texttt{Change\ point} and \texttt{Cluster\ mass} methods
+were the most conservative methods, identifying a time window from
+approximately +60ms to +500ms. The \texttt{Holm}, \texttt{TFCE},
+\texttt{brms}, and \texttt{brms\_full} methods produced somewhat similar
+estimates of onset and offset, from approximately +60ms to
++650ms.\footnote{It should be noted that although each method can
+  produce several ``clusters'' of timesteps, we only considered the
+  first (onset) and last (offset) timesteps identified by each method to
+  compute the error (difference).}
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-onset-offset}Group-level average decoding
+performance through time with onset and offset estimates for each method
+(data from Nalborczyk et al., in preparation).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-offset-1.pdf}
+
+}
+
+\end{figure}%
+
+Figure~\ref{fig-reliability} shows two properties of each method: i) the
+median difference between the onset and offset estimates from each data
+split and the onset and offset estimates from the full dataset (x-axis)
+and ii) and the variance of its onset and offset estimates across data
+splits (y-axis). This figure reveals that the \texttt{brms} \emph{onset
+and offset} estimates on each split are the closest to the estimates
+from the full dataset on average (0ms difference for the onset estimate
+and 5ms difference for the offset estimate). The \texttt{Raw\ p-value}
+method has similar performance, but given the aberrant estimates it
+produces (cf. Figure~\ref{fig-onset-offset}), the result that it is
+consistent between data splits and the full dataset is not convincing on
+its own. Overall, the figure reveals that for all other methods, split
+datasets produce later onset estimates and earlier offset estimates (as
+compared to the estimates from the model fitted on the full dataset).
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-reliability}Median absolute error and variance of
+onset (left) and offset (right) estimates for each method.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-reliability-1.pdf}
+
+}
+
+\end{figure}%
 
@@
 
-\ldots{}
+Overall, before concluding on the onset/offset of effect based on the
+model, we need to ensure that the model provides a faithful description
+of the data-generating process (e.g., via posterior predictive checks
+etc)\ldots{}
 
@@
 
+As in previous simulation work (e.g.,
+\citeproc{ref-rousselet2008}{Rousselet et al., 2008};
+\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}), the
+present simulation results depend on various choices such as the
+specific cluster-forming algorithm and threshold, signal-to-noise ratio,
+negative impact of preprocessing steps (e.g., low-pass filter) on
+temporal resolution\ldots{} note however, that the same caveats apply to
+all methods\ldots{}
+
+The error properties depend on the threshold parameter, a value of 10 or
+20 seems to be a reasonable default, but the optimal threshold parameter
+can be adjusted using split-half reliability assessment\ldots{}
+
 Can be applied to any 1D timeseries (e.g., pupillometry,
@@
 word)\ldots{} but also continuous predictors at the trial level?
-
-The error properties depend on the threshold parameter, a value of 10 or
-20 seems to be a reasonable default, but the optimal threshold parameter
-can be adjusted using split-half reliability assessment\ldots{}
 
@@
 
+\bibitem[\citeproctext]{ref-baayen2018}
+Baayen, R. H., Rij, J. van, Cat, C. de, \& Wood, S. (2018).
+\emph{Autocorrelated errors in experimental data in the language
+sciences: Some solutions offered by generalized additive mixed models}
+(pp. 49--69). Springer International Publishing.
+\url{https://doi.org/10.1007/978-3-319-69830-4_4}
+
 \bibitem[\citeproctext]{ref-benjamini1995}
@@
 
+\bibitem[\citeproctext]{ref-coretta2025}
+Coretta, S., \& Bürkner, P.-. C. (2025). \emph{Bayesian beta regressions
+with brms in r: A tutorial for phoneticians}.
+\url{http://dx.doi.org/10.31219/osf.io/f9rqg_v1}
+
 \bibitem[\citeproctext]{ref-dimigen2021}
@@
 
+\bibitem[\citeproctext]{ref-kruschke2017}
+Kruschke, J. K., \& Liddell, T. M. (2017). The Bayesian New Statistics:
+Hypothesis testing, estimation, meta-analysis, and power analysis from a
+Bayesian perspective. \emph{Psychonomic Bulletin \& Review},
+\emph{25}(1), 178--206. \url{https://doi.org/10.3758/s13423-016-1221-4}
+
 \bibitem[\citeproctext]{ref-kryuchkova2012}
@@
 
+\bibitem[\citeproctext]{ref-meulman2023}
+Meulman, N., Sprenger, S. A., Schmid, M. S., \& Wieling, M. (2023).
+GAM-based individual difference measures for L2 ERP studies.
+\emph{Research Methods in Applied Linguistics}, \emph{2}(3), 100079.
+\url{https://doi.org/10.1016/j.rmal.2023.100079}
+
 \bibitem[\citeproctext]{ref-meulman2015}
@@
 \url{https://doi.org/10.1371/journal.pone.0143328}
+
+\bibitem[\citeproctext]{ref-miller2025}
+Miller, D. L. (2025). Bayesian views of generalized additive modelling.
+\emph{Methods in Ecology and Evolution}.
+\url{https://doi.org/10.1111/2041-210x.14498}
 
@@
 
+\bibitem[\citeproctext]{ref-rigby2005}
+Rigby, R. A., \& Stasinopoulos, D. M. (2005). Generalized Additive
+Models for Location, Scale and Shape. \emph{Journal of the Royal
+Statistical Society Series C: Applied Statistics}, \emph{54}(3),
+507--554. \url{https://doi.org/10.1111/j.1467-9876.2005.00510.x}
+
 \bibitem[\citeproctext]{ref-vanrij2019}
@@
 
+\bibitem[\citeproctext]{ref-smith2014a}
+Smith, N. J., \& Kutas, M. (2014a). Regression{-}based estimation of ERP
+waveforms: I. The rERP framework. \emph{Psychophysiology}, \emph{52}(2),
+157--168. \url{https://doi.org/10.1111/psyp.12317}
+
+\bibitem[\citeproctext]{ref-smith2014b}
+Smith, N. J., \& Kutas, M. (2014b). Regression{-}based estimation of ERP
+waveforms: II. Nonlinear effects, overlap correction, and practical
+considerations. \emph{Psychophysiology}, \emph{52}(2), 169--181.
+\url{https://doi.org/10.1111/psyp.12320}
+
 \bibitem[\citeproctext]{ref-smith2009}
@@
 
+\bibitem[\citeproctext]{ref-suxf3skuthy2021}
+Sóskuthy, M. (2021). Evaluating generalised additive mixed modelling
+strategies for dynamic speech analysis. \emph{Journal of Phonetics},
+\emph{84}, 101017. \url{https://doi.org/10.1016/j.wocn.2020.101017}
+
 \bibitem[\citeproctext]{ref-teichmann2022}
@@
 \url{https://doi.org/10.52294/apertureneuro.2022.2.maoc6465}
+
+\bibitem[\citeproctext]{ref-tremblay2014}
+Tremblay, A., \& Newman, A. J. (2014). Modeling nonlinear relationships
+in ERP data using mixed{-}effects regression with R examples.
+\emph{Psychophysiology}, \emph{52}(1), 124--139.
+\url{https://doi.org/10.1111/psyp.12299}
+
+\bibitem[\citeproctext]{ref-umlauf2018}
+Umlauf, N., Klein, N., \& Zeileis, A. (2018). BAMLSS: Bayesian Additive
+Models for Location, Scale, and Shape (and Beyond). \emph{Journal of
+Computational and Graphical Statistics}, \emph{27}(3), 612--627.
+\url{https://doi.org/10.1080/10618600.2017.1407325}
 
@@
 
-\section{Application to time-resolved decoding results (accuracy over
-time)}\label{application-to-time-resolved-decoding-results-accuracy-over-time}
-
-We conducted time-resolved multivariate pattern analysis (MVPA), also
-known as decoding. As a result, we have a timecourse of decoding
-accuracies (e.g., ROC AUC), bounded between 0 and 1, per participant
-(Figure~\ref{fig-mne-decoding})\ldots{}
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-mne-decoding}Exemplary average timecourse of binary
-decoding accuracy (ROC AUC) for each participant. Group-level average
-decoding accuracy is depicted as a grey background density in each
-panel.}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-mne-decoding-1.pdf}
-
-}
-
-\end{figure}%
-
-Now, we want to \emph{test} whether the group-level average decoding
-accuracy is above chance (i.e., 0.5) at each timestep. We use a similar
-GAM/GP as previously, but we replace the \(\mathrm{Normal}\) likelihood
-function by a \(\mathrm{Beta}\) one to account for the bounded nature of
-AUC values (between 0 and 1).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# fitting the GAM}
-\NormalTok{decoding\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
-    \AttributeTok{data =}\NormalTok{ decoding\_data,}
-    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/decoding\_gam.rds"}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# fitting the GP}
-\NormalTok{decoding\_gp }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{gp}\NormalTok{(time, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
-    \AttributeTok{data =}\NormalTok{ decoding\_data,}
-    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.99}\NormalTok{),}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/decoding\_gp.rds"}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[H]
-
-\caption{Posterior predictions of the GAM (left) and GP (right) fitted
-on decoding accuracy over time.}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/decoding-preds-1.pdf}
-
-}
-
-\end{figure}%
-
-Next, we plot the posterior probability of decoding accuracy being above
-chance level (plus some epsilon) (Figure~\ref{fig-decoding-post}).
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-decoding-post}Posterior probability of decoding
-accuracy being above chance level according to the GAM (left) or the GP
-(right).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-post-1.pdf}
-
-}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-decoding-ratio}Ratio of posterior probabilities of
-decoding accuracy being above chance level according to the GAM (left)
-or the GP (right).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-ratio-1.pdf}
-
-}
-
-\end{figure}%
-
-\newpage
-
 \section{Application to 2D time-resolved decoding results
@@
     \CommentTok{\# 2D thin{-}plate spline (tp)}
-    \CommentTok{\# auc \textasciitilde{} s(train\_time, test\_time, bs = "tp", k = 10),}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{10}\NormalTok{),}
+\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
     \AttributeTok{data =}\NormalTok{ timegen\_data,}
@@
 
+Could be extended to spatial and temporal dimensions with formulas such
+as \texttt{te(x,\ y,\ Time,\ d\ =\ c(2,\ 1)\ )}\ldots{}
+
 \newpage
@@
 
-\section{Threshold-free cluster
-enhancement}\label{threshold-free-cluster-enhancement}
-
-Cluster-based permutation approaches require defining a cluster-forming
-threshold (e.g., a t- or f-value) as the initial step of the algorithm.
-As different cluster-forming thresholds lead to clusters with different
-spatial or temporal extent, this threshold modulates the sensitivity of
-the subsequent permutation test. The threshold-free cluster enhancement
-method (TFCE) was introduced by Smith \& Nichols
-(\citeproc{ref-smith2009}{2009}) to overcome this arbitrary threshold.
-
-In brief, the TFCE method works as follows. Instead of picking an
-arbitrary cluster-forming threshold (e.g., \(t=2\)), we try all (or
-many) possible thresholds in a given range and check whether a given
-timestep/voxel belongs to a significant cluster under any of the set of
-thresholds\ldots{} Then, instead of using cluster mass, we use a
-weighted average between the cluster extend (\(e\), how broad is the
-cluster, that is, how many connected samples it contains) and the
-cluster height (\(h\), how high is the cluster, that is, how large is
-the test statistic) according to the formula:
-
-\[
-\text{TFCE} = \int_{h} e(h)^{E} h^{H} \mathrm{d}h
-\]
-
-Where\ldots{} the parameters \(E\) and \(H\) are set a priori and
-control the influence of the extend and height on the TFCE. Then,
-p-value for timestep/voxel \(i\) is computed by comparing it TFCE with
-the null distribution of TFCE values. For each permuted signal, we keep
-the maximal value over the whole signal for the null distribution of the
-TFCE\ldots. But see Sassenhagen \& Draschkow
-(\citeproc{ref-sassenhagen2019}{2019})\ldots{}
-
-\newpage
-
 \section{\texorpdfstring{Integration with
@@
 \begin{Highlighting}[]
-\CommentTok{\# suggest a package name}
-\FunctionTok{library}\NormalTok{(available)}
-\FunctionTok{suggest}\NormalTok{(}
-    \AttributeTok{text =} \StringTok{"Time{-}resolved electrophysiological measurements such as those offered by magneto{-} or electro{-}encephalography (M/EEG) provide a unique window onto neural activity underlying cognitive processes.
-Typically, researchers are interested in testing whether and when such measures differ across conditions and/or groups.
-The conventional approach consists in conducting mass{-}univariate statistics through time followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster{-}based inference.
-However, these cluster{-}based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the onset or offset of effects (e.g., differences across conditions or groups).
-Here, we introduce a *model{-}based* approch for analysing M/EEG timeseries such as ERPs or timecourses of decoding performance and their differences across conditions or groups.
-This approach relies on Bayesian generalised additive multilevel models, which output the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between{-}subject variability present in such data.
-Using both simulation and actual EEG data, we show that the proposed approach largely outperforms conventional methods in determining both the onset and offset of M/EEG effects (e.g., ERPs difference, decoding performance), producting more precise and more reliable estimates.
-We provide an R package implementing the approach and illustrate how to integrate it into M/EEG statistical pipelines in MNE{-}Python."}
-\NormalTok{    )}
-
-\NormalTok{available}\SpecialCharTok{:::}\FunctionTok{namr}\NormalTok{(}
-    \AttributeTok{title =} \StringTok{"Precise temporal localisation of M/EEG effects with Bayesian generalised additive multilevel models"}\NormalTok{,}
-    \AttributeTok{verb =} \ConstantTok{TRUE}
-\NormalTok{    )}
-
-\CommentTok{\# to{-}do adding some code here...}
+\CommentTok{\# TO{-}DO: adding some code here...}
 \end{Highlighting}

### 2025-02-20 code/errors_simulation.R

@@ -164,7 +164,7 @@ freq_stats_cluster_matrix <- function (
         # using the first slice of the 3D array
         p_values_ <- np$transpose(np$ones_like(X_np[1, , drop = FALSE]) )
         
-        # defining the (main) cluster onsef and offset
+        # defining the (main) cluster onset and offset
         main_cluster_onset <- clusters[which(p_values < alpha_level)][[1]][[1]]$start+1
         main_cluster_offset <- clusters[which(p_values < alpha_level)][[1]][[1]]$stop
         

### 2025-02-20 code/split_half_reliability.R

@@ -1,110 +1,91 @@
 ###########################################################
-# Monte-Carlo simulation of onset/offset error properties #
+# Split-half reliability of onset/offset estimates        #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on February 13, 2025                       #
+# Last updated on February 20, 2025                       #
 ###########################################################
 
 # importing R packages
 library(changepoint)
 library(reticulate)
+library(easystats)
 library(tidyverse)
 library(tidybayes)
+library(scales)
 library(brms)
+
+# defining the default ggplot2 theme
+theme_set(theme_light(base_size = 12, base_family = "Open Sans") )
 
 # importing Python modules
 np <- import("numpy")
 mne <- import("mne")
 
-# importing R version of Matlab code from Yeung et al. (2004)
-source("code/eeg_noise.R")
+# reshaping the decoding results
+# read.csv(file = "code/decoding_results.csv") %>%
+#     filter(crossval_strategy == "4") %>%
+#     select(participant, cv_fold = fold, time, auc = score) %>%
+#     mutate(cv_fold = cv_fold + 1) %>%
+#     mutate(participant = str_trunc(string = participant, width = 12, ellipsis = "") ) %>%
+#     mutate(participant = str_replace(string = participant, pattern = "tickertape", replacement = "participant") ) %>%
+#     data.frame() %>%
+#     filter(time >= -0.2) %>%
+#     write.csv(file = "code/decoding_results_reshaped.csv", row.names = FALSE)
 
-# importing the ERP template with true onset = 160 ms, F=81, and max at F=126
-source("code/erp_template.R")
+# importing the reshaped MEG decoding results
+raw_df <- read.csv(file = "code/decoding_results_reshaped.csv") %>%
+    # removing the last participant
+    mutate(participant_id = cur_group_id(), .by = participant) %>%
+    filter(participant_id < 33) %>%
+    select(-participant_id)
 
-# importing helper functions
-source("code/functions.R")
+# averaging across participants
+summary_df <- raw_df %>%
+    summarise(
+        auc_mean = mean(auc),
+        auc_sd = sd(auc),
+        auc_se = sd(auc) / sqrt(n() ),
+        .by = c(participant, time)
+        )
 
[…]

### 2025-04-02 brms_meeg.qmd
minor updates to cache

@@
     # financial-support: "This study was supported by Grant [Grant Number] from [Funding Source]."
-abstract: "Time-resolved electrophysiological measurements such as those offered by magneto- or electro-encephalography (M/EEG) provide a unique window onto neural activity underlying cognitive processes.
-Typically, researchers are interested in testing whether and when such measures differ across conditions and/or groups.
-The conventional approach consists in conducting mass-univariate statistics through time followed by some form of multiplicity correction (e.g., FDR, FWER) or cluster-based inference.
-However, while allowing efficient error-rates control, cluster-based methods have an important downside: they shift the focus of inference from the timepoint to the cluster level, thus preventing any conclusion to be made about the precise onset or offset of effects (e.g., differences across conditions or groups).
-Here, we introduce a *model-based* approch for analysing M/EEG timeseriesn such as ERPs or decoding performance through time, and their differences across conditions or groups.
-This approach relies on Bayesian generalised additive multilevel models and outputs the posterior probabilility of the effect being above 0 (or above chance) at every timestep, while naturally taking into account the temporal dependencies and between-subject variability present in such data.
-Using both simulated and actual M/EEG data, we show that the proposed approach largely outperforms conventional methods in estimating the onset and offset of M/EEG effects, producing more precise and more reliable estimates.
-We provide an R package implementing the approach and illustrate how to integrate it into M/EEG statistical pipelines in MNE-Python."
+abstract: "Time-resolved electrophysiological measurements such as those obtained through magneto- or electro-encephalography (M/EEG) offer a unique window into the neural activity underlying cognitive processes.
+Researchers are often interested in determining whether and when these signals differ across experimental conditions or participant groups.
+The conventional approach involves mass-univariate statistical testing across time, followed by corrections for multiple comparisons (e.g., FDR, FWER) or cluster-based inference.
+While effective for controlling error rates, cluster-based methods come with a significant limitation: they shift the focus of inference from individual time points to clusters, making it difficult to draw precise conclusions about the onset or offset of observed effects.
+Here, we introduce a *model-based* approch for analysing M/EEG timeseries such as event-related potentials (ERPs) or decoding performance over time.
+Our method leverages Bayesian generalised additive multilevel models (GAMMs), providing posterior probabilities that an effect is above zero (or above chance) at each time point, while naturally accounting for temporal dependencies and between-subject variability.
+Using both simulated and actual M/EEG datasets, we demonstrate that this approach substantially outperforms conventional methods in estimating the onset and offset of neural effects, yielding more precise and reliable results.
+We also provide an R package implementing the method and show how it can be seamlessly integrated into M/EEG analysis pipelines using MNE-Python."
 keywords: [EEG, MEG, generalised additive models, mixed-effects models, multilevel models, Bayesian statistics, brms]
@@
 library(MetBrewer)
+library(tidytext)
 library(grateful)
@@
 # setting a default ggplot2 theme
-theme_set(theme_light(base_size = 10, base_family = "Open Sans") )
+# theme_set(theme_light(base_size = 10, base_family = "Open Sans") )
+theme_set(theme_bw(base_size = 10, base_family = "Open Sans") )
 
@@
 
+Understanding the temporal dynamics of cognitive processes requires methods that can capture fast-changing neural activity with high temporal resolution.
+Magnetoencephalography (MEG) and electroencephalography (EEG) are two such methods, widely used in cognitive neuroscience for their ability to track brain activity at the millisecond scale.
+These techniques provide rich time series data that reflect how neural responses unfold in response to stimuli or tasks.
+A central goal in many M/EEG studies is to determine whether—and when—neural responses differ across experimental conditions or participant groups.
+
 Here are some useful references to be discussed [@combrisson_exceeding_2015; @ehinger_unfold_2019; @hayasaka_validating_2003; @luck_how_2017; @pedersen_hierarchical_2019; @frossard2021; @frossard2022; @gramfort2013; @maris2007; @pernet2015]...
@@
 Description of cluster-based approaches [see @sassenhagen2019]...
-As pointed by the original authors themselves [@maris2007], "there There is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA \[false alarm\] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect" [@maris2007; @sassenhagen2019]...
-In contrast, the proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the probability of effects at the level of timesteps, voxels, sensors (etc), while naturally taking into account spatiotemporal dependencies.
+
+However, as pointed by the original authors themselves [@maris2007], "there is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA \[false alarm\] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect" [@maris2007; @sassenhagen2019].
+In contrast, the proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the posterior probability of effects being above chance at the level of timesteps, voxels, sensors (etc), while naturally taking into account spatiotemporal dependencies present in M/EEG timeseries.
 
@@
 where $\beta_{jk}$ is the weight (coefficient) associated with the $k$th basis function $b_{jk}()$ evaluated at the covariate value $x_{ij}$ for the $j$th smooth function $f_{j}$.
-Splines coefficients are penalised (usually through the squared of the smooth functions' second derivative) in a way that can be interpreted as a prior on the "wiggliness" of the function...
-
-Terminology: *splines* are functions composed of simpler functions.
+Splines coefficients are penalised (usually through the squared of the smooth functions' second derivative) in a way that can be interpreted, in Bayesian terms, as a prior on the "wiggliness" of the function.
+In other words, more complex (wiggly) basis function are penalised.
+To clarify the terminology at this stage: *splines* are functions composed of simpler functions.
 These simpler functions are basis functions (e.g., a cubic polynomial) and the set of basis functions is a *basis*.
@@
 
-Instead of averaring, obtain the smooth ERP signal from multilevel GAM... less susceptible to outliers (Meulman et al., 2023)...
-
-```{=html}
-<!--
-
-## Gaussian process regression
-
-A Gaussian process (GP) is a stochastic process that defines the distribution over a collection of random variables indexed by a continuous variable, that is $\{f(t): t \in \mathcal{T}\}$ for some index set $\mathcal{T}$ [@riutort-mayol_practical_2023; @rasmussen2005].
-Whereas Bayesian linear regression outputs a distribution over the parameters of some predefind parametric model, the GP approach, in contrast, is a non-parametric approach, in that it finds a distribution over the possible functions that are consistent with the observed data.
-However, note that nonparametric does not mean there aren't parameters, it means that there are infinitely many parameters.
-
-From [brms documentation](https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp): A GP is a stochastic process, which describes the relation between one or more predictors $x=\left(x_{1}, \ldots, x_{d}\right)$ and a response $f(x)$, where $d$ is the number of predictors.
-A GP is the generalization of the multivariate normal distribution to an infinite number of dimensions.
-Thus, it can be interpreted as a prior over functions.
-The values of $f()$ at any finite set of locations are jointly multivariate normal, with a covariance matrix defined by the covariance kernel $k_p\left(x_i, x_j\right)$, where $p$ is the vector of parameters of the GP:
-
-$$
-\left(f\left(x_{1}\right), \ldots f\left(x_{n}\right) \sim \operatorname{MVN}\left(0, \left(k_p\left(x_{i}, x_{j}\right)\right)_{i, j=1}^{n}\right)\right.
-$$
-
-The smoothness and general behaviour of the function $f$ depends only on the choice of covariance kernel, which ensures that values that are close together in the input space will be mapped to similar output values...
-
-From this perspective, $f$ is a realisation of an infinite dimensional normal distribution:
-
-$$
-f \sim \mathrm{Normal}(0, C(\lambda))
-$$
-
-where $C$ is a covariance kernel with hyperparameters $\lambda$ that defines the covariance between two function values $f\left(t_1\right)$ and $f\left(t_2\right)$ for two time points $t_1$ and $t_2$ [@rasmussen2005].
-Similar to the different choices of the basis function for splines, different choices of the covariance kernel lead to different GPs.
-In this article, we consider the squared-exponential (a.k.a.
-radial basis function) kernel, which computes the squared distance between points and converts it into a measure of similarity.
-It is defined as:
-
-$$
-C(\lambda) := C\left(t_1, t_2, \sigma, \gamma\right) := \sigma^2 \exp \left(-\frac{||t_1-t_2||^{2}}{2 \gamma^2}\right)
-$$
-
-with hyperparameters $\lambda = (\sigma, \gamma)$, expressing the overall scale of GP and the length-scale, respectively [@rasmussen2005].
-The advantages of this kernel are that it is computationally efficient and (infinitely) smooth making it a reasonable choice for the purposes of the present article.
-Here again, $\lambda$ hyperparameters are estimated from the data, along with all other model parameters.
-
-Taken from <https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10c-Gaussian-processes.html>: For a given vector $\mathbf{x}$, we can use the kernel to construct finite multi-variate normal distribution associated with it like so:
-
-$$
-\mathbf{x} \mapsto_{G P} \operatorname{MVNormal}(m(\mathbf{x}), k(\mathbf{x}, \mathbf{x}))
-$$
-
-where $m$ is a function that specifies the mean for the distribution associated with $\mathbf{x}$.
-This mapping is essentially the Gaussian process: a systematic association of vectors of arbitrary length with a suitable multi-variate normal distribution.
-
-See also <https://jtimonen.github.io/lgpr-usage/articles/math.html>...
-
-Low-rank approximate Gaussian processes are of main interest in machine learning and statistics due to the high computational demands of exact Gaussian process models [@riutort-mayol_practical_2023]...
-
--->
-```
+Instead of averaging, obtain the smooth ERP signal from multilevel GAM... less susceptible to outliers [@meulman2023]...
 
@@
 The `brms` package uses the same syntax as `r pkrt("mgcv")` for specifying smooth effects.
-@fig-post-prob-test shows the predictions of this model together with the raw data.
+@fig-plot-post-slope shows the predictions of this model together with the raw data.
 
@@
 
+# timing the model fitting
+# around 600s for 2000 iterations on 4 chains and 4 cores
+# around 190s for 2000 iterations on 4 chains and 4 cores + cmdstanr and 01
+# around 440s for 2000 iterations on 4 chains and 4 cores + threading(2), cmdstanr and 01
+# library(tictoc)
+# tic()
+
 # fitting the GAM
@@
     # cubic regression splines with k-1 basis functions
-    eeg ~ condition + s(time, bs = "cr", k = 20, by = condition),
+    # eeg ~ condition + s(time, bs = "cr", k = 20, by = condition),
+    # thin-plate regression splines with k-1 basis functions
+    eeg ~ condition + s(time, bs = "tp", k = 20, by = condition),
     data = ppt_df,
@@
     file = "models/gam.rds"
-    )
+    #################################
+    # speeding up the model fitting #
+    #################################
+    # backend = "cmdstanr",
+    # threads = threading(2),
+    # stan_model_args = list(stanc_options = list("O1") )
+    )
+
+# stop timing
+# toc()
 ```
@@
     file = "models/gamm.rds"
-    )
-```
-
-```{r gp, echo = FALSE, eval = FALSE, message = FALSE}
-# Now we fit the GP regression model...
-# computing eeg_diff and averaging across participants
-# group_df <- raw_df %>%
-#     group_by(participant, time) %>%
-#     pivot_wider(names_from = condition, values_from = eeg, values_fn = mean) %>%
-#     mutate(eeg_diff = cond2 - cond1) %>%
-#     summarise(eeg_diff = mean(eeg_diff) ) %>%
-#     ungroup()
-
-# see https://betanalpha.github.io/assets/case_studies/gp_part3/part3.html
-# and https://betanalpha.github.io/assets/case_studies/gaussian_processes.html
-gp_priors <- c(
-    set_prior("normal(0, 1)", class = "Intercept"),
-    set_prior("normal(0, 1)", class = "b"),
-    set_prior("exponential(1)", class = "sigma"),
-    set_prior("exponential(1)", class = "sdgp")
-    )
-
-gp_model <- brm(
-    # k refers to the number of basis functions for
-    # computing Hilbert-space approximate GPs
-    # if k = NA (default), exact GPs are computed
-    # eeg ~ gp(time, by = condition),
-    eeg ~ condition + gp(time, by = condition, k = 20, cov = "exp_quad"),
-    data = ppt_df,
-    family = gaussian(),
-    # prior = gp_priors,
-    control = list(adapt_delta = 0.99, max_treedepth = 15),
-    backend = "cmdstanr",
-    iter = 2000,
-    chains = 4,
-    cores = 4,
-    file = "models/gp.rds"
     )
@@
 
-We then compute the posterior probability of the slope for `condition` being above $0+\epsilon$ (@fig-post-prob-test), with $\epsilon := 0.05$, which can be interpreted as the smallest effect size of interest.
+We then compute the posterior probability of the slope for `condition` being above $0+\epsilon$ (@fig-post-prob-ratio, left), with $\epsilon := 0.05$, which can be interpreted as the smallest effect size of interest.
+We can also express this as the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio, right).
 
@@
 # using the function with the two models
-plot_post_test(model = gam, eps = 0.1)
-```
-
-We can also express this as the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio).
-
-```{=html}
-<!--
-
-Maybe better to use the p-MAP (i.e., density ratio between the null and the MAP)?
-And/or look at the derivative (e.g., <https://fromthebottomoftheheap.net/2014/05/15/identifying-periods-of-change-with-gams/>)?
-
--->
-```
-
-```{r fig-post-prob-ratio, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Ratio of posterior probability according to the GAM (on a log10 scale).
+# plot_post_test(model = gam, eps = 0.1)
+```
+
+```{r tests, echo = FALSE, eval = FALSE}
+# compute percentage in ROPE
+library(easystats)
+# rope(
+#     x = gam,
+#     range = "default",
+#     ci = 0.95,
+#     ci_method = "ETI",
+#     effects = "fixed",
+#     # component = c("conditional", "zi", "zero_inflated", "all"),
+#     # parameters = NULL,
+#     verbose = TRUE
+#     )
+# rope_range(x = gam)
+
+# retrieving posterior draws for EEG over time
+gam <- full_model
+post_samples <- posterior_epred(object = gam, newdata = ppt_df, allow_new_levels = TRUE)
+
+# defining the ROPE as baseline SD
+# rope_range <- c(-0.1, 0.1)
+baseline_sd <- ppt_df %>%
+    pivot_wider(names_from = condition, values_from = eeg) %>%
+    mutate(cond_diff = abs(cond1 - cond2) ) %>%
+    filter(time < 0.1) %>%
+    # head()
+    summarise(baseline_sd = sd(cond_diff) ) %>%
+    data.frame()
+
+rope_range <- c(-baseline_sd, +baseline_sd)
+
+# computing the HDI for each time step
+hdi_values <- apply(post_samples, 2, function (x) hdi(x, ci = 0.95) )
+
+# computing the  proportion of the HDI that falls within the ROPE
+prop_in_rope <- sapply(1:ncol(post_samples), function (i) {
+    lower_hdi <- hdi_values[[i]][1]
+    upper_hdi <- hdi_values[[i]][2]
+    # computing the  proportion inside ROPE
+    mean(lower_hdi >= rope_range[1] & upper_hdi <= rope_range[2])
+    })
+
+# storing the results in a dataframe
+rope_results <- data.frame(time = ppt_df$time, prop_in_rope = prop_in_rope)
+
+# plotting the results
+rope_results %>%
+    summarise(prop_in_rope = mean(prop_in_rope), .by = time) %>%
+    ggplot(aes(x = time, y = prop_in_rope) ) +
+    geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = true_offset, linetype = 2) +
+    geom_point()
+
+# loading or fitting the full model
+# full_model <- gam
+full_model <- readRDS("models/gam.rds")
+
+# fitting the null model (no 'condition' effect)
+null_model <- brm(
+    eeg ~ s(time, bs = "tp", k = 20),
+    data = ppt_df,
+    family = gaussian(),
+    warmup = 2000,
+    iter = 5000,
+    chains = 8,
+    cores = 8,
+    file = "models/null_gam.rds"
+    )
+
+# computing the marginal likelihoods using bridge sampling
+# library(bridgesampling)
+# full_bridge <- bridge_sampler(full_model)
+# null_bridge <- bridge_sampler(null_model)
+
+# computing the Bayes Factor in favor of full model
+# BF_10 <- bf(full_bridge, null_bridge)
+# print(BF_10)
+
+# extracting log-likelihoods for each model
+log_lik_full <- log_lik(full_model)
+log_lik_null <- log_lik(null_model)
+
+# computing the pointwise log pointwise predictive density (LPD)
+# averaging the log-likelihood across posterior draws
+lpd_full <- colMeans(log_lik_full)
+lpd_null <- colMeans(log_lik_null)
+
+# computing the LPD difference (full model - null model)
+lpd_diff <- lpd_full - lpd_null
+
+# computing pointwise Bayes Factors as exp(lpd_diff)
+BF_10_pointwise <- exp(lpd_diff)
+
+# storing the results in a dataframe
+ppt_df <- full_model$data
+lpd_results <- data.frame(time = ppt_df$time, LPD_diff = lpd_diff, BF_10 = BF_10_pointwise)
+
+# viewing the results
+head(lpd_results)
+
+# plotting it
+lpd_results %>%
+    # pivot_longer(cols = LPD_diff:BF_10) %>%
+    summarise(BF_10 = mean(LPD_diff), .by = time) %>%
+    # ggplot(aes(x = time, y = value, colour = name) ) +
+    ggplot(aes(x = time, y = BF_10) ) +
+    geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = true_offset, linetype = 2) +
+    geom_line() +
+    geom_point()
+```
+
+```{r fig-post-prob-ratio, echo = FALSE, out.width = "100%", fig.width = 9, fig.asp = 0.5, fig.cap = "Left: Posterior probability of the ERP difference (slope) being above 0 according to the GAM.
+Right: Ratio of posterior probability according to the GAM (on a log10 scale).
 Timesteps above threshold (10) are highlighted in green."}
@@
 # computing the posterior probability that y is above 0
-n_posterior_samples <- 24000
+n_posterior_samples <- ndraws(gam)
 prob_y_above_0 <- time_seq %>%
@@
     ggplot(aes(x = time, y = prob_ratio) ) +
-    geom_raster(
-        data = bg_data,
-        aes(x = time, y = prob_ratio, fill = fill_value),
-        alpha = 0.5,
-        interpolate = FALSE,
-        inherit.aes = FALSE,
-        show.legend = FALSE
-        ) +
-    scale_fill_scico(palette = "vik", midpoint = log10(1) ) +
+    # geom_raster(
+    #     data = bg_data,
+    #     aes(x = time, y = prob_ratio, fill = fill_value),
+    #     alpha = 0.5,
+    #     interpolate = FALSE,
+    #     inherit.aes = FALSE,
+    #     show.legend = FALSE
+    #     ) +
+    # scale_fill_scico(palette = "vik", midpoint = log10(1) ) +
     geom_hline(yintercept = 1, linetype = "dashed") +
@@
 # or plotting only the GAM
-gam_ratio_plot
+plot_post_test(model = gam, eps = 0.1) + gam_ratio_plot
 
@@
 #     relocate(model, .before = cluster_onset)
-
 ```
@@
 
-Next we fit a hierarchical/multilevel GAM using summary statistics of ERPs (mean and SD) at the participant level (similar to what is done in meta-analysis).
-Although it is possible to fit a GAMM at the single-trial level, we present a resource-efficient version of the model that is fitted directly on the by-participant summary statistics (mean and SD)...
+Next we fit a multilevel version of the GAM.
+Although it is possible to fit a GAMM at the single-trial level, we present a resource-efficient version of the model that is fitted directly on the by-participant summary statistics (mean and SD, similar to what is done in meta-analysis...
 
@@
     file = "models/meta_gam.rds"
-    )
-```
-
-```{r meta-gp, echo = FALSE, eval = FALSE, message = FALSE}
-# fitting the GP
-meta_gp <- brm(
-    # using by-participant SD of ERPs across trials
-    eeg_mean | se(eeg_sd) ~
-        condition + gp(time, k = 20, by = condition) +
-        (1 | participant),
-    data = summary_df,
-    family = gaussian(),
-    control = list(adapt_delta = 0.99),
-    iter = 5000,
-    chains = 4,
-    cores = 4,
-    file = "models/meta_gp.rds"
     )
@@
 
-We then computed the difference between the true and estimated onset/offset of the ERP difference ($\text{error}:=|\hat{\theta}-\theta|$), according to various `eps` and `threshold` values.
+We then computed the difference between the true and estimated onset/offset of the ERP difference ($\text{error}:=|\hat{\theta}-\theta|$), according to various `k` (GAM basis dimension) and `threshold` values.
 Remember that the signal is generated from a truncated Gaussian with an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
-@fig-onset-error shows that the hierarchical GAM can *exactly* recover the true onset and offset values, given some reasonable choice of `eps` and `threshold` values (e.g., a threshold of 20).
-
-```{r fig-onset-error, echo = FALSE, dpi = 300, fig.width = 8, fig.asp = 0.5, fig.cap = "Error function of onset (left) and offset (right) estimation according to various eps and threshold values (according to the hierarchical GAM).
-Minimum error values are indicated by red crosses."}
-# defining values of eps and threshold to test
-eps_values <- seq(from = 0, to = 0.1, by = 0.01)
-threshold_values <- seq(from = 1, to = 30, by = 1)
-
-# using as many values of time as in the original data
-# n_approx <- n_distinct(ppt_df$time)
+@fig-onset-error shows that the hierarchical GAM can *exactly* recover the true onset and offset values, given some reasonable choice of `k` and `threshold` values (e.g., a threshold of 20).
+We provide more detailed recommendations on how to set `k` in @sec-basis.
+
+```{r fig-onset-error, eval = TRUE, echo = FALSE, results = "hide", dpi = 300, fig.width = 9, fig.asp = 0.5, fig.cap = "Error function of onset (left) and offset (right) estimation according to various basis dimension and threshold values (according to the multilevel GAM)."}
+# defining values of k and threshold to test
+# threshold_values <- seq(from = 1, to = 50, by = 1)
+# kvalues <- c(5, 10, 15, 20, 25, 30, 35, 40)
+
+# number of simulations to run per combination
+# nsims <- 10
 
 # number of posterior samples to use
-n_post_samples <- 1e3
+# n_post_samples <- 1e3
+
+# importing the data generation functions 
+# source(file = "code/generate_eeg_data.R")
 
 # initialising results dataframe
-results <- crossing(eps = eps_values, threshold = threshold_values) %>%
-    mutate(
-        estimated_onset = NA, estimated_peak = NA, estimated_offset = NA,
-        error_onset = NA, error_peak = NA, error_offset = NA
-        )
-
-# which model should we use?
-error_model <- meta_gam
-
-# generating a grid of time values to compute predictions for
-# time_seq <- crossing(
-#     condition = c("cond1", "cond2"),
-#     time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx*2-1),
-#     ) %>%
-#     arrange(condition)
-
-# eps <- 0.05
-# preds <- meta_gam$data %>%
-#     add_epred_draws(meta_gam, ndraws = 1e3) %>%
+# results <- crossing(kvalue = kvalues, threshold = threshold_values) %>%
+# results <- crossing(kvalue = kvalues, sim_id = 1:nsims, threshold = threshold_values) %>%
+#     mutate(
+#         estimated_onset = NA,
+#         estimated_offset = NA,
+#         error_onset = NA,
+#         error_offset = NA
+#         )
+
+# looping over different values of kvalue and threshold
+# for (i in seq_len(nrow(results) ) ) {
+# 
+#     # retrieving current kvalue and threshold
+#     kvalue <- results$kvalue[i]
+#     threshold <- results$threshold[i]
+# 
+#     # printing progress
+#     cat(
+#         "\nAssessing combination:", i, "out of", nrow(results), "combinations...\n",
+#         "k-value:", kvalue, "\n", "threshold:", threshold, "\n"
+#         )
+# 
+#     # when we assess a new k-value
+#     if (threshold == 1) {
+#         
+#         # generate some data
+#         raw_df <- generate_data(
+#             n_trials = n_trials, n_ppt = n_ppt, outvar = outvar,
+#             srate = srate, ronset = ronset
+#             )
+#         
+#         # sanity check
+#         head(raw_df)
+#         
+#         # averaging across participants
+#         summary_df <- raw_df %>%
+#             summarise(
+#                 eeg_mean = mean(eeg),
+#                 eeg_sd = sd(eeg),
+#                 .by = c(participant, condition, time)
+#                 )
+#         
+#         # defining a contrast for condition
+#         contrasts(summary_df$condition) <- c(-0.5, 0.5)
+# 
+#         # fitting the GAM
+#         error_model <- brm(
+#             # using by-participant SD of EEG across trials
+#             eeg_mean | se(eeg_sd) ~
+#                 condition + s(time, bs = "cr", k = kvalue, by = condition) +
+#                 (1 | participant),
+#             data = summary_df,
+#             family = gaussian(),
+#             warmup = 2000,
+#             iter = 5000,
+#             chains = 8,
+#             cores = 8
+#             )
+# 
+#     }
+# 
+#     prob_y_above_0 <- error_model$data %>%
+#         add_epred_draws(object = error_model, ndraws = n_post_samples) %>%
+#         # add_epred_draws(object = meta_gam) %>%
 #         data.frame() %>%
 #         dplyr::select(participant, time, condition, .epred, .draw) %>%
-#     group_by(participant, time) %>%
+#         # dplyr::select(time, condition, .epred, .draw) %>%
 #         pivot_wider(names_from = condition, values_from = .epred) %>%
 #         mutate(epred_diff = cond2 - cond1) %>%
-#     summarise(m = mean(epred_diff > eps) ) %>%
+#         # computing mean posterior probability at the participant level
+#         # group_by(participant, time) %>%
+#         # summarise(m = mean(epred_diff > 0) ) %>%
+#         # ungroup() %>%
+#         # computing mean posterior probability at the group level
+#         group_by(time) %>%
+#         # summarise(m = mean(m) ) %>%
+#         summarise(m = mean(epred_diff > 0) ) %>%
 #         mutate(prob_ratio = m / (1 - m) ) %>%
 #         ungroup()
-
-# plotting these predictions
-# preds %>%
-#     ggplot(aes(x = time, y = m) ) +
-#     geom_line() +
-#     facet_wrap(~participant)
-
-# retrieving posterior samples
-# posterior_samples <- posterior_epred(object = gp_model, newdata = time_seq)
-# posterior_samples <- posterior_epred(object = gam, newdata = time_seq)
-# library(modelr)
-# meta_gam$data %>%
-#     add_epred_draws(meta_gam, ndraws = 20, re_formula = NA) %>%
+# 
+#     # finding onset, offset, and peak for the current threshold
+#     exceeding_times <- prob_y_above_0 %>%
+#         dplyr::filter(prob_ratio > threshold) %>%
+#         summarise(
+#             cluster_onset = min(time, na.rm = TRUE),
+#             cluster_offset = max(time, na.rm = TRUE)
+#             ) %>%
+#         mutate(true_onset = true_onset, true_offset = true_offset)
+# 
+#     # storing the results in the dataframe
+#     if (nrow(exceeding_times) > 0) {
+# 
+#         results$estimated_onset[i] <- exceeding_times$cluster_onset
+#         results$estimated_offset[i] <- exceeding_times$cluster_offset
+# 
+#         # computing errors
+#         results$error_onset[i] <- abs(
+#             exceeding_times$cluster_onset - exceeding_times$true_onset
+#             )
+#         results$error_offset[i] <- abs(
+#             exceeding_times$cluster_offset - exceeding_times$true_offset
+#             )
+# 
+#     }
+# 
+# }
+
+# saving the simulation results
+# saveRDS(object = results, file = "results/meta_gam_error_properties.rds")
+
+# loading the simulation results
+results <- readRDS(file = "results/meta_gam_error_properties.rds")
+
+results %>%
+    group_by(kvalue, threshold) %>%
+    summarise(across(estimated_onset:error_offset, mean) ) %>%
+    ungroup() %>%
+    pivot_longer(cols = error_onset:error_offset) %>%
+    mutate(name = factor(x = name, levels = c("error_onset", "error_offset") ) ) %>%
+    ggplot(
+        aes(
+            x = as.factor(kvalue),
+            y = as.factor(threshold),
+            fill = value
+            )
+        ) +
+    # geom_raster(interpolate = FALSE) +
+    geom_tile() +
+    # geom_point(
+    #     data = results %>%
+    #         group_by(kvalue, threshold) %>%
+    #         summarise(across(estimated_onset:error_offset, mean) ) %>%
+    #         ungroup() %>%
+    #         na.omit() %>%
+    #         filter(error_onset == min(error_onset) ),
+    #     # aes(x = kvalue, y = threshold),
+    #     aes(x = kvalue, y = value),
+    #     color = "orangered", size = 4, shape = 4,
+    #     # show.legend = FALSE
+    #     show.legend = TRUE
+    #     ) +
+    # scale_x_continuous(expand = c(0, 0) ) +
+    # scale_y_continuous(expand = c(0, 0) ) +
+    facet_wrap(~name) +
+    scale_x_discrete(expand = c(0, 0) ) +
+    scale_y_discrete(expand = c(0, 0), breaks = c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50) ) +
+    # scale_y_continuous(n.breaks=10) +
+    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
+    # scale_fill_viridis_c(direction = 1) +
+    labs(
+        x = "GAM basis dimension (k)",
+        y = "Threshold",
+        fill = "Error (s)"
+        )
+
+# results %>%
+#     group_by(kvalue, threshold) %>%
+#     summarise(across(estimated_onset:error_offset, mean) ) %>%
+#     ungroup() %>%
+#     # head()
+#     pivot_longer(cols = error_onset:error_offset) %>%
+#     mutate(name = factor(x = name, levels = c("error_onset", "error_offset") ) ) %>%
 #     ggplot(
 #         aes(
-#             x = time, y = .epred,
-#             colour = condition,
-#             group = interaction(participant, condition, .draw)
+#             x = threshold,
+#             y = value,
+#             colour = kvalue,
+#             fill = kvalue
 #             )
 #         ) +
-#     geom_hline(yintercept = 0, linetype = 2) +
-#     geom_vline(xintercept = true_onset, linetype = 2) +
-#     geom_vline(xintercept = true_offset, linetype = 2) +
-#     geom_line(show.legend = TRUE, alpha = 0.5, linewidth = 0.5) +
-#     facet_wrap(~participant) +
-#     labs(x = "Time (s)", y = "EEG signal (a.u.)")
-
-# computing the difference between cond2 and cond1
-# posterior_samples <- posterior_samples[, (ncol(posterior_samples)/2):ncol(posterior_samples)] -
-#     posterior_samples[, 1:(ncol(posterior_samples)/2+1)]
-
-# looping over different values of eps and threshold
-for (i in seq_len(nrow(results) ) ) {
-
-    # printing progress
-    # cat("Assessing combination:", i, "out of", nrow(results), "combinations...")
-
-    # retrieving current eps and threshold values
-    eps <- results$eps[i]
-    threshold <- results$threshold[i]
-
-    # computing probability metrics with the current eps value
-    # prob_y_above_0 <- data.frame(
-    #     time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx*2-1)
-    #     ) %>%
-    # mutate(post_prob = colMeans(posterior_samples) ) %>%
-    # mutate(m = colMeans(posterior_samples > eps) ) %>%
-    # mutate(prob_ratio = m / (1 - m) )
-
-    prob_y_above_0 <- error_model$data %>%
-        add_epred_draws(object = error_model, ndraws = n_post_samples) %>%
-        data.frame() %>%
-        dplyr::select(participant, time, condition, .epred, .draw) %>%
-        # dplyr::select(time, condition, .epred, .draw) %>%
-        pivot_wider(names_from = condition, values_from = .epred) %>%
-        mutate(epred_diff = cond2 - cond1) %>%
-        # computing mean posterior prob at the participant level
-        group_by(participant, time) %>%
-        summarise(m = mean(epred_diff > eps) ) %>%
-        ungroup() %>%
-        # computing mean posterior prob at the group level
-        group_by(time) %>%
-        summarise(m = mean(m) ) %>%
-        mutate(prob_ratio = m / (1 - m) ) %>%
-        ungroup()
-
-    # finding onset, offset, and peak for the current threshold
-    exceeding_times <- prob_y_above_0 %>%
-        dplyr::filter(prob_ratio > threshold) %>%
-        summarise(
-            cluster_onset = min(time, na.rm = TRUE),
-            cluster_offset = max(time, na.rm = TRUE)
-            ) %>%
-        # mutate(
-        #     cluster_peak = prob_y_above_0 %>%
-        #     dplyr::filter(post_prob == max(post_prob, na.rm = TRUE) ) %>%
-        #     pull(time)
-        #     ) %>%
-        # mutate(true_onset = true_onset, true_peak = true_peak, true_offset = true_offset)
-        mutate(true_onset = true_onset, true_offset = true_offset)
-
-    # storing the results in the dataframe
-    if (nrow(exceeding_times) > 0) {
-
-        results$estimated_onset[i] <- exceeding_times$cluster_onset
-        # results$estimated_peak[i] <- exceeding_times$cluster_peak
-        results$estimated_offset[i] <- exceeding_times$cluster_offset
-
-        # computing errors
-        results$error_onset[i] <- abs(
-            exceeding_times$cluster_onset - exceeding_times$true_onset
-            )
-        # results$error_peak[i] <- abs(
-        #     exceeding_times$cluster_peak - exceeding_times$true_peak
+#     # geom_raster(interpolate = FALSE) +
+#     # geom_tile() +
+#     # geom_line() +
+#     geom_point() +
+#     # geom_point(
+#     #     data = results %>% dplyr::filter(error_onset == min(error_onset, na.rm = TRUE) ),
+#     #     # aes(x = kvalue, y = threshold),
+#     #     aes(x = as.factor(kvalue), y = threshold),
+#     #     color = "orangered", size = 4, shape = 4,
+#     #     # show.legend = FALSE
+#     #     show.legend = TRUE
+#     #     ) +
+#     # scale_x_continuous(expand = c(0, 0) ) +
+#     # scale_y_continuous(expand = c(0, 0) ) +
+#     facet_wrap(~name) +
+#     scale_colour_viridis_c() +
+#     # scale_x_discrete(expand = c(0, 0) ) +
+#     # scale_y_discrete(expand = c(0, 0) ) +
+#     # scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) )
+#     labs(
+#         x = "GAM basis dimension (k)",
+#         y = "Threshold",
+#         fill = "Error (s)"
 #         )
-        results$error_offset[i] <- abs(
-            exceeding_times$cluster_offset - exceeding_times$true_offset
-            )
-
-    }
-
-}
 
 # plotting the results
-onset_plot <- results %>%
-    ggplot(aes(x = eps, y = threshold, fill = error_onset) ) +
-    geom_tile(show.legend = FALSE) +
-    geom_point(
-        data = results %>% dplyr::filter(error_onset == min(error_onset, na.rm = TRUE) ),
-        aes(x = eps, y = threshold),
-        color = "orangered", size = 2, shape = 4,
-        show.legend = FALSE
-        ) +
-    scale_x_continuous(expand = c(0, 0) ) +
-    scale_y_continuous(expand = c(0, 0) ) +
-    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-    labs(
-        title = "Onset error",
-        x = "eps",
-        y = "Threshold",
-        fill = "Error"
-        )
-
-offset_plot <- results %>%
-    ggplot(aes(x = eps, y = threshold, fill = error_offset) ) +
-    geom_tile(show.legend = FALSE) +
-    geom_point(
-        data = results %>% dplyr::filter(error_offset == min(error_offset, na.rm = TRUE) ),
-        aes(x = eps, y = threshold),
-        color = "orangered", size = 2, shape = 4,
-        show.legend = FALSE
-        ) +
-    scale_x_continuous(expand = c(0, 0) ) +
-    scale_y_continuous(expand = c(0, 0) ) +
-    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-    labs(
-        title = "Offset error",
-        x = "eps",
-        y = "Threshold",
-        fill = "Error"
-        )
+# onset_plot <- results %>%
+#     # ggplot(aes(x = eps, y = threshold, fill = error_onset) ) +
+#     # ggplot(aes(x = as.factor(kvalue), y = threshold, fill = error_onset) ) +
+#     ggplot(aes(x = as.factor(kvalue), y = as.factor(threshold), fill = error_onset) ) +
+#     geom_tile(show.legend = FALSE) +
+#     geom_point(
+#         data = results %>% dplyr::filter(error_onset == min(error_onset, na.rm = TRUE) ),
+#         # aes(x = kvalue, y = threshold),
+#         aes(x = as.factor(kvalue), y = threshold),
+#         color = "orangered", size = 4, shape = 4,
+#         # show.legend = FALSE
+#         show.legend = TRUE
+#         ) +
+#     # scale_x_continuous(expand = c(0, 0) ) +
+#     # scale_y_continuous(expand = c(0, 0) ) +
+#     scale_x_discrete(expand = c(0, 0) ) +
+#     scale_y_discrete(expand = c(0, 0) ) +
+#     scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
+#     labs(
+#         title = "Onset error",
+#         # x = "eps",
+#         x = "GAM basis dimension (k)",
+#         y = "Threshold",
+#         fill = "Error (s)"
+#         )
+# 
+# offset_plot <- results %>%
+#     # ggplot(aes(x = eps, y = threshold, fill = error_offset) ) +
+#     # ggplot(aes(x = kvalue, y = threshold, fill = error_offset) ) +
+#     ggplot(aes(x = as.factor(kvalue), y = as.factor(threshold), fill = error_offset) ) +
+#     geom_tile(show.legend = FALSE) +
+#     geom_point(
+#         data = results %>% dplyr::filter(error_offset == min(error_offset, na.rm = TRUE) ),
+#         # aes(x = kvalue, y = threshold),
+#         aes(x = as.factor(kvalue), y = threshold),
+#         color = "orangered", size = 4, shape = 4,
+#         # show.legend = FALSE
+#         show.legend = TRUE
+#         ) +
+#     # scale_x_continuous(expand = c(0, 0) ) +
+#     # scale_y_continuous(expand = c(0, 0) ) +
+#     scale_x_discrete(expand = c(0, 0) ) +
+#     scale_y_discrete(expand = c(0, 0) ) +
+#     scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
+#     labs(
+#         title = "Offset error",
+#         # x = "eps",
+#         x = "GAM basis dimension (k)",
+#         y = "Threshold",
+#         fill = "Error (s)"
+#         )
 
 # combining the plots
-onset_plot + offset_plot
+# onset_plot + offset_plot + plot_layout(guides = "collect")
 
@@
 # displaying best parameters values (according to both onset and offset error)
-results %>%
-    dplyr::select(-estimated_peak, -error_peak) %>%
-    arrange(error_onset, error_offset) %>%
-    data.frame() %>%
-    head()
+# results %>%
+#     group_by(kvalue, threshold) %>%
+#     summarise(across(estimated_onset:error_offset, mean) ) %>%
+#     ungroup() %>%
+#     arrange(error_onset, error_offset) %>%
+#     data.frame() %>%
+#     head(40)
 ```
@@
 
-We compared the ability of the GAMM to correctly estimate the onset and offset of the ERP difference to widely used methods.
-First, we conducted mass-univariate t-tests (thus treating each timestep independently) and identified the onset and offset of the ERP difference as the first and last values crossing an arbitrary significance threshold ($\alpha = 0.01$).
-We then followed the same approach but after applying different forms of multiplicity correction the the p-values.
-We compared two methods that control the false discovery rate (FDR) [i.e., BH95, @benjamini1995; and BY01, @benjamini2001], one method that controls the familywise error rate (FWER) [i.e., Holm–Bonferroni method, @holm1979], and two cluster-based methods [permutation with a single threshold and TFCE, @smith2009].
+To provide a more exhaustive assessment of the performance of the proposed approach, we compared the ability of the GAMM to correctly estimate the onset and offset of the ERP difference to widely-used methods.
+First, we conducted mass-univariate t-tests (thus treating each timestep independently) and identified the onset and offset of the ERP difference as the first and last values crossing an arbitrary significance threshold ($\alpha = 0.05$).
+We then followed the same approach but after applying different forms of multiplicity correction to the p-values.
+We compared two methods that control the false discovery rate (FDR) [i.e., BH95, @benjamini1995; and BY01, @benjamini2001], one method that controls the family-wise error rate (FWER) [i.e., Holm–Bonferroni method, @holm1979], and two cluster-based permutation methods [permutation with a single cluster-forming threshold and threshold-free cluster enhancement, TFCE, @smith2009].
 The BH95, BY01, and Holm corrections were applied to the p-values using the `p.adjust()` function in `R`.
@@
 We also compared these estimates to the onset and offset as estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared t-values [as in @rousselet_using_2025].
+@fig-corrections illustrates the onsets and offsets estimated by each method on a single simulated dataset and shows that all methods over-estimate the true onset and under-estimate the true offset.
+
+<!--
+
 For visualisation and interpretability purposes, we converted p-values to s-values, which can be interpreted as bits of surprising information, assuming a null effect [@greenland2019] (@fig-corrections).
 
-```{r permutations-1d, echo = FALSE, eval = FALSE, fig.cap = "T-values timecourse with onset identified using the permuco package (clustermass method)."}
-# centering AUC relative to chance level (0.5)
-decoding_data_perm <- decoding_data %>%
-    mutate(auc_centered = auc - 0.5) %>%
-    dplyr::select(-auc)
-
-# formatting the AUC data to be used with permuco
-dv <- decoding_data_perm %>%
-    pivot_wider(names_from = time, values_from = auc_centered) %>%
-    dplyr::select(-participant) %>%
-    data.frame()
-
-# formatting the intercept to be used with permuco
-iv <- decoding_data %>%
-    dplyr::select(participant) %>%
-    distinct() %>%
-    mutate(participant = as.factor(participant) ) %>%
-    data.frame() %>%
-    mutate(intercept = 1)
-
-# performing cluster-based permutation test
-cluster_results <- clusterlm(
-    formula = dv ~ intercept,
-    data = iv,
-    # data = decoding_data_perm,
-    # formula = auc_centered ~ participant,
-    # number of permutations (increase for accuracy)
-    # np = 2000,
-    # type of transformation
-    type = "signflip"
-    # using a t-test
-    # test = "t"
-    # multiple comparison correction
-    # multcomp = "clustermass"
-    # multcomp = "clusterdepth_head",
-    # multcomp = "tfce"
-    )
-
-# printing results
-print(cluster_results)
-
-# plotting significant clusters
-plot(cluster_results)
-
-# retrieving t and p-values
-f_values <- cluster_results$multiple_comparison$intercept$clustermass$main[, 1]
-p_values <- cluster_results$multiple_comparison$intercept$clustermass$main[, 2]
-```
-
-```{r fig-corrections, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Timecourse of squared t-values and s-values, with true onset (black dashed line) and onsets identified using the raw (uncorrected) p-values or the corrected p-values (BH, BY, Holm)."}
+-->
+
+```{r fig-corrections, echo = FALSE, results = "hide", fig.width = 8, fig.asp = 0.5, fig.cap = "Exemplary timecourse of squared t-values with true onset and offset (vertical black dashed lines) and onsets/offsets identified using the raw p-values, the corrected p-values (BH95, BY01, Holm), the cluster-based methods (Cluster-mass, TFCE), or using the binary segmentation method (Change point)."}
 # defining an arbitrary alpha threshold
-aath <- 0.01
+aath <- 0.05
 
@@
 
-# finding onsets
-from_timestep <- 1
-onset_p <- find_onset(
-    tests_results$pval[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$time[from_timestep:nrow(tests_results)]
-    )
-onset_bh <- find_onset(
-    tests_results$pval_bh[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$time[from_timestep:nrow(tests_results)]
-    )
-onset_by <- find_onset(
-    tests_results$pval_by[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$time[from_timestep:nrow(tests_results)]
-    )
-onset_holm <- find_onset(
-    tests_results$pval_holm[from_timestep:nrow(tests_results)] <= aath,
-    tests_results$time[from_timestep:nrow(tests_results)]
+# defining a function to find onset and offset in timeseries
+find_clusters <- function (df, threshold = aath) {
+
+    df %>%
+        select(-tval) %>%
+        pivot_longer(cols = -time) %>%
+        group_by(name) %>%
+        mutate(above_threshold = value <= threshold) %>%
+        mutate(cluster_change = c(TRUE, diff(above_threshold) != 0) ) %>%
+        filter(above_threshold) %>%
+        mutate(cluster_id = cumsum(cluster_change) ) %>%
+        group_by(name, cluster_id) %>%
+        summarise(cluster_onset = first(time), cluster_offset = last(time) ) %>%
+        select(cluster_onset, cluster_offset)
+
+}
+
+# estimating onset and offset using the binary segmentation method
+res <- cpt.meanvar(data = tests_results$tval, method = "BinSeg", Q = 2)
+onset_cpt <- tests_results$time[res@cpts[1]]
+offset_cpt <- tests_results$time[res@cpts[2]]
+
+# running the cluster-based permutation tests
+source(file = "code/mne_clusters.R")
+
+# joining results in a common dataframe
+methods_offset = data.frame(
+    method = c("pval", "pval_bh", "pval_by", "pval_holm", "cluster_mass", "cluster_tfce", "cpt"),
+    y_offset = rev(seq(from = -150, to = -20, length.out = 7) )
+    )
+
+clusters <- find_clusters(df = tests_results) %>%
+    rename(method = name, onset = cluster_onset, offset = cluster_offset) %>%
+    bind_rows(
+        data.frame(
+            method = "cpt",
+            onset = onset_cpt,
+            offset = offset_cpt
+            )
+        ) %>%
+    data.frame() %>%
+    bind_rows(data.frame(method = "cluster_mass", onset = onset_cluster_mass, offset = offset_cluster_mass) ) %>%
+    bind_rows(data.frame(method = "cluster_tfce", onset = onset_cluster_mass_tfce, offset = offset_cluster_mass_tfce) ) %>%
+    left_join(methods_offset, by = "method") %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c("pval", "pval_bh", "pval_by", "pval_holm", "cluster_mass", "cluster_tfce", "cpt"),
+            labels = c(
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass", "TFCE", "Change point"
+                ),
+            )
         )
@@
 # plotting the results
-default_colours <- met.brewer(name = "Johnson", n = 4)
-tvals_plot <- tests_results %>%
+tests_results %>%
     ggplot(aes(x = time, y = tval) ) +
+    geom_hline(yintercept = 0, linetype = 1) +
     geom_area(position = "identity") +
     geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = onset_p, linetype = 2, color = default_colours[1]) +
-    geom_vline(xintercept = onset_bh, linetype = 2, color = default_colours[2]) +
-    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[3]) +
-    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[4]) +
-    labs(x = "Time (s)", y = bquote(t^2) )
-
-plot_title <- paste0(
-    "True onset: ", true_onset, "s, ",
-    "Uncorrected onset: ", round(onset_p, 3), "s, ",
-    "BH onset: ", round(onset_bh, 3), "s, ",
-    "BY onset: ", round(onset_by, 3), "s, ",
-    "Holm onset: ", round(onset_holm, 3),  "s"
-    )
+    geom_vline(xintercept = true_offset, linetype = 2) +
+    geom_segment(
+        # data = all_methods_onset_offset,
+        data = clusters,
+        aes(
+            x = onset,
+            xend = offset,
+            y = y_offset,
+            yend = y_offset,
+            colour = method,
+            ),
+        linewidth = 1,
+        lineend = "round",
+        inherit.aes = FALSE
+        ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    labs(x = "Time (s)", y = "Squared t-value", colour = NULL)
 
@@
 # getting all heads on 4 fair coin tosses
-pvals_plot <- tests_results %>%
-    pivot_longer(cols = pval:pval_holm) %>%
-    mutate(sval = -log2(value) ) %>%
-    ggplot(aes(x = time, y = sval, colour = name) ) +
-    geom_line(linewidth = 0.5) +
-    geom_hline(yintercept = -log2(aath), linetype = 2) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = onset_p, linetype = 2, color = default_colours[1]) +
-    geom_vline(xintercept = onset_bh, linetype = 2, color = default_colours[2]) +
-    geom_vline(xintercept = onset_by, linetype = 2, color = default_colours[3]) +
-    geom_vline(xintercept = onset_holm, linetype = 2, color = default_colours[4]) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 4) ) +
-    labs(
-        x = "Time (s)",
-        y = "-log2(p-value)",
-        colour = "correction"
-        )
+# pvals_plot <- tests_results %>%
+#     pivot_longer(cols = pval:pval_holm) %>%
+#     mutate(sval = -log2(value) ) %>%
+#     ggplot(aes(x = time, y = sval, colour = name) ) +
+#     geom_density(
+#         aes(fill = name),
+#         stat = "identity",
+#         alpha = 0.8,
+#         show.legend = FALSE
+#         ) +
+#     geom_hline(yintercept = -log2(aath), linetype = 2) +
+#     geom_vline(xintercept = true_onset, linetype = 2) +
+#     geom_vline(xintercept = true_offset, linetype = 2) +
+#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 5) ) +
+#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 5) ) +
+#     labs(
+#         x = "Time (s)",
+#         y = "-log2(p-value)"
+#         )
 
 # combining the two plots
-tvals_plot + pvals_plot +
-    plot_annotation(title = plot_title) & theme(plot.title = element_text(hjust = 0.5) )
-```
-
-```{r changepoint, echo = FALSE, eval = FALSE, out.width = "75%", fig.cap = "T-values timecourse with true onset/offset and onset/offset identified using the changepoint package (binary segmentation method, in green)."}
-# using the changepoint package to identify onsets and offsets
-res <- cpt.meanvar(data = tests_results$tval, method = "BinSeg", Q = 2)
-
-# plotting the results
-tests_results %>%
-    ggplot(aes(x = time, y = tval) ) +
-    geom_line(linewidth = 1) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = true_offset, linetype = 2) +
-    geom_vline(
-        xintercept = tests_results$time[res@cpts[1]],
-        linetype = 3, color = "darkgreen"
-        ) +
-    geom_vline(
-        xintercept = tests_results$time[res@cpts[2]],
-        linetype = 3, color = "darkgreen"
-        ) +
-    labs(x = "Time (s)", y = bquote(t^2) ) +
-    ggtitle(
-        paste0(
-            "True onset: ", true_onset, "s, ", "Change point onset: ",
-            round(tests_results$time[res@cpts[1]], 3), "s, ",
-            "True offset: ", true_offset, "s, ", "Change point offset: ",
-            round(tests_results$time[res@cpts[2]], 3), "s"
-            )
-        )
-```
-
-```{r mne-cluster, echo = FALSE, eval = FALSE, fig.cap = "Cluster-based permutation tests via MNE-Python."}
-# loading the reticulate package
-library(reticulate)
-
-# importing the numpy and mne python modules
-np <- import("numpy")
-mne <- import("mne")
-
-# defining the function in R (it will be executed in Python)
-freq_stats_gat_matrix <- function (X) {
-
-    # converting R matrix to NumPy array
-    X_np <- np$array(X, dtype = "float64")
-
-    # running the statistical test
-    results <- mne$stats$spatio_temporal_cluster_1samp_test(
-        X_np,
-        out_type = "mask",
-        n_permutations = as.integer(1000),
-        n_jobs = as.integer(4),
-        verbose = TRUE
-        )
-
-    # extracting results
-    T_obs_ <- results[[1]]
-    clusters <- results[[2]]
-    p_values <- results[[3]]
-
-    # using the first slice of the 3D array
-    p_values_ <- np$transpose(np$ones_like(X_np[1, , drop = FALSE]) )
-
-    # assigning p-values to clusters
-    for (i in seq_along(clusters) ) {
-
-        # retrieving the current cluster
-        cluster_mask <- clusters[[i]][[1]]
-        idx_start <- cluster_mask$start+1
-        idx_stop <- cluster_mask$stop
-
-        # retrieving the p-value for this cluster
-        pval <- p_values[[i]]
-
-        # assigning this p-value to timesteps belonging
-        # to the current cluster
-        p_values_[idx_start:idx_stop] <- pval
-
-    }
-
-    # converting result back to R format
-    return (as.matrix(np$squeeze(np$transpose(p_values_) ) ) )
-
-}
-
-# converting decoding_data to a matrix
-data_matrix <- matrix(
-    decoding_data$auc,
-    ncol = length(unique(decoding_data$time) ),
-    byrow = TRUE
-    )
-
-# running the function
-chance_level <- 0.5
-p_values_results <- freq_stats_gat_matrix(data_matrix-chance_level)
-
-# converting back to a dataframe if needed
-p_values_df <- data.frame(
-    pval = p_values_results,
-    time = decoding_data$time
-    )
-
-# saving the scores
-saveRDS(object = p_values_df, file = "results/mne_permutation_decoding_scores.rds")
-```
-
-```{r fig-mne-cluster, echo = FALSE, eval = FALSE, out.width = "75%", fig.cap = "Cluster-based permutation tests via MNE-Python."}
-# importing the decoding scores
-p_values_df <- readRDS(file = "results/mne_permutation_decoding_scores.rds")
-
-# plotting the s-values (-log2(p-values))
-p_values_df %>%
-    mutate(sval = -log2(pval) ) %>%
-    ggplot(aes(x = time, y = sval) ) +
-    geom_line(linewidth = 0.5) +
-    geom_hline(yintercept = -log2(0.05), linetype = 2) +
-    labs(
-        x = "Time (s)",
-        y = "-log2(p-value)"
-        )
+# tvals_plot + pvals_plot +
+#     # plot_annotation(title = plot_title, tag_levels = "A") &
+#     # theme(plot.title = element_text(hjust = 0.5) )
+#     plot_annotation(tag_levels = "A") + plot_layout(guides = "collect")
 ```
@@
 
-Onset/offset estimation methods were assessed using the median absolute error (MAE) and variance of 10.000 simulated datasets...
+The various methods were assessed using the median error and variance of onset/offset estimates computed on 10.000 simulated datasets...
 
@@
 
-Next, we assessed the performance of the proposed approach on actual MEG data (decoding results of Nalborczyk et al., in preparation).
+Next, we assessed the performance of the proposed approach on actual MEG data [decoding results from @nalborczyk:inprep].
 We conducted time-resolved multivariate pattern analysis (MVPA), also known as decoding...
@@
 We fitted a similar GAM as discussed previously, but we replaced the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one to account for the bounded nature of AUC values (between 0 and 1) [for a tutorial on Beta regression, see @coretta2025].
-Note that although we chose a basis dimension of $k=50$, which seems appropriate for the present data, this choice should be adapted according to the properties of the modelled data (e.g., signal-to-noise ratio, prior low-pass filtering, sampling rate, etc) and should be assessed by the usual model checking tools (e.g., posterior predictive checks).
-We also define a region of practical equivalence [ROPE, @kruschke2017], defined as the chance level plus the standard deviation of the (group-level average) decoding performance during the baseline period.
-This ensures that...[^1]
-
-[^1]: Could an alternative approach be to include a predictor for "baseline vs. after baseline"?
+Note that although we chose a basis dimension of $k=50$, which seems appropriate for the present data, this choice should be adapted according to the properties of the modelled data (e.g., signal-to-noise ratio, prior low-pass filtering, sampling rate, etc) and should be assessed by the usual model checking tools (e.g., posterior predictive checks, see also @sec-basis).
+To distinguish signal from noise, we also defined a region of practical equivalence [ROPE, @kruschke2017], defined as the chance level plus the standard deviation of the (group-level average) decoding performance during the baseline period.
 
@@
 
-We assessed the reliability of the proposed approach using a form of permutation-based split-half reliability [e.g., @rosenblatt2018], which consisted of the following steps:
-
--   Create many (e.g., 10 or 20) half train/test splits of the data
--   For each fold, estimate the onset/offset on both splits using all methods
--   Then summarise the distribution of onset/offset with the median and variance across folds
-
-This will allow checking that the proposed approach produces reliable onset/offset estimates.
-
-```{r fig-decoding-data, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Group-level average decoding performance (N=32) superimposed with the GAM predictions (in blue) and the region of practical equivalence (ROPE, in orange) computed from the baseline period.
-The blue horizontal line indicates the timesteps at which the posterior probability ratio is equal to or greater than 20."}
+We assessed the reliability of the proposed approach using a form of permutation-based split-half reliability [as for instance in @rosenblatt2018], which consisted of the following steps.
+First, we created 1000 split halves of the data (i.e., with half the participants in the original data, that is, 16 participants).
+For each split, we estimated the onset/offset using all methods described previously.
+Third, we summarised the distribution of onset/offset estimates using the median "error" (i.e., difference between the split estimate and the estimate obtained using the full dataset) and the variance across splits.
+
+<!--
+
+OR, we could compute the onset/offset at the participant-level and correlate estimates across splits, BUT, individual-level estimates are not available for all methods...
+
+To test consistency between participants we calculated the split-half consistency across 1000 permutations.
+For each permutation, the participants were randomly split into two halves.
+Onset and offset were estimated at the participant level.
+Then, the Spearman rank correlation between the halves was calculated, corrected by the Spearman-Brown split-half reliability correction.
+A null distribution was calculated by randomly onset/offset values between words within one of the participant halves.
+
+-->
+
+```{r fig-decoding-data, cache = FALSE, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Group-level average decoding performance (N=32) superimposed with the GAM predictions (in blue) and the region of practical equivalence (ROPE, in orange) computed from the baseline period.
+The blue horizontal markers indicate the timesteps at which the posterior probability ratio exceeds 20."}
 # importing the reshaped MEG decoding results
@@
     summarise(sesoi = sd(auc_mean) ) %>%
-    # summarise(sesoi = sd(auc_mean) / sqrt(n_distinct(decoding_summary_df$participant) ) ) %>%
     pull(sesoi)
@@
     )
+
+# fitting the Beta GAMM
+# meg_decoding_gam <- brm(
+#     # auc_mean ~ s(time, bs = "cr", k = 50),
+#     auc_mean | se(auc_sd) ~ s(time, bs = "cr", k = 10) +
+#         s(time, participant, bs = "fs", xt = "cr", k = 10, m = 1),
+#     data = decoding_summary_df,
+#     # family = Beta(),
+#     # warmup = 2000,
+#     # iter = 5000,
+#     chains = 4,
+#     cores = 4,
+#     file = "models/meg_decoding_gamm.rds"
+#     )
 
@@
 # checking model's predictions against raw data
+brms_color <- met.brewer(name = "Johnson", n = 8)[8]
 plot(
     conditional_effects(x = meg_decoding_gam),
-    line_args = list(colour = "steelblue", fill = "steelblue", alpha = 0.3),
+    # line_args = list(colour = "steelblue", fill = "steelblue", alpha = 0.3),
+    line_args = list(colour = brms_color, fill = brms_color, alpha = 0.3),
     points = FALSE, plot = FALSE
@@
         fill ="orangered",
+        # fill ="darkred",
         alpha = 0.2
@@
             ),
-        colour = "steelblue",
+        # colour = "steelblue",
+        colour = brms_color,
         inherit.aes = FALSE,
-        lineend = "round", linewidth = 2
+        lineend = "round",
+        linewidth = 2
         ) +
@@
 First, we present the results from the simulation study, assessing the bias and variance of each method when applied to simulated data in which the ground truth is known.
-Second, we present the results obtained when applying the different methods to actual MEG data (decoding performance through time), assessing the reliability of each method and the stability of its estimates.
+Second, we present the results obtained when applying the different methods to actual MEG data (decoding performance through time), assessing the reliability of the estimates provided by each method.
 
@@
 
-@fig-simulation-mae-variance shows a summary of the simulation results, revealing that the proposed approach (`brms`) has the lowest MAE and variance for both the onset and offset estimates...
-
-```{r fig-simulation-mae-variance, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Median absolute error and variance of onset and offset estimates for each method."}
+@fig-simulation-mae-variance shows a summary of the simulation results, revealing that the proposed approach (`BGAM`) has the lowest median error and variance for both the onset and offset estimates.
+
+```{r fig-simulation-mae-variance, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Median error and variance of onset and offset estimates for each method.
+Variance is plotted on a log10 scale for visual purposes."}
 # loading the simulation results
@@
         error = case_when(
-            grepl("onset", measure) ~ abs(value - true_onset),
-            grepl("offset", measure) ~ abs(value - true_offset)
-            # grepl("onset", measure) ~ value - true_onset,
-            # grepl("offset", measure) ~ value - true_offset
+            # grepl("onset", measure) ~ abs(value - true_onset),
+            # grepl("offset", measure) ~ abs(value - true_offset)
+            grepl("onset", measure) ~ value - true_onset,
+            grepl("offset", measure) ~ value - true_offset
             )
@@
                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
-                "TFCE", "Change point", "brms"
+                "TFCE", "Change point", "BGAM (new method)"
                 ),
@@
         ) +
-    # geom_vline(xintercept = 0, linetype = 2) +
+    geom_vline(xintercept = 0, linetype = 2) +
     geom_point(
@@
     facet_wrap(~type) +
-    scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
-    labs(x = "Median absolute error (s)", y = "Variance (log-scale)")
-```
-
-```{r fig-simulation-distribution, echo = FALSE, fig.width = 9, fig.cap = "Distributions of onset and offset estimates for each method."}
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    labs(x = "Median error (s)", y = "Variance (log-scale)")
+```
+
+```{r fig-simulation-distributions, echo = FALSE, eval = FALSE, fig.width = 9, fig.cap = "Distributions of onset and offset estimates for each method."}
 sim_results %>%
@@
                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm",
-                "Cluster-mass", "TFCE", "Change point", "brms"
+                "Cluster-mass", "TFCE", "Change point", "BGAM (new method)"
                 ),
@@
         ) +
-    geom_vline(xintercept = true_onset) +
-    geom_vline(xintercept = true_offset) +
+    geom_vline(xintercept = true_onset, linetype = 2) +
+    geom_vline(xintercept = true_offset, linetype = 2) +
     geom_density(aes(fill = NULL) ) +
@@
     theme_light(base_size = 12, base_family = "Open Sans") +
-    scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
     labs(x = "Onset/Offset (s)", y = "Density")
@@
 
+```{r sim-errors, eval = FALSE, echo = FALSE}
+# plotting the results
+onset_plot <- sim_results %>%
+    select(simulation_id:offset_brms) %>%
+    mutate(
+        error_onset = abs(onset_brms - true_onset),
+        error_offset = abs(offset_brms - true_offset)
+        ) %>%
+    summarise(
+        error_onset = mean(error_onset),
+        error_offset = mean(error_offset),
+        .by = c(eps, threshold)
+        ) %>%
+    # arrange(error_onset) %>%
+    # head(10)
+    ggplot(aes(x = eps, y = threshold, fill = error_onset) ) +
+    geom_tile(show.legend = FALSE) +
+    geom_point(
+        data = . %>% dplyr::filter(error_onset == min(error_onset, na.rm = TRUE) ),
+        aes(x = eps, y = threshold),
+        color = "orangered", size = 2, shape = 4,
+        show.legend = FALSE
+        ) +
+    scale_x_continuous(expand = c(0, 0) ) +
+    scale_y_continuous(expand = c(0, 0) ) +
+    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
+    labs(
+        title = "Onset error",
+        x = "eps",
+        y = "Threshold",
+        fill = "Error"
+        )
+
+offset_plot <- sim_results %>%
+    select(simulation_id:offset_brms) %>%
+    mutate(
+        error_onset = abs(onset_brms - true_onset),
+        error_offset = abs(offset_brms - true_offset)
+        ) %>%
+    summarise(
+        error_onset = mean(error_onset),
+        error_offset = mean(error_offset),
+        .by = c(eps, threshold)
+        ) %>%
+    # arrange(error_offset) %>%
+    # head(10)
+    ggplot(aes(x = eps, y = threshold, fill = error_offset) ) +
+    geom_tile(show.legend = FALSE) +
+    geom_point(
+        data = . %>% dplyr::filter(error_offset == min(error_offset, na.rm = TRUE) ),
+        aes(x = eps, y = threshold),
+        color = "orangered", size = 2, shape = 4,
+        show.legend = FALSE
+        ) +
+    scale_x_continuous(expand = c(0, 0) ) +
+    scale_y_continuous(expand = c(0, 0) ) +
+    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
+    labs(
+        title = "Offset error",
+        x = "eps",
+        y = "Threshold",
+        fill = "Error"
+        )
+
+# combining the plots
+onset_plot + offset_plot
+```
+
 \newpage
@@
 @fig-onset-offset shows the group-level average decoding performance through time with onset and offset estimates for each method.
-The `brms_full` method is similar to the `brms` method except that the ROPE is defined on the entire dataset rather than on the split dataset.
 Overall, this figure shows that both the `Raw p-value` and `FDR BH95` methods were extremely lenient, considering that the decoding performance was above chance before the onset of the stimulus (false positive) and until the end of the trial.
 The `Change point` and `Cluster mass` methods were the most conservative methods, identifying a time window from approximately +60ms to +500ms.
-The `Holm`, `TFCE`, `brms`, and `brms_full` methods produced somewhat similar estimates of onset and offset, from approximately +60ms to +650ms.^[It should be noted that although each method can produce several "clusters" of timesteps, we only considered the first (onset) and last (offset) timesteps identified by each method to compute the error (difference).]
-
-```{r fig-onset-offset, echo = FALSE, fig.width = 9, fig.asp = 0.7, fig.cap = "Group-level average decoding performance through time with onset and offset estimates for each method (data from Nalborczyk et al., in preparation)."}
+The `Holm`, `TFCE`, and `BGAM` methods produced similar estimates of onset and offset, ranging from approximately +60ms to +650ms.^[It should be noted that although each method can produce several "clusters" of "significant" timesteps, we only considered the first (onset) and last (offset) timesteps identified by each method to compute the error (difference).]
+
+```{r fig-onset-offset, echo = FALSE, fig.width = 9, fig.asp = 0.7, fig.cap = "Group-level average decoding performance through time with onset and offset estimates for each method [data from @nalborczyk:inprep]."}
 # loading the results
@@
     pivot_wider(names_from = type, values_from = value) %>%
+    filter(method != "brms_full") %>%
     mutate(
@@
                 "p", "bh", "by", "holm", "cluster_mass",
-                "cluster_tfce", "cpt", "brms", "brms_full"
+                "cluster_tfce", "cpt", "brms"#, "brms_full"
                 ),
@@
                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
-                "TFCE", "Change point", "brms", "brms_full"
+                "TFCE", "Change point", "BGAM (new method)"#, "brms_full"
                 )
@@
         ) +
-    scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
     labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)", colour = NULL)
@@
 # saving the plot
-# ggsave(
-#     filename = "figures/meg_decoding_results_all_methods.png",
-#     width = 12, height = 6, dpi = 300,
-#     device = "png"
-#     )
-```
-
-@fig-reliability shows two properties of each method: i) the median difference between the onset and offset estimates from each data split and the onset and offset estimates from the full dataset (x-axis) and ii) and the variance of its onset and offset estimates across data splits (y-axis).
-This figure reveals that the `brms` *onset and offset* estimates on each split are the closest to the estimates from the full dataset on average (0ms difference for the onset estimate and 5ms difference for the offset estimate).
+ggsave(
+    filename = "figures/meg_decoding_results_all_methods.png",
+    width = 12, height = 6, dpi = 300,
+    device = "png"
+    )
+```
+
+@fig-reliability shows the median difference between the onset and offset estimates from each data split and the onset and offset estimates from the full dataset (x-axis) along with the variance of its onset and offset estimates across data splits (error bar).
+This figure reveals that the `BGAM` *onset and offset* estimates on each split are the closest to the estimates from the full dataset on average (0ms difference for the onset estimate and 5ms difference for the offset estimate).
 The `Raw p-value` method has similar performance, but given the aberrant estimates it produces (cf.
-@fig-onset-offset), the result that it is consistent between data splits and the full dataset is not convincing on its own.
+@fig-onset-offset), the fact that it is consistent between data splits and the full dataset is not convincing on its own.
+The `Change point` method also has a very good performance (i.e., very low difference between split estimates and full estimates), but produces too short cluster of significant decoding performance (cf.
+@fig-onset-offset).
 Overall, the figure reveals that for all other methods, split datasets produce later onset estimates and earlier offset estimates (as compared to the estimates from the model fitted on the full dataset).
-
-```{r fig-reliability, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Median absolute error and variance of onset (left) and offset (right) estimates for each method."}
+These results highlight that the results of any method should be assessed using both i) good asymptotic properties on simulated data, ii) sensible identified clusters in actual data, and iii) reliable/stable estimates on actual data.
+
+```{r fig-reliability, echo = FALSE, cache = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Median absolute error and variance of onset (left) and offset (right) estimates for each method."}
 # extract full data estimates
@@
     pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
+    # filter(method != "brms_full") %>%
     data.frame() %>%
@@
 # analysing the split results
+# reliability_results %>%
+#     slice(1:nrow(.)-1) %>%
+#     pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
+#     data.frame() %>%
+#     left_join(
+#         full_reliability_results,
+#         by = "measure",
+#         suffix = c("_split", "_full")
+#         ) %>%
+#     # mutate(error = abs(value_split - value_full) ) %>%
+#     mutate(error = value_split - value_full) %>%
+#     # mutate(error = (value_split - value_full)^2) %>%
+#     separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
+#     group_by(type, method) %>%
+#     summarise(
+#         MAE = median(error, na.rm = TRUE),
+#         # variance = var(error, na.rm = TRUE),
+#         variance = var(value_split, na.rm = TRUE)
+#         # variance = sum(error, na.rm = TRUE)
+#         # .groups = "drop"
+#         ) %>%
+#     ungroup() %>%
+#     # now plotting
+#     mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
+#     mutate(
+#         method = factor(
+#             x = method,
+#             levels = c(
+#                 "p", "bh", "by", "holm", "cluster_mass",
+#                 "cluster_tfce", "cpt", "brms", "brms_full"
+#                 ),
+#             labels = c(
+#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
+#                 "TFCE", "Change point", "brms", "brms_full"
+#                 ),
+#             )
+#         ) %>%
+#     ggplot(
+#         aes(
+#             x = MAE,
+#             y = variance,
+#             colour = method,
+#             fill = method
+#             )
+#         ) +
+#     geom_vline(xintercept = 0, linetype = 2) +
+#     geom_point(
+#         size = 2,
+#         pch = 21,
+#         colour = "white",
+#         show.legend = FALSE
+#         ) +
+#     geom_label_repel(
+#         aes(label = method),
+#         colour = "white",
+#         size = 3,
+#         segment.color = NA,
+#         max.overlaps = 20,
+#         show.legend = FALSE
+#         ) +
+#     scale_y_log10() +
+#     facet_wrap(~type, scales = "free") +
+#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+#     labs(
+#         x = "Average (median) difference to full dataset (s)",
+#         y = "Variance across splits"
+#         )
+
+# defining a consistent color palette using original method levels
+original_methods <- c(
+    "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
+    "TFCE", "Change point", "BGAM"#, "BGAM_full"
+    )
+
+color_palette <- setNames(
+    met.brewer(name = "Johnson", n = length(original_methods) ),
+    original_methods
+    )
+
+# or plotting only MAE in descending order
 reliability_results %>%
@@
     summarise(
-        MAE = median(error, na.rm = TRUE),
+        # MAE = median(error, na.rm = TRUE),
+        MAE = mean(error, na.rm = TRUE),
         # variance = var(error, na.rm = TRUE),
-        variance = var(value_split, na.rm = TRUE)
+        # variance = var(value_split, na.rm = TRUE)
+        variance = sd(error, na.rm = TRUE) / sqrt(n() )
         # variance = sum(error, na.rm = TRUE)
@@
     mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
+    filter(method != "brms_full") %>%
     mutate(
@@
                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
-                "TFCE", "Change point", "brms", "brms_full"
+                "TFCE", "Change point", "BGAM"#, "BGAM_full"
                 ),
@@
         ) %>%
+    # group_by(type) %>%
+    # arrange(abs(MAE), .by_group = TRUE) %>%
+    # arrange(abs(MAE), .by_group = TRUE) %>%
+    # ungroup() %>%
+    # mutate(method = forcats::fct_reorder(method, MAE, .desc = ) ) %>%
+    # mutate(method = forcats::fct_reorder2(method, type, MAE, .desc = FALSE) ) %>%
+    # head(20)
+    mutate(
+        method_reordered = tidytext::reorder_within(
+            x = method, by = 1/abs(MAE), within = type
+            )
+        ) %>%
+    # head()
     ggplot(
@@
             x = MAE,
-            y = variance,
+            xmin = MAE - variance,
+            xmax = MAE + variance,
+            # y = method,
+            y = method_reordered,
             colour = method,
@@
     geom_vline(xintercept = 0, linetype = 2) +
-    geom_point(
-        size = 2,
-        pch = 21,
-        colour = "white",
+    geom_pointrange(
+        size = 0.25,
+        linewidth = 0.75,
         show.legend = FALSE
         ) +
-    geom_label_repel(
-        aes(label = method),
-        colour = "white",
-        size = 3,
-        segment.color = NA,
-        max.overlaps = 20,
-        show.legend = FALSE
-        ) +
-    # scale_y_log10() +
     facet_wrap(~type, scales = "free") +
+    scale_y_reordered() +
     scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
     scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
+    # scale_fill_manual(values = color_palette) +
+    # scale_colour_manual(values = color_palette) +
     labs(
         x = "Average (median) difference to full dataset (s)",
-        y = "Variance across splits"
+        y = "Method"
         )
@@
                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
-                "TFCE", "Change point", "brms", "brms_full"
+                "TFCE", "Change point", "BGAM"#, "brms_full"
                 ),
@@
             method %in% c("Cluster-mass", "TFCE") ~ 3,
-            method %in% c("Change point", "brms") ~ 4
+            method %in% c("Change point", "BGAM") ~ 4
             )
@@
 # Discussion
-
-...
 
@@
 ```
+
+# Acknolwedgements
+
+Centre de Calcul Intensif d’Aix-Marseille is acknowledged for granting access to its high performance computing resources.
 
@@
 
+<!--
+
 Now, we want to test whether and when decoding performance is above chance level (0.5 for a binary decoding task).
 These two models are computationally heavier to fit (more observations and 2D smooth functions)...
+
+See https://bookdown.org/epeterson_2010/bios526_book/Module_5_3.html#bivariate-splines
+
+-->
+
+To model cross-temporal generalisation matrices of decoding performance (ROC AUC), we extended the initial (decoding) GAM to take into account the bivariate temporal distribution of AUC values, thus producing naturally smoothed estimates (timecourses) of AUC values and posterior probabilities.
+This model can be written as follows:
+
+$$
+\begin{aligned}
+\text{AUC}_{i} &\sim \mathrm{Beta}(\mu_{i}, \phi)\\
+g(\mu_{i}) &= f \left(\text{train}_{i}, \text{test}_{i} \right)\\
+\end{aligned}
+$$
+
+where we assume that AUC values come from a $\mathrm{Beta}$ distribution with two parameters $\mu$ and $\phi$.
+We can think of $f \left(\text{train}_{i}, \text{test}_{i} \right)$ as a surface (a smooth function of two variables) that we can model using a 2-dimensional splines.
+Let $\mathbf{s}_{i} = \left(\text{train}_{i}, \text{test}_{i} \right)$ be some pair of training and testing samples, and let $\mathbf{k}_{m} = \left(\text{train}_{m}, \text{test}_{m} \right)$ denote the $m^{\text{th}}$ knot in the domain of $\text{train}_{i}$ and $\text{test}_{i}$.
+We can then express the smooth function as:
+
+$$
+f \left(\text{train}_{i}, \text{test}_{i} \right) = \alpha + \sum_{m=1}^M \beta_{m} b_{m} \left(\tilde{s}_{i}, \tilde{k}_{m} \right)
+$$
+
+Note that $b_{m}(,)$ is a basis function that maps $R \times R \rightarrow R$.
+A popular bivariate basis function uses *thin-plate splines* [@wood2003], which extend to $\mathbf{s}_{i} \in \mathbb{R}^{d}$ and $\partial l_{g}$ penalties.
+These splines are designed to interpolate and approximate smooth surfaces over two dimensions (hence the "bivariate" term).
+For $d=2$ dimensions and $l=2$ (smoothness penalty involving second order derivative):
+
+$$
+f \left(\tilde{s}_{i} \right) = \alpha + \beta_{1} x_{i} + \beta_{2} z_{i} +\sum_{m=1}^{M} \beta_{2+m} b_m\left(\tilde{s}_i, \tilde{k}_m\right)
+$$
+
+using the the radial basis function given by:
+
+$$
+b_m\left(\tilde{s}_i, \tilde{k}_m\right)=\left\|\tilde{s}_i-\tilde{k}_m\right\|^2 \log \left\|\tilde{s}_i-\tilde{k}_m\right\|
+$$
+
+where $\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|$ is the Euclidean distance between the covariate $\mathbf{s}_{i}$ and the knot location $\mathbf{k}_{m}$. We fitted this model using `brms`...
 
@@
     family = Beta(),
-    iter = 5000,
-    chains = 4,
-    cores = 4,
+    warmup = 1000,
+    iter = 2000,
+    chains = 8,
+    cores = 8,
     file = "models/timegen_gam_t2.rds"
@@
 
-```{r gam-timegen-post-preds, echo = FALSE, fig.width = 8, fig.asp = 0.6, fig.cap = "Posterior probability of decoding accuracy being above chance level (2D GAM)."}
+```{r gam-timegen-post-preds, echo = FALSE, fig.width = 8, fig.asp = 0.6, fig.cap = "Predicted AUC values (left) and posterior probabilies of decoding accuracy being above chance level (right) according to the bivariate GAM."}
 # defining a function to plot posterior prob of slope above 0
@@
 # using the function with the decoding GAM
-plot_post_timegen(model = timegen_gam, data = timegen_data, chance_level = 0.5, eps = 0.01)
-```
-
-```{r gp-timegen-preds, eval = FALSE, echo = FALSE, fig.cap = "Posterior predictions of the GP fitted on decoding accuracy over time."}
-# plotting the posterior predictions
-plot(
-    conditional_effects(x = timegen_gp, prob = 0.95),
-    points = FALSE,
-    method = "posterior_epred",
-    surface = TRUE,
-    theme = theme_light(),
-    plot = TRUE
-    )
-
-# posterior predictive checks
-pp_check(timegen_gp)
-
-# checking timecourse of timegen_data's diagonal
-timegen_data %>%
-    filter(train_time == test_time) %>%
-    summarise(auc = mean(auc), .by = train_time) %>%
-    ggplot(aes(x = train_time, y = auc) ) +
-    geom_hline(yintercept = 0.5, linetype = 2) +
-    geom_line() +
-    labs(x = "Time (s)", y = "AUC")
+plot_post_timegen(
+    model = timegen_gam, data = timegen_data,
+    chance_level = 0.5, eps = 0.01
+    )
 ```
@@
 Could be extended to spatial and temporal dimensions with formulas such as `te(x, y, Time, d = c(2, 1) )`...
-
-\newpage
-
-# Mathematical formulation of the bivariate GAM
-
-```{=html}
-<!--
-
-See https://bookdown.org/epeterson_2010/bios526_book/Module_5_3.html#bivariate-splines
-
--->
-```
-
-To model cross-temporal generalisation matrices of decoding performance (ROC AUC), we extended the initial (decoding) GAM to take into account the bivariate temporal distribution of AUC values, thus producing naturally smoothed estimates (timecourses) of AUC values and posterior probabilities.
-This model can be written as follows:
-
-$$
-\begin{aligned}
-\text{AUC}_{i} &\sim \mathrm{Beta}(\mu_{i}, \phi)\\
-g(\mu_{i}) &= f \left(\text{train}_{i}, \text{test}_{i} \right)\\
-\end{aligned}
-$$
-
-where we assume that AUC values come from a $\mathrm{Beta}$ distribution with two parameters $\mu$ and $\phi$.
-We can think of $f \left(\text{train}_{i}, \text{test}_{i} \right)$ as a surface (a smooth function of two variables) that we can model using a 2-dimensional splines.
-Let $\mathbf{s}_{i} = \left(\text{train}_{i}, \text{test}_{i} \right)$ be some pair of training and testing samples, and let $\mathbf{k}_{m} = \left(\text{train}_{m}, \text{test}_{m} \right)$ denote the $m^{\text{th}}$ knot in the domain of $\text{train}_{i}$ and $\text{test}_{i}$.
-We can then express the smooth function as:
-
-$$
-f \left(\text{train}_{i}, \text{test}_{i} \right) = \alpha + \sum_{m=1}^M \beta_{m} b_{m} \left(\tilde{s}_{i}, \tilde{k}_{m} \right)
-$$
-
-Note that $b_{m}(,)$ is a basis function that maps $R \times R \rightarrow R$.
-A popular bivariate basis function uses *thin-plate splines* [@wood2003], which extend to $\mathbf{s}_{i} \in \mathbb{R}^{d}$ and $\partial l_{g}$ penalties.
-These splines are designed to interpolate and approximate smooth surfaces over two dimensions (hence the "bivariate" term).
-For $d=2$ dimensions and $l=2$ (smoothness penalty involving second order derivative):
-
-$$
-f \left(\tilde{s}_{i} \right) = \alpha + \beta_{1} x_{i} + \beta_{2} z_{i} +\sum_{m=1}^{M} \beta_{2+m} b_m\left(\tilde{s}_i, \tilde{k}_m\right)
-$$
-
-using the the radial basis function given by:
-
-$$
-b_m\left(\tilde{s}_i, \tilde{k}_m\right)=\left\|\tilde{s}_i-\tilde{k}_m\right\|^2 \log \left\|\tilde{s}_i-\tilde{k}_m\right\|
-$$
-
-where $\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|$ is the Euclidean distance between the covariate $\mathbf{s}_{i}$ and the knot location $\mathbf{k}_{m}$.
 
@@
 
+# Alternative to GAMs: Approximate Gaussian Process regression
+
+A Gaussian process (GP) is a stochastic process that defines the distribution over a collection of random variables indexed by a continuous variable, that is $\{f(t): t \in \mathcal{T}\}$ for some index set $\mathcal{T}$ [@riutort-mayol_practical_2023; @rasmussen2005].
+Whereas Bayesian linear regression outputs a distribution over the parameters of some predefined parametric model, the GP approach, in contrast, is a non-parametric approach, in that it finds a distribution over the possible functions that are consistent with the observed data.
+However, note that nonparametric does not mean there aren't parameters, it means that there are infinitely many parameters.
+
+From [brms documentation](https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp): A GP is a stochastic process, which describes the relation between one or more predictors $x=\left(x_{1}, \ldots, x_{d}\right)$ and a response $f(x)$, where $d$ is the number of predictors.
+A GP is the generalization of the multivariate normal distribution to an infinite number of dimensions.
+Thus, it can be interpreted as a prior over functions.
+The values of $f()$ at any finite set of locations are jointly multivariate normal, with a covariance matrix defined by the covariance kernel $k_p\left(x_i, x_j\right)$, where $p$ is the vector of parameters of the GP:
+
+$$
+\left(f\left(x_{1}\right), \ldots f\left(x_{n}\right) \sim \operatorname{MVN}\left(0, \left(k_p\left(x_{i}, x_{j}\right)\right)_{i, j=1}^{n}\right)\right.
+$$
+
+The smoothness and general behaviour of the function $f$ depends only on the choice of covariance kernel, which ensures that values that are close together in the input space will be mapped to similar output values...
+
+From this perspective, $f$ is a realisation of an infinite dimensional normal distribution:
+
+$$
+f \sim \mathrm{Normal}(0, C(\lambda))
+$$
+
+where $C$ is a covariance kernel with hyperparameters $\lambda$ that defines the covariance between two function values $f\left(t_1\right)$ and $f\left(t_2\right)$ for two time points $t_1$ and $t_2$ [@rasmussen2005].
+Similar to the different choices of the basis function for splines, different choices of the covariance kernel lead to different GPs.
+In this article, we consider the squared-exponential (a.k.a.
+radial basis function) kernel, which computes the squared distance between points and converts it into a measure of similarity.
+It is defined as:
+
+$$
+C(\lambda) := C\left(t_1, t_2, \sigma, \gamma\right) := \sigma^2 \exp \left(-\frac{||t_1-t_2||^{2}}{2 \gamma^2}\right)
+$$
+
+with hyperparameters $\lambda = (\sigma, \gamma)$, expressing the overall scale of GP and the length-scale, respectively [@rasmussen2005].
+The advantages of this kernel are that it is computationally efficient and (infinitely) smooth making it a reasonable choice for the purposes of the present article.
+Here again, $\lambda$ hyperparameters are estimated from the data, along with all other model parameters.
+
+Taken from <https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10c-Gaussian-processes.html>: For a given vector $\mathbf{x}$, we can use the kernel to construct finite multi-variate normal distribution associated with it like so:
+
+$$
+\mathbf{x} \mapsto_{G P} \operatorname{MVNormal}(m(\mathbf{x}), k(\mathbf{x}, \mathbf{x}))
+$$
+
+where $m$ is a function that specifies the mean for the distribution associated with $\mathbf{x}$.
+This mapping is essentially the Gaussian process: a systematic association of vectors of arbitrary length with a suitable multi-variate normal distribution.
+
+Low-rank approximate Gaussian processes are of main interest in machine learning and statistics due to the high computational demands of exact Gaussian process models [@riutort-mayol_practical_2023]...
+
+```{r gp, echo = FALSE, eval = FALSE, message = FALSE}
+# See also <https://jtimonen.github.io/lgpr-usage/articles/math.html>...
+# Now we fit the GP regression model...
+# computing eeg_diff and averaging across participants
+# group_df <- raw_df %>%
+#     group_by(participant, time) %>%
+#     pivot_wider(names_from = condition, values_from = eeg, values_fn = mean) %>%
+#     mutate(eeg_diff = cond2 - cond1) %>%
+#     summarise(eeg_diff = mean(eeg_diff) ) %>%
+#     ungroup()
+
+# see https://betanalpha.github.io/assets/case_studies/gp_part3/part3.html
+# and https://betanalpha.github.io/assets/case_studies/gaussian_processes.html
+gp_priors <- c(
+    set_prior("normal(0, 1)", class = "Intercept"),
+    set_prior("normal(0, 1)", class = "b"),
+    set_prior("exponential(1)", class = "sigma"),
+    set_prior("exponential(1)", class = "sdgp")
+    )
+
+gp_model <- brm(
+    # k refers to the number of basis functions for
+    # computing Hilbert-space approximate GPs
+    # if k = NA (default), exact GPs are computed
+    # eeg ~ gp(time, by = condition),
+    # eeg ~ condition + gp(time, by = condition, k = 20, cov = "exp_quad"),
+    eeg ~ condition + gp(time, by = condition, k = 20, cov = "matern32"),
+    data = ppt_df,
+    family = gaussian(),
+    # prior = gp_priors,
+    control = list(adapt_delta = 0.99, max_treedepth = 15),
+    iter = 2000,
+    chains = 4,
+    cores = 4,
+    file = "models/gp.rds"
+    )
+```
+
+```{r meta-gp, echo = FALSE, eval = FALSE, message = FALSE}
+# fitting the GP
+meta_gp <- brm(
+    # using by-participant SD of ERPs across trials
+    eeg_mean | se(eeg_sd) ~
+        condition + gp(time, k = 20, by = condition) +
+        (1 | participant),
+    data = summary_df,
+    family = gaussian(),
+    control = list(adapt_delta = 0.99),
+    iter = 5000,
+    chains = 4,
+    cores = 4,
+    file = "models/meta_gp.rds"
+    )
+```
+
+\newpage
+
+# How to choose the GAM basis dimension? {#sec-basis}
+
+Here provide recommendation about how to define `k`.
+An option is to vary `k` and examine the predictions and posterior predictive checks (PPCs) of each model...
+In this example (@fig-choose-k)...
+However, it is not possible to provide general recommendations, as the optimal `k` depends on the sampling rate, the preprocessing steps (e.g., signal-to-noise ratio, low-pass filtering, etc), and the neural dynamics of the phenomenon under study.
+
+```{r fig-choose-k, results = "hide", echo = FALSE, fig.width = 9, fig.asp = 1, fig.cap = "Posterior predictions and posterior predictive checks for the GAM with varying k (in rows)."}
+# defining a list of possible k values
+# k_values <- c(5, 10, 20, 50)
+
+# initialising lists to store models and plots
+# gam_models <- list()
+# post_pred_plots <- list()
+# ppc_dens_plots <- list()
+# ppc_ecdf_plots <- list()
+# ppc_stat2d_plots <- list()
+# 
+# # looping over k values
+# for (k_value in k_values) {
+#     
+#     # printing progress
+#     cat("Processing k =", k_value, "\n")
+#     
+#     # fitting the GAM
+#     gam <- brm(
+#         eeg ~ condition + s(time, bs = "tp", k = k_value, by = condition),
+#         data = ppt_df,
+#         family = gaussian(),
+#         warmup = 2000,
+#         iter = 5000,
+#         chains = 4,
+#         cores = 4
+#         )
+#     
+#     # storing the model
+#     gam_models[[as.character(k_value)]] <- gam
+#     
+#     # generating posterior predictions
+#     post_pred_plots[[as.character(k_value)]] <- plot_post_preds(model = gam) +
+#         ggtitle(paste("Posterior predictions (k =", k_value, ")") )
+#     
+#     # generating posterior predictive checks
+#     ppc_dens_plots[[as.character(k_value)]] <- pp_check(gam, ndraws = 100, type = "dens_overlay") + 
+#         labs(x = "EEG signal", y = "Density") +
+#         ggtitle(paste("Density overlay (k =", k_value, ")") ) +
+#         theme(legend.position = "none")
+#     
+#     ppc_ecdf_plots[[as.character(k_value)]] <- pp_check(gam, ndraws = 100, type = "ecdf_overlay") + 
+#         labs(x = "EEG signal", y = "Cumulative density") +
+#         ggtitle(paste("ECDF overlay (k =", k_value, ")") ) +
+#         theme(legend.position = "none")
+#     
+#     ppc_stat2d_plots[[as.character(k_value)]] <- pp_check(gam, ndraws = 500, type = "stat_2d", stat = c("mean", "sd") ) + 
+#         labs(x = "Mean", y = "SD") +
+#         ggtitle(paste("Mean vs. SD (k =", k_value, ")") )
+# 
+# }
+
+# saving the results
+# saveRDS(object = post_pred_plots, file = "results/post_pred_plots.rds")
+# saveRDS(object = ppc_dens_plots, file = "results/ppc_dens_plots.rds")
+# saveRDS(object = ppc_ecdf_plots, file = "results/ppc_ecdf_plots.rds")
+# saveRDS(object = ppc_stat2d_plots, file = "results/ppc_stat2d_plots.rds")
+
+# importing the results
+post_pred_plots <- readRDS(file = "results/post_pred_plots.rds")
+ppc_dens_plots <- readRDS(file = "results/ppc_dens_plots.rds")
+ppc_ecdf_plots <- readRDS(file = "results/ppc_ecdf_plots.rds")
+ppc_stat2d_plots <- readRDS(file = "results/ppc_stat2d_plots.rds")
+
+# importing the results
+
+# pp_check(gam, ndraws = 100, type = "ecdf_overlay") + 
+#     labs(x = "EEG signal", y = "Cumulative density") + ggtitle(paste("ECDF overlay (k =", k_value, ")") ) + theme(legend.position = "none")
+# pp_check(gam, ndraws = 200, type = "stat_2d", stat = c("median", "mad") ) + 
+#     labs(x = "Mean", y = "SD") + ggtitle(paste("Mean vs. SD (k =", k_value, ")") )
+
+# arranging all plots in a 4x4 grid using patchwork
+# final_plot <- (
+#     post_pred_plots[["5"]] + ppc_dens_plots[["5"]] + ppc_ecdf_plots[["5"]] + ppc_stat2d_plots[["5"]] +
+#     post_pred_plots[["10"]] + ppc_dens_plots[["10"]] + ppc_ecdf_plots[["10"]] + ppc_stat2d_plots[["10"]] +
+#     post_pred_plots[["20"]] + ppc_dens_plots[["20"]] + ppc_ecdf_plots[["20"]] + ppc_stat2d_plots[["20"]] +
+#     post_pred_plots[["50"]] + ppc_dens_plots[["50"]] + ppc_ecdf_plots[["50"]] + ppc_stat2d_plots[["50"]]
+#     ) + plot_layout(ncol = 4, nrow = 4)
+final_plot <- (
+    post_pred_plots[["5"]] + ppc_dens_plots[["5"]] + ppc_stat2d_plots[["5"]] +
+    post_pred_plots[["10"]] + ppc_dens_plots[["10"]] + ppc_stat2d_plots[["10"]] +
+    post_pred_plots[["20"]] + ppc_dens_plots[["20"]] + ppc_stat2d_plots[["20"]] +
+    post_pred_plots[["50"]] + ppc_dens_plots[["50"]] + ppc_stat2d_plots[["50"]]
+    ) + plot_layout(ncol = 3, nrow = 4)
+
+# displaying the final plot
+print(final_plot)
+
+# saving the plot
+# ggsave(
+#     filename = "figures/k_value_ppc.png",
+#     width = 20, height = 15, dpi = 200,
+#     device = "png"
+#     )
+
+# combining posterior predictions and some posterior predictive checks
+# post_preds <- plot_post_preds(model = gam)
+# ppc_dens <- pp_check(object = gam, ndraws = 100, type = "dens_overlay") + labs(x = "EEG signal", y = "Density")
+# ppc_ecdf <- pp_check(object = gam, ndraws = 100, type = "ecdf_overlay") + labs(x = "EEG signal", y = "CDF")
+# ppc_stat2d <- pp_check(object = gam, ndraws = 200, type = "stat_2d", stat = c("mean", "sd") ) + labs(x = "Mean", y = "SD")
+# post_preds + ppc_dens + ppc_ecdf + ppc_stat2d
+
+# some posterior predictive checks
+# pp_check(gam, prefix = "ppd")
+# pp_check(gam, type = "error_hist")
+# pp_check(object = gam, ndraws = 10, type = "intervals", x = "time")
+# pp_check(object = gam, ndraws = 10, type = "scatter_avg")
+# pp_check(object = gam, ndraws = 20, type = "hist")
+# pp_check(object = gam, ndraws = 50, type = "stat_2d")
+# pp_check(object = gam, ndraws = 50, type = "ecdf_overlay")
+# pp_check(object = gam, ndraws = 200, type = "stat", stat = "mean")
+# pp_check(object = gam, ndraws = 200, type = "stat", stat = "median")
+# pp_check(object = gam, ndraws = 200, type = "stat_2d", stat = c("mean", "sd") )
+# pp_check(object = gam, ndraws = 200, type = "stat_2d", stat = c("median", "mad") )
+```
+
+\newpage
+
 # Integration with `MNE-Python`
 
-Explain how to use the `R` package with `MNE` epochs...
-
-```{r r-python, eval = FALSE}
-# TO-DO: adding some code here...
-```
+For users who are already familiar with `brms`, the recommended pipeline is to import ERPs or decoding results in `R` and analyse these data using the code provided in the main paper.
+However, it is also possible to simply call functions from the `neurogam R` package (available at <https://github.com/lnalborczyk/neurogam>), which come with sensible defaults.
+
+```{r, eval = FALSE, echo = TRUE}
+# installing (if needed) and loading the neurogam R package
+# remotes::install_github("https://github.com/lnalborczyk/neurogam")
+library(neurogam)
+
+# using the testing_through_time() function from the neurogam package
+gam_onset_offset <- testing_through_time(
+    data = raw_df,
+    participant_id = "participant", meeg_id = "eeg",
+    time_id = "time", predictor_id = "condition",
+    warmup = 1000, iter = 5000, chains = 4, cores = 4
+    )
+
+# displaying the results
+gam_onset_offset$clusters
+```
+
+The `neurogam` package can also be called from `Python` using the `rpy2` module, and can easily be integrated into `MNE-Python` pipelines.
+For example, we use it below to estimate the onset and offset of effects for one channel from a MNE evoked object.
+Note that the code used to reshape the `sample MNE` dataset is provided later.
+
+```{python, eval = FALSE, echo = TRUE}
+# loading the Python modules
+import rpy2.robjects as robjects
+from rpy2.robjects.packages import importr
+from rpy2.robjects import pandas2ri
+from rpy2.robjects.conversion import localconverter
+
+# importing the "neurogam" R package
+neurogam = importr("neurogam")
+
+# activating automatic pandas-R conversion
+pandas2ri.activate()
+
+# assuming reshaped_df is some M/EEG data reshaped in long format
+with localconverter(robjects.default_converter + pandas2ri.converter):
+
+    reshaped_df_r = robjects.conversion.py2rpy(reshaped_df)
+
+# using the testing_through_time() function from the neurogam R package
+gam_onset_offset = neurogam.testing_through_time(
+    data=reshaped_df_r,
+    post_prob_ratio_threshold=3,
+    multilevel=False
+    )
+
+# displaying the results
+print(list(gam_onset_offset) )
+```
+
+```{python, eval = FALSE, echo = TRUE}
+# loading the Python modules
+import mne
+from mne.datasets import sample
+import numpy as np
+import pandas as pd
+
+# defining the path to the "sample" dataset
+path = sample.data_path()
+
+# loading the evoked data
+evokeds = mne.read_evokeds(path / "MEG" / "sample" / "sample_audvis-ave.fif")
+
+# defining a function to reshape the data
+def reshape_eeg_channels_as_participants(evokeds, condition_labels):
+
+    all_dfs = []
+
+    for evoked, condition in zip(evokeds, condition_labels):
+
+        # selecting only EEG channels
+        picks = mne.pick_types(evoked.info, meg=False, eeg=True)
+        data = evoked.data[picks, :]
+        channel_names = np.array(evoked.ch_names)[picks]
+
+        n_channels, n_times = data.shape
+        # repeating time for each channel
+        times = np.tile(evoked.times, n_channels)
+        meeg_values = data.flatten()
+        pseudo_participant_ids = np.repeat(channel_names, n_times)
+
+        # converting to dataframe
+        df = pd.DataFrame({
+            "participant": pseudo_participant_ids,
+            "time": times,
+            "eeg": meeg_values,
+            "condition": condition
+            })
+
+        all_dfs.append(df)
+
+    return pd.concat(all_dfs, ignore_index=True)
+
+# picking two evoked conditions
+faces = evokeds[0]
+scrambled = evokeds[1]
+
+# reshaping data by pretending each EEG channel is a participant
+reshaped_df = reshape_eeg_channels_as_participants(
+    evokeds=[faces, scrambled],
+    condition_labels=["faces", "scrambled"]
+    )
+```

### 2025-04-02 brms_meeg.tex
minor updates to cache

@@
 \abstract{Time-resolved electrophysiological measurements such as those
-offered by magneto- or electro-encephalography (M/EEG) provide a unique
-window onto neural activity underlying cognitive processes. Typically,
-researchers are interested in testing whether and when such measures
-differ across conditions and/or groups. The conventional approach
-consists in conducting mass-univariate statistics through time followed
-by some form of multiplicity correction (e.g., FDR, FWER) or
-cluster-based inference. However, while allowing efficient error-rates
-control, cluster-based methods have an important downside: they shift
-the focus of inference from the timepoint to the cluster level, thus
-preventing any conclusion to be made about the precise onset or offset
-of effects (e.g., differences across conditions or groups). Here, we
-introduce a \emph{model-based} approch for analysing M/EEG timeseriesn
-such as ERPs or decoding performance through time, and their differences
-across conditions or groups. This approach relies on Bayesian
-generalised additive multilevel models and outputs the posterior
-probabilility of the effect being above 0 (or above chance) at every
-timestep, while naturally taking into account the temporal dependencies
-and between-subject variability present in such data. Using both
-simulated and actual M/EEG data, we show that the proposed approach
-largely outperforms conventional methods in estimating the onset and
-offset of M/EEG effects, producing more precise and more reliable
-estimates. We provide an R package implementing the approach and
-illustrate how to integrate it into M/EEG statistical pipelines in
+obtained through magneto- or electro-encephalography (M/EEG) offer a
+unique window into the neural activity underlying cognitive processes.
+Researchers are often interested in determining whether and when these
+signals differ across experimental conditions or participant groups. The
+conventional approach involves mass-univariate statistical testing
+across time, followed by corrections for multiple comparisons (e.g.,
+FDR, FWER) or cluster-based inference. While effective for controlling
+error rates, cluster-based methods come with a significant limitation:
+they shift the focus of inference from individual time points to
+clusters, making it difficult to draw precise conclusions about the
+onset or offset of observed effects. Here, we introduce a
+\emph{model-based} approch for analysing M/EEG timeseries such as
+event-related potentials (ERPs) or decoding performance over time. Our
+method leverages Bayesian generalised additive multilevel models
+(GAMMs), providing posterior probabilities that an effect is above zero
+(or above chance) at each time point, while naturally accounting for
+temporal dependencies and between-subject variability. Using both
+simulated and actual M/EEG datasets, we demonstrate that this approach
+substantially outperforms conventional methods in estimating the onset
+and offset of neural effects, yielding more precise and reliable
+results. We also provide an R package implementing the method and show
+how it can be seamlessly integrated into M/EEG analysis pipelines using
 MNE-Python.}
@@
 
+Understanding the temporal dynamics of cognitive processes requires
+methods that can capture fast-changing neural activity with high
+temporal resolution. Magnetoencephalography (MEG) and
+electroencephalography (EEG) are two such methods, widely used in
+cognitive neuroscience for their ability to track brain activity at the
+millisecond scale. These techniques provide rich time series data that
+reflect how neural responses unfold in response to stimuli or tasks. A
+central goal in many M/EEG studies is to determine whether---and
+when---neural responses differ across experimental conditions or
+participant groups.
+
 Here are some useful references to be discussed
@@
 \citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019})\ldots{}
-As pointed by the original authors themselves
-(\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007}), ``there There is
-a conflict between this interest in localized effects and our choice for
-a global null hypothesis: by controlling the FA {[}false alarm{]} rate
+
+However, as pointed by the original authors themselves
+(\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007}), ``there is a
+conflict between this interest in localized effects and our choice for a
+global null hypothesis: by controlling the FA {[}false alarm{]} rate
 under this global null hypothesis, one cannot quantify the uncertainty
@@
 (\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
-\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019})\ldots{}
-In contrast, the proposed approach, based on Bayesian generalised
-additive multilevel models, allows quantifying the probability of
-effects at the level of timesteps, voxels, sensors (etc), while
-naturally taking into account spatiotemporal dependencies.
+\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}). In
+contrast, the proposed approach, based on Bayesian generalised additive
+multilevel models, allows quantifying the posterior probability of
+effects being above chance at the level of timesteps, voxels, sensors
+(etc), while naturally taking into account spatiotemporal dependencies
+present in M/EEG timeseries.
 
@@
 coefficients are penalised (usually through the squared of the smooth
-functions' second derivative) in a way that can be interpreted as a
-prior on the ``wiggliness'' of the function\ldots{}
-
-Terminology: \emph{splines} are functions composed of simpler functions.
-These simpler functions are basis functions (e.g., a cubic polynomial)
-and the set of basis functions is a \emph{basis}. Each basis function
-gets its coefficient and the resultant spline is the sum of these
-weighted basis functions.
+functions' second derivative) in a way that can be interpreted, in
+Bayesian terms, as a prior on the ``wiggliness'' of the function. In
+other words, more complex (wiggly) basis function are penalised. To
+clarify the terminology at this stage: \emph{splines} are functions
+composed of simpler functions. These simpler functions are basis
+functions (e.g., a cubic polynomial) and the set of basis functions is a
+\emph{basis}. Each basis function gets its coefficient and the resultant
+spline is the sum of these weighted basis functions.
 
@@
 
-Instead of averaring, obtain the smooth ERP signal from multilevel
-GAM\ldots{} less susceptible to outliers (Meulman et al., 2023)\ldots{}
+Instead of averaging, obtain the smooth ERP signal from multilevel
+GAM\ldots{} less susceptible to outliers
+(\citeproc{ref-meulman2023}{Meulman et al., 2023})\ldots{}
 
@@
 \texttt{R} package \texttt{mgcv} v 1.9-1 (\citeproc{ref-mgcv}{Wood,
-2017b}) for specifying smooth effects. Figure~\ref{fig-post-prob-test}
+2017b}) for specifying smooth effects. Figure~\ref{fig-plot-post-slope}
 shows the predictions of this model together with the raw data.
@@
 
+\CommentTok{\# timing the model fitting}
+\CommentTok{\# around 600s for 2000 iterations on 4 chains and 4 cores}
+\CommentTok{\# around 190s for 2000 iterations on 4 chains and 4 cores + cmdstanr and 01}
+\CommentTok{\# around 440s for 2000 iterations on 4 chains and 4 cores + threading(2), cmdstanr and 01}
+\CommentTok{\# library(tictoc)}
+\CommentTok{\# tic()}
+
 \CommentTok{\# fitting the GAM}
@@
     \CommentTok{\# cubic regression splines with k{-}1 basis functions}
-\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
+    \CommentTok{\# eeg \textasciitilde{} condition + s(time, bs = "cr", k = 20, by = condition),}
+    \CommentTok{\# thin{-}plate regression splines with k{-}1 basis functions}
+\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
     \AttributeTok{data =}\NormalTok{ ppt\_df,}
@@
     \AttributeTok{file =} \StringTok{"models/gam.rds"}
+    \DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
+    \CommentTok{\# speeding up the model fitting \#}
+    \DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
+    \CommentTok{\# backend = "cmdstanr",}
+    \CommentTok{\# threads = threading(2),}
+    \CommentTok{\# stan\_model\_args = list(stanc\_options = list("O1") )}
 \NormalTok{    )}
+
+\CommentTok{\# stop timing}
+\CommentTok{\# toc()}
 \end{Highlighting}
@@
 \texttt{condition} being above \(0+\epsilon\)
-(Figure~\ref{fig-post-prob-test}), with \(\epsilon := 0.05\), which can
-be interpreted as the smallest effect size of interest.
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-post-prob-test}Posterior probability of the ERP
-difference (slope) being above 0 according to the GAM.}
-
-\centering{
-
-\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-test-1.pdf}
-
-}
-
-\end{figure}%
-
-We can also express this as the ratio of posterior probabilities (i.e.,
+(Figure~\ref{fig-post-prob-ratio}, left), with \(\epsilon := 0.05\),
+which can be interpreted as the smallest effect size of interest. We can
+also express this as the ratio of posterior probabilities (i.e.,
 \(p/(1-p)\)) and visualise the timecourse of this ratio superimposed
 with the conventional thresholds on evidence ratios
-(Figure~\ref{fig-post-prob-ratio}).
+(Figure~\ref{fig-post-prob-ratio}, right).
 
@@
 
-\caption{\label{fig-post-prob-ratio}Ratio of posterior probability
-according to the GAM (on a log10 scale). Timesteps above threshold (10)
-are highlighted in green.}
+\caption{\label{fig-post-prob-ratio}Left: Posterior probability of the
+ERP difference (slope) being above 0 according to the GAM. Right: Ratio
+of posterior probability according to the GAM (on a log10 scale).
+Timesteps above threshold (10) are highlighted in green.}
 
@@
 
-\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.pdf}
 
@@
 
-Next we fit a hierarchical/multilevel GAM using summary statistics of
-ERPs (mean and SD) at the participant level (similar to what is done in
-meta-analysis). Although it is possible to fit a GAMM at the
-single-trial level, we present a resource-efficient version of the model
-that is fitted directly on the by-participant summary statistics (mean
-and SD)\ldots{}
+Next we fit a multilevel version of the GAM. Although it is possible to
+fit a GAMM at the single-trial level, we present a resource-efficient
+version of the model that is fitted directly on the by-participant
+summary statistics (mean and SD, similar to what is done in
+meta-analysis\ldots{}
 
@@
 (\(\text{error}:=|\hat{\theta}-\theta|\)), according to various
-\texttt{eps} and \texttt{threshold} values. Remember that the signal is
-generated from a truncated Gaussian with an objective onset at 160 ms, a
-maximum at 250 ms, and an offset at 342 ms. Figure~\ref{fig-onset-error}
-shows that the hierarchical GAM can \emph{exactly} recover the true
-onset and offset values, given some reasonable choice of \texttt{eps}
-and \texttt{threshold} values (e.g., a threshold of 20).
-
-\begin{verbatim}
-   eps threshold estimated_onset estimated_offset error_onset error_offset
-1 0.00        13            0.16             0.34           0        0.002
-2 0.00        14            0.16             0.34           0        0.002
-3 0.00        15            0.16             0.34           0        0.002
-4 0.00        16            0.16             0.34           0        0.002
-5 0.00        17            0.16             0.34           0        0.002
-6 0.01        10            0.16             0.34           0        0.002
-\end{verbatim}
+\texttt{k} (GAM basis dimension) and \texttt{threshold} values. Remember
+that the signal is generated from a truncated Gaussian with an objective
+onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
+Figure~\ref{fig-onset-error} shows that the hierarchical GAM can
+\emph{exactly} recover the true onset and offset values, given some
+reasonable choice of \texttt{k} and \texttt{threshold} values (e.g., a
+threshold of 20). We provide more detailed recommendations on how to set
+\texttt{k} in Section~\ref{sec-basis}.
 
@@
 \caption{\label{fig-onset-error}Error function of onset (left) and
-offset (right) estimation according to various eps and threshold values
-(according to the hierarchical GAM). Minimum error values are indicated
-by red crosses.}
+offset (right) estimation according to various basis dimension and
+threshold values (according to the multilevel GAM).}
 
@@
 
-We compared the ability of the GAMM to correctly estimate the onset and
-offset of the ERP difference to widely used methods. First, we conducted
-mass-univariate t-tests (thus treating each timestep independently) and
-identified the onset and offset of the ERP difference as the first and
-last values crossing an arbitrary significance threshold
-(\(\alpha = 0.01\)). We then followed the same approach but after
-applying different forms of multiplicity correction the the p-values. We
-compared two methods that control the false discovery rate (FDR) (i.e.,
-BH95, \citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995}; and
-BY01, \citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001}), one
-method that controls the familywise error rate (FWER) (i.e.,
-Holm--Bonferroni method, \citeproc{ref-holm1979}{Holm, 1979}), and two
-cluster-based methods (permutation with a single threshold and TFCE,
-\citeproc{ref-smith2009}{S. Smith \& Nichols, 2009}). The BH95, BY01,
-and Holm corrections were applied to the p-values using the
+To provide a more exhaustive assessment of the performance of the
+proposed approach, we compared the ability of the GAMM to correctly
+estimate the onset and offset of the ERP difference to widely-used
+methods. First, we conducted mass-univariate t-tests (thus treating each
+timestep independently) and identified the onset and offset of the ERP
+difference as the first and last values crossing an arbitrary
+significance threshold (\(\alpha = 0.05\)). We then followed the same
+approach but after applying different forms of multiplicity correction
+to the p-values. We compared two methods that control the false
+discovery rate (FDR) (i.e., BH95, \citeproc{ref-benjamini1995}{Benjamini
+\& Hochberg, 1995}; and BY01, \citeproc{ref-benjamini2001}{Benjamini \&
+Yekutieli, 2001}), one method that controls the family-wise error rate
+(FWER) (i.e., Holm--Bonferroni method, \citeproc{ref-holm1979}{Holm,
+1979}), and two cluster-based permutation methods (permutation with a
+single cluster-forming threshold and threshold-free cluster enhancement,
+TFCE, \citeproc{ref-smith2009}{S. Smith \& Nichols, 2009}). The BH95,
+BY01, and Holm corrections were applied to the p-values using the
 \texttt{p.adjust()} function in \texttt{R}. The cluster-based inference
@@
 (\citeproc{ref-gramfort2013}{Gramfort, 2013}), called via the \texttt{R}
-package \texttt{reticulate} v 1.35.0 (\citeproc{ref-reticulate}{Ushey et
+package \texttt{reticulate} v 1.42.0 (\citeproc{ref-reticulate}{Ushey et
 al., 2024}). We also compared these estimates to the onset and offset as
 estimated using the binary segmentation algorithm, as implemented in the
-\texttt{R} package \texttt{changepoint} v 2.2.4
+\texttt{R} package \texttt{changepoint} v 2.3
 (\citeproc{ref-changepoint}{Killick et al., 2022a}), and applied
 directly to the squared t-values (as in
-\citeproc{ref-rousselet_using_2025}{Rousselet, 2025}). For visualisation
-and interpretability purposes, we converted p-values to s-values, which
-can be interpreted as bits of surprising information, assuming a null
-effect (\citeproc{ref-greenland2019}{Greenland, 2019})
-(Figure~\ref{fig-corrections}).
+\citeproc{ref-rousselet_using_2025}{Rousselet, 2025}).
+Figure~\ref{fig-corrections} illustrates the onsets and offsets
+estimated by each method on a single simulated dataset and shows that
+all methods over-estimate the true onset and under-estimate the true
+offset.
 
@@
 
-\caption{\label{fig-corrections}Timecourse of squared t-values and
-s-values, with true onset (black dashed line) and onsets identified
-using the raw (uncorrected) p-values or the corrected p-values (BH, BY,
-Holm).}
+\caption{\label{fig-corrections}Exemplary timecourse of squared t-values
+with true onset and offset (vertical black dashed lines) and
+onsets/offsets identified using the raw p-values, the corrected p-values
+(BH95, BY01, Holm), the cluster-based methods (Cluster-mass, TFCE), or
+using the binary segmentation method (Change point).}
 
@@
 
-Onset/offset estimation methods were assessed using the median absolute
-error (MAE) and variance of 10.000 simulated datasets\ldots{}
+The various methods were assessed using the median error and variance of
+onset/offset estimates computed on 10.000 simulated datasets\ldots{}
 
@@
 Next, we assessed the performance of the proposed approach on actual MEG
-data (decoding results of Nalborczyk et al., in preparation). We
-conducted time-resolved multivariate pattern analysis (MVPA), also known
-as decoding\ldots{} As a result, we have a timecourse of decoding
-performance (ROC AUC), bounded between 0 and 1, for each participant
-(for a total of 32 participants). Now, we want to \emph{test} whether
-the group-level average decoding accuracy is above chance (i.e., 0.5) at
-each timestep (Figure~\ref{fig-decoding-data}). We fitted a similar GAM
-as discussed previously, but we replaced the \(\mathrm{Normal}\)
-likelihood function by a \(\mathrm{Beta}\) one to account for the
-bounded nature of AUC values (between 0 and 1) (for a tutorial on Beta
-regression, see \citeproc{ref-coretta2025}{Coretta \& Bürkner, 2025}).
-Note that although we chose a basis dimension of \(k=50\), which seems
-appropriate for the present data, this choice should be adapted
-according to the properties of the modelled data (e.g., signal-to-noise
-ratio, prior low-pass filtering, sampling rate, etc) and should be
-assessed by the usual model checking tools (e.g., posterior predictive
-checks). We also define a region of practical equivalence (ROPE,
+data (decoding results from \citeproc{ref-nalborczyk:inprep}{Nalborczyk
+et al., in preparation}). We conducted time-resolved multivariate
+pattern analysis (MVPA), also known as decoding\ldots{} As a result, we
+have a timecourse of decoding performance (ROC AUC), bounded between 0
+and 1, for each participant (for a total of 32 participants). Now, we
+want to \emph{test} whether the group-level average decoding accuracy is
+above chance (i.e., 0.5) at each timestep
+(Figure~\ref{fig-decoding-data}). We fitted a similar GAM as discussed
+previously, but we replaced the \(\mathrm{Normal}\) likelihood function
+by a \(\mathrm{Beta}\) one to account for the bounded nature of AUC
+values (between 0 and 1) (for a tutorial on Beta regression, see
+\citeproc{ref-coretta2025}{Coretta \& Bürkner, 2025}). Note that
+although we chose a basis dimension of \(k=50\), which seems appropriate
+for the present data, this choice should be adapted according to the
+properties of the modelled data (e.g., signal-to-noise ratio, prior
+low-pass filtering, sampling rate, etc) and should be assessed by the
+usual model checking tools (e.g., posterior predictive checks, see also
+Section~\ref{sec-basis}). To distinguish signal from noise, we also
+defined a region of practical equivalence (ROPE,
 \citeproc{ref-kruschke2017}{Kruschke \& Liddell, 2017}), defined as the
 chance level plus the standard deviation of the (group-level average)
-decoding performance during the baseline period. This ensures
-that\ldots{}\footnote{Could an alternative approach be to include a
-  predictor for ``baseline vs.~after baseline''?}
+decoding performance during the baseline period.
 
@@
 We assessed the reliability of the proposed approach using a form of
-permutation-based split-half reliability (e.g.,
+permutation-based split-half reliability (as for instance in
 \citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}), which consisted
-of the following steps:
-
-\begin{itemize}
-\tightlist
-\item
-  Create many (e.g., 10 or 20) half train/test splits of the data
-\item
-  For each fold, estimate the onset/offset on both splits using all
-  methods
-\item
-  Then summarise the distribution of onset/offset with the median and
-  variance across folds
-\end{itemize}
-
-This will allow checking that the proposed approach produces reliable
-onset/offset estimates.
+of the following steps. First, we created 1000 split halves of the data
+(i.e., with half the participants in the original data, that is, 16
+participants). For each split, we estimated the onset/offset using all
+methods described previously. Third, we summarised the distribution of
+onset/offset estimates using the median ``error'' (i.e., difference
+between the split estimate and the estimate obtained using the full
+dataset) and the variance across splits.
 
@@
 the region of practical equivalence (ROPE, in orange) computed from the
-baseline period. The blue horizontal line indicates the timesteps at
-which the posterior probability ratio is equal to or greater than 20.}
+baseline period. The blue horizontal markers indicate the timesteps at
+which the posterior probability ratio exceeds 20.}
 
@@
 methods to actual MEG data (decoding performance through time),
-assessing the reliability of each method and the stability of its
-estimates.
+assessing the reliability of the estimates provided by each method.
 
@@
 Figure~\ref{fig-simulation-mae-variance} shows a summary of the
-simulation results, revealing that the proposed approach (\texttt{brms})
-has the lowest MAE and variance for both the onset and offset
-estimates\ldots{}
+simulation results, revealing that the proposed approach (\texttt{BGAM})
+has the lowest median error and variance for both the onset and offset
+estimates.
 
@@
 
-\caption{\label{fig-simulation-mae-variance}Median absolute error and
-variance of onset and offset estimates for each method.}
+\caption{\label{fig-simulation-mae-variance}Median error and variance of
+onset and offset estimates for each method. Variance is plotted on a
+log10 scale for visual purposes.}
 
@@
 \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-simulation-mae-variance-1.pdf}
-
-}
-
-\end{figure}%
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-simulation-distribution}Distributions of onset and
-offset estimates for each method.}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-simulation-distribution-1.pdf}
 
@@
 performance through time with onset and offset estimates for each
-method. The \texttt{brms\_full} method is similar to the \texttt{brms}
-method except that the ROPE is defined on the entire dataset rather than
-on the split dataset. Overall, this figure shows that both the
-\texttt{Raw\ p-value} and \texttt{FDR\ BH95} methods were extremely
-lenient, considering that the decoding performance was above chance
-before the onset of the stimulus (false positive) and until the end of
-the trial. The \texttt{Change\ point} and \texttt{Cluster\ mass} methods
-were the most conservative methods, identifying a time window from
-approximately +60ms to +500ms. The \texttt{Holm}, \texttt{TFCE},
-\texttt{brms}, and \texttt{brms\_full} methods produced somewhat similar
-estimates of onset and offset, from approximately +60ms to
-+650ms.\footnote{It should be noted that although each method can
-  produce several ``clusters'' of timesteps, we only considered the
-  first (onset) and last (offset) timesteps identified by each method to
-  compute the error (difference).}
+method. Overall, this figure shows that both the \texttt{Raw\ p-value}
+and \texttt{FDR\ BH95} methods were extremely lenient, considering that
+the decoding performance was above chance before the onset of the
+stimulus (false positive) and until the end of the trial. The
+\texttt{Change\ point} and \texttt{Cluster\ mass} methods were the most
+conservative methods, identifying a time window from approximately +60ms
+to +500ms. The \texttt{Holm}, \texttt{TFCE}, and \texttt{BGAM} methods
+produced similar estimates of onset and offset, ranging from
+approximately +60ms to +650ms.\footnote{It should be noted that although
+  each method can produce several ``clusters'' of ``significant''
+  timesteps, we only considered the first (onset) and last (offset)
+  timesteps identified by each method to compute the error (difference).}
 
@@
 performance through time with onset and offset estimates for each method
-(data from Nalborczyk et al., in preparation).}
+(data from \citeproc{ref-nalborczyk:inprep}{Nalborczyk et al., in
+preparation}).}
 
@@
 
-Figure~\ref{fig-reliability} shows two properties of each method: i) the
-median difference between the onset and offset estimates from each data
-split and the onset and offset estimates from the full dataset (x-axis)
-and ii) and the variance of its onset and offset estimates across data
-splits (y-axis). This figure reveals that the \texttt{brms} \emph{onset
-and offset} estimates on each split are the closest to the estimates
-from the full dataset on average (0ms difference for the onset estimate
-and 5ms difference for the offset estimate). The \texttt{Raw\ p-value}
-method has similar performance, but given the aberrant estimates it
-produces (cf. Figure~\ref{fig-onset-offset}), the result that it is
-consistent between data splits and the full dataset is not convincing on
-its own. Overall, the figure reveals that for all other methods, split
-datasets produce later onset estimates and earlier offset estimates (as
-compared to the estimates from the model fitted on the full dataset).
+Figure~\ref{fig-reliability} shows the median difference between the
+onset and offset estimates from each data split and the onset and offset
+estimates from the full dataset (x-axis) along with the variance of its
+onset and offset estimates across data splits (error bar). This figure
+reveals that the \texttt{BGAM} \emph{onset and offset} estimates on each
+split are the closest to the estimates from the full dataset on average
+(0ms difference for the onset estimate and 5ms difference for the offset
+estimate). The \texttt{Raw\ p-value} method has similar performance, but
+given the aberrant estimates it produces (cf.
+Figure~\ref{fig-onset-offset}), the fact that it is consistent between
+data splits and the full dataset is not convincing on its own. The
+\texttt{Change\ point} method also has a very good performance (i.e.,
+very low difference between split estimates and full estimates), but
+produces too short cluster of significant decoding performance (cf.
+Figure~\ref{fig-onset-offset}). Overall, the figure reveals that for all
+other methods, split datasets produce later onset estimates and earlier
+offset estimates (as compared to the estimates from the model fitted on
+the full dataset). These results highlight that the results of any
+method should be assessed using both i) good asymptotic properties on
+simulated data, ii) sensible identified clusters in actual data, and
+iii) reliable/stable estimates on actual data.
 
@@
 \section{Discussion}\label{discussion}
-
-\ldots{}
 
@@
 
+\section{Acknolwedgements}\label{acknolwedgements}
+
+Centre de Calcul Intensif d'Aix-Marseille is acknowledged for granting
+access to its high performance computing resources.
+
 \newpage
@@
 
-\bibitem[\citeproctext]{ref-greenland2019}
-Greenland, S. (2019). Valid {\emph{P}} -Values Behave Exactly as They
-Should: Some Misleading Criticisms of {\emph{P}} -Values and Their
-Resolution With {\emph{S}} -Values. \emph{The American Statistician},
-\emph{73}(sup1), 106--114.
-\url{https://doi.org/10.1080/00031305.2018.1529625}
-
 \bibitem[\citeproctext]{ref-hastie2017}
@@
 
+\bibitem[\citeproctext]{ref-nalborczyk:inprep}
+Nalborczyk, L., Hauw, F., Torcy, H. de, Dehaene, S., \& Cohen, L. (in
+preparation). \emph{Neural and representational dynamics of tickertape
+synesthesia}.
+
 \bibitem[\citeproctext]{ref-pedersen_hierarchical_2019}
@@
 
+\bibitem[\citeproctext]{ref-rasmussen2005}
+Rasmussen, C. E., \& Williams, C. K. I. (2005). \emph{Gaussian Processes
+for Machine Learning}.
+\url{https://doi.org/10.7551/mitpress/3206.001.0001}
+
 \bibitem[\citeproctext]{ref-rigby2005}
@@
 Hearing}, \emph{23}. \url{https://doi.org/10.1177/2331216519832483}
+
+\bibitem[\citeproctext]{ref-riutort-mayol_practical_2023}
+Riutort-Mayol, G., Bürkner, P.-C., Andersen, M. R., Solin, A., \&
+Vehtari, A. (2023). Practical {Hilbert} space approximate {Bayesian
+Gaussian} processes for probabilistic programming. \emph{Statistics and
+Computing}, \emph{33}(1), 17.
+\url{https://doi.org/10.1007/s11222-022-10167-2}
 
@@
 
-Now, we want to test whether and when decoding performance is above
-chance level (0.5 for a binary decoding task). These two models are
-computationally heavier to fit (more observations and 2D smooth
-functions)\ldots{}
+To model cross-temporal generalisation matrices of decoding performance
+(ROC AUC), we extended the initial (decoding) GAM to take into account
+the bivariate temporal distribution of AUC values, thus producing
+naturally smoothed estimates (timecourses) of AUC values and posterior
+probabilities. This model can be written as follows:
+
+\[
+\begin{aligned}
+\text{AUC}_{i} &\sim \mathrm{Beta}(\mu_{i}, \phi)\\
+g(\mu_{i}) &= f \left(\text{train}_{i}, \text{test}_{i} \right)\\
+\end{aligned}
+\]
+
+where we assume that AUC values come from a \(\mathrm{Beta}\)
+distribution with two parameters \(\mu\) and \(\phi\). We can think of
+\(f \left(\text{train}_{i}, \text{test}_{i} \right)\) as a surface (a
+smooth function of two variables) that we can model using a
+2-dimensional splines. Let
+\(\mathbf{s}_{i} = \left(\text{train}_{i}, \text{test}_{i} \right)\) be
+some pair of training and testing samples, and let
+\(\mathbf{k}_{m} = \left(\text{train}_{m}, \text{test}_{m} \right)\)
+denote the \(m^{\text{th}}\) knot in the domain of \(\text{train}_{i}\)
+and \(\text{test}_{i}\). We can then express the smooth function as:
+
+\[
+f \left(\text{train}_{i}, \text{test}_{i} \right) = \alpha + \sum_{m=1}^M \beta_{m} b_{m} \left(\tilde{s}_{i}, \tilde{k}_{m} \right)
+\]
+
+Note that \(b_{m}(,)\) is a basis function that maps
+\(R \times R \rightarrow R\). A popular bivariate basis function uses
+\emph{thin-plate splines} (\citeproc{ref-wood2003}{Wood, 2003}), which
+extend to \(\mathbf{s}_{i} \in \mathbb{R}^{d}\) and \(\partial l_{g}\)
+penalties. These splines are designed to interpolate and approximate
+smooth surfaces over two dimensions (hence the ``bivariate'' term). For
+\(d=2\) dimensions and \(l=2\) (smoothness penalty involving second
+order derivative):
+
+\[
+f \left(\tilde{s}_{i} \right) = \alpha + \beta_{1} x_{i} + \beta_{2} z_{i} +\sum_{m=1}^{M} \beta_{2+m} b_m\left(\tilde{s}_i, \tilde{k}_m\right)
+\]
+
+using the the radial basis function given by:
+
+\[
+b_m\left(\tilde{s}_i, \tilde{k}_m\right)=\left\|\tilde{s}_i-\tilde{k}_m\right\|^2 \log \left\|\tilde{s}_i-\tilde{k}_m\right\|
+\]
+
+where \(\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|\) is the Euclidean
+distance between the covariate \(\mathbf{s}_{i}\) and the knot location
+\(\mathbf{k}_{m}\). We fitted this model using \texttt{brms}\ldots{}
 
@@
     \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
+    \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
+    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
+    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
+    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
     \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"}
@@
 
-\caption{Posterior probability of decoding accuracy being above chance
-level (2D GAM).}
+\caption{Predicted AUC values (left) and posterior probabilies of
+decoding accuracy being above chance level (right) according to the
+bivariate GAM.}
 
@@
 
-\section{Mathematical formulation of the bivariate
-GAM}\label{mathematical-formulation-of-the-bivariate-gam}
-
-To model cross-temporal generalisation matrices of decoding performance
-(ROC AUC), we extended the initial (decoding) GAM to take into account
-the bivariate temporal distribution of AUC values, thus producing
-naturally smoothed estimates (timecourses) of AUC values and posterior
-probabilities. This model can be written as follows:
+\section{Alternative to GAMs: Approximate Gaussian Process
+regression}\label{alternative-to-gams-approximate-gaussian-process-regression}
+
+A Gaussian process (GP) is a stochastic process that defines the
+distribution over a collection of random variables indexed by a
+continuous variable, that is \(\{f(t): t \in \mathcal{T}\}\) for some
+index set \(\mathcal{T}\) (\citeproc{ref-rasmussen2005}{Rasmussen \&
+Williams, 2005};
+\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
+2023}). Whereas Bayesian linear regression outputs a distribution over
+the parameters of some predefined parametric model, the GP approach, in
+contrast, is a non-parametric approach, in that it finds a distribution
+over the possible functions that are consistent with the observed data.
+However, note that nonparametric does not mean there aren't parameters,
+it means that there are infinitely many parameters.
+
+From
+\href{https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp}{brms
+documentation}: A GP is a stochastic process, which describes the
+relation between one or more predictors
+\(x=\left(x_{1}, \ldots, x_{d}\right)\) and a response \(f(x)\), where
+\(d\) is the number of predictors. A GP is the generalization of the
+multivariate normal distribution to an infinite number of dimensions.
+Thus, it can be interpreted as a prior over functions. The values of
+\(f()\) at any finite set of locations are jointly multivariate normal,
+with a covariance matrix defined by the covariance kernel
+\(k_p\left(x_i, x_j\right)\), where \(p\) is the vector of parameters of
+the GP:
 
 \[
-\begin{aligned}
-\text{AUC}_{i} &\sim \mathrm{Beta}(\mu_{i}, \phi)\\
-g(\mu_{i}) &= f \left(\text{train}_{i}, \text{test}_{i} \right)\\
-\end{aligned}
+\left(f\left(x_{1}\right), \ldots f\left(x_{n}\right) \sim \operatorname{MVN}\left(0, \left(k_p\left(x_{i}, x_{j}\right)\right)_{i, j=1}^{n}\right)\right.
 \]
 
-where we assume that AUC values come from a \(\mathrm{Beta}\)
-distribution with two parameters \(\mu\) and \(\phi\). We can think of
-\(f \left(\text{train}_{i}, \text{test}_{i} \right)\) as a surface (a
-smooth function of two variables) that we can model using a
-2-dimensional splines. Let
-\(\mathbf{s}_{i} = \left(\text{train}_{i}, \text{test}_{i} \right)\) be
-some pair of training and testing samples, and let
-\(\mathbf{k}_{m} = \left(\text{train}_{m}, \text{test}_{m} \right)\)
-denote the \(m^{\text{th}}\) knot in the domain of \(\text{train}_{i}\)
-and \(\text{test}_{i}\). We can then express the smooth function as:
+The smoothness and general behaviour of the function \(f\) depends only
+on the choice of covariance kernel, which ensures that values that are
+close together in the input space will be mapped to similar output
+values\ldots{}
+
+From this perspective, \(f\) is a realisation of an infinite dimensional
+normal distribution:
 
 \[
-f \left(\text{train}_{i}, \text{test}_{i} \right) = \alpha + \sum_{m=1}^M \beta_{m} b_{m} \left(\tilde{s}_{i}, \tilde{k}_{m} \right)
+f \sim \mathrm{Normal}(0, C(\lambda))
 \]
 
-Note that \(b_{m}(,)\) is a basis function that maps
-\(R \times R \rightarrow R\). A popular bivariate basis function uses
-\emph{thin-plate splines} (\citeproc{ref-wood2003}{Wood, 2003}), which
-extend to \(\mathbf{s}_{i} \in \mathbb{R}^{d}\) and \(\partial l_{g}\)
-penalties. These splines are designed to interpolate and approximate
-smooth surfaces over two dimensions (hence the ``bivariate'' term). For
-\(d=2\) dimensions and \(l=2\) (smoothness penalty involving second
-order derivative):
+where \(C\) is a covariance kernel with hyperparameters \(\lambda\) that
+defines the covariance between two function values \(f\left(t_1\right)\)
+and \(f\left(t_2\right)\) for two time points \(t_1\) and \(t_2\)
+(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). Similar to
+the different choices of the basis function for splines, different
+choices of the covariance kernel lead to different GPs. In this article,
+we consider the squared-exponential (a.k.a. radial basis function)
+kernel, which computes the squared distance between points and converts
+it into a measure of similarity. It is defined as:
 
 \[
-f \left(\tilde{s}_{i} \right) = \alpha + \beta_{1} x_{i} + \beta_{2} z_{i} +\sum_{m=1}^{M} \beta_{2+m} b_m\left(\tilde{s}_i, \tilde{k}_m\right)
+C(\lambda) := C\left(t_1, t_2, \sigma, \gamma\right) := \sigma^2 \exp \left(-\frac{||t_1-t_2||^{2}}{2 \gamma^2}\right)
 \]
 
-using the the radial basis function given by:
+with hyperparameters \(\lambda = (\sigma, \gamma)\), expressing the
+overall scale of GP and the length-scale, respectively
+(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). The
+advantages of this kernel are that it is computationally efficient and
+(infinitely) smooth making it a reasonable choice for the purposes of
+the present article. Here again, \(\lambda\) hyperparameters are
+estimated from the data, along with all other model parameters.
+
+Taken from
+\url{https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10c-Gaussian-processes.html}:
+For a given vector \(\mathbf{x}\), we can use the kernel to construct
+finite multi-variate normal distribution associated with it like so:
 
 \[
-b_m\left(\tilde{s}_i, \tilde{k}_m\right)=\left\|\tilde{s}_i-\tilde{k}_m\right\|^2 \log \left\|\tilde{s}_i-\tilde{k}_m\right\|
+\mathbf{x} \mapsto_{G P} \operatorname{MVNormal}(m(\mathbf{x}), k(\mathbf{x}, \mathbf{x}))
 \]
 
-where \(\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|\) is the Euclidean
-distance between the covariate \(\mathbf{s}_{i}\) and the knot location
-\(\mathbf{k}_{m}\).
+where \(m\) is a function that specifies the mean for the distribution
+associated with \(\mathbf{x}\). This mapping is essentially the Gaussian
+process: a systematic association of vectors of arbitrary length with a
+suitable multi-variate normal distribution.
+
+Low-rank approximate Gaussian processes are of main interest in machine
+learning and statistics due to the high computational demands of exact
+Gaussian process models
+(\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
+2023})\ldots{}
+
+\newpage
+
+\section{How to choose the GAM basis dimension?}\label{sec-basis}
+
+Here provide recommendation about how to define \texttt{k}. An option is
+to vary \texttt{k} and examine the predictions and posterior predictive
+checks (PPCs) of each model\ldots{} In this example
+(Figure~\ref{fig-choose-k})\ldots{} However, it is not possible to
+provide general recommendations, as the optimal \texttt{k} depends on
+the sampling rate, the preprocessing steps (e.g., signal-to-noise ratio,
+low-pass filtering, etc), and the neural dynamics of the phenomenon
+under study.
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-choose-k}Posterior predictions and posterior
+predictive checks for the GAM with varying k (in rows).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-choose-k-1.pdf}
+
+}
+
+\end{figure}%
 
@@
 
-Explain how to use the \texttt{R} package with \texttt{MNE}
-epochs\ldots{}
+For users who are already familiar with \texttt{brms}, the recommended
+pipeline is to import ERPs or decoding results in \texttt{R} and analyse
+these data using the code provided in the main paper. However, it is
+also possible to simply call functions from the \texttt{neurogam\ R}
+package (available at \url{https://github.com/lnalborczyk/neurogam}),
+which come with sensible defaults.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# TO{-}DO: adding some code here...}
+\CommentTok{\# installing (if needed) and loading the neurogam R package}
+\CommentTok{\# remotes::install\_github("https://github.com/lnalborczyk/neurogam")}
+\FunctionTok{library}\NormalTok{(neurogam)}
+
+\CommentTok{\# using the testing\_through\_time() function from the neurogam package}
+\NormalTok{gam\_onset\_offset }\OtherTok{\textless{}{-}} \FunctionTok{testing\_through\_time}\NormalTok{(}
+    \AttributeTok{data =}\NormalTok{ raw\_df,}
+    \AttributeTok{participant\_id =} \StringTok{"participant"}\NormalTok{, }\AttributeTok{meeg\_id =} \StringTok{"eeg"}\NormalTok{,}
+    \AttributeTok{time\_id =} \StringTok{"time"}\NormalTok{, }\AttributeTok{predictor\_id =} \StringTok{"condition"}\NormalTok{,}
+    \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{, }\AttributeTok{iter =} \DecValTok{5000}\NormalTok{, }\AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}
+\NormalTok{    )}
+
+\CommentTok{\# displaying the results}
+\NormalTok{gam\_onset\_offset}\SpecialCharTok{$}\NormalTok{clusters}
 \end{Highlighting}
@@
 
+The \texttt{neurogam} package can also be called from \texttt{Python}
+using the \texttt{rpy2} module, and can easily be integrated into
+\texttt{MNE-Python} pipelines. For example, we use it below to estimate
+the onset and offset of effects for one channel from a MNE evoked
+object. Note that the code used to reshape the \texttt{sample\ MNE}
+dataset is provided later.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# loading the Python modules}
+\ImportTok{import}\NormalTok{ rpy2.robjects }\ImportTok{as}\NormalTok{ robjects}
+\ImportTok{from}\NormalTok{ rpy2.robjects.packages }\ImportTok{import}\NormalTok{ importr}
+\ImportTok{from}\NormalTok{ rpy2.robjects }\ImportTok{import}\NormalTok{ pandas2ri}
+\ImportTok{from}\NormalTok{ rpy2.robjects.conversion }\ImportTok{import}\NormalTok{ localconverter}
+
+\CommentTok{\# importing the "neurogam" R package}
+\NormalTok{neurogam }\OperatorTok{=}\NormalTok{ importr(}\StringTok{"neurogam"}\NormalTok{)}
+
+\CommentTok{\# activating automatic pandas{-}R conversion}
+\NormalTok{pandas2ri.activate()}
+
+\CommentTok{\# assuming reshaped\_df is some M/EEG data reshaped in long format}
+\ControlFlowTok{with}\NormalTok{ localconverter(robjects.default\_converter }\OperatorTok{+}\NormalTok{ pandas2ri.converter):}
+
+\NormalTok{    reshaped\_df\_r }\OperatorTok{=}\NormalTok{ robjects.conversion.py2rpy(reshaped\_df)}
+
+\CommentTok{\# using the testing\_through\_time() function from the neurogam R package}
+\NormalTok{gam\_onset\_offset }\OperatorTok{=}\NormalTok{ neurogam.testing\_through\_time(}
+\NormalTok{    data}\OperatorTok{=}\NormalTok{reshaped\_df\_r,}
+\NormalTok{    post\_prob\_ratio\_threshold}\OperatorTok{=}\DecValTok{3}\NormalTok{,}
+\NormalTok{    multilevel}\OperatorTok{=}\VariableTok{False}
+\NormalTok{    )}
+
+\CommentTok{\# displaying the results}
+\BuiltInTok{print}\NormalTok{(}\BuiltInTok{list}\NormalTok{(gam\_onset\_offset) )}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# loading the Python modules}
+\ImportTok{import}\NormalTok{ mne}
+\ImportTok{from}\NormalTok{ mne.datasets }\ImportTok{import}\NormalTok{ sample}
+\ImportTok{import}\NormalTok{ numpy }\ImportTok{as}\NormalTok{ np}
+\ImportTok{import}\NormalTok{ pandas }\ImportTok{as}\NormalTok{ pd}
+
+\CommentTok{\# defining the path to the "sample" dataset}
+\NormalTok{path }\OperatorTok{=}\NormalTok{ sample.data\_path()}
+
+\CommentTok{\# loading the evoked data}
+\NormalTok{evokeds }\OperatorTok{=}\NormalTok{ mne.read\_evokeds(path }\OperatorTok{/} \StringTok{"MEG"} \OperatorTok{/} \StringTok{"sample"} \OperatorTok{/} \StringTok{"sample\_audvis{-}ave.fif"}\NormalTok{)}
+
+\CommentTok{\# defining a function to reshape the data}
+\KeywordTok{def}\NormalTok{ reshape\_eeg\_channels\_as\_participants(evokeds, condition\_labels):}
+
+\NormalTok{    all\_dfs }\OperatorTok{=}\NormalTok{ []}
+
+    \ControlFlowTok{for}\NormalTok{ evoked, condition }\KeywordTok{in} \BuiltInTok{zip}\NormalTok{(evokeds, condition\_labels):}
+
+        \CommentTok{\# selecting only EEG channels}
+\NormalTok{        picks }\OperatorTok{=}\NormalTok{ mne.pick\_types(evoked.info, meg}\OperatorTok{=}\VariableTok{False}\NormalTok{, eeg}\OperatorTok{=}\VariableTok{True}\NormalTok{)}
+\NormalTok{        data }\OperatorTok{=}\NormalTok{ evoked.data[picks, :]}
+\NormalTok{        channel\_names }\OperatorTok{=}\NormalTok{ np.array(evoked.ch\_names)[picks]}
+
+\NormalTok{        n\_channels, n\_times }\OperatorTok{=}\NormalTok{ data.shape}
+        \CommentTok{\# repeating time for each channel}
+\NormalTok{        times }\OperatorTok{=}\NormalTok{ np.tile(evoked.times, n\_channels)}
+\NormalTok{        meeg\_values }\OperatorTok{=}\NormalTok{ data.flatten()}
+\NormalTok{        pseudo\_participant\_ids }\OperatorTok{=}\NormalTok{ np.repeat(channel\_names, n\_times)}
+
+        \CommentTok{\# converting to dataframe}
+\NormalTok{        df }\OperatorTok{=}\NormalTok{ pd.DataFrame(\{}
+            \StringTok{"participant"}\NormalTok{: pseudo\_participant\_ids,}
+            \StringTok{"time"}\NormalTok{: times,}
+            \StringTok{"eeg"}\NormalTok{: meeg\_values,}
+            \StringTok{"condition"}\NormalTok{: condition}
+\NormalTok{            \})}
+
+\NormalTok{        all\_dfs.append(df)}
+
+    \ControlFlowTok{return}\NormalTok{ pd.concat(all\_dfs, ignore\_index}\OperatorTok{=}\VariableTok{True}\NormalTok{)}
+
+\CommentTok{\# picking two evoked conditions}
+\NormalTok{faces }\OperatorTok{=}\NormalTok{ evokeds[}\DecValTok{0}\NormalTok{]}
+\NormalTok{scrambled }\OperatorTok{=}\NormalTok{ evokeds[}\DecValTok{1}\NormalTok{]}
+
+\CommentTok{\# reshaping data by pretending each EEG channel is a participant}
+\NormalTok{reshaped\_df }\OperatorTok{=}\NormalTok{ reshape\_eeg\_channels\_as\_participants(}
+\NormalTok{    evokeds}\OperatorTok{=}\NormalTok{[faces, scrambled],}
+\NormalTok{    condition\_labels}\OperatorTok{=}\NormalTok{[}\StringTok{"faces"}\NormalTok{, }\StringTok{"scrambled"}\NormalTok{]}
+\NormalTok{    )}
+\end{Highlighting}
+\end{Shaded}
+
 \end{document}

### 2025-04-02 code/errors_simulation.R

@@ -2,7 +2,7 @@
 # Monte-Carlo simulation of onset/offset error properties #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on February 13, 2025                       #
+# Last updated on March 31, 2025                          #
 ###########################################################
 
 # importing R packages
@@ -13,6 +13,9 @@ library(tidybayes)
 library(brms)
 
 # importing Python modules
+# use_condaenv("/Users/ladislas/opt/anaconda3/envs/r-reticulate/bin/python", required = TRUE)
+# use_condaenv("r-reticulate", required = TRUE)
+use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
 np <- import("numpy")
 mne <- import("mne")
 
@@ -29,7 +32,7 @@ source("code/functions.R")
 meanpower <- unlist(read.table("code/meanpower.txt") )
 
 # defining the number of simulations to be performed
-nsims <- 1e2
+nsims <- 100
 
 # defining simulation parameters
 n_trials <- 50 # number of trials
@@ -220,7 +223,7 @@ freq_stats_cluster_matrix <- function (
 #     srate = srate, ronset = ronset
 #     )
 
-# testing the find_onset_offset function
+# testing the find_onset_offset() function
 # find_onset_offset(
 #     mask = tests_results$pval[1:nrow(tests_results)] <= aath,
 #     timeseries = tests_results$time[1:nrow(tests_results)]
@@ -233,7 +236,7 @@ sim_results <- data.frame()
 for (i in seq(from = 1, to = nsims, by = 1) ) {
     
     # printing progress
-    cat("Simulation number:", i, "out of", nsims, "simulations...\n")
+    cat("\n-------- Simulation number:", i, "out of", nsims, "simulations...\n")
     
     # simulating some EEG data
     raw_df <- generate_data(
@@ -262,7 +265,7 @@ for (i in seq(from = 1, to = nsims, by = 1) ) {
     meta_gam <- brm(
         # using by-participant SD of ERPs across trials
         eeg_mean | se(eeg_sd) ~
-            condition + s(time, bs = "cr", k = 10, by = condition) +
+            condition + s(time, bs = "cr", k = 20, by = condition) +
             (1 | participant),
         data = summary_df,
         family = gaussian(),
[…]

### 2025-04-02 code/errors_simulation_cluster.R

@@ -1,0 +1,489 @@
+###########################################################
+# Monte-Carlo simulation of onset/offset error properties #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on March 31, 2025                          #
+###########################################################
+
+# importing R packages
+library(changepoint)
+library(reticulate)
+library(tidyverse)
+library(tidybayes)
+library(brms)
+
+# for setting up the cluster
+library(doParallel)
+library(foreach)
+
+# total number of total CPU cores available on the HPC node
+total_cores <- 64
+
+# number of cores per brms model
+cores_per_model <- 8
+
+# number of parallel simulations to run at once
+n_parallel <- total_cores / cores_per_model
+
+# registering parallel backend
+cl <- makeCluster(n_parallel)
+registerDoParallel(cl)
+
+# importing Python modules
+# use_condaenv("/Users/ladislas/opt/anaconda3/envs/r-reticulate/bin/python", required = TRUE)
+# use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
+use_condaenv("r-reticulate", required = TRUE)
+np <- import("numpy")
+mne <- import("mne")
+
+# importing R version of Matlab code from Yeung et al. (2004)
+source("code/eeg_noise.R")
+
+# importing the ERP template with true onset = 160 ms, F=81, and max at F=126
+source("code/erp_template.R")
+
+# importing helper functions
+source("code/functions.R")
+
+# to use with the eeg_noise function
+meanpower <- unlist(read.table("code/meanpower.txt") )
+
+# defining the number of simulations to be performed
+nsims <- 1e4
+
+# defining simulation parameters
+n_trials <- 50 # number of trials
+n_ppt <- 20 # number of participants
+outvar <- 1 # noise variance
+srate <- 500 # sampling rate in Hz
+ronset <- seq(from = 150, to = 170, by = 2) # random onset for each participant
+
+# defining the true onset and offset in seconds
+true_onset <- 0.160
+true_offset <- 0.342
+
+# defining significance (alpha) thresholds
+significance_thresholds <- c(0.05, 0.01, 0.005, 0.001)
+
+# defining a function to generate a dataframe
+generate_data <- function (n_trials, n_ppt, outvar, srate, ronset) {
+    
+    for (P in 1:n_ppt) { # for each participant
+        
+        # get random onset
[…]

### 2025-04-02 code/generate_eeg_data.R

@@ -1,0 +1,88 @@
+# importing R version of Matlab code from Yeung et al. (2004)
+source("code/eeg_noise.R")
+
+# importing the ERP template with true onset = 160 ms, F=81, and max at F=126
+source("code/erp_template.R")
+
+# importing helper functions
+source("code/functions.R")
+
+# to use with the eeg_noise function
+meanpower <- unlist(read.table("code/meanpower.txt") )
+
+# defining the number of simulations to be performed
+# nsims <- 100
+
+# defining simulation parameters
+n_trials <- 50 # number of trials
+n_ppt <- 20 # number of participants
+outvar <- 1 # noise variance
+srate <- 500 # sampling rate in Hz
+ronset <- seq(from = 150, to = 170, by = 2) # random onset for each participant
+
+# defining the true onset and offset in seconds
+true_onset <- 0.160
+true_offset <- 0.342
+
+# defining significance (alpha) thresholds
+# significance_thresholds <- c(0.05, 0.01, 0.005, 0.001)
+
+# defining a function to generate a dataframe
+generate_data <- function (n_trials, n_ppt, outvar, srate, ronset) {
+    
+    for (P in 1:n_ppt) { # for each participant
+        
+        # get random onset
+        ponset <- sample(x = ronset, size = 1)
+        
+        # find starting point
+        st <- which(Xf == ponset)
+        
+        # pad vector
+        temp2 <- c(rep(0, st - 2), erp, rep(0, Nf - st - length(erp) + 2) )
+        
+        # initialising empty conditions
+        cond1 <- matrix(0, nrow = n_trials, ncol = Nf)
+        cond2 <- matrix(0, nrow = n_trials, ncol = Nf)
+        
+        for (trial in 1:n_trials) { # for each trial
+            
+            cond1[trial, ] <- temp1 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
+            cond2[trial, ] <- temp2 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
+            
+        }
+        
+        # converting results to dataframe
+        temp_df <- data.frame(
+            x = rep(Xf, 2 * nrow(cond1) ),
+            y = c(c(t(cond1) ), c(t(cond2) ) ),
+            trial = c(rep(1:n_trials, each = length(Xf) ), rep(1:n_trials, each = length(Xf) ) ),
[…]

### 2025-04-02 code/mne_clusters.R

@@ -1,0 +1,135 @@
+###########################################################
+# Split-half reliability of onset/offset estimates        #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on March 31, 2025                          #
+###########################################################
+
+# loading the reticulate package
+library(reticulate)
+
+# importing the numpy and mne python modules
+use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
+np <- import("numpy")
+mne <- import("mne")
+
+# defining the function in R (it will be executed in Python)
+freq_stats_cluster_matrix <- function (
+        X, timesteps,
+        cluster_type = c("mass", "tfce"),
+        alpha_level = 0.05
+        ) {
+    
+    # converting R matrix to NumPy array
+    X_np <- np$array(X, dtype = "float64")
+    
+    # defining the type of cluster-based method to use
+    if (cluster_type == "mass") {
+        
+        threshold <- py_none()
+        
+    } else {
+        
+        threshold <- dict(start = 0, step = 0.2)
+        
+    }
+    
+    # running the statistical test
+    results <- mne$stats$spatio_temporal_cluster_1samp_test(
+        X = X_np,
+        out_type = "mask",
+        threshold = threshold,
+        n_permutations = as.integer(2^12),
+        n_jobs = as.integer(4),
+        verbose = TRUE
+        )
+    
+    # extracting results
+    T_obs_ <- results[[1]]
+    clusters <- results[[2]]
+    p_values <- results[[3]]
+    
+    # retrieving significant clusters
+    if (cluster_type == "mass") {
+        
+        # using the first slice of the 3D array
+        p_values_ <- np$transpose(np$ones_like(X_np[1, , drop = FALSE]) )
+        
+        # defining the (main) cluster onset and offset
+        main_cluster_onset <- clusters[which(p_values < alpha_level)][[1]][[1]]$start+1
+        main_cluster_offset <- clusters[which(p_values < alpha_level)][[1]][[1]]$stop
+        
+        # assigning p-values to clusters
+        for (i in seq_along(clusters) ) {
+            
+            # retrieving the current cluster
+            cluster_mask <- clusters[[i]][[1]]
+            idx_start <- cluster_mask$start+1
+            idx_stop <- cluster_mask$stop
+            
[…]

### 2025-04-02 code/split_half_reliability.R

@@ -2,7 +2,7 @@
 # Split-half reliability of onset/offset estimates        #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on February 20, 2025                       #
+# Last updated on April 2, 2025                           #
 ###########################################################
 
 # importing R packages
@@ -18,6 +18,7 @@ library(brms)
 theme_set(theme_light(base_size = 12, base_family = "Open Sans") )
 
 # importing Python modules
+use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
 np <- import("numpy")
 mne <- import("mne")
 
@@ -34,7 +35,7 @@ mne <- import("mne")
 
 # importing the reshaped MEG decoding results
 raw_df <- read.csv(file = "code/decoding_results_reshaped.csv") %>%
-    # removing the last participant
+    # removing the last participant (to get a round number of participants)
     mutate(participant_id = cur_group_id(), .by = participant) %>%
     filter(participant_id < 33) %>%
     select(-participant_id)
@@ -81,28 +82,36 @@ full_sesoi <- summary_df %>%
     summarise(sesoi = sd(auc_mean) ) %>%
     pull(sesoi)
 
-# computing SD and max range of decoding performance during baseline
-# raw_df %>%
-#     filter(time < 0) %>%
-#     summarise(auc = mean(auc), .by = time) %>%
-#     summarise(auc_baseline_sd = sd(auc), max_baseline_range = max(abs(range(auc)-0.5) ) )
+# getting unique participants
+unique_participants <- unique(raw_df$participant)
 
-# defining a function to find onset and offset in timeseries
-find_onset_offset <- function (mask, timeseries) {
-    
-    # initialising onset and offset values
-    onset <- NA
-    offset <- NA
-    
-    # identifying the onset and offset
-    masked_timesteps <- try(timeseries[which(mask)], silent = TRUE)
-    onset <- head(x = masked_timesteps, n = 1)
-    offset <- tail(x = masked_timesteps, n = 1)
-    
-    # returning the onset and offset
-    return (c(onset, offset) )
-    
-}
+# getting the number of participants
+n_participants <- as.numeric(length(unique_participants) )
+
+# defining the number of splits
+n_splits <- 100
+
[…]

### 2025-04-02 brms_meeg.qmd
minor updates to readme

@@
     conflict-of-interest: "The authors have no conflicts of interest to disclose."
-    # financial-support: "This study was supported by Grant [Grant Number] from [Funding Source]."
 abstract: "Time-resolved electrophysiological measurements such as those obtained through magneto- or electro-encephalography (M/EEG) offer a unique window into the neural activity underlying cognitive processes.

### 2025-04-07 brms_meeg.qmd
updating the simulation code (now faster and more parallelised)

@@
     fig.pos = "!htb",
-    dev = "cairo_pdf"
+    # dev = "cairo_pdf",
+    dev = "png",
+    dpi = 300
     )
@@
 These techniques provide rich time series data that reflect how neural responses unfold in response to stimuli or tasks.
-A central goal in many M/EEG studies is to determine whether—and when—neural responses differ across experimental conditions or participant groups.
-
-Here are some useful references to be discussed [@combrisson_exceeding_2015; @ehinger_unfold_2019; @hayasaka_validating_2003; @luck_how_2017; @pedersen_hierarchical_2019; @frossard2021; @frossard2022; @gramfort2013; @maris2007; @pernet2015]...
+A central goal in many M/EEG studies is to determine whether and when neural responses differ across experimental conditions or participant groups.
+
+Here are some useful references to be discussed [@combrisson_exceeding_2015; @ehinger_unfold_2019; @hayasaka_validating_2003; @luck_how_2017; @frossard2021; @frossard2022; @gramfort2013; @maris2007; @pernet2015]...
 See also @maris2011 and @rosenblatt2018 (history of cluster-based approaches)...
@@
 
-## Previous work
+## Related work
 
@@
 where $y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)$ denotes that the observations $y_{i}$ are distributed as some member of the exponential family of distributions (e.g., Gaussian, Gamma, Beta, Poisson) with mean $\mu_{i}$ and scale parameter $\phi$; $g(\cdot)$ is the link function, $\mathbf{A}_{i}$ is the $i$th row of a known parametric model matrix, $\gamma$ is a vector of parameters for the parametric terms (to be estimated), $f_{j}$ is a smooth function of covariate $x_{j}$ (to be estimated as well).
-The smooth functions $f_{j}$ are represented in the model via penalised splines basis expansions of the covariates, that are a weighted sum of basis functions:
+The smooth functions $f_{j}$ are represented in the model via penalised splines basis expansions of the covariates, that are a weighted sum of $K$ simpler, basis functions:
 
@@
 where $\beta_{jk}$ is the weight (coefficient) associated with the $k$th basis function $b_{jk}()$ evaluated at the covariate value $x_{ij}$ for the $j$th smooth function $f_{j}$.
-Splines coefficients are penalised (usually through the squared of the smooth functions' second derivative) in a way that can be interpreted, in Bayesian terms, as a prior on the "wiggliness" of the function.
-In other words, more complex (wiggly) basis function are penalised.
 To clarify the terminology at this stage: *splines* are functions composed of simpler functions.
@@
 Each basis function gets its coefficient and the resultant spline is the sum of these weighted basis functions.
+Splines coefficients are penalised (usually through the squared of the smooth functions' second derivative) in a way that can be interpreted, in Bayesian terms, as a prior on the "wiggliness" of the function.
+In other words, more complex (wiggly) basis function are penalised.
 
 ## Bayesian generalised additive multilevel models
+
+Introduction to multilevel GAMs [@pedersen_hierarchical_2019]...
 
@@
 
-Focusing on identifying onset and offset of effects (as assessed by ERP differences or decoding performance)...
-Assessing the performance of a model-based approach (i.e., Bayesian GAMMs) to conventional methods (multiplicity corrections or cluster-based permutation)...
+Given the previously reported limitations of conventional methods to precisely identify the onset and offset of M/EEG effects (e.g., ERPs, decoding performance), we aimed to provide a model-based approach for estimating the onset and offset of such effects.
+To achieve this, we leveraged Bayesian generalised additive multilevel models (BGAMMs) fitted via the `brms R` package and compared the performance of this approach to conventional methods on both simulated and actual M/EEG data.
 
@@
 Following the approach of @sassenhagen2019 and @rousselet_using_2025, we simulated EEG data stemming from two conditions, one with noise only, and the other with noise + signal.
-As in previous studies, the noise was generated by superimposing 50 sinusoids at different frequencies, following an EEG-like spectrum [see details and code in @yeung2004].
+As in previous studies, the noise was generated by superimposing 50 sinusoids at different frequencies, following an EEG-like spectrum [see code in the online supplementary materials and details in @yeung2004].
 As in @rousselet_using_2025, the signal was generated from truncated Gaussian with an objective onset at 160 ms, a peak at 250 ms, and an offset at 342 ms.
 We simulated this signal for 250 timesteps between 0 and 0.5s, akin to a 500 Hz sampling rate.
-We simulated such data for a group of 20 participants with 50 trials per participant and condition (@fig-eeg).
-
-```{r fig-eeg, echo = FALSE, fig.width = 9, fig.cap = "Averaged (mean) simulated EEG activity in two conditions with 50 trials each, for a group of 20 participants.
-The error band represents the mean plus/minus 1 standard error of the mean."}
+We simulated data for a group of 20 participants with 50 trials per participant and condition (@fig-eeg).
+
+```{r fig-eeg, echo = FALSE, fig.width = 9, fig.cap = "Mean simulated EEG activity in two conditions with 50 trials each, for a group of 20 participants.
+The error band represents the mean +/- 1 standard error of the mean."}
 # importing R version of Matlab code from Yeung et al. (2004)
@@
 # importing helper functions
-source("code/functions.R")
-
-# to use with the eeg_noise function
+# source("code/functions.R")
+
+# EEG power information to use with the eeg_noise() function
 meanpower <- unlist(read.table("code/meanpower.txt") )
@@
 
-# removing raw_df
+# removing raw_df if it already exists
 if (exists("raw_df") ) rm(raw_df)
@@
 
+````{=html}
+<!--
+
 We computed the average of the ERP difference (@fig-erp)...
 
-```{r fig-erp, echo = FALSE, out.width = "75%", fig.cap = "Group-level average difference between conditions (mean +/- standard error of the mean).
+```{r fig-erp, eval = FALSE, echo = FALSE, out.width = "75%", fig.cap = "Group-level average difference between conditions (mean +/- standard error of the mean).
 The 'true' onset and offset are indicated by the vertical dashed lines."}
@@
 
+-->
+````
+
 ## Model fitting
@@
 We ran eight Markov Chain Monte-Carlo (MCMC) to approximate the posterior distribution, including each 5000 iterations and a warmup of 2000 iterations, yielding a total of $8 \times (5000-2000) = 24000$ posterior samples to use for inference.
-Posterior convergence was assessed examining trace plots as well as the Gelman–Rubin statistic $\hat{R}$.
+Posterior convergence was assessed examining trace plots as well as the Gelman–Rubin statistic $\hat{R}$ [@gabry2019; @gelman2020].
 The `brms` package uses the same syntax as `r pkrt("mgcv")` for specifying smooth effects.
@@
 
-```{r gam, message = FALSE}
+```{r gam, results = "hide"}
 # averaging across participants
@@
 
-```{r gamm, echo = FALSE, eval = FALSE, message = FALSE}
+```{r gamm, echo = FALSE, eval = FALSE, results = "hide"}
 # defining a contrast for condition
@@
 
-## Posterior probability of difference above 0
-
-We then plot the posterior predictions together with the posterior estimate of the slope for `condition` at each timestep (@fig-plot-post-slope).
-
-```{r fig-plot-post-slope, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior estimate of the ERP in each condition (left) or directly for the difference of ERPs (right) according to the GAM."}
+We depict the posterior predictions together with the posterior estimate of the slope for `condition` at each timestep (@fig-plot-post-slope).
+
+```{r fig-plot-post-slope, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior estimate of the EEG activity in each condition (left) and posterior estimate of the difference in EEG activity (right) according to the GAM."}
 # defining a function to plot posterior slope
@@
 
-We then compute the posterior probability of the slope for `condition` being above $0+\epsilon$ (@fig-post-prob-ratio, left), with $\epsilon := 0.05$, which can be interpreted as the smallest effect size of interest.
+We then compute the posterior probability of the slope for `condition` being above $0$ (@fig-post-prob-ratio, left).
 We can also express this as the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio, right).
-
-```{r fig-post-prob-test, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Posterior probability of the ERP difference (slope) being above 0 according to the GAM."}
+A ratio of 10 means that the probability of the difference being above 0 is 10 times higher than the probability of the difference not being above 0, given the data, the priors, and other model's assumptions.
+
+```{r fig-post-prob-test, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Posterior probability of the EEG difference (slope) being above 0 according to the GAM."}
 # defining a function to plot posterior prob of slope above 0
@@
 
-```{r fig-post-prob-ratio, echo = FALSE, out.width = "100%", fig.width = 9, fig.asp = 0.5, fig.cap = "Left: Posterior probability of the ERP difference (slope) being above 0 according to the GAM.
+```{r fig-post-prob-ratio, echo = FALSE, out.width = "100%", fig.width = 9, fig.asp = 0.5, fig.cap = "Left: Posterior probability of the EEG difference (slope) being above 0 according to the GAM.
 Right: Ratio of posterior probability according to the GAM (on a log10 scale).
@@
 
-## Multilevel modelling using ERP summary statistics
-
-Next we fit a multilevel version of the GAM.
-Although it is possible to fit a GAMM at the single-trial level, we present a resource-efficient version of the model that is fitted directly on the by-participant summary statistics (mean and SD, similar to what is done in meta-analysis...
+## Multilevel modelling using summary statistics
+
+The previous model only included constant (fixed) effects, thus not properly accounting for between-participant variability.
+We next fit a multilevel version of the GAM.
+Although it is possible to fit a GAMM at the single-trial level, we present a resource-efficient version of the model that is fitted directly on the by-participant summary statistics (mean and SD), similar to what is done in meta-analysis.
 
@@
 
-We then computed the difference between the true and estimated onset/offset of the ERP difference ($\text{error}:=|\hat{\theta}-\theta|$), according to various `k` (GAM basis dimension) and `threshold` values.
-Remember that the signal is generated from a truncated Gaussian with an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
-@fig-onset-error shows that the hierarchical GAM can *exactly* recover the true onset and offset values, given some reasonable choice of `k` and `threshold` values (e.g., a threshold of 20).
+We then compute the difference between the true and estimated onset/offset of the EEG difference according to various `k` (GAM basis dimension) and `threshold` values.
+Remember that the EEG signal was generated from a truncated Gaussian with an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
+@fig-onset-error shows that the multilevel GAM can *almost exactly* recover the true onset and offset values, given some reasonable choice of `k` and `threshold` values (e.g., a threshold of 20).
 We provide more detailed recommendations on how to set `k` in @sec-basis.
 
-```{r fig-onset-error, eval = TRUE, echo = FALSE, results = "hide", dpi = 300, fig.width = 9, fig.asp = 0.5, fig.cap = "Error function of onset (left) and offset (right) estimation according to various basis dimension and threshold values (according to the multilevel GAM)."}
+```{r fig-onset-error, eval = TRUE, echo = FALSE, results = "hide", dev = "png", dpi = 300, fig.width = 9, fig.asp = 0.5, fig.cap = "Average estimation error for the onset (left) and offset (right) according to various basis dimension and threshold values (according to the multilevel GAM)."}
 # defining values of k and threshold to test
@@
 # number of posterior samples to use
-# n_post_samples <- 1e3
+# n_post_samples <- 1e4
 
@@
 # initialising results dataframe
-# results <- crossing(kvalue = kvalues, threshold = threshold_values) %>%
 # results <- crossing(kvalue = kvalues, sim_id = 1:nsims, threshold = threshold_values) %>%
@@
 
-# looping over different values of kvalue and threshold
+# looping over different values of k and threshold
 # for (i in seq_len(nrow(results) ) ) {
@@
 # 
-#     # when we assess a new k-value
+#     # when we assess a new k-value or new sim_id, we generate new data
+#     # and fit a new model...
 #     if (threshold == 1) {
@@
 #         # sanity check
-#         head(raw_df)
+#         # head(raw_df)
 # 
@@
 #             chains = 8,
-#             cores = 8
+#             cores = 8,
+#             ###################################
+#             # when running on the HPC cluster #
+#             ###################################
+#             backend = "cmdstanr",
+#             stan_model_args = list(stanc_options = list("O1") )
 #             )
 # 
-#     }
-# 
 #         prob_y_above_0 <- error_model$data %>%
-#         add_epred_draws(object = error_model, ndraws = n_post_samples) %>%
-#         # add_epred_draws(object = meta_gam) %>%
+#             # add_epred_draws(object = error_model, ndraws = n_post_samples) %>%
+#             add_epred_draws(object = error_model) %>%
 #             data.frame() %>%
 #             dplyr::select(participant, time, condition, .epred, .draw) %>%
-#         # dplyr::select(time, condition, .epred, .draw) %>%
 #             pivot_wider(names_from = condition, values_from = .epred) %>%
 #             mutate(epred_diff = cond2 - cond1) %>%
-#         # computing mean posterior probability at the participant level
-#         # group_by(participant, time) %>%
-#         # summarise(m = mean(epred_diff > 0) ) %>%
-#         # ungroup() %>%
 #             # computing mean posterior probability at the group level
 #             group_by(time) %>%
-#         # summarise(m = mean(m) ) %>%
 #             summarise(m = mean(epred_diff > 0) ) %>%
 #             mutate(prob_ratio = m / (1 - m) ) %>%
-#         ungroup()
+#             ungroup() %>%
+#             # ensuring there is no 0 or +Inf values
+#             mutate(
+#                 prob_ratio = ifelse(
+#                     is.infinite(prob_ratio),
+#                     ndraws(error_model),
+#                     prob_ratio
+#                     )
+#                 ) %>%
+#             mutate(
+#                 prob_ratio = ifelse(
+#                     prob_ratio == 0,
+#                     1 / ndraws(error_model),
+#                     prob_ratio
+#                     )
+#                 )
+# 
+#     }
 # 
@@
 
+# running the simulation on the HPC cluster
+# source("code/meta_gam_errors_cluster.R")
+
 # loading the simulation results
-results <- readRDS(file = "results/meta_gam_error_properties.rds")
-
+# results <- readRDS(file = "results/meta_gam_error_properties.rds")
+results <- readRDS(file = "results/meta_gam_error_properties_cluster.rds")
+
+# plotting the results
 results %>%
+    group_by(kvalue, threshold, sim_id) %>%
+    # mutate(error_onset = ) %>%
+    summarise(
+        # value = if (onset_offset[1] == "onset") min(value) else max(value),
+        cluster_onset = min(cluster_onset),
+        cluster_offset = max(cluster_offset)
+        # .groups = "drop"
+        ) %>%
+    ungroup() %>%
     group_by(kvalue, threshold) %>%
-    summarise(across(estimated_onset:error_offset, mean) ) %>%
+    # summarise(across(estimated_onset:error_offset, median) ) %>%
+    # summarise(across(estimated_onset:error_offset, mean) ) %>%
+    mutate(
+        error_onset = abs(cluster_onset - true_onset),
+        error_offset = abs(cluster_offset - true_offset)
+        ) %>%
+    summarise(across(error_onset:error_offset, mean) ) %>%
     ungroup() %>%
@@
         ) +
-    # geom_raster(interpolate = FALSE) +
-    geom_tile() +
+    geom_raster(interpolate = FALSE) +
+    # geom_tile() +
     # geom_point(
@@
 #     summarise(across(estimated_onset:error_offset, mean) ) %>%
+#     # summarise(across(estimated_onset:error_offset, mode) ) %>%
 #     ungroup() %>%
@@
 
-To provide a more exhaustive assessment of the performance of the proposed approach, we compared the ability of the GAMM to correctly estimate the onset and offset of the ERP difference to widely-used methods.
+We then compared the ability of the GAMM to correctly estimate the onset and offset of the ERP difference to other widely-used methods.
 First, we conducted mass-univariate t-tests (thus treating each timestep independently) and identified the onset and offset of the ERP difference as the first and last values crossing an arbitrary significance threshold ($\alpha = 0.05$).
@@
 -->
-
-```{r fig-corrections, echo = FALSE, results = "hide", fig.width = 8, fig.asp = 0.5, fig.cap = "Exemplary timecourse of squared t-values with true onset and offset (vertical black dashed lines) and onsets/offsets identified using the raw p-values, the corrected p-values (BH95, BY01, Holm), the cluster-based methods (Cluster-mass, TFCE), or using the binary segmentation method (Change point)."}
+```
+
+```{r fig-corrections, echo = FALSE, results = "hide", fig.width = 8, fig.asp = 0.5, fig.cap = "Exemplary timecourse of squared t-values with true onset and offset (vertical black dashed lines) and onsets/offsets identified using the raw p-values, the corrected p-values (BH95, BY01, Holm), the cluster-based methods (Cluster mass, TFCE), or using the binary segmentation method (Change point)."}
 # defining an arbitrary alpha threshold
@@
             x = method,
-            levels = c("pval", "pval_bh", "pval_by", "pval_holm", "cluster_mass", "cluster_tfce", "cpt"),
+            levels = c(
+                "pval", "pval_bh", "pval_by", "pval_holm",
+                "cluster_mass", "cluster_tfce", "cpt"
+                ),
             labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass", "TFCE", "Change point"
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm",
+                "Cluster mass", "TFCE", "Change point"
                 ),
@@
 
-The various methods were assessed using the median error and variance of onset/offset estimates computed on 10.000 simulated datasets...
+The various methods were compared using the median error (i.e., true-estimated) and variance of onset/offset estimates computed on 10.000 simulated datasets.
 
@@
 
-Next, we assessed the performance of the proposed approach on actual MEG data [decoding results from @nalborczyk:inprep].
-We conducted time-resolved multivariate pattern analysis (MVPA), also known as decoding...
-As a result, we have a timecourse of decoding performance (ROC AUC), bounded between 0 and 1, for each participant (for a total of 32 participants).
-Now, we want to *test* whether the group-level average decoding accuracy is above chance (i.e., 0.5) at each timestep (@fig-decoding-data).
-We fitted a similar GAM as discussed previously, but we replaced the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one to account for the bounded nature of AUC values (between 0 and 1) [for a tutorial on Beta regression, see @coretta2025].
+To provide a more exhaustive assessment of these methods, we  also evaluated the performance of the proposed approach on actual MEG data [decoding results from @nalborczyk:inprep].
+In this study, we conducted time-resolved multivariate pattern analysis (MVPA, also known as decoding) of MEG data during reading tasks.
+As a result, we obtain a timecourse of decoding performance (ROC AUC), bounded between 0 and 1, for each participant (for a total of 32 participants).
+Next, we want to *test* whether the group-level average decoding accuracy is above chance (i.e., 0.5) at each timestep (@fig-decoding-data).
+To achieve this, we fitted a similar GAM as discussed previously, but we replaced the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one to account for the bounded nature of AUC values (between 0 and 1) [for a tutorial on Beta regression, see @coretta2025].
+
 Note that although we chose a basis dimension of $k=50$, which seems appropriate for the present data, this choice should be adapted according to the properties of the modelled data (e.g., signal-to-noise ratio, prior low-pass filtering, sampling rate, etc) and should be assessed by the usual model checking tools (e.g., posterior predictive checks, see also @sec-basis).
-To distinguish signal from noise, we also defined a region of practical equivalence [ROPE, @kruschke2017], defined as the chance level plus the standard deviation of the (group-level average) decoding performance during the baseline period.
+To better distinguish signal from noise, we also defined a region of practical equivalence [ROPE, @kruschke2017], defined as the chance level plus the standard deviation of the (group-level average) decoding performance during the baseline period.
 
@@
 Third, we summarised the distribution of onset/offset estimates using the median "error" (i.e., difference between the split estimate and the estimate obtained using the full dataset) and the variance across splits.
-
+This approach allows assessing how similar the estimate of each half split is to the full dataset (thus acting as a proxy for the population) and how variable the estimates are (across split halves).
+
+```{=html}
 <!--
@@
 # importing the reshaped MEG decoding results
-decoding_df <- read.csv(file = "code/decoding_results_reshaped.csv") %>%
+decoding_df <- read.csv(file = "data/decoding_results_reshaped.csv") %>%
     # removing some participants
@@
 
-## Simulation results (bias and variance)
-
-@fig-simulation-mae-variance shows a summary of the simulation results, revealing that the proposed approach (`BGAM`) has the lowest median error and variance for both the onset and offset estimates.
-
-```{r fig-simulation-mae-variance, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Median error and variance of onset and offset estimates for each method.
+## Simulation study (bias and variance)
+
+@fig-simulation-results shows a summary of the simulation results, revealing that the proposed approach (`BGAM`) has the lowest median error and variance for both the onset and offset estimates.
+The `Cluster mass`, `Holm`, and `Change point` also have good performance, but surprisingly, the `TFCE` method has relatively bad performance for estimating the effect offset.
+Unsurprisingly, the `Raw p-value`, `FDR BH95`, and `FDR BY01` methods show the worst performance.
+
+```{r table-simulation-results, eval = FALSE, echo = FALSE}
+# loading the simulation results
+# Here are a couple of example tables (@table-simulation-results)....
+sim_results <- readRDS(file = "results/errors_results_100sims.rds")
+
+# summarising the results
+sim_results_summary <- sim_results %>%
+    # keeping only threshold = 20 for BGAM and threshold = 0.05 for frequentist results
+    filter(threshold %in% c(0.05, 20) ) %>%
+    mutate(
+        error = case_when(
+            grepl("onset", onset_offset) ~ value - true_onset,
+            grepl("offset", onset_offset) ~ value - true_offset
+            )
+        ) %>%
+    group_by(method, onset_offset) %>%
+    summarise(
+        mean_error = mean(error, na.rm = TRUE),
+        median_error = median(error, na.rm = TRUE),
+        median_abs_error = median(abs(error), na.rm = TRUE),
+        rmse = sqrt(mean(error)^2),
+        variance = var(error, na.rm = TRUE),
+        mad = mad(error, na.rm = TRUE),
+        .groups = "drop"
+        ) %>%
+    ungroup() %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c(
+                "raw_p", "pval_bh", "pval_by", "pval_holm",
+                "cluster_mass", "cluster_tfce", "cpt", "brms"
+                ),
+            labels = c(
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+                "TFCE", "Change point", "BGAM (new method)"
+                ),
+            )
+        ) %>%
+    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
+    data.frame()
+
+# creating the table
+# library(tinytable)
+# tt(
+#     x = sim_results_summary,
+#     digits = 2
+#     )
+```
+
+```{r fig-simulation-results, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median error and variance of onset and offset estimates for each method.
+The vertical dashed line represents an unbiased estimate of the true onset/ofsset.
 Variance is plotted on a log10 scale for visual purposes."}
 # loading the simulation results
-sim_results <- readRDS(file = "results/sim_results.rds")
+sim_results <- readRDS(file = "results/errors_results_100sims.rds")
 
 # identifying the best brms results
-# eps=0 and threshold = 10 or 20 seems pretty good...
 # sim_results %>%
-#     select(simulation_id:offset_brms) %>%
+#     filter(method == "brms") %>%
+#     pivot_wider(names_from = onset_offset, values_from = value) %>%
 #     summarise(
-#         MAE_onset = median(abs(onset_brms - true_onset) ),
-#         variance_onset = var(onset_brms),
-#         MAE_offset = median(abs(offset_brms - true_offset) ),
-#         variance_offset = var(offset_brms),
-#         .by = c(eps, threshold)
+#         MAE_onset = median(abs(onset - true_onset) ),
+#         variance_onset = var(onset),
+#         MAE_offset = median(abs(offset - true_offset) ),
+#         variance_offset = var(offset),
+#         .by = threshold
 #         ) %>%
-#     arrange(MAE_onset, MAE_offset) %>%
-#     head(10)
-
-# computing and plotting bias, MAE, and variance
+#     arrange(MAE_onset, MAE_offset)
+
+# computing and plotting bias and variance
 # bias := true_onset - median(sim_onset_distrib)
-# MAE :=  mean(abs(sim_onset-true_onset))
 # variance := variance(sim_onset_distrib)
 sim_results %>%
-    # keeping only the best brms results
-    filter(eps == 0 & threshold == 20) %>%
-    select(-eps,-threshold) %>%
-    # filtering some alpha_level
-    filter(alpha_level == 0.05) %>%
-    select(-alpha_level) %>%
-    pivot_longer(cols = -simulation_id, names_to = "measure", values_to = "value") %>%
-    data.frame() %>%
+    # keeping only threshold = 20 for BGAM and threshold = 0.05 for frequentist results
+    filter(threshold %in% c(0.05, 20) ) %>%
     mutate(
@@
             # grepl("offset", measure) ~ abs(value - true_offset)
-            grepl("onset", measure) ~ value - true_onset,
-            grepl("offset", measure) ~ value - true_offset
-            )
-        ) %>%
-    separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
-    group_by(type, method) %>%
+            grepl("onset", onset_offset) ~ value - true_onset,
+            grepl("offset", onset_offset) ~ value - true_offset
+            )
+        ) %>%
+    group_by(method, onset_offset) %>%
     summarise(
-        MAE = median(error, na.rm = TRUE),
+        median_error = median(error, na.rm = TRUE),
         variance = var(error, na.rm = TRUE),
+        # variance = mad(error, na.rm = TRUE),
         .groups = "drop"
@@
     ungroup() %>%
-    mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
+    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
     mutate(
@@
             levels = c(
-                "p", "bh", "by", "holm", "cluster_mass",
-                "cluster_tfce", "cpt", "brms"
+                "raw_p", "pval_bh", "pval_by", "pval_holm",
+                "cluster_mass", "cluster_tfce", "cpt", "brms"
                 ),
             labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
                 "TFCE", "Change point", "BGAM (new method)"
@@
         aes(
-            x = MAE,
+            x = median_error,
             y = variance,
@@
     scale_y_log10() +
-    facet_wrap(~type) +
+    facet_wrap(~onset_offset) +
     scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
@@
     labs(x = "Median error (s)", y = "Variance (log-scale)")
-```
-
-```{r fig-simulation-distributions, echo = FALSE, eval = FALSE, fig.width = 9, fig.cap = "Distributions of onset and offset estimates for each method."}
-sim_results %>%
-    # keeping only the best brms results
-    filter(eps == 0 & threshold == 20) %>%
-    # filtering some alpha_level
-    filter(alpha_level == 0.05) %>%
-    select(-alpha_level) %>%
-    select(-eps,-threshold) %>%
-    pivot_longer(
-        cols = -simulation_id, 
-        names_to = "measure",
-        values_to = "value"
-        ) %>%
-    separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
-    data.frame() %>%
-    mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
+
+# saving the plot
+ggsave(
+    filename = "figures/simulation_results_median_error_variance.png",
+    width = 12, height = 6, dpi = 300,
+    device = "png"
+    )
+```
+
+\newpage
+
+## Application to actual MEG data (reliability)
+
+@fig-onset-offset shows the group-level average decoding performance through time with onset and offset estimates for each method.
+Overall, this figure shows that both the `Raw p-value` and `FDR BH95` methods are extremely lenient, considering that the decoding performance is above chance before the onset of the stimulus (false positive) and until the end of the trial.
+The `Change point` and `Cluster mass` methods seem the most conservative methods, identifying a time window from approximately +60ms to +500ms.
+The `Holm`, `TFCE`, and `BGAM` methods produce similar estimates of onset and offset, ranging from approximately +60ms to +650ms, although the `BGAM` method seems to result in fewer clusters.[^1]
+
+[^1]: It should be noted that although each method can produce several "clusters" of "significant" timesteps, we only considered the first (onset) and last (offset) timesteps identified by each method to compute the estimation error.
+
+```{r fig-onset-offset, echo = FALSE, fig.width = 8, fig.asp = 0.6, fig.cap = "Group-level average decoding performance through time with onset and offset estimates for each method [data from @nalborczyk:inprep]."}
+# loading the results
+# reliability_results <- readRDS(file = "results/reliability_results.rds")
+# reliability_results <- readRDS(file = "results/reliability_results_cluster.rds")
+reliability_results <- readRDS(file = "results/reliability_results_100splits.rds")
+
+# extracting the full data estimates
+full_reliability_results <- reliability_results %>%
+    filter(split_id == max(split_id) ) %>%
+    # keeping only the first and last timestep
+    # group_by(split_id, method, onset_offset) %>%
+    # summarise(
+    #     value = if (onset_offset[1] == "onset") min(value) else max(value),
+    #     .groups = "drop"
+    #     ) %>%
+    # ungroup() %>%
+    # pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
+    # data.frame() %>%
+    # select(-split_id) %>%
+    # separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
+    pivot_wider(names_from = onset_offset, values_from = value) %>%
     mutate(
@@
             levels = c(
-                "p", "bh", "by", "holm", "cluster_mass",
+                "raw_p", "pval_bh", "pval_by", "pval_holm", "cluster_mass",
                 "cluster_tfce", "cpt", "brms"
@@
             labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm",
-                "Cluster-mass", "TFCE", "Change point", "BGAM (new method)"
-                ),
-            )
-        ) %>%
-    # mutate(
-    #     plot_row = ifelse(
-    #         method %in% c("p-value", "FDR BH95", "FDR BY01", "Holm"), 0, 1
-    #         )
-    #     ) %>%
-    mutate(
-        plot_row = case_when(
-            method %in% c("Raw p-value", "Holm") ~ 1,
-            method %in% c("FDR BH95", "FDR BY01") ~ 2,
-            method %in% c("Cluster-mass", "TFCE") ~ 3,
-            method %in% c("Change point", "brms") ~ 4
-            )
-        ) %>%
-    ggplot(
-        aes(
-            x = value,
-            colour = method,
-            fill = method
-            )
-        ) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = true_offset, linetype = 2) +
-    geom_density(aes(fill = NULL) ) +
-    facet_grid(plot_row~type, scales = "free_x") +
-    theme_light(base_size = 12, base_family = "Open Sans") +
-    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    labs(x = "Onset/Offset (s)", y = "Density")
-```
-
-```{r sim-errors, eval = FALSE, echo = FALSE}
-# plotting the results
-onset_plot <- sim_results %>%
-    select(simulation_id:offset_brms) %>%
-    mutate(
-        error_onset = abs(onset_brms - true_onset),
-        error_offset = abs(offset_brms - true_offset)
-        ) %>%
-    summarise(
-        error_onset = mean(error_onset),
-        error_offset = mean(error_offset),
-        .by = c(eps, threshold)
-        ) %>%
-    # arrange(error_onset) %>%
-    # head(10)
-    ggplot(aes(x = eps, y = threshold, fill = error_onset) ) +
-    geom_tile(show.legend = FALSE) +
-    geom_point(
-        data = . %>% dplyr::filter(error_onset == min(error_onset, na.rm = TRUE) ),
-        aes(x = eps, y = threshold),
-        color = "orangered", size = 2, shape = 4,
-        show.legend = FALSE
-        ) +
-    scale_x_continuous(expand = c(0, 0) ) +
-    scale_y_continuous(expand = c(0, 0) ) +
-    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-    labs(
-        title = "Onset error",
-        x = "eps",
-        y = "Threshold",
-        fill = "Error"
-        )
-
-offset_plot <- sim_results %>%
-    select(simulation_id:offset_brms) %>%
-    mutate(
-        error_onset = abs(onset_brms - true_onset),
-        error_offset = abs(offset_brms - true_offset)
-        ) %>%
-    summarise(
-        error_onset = mean(error_onset),
-        error_offset = mean(error_offset),
-        .by = c(eps, threshold)
-        ) %>%
-    # arrange(error_offset) %>%
-    # head(10)
-    ggplot(aes(x = eps, y = threshold, fill = error_offset) ) +
-    geom_tile(show.legend = FALSE) +
-    geom_point(
-        data = . %>% dplyr::filter(error_offset == min(error_offset, na.rm = TRUE) ),
-        aes(x = eps, y = threshold),
-        color = "orangered", size = 2, shape = 4,
-        show.legend = FALSE
-        ) +
-    scale_x_continuous(expand = c(0, 0) ) +
-    scale_y_continuous(expand = c(0, 0) ) +
-    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-    labs(
-        title = "Offset error",
-        x = "eps",
-        y = "Threshold",
-        fill = "Error"
-        )
-
-# combining the plots
-onset_plot + offset_plot
-```
-
-\newpage
-
-## Application to actual MEG data (reliability)
-
-@fig-onset-offset shows the group-level average decoding performance through time with onset and offset estimates for each method.
-Overall, this figure shows that both the `Raw p-value` and `FDR BH95` methods were extremely lenient, considering that the decoding performance was above chance before the onset of the stimulus (false positive) and until the end of the trial.
-The `Change point` and `Cluster mass` methods were the most conservative methods, identifying a time window from approximately +60ms to +500ms.
-The `Holm`, `TFCE`, and `BGAM` methods produced similar estimates of onset and offset, ranging from approximately +60ms to +650ms.^[It should be noted that although each method can produce several "clusters" of "significant" timesteps, we only considered the first (onset) and last (offset) timesteps identified by each method to compute the error (difference).]
-
-```{r fig-onset-offset, echo = FALSE, fig.width = 9, fig.asp = 0.7, fig.cap = "Group-level average decoding performance through time with onset and offset estimates for each method [data from @nalborczyk:inprep]."}
-# loading the results
-reliability_results <- readRDS(file = "results/reliability_results.rds")
-
-# extract full data estimates
-full_reliability_results <- reliability_results %>%
-    tail(1) %>%
-    pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
-    data.frame() %>%
-    select(-split_id) %>%
-    separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
-    pivot_wider(names_from = type, values_from = value) %>%
-    filter(method != "brms_full") %>%
-    mutate(
-        method = factor(
-            x = method,
-            levels = c(
-                "p", "bh", "by", "holm", "cluster_mass",
-                "cluster_tfce", "cpt", "brms"#, "brms_full"
-                ),
-            labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
-                "TFCE", "Change point", "BGAM (new method)"#, "brms_full"
-                )
-            )
-        ) %>%
-    # arrange(method) %>%
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+                "TFCE", "Change point", "BGAM (new method)"
+                )
+            )
+        ) %>%
     mutate(y_offset = 0.5 - as.numeric(factor(method) ) * 0.01 - 0.01)
@@
         ) %>%
-    # summarise(auc_mean = mean(auc_mean), .by = time) %>%
     ggplot(aes(x = time, y = auc_mean) ) +
@@
         inherit.aes = FALSE,
-        lineend = "round", linewidth = 2
+        lineend = "round",
+        linewidth = 1.5
         ) +
@@
 The `Change point` method also has a very good performance (i.e., very low difference between split estimates and full estimates), but produces too short cluster of significant decoding performance (cf.
-@fig-onset-offset).
-Overall, the figure reveals that for all other methods, split datasets produce later onset estimates and earlier offset estimates (as compared to the estimates from the model fitted on the full dataset).
+@fig-onset-offset).[^2] Overall, the figure reveals that for all other methods, split datasets produce later onset estimates and earlier offset estimates (as compared to the estimates from the model fitted on the full dataset).
 These results highlight that the results of any method should be assessed using both i) good asymptotic properties on simulated data, ii) sensible identified clusters in actual data, and iii) reliable/stable estimates on actual data.
 
-```{r fig-reliability, echo = FALSE, cache = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Median absolute error and variance of onset (left) and offset (right) estimates for each method."}
-# extract full data estimates
+[^2]: As in @rousselet_using_2025, we fixed the number of expected change points to two in the binary segmentation algorithm, thus producing always one cluster.
+
+```{r fig-reliability, echo = FALSE, cache = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median error and median absolute deviation of the error for the onset (left) and offset (right) estimates according to each method.
+Methods are ordered from lowest (top) to highest (bottom) median absolute error (separately for the onset and offset estimates)."}
+# extracting the full data estimates
 full_reliability_results <- reliability_results %>%
-    tail(1) %>%
-    pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
-    # filter(method != "brms_full") %>%
+    filter(split_id == max(split_id) ) %>%
+    # keeping only the first and last timestep
+    group_by(split_id, method, onset_offset) %>%
+    summarise(
+        value = if (onset_offset[1] == "onset") min(value) else max(value),
+        .groups = "drop"
+        ) %>%
+    ungroup() %>%
     data.frame() %>%
@@
 #             labels = c(
-#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
+#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
 #                 "TFCE", "Change point", "brms", "brms_full"
@@
 
-# defining a consistent color palette using original method levels
+# defining a consistent colour palette using original method levels
 original_methods <- c(
-    "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
-    "TFCE", "Change point", "BGAM"#, "BGAM_full"
+    "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+    "TFCE", "Change point", "BGAM (new method)"
     )
@@
 
-# or plotting only MAE in descending order
+# or plotting only the error in descending order
 reliability_results %>%
-    slice(1:nrow(.)-1) %>%
-    pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
+    filter(split_id < max(split_id) ) %>%
     data.frame() %>%
+    # keeping only the first and last timestep
+    group_by(split_id, method, onset_offset) %>%
+    summarise(
+        value = if (onset_offset[1] == "onset") min(value) else max(value),
+        .groups = "drop"
+        ) %>%
+    ungroup() %>%
+    select(-split_id) %>%
     left_join(
         full_reliability_results,
-        by = "measure",
+        # by = "measure",
+        by = c("method", "onset_offset"),
         suffix = c("_split", "_full")
         ) %>%
+    group_by(method, onset_offset) %>%
     # mutate(error = abs(value_split - value_full) ) %>%
     mutate(error = value_split - value_full) %>%
-    # mutate(error = (value_split - value_full)^2) %>%
-    separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
-    group_by(type, method) %>%
     summarise(
-        # MAE = median(error, na.rm = TRUE),
-        MAE = mean(error, na.rm = TRUE),
+        # mean_error = mean(error, na.rm = TRUE),
         # variance = var(error, na.rm = TRUE),
         # variance = var(value_split, na.rm = TRUE)
-        variance = sd(error, na.rm = TRUE) / sqrt(n() )
+        # variance = sd(error, na.rm = TRUE) / sqrt(n() )
         # variance = sum(error, na.rm = TRUE)
+        median_error = median(error, na.rm = TRUE),
+        # variance = IQR(error, na.rm = TRUE)
+        variance = mad(error, na.rm = TRUE)
         # .groups = "drop"
@@
     ungroup() %>%
-    # now plotting
-    mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
-    filter(method != "brms_full") %>%
+    # head()
+    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
     mutate(
@@
             levels = c(
-                "p", "bh", "by", "holm", "cluster_mass",
-                "cluster_tfce", "cpt", "brms"#, "brms_full"
+                "raw_p", "pval_bh", "pval_by", "pval_holm", "cluster_mass",
+                "cluster_tfce", "cpt", "brms"
                 ),
             labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
-                "TFCE", "Change point", "BGAM"#, "BGAM_full"
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+                "TFCE", "Change point", "BGAM (new method)"
                 ),
@@
         ) %>%
-    # group_by(type) %>%
-    # arrange(abs(MAE), .by_group = TRUE) %>%
-    # arrange(abs(MAE), .by_group = TRUE) %>%
-    # ungroup() %>%
-    # mutate(method = forcats::fct_reorder(method, MAE, .desc = ) ) %>%
-    # mutate(method = forcats::fct_reorder2(method, type, MAE, .desc = FALSE) ) %>%
-    # head(20)
     mutate(
         method_reordered = tidytext::reorder_within(
-            x = method, by = 1/abs(MAE), within = type
-            )
-        ) %>%
-    # head()
+            x = method,
+            by = 1 / abs(median_error),
+            within = onset_offset
+            )
+        ) %>%
     ggplot(
         aes(
-            x = MAE,
-            xmin = MAE - variance,
-            xmax = MAE + variance,
-            # y = method,
+            x = median_error,
+            xmin = median_error - variance,
+            xmax = median_error + variance,
             y = method_reordered,
@@
         ) +
-    facet_wrap(~type, scales = "free") +
+    facet_wrap(~onset_offset, scales = "free") +
     scale_y_reordered() +
-    scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
-    # scale_fill_manual(values = color_palette) +
-    # scale_colour_manual(values = color_palette) +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
     labs(
-        x = "Average (median) difference to full dataset (s)",
+        x = "Median difference to full dataset (s)",
         y = "Method"
@@
                 "p", "bh", "by", "holm", "cluster_mass",
-                "cluster_tfce", "cpt", "brms"#, "brms_full"
+                "cluster_tfce", "cpt", "brms"
                 ),
             labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster-mass",
-                "TFCE", "Change point", "BGAM"#, "brms_full"
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+                "TFCE", "Change point", "BGAM"
                 ),
@@
             method %in% c("FDR BH95", "FDR BY01") ~ 2,
-            method %in% c("Cluster-mass", "TFCE") ~ 3,
+            method %in% c("Cluster mass", "TFCE") ~ 3,
             method %in% c("Change point", "BGAM") ~ 4
@@
 
-## Increasing potential usage
-
-Prepare a wrapper `R` package and show how to call it in `Python` and integrate it with `MNE-Python` [@gramfort2013] pipelines...
-
 ## Limitations and future directions
@@
 
-## Conclusions
-
-...
-
 \newpage
@@
 The simulation results as well as the `R` code to reproduce the simulations are available on GitHub: <https://github.com/lnalborczyk/brms_meeg>.
+The `neurogam R` package is available at <https://github.com/lnalborczyk/neurogam>.
 
@@
 # Appendix
-
-````{=html}
-<!--
-
-# Application to time-resolved decoding results (accuracy over time)
-
-We conducted time-resolved multivariate pattern analysis (MVPA), also known as decoding.
-As a result, we have a timecourse of decoding accuracies (e.g., ROC AUC), bounded between 0 and 1, per participant (@fig-mne-decoding)...
-
-```{r fig-sim-decoding, echo = FALSE, eval = TRUE, out.width = "75%", fig.cap = "Exemplary (simulated) group-level average timecourse of binary decoding accuracy (ROC AUC)."}
-# simulating decoding accuracies (ROC AUC)
-n_participants <- 20
-n_timepoints <- 240
-timepoints <- seq(from = 0.1, to = 1.3, length.out = n_timepoints)
-
-# defining a function to generate AUC using a Beta distribution
-simulate_auc <- function (
-        timepoints, onset = 0.1,
-        peak_time = 0.4, sigma = 0.5, amplitude = 30
-        ) {
-
-    # ensuring timepoints - onset > 0
-    adjusted_time <- timepoints - onset + 0.001
-
-    # log-normal peak function with right skew
-    log_peak <- log(peak_time - onset + 0.001)
-    log_t <- log(adjusted_time)
-
-    # computing right-skewed alpha with log-normal peak
-    alpha_t <- 10 + amplitude * exp(-((log_t - log_peak)^2) / (2 * sigma^2) )
-
-    # keeping baseline at chance level (0.5)
-    beta_t <- 10
-
-    # simulate Beta-distributed AUC values
-    auc <- rbeta(length(timepoints), shape1 = alpha_t, shape2 = beta_t)
-
-    # returning the AUC
-    return (auc)
-
-}
-
-# simulating data for all participants
-decoding_data <- data.frame(
-    participant = rep(1:n_participants, each = n_timepoints),
-    time = rep(timepoints, times = n_participants),
-    auc = unlist(lapply(1:n_participants, function (i) {
-        simulate_auc(
-            timepoints,
-            onset = 0.1,
-            # variable peak time
-            peak_time = runif(1, 0.6, 0.7),
-            # variable spread
-            sigma = runif(1, 0.1, 0.3),
-            # variable peak amplitude
-            amplitude = runif(1, 30, 40)
-            )})
-        )
-    ) %>%
-    mutate(time = time - 0.3)
-
-# computing group mean and 95% quantile interval
-group_data <- decoding_data %>%
-    group_by(time) %>%
-    summarise(
-        mean_auc = mean(auc),
-        lower = quantile(auc, 0.025),
-        upper = quantile(auc, 0.975)
-        )
-
-# plotting it
-# group_data %>%
-#     ggplot(aes(x = time, y = mean_auc) ) +
-#     geom_hline(yintercept = 0.5, linetype = "dashed") +
-#     geom_vline(xintercept = 0.0, linetype = "dashed") +
-#     geom_ribbon(aes(ymin = lower, ymax = upper), fill = "steelblue", alpha = 0.2) +
-#     geom_line(color = "steelblue", size = 1) +
-#     ylim(0, 1) +
-#     labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
-```
-
-```{r mne-decoding, eval = FALSE, echo = FALSE}
-# loading the reticulate package
-library(reticulate)
-
-# importing the numpy and mne python modules
-np <- import("numpy")
-mne <- import("mne")
-sklearn <- import("sklearn")
-
-# defining the function in R (it will be executed in Python)
-mne_decoding <- function (X, labels, ncores = 8) {
-
-    # converting R dataframe to NumPy array (reshaping if needed)
-    # X should be a matrix before conversion
-    X_np <- np$array(X)
-
-    if (length(dim(X_np) ) == 2) {
-
-        # adding a second dimension (channels) if missing
-        X_np <- np$expand_dims(X_np, axis = as.integer(1) )
-
-    }
-
-    # defining the classifier
-    clf <- sklearn$linear_model$LogisticRegression(solver = "liblinear")
-
-    # sliding the estimator on all time frames
-    time_decod <- mne$decoding$SlidingEstimator(
-        clf, n_jobs = as.integer(ncores),
-        scoring = "roc_auc", verbose = TRUE
-        )
-
-    # or using N-fold cross-validation
-    scores <- mne$decoding$cross_val_multiscore(
-        time_decod,
-        X_np,
-        labels,
-        cv = as.integer(4),
-        n_jobs = as.integer(ncores),
-        verbose = TRUE
-        )
-
-    # returning the scores (averaged over CV folds)
-    return (scores)
-
-}
-
-# listing all participants
-participants <- unique(raw_df$participant)
-
-# initialising empty decoding results
-group_decoding_scores <- data.frame()
-
-# running decoding for each participant
-for (ppt in participants) {
-
-    # printing progress
-    print(ppt)
-
-    # retrieve data from one participant
-    ppt_data <- raw_df %>%
-        filter(participant == ppt) %>%
-        select(-participant) %>%
-        pivot_wider(names_from = time, values_from = eeg) %>%
-        select(-condition, -trial)
-
-    # extracting the labels
-    labels <- raw_df %>%
-        filter(participant == ppt) %>%
-        select(-participant) %>%
-        pivot_wider(names_from = time, values_from = eeg) %>%
-        pull(condition) %>%
-        as.numeric()
-
-    # extracting the timesteps
-    timesteps <- raw_df %>%
-        filter(participant == ppt) %>%
-        select(-participant) %>%
-        pull(time) %>%
-        unique()
-
-    # running the decoding
-    decoding_scores <- data.frame(
-        mne_decoding(X = ppt_data, labels = labels-1)
-        ) %>%
-        # computing the average over CV folds
-        summarise(across(where(is.numeric), mean) )
-
-    # appending to previous results
-    group_decoding_scores <- bind_rows(group_decoding_scores, decoding_scores)
-
-}
-
-# saving the scores
-saveRDS(object = group_decoding_scores, file = "results/decoding_scores.rds")
-```
-
-```{r fig-mne-decoding, echo = FALSE, fig.width = 9, fig.cap = "Exemplary average timecourse of binary decoding accuracy (ROC AUC) for each participant.
-Group-level average decoding accuracy is depicted as a grey background density in each panel."}
-# importing the decoding scores
-group_decoding_scores <- readRDS(file = "results/decoding_scores.rds")
-
-# extracting the timesteps
-timesteps <- raw_df %>%
-    pull(time) %>%
-    unique()
-
-# plotting it
-decoding_summary <- group_decoding_scores %>%
-    t() %>%
-    data.frame() %>%
-    mutate(
-        mean = rowMeans(across(1:20) ),
-        se = apply(across(1:20), 1, function (x) sd(x) / sqrt(length(x) ) )
-        ) %>%
-    mutate(lower = mean - se, upper = mean + se) %>%
-    mutate(time = timesteps) %>%
-    mutate(
-        ymin = pmin(mean, 0.5),
-        ymax = pmax(mean, 0.5)
-        )
-
-# reshaping the decoding data
-decoding_data <- group_decoding_scores %>%
-    t() %>%
-    data.frame() %>%
-    mutate(time = timesteps) %>%
-    pivot_longer(cols = 1:20) %>%
-    mutate(
-        participant = rep(
-            unique(raw_df$participant),
-            n_distinct(raw_df$time)
-            )
-        ) %>%
-    select(participant, time, auc = value)
-
-# plotting it
-# group_decoding_scores %>%
-#     t() %>%
-#     data.frame() %>%
-#     mutate(time = timesteps) %>%
-#     pivot_longer(cols = 1:20) %>%
-#     mutate(
-#         participant = rep(
-#             unique(raw_df$participant),
-#             n_distinct(raw_df$time)
-#             )
-#         ) %>%
-decoding_data %>%
-    ggplot(aes(x = time) ) +
-    geom_hline(yintercept = 0.5, linetype = "dashed") +
-    geom_vline(xintercept = true_onset, linetype = "dashed") +
-    geom_line(aes(y = auc), linewidth = 0.5) +
-    facet_wrap(~participant) +
-    geom_ribbon(
-        data = decoding_summary,
-        aes(x = time, ymin = ymin, ymax = ymax),
-        alpha = 0.5
-        ) +
-    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
-```
-
-Now, we want to *test* whether the group-level average decoding accuracy is above chance (i.e., 0.5) at each timestep.
-We use a similar GAM/GP as previously, but we replace the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one to account for the bounded nature of AUC values (between 0 and 1).
-
-```{r gam-decoding, message = FALSE}
-# fitting the GAM
-decoding_gam <- brm(
-    auc ~ s(time, bs = "cr", k = 10),
-    data = decoding_data,
-    family = Beta(),
-    iter = 5000,
-    chains = 4,
-    cores = 4,
-    file = "models/decoding_gam.rds"
-    )
-```
-
-```{r gp-decoding, message = FALSE}
-# fitting the GP
-decoding_gp <- brm(
-    auc ~ gp(time, k = 20),
-    data = decoding_data,
-    family = Beta(),
-    control = list(adapt_delta = 0.99),
-    iter = 5000,
-    chains = 4,
-    cores = 4,
-    file = "models/decoding_gp.rds"
-    )
-```
-
-```{r decoding-preds, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Posterior predictions of the GAM (left) and GP (right) fitted on decoding accuracy over time."}
-# plotting the posterior predictions
-gam_decoding_preds <- plot(
-    conditional_effects(x = decoding_gam, prob = 0.99),
-    points = TRUE,
-    point_args = list(size = 1, pch = 21, colour = "grey"),
-    line_args = list(colour = "black"),
-    theme = theme_light(), plot = FALSE
-    )[[1]] +
-    geom_hline(yintercept = 0.5, linetype = 2) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    ylim(0, 1) +
-    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
-
-gp_decoding_preds <- plot(
-    conditional_effects(x = decoding_gp, prob = 0.99),
-    points = TRUE,
-    point_args = list(size = 1, pch = 21, colour = "grey"),
-    line_args = list(colour = "black"),
-    theme = theme_light(), plot = FALSE
-    )[[1]] +
-    geom_hline(yintercept = 0.5, linetype = 2) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    ylim(0, 1) +
-    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
-
-# combining the two plots
-gam_decoding_preds + gp_decoding_preds
-```
-
-Next, we plot the posterior probability of decoding accuracy being above chance level (plus some epsilon) (@fig-decoding-post).
-
-```{r fig-decoding-post, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Posterior probability of decoding accuracy being above chance level according to the GAM (left) or the GP (right)."}
-# defining a function to plot posterior prob of slope above 0
-plot_post_decoding <- function (model, data, chance_level = 0.5, eps = 0.1) {
-
-    # defining a sequence of time values to make predictions
-    time_seq <- crossing(time = seq(min(data$time), max(data$time), length.out = 100) )
-
-    # retrieving posterior samples
-    posterior_samples <- posterior_epred(object = model, newdata = time_seq)
-
-    # computing the probability that y is above 0
-    prob_y_above_0 <- data.frame(
-        time = time_seq$time,
-        m = colMeans(posterior_samples > (chance_level + eps) ),
-        lower = apply(
-            X = posterior_samples > (chance_level + eps),
-            MARGIN = 2, FUN = quantile, probs = 0.025
-            ),
-        upper = apply(
-            X = posterior_samples > (chance_level + eps),
-            MARGIN = 2, FUN = quantile, probs = 0.975
-            )
-        )
-
-    # plotting it
-    prob_y_above_0 %>%
-    ggplot(aes(x = time, y = m) ) +
-    geom_hline(yintercept = chance_level, linetype = 2) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_line() +
-    labs(x = "Time (s)", y = expression(Pr(AUC>chance~"|"~data, model) ) )
-
-}
-
-# using the function with the two models
-plot_post_decoding(
-    model = decoding_gam, data = decoding_data,
-    chance_level = 0.5, eps = 0.1
-    ) +
-    plot_post_decoding(
-        model = decoding_gp, data = decoding_data,
-        chance_level = 0.5, eps = 0.1
-        ) 
-```
-
-```{r fig-decoding-ratio, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Ratio of posterior probabilities of decoding accuracy being above chance level according to the GAM (left) or the GP (right)."}
-# defining a function to plot posterior prob of slope above 0
-plot_post_decoding_ratio <- function (model, data, chance_level = 0.5, eps = 0.1) {
-
-    # defining a sequence of time values to make predictions
-    time_seq <- data.frame(time = data$time)
-
-    # retrieving posterior samples
-    posterior_samples <- posterior_epred(object = model, newdata = time_seq)
-
-    # computing the probability that y is above 0
-    prob_y_above_0 <- data.frame(time = data$time) %>%
-        mutate(m = colMeans(posterior_samples > (chance_level + eps) ) ) %>%
-        mutate(prob_ratio = m / (1 - m) )
-
-    # printing the identified clusters
-    threshold <- 10
-    exceeding_times <- prob_y_above_0 %>%
-        dplyr::filter(prob_ratio > threshold) %>%
-        summarise(cluster_onset = min(time), cluster_offset = max(time) )
-
-    # printing the identified clusters
-    # print(exceeding_times)
-
-    # plotting it
-    prob_y_above_0 %>%
-        mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
-        ggplot(aes(x = time, y = prob_ratio) ) +
-        geom_hline(yintercept = 1, linetype = "dashed") +
-        geom_hline(yintercept = 3, linetype = "dashed", color = "darkred") +
-        geom_hline(yintercept = 1/3, linetype = "dashed", color = "darkred") +
-        geom_hline(yintercept = 10, linetype = "dashed", color = "red") +
-        geom_hline(yintercept = 1/10, linetype = "dashed", color = "red") +
-        geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
-        geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
-        geom_vline(xintercept = true_onset, linetype = 2) +
-        geom_line() +
-        geom_point(
-            data = . %>% dplyr::filter(!is.na(above_thres) ),
-            aes(y = threshold),
-            colour = "darkgreen",
-            shape = 15,
-            size = 3,
-            na.rm = TRUE
-            ) +
-        scale_y_log10(labels = label_log(digits = 2), limits = c(NA, 1e4) ) +
-        labs(
-            x = "Time (s)",
-            y = expression(
-                Pr(AUC>chance~"|"~data) / (1 - Pr(AUC>chance~"|"~data) )
-                )
-            )
-
-}
-
-# using the function with the two models
-plot_post_decoding_ratio(
-    model = decoding_gam, data = decoding_data,
-    chance_level = 0.5, eps = 0.05
-    ) +
-    plot_post_decoding_ratio(
-        model = decoding_gp, data = decoding_data,
-        chance_level = 0.5, eps = 0.05
-        )
-```
-
-\newpage
-
--->
-````
 
@@
     data = raw_df,
+    threshold = 10,
     participant_id = "participant", meeg_id = "eeg",
@@
 For example, we use it below to estimate the onset and offset of effects for one channel from a MNE evoked object.
-Note that the code used to reshape the `sample MNE` dataset is provided later.
+Note that the code used to reshape the `sample MNE` dataset is provided later, and we refer to the [MNE documentation](https://mne.tools/stable/auto_tutorials/epochs/50_epochs_to_data_frame.html) about converting `MNE` epochs to `Pandas` dataframes in long format.
 
@@
     data=reshaped_df_r,
-    post_prob_ratio_threshold=3,
+    threshold=10,
     multilevel=False

### 2025-04-07 brms_meeg.tex
updating the simulation code (now faster and more parallelised)

@@
 reflect how neural responses unfold in response to stimuli or tasks. A
-central goal in many M/EEG studies is to determine whether---and
-when---neural responses differ across experimental conditions or
-participant groups.
+central goal in many M/EEG studies is to determine whether and when
+neural responses differ across experimental conditions or participant
+groups.
 
@@
 \citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
-\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al., 2019};
 \citeproc{ref-pernet2015}{Pernet et al., 2015})\ldots{} See also Maris
@@
 
-\subsection{Previous work}\label{previous-work}
+\subsection{Related work}\label{related-work}
 
@@
 functions \(f_{j}\) are represented in the model via penalised splines
-basis expansions of the covariates, that are a weighted sum of basis
-functions:
+basis expansions of the covariates, that are a weighted sum of \(K\)
+simpler, basis functions:
 
@@
 \(k\)th basis function \(b_{jk}()\) evaluated at the covariate value
-\(x_{ij}\) for the \(j\)th smooth function \(f_{j}\). Splines
-coefficients are penalised (usually through the squared of the smooth
-functions' second derivative) in a way that can be interpreted, in
-Bayesian terms, as a prior on the ``wiggliness'' of the function. In
-other words, more complex (wiggly) basis function are penalised. To
-clarify the terminology at this stage: \emph{splines} are functions
-composed of simpler functions. These simpler functions are basis
-functions (e.g., a cubic polynomial) and the set of basis functions is a
-\emph{basis}. Each basis function gets its coefficient and the resultant
-spline is the sum of these weighted basis functions.
+\(x_{ij}\) for the \(j\)th smooth function \(f_{j}\). To clarify the
+terminology at this stage: \emph{splines} are functions composed of
+simpler functions. These simpler functions are basis functions (e.g., a
+cubic polynomial) and the set of basis functions is a \emph{basis}. Each
+basis function gets its coefficient and the resultant spline is the sum
+of these weighted basis functions. Splines coefficients are penalised
+(usually through the squared of the smooth functions' second derivative)
+in a way that can be interpreted, in Bayesian terms, as a prior on the
+``wiggliness'' of the function. In other words, more complex (wiggly)
+basis function are penalised.
 
@@
 models}\label{bayesian-generalised-additive-multilevel-models}
+
+Introduction to multilevel GAMs
+(\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al.,
+2019})\ldots{}
 
@@
 
-Focusing on identifying onset and offset of effects (as assessed by ERP
-differences or decoding performance)\ldots{} Assessing the performance
-of a model-based approach (i.e., Bayesian GAMMs) to conventional methods
-(multiplicity corrections or cluster-based permutation)\ldots{}
+Given the previously reported limitations of conventional methods to
+precisely identify the onset and offset of M/EEG effects (e.g., ERPs,
+decoding performance), we aimed to provide a model-based approach for
+estimating the onset and offset of such effects. To achieve this, we
+leveraged Bayesian generalised additive multilevel models (BGAMMs)
+fitted via the \texttt{brms\ R} package and compared the performance of
+this approach to conventional methods on both simulated and actual M/EEG
+data.
 
@@
 superimposing 50 sinusoids at different frequencies, following an
-EEG-like spectrum (see details and code in
-\citeproc{ref-yeung2004}{Yeung et al., 2004}). As in Rousselet
-(\citeproc{ref-rousselet_using_2025}{2025}), the signal was generated
-from truncated Gaussian with an objective onset at 160 ms, a peak at 250
-ms, and an offset at 342 ms. We simulated this signal for 250 timesteps
-between 0 and 0.5s, akin to a 500 Hz sampling rate. We simulated such
-data for a group of 20 participants with 50 trials per participant and
-condition (Figure~\ref{fig-eeg}).
+EEG-like spectrum (see code in the online supplementary materials and
+details in \citeproc{ref-yeung2004}{Yeung et al., 2004}). As in
+Rousselet (\citeproc{ref-rousselet_using_2025}{2025}), the signal was
+generated from truncated Gaussian with an objective onset at 160 ms, a
+peak at 250 ms, and an offset at 342 ms. We simulated this signal for
+250 timesteps between 0 and 0.5s, akin to a 500 Hz sampling rate. We
+simulated data for a group of 20 participants with 50 trials per
+participant and condition (Figure~\ref{fig-eeg}).
 
@@
 
-\caption{\label{fig-eeg}Averaged (mean) simulated EEG activity in two
-conditions with 50 trials each, for a group of 20 participants. The
-error band represents the mean plus/minus 1 standard error of the mean.}
+\caption{\label{fig-eeg}Mean simulated EEG activity in two conditions
+with 50 trials each, for a group of 20 participants. The error band
+represents the mean +/- 1 standard error of the mean.}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-eeg-1.pdf}
-
-}
-
-\end{figure}%
-
-We computed the average of the ERP difference
-(Figure~\ref{fig-erp})\ldots{}
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-erp}Group-level average difference between
-conditions (mean +/- standard error of the mean). The `true' onset and
-offset are indicated by the vertical dashed lines.}
-
-\centering{
-
-\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-erp-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-eeg-1.png}
 
@@
 assessed examining trace plots as well as the Gelman--Rubin statistic
-\(\hat{R}\). The \texttt{brms} package uses the same syntax as the
-\texttt{R} package \texttt{mgcv} v 1.9-1 (\citeproc{ref-mgcv}{Wood,
-2017b}) for specifying smooth effects. Figure~\ref{fig-plot-post-slope}
-shows the predictions of this model together with the raw data.
+\(\hat{R}\) (\citeproc{ref-gabry2019}{Gabry et al., 2019};
+\citeproc{ref-gelman2020}{Gelman et al., 2020}). The \texttt{brms}
+package uses the same syntax as the \texttt{R} package \texttt{mgcv} v
+1.9-1 (\citeproc{ref-mgcv}{Wood, 2017b}) for specifying smooth effects.
+Figure~\ref{fig-plot-post-slope} shows the predictions of this model
+together with the raw data.
 
@@
 
-\subsection{Posterior probability of difference above
-0}\label{posterior-probability-of-difference-above-0}
-
-We then plot the posterior predictions together with the posterior
-estimate of the slope for \texttt{condition} at each timestep
+We depict the posterior predictions together with the posterior estimate
+of the slope for \texttt{condition} at each timestep
 (Figure~\ref{fig-plot-post-slope}).
@@
 
-\caption{\label{fig-plot-post-slope}Posterior estimate of the ERP in
-each condition (left) or directly for the difference of ERPs (right)
-according to the GAM.}
+\caption{\label{fig-plot-post-slope}Posterior estimate of the EEG
+activity in each condition (left) and posterior estimate of the
+difference in EEG activity (right) according to the GAM.}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-plot-post-slope-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-plot-post-slope-1.png}
 
@@
 We then compute the posterior probability of the slope for
-\texttt{condition} being above \(0+\epsilon\)
-(Figure~\ref{fig-post-prob-ratio}, left), with \(\epsilon := 0.05\),
-which can be interpreted as the smallest effect size of interest. We can
-also express this as the ratio of posterior probabilities (i.e.,
-\(p/(1-p)\)) and visualise the timecourse of this ratio superimposed
-with the conventional thresholds on evidence ratios
-(Figure~\ref{fig-post-prob-ratio}, right).
+\texttt{condition} being above \(0\) (Figure~\ref{fig-post-prob-ratio},
+left). We can also express this as the ratio of posterior probabilities
+(i.e., \(p/(1-p)\)) and visualise the timecourse of this ratio
+superimposed with the conventional thresholds on evidence ratios
+(Figure~\ref{fig-post-prob-ratio}, right). A ratio of 10 means that the
+probability of the difference being above 0 is 10 times higher than the
+probability of the difference not being above 0, given the data, the
+priors, and other model's assumptions.
 
@@
 \caption{\label{fig-post-prob-ratio}Left: Posterior probability of the
-ERP difference (slope) being above 0 according to the GAM. Right: Ratio
+EEG difference (slope) being above 0 according to the GAM. Right: Ratio
 of posterior probability according to the GAM (on a log10 scale).
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.png}
 
@@
 
-\subsection{Multilevel modelling using ERP summary
-statistics}\label{multilevel-modelling-using-erp-summary-statistics}
-
-Next we fit a multilevel version of the GAM. Although it is possible to
-fit a GAMM at the single-trial level, we present a resource-efficient
-version of the model that is fitted directly on the by-participant
-summary statistics (mean and SD, similar to what is done in
-meta-analysis\ldots{}
+\subsection{Multilevel modelling using summary
+statistics}\label{multilevel-modelling-using-summary-statistics}
+
+The previous model only included constant (fixed) effects, thus not
+properly accounting for between-participant variability. We next fit a
+multilevel version of the GAM. Although it is possible to fit a GAMM at
+the single-trial level, we present a resource-efficient version of the
+model that is fitted directly on the by-participant summary statistics
+(mean and SD), similar to what is done in meta-analysis.
 
@@
 
-We then computed the difference between the true and estimated
-onset/offset of the ERP difference
-(\(\text{error}:=|\hat{\theta}-\theta|\)), according to various
-\texttt{k} (GAM basis dimension) and \texttt{threshold} values. Remember
-that the signal is generated from a truncated Gaussian with an objective
-onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
-Figure~\ref{fig-onset-error} shows that the hierarchical GAM can
-\emph{exactly} recover the true onset and offset values, given some
-reasonable choice of \texttt{k} and \texttt{threshold} values (e.g., a
-threshold of 20). We provide more detailed recommendations on how to set
-\texttt{k} in Section~\ref{sec-basis}.
+We then compute the difference between the true and estimated
+onset/offset of the EEG difference according to various \texttt{k} (GAM
+basis dimension) and \texttt{threshold} values. Remember that the EEG
+signal was generated from a truncated Gaussian with an objective onset
+at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
+Figure~\ref{fig-onset-error} shows that the multilevel GAM can
+\emph{almost exactly} recover the true onset and offset values, given
+some reasonable choice of \texttt{k} and \texttt{threshold} values
+(e.g., a threshold of 20). We provide more detailed recommendations on
+how to set \texttt{k} in Section~\ref{sec-basis}.
 
@@
 
-\caption{\label{fig-onset-error}Error function of onset (left) and
-offset (right) estimation according to various basis dimension and
+\caption{\label{fig-onset-error}Average estimation error for the onset
+(left) and offset (right) according to various basis dimension and
 threshold values (according to the multilevel GAM).}
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-error-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-error-1.png}
 
@@
 
-To provide a more exhaustive assessment of the performance of the
-proposed approach, we compared the ability of the GAMM to correctly
-estimate the onset and offset of the ERP difference to widely-used
-methods. First, we conducted mass-univariate t-tests (thus treating each
-timestep independently) and identified the onset and offset of the ERP
-difference as the first and last values crossing an arbitrary
-significance threshold (\(\alpha = 0.05\)). We then followed the same
-approach but after applying different forms of multiplicity correction
-to the p-values. We compared two methods that control the false
-discovery rate (FDR) (i.e., BH95, \citeproc{ref-benjamini1995}{Benjamini
-\& Hochberg, 1995}; and BY01, \citeproc{ref-benjamini2001}{Benjamini \&
-Yekutieli, 2001}), one method that controls the family-wise error rate
-(FWER) (i.e., Holm--Bonferroni method, \citeproc{ref-holm1979}{Holm,
-1979}), and two cluster-based permutation methods (permutation with a
-single cluster-forming threshold and threshold-free cluster enhancement,
-TFCE, \citeproc{ref-smith2009}{S. Smith \& Nichols, 2009}). The BH95,
-BY01, and Holm corrections were applied to the p-values using the
+We then compared the ability of the GAMM to correctly estimate the onset
+and offset of the ERP difference to other widely-used methods. First, we
+conducted mass-univariate t-tests (thus treating each timestep
+independently) and identified the onset and offset of the ERP difference
+as the first and last values crossing an arbitrary significance
+threshold (\(\alpha = 0.05\)). We then followed the same approach but
+after applying different forms of multiplicity correction to the
+p-values. We compared two methods that control the false discovery rate
+(FDR) (i.e., BH95, \citeproc{ref-benjamini1995}{Benjamini \& Hochberg,
+1995}; and BY01, \citeproc{ref-benjamini2001}{Benjamini \& Yekutieli,
+2001}), one method that controls the family-wise error rate (FWER)
+(i.e., Holm--Bonferroni method, \citeproc{ref-holm1979}{Holm, 1979}),
+and two cluster-based permutation methods (permutation with a single
+cluster-forming threshold and threshold-free cluster enhancement, TFCE,
+\citeproc{ref-smith2009}{S. Smith \& Nichols, 2009}). The BH95, BY01,
+and Holm corrections were applied to the p-values using the
 \texttt{p.adjust()} function in \texttt{R}. The cluster-based inference
@@
 \texttt{R} package \texttt{changepoint} v 2.3
-(\citeproc{ref-changepoint}{Killick et al., 2022a}), and applied
-directly to the squared t-values (as in
+(\citeproc{ref-changepoint}{Killick et al., 2022}), and applied directly
+to the squared t-values (as in
 \citeproc{ref-rousselet_using_2025}{Rousselet, 2025}).
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-corrections-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-corrections-1.png}
 
@@
 
-The various methods were assessed using the median error and variance of
-onset/offset estimates computed on 10.000 simulated datasets\ldots{}
+The various methods were compared using the median error (i.e.,
+true-estimated) and variance of onset/offset estimates computed on
+10.000 simulated datasets.
 
@@
 
-Next, we assessed the performance of the proposed approach on actual MEG
-data (decoding results from \citeproc{ref-nalborczyk:inprep}{Nalborczyk
-et al., in preparation}). We conducted time-resolved multivariate
-pattern analysis (MVPA), also known as decoding\ldots{} As a result, we
-have a timecourse of decoding performance (ROC AUC), bounded between 0
-and 1, for each participant (for a total of 32 participants). Now, we
-want to \emph{test} whether the group-level average decoding accuracy is
-above chance (i.e., 0.5) at each timestep
-(Figure~\ref{fig-decoding-data}). We fitted a similar GAM as discussed
-previously, but we replaced the \(\mathrm{Normal}\) likelihood function
-by a \(\mathrm{Beta}\) one to account for the bounded nature of AUC
-values (between 0 and 1) (for a tutorial on Beta regression, see
-\citeproc{ref-coretta2025}{Coretta \& Bürkner, 2025}). Note that
-although we chose a basis dimension of \(k=50\), which seems appropriate
-for the present data, this choice should be adapted according to the
-properties of the modelled data (e.g., signal-to-noise ratio, prior
-low-pass filtering, sampling rate, etc) and should be assessed by the
-usual model checking tools (e.g., posterior predictive checks, see also
-Section~\ref{sec-basis}). To distinguish signal from noise, we also
-defined a region of practical equivalence (ROPE,
+To provide a more exhaustive assessment of these methods, we also
+evaluated the performance of the proposed approach on actual MEG data
+(decoding results from \citeproc{ref-nalborczyk:inprep}{Nalborczyk et
+al., in preparation}). In this study, we conducted time-resolved
+multivariate pattern analysis (MVPA, also known as decoding) of MEG data
+during reading tasks. As a result, we obtain a timecourse of decoding
+performance (ROC AUC), bounded between 0 and 1, for each participant
+(for a total of 32 participants). Next, we want to \emph{test} whether
+the group-level average decoding accuracy is above chance (i.e., 0.5) at
+each timestep (Figure~\ref{fig-decoding-data}). To achieve this, we
+fitted a similar GAM as discussed previously, but we replaced the
+\(\mathrm{Normal}\) likelihood function by a \(\mathrm{Beta}\) one to
+account for the bounded nature of AUC values (between 0 and 1) (for a
+tutorial on Beta regression, see \citeproc{ref-coretta2025}{Coretta \&
+Bürkner, 2025}).
+
+Note that although we chose a basis dimension of \(k=50\), which seems
+appropriate for the present data, this choice should be adapted
+according to the properties of the modelled data (e.g., signal-to-noise
+ratio, prior low-pass filtering, sampling rate, etc) and should be
+assessed by the usual model checking tools (e.g., posterior predictive
+checks, see also Section~\ref{sec-basis}). To better distinguish signal
+from noise, we also defined a region of practical equivalence (ROPE,
 \citeproc{ref-kruschke2017}{Kruschke \& Liddell, 2017}), defined as the
@@
 between the split estimate and the estimate obtained using the full
-dataset) and the variance across splits.
+dataset) and the variance across splits. This approach allows assessing
+how similar the estimate of each half split is to the full dataset (thus
+acting as a proxy for the population) and how variable the estimates are
+(across split halves).
 
@@
 
-\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-data-1.pdf}
+\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-data-1.png}
 
@@
 
-\subsection{Simulation results (bias and
-variance)}\label{simulation-results-bias-and-variance}
-
-Figure~\ref{fig-simulation-mae-variance} shows a summary of the
-simulation results, revealing that the proposed approach (\texttt{BGAM})
-has the lowest median error and variance for both the onset and offset
-estimates.
+\subsection{Simulation study (bias and
+variance)}\label{simulation-study-bias-and-variance}
+
+Figure~\ref{fig-simulation-results} shows a summary of the simulation
+results, revealing that the proposed approach (\texttt{BGAM}) has the
+lowest median error and variance for both the onset and offset
+estimates. The \texttt{Cluster-mass}, \texttt{Holm}, and
+\texttt{Change\ point} also have good performance, but surprisingly, the
+\texttt{TFCE} method has relatively bad performance for estimating the
+effect offset. Unsurprisingly, the \texttt{Raw\ p-value},
+\texttt{FDR\ BH95}, and \texttt{FDR\ BY01} methods show the worst
+performance.
 
@@
 
-\caption{\label{fig-simulation-mae-variance}Median error and variance of
-onset and offset estimates for each method. Variance is plotted on a
-log10 scale for visual purposes.}
+\caption{\label{fig-simulation-results}Median error and variance of
+onset and offset estimates for each method. The vertical dashed line
+represents an unbiased estimate of the true onset/ofsset. Variance is
+plotted on a log10 scale for visual purposes.}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-simulation-mae-variance-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-simulation-results-1.png}
 
@@
 method. Overall, this figure shows that both the \texttt{Raw\ p-value}
-and \texttt{FDR\ BH95} methods were extremely lenient, considering that
-the decoding performance was above chance before the onset of the
+and \texttt{FDR\ BH95} methods are extremely lenient, considering that
+the decoding performance is above chance before the onset of the
 stimulus (false positive) and until the end of the trial. The
-\texttt{Change\ point} and \texttt{Cluster\ mass} methods were the most
+\texttt{Change\ point} and \texttt{Cluster\ mass} methods seem the most
 conservative methods, identifying a time window from approximately +60ms
 to +500ms. The \texttt{Holm}, \texttt{TFCE}, and \texttt{BGAM} methods
-produced similar estimates of onset and offset, ranging from
-approximately +60ms to +650ms.\footnote{It should be noted that although
+produce similar estimates of onset and offset, ranging from
+approximately +60ms to +650ms, although the \texttt{BGAM} method seems
+to result in fewer clusters.\footnote{It should be noted that although
   each method can produce several ``clusters'' of ``significant''
   timesteps, we only considered the first (onset) and last (offset)
-  timesteps identified by each method to compute the error (difference).}
+  timesteps identified by each method to compute the estimation error.}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-offset-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-offset-1.png}
 
@@
 produces too short cluster of significant decoding performance (cf.
-Figure~\ref{fig-onset-offset}). Overall, the figure reveals that for all
-other methods, split datasets produce later onset estimates and earlier
-offset estimates (as compared to the estimates from the model fitted on
-the full dataset). These results highlight that the results of any
-method should be assessed using both i) good asymptotic properties on
-simulated data, ii) sensible identified clusters in actual data, and
+Figure~\ref{fig-onset-offset}).\footnote{As in Rousselet
+  (\citeproc{ref-rousselet_using_2025}{2025}), we fixed the number of
+  expected change points to two in the binary segmentation algorithm,
+  thus producing always one cluster.} Overall, the figure reveals that
+for all other methods, split datasets produce later onset estimates and
+earlier offset estimates (as compared to the estimates from the model
+fitted on the full dataset). These results highlight that the results of
+any method should be assessed using both i) good asymptotic properties
+on simulated data, ii) sensible identified clusters in actual data, and
 iii) reliable/stable estimates on actual data.
@@
 
-\caption{\label{fig-reliability}Median absolute error and variance of
-onset (left) and offset (right) estimates for each method.}
+\caption{\label{fig-reliability}Median error and median absolute
+deviation of the error for the onset (left) and offset (right) estimates
+according to each method. Methods are ordered from lowest (top) to
+highest (bottom) median absolute error (separately for the onset and
+offset estimates).}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-reliability-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-reliability-1.png}
 
@@
 etc)\ldots{}
-
-\subsection{Increasing potential
-usage}\label{increasing-potential-usage}
-
-Prepare a wrapper \texttt{R} package and show how to call it in
-\texttt{Python} and integrate it with \texttt{MNE-Python}
-(\citeproc{ref-gramfort2013}{Gramfort, 2013}) pipelines\ldots{}
 
@@
 
-\subsection{Conclusions}\label{conclusions}
-
-\ldots{}
-
 \newpage
@@
 simulations are available on GitHub:
-\url{https://github.com/lnalborczyk/brms_meeg}.
+\url{https://github.com/lnalborczyk/brms_meeg}. The \texttt{neurogam\ R}
+package is available at \url{https://github.com/lnalborczyk/neurogam}.
 
@@
 
-We used R version 4.2.3 (\citeproc{ref-base}{R Core Team, 2023}) and the
-following R packages: brms v. 2.22.0 (\citeproc{ref-brms2017}{Bürkner,
-2017}, \citeproc{ref-brms2018}{2018}, \citeproc{ref-brms2021}{2021}),
-changepoint v. 2.2.4 (\citeproc{ref-changepoint2022}{Killick et al.,
-2022b}; \citeproc{ref-changepoint2014}{Killick \& Eckley, 2014}),
-grateful v. 0.2.10 (\citeproc{ref-grateful}{Rodriguez-Sanchez \&
-Jackson, 2023}), knitr v. 1.45 (\citeproc{ref-knitr2014}{Xie, 2014},
-\citeproc{ref-knitr2015}{2015}, \citeproc{ref-knitr2023}{2023}),
-MetBrewer v. 0.2.0 (\citeproc{ref-MetBrewer}{Mills, 2022}), pakret v.
-0.2.2 (\citeproc{ref-pakret}{Gallou, 2024}), patchwork v. 1.2.0
+We used R version 4.4.3 (\citeproc{ref-base}{R Core Team, 2025}) and the
+following R packages: assertthat v. 0.2.1
+(\citeproc{ref-assertthat}{Wickham, 2019}), brms v. 2.22.0
+(\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018},
+\citeproc{ref-brms2021}{2021}), changepoint v. 2.3
+(\citeproc{ref-changepoint2024}{Killick et al., 2024};
+\citeproc{ref-changepoint2014}{Killick \& Eckley, 2014}), doParallel v.
+1.0.17 (\citeproc{ref-doParallel}{Corporation \& Weston, 2022}),
+easystats v. 0.7.4 (\citeproc{ref-easystats}{Lüdecke et al., 2022}),
+foreach v. 1.5.2 (\citeproc{ref-foreach}{Microsoft \& Weston, 2022}),
+furrr v. 0.3.1 (\citeproc{ref-furrr}{Vaughan \& Dancho, 2022}), future
+v. 1.34.0 (\citeproc{ref-RJ-2021-048}{Bengtsson, 2021}), ggrepel v.
+0.9.6 (\citeproc{ref-ggrepel}{Slowikowski, 2024}), glue v. 1.8.0
+(\citeproc{ref-glue}{Hester \& Bryan, 2024}), grateful v. 0.2.11
+(\citeproc{ref-grateful}{Rodriguez-Sanchez \& Jackson, 2024}), knitr v.
+1.50 (\citeproc{ref-knitr2014}{Xie, 2014},
+\citeproc{ref-knitr2015}{2015}, \citeproc{ref-knitr2025}{2025}),
+MetBrewer v. 0.2.0 (\citeproc{ref-MetBrewer}{Mills, 2022}), neurogam v.
+0.0.1 (\citeproc{ref-neurogam}{Nalborczyk, 2025}), pakret v. 0.2.2
+(\citeproc{ref-pakret}{Gallou, 2024}), patchwork v. 1.3.0
 (\citeproc{ref-patchwork}{T. L. Pedersen, 2024}), rmarkdown v. 2.29
@@
 (\citeproc{ref-scico}{T. L. Pedersen \& Crameri, 2023}), tidybayes v.
-3.0.6 (\citeproc{ref-tidybayes}{Kay, 2023}), tidyverse v. 2.0.0
+3.0.7 (\citeproc{ref-tidybayes}{Kay, 2024}), tidytext v. 0.4.2
+(\citeproc{ref-tidytext}{Silge \& Robinson, 2016}), tidyverse v. 2.0.0
 (\citeproc{ref-tidyverse}{Wickham et al., 2019}).
@@
 
+\bibitem[\citeproctext]{ref-RJ-2021-048}
+Bengtsson, H. (2021). A unifying framework for parallel and distributed
+processing in r using futures. \emph{The R Journal}, \emph{13}(2),
+208--227. \url{https://doi.org/10.32614/RJ-2021-048}
+
 \bibitem[\citeproctext]{ref-benjamini1995}
@@
 
+\bibitem[\citeproctext]{ref-doParallel}
+Corporation, M., \& Weston, S. (2022). \emph{{doParallel}: Foreach
+parallel adaptor for the {``{parallel}''} package}.
+\url{https://CRAN.R-project.org/package=doParallel}
+
 \bibitem[\citeproctext]{ref-dimigen2021}
@@
 
+\bibitem[\citeproctext]{ref-gabry2019}
+Gabry, J., Simpson, D., Vehtari, A., Betancourt, M., \& Gelman, A.
+(2019). Visualization in Bayesian work{fl}ow. \emph{Journal of the Royal
+Statistical Society: Series A (Statistics in Society)}, \emph{182}(2),
+389--402. \url{https://doi.org/10.1111/rssa.12378}
+
 \bibitem[\citeproctext]{ref-pakret}
@@
 \url{https://CRAN.R-project.org/package=pakret}
+
+\bibitem[\citeproctext]{ref-gelman2020}
+Gelman, A., Vehtari, A., Simpson, D., Margossian, C. C., Carpenter, B.,
+Yao, Y., Kennedy, L., Gabry, J., Bürkner, P.-C., \& Modrák, M. (2020).
+Bayesian workflow. \emph{arXiv:2011.01808 {[}Stat{]}}.
+\url{http://arxiv.org/abs/2011.01808}
 
@@
 
+\bibitem[\citeproctext]{ref-glue}
+Hester, J., \& Bryan, J. (2024). \emph{{glue}: Interpreted string
+literals}. \url{https://CRAN.R-project.org/package=glue}
+
 \bibitem[\citeproctext]{ref-holm1979}
@@
 \bibitem[\citeproctext]{ref-tidybayes}
-Kay, M. (2023). \emph{{tidybayes}: Tidy data and geoms for {Bayesian}
+Kay, M. (2024). \emph{{tidybayes}: Tidy data and geoms for {Bayesian}
 models}. \url{https://doi.org/10.5281/zenodo.1308151}
@@
 \bibitem[\citeproctext]{ref-changepoint}
-Killick, R., Haynes, K., \& Eckley, I. A. (2022a). \emph{{changepoint}:
+Killick, R., Haynes, K., \& Eckley, I. A. (2022). \emph{{changepoint}:
 An {R} package for changepoint analysis}.
@@
 
-\bibitem[\citeproctext]{ref-changepoint2022}
-Killick, R., Haynes, K., \& Eckley, I. A. (2022b). \emph{{changepoint}:
+\bibitem[\citeproctext]{ref-changepoint2024}
+Killick, R., Haynes, K., \& Eckley, I. A. (2024). \emph{{changepoint}:
 An {R} package for changepoint analysis}.
@@
 
+\bibitem[\citeproctext]{ref-easystats}
+Lüdecke, D., Ben-Shachar, M. S., Patil, I., Wiernik, B. M., Bacher, E.,
+Thériault, R., \& Makowski, D. (2022). {easystats}: Framework for easy
+statistical modeling, visualization, and reporting. \emph{CRAN}.
+\url{https://doi.org/10.32614/CRAN.package.easystats}
+
 \bibitem[\citeproctext]{ref-maris2011}
@@
 
+\bibitem[\citeproctext]{ref-foreach}
+Microsoft, \& Weston, S. (2022). \emph{{foreach}: Provides foreach
+looping construct}. \url{https://CRAN.R-project.org/package=foreach}
+
 \bibitem[\citeproctext]{ref-miller2025}
@@
 \url{https://CRAN.R-project.org/package=MetBrewer}
+
+\bibitem[\citeproctext]{ref-neurogam}
+Nalborczyk, L. (2025). \emph{{neurogam}: Precise temporal localisation
+of m/EEG effects with bayesian generalised additive multilevel models}.
 
@@
 \bibitem[\citeproctext]{ref-base}
-R Core Team. (2023). \emph{{R}: A language and environment for
+R Core Team. (2025). \emph{{R}: A language and environment for
 statistical computing}. R Foundation for Statistical Computing.
@@
 \bibitem[\citeproctext]{ref-grateful}
-Rodriguez-Sanchez, F., \& Jackson, C. P. (2023). \emph{{grateful}:
-Facilitate citation of r packages}.
+Rodriguez-Sanchez, F., \& Jackson, C. P. (2024). \emph{{grateful}:
+Facilitate citation of {R} packages}.
 \url{https://pakillo.github.io/grateful/}
@@
 
+\bibitem[\citeproctext]{ref-tidytext}
+Silge, J., \& Robinson, D. (2016). {tidytext}: Text mining and analysis
+using tidy data principles in r. \emph{JOSS}, \emph{1}(3).
+\url{https://doi.org/10.21105/joss.00037}
+
 \bibitem[\citeproctext]{ref-skukies_modelling_2021}
@@
 challenges}. \url{https://doi.org/10.1101/2024.12.05.626938}
+
+\bibitem[\citeproctext]{ref-ggrepel}
+Slowikowski, K. (2024). \emph{{ggrepel}: Automatically position
+non-overlapping text labels with {``{ggplot2}''}}.
+\url{https://CRAN.R-project.org/package=ggrepel}
 
@@
 to 'python'}. \url{https://CRAN.R-project.org/package=reticulate}
+
+\bibitem[\citeproctext]{ref-furrr}
+Vaughan, D., \& Dancho, M. (2022). \emph{{furrr}: Apply mapping
+functions in parallel using futures}.
+\url{https://CRAN.R-project.org/package=furrr}
+
+\bibitem[\citeproctext]{ref-assertthat}
+Wickham, H. (2019). \emph{{assertthat}: Easy pre and post assertions}.
+\url{https://CRAN.R-project.org/package=assertthat}
 
@@
 
-\bibitem[\citeproctext]{ref-knitr2023}
-Xie, Y. (2023). \emph{{knitr}: A general-purpose package for dynamic
-report generation in r}. \url{https://yihui.org/knitr/}
+\bibitem[\citeproctext]{ref-knitr2025}
+Xie, Y. (2025). \emph{{knitr}: A general-purpose package for dynamic
+report generation in {R}}. \url{https://yihui.org/knitr/}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-timegen-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-timegen-1.png}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-timegen-post-preds-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-timegen-post-preds-1.png}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-choose-k-1.pdf}
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-choose-k-1.png}
 
@@
     \AttributeTok{data =}\NormalTok{ raw\_df,}
+    \AttributeTok{threshold =} \DecValTok{10}\NormalTok{,}
     \AttributeTok{participant\_id =} \StringTok{"participant"}\NormalTok{, }\AttributeTok{meeg\_id =} \StringTok{"eeg"}\NormalTok{,}
@@
 object. Note that the code used to reshape the \texttt{sample\ MNE}
-dataset is provided later.
+dataset is provided later, and we refer to the
+\href{https://mne.tools/stable/auto_tutorials/epochs/50_epochs_to_data_frame.html}{MNE
+documentation} about converting \texttt{MNE} epochs to \texttt{Pandas}
+dataframes in long format.
 
@@
 \NormalTok{    data}\OperatorTok{=}\NormalTok{reshaped\_df\_r,}
-\NormalTok{    post\_prob\_ratio\_threshold}\OperatorTok{=}\DecValTok{3}\NormalTok{,}
+\NormalTok{    threshold}\OperatorTok{=}\DecValTok{10}\NormalTok{,}
 \NormalTok{    multilevel}\OperatorTok{=}\VariableTok{False}

### 2025-04-07 code/errors_simulation_cluster.R

@@ -2,14 +2,15 @@
 # Monte-Carlo simulation of onset/offset error properties #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on March 31, 2025                          #
+# Last updated on April 7, 2025                           #
 ###########################################################
 
 # importing R packages
 library(changepoint)
 library(reticulate)
-library(tidyverse)
 library(tidybayes)
+library(dplyr)
+library(tidyr)
 library(brms)
 
 # for setting up the cluster
@@ -17,10 +18,10 @@ library(doParallel)
 library(foreach)
 
 # total number of total CPU cores available on the HPC node
-total_cores <- 64
+total_cores <- 12
 
 # number of cores per brms model
-cores_per_model <- 8
+cores_per_model <- 4
 
 # number of parallel simulations to run at once
 n_parallel <- total_cores / cores_per_model
@@ -29,27 +30,23 @@ n_parallel <- total_cores / cores_per_model
 cl <- makeCluster(n_parallel)
 registerDoParallel(cl)
 
-# importing Python modules
-# use_condaenv("/Users/ladislas/opt/anaconda3/envs/r-reticulate/bin/python", required = TRUE)
-# use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
-use_condaenv("r-reticulate", required = TRUE)
-np <- import("numpy")
-mne <- import("mne")
-
 # importing R version of Matlab code from Yeung et al. (2004)
 source("code/eeg_noise.R")
 
-# importing the ERP template with true onset = 160 ms, F=81, and max at F=126
+# importing the ERP template with true onset = 160 ms, F = 81, and max at F = 126
 source("code/erp_template.R")
 
-# importing helper functions
-source("code/functions.R")
+# importing helper functions from Guillaume Rouselet
+# source("code/functions.R")
+
+# importing home-made helper functions
+source("code/utils.R")
 
 # to use with the eeg_noise function
 meanpower <- unlist(read.table("code/meanpower.txt") )
 
-# defining the number of simulations to be performed
-nsims <- 1e4
+# defining the total number of simulations to be performed
+nsims <- 100
 
 # defining simulation parameters
 n_trials <- 50 # number of trials
@@ -65,425 +62,282 @@ true_offset <- 0.342
 # defining significance (alpha) thresholds
 significance_thresholds <- c(0.05, 0.01, 0.005, 0.001)
 
[…]

### 2025-04-07 code/functions.R

@@ -1,219 +1,0 @@
-######################################################################################
-# Functions from https://github.com/GRousselet/onsetsim/blob/main/code/functions.R
-# Generate permutation distributions of t values for independent groups
-# cond1 and cond2 are Nt trials x Nf time frames matrices
-######################################################################################
-
-# permtdist <- function(cond1, cond2, Nt, Nf, nboot = 2000){
-#   
-#   st2 <- Nt+1 # index to start condition 2 
-#   en2 <- Nt+Nt # index to end condition 2
-#   all <- rbind(cond1, cond2)
-#   
-#   perm.tvals <- matrix(data = 0, nrow = nboot, ncol = Nf)
-#   # perm.pvals <- matrix(data = 0, nrow = nboot, ncol = Nf)
-#   
-#   for(B in 1:nboot){
-#     
-#     permdata <- all[sample(1:en2, size = en2, replace = FALSE),]
-#     permcond1 <- permdata[1:Nt,]
-#     permcond2 <- permdata[st2:en2,]
-#     
-#     perm.tvals[B,] <- Rfast::ttests(permcond1, permcond2, paired=FALSE)[,1]    
-#     # for(F in 1:Nf){ # for each time frame:
-#     #   perm.tvals[B,F] <- t.test(permcond1[,F], permcond2[,F])$statistic
-#     # }
-#   }
-#   # list(tvals = perm.tvals, pvals = perm.pvals)
-#   perm.tvals
-# }
-# 
-# # Generate permutation distributions of t values for one group
-# ## diff is a matrix of pairwise differences Np participants x Nf time frames matrices
-# permtdist.onesamp <- function(diff, Np, Nf, nboot = 2000){
-#   
-#   perm.tvals <- matrix(data = 0, nrow = nboot, ncol = Nf)
-#   permdiff <- diff
-#   
-#   for(B in 1:nboot){
-#     
-#     for(P in 1:Np){
-#       newsign <- sample(c(-1,1),1) # flip a coin
-#       permdiff[P,] <- diff[P,] * newsign 
-#     }
-#     
-#     for(F in 1:Nf){ # for each time frame:
-#       perm.tvals[B,F] <- t.test(permdiff[,F])$statistic
-#     }
-#     
-#   }
-#   perm.tvals
-# }
-
-# Cluster correction
[…]

### 2025-04-07 code/generate_eeg_data.R

@@ -5,7 +5,7 @@ source("code/eeg_noise.R")
 source("code/erp_template.R")
 
 # importing helper functions
-source("code/functions.R")
+# source("code/functions.R")
 
 # to use with the eeg_noise function
 meanpower <- unlist(read.table("code/meanpower.txt") )

### 2025-04-07 code/meta_gam_errors_cluster.R

@@ -1,0 +1,184 @@
+###########################################################
+# Monte-Carlo simulation of onset/offset error properties #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on April 7, 2025                           #
+###########################################################
+
+# importing R packages
+library(tidybayes)
+library(dplyr)
+library(tidyr)
+library(brms)
+
+# setting up parallel workers
+library(purrr)
+library(furrr)
+
+# total number of total CPU cores available on the HPC node
+total_cores <- 12
+
+# number of cores per brms model
+cores_per_model <- 4
+
+# number of parallel simulations to run at once
+n_parallel <- total_cores / cores_per_model
+
+# setting up parallel workers
+future::plan(future::multisession(workers = n_parallel) )
+
+# number of simulations to run per combination
+nsims <- 100
+
+# importing the data generation functions 
+source(file = "code/generate_eeg_data.R")
+source(file = "code/utils.R")
+
+# defining a function to run a single iteration of the simulation
+single_iteration <- function (i) {
+    
+    # sanity check
+    system(
+        sprintf(
+            "echo 'Running simulation %d on PID %d with %s cores' >> ./sim_log.txt",
+            i, Sys.getpid(), cores_per_model
+            )
+        )
+    
+    # initialising an empty dataframe
+    temp_results <- data.frame()
+    
+    # number of posterior samples to use to compute the posterior probability ratio
+    n_post_samples <- 1e4
+    
+    # defining the parameter grid
+    param_grid <- crossing(
+        kvalue = c(5, 10, 15, 20, 25, 30, 35, 40),
+        threshold = seq(from = 1, to = 50, by = 1)
+        )
+    
+    for (j in seq_len(nrow(param_grid) ) ) {
+        
+        # retrieving the current kvalue and threshold
+        kvalue <- param_grid$kvalue[j]
+        threshold <- param_grid$threshold[j]
+        
+        # printing progress
+        message_parallel(sprintf("Running sim %d, k = %d, threshold = %d", i, kvalue, threshold) )
+        
+        # when we assess a new sim_id, we generate new data
+        if (kvalue == 5 & threshold == 1) {
+    
+            # generate some data
+            message_parallel("\nSimulating new data...\n")
+            raw_df <- generate_data(
+                n_trials = n_trials, n_ppt = n_ppt, outvar = outvar,
+                srate = srate, ronset = ronset
+                )
+    
[…]

### 2025-04-07 code/old/cluster_reticulate_test.R

@@ -1,0 +1,25 @@
+library(reticulate)
+
+# use_virtualenv("./reticulate", required = TRUE)
+py_require(packages = c("numpy", "mne"), python_version = "3.12.0")
+np <- import("numpy")
+mne <- import("mne")
+X <- np$random$randn(as.integer(10), as.integer(50) )
+mne$set_log_level("WARNING") # or "ERROR"
+
+results <- tryCatch({
+    mne$stats$spatio_temporal_cluster_1samp_test(
+        X = X,
+        out_type = "mask",
+        threshold = NULL,
+        n_permutations = as.integer(2^12),
+        n_jobs = as.integer(1),
+        verbose = TRUE
+        )
+    }, error = function(e) {
+    message("Error: ", e$message)
+    reticulate::py_last_error()
+    return(NULL)
+    })
+
+print(results)

### 2025-04-07 code/old/functions.R

@@ -1,0 +1,173 @@
+######################################################################################
+# Functions from https://github.com/GRousselet/onsetsim/blob/main/code/functions.R
+# Generate permutation distributions of t values for independent groups
+# cond1 and cond2 are Nt trials x Nf time frames matrices
+######################################################################################
+
+# Cluster correction
+# Form clusters using binary vector: pvals < alpha
+# clusters must be at least 2 time frames
+cluster.make <- function (x) {
+    
+    y <- rle(x)
+    cmap <- vector(mode = "numeric", length = 0)
+    nC <- length(y$values) # number of clusters
+    indx <- 0 # cluster counter
+    
+    for (CL in 1:nC) {
+        
+    if(y$values[CL] == 0 || y$lengths[CL] == 1){
+      val <- 0
+    } else {
+      indx <- indx + 1
+      val <- indx
+    }
+    cmap <- c(cmap, rep(val, y$lengths[CL]))
+    }
+    
+    cmap
+    
+}
+
+# ## Save sum for each cluster
+# # values = statistics (t values)
+# # cmap = output from cluster.make
+# cluster.sum <- function(values, cmap){
+#   csum <- vector(mode = "numeric", length = max(cmap))
+#   if(max(cmap)>0){
+#     for(CL in 1:max(cmap)){
+#       csum[CL] <- sum(values[cmap==CL])
+#     }
+#   } else {
+#     csum <- 0
+#   }
+#   csum
+# }
+# 
+# # Cluster test
+# # values = statistics (t values)
+# # cmap = output from cluster.make
+# # boot.th = bootstrap quantile of distribution of max cluster sums
+# cluster.test <- function(values, cmap, boot.th){
+#   csig <- vector(mode = "logical", length = length(cmap))
+#   if(max(cmap)>0){
+#     for(CL in 1:max(cmap)){
+#       csig[cmap==CL] <- sum(values[cmap==CL]) > boot.th
+#     }
+#   }
+#   csig
+# }
+
+sim.counter <- function(S, nsim, inc){
+  if(S == 1){
+    # print(paste(nsim,"iterations:",S))
+    cat(nsim,"iterations:",S)
+    #beep(2)
+  }
+  if(S %% inc == 0){
[…]

### 2025-04-07 code/split_half_reliability_cluster.R

@@ -1,0 +1,493 @@
+###########################################################
+# Split-half reliability of onset/offset estimates        #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on April 7, 2025                           #
+###########################################################
+
+# importing R packages
+library(changepoint)
+library(reticulate)
+library(tidybayes)
+library(dplyr)
+library(tidyr)
+library(brms)
+
+# for setting up the cluster
+library(doParallel)
+library(foreach)
+
+# total number of total CPU cores available on the HPC node
+total_cores <- 12
+
+# number of cores per brms model
+cores_per_model <- 4
+
+# number of parallel simulations to run at once
+n_parallel <- total_cores / cores_per_model
+
+# registering parallel backend
+cl <- makeCluster(n_parallel)
+registerDoParallel(cl)
+
+# importing home-made helper functions
+source("code/utils.R")
+
+# importing the reshaped MEG decoding results
+raw_df <- read.csv(file = "data/decoding_results_reshaped.csv") |>
+    # removing the last participant (to get a round number of participants)
+    mutate(participant_id = cur_group_id(), .by = participant) |>
+    filter(participant_id < 33) |>
+    select(-participant_id)
+
+# averaging across participants
+summary_df <- raw_df |>
+    summarise(
+        auc_mean = mean(auc),
+        auc_sd = sd(auc),
+        auc_se = sd(auc) / sqrt(n() ),
+        .by = c(participant, time)
+        )
+
+# computing the smallest effect size of interest (SESOI)
+# as the SD of decoding performance during the baseline
+full_sesoi <- summary_df %>%
+    filter(time < 0) %>%
+    summarise(auc_mean = mean(auc_mean), .by = time) %>%
+    summarise(sesoi = sd(auc_mean) ) %>%
+    pull(sesoi)
+
+# getting unique participants
+unique_participants <- unique(raw_df$participant)
+
+# getting the number of participants
+n_participants <- as.numeric(length(unique_participants) )
+
+# defining the number of splits
+n_splits <- 100
+
+# number of posterior samples to use to compute the posterior probability ratio
+n_post_samples <- 1e4
+
+# creating a dataframe where each participant appears in every split
+split_df <- crossing(participant = unique_participants, split_number = 1:n_splits)
+
+# assigning alternating groups (A or B) within each split for each participant
+split_df <- split_df |>
[…]

### 2025-04-07 code/utils.R

@@ -1,0 +1,237 @@
+###########################################################
+# Defining some utility functions                         #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on April 4, 2025                           #
+###########################################################
+
+# defining a function to generate a dataframe
+generate_data <- function (n_trials, n_ppt, outvar, srate, ronset) {
+    
+    for (P in 1:n_ppt) { # for each participant
+        
+        # get random onset
+        ponset <- sample(x = ronset, size = 1)
+        
+        # find starting point
+        st <- which(Xf == ponset)
+        
+        # pad vector
+        temp2 <- c(rep(0, st - 2), erp, rep(0, Nf - st - length(erp) + 2) )
+        
+        # initialising empty conditions
+        cond1 <- matrix(0, nrow = n_trials, ncol = Nf)
+        cond2 <- matrix(0, nrow = n_trials, ncol = Nf)
+        
+        for (trial in 1:n_trials) { # for each trial
+            
+            cond1[trial, ] <- temp1 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
+            cond2[trial, ] <- temp2 + eeg_noise(frames = Nf, srate = srate, outvar = outvar, meanpower)
+            
+        }
+        
+        # converting results to dataframe
+        temp_df <- data.frame(
+            x = rep(Xf, 2 * nrow(cond1) ),
+            y = c(c(t(cond1) ), c(t(cond2) ) ),
+            trial = c(rep(1:n_trials, each = length(Xf) ), rep(1:n_trials, each = length(Xf) ) ),
+            condition = factor(rep(c("cond1", "cond2"), each = Nf * n_trials) ),
+            participant = paste0("participant_", sprintf("%02d", P) )
+            ) %>%
+            select(participant, condition, trial, x, y)
+        
+        # and appending it to previous participants
+        if (exists("raw_data") ) {
+            
+            raw_data <- bind_rows(raw_data, temp_df)
+            
+        } else {
+            
+            raw_data <- temp_df
+            
+        }
+        
+    }
+    
+    # converting time from ms to seconds
+    raw_data <- raw_data %>%
+        # converting time from ms to seconds
+        mutate(x = x / 1000) %>%
+        # renaming columns
+        rename(time = x, eeg = y)
+    
+    # returning the simulated data
+    return (raw_data)
+    
+}
+
[…]

### 2025-04-10 brms_meeg.qmd
updating the manuscript (intro)

@@
 Researchers are often interested in determining whether and when these signals differ across experimental conditions or participant groups.
-The conventional approach involves mass-univariate statistical testing across time, followed by corrections for multiple comparisons (e.g., FDR, FWER) or cluster-based inference.
-While effective for controlling error rates, cluster-based methods come with a significant limitation: they shift the focus of inference from individual time points to clusters, making it difficult to draw precise conclusions about the onset or offset of observed effects.
-Here, we introduce a *model-based* approch for analysing M/EEG timeseries such as event-related potentials (ERPs) or decoding performance over time.
+The conventional approach involves mass-univariate statistical testing across time and space, followed by corrections for multiple comparisons such as cluster-based inference.
+While effective for controlling error rates at the cluster-level, cluster-based inference comes with a significant limitation: by shifting the focus of inference from individual time points to clusters, it makes difficult to draw precise conclusions about the onset or offset of observed effects.
+Here, we introduce a *model-based* approach for analysing M/EEG timeseries such as event-related potentials (ERPs) or decoding performance over time.
 Our method leverages Bayesian generalised additive multilevel models (GAMMs), providing posterior probabilities that an effect is above zero (or above chance) at each time point, while naturally accounting for temporal dependencies and between-subject variability.
 Using both simulated and actual M/EEG datasets, we demonstrate that this approach substantially outperforms conventional methods in estimating the onset and offset of neural effects, yielding more precise and reliable results.
-We also provide an R package implementing the method and show how it can be seamlessly integrated into M/EEG analysis pipelines using MNE-Python."
-keywords: [EEG, MEG, generalised additive models, mixed-effects models, multilevel models, Bayesian statistics, brms]
+We provide an R package implementing the method and show how it can be seamlessly integrated into M/EEG analysis pipelines using MNE-Python."
+keywords: [EEG, MEG, cluster-based inference, simulation, multiple comparisons, generalised additive models, mixed-effects models, multilevel models, Bayesian statistics, brms]
 floatsintext: true
@@
 format:
-  # apaquarto-html: default
   apaquarto-pdf:
@@
           \usepackage{mathtools}
+          \usepackage{booktabs, caption, longtable, colortbl, array}
 ---
@@
 library(brms)
+library(mgcv)
+library(gt)
 
@@
     fig.pos = "!htb",
-    # dev = "cairo_pdf",
     dev = "png",
@@
 # setting a default ggplot2 theme
-# theme_set(theme_light(base_size = 10, base_family = "Open Sans") )
 theme_set(theme_bw(base_size = 10, base_family = "Open Sans") )
@@
 # template for citing R packages
-# pkrt_set(pkg = "the `R` package `:pkg` [:ref]")
 pkrt_set(pkg = "the `R` package `:pkg` v :ver [:ref]")
@@
 
+## Problem statement
+
 Understanding the temporal dynamics of cognitive processes requires methods that can capture fast-changing neural activity with high temporal resolution.
-Magnetoencephalography (MEG) and electroencephalography (EEG) are two such methods, widely used in cognitive neuroscience for their ability to track brain activity at the millisecond scale.
+Magnetoencephalography and electroencephalography (M/EEG) are two such methods, widely used in cognitive neuroscience for their ability to track brain activity at the millisecond scale.
 These techniques provide rich time series data that reflect how neural responses unfold in response to stimuli or tasks.
-A central goal in many M/EEG studies is to determine whether and when neural responses differ across experimental conditions or participant groups.
-
-Here are some useful references to be discussed [@combrisson_exceeding_2015; @ehinger_unfold_2019; @hayasaka_validating_2003; @luck_how_2017; @frossard2021; @frossard2022; @gramfort2013; @maris2007; @pernet2015]...
-See also @maris2011 and @rosenblatt2018 (history of cluster-based approaches)...
-Clusters failures [@eklund2016]...
-
-## Problem statement
-
+A central goal in many M/EEG studies is to determine whether, when, and where neural responses differ across experimental conditions or participant groups.
+
+The conventional approach involves mass-univariate statistical testing with a unique threshold, followed by corrections for multiple comparisons with the goal of maintaining the familywise error rate (FWER) or false discovery rate (FDR) at the nominal level.
+Cluster-based inference is the most common way of achieving this sort of error control in the M/EEG literature, being the default option in several software programs.
+While effective for controlling error rates, cluster-based inference comes with a significant limitation: they shift the focus of inference from individual time points to clusters, making it difficult to draw precise conclusions about the onset or offset of observed effects [@maris2007; @sassenhagen2019].
+As pointed by @maris2007; "there is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA \[false alarm\] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect".
+Even worse, as noted by @rosenblatt2018, cluster-based inference suffers from low spatial resolution: "Since discovering a cluster means that 'there exists at least one voxel with an evoked response in the cluster', and not that 'all the voxels in the cluster have an evoked response', it follows that the larger the detected cluster, the less information we have on the location of the activation." As a consequence, cluster-based inferenc has been shown...
+
+To overcome the limitations of cluster-based inference, we introduce a novel model-based approach for precisely localising M/EEG effects in time (space, and other dimensions).
+The proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the posterior probability of effects being above chance at the level of timesteps, sensors, voxels, etc., while naturally taking into account spatiotemporal dependencies present in M/EEG timeseries.
+We compare the performance of the proposed approach to well-established alternative methods using both simulated and actual M/EEG data and show that it significantly outperforms alternative methods in estimating the onset and offset of M/EEG effects.
+
+## Cluster-based inference
+
+Different methods exist to control the family-wiser error rate (FWER), defined as the type-1 error rate (false positive) over an ensemble (family) of tests...
+for instance the Bonferroni correction [@dunn1961], however this method is generally overconservative as it assumes statistical independence of tests, an assumptions that is clearly violated in the context of M/EEG timeseries with spatiotemporal dependencies...
+or the false discovery rate (FDR), defined as the proportion of false positive among positive tests [e.g., @benjamini1995; @benjamini2001]...
+
+<!--
+
+Here are some useful references to be discussed [@ehinger_unfold_2019; @luck_how_2017; @frossard2021; @frossard2022]...
+
+-->
+
+On popular technique to account for spatiotemporal dependencies while controlling the FWER is cluster-based inference.
 Description of cluster-based approaches [see @sassenhagen2019]...
-
-However, as pointed by the original authors themselves [@maris2007], "there is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA \[false alarm\] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect" [@maris2007; @sassenhagen2019].
-In contrast, the proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the posterior probability of effects being above chance at the level of timesteps, voxels, sensors (etc), while naturally taking into account spatiotemporal dependencies present in M/EEG timeseries.
-
-## Related work
+A typical cluster-based inference consists of two successive steps.
+First, clusters are defined as sets of contiguous voxels, channels, and/or timesteps, whose intensity/activity exceeds some predefined threshold.
+Clusters are then characterised by their height (i.e., maximal value), extent (number of constituent elements), or some combination of both, such at its "mass", for instance by summing the statistics within a cluster, an approach refereed to as "cluster mass" [@maris2007; @pernet2015].
+Then, the null hypothesis is tested by assessing the probability..
+As different cluster-forming thresholds lead to clusters with different spatial or temporal extent, this initial threshold modulates the sensitivity of the subsequent permutation test.
+The threshold-free cluster enhancement method (TFCE) was introduced by @smith2009 to overcome this choice of an arbitrary threshold.
+
+In brief, the TFCE method works as follows.
+Instead of picking an arbitrary cluster-forming threshold (e.g., $t=2$), we try all (or many) possible thresholds in a given range and check whether a given timestep/voxel belongs to a significant cluster under any of the set of thresholds.
+Then, instead of using cluster mass (e.g., the sum of squared t-values within the cluster), we use a weighted average between the cluster extend ($e$, how broad is the cluster, that is, how many connected samples it contains) and the cluster height ($h$, how high is the cluster, that is, how large is the test statistic).
+The TFCE score at each timestep/voxel $t$ is given by:
+
+$$
+\text{TFCE(t)} = \int_{h=h_{0}}^{h_{t}} e(h)^{E} h^{H} \mathrm{d}h
+$$
+
+where $h_{0}$ is typically $0$ and parameters $E$ and $H$ are set a priori (typically to 0.5 and 2, respectively) and control the influence of the extend and height on the TFCE.
+Then, p-value for timestep/voxel $t$ is computed by comparing it TFCE with the null distribution of TFCE values.
+For each permuted signal, we keep the maximal value over the whole signal for the null distribution of the TFCE.
+The TFCE combined with permutation (assuming a large enough number of permutations) has been shown to provide accurate type 1 FWER [e.g., @pernet2015].
+However, previous simulation work showed that cluster-based methods have indeed poor performance in localising the onset of M/EEG effects [e.g., @rousselet_using_2025; @sassenhagen2019].
+
+To sum up this section, cluster-based inference main limitations is that it provides inference at the cluster level only, not allowing inference at the level of timesteps, sensors, etc.
+Thus, it does not allow inferring the localisation of effects [for more details on cluster-based inference, see for instance @maris2007; @maris2011; @sassenhagen2019].
+In the following, we briefly review previous modelling work.
+Then, we provide a brief introduction to Bayesian statistical modelling, before moving to generalised additive models (GAMs) and generalised additive multilevel models (GAMMs) to illustrate how they can be used to precisely localise the onset and offset of M/EEG effects.
+
+## Previous work on modelling M/EEG data
 
@@
 
+As put by @rousselet_using_2025, concluding on the onset of effect based on a series of univariate tests...
+commits to three fallacies...
+here, we want to avoid these by introducing a model-based approach, which naturally take into account the temporal dependencies in the data to output a series of posterior probabilities...
+
 See also the rERP framework [@smith2014a; @smith2014b] and @tremblay2014...
@@
 Recently, @teichmann2022 provided a detailed tutorial on using Bayes factors (BFs) to analyse the 1D or 2D output from MVPA, that is, for testing, at every timestep, whether decoding performance is above chance level.
-However, this approach provides timeseries of BFs that ignores temporal dependencies..
-
-## Bayesian regression modelling
-
-Short intro/recap about Bayesian (linear and generalised) regression models...
+However, this approach provides timeseries of BFs that ignores temporal dependencies...
 
@@
 To clarify the terminology at this stage: *splines* are functions composed of simpler functions.
-These simpler functions are basis functions (e.g., a cubic polynomial) and the set of basis functions is a *basis*.
-Each basis function gets its coefficient and the resultant spline is the sum of these weighted basis functions.
+These simpler functions are basis functions (e.g., cubic polynomial, thin-plate) and the set of basis functions is a *basis*.
+Each basis function gets its coefficient and the resultant spline is the sum of these weighted basis functions (@fig-intro-gam).
 Splines coefficients are penalised (usually through the squared of the smooth functions' second derivative) in a way that can be interpreted, in Bayesian terms, as a prior on the "wiggliness" of the function.
@@
 
+The Bayesian approach to statistical modelling is characterised by its reliance on probability theory to [@gelman2020]...
+In this framework, all unknown entities are assigned probability distributions reflecting the uncertainty...
+These probability distributions are commonly referred to as "priors" and represent some state of knowledge about unknown quantities before seeing any data.
+There are debates among Bayesian practitioners as to whether prior distributions should encore subjective (personal) beliefs or...
+but these debates are outside the scope of the present paper and we therefore the interested reader to dedicated work (e.g., XX; YY)...
+In practice, weakly informative priors are often used as default priors in situations in which subjective priors are difficult to define/elicit...
+Bayesian models are then fitted on empirical (actual or simulated) data to update prior states of knowledge to posterior states of knowledge using Bayes theorem, or in practise, sampling-based approximations of the posterior distribution...
+
+```{r fig-intro-gam, echo = FALSE, fig.width = 9, fig.cap = "Different types of GAM(M)s.
+**A**: GAMs predictions are computed as the weigthed sum (in black) of basis functions (here thin-plate basis functions, in grey).
+**B**: Constant-effect GAM, with 5 participants in colours and the group-level prediction in black.
+**C**: Varying-intercept + varying-slope GAMM (with common smoother).
+**D**: Varying-intercept + varying-slope + varying-smoother GAMM.
+In this model, each participant gets its own intercept, slope, and degree of 'wiggliness' (smoother)."}
+########################
+# simulating som edata #
+########################
+
+set.seed(666)
+n_subjects <- 5
+n_per_subject <- 50
+subject <- rep(paste0("S", 1:n_subjects), each = n_per_subject)
+x <- rep(seq(0, 10, length.out = n_per_subject), times = n_subjects)
+
+# subject-level varying effects
+intercepts <- rnorm(n_subjects, 0, 1)
+slopes <- rnorm(n_subjects, 0, 0.3)
+wiggles <- rnorm(n_subjects, 0, 0.2)
+subject_df <- tibble(
+    subject = paste0("S", 1:n_subjects),
+    intercept = intercepts,
+    slope = slopes,
+    wiggle = wiggles
+    )
+
+df <- tibble(x = x, subject = subject) %>%
+    left_join(subject_df, by = "subject") %>%
+    mutate(
+        subject = as.factor(subject),
+        y = sin(x) + intercept + slope * x + wiggle * sin(2 * x) + rnorm(n(), sd = 0.3)
+        )
+
+######################
+# fitting the models #
+######################
+
+# fixed-effect GAM
+gam_fixed <- gam(
+    y ~ s(x),
+    data = df
+    )
+
+# random intercept + slope
+gam_int_slope <- gam(
+    y ~ s(x) +
+        s(subject, bs = "re") +
+        s(x, subject, bs = "re"),
+    data = df
+    )
+
+# random smooth per subject + global smooth
+gam_rand_smooth <- gam(
+    y ~ s(x) +
+        s(subject, bs = "re") +
+        s(x, subject, bs = "fs", m = 1),
+    data = df
+    )
+
+# defining plotting functions
+plot_basis_functions <- function (df, model) {
+
+    # retrieving design matrix and coefficients
+    Xp <- predict(model, type = "lpmatrix")
+    coefs <- coef(model)
+
+    # basis dataframe
+    basis_df <- as.data.frame(Xp)
+    colnames(basis_df) <- paste0("b", 1:ncol(basis_df) )
+    basis_df$x <- df$x
+    basis_long <- pivot_longer(basis_df, starts_with("b"), names_to = "basis", values_to = "value")
+
+    # matching basis to weights
+    weights <- data.frame(
+        basis = paste0("b", seq_along(coefs) ),
+        weight = coefs
+        ) %>%
+        mutate(label = paste0(basis, " × ", sprintf("%.2f", weight) ) )
+
+    # adding labels to the end of each basis line
+    label_df <- basis_long %>%
+        group_by(basis) %>%
+        slice_tail(n = 1) %>% # get last x point per basis
+        left_join(weights, by = "basis") %>%
+        mutate(label = paste0(basis, " × ", sprintf("%.2f", weight) ) )
+
+    # fitted curve at the group level
+    df$fit <- as.vector(Xp %*% coefs)
+
+    # plotting everything
+    ggplot() +
+        geom_hline(yintercept = 0, linetype = 2) +
+        geom_line(
+            # data = basis_long, aes(x = x, y = value, color = basis),
+            data = basis_long, aes(x = x, y = value, group = basis),
+            colour = "grey50",
+            alpha = 0.5,
+            linewidth = 0.5
+            ) +
+        geom_line(
+            data = df,
+            aes(x = x, y = fit),
+            color = "black",
+            linewidth = 1.0
+            ) +
+        # geom_text(
+        #     data = label_df,
+        #     aes(x = x, y = value, label = label, color = basis),
+        #     hjust = 0, nudge_x = 0.1, size = 2
+        #     ) +
+        # ggrepel::geom_text_repel(
+        #     data = label_df,
+        #     aes(x = x, y = value, label = label, color = basis),
+        #     hjust = 0, nudge_x = 1,
+        #     size = 2, direction = "y"
+        #     ) +
+        labs(x = "Predictor", y = "Response") +
+        ylim(-2, +2) +
+        # scale_colour_manual(values = met.brewer(name = "Johnson", n = 10) ) +
+        theme(legend.position = "none")
+
+    }
+
+plot_model <- function (df, model) {
+
+    # predictions fo the full model: includes all terms (fixed + random)
+    df$fit_subject <- predict(model, newdata = df)
+
+    # identify which smooth terms to exclude (those involving 'subject')
+    smooth_labels <- sapply(model$smooth, function(s) s$label)
+    subject_terms <- smooth_labels[grepl("subject", smooth_labels)]
+
+    # group-level prediction + standard error (excludes subject-specific terms)
+    pred_global <- predict(model, newdata = df, exclude = subject_terms, se.fit = TRUE)
+    df$fit_global <- pred_global$fit
+    df$fit_global_upper <- pred_global$fit + pred_global$se.fit
+    df$fit_global_lower <- pred_global$fit - pred_global$se.fit
+
+    # plotting the predictions
+    df %>%
+        ggplot(aes(x = x, y = y, colour = subject) ) +
+        geom_hline(yintercept = 0, linetype = 2) +
+        geom_point(
+            size = 0.5,
+            show.legend = FALSE
+            ) +
+        geom_line(
+            aes(y = fit_subject, group = subject),
+            alpha = 0.5,
+            linewidth = 0.5,
+            show.legend = FALSE
+            ) +
+        geom_ribbon(
+            aes(x = x, y = y, ymin = fit_global_lower, ymax = fit_global_upper),
+            fill = "black",
+            alpha = 0.2,
+            inherit.aes = FALSE
+            ) +
+        geom_line(
+            aes(y = fit_global, group = 1),
+            color = "black",
+            linewidth = 1
+            ) +
+        ylim(-5, 5) +
+        scale_colour_manual(values = met.brewer(name = "Johnson", n = 5) ) +
+        labs(x = "Predictor", y = "Response")
+
+}
+
+# plotting, note that the basis plot uses the fixed-effect model for simplicity
+p1 <- plot_basis_functions(df, gam_fixed)
+p2 <- plot_model(df, gam_fixed)
+p3 <- plot_model(df, gam_int_slope)
+p4 <- plot_model(df, gam_rand_smooth)
+
+# combining plots with labels
+(p1 | p2) / (p3 | p4) + plot_annotation(tag_levels = "A")
+```
+
+See @fig-intro-gam...
 Introduction to multilevel GAMs [@pedersen_hierarchical_2019]...
-
 Now describe the Bayesian GAMM [@miller2025]...
@@
 
-Given the previously reported limitations of conventional methods to precisely identify the onset and offset of M/EEG effects (e.g., ERPs, decoding performance), we aimed to provide a model-based approach for estimating the onset and offset of such effects.
-To achieve this, we leveraged Bayesian generalised additive multilevel models (BGAMMs) fitted via the `brms R` package and compared the performance of this approach to conventional methods on both simulated and actual M/EEG data.
+Given the previously reported limitations of conventional methods to precisely identify the onset and offset of M/EEG effects (e.g., ERPs, decoding performance), we developed a model-based approach for estimating the onset and offset of such effects.
+To achieve this, we leveraged Bayesian generalised additive multilevel models (BGAMMs) fitted in `R` via the `brms` package and compared the performance of this approach to conventional methods on both simulated and actual M/EEG data.
 
@@
 As in previous studies, the noise was generated by superimposing 50 sinusoids at different frequencies, following an EEG-like spectrum [see code in the online supplementary materials and details in @yeung2004].
-As in @rousselet_using_2025, the signal was generated from truncated Gaussian with an objective onset at 160 ms, a peak at 250 ms, and an offset at 342 ms.
+As in @rousselet_using_2025, the signal was generated from a truncated Gaussian distribution with an objective onset at 160 ms, a peak at 250 ms, and an offset at 342 ms.
 We simulated this signal for 250 timesteps between 0 and 0.5s, akin to a 500 Hz sampling rate.
-We simulated data for a group of 20 participants with 50 trials per participant and condition (@fig-eeg).
+We simulated data for a group of 20 participants (with variable true onset) with 50 trials per participant and condition (@fig-eeg).
 
@@
 source("code/erp_template.R")
-
-# importing helper functions
-# source("code/functions.R")
 
@@
     scale_colour_met_d(name = "Johnson") +
-    labs(x = "Time (s)", y = "EEG signal (a.u.)")
-```
-
-````{=html}
-<!--
-
-We computed the average of the ERP difference (@fig-erp)...
-
-```{r fig-erp, eval = FALSE, echo = FALSE, out.width = "75%", fig.cap = "Group-level average difference between conditions (mean +/- standard error of the mean).
-The 'true' onset and offset are indicated by the vertical dashed lines."}
-# averaging across participants
-raw_df %>%
-    pivot_wider(names_from = condition, values_from = eeg, values_fn = mean) %>%
-    mutate(eeg_diff = cond2 - cond1) %>%
-    summarise(
-        eeg_mean = mean(eeg_diff),
-        eeg_sem = sd(eeg_diff) / sqrt(n() ),
-        .by = time
-        ) %>%
-    mutate(lower = eeg_mean - eeg_sem, upper = eeg_mean + eeg_sem) %>%
-    # plotting the data
-    ggplot(aes(x = time, y = eeg_mean) ) +
-    geom_hline(yintercept = 0, linetype = 2) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = true_offset, linetype = 2) +
-    geom_ribbon(
-        aes(ymin = lower, ymax = upper, colour = NULL),
-        alpha = 0.3, show.legend = FALSE
-        ) +
-    geom_line(show.legend = FALSE) +
-    labs(x = "Time (s)", y = "EEG difference (a.u.)")
-```
-
--->
-````
-
-## Model fitting
-
-We then fitted a Bayesian GAM using the `brms` package [@brms2017; @brms2018; @nalborczyk2019].
+    labs(x = "Time (s)", y = "Simulated EEG signal (a.u.)")
+```
+
+## Model description and model fitting
+
+We then fitted a Bayesian GAM (BGAM) using the `brms` package [@brms2017; @brms2018; @nalborczyk2019].
 We used the default priors in `brms` (i.e., weakly informative priors).
@@
 
-# timing the model fitting
-# around 600s for 2000 iterations on 4 chains and 4 cores
-# around 190s for 2000 iterations on 4 chains and 4 cores + cmdstanr and 01
-# around 440s for 2000 iterations on 4 chains and 4 cores + threading(2), cmdstanr and 01
-# library(tictoc)
-# tic()
-
-# fitting the GAM
+# fitting the BGAM
 gam <- brm(
@@
     file = "models/gam.rds"
-    #################################
-    # speeding up the model fitting #
-    #################################
-    # backend = "cmdstanr",
-    # threads = threading(2),
-    # stan_model_args = list(stanc_options = list("O1") )
-    )
-
-# stop timing
-# toc()
+    )
+```
+
+However, the previous model only included constant (fixed) effects, thus not properly accounting for between-participant variability.
+We next fit a multilevel version of the BGAM (BGAMM).
+Although it is possible to fit a BGAMM at the single-trial level, we present a resource-efficient version of the model that is fitted directly on the by-participant summary statistics (mean and SD), similar to what is done in meta-analysis.
+
+```{r meta-gam, results = "hide"}
+# averaging across participants
+summary_df <- raw_df %>%
+    summarise(
+        eeg_mean = mean(eeg),
+        eeg_sd = sd(eeg),
+        .by = c(participant, condition, time)
+        )
+
+# defining a contrast for condition
+contrasts(summary_df$condition) <- c(-0.5, 0.5)
+
+# fitting the BGAMM
+meta_gam <- brm(
+    # using by-participant SD of ERPs across trials
+    eeg_mean | se(eeg_sd) ~
+        condition + s(time, bs = "cr", k = 20, by = condition) +
+        (1 | participant),
+    data = summary_df,
+    family = gaussian(),
+    warmup = 2000,
+    iter = 5000,
+    chains = 8,
+    cores = 8,
+    file = "models/meta_gam.rds"
+    )
 ```
@@
 
-```{r model-preds, echo = FALSE}
+We depict the posterior predictions together with the posterior estimate of the slope for `condition` at each timestep (@fig-plot-post-slope).
+This figure suggests that the BGAMM provides an adequate description of the simulated data.
+
+```{r fig-plot-post-slope, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior estimate of the EEG activity in each condition (left) and posterior estimate of the difference in EEG activity (right) according to the BGAMM."}
 # defining a function to plot posterior predictions
@@
     post_preds %>%
-        ggplot(aes(x = time, y = eeg, colour = condition, fill = condition) ) +
+        ggplot(
+            aes(
+                x = time,
+                y = eeg_mean,
+                colour = condition, fill = condition
+                )
+            ) +
         geom_vline(xintercept = true_onset, linetype = 2) +
@@
         geom_point(
-            data = data %>% summarise(eeg = mean(eeg), .by = c(time, condition) ),
+            data = data %>% summarise(eeg_mean = mean(eeg_mean), .by = c(time, condition) ),
             aes(fill = NULL),
@@
 }
-```
-
-```{r gam-preds, echo = FALSE, eval = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior predictions of the GAM (left) and GP (right) models."}
-# plotting the posterior predictions
-plot_post_preds(model = gam) + plot_post_preds(model = gp_model)
-```
-
-We depict the posterior predictions together with the posterior estimate of the slope for `condition` at each timestep (@fig-plot-post-slope).
-
-```{r fig-plot-post-slope, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior estimate of the EEG activity in each condition (left) and posterior estimate of the difference in EEG activity (right) according to the GAM."}
+
 # defining a function to plot posterior slope
@@
 
-    # defining a sequence of time values to make predictions
-    time_seq <- crossing(
-        time = seq(min(data$time), max(data$time), length.out = 100),
-        condition = c("cond1", "cond2")
-        ) %>%
-        arrange(condition)
-
-    # retrieving posterior samples
-    posterior_samples <- posterior_epred(object = model, newdata = time_seq)
-
-    # computing the difference between cond2 and cond1
-    posterior_samples <- posterior_samples[, 101:200] - posterior_samples[, 1:100]
-
-    # computing the probability that y is above 0
-    prob_y_above_0 <- data.frame(
-        time = seq(min(data$time), max(data$time), length.out = 100)
-        ) %>%
-        mutate(
-          m = colMeans(posterior_samples),
-          lower = apply(
-              X = posterior_samples,
-              MARGIN = 2, FUN = quantile, probs = 0.025
-              ),
-          upper = apply(
-              X = posterior_samples,
-              MARGIN = 2, FUN = quantile, probs = 0.975
-              )
-          )
+    # computing the posterior probability for the slope over time
+    post_prob_slope <- model$data |>
+        add_epred_draws(object = model) |>
+        data.frame() |>
+        dplyr::select(participant, time, condition, .epred, .draw) |>
+        pivot_wider(names_from = condition, values_from = .epred) |>
+        mutate(epred_diff = cond2 - cond1) |>
+        # computing mean posterior probability at the group level
+        group_by(time) |>
+        summarise(
+            post_prob = quantile(x = epred_diff, probs = 0.5),
+            lower = quantile(x = epred_diff, probs = 0.025),
+            upper = quantile(x = epred_diff, probs = 0.975)
+            ) %>%
+        ungroup()
 
     # plotting it
-    prob_y_above_0 %>%
-        ggplot(aes(x = time, y = m) ) +
+    post_prob_slope %>%
+        ggplot(aes(x = time, y = post_prob) ) +
         geom_hline(yintercept = 0, linetype = 2) +
@@
         geom_vline(xintercept = true_offset, linetype = 2) +
+        geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
         geom_line() +
-        geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
-        scale_colour_met_d(name = "Johnson") +
-        scale_fill_met_d(name = "Johnson") +
         labs(
@@
 # plotting it
-# plot_post_slope(model = gam, eps = 0.1) + plot_post_slope(model = gp_model, eps = 0.1)
-plot_post_preds(model = gam) + plot_post_slope(model = gam)
+plot_post_preds(model = meta_gam) + plot_post_slope(model = meta_gam)
 ```
@@
 We can also express this as the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio, right).
-A ratio of 10 means that the probability of the difference being above 0 is 10 times higher than the probability of the difference not being above 0, given the data, the priors, and other model's assumptions.
-
-```{r fig-post-prob-test, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Posterior probability of the EEG difference (slope) being above 0 according to the GAM."}
-# defining a function to plot posterior prob of slope above 0
-plot_post_test <- function (model, data = NULL, eps = 0.1, n_approx = 1e3, n_draws = 1e3) {
-
-    # if no data is specified, use model$data
-    if (is.null(data) ) data <- model$data
-
-    # defining a sequence of time values to make predictions
-    time_seq <- crossing(
-        time = seq(min(data$time), max(data$time), length.out = n_approx),
-        # time = unique(data$time),
-        condition = c("cond1", "cond2")
-        ) %>%
-        arrange(condition)
-
-    # retrieving posterior samples
-    # posterior_samples <- posterior_epred(object = model, newdata = time_seq)
-
-    # computing the difference between cond2 and cond1
-    # posterior_samples <- posterior_samples[, (nrow(time_seq)/2+1):nrow(time_seq)] -
-    #     posterior_samples[, 1:(nrow(time_seq)/2)]
-
-    # computing the probability that y is above 0
-    # prob_y_above_0 <- data.frame(
-    #     time = time_seq$time,
-    #     m = colMeans(posterior_samples > (0 + eps) )
-    #     )
-
-    # computing probability metrics with the current eps value
-    prob_y_above_0 <- time_seq %>%
-        add_epred_draws(object = model, ndraws = n_draws) %>%
-        data.frame() %>%
-        dplyr::select(time, condition, .epred, .draw) %>%
-        pivot_wider(names_from = condition, values_from = .epred) %>%
-        mutate(epred_diff = cond2 - cond1) %>%
-        group_by(time) %>%
-        summarise(m = mean(epred_diff > 0 + eps) ) %>%
-        ungroup()
-
-    # plotting it
-    prob_y_above_0 %>%
-        ggplot(aes(x = time, y = m) ) +
-        geom_hline(yintercept = 0.5, linetype = 2) +
-        geom_vline(xintercept = true_onset, linetype = 2) +
-        geom_vline(xintercept = true_offset, linetype = 2) +
-        geom_line() +
-        ylim(c(0, 1) ) +
-        labs(x = "Time (s)", y = expression(Pr(beta>0~"|"~data, model) ) )
-
-}
-
-# using the function with the two models
-# plot_post_test(model = gam, eps = 0.1)
-```
+Note that a ratio of 10 means that the probability of the difference being above 0 is 10 times higher than the probability of the difference not being above 0, given the data, the priors, and other model's assumptions.
+Thresholding the posterior probability ratio thus provides a model-based approach for estimating the onset and offset of M/EEG effects.
 
@@
 
-```{r fig-post-prob-ratio, echo = FALSE, out.width = "100%", fig.width = 9, fig.asp = 0.5, fig.cap = "Left: Posterior probability of the EEG difference (slope) being above 0 according to the GAM.
-Right: Ratio of posterior probability according to the GAM (on a log10 scale).
+```{r fig-post-prob-ratio, echo = FALSE, out.width = "100%", fig.width = 9, fig.asp = 0.5, fig.cap = "Left: Posterior probability of the EEG difference (slope) being above 0 according to the BGAMM.
+Right: Ratio of posterior probability according to the BGAMM (on a log10 scale).
 Timesteps above threshold (10) are highlighted in green."}
-# defining some parameters
-n_approx <- 1e3
-eps <- 0.05
-threshold <- 10
-
-# defining a sequence of x values to make predictions
-time_seq <- crossing(
-    time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx),
-    # time = unique(ppt_df$time),
-    condition = c("cond1", "cond2")
-    ) %>%
-    arrange(condition)
-
-# computing the posterior probability that y is above 0
-n_posterior_samples <- ndraws(gam)
-prob_y_above_0 <- time_seq %>%
-    add_epred_draws(object = gam) %>%
+# defining a function to compute and plot the posterior probability ratio
+plot_post_test_ratio <- function (model, sesoi = 0, threshold = 10) {
+
+    n_posterior_samples <- ndraws(model)
+    prob_y_above_0 <- model$data %>%
+        add_epred_draws(object = model) %>%
         data.frame() %>%
-    dplyr::select(time, condition, .epred, .draw) %>%
+        dplyr::select(participant, time, condition, .epred, .draw) %>%
         pivot_wider(names_from = condition, values_from = .epred) %>%
         mutate(epred_diff = cond2 - cond1) %>%
+        # computing mean posterior probability at the group level
         group_by(time) %>%
-    summarise(m = mean(epred_diff > 0 + eps) ) %>%
+        summarise(m = mean(epred_diff > 0 + sesoi) ) %>%
+        mutate(prob_ratio = m / (1 - m) ) %>%
         ungroup() %>%
-    mutate(prob_ratio = m / (1 - m) ) %>%
-    # replacing infinite values by maximum possible values (number of posterior samples)
+        # ensuring there is no 0 or +Inf values
         mutate(prob_ratio = ifelse(is.infinite(prob_ratio), n_posterior_samples, prob_ratio) ) %>%
-    mutate(prob_ratio = ifelse(prob_ratio == 0, 1/n_posterior_samples, prob_ratio) )
-
-# printing the identified clusters
-# the signal was a truncated Gaussian defining an objective onset at 160 ms,
-# a maximum at 250 ms, and an offset at 342 ms.
-# exceeding_times_gam <- prob_y_above_0 %>%
-#     dplyr::filter(prob_ratio > threshold) %>%
-#     summarise(cluster_onset = min(time), cluster_offset = max(time) ) %>%
-#     # mutate(
-#     #     cluster_peak = prob_y_above_0 %>%
-#     #         dplyr::filter(post_prob == max(post_prob) ) %>%
-#     #         pull(time)
-#     #     ) %>%
-#     # mutate(true_onset = true_onset, true_peak = true_peak, true_offset = true_offset) %>%
-#     mutate(true_onset = true_onset, true_offset = true_offset) %>%
-#     mutate(model = "GAM")
-
-# creating a grid covering the range of time and log-transformed y-values
-bg_data <- expand.grid(
-    time = seq(min(prob_y_above_0$time), max(prob_y_above_0$time), length.out = 100),
-    # prob_ratio = 10^seq(log10(0.0001), log10(1e4), length.out = 100)
-    prob_ratio = 10^seq(log10(min(prob_y_above_0$prob_ratio) ), log10(max(prob_y_above_0$prob_ratio) ), length.out = 100)
-    )
-
-# assigning a fill value, for example, using prob_ratio or another variable
-bg_data$fill_value <- log10(bg_data$prob_ratio)
+        mutate(prob_ratio = ifelse(prob_ratio == 0, 1 / n_posterior_samples, prob_ratio) )
 
     # plotting it
-gam_ratio_plot <- prob_y_above_0 %>%
-    mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
+    post_plot <- prob_y_above_0 %>%
+        ggplot(aes(x = time, y = m) ) +
+        geom_hline(yintercept = 0.5, linetype = 2) +
+        geom_vline(xintercept = true_onset, linetype = 2) +
+        geom_vline(xintercept = true_offset, linetype = 2) +
+        geom_line() +
+        ylim(c(0, 1) ) +
+        labs(x = "Time (s)", y = expression(Pr(beta>0~"|"~data, model) ) )
+
+    # plotting it
+    ratio_plot <- prob_y_above_0 %>%
+        mutate(above_thres = ifelse(prob_ratio >= threshold, 1, NA) ) %>%
         ggplot(aes(x = time, y = prob_ratio) ) +
-    # geom_raster(
-    #     data = bg_data,
-    #     aes(x = time, y = prob_ratio, fill = fill_value),
-    #     alpha = 0.5,
-    #     interpolate = FALSE,
-    #     inherit.aes = FALSE,
-    #     show.legend = FALSE
-    #     ) +
-    # scale_fill_scico(palette = "vik", midpoint = log10(1) ) +
         geom_hline(yintercept = 1, linetype = "dashed") +
-    geom_hline(yintercept = 3, linetype = "dashed", color = "darkred") +
-    geom_hline(yintercept = 1/3, linetype = "dashed", color = "darkred") +
-    geom_hline(yintercept = 10, linetype = "dashed", color = "red") +
-    geom_hline(yintercept = 1/10, linetype = "dashed", color = "red") +
+        geom_hline(yintercept = 10, linetype = "dashed", color = "darkred") +
+        geom_hline(yintercept = 1/10, linetype = "dashed", color = "darkred") +
         geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
@@
             shape = 16,
-        size = 3,
+            size = 2,
             na.rm = TRUE
@@
             labels = label_log(digits = 2),
+            limits = c(1 / n_posterior_samples, n_posterior_samples),
             breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
-        # limits = c(0.0001, 1e4)
             ) +
-    coord_cartesian(expand = FALSE) +
         labs(
@@
 
-# retrieving posterior samples
-# posterior_samples <- posterior_epred(object = gp_model, newdata = time_seq)
-
-# computing the difference between cond2 and cond1
-# posterior_samples <- posterior_samples[, (ncol(posterior_samples)/2+1):ncol(posterior_samples)] -
-#     posterior_samples[, 1:(ncol(posterior_samples)/2)]
-
-# computing the probability that y is above 0
-# prob_y_above_0 <- data.frame(
-#     time = seq(min(ppt_df$time), max(ppt_df$time), length.out = n_approx)
-#     ) %>%
-#     mutate(post_prob = colMeans(posterior_samples) ) %>%
-#     mutate(m = colMeans(posterior_samples > eps) ) %>%
-#     mutate(prob_ratio = m / (1 - m) )
-
-# printing the identified clusters
-# the signal was a truncated Gaussian defining an objective onset at 160 ms,
-# a maximum at 250 ms, and an offset at 342 ms
-# exceeding_times_gp <- prob_y_above_0 %>%
-#     dplyr::filter(prob_ratio > threshold) %>%
-#     summarise(cluster_onset = min(time), cluster_offset = max(time) ) %>%
-#     mutate(
-#         cluster_peak = prob_y_above_0 %>%
-#             dplyr::filter(post_prob == max(post_prob) ) %>%
-#             pull(time)
-#         ) %>%
-#     mutate(true_onset = true_onset, true_peak = true_peak, true_offset = true_offset) %>%
-#     mutate(model = "GP")
-
-# plotting it
-# gp_ratio_plot <- prob_y_above_0 %>%
-#     mutate(above_thres = ifelse(prob_ratio > threshold, 1, NA) ) %>%
-#     ggplot(aes(x = time, y = prob_ratio) ) +
-#     geom_hline(yintercept = 1, linetype = "dashed") +
-#     geom_hline(yintercept = 3, linetype = "dashed", color = "darkred") +
-#     geom_hline(yintercept = 1/3, linetype = "dashed", color = "darkred") +
-#     geom_hline(yintercept = 10, linetype = "dashed", color = "red") +
-#     geom_hline(yintercept = 1/10, linetype = "dashed", color = "red") +
-#     geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
-#     geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
-#     geom_vline(xintercept = true_onset, linetype = 2) +
-#     geom_vline(xintercept = true_offset, linetype = 2) +
-#     geom_line() +
-#     geom_point(
-#         data = . %>% dplyr::filter(!is.na(above_thres) ),
-#         aes(y = threshold),
-#         colour = "darkgreen",
-#         shape = 15,
-#         size = 3,
-#         na.rm = TRUE
-#         ) +
-#     scale_y_log10(labels = label_log(digits = 2), limits = c(NA, 1e4) ) +
-#     labs(
-#         x = "Time (s)",
-#         y = expression(Pr(beta>0~"|"~data) / (1 - Pr(beta>0~"|"~data) ) )
-#         )
-
-# plotting results for the two models
-# gam_ratio_plot + gp_ratio_plot
-
-# or plotting only the GAM
-plot_post_test(model = gam, eps = 0.1) + gam_ratio_plot
-
-# printing the results
-# bind_rows(exceeding_times_gam, exceeding_times_gp) %>%
-#     mutate(across(.cols = where(is.numeric), .fns = ~round(.x, 2) ) ) %>%
-#     relocate(model, .before = cluster_onset)
+    # plotting both the posterior probability and the ratio...
+    post_plot + ratio_plot
+
+}
+
+# plotting the meta_gam predictions
+plot_post_test_ratio(model = meta_gam, threshold = 10)
 ```
@@
 
-## Multilevel modelling using summary statistics
-
-The previous model only included constant (fixed) effects, thus not properly accounting for between-participant variability.
-We next fit a multilevel version of the GAM.
-Although it is possible to fit a GAMM at the single-trial level, we present a resource-efficient version of the model that is fitted directly on the by-participant summary statistics (mean and SD), similar to what is done in meta-analysis.
-
-```{r meta-gam, message = FALSE}
-# averaging across participants
-summary_df <- raw_df %>%
-    summarise(
-        eeg_mean = mean(eeg),
-        eeg_sd = sd(eeg),
-        .by = c(participant, condition, time)
-        )
-
-# defining a contrast for condition
-contrasts(summary_df$condition) <- c(-0.5, 0.5)
-
-# fitting the GAM
-meta_gam <- brm(
-    # using by-participant SD of ERPs across trials
-    eeg_mean | se(eeg_sd) ~
-        condition + s(time, bs = "cr", k = 20, by = condition) +
-        (1 | participant),
-    data = summary_df,
-    family = gaussian(),
-    warmup = 2000,
-    iter = 5000,
-    chains = 8,
-    cores = 8,
-    file = "models/meta_gam.rds"
-    )
-```
-
-```{r meta-preds, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Posterior predictions for the hierarchical GAM."}
-# plotting the posterior predictions
-meta_gam_preds <- plot(
-    conditional_effects(x = meta_gam, effect = "time:condition"),
-    points = FALSE, theme = theme_light(), plot = FALSE
-    )[[1]] +
-    scale_colour_met_d(name = "Johnson") +
-    scale_fill_met_d(name = "Johnson") +
-    labs(x = "Time (s)", y = "EEG signal (a.u.)")
-# meta_gp_preds <- plot(
-#     conditional_effects(x = meta_gp, effect = "time:condition"),
-#     points = FALSE, theme = theme_light(), plot = FALSE
-#     )[[1]] +
-#     scale_colour_met_d(name = "Johnson") +
-#     scale_fill_met_d(name = "Johnson") +
-#     labs(x = "Time (s)", y = "EEG signal (a.u.)")
-
-# combining the two plots
-# meta_gam_preds + meta_gp_preds
-```
-
 ## Error properties of the proposed approach
 
-We then compute the difference between the true and estimated onset/offset of the EEG difference according to various `k` (GAM basis dimension) and `threshold` values.
+We then assess the performance of the proposed approach by computing the difference between the true and estimated onset/offset of the EEG difference according to various `k` (BGAM basis dimension) and `threshold` values.
 Remember that the EEG signal was generated from a truncated Gaussian with an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
-@fig-onset-error shows that the multilevel GAM can *almost exactly* recover the true onset and offset values, given some reasonable choice of `k` and `threshold` values (e.g., a threshold of 20).
+@fig-onset-error shows that the multilevel GAM can almost exactly recover the true onset and offset values, given some reasonable choice of `k` and `threshold` values.
 We provide more detailed recommendations on how to set `k` in @sec-basis.
-
-```{r fig-onset-error, eval = TRUE, echo = FALSE, results = "hide", dev = "png", dpi = 300, fig.width = 9, fig.asp = 0.5, fig.cap = "Average estimation error for the onset (left) and offset (right) according to various basis dimension and threshold values (according to the multilevel GAM)."}
-# defining values of k and threshold to test
-# threshold_values <- seq(from = 1, to = 50, by = 1)
-# kvalues <- c(5, 10, 15, 20, 25, 30, 35, 40)
-
-# number of simulations to run per combination
-# nsims <- 10
-
-# number of posterior samples to use
-# n_post_samples <- 1e4
-
-# importing the data generation functions 
-# source(file = "code/generate_eeg_data.R")
-
-# initialising results dataframe
-# results <- crossing(kvalue = kvalues, sim_id = 1:nsims, threshold = threshold_values) %>%
-#     mutate(
-#         estimated_onset = NA,
-#         estimated_offset = NA,
-#         error_onset = NA,
-#         error_offset = NA
-#         )
-
-# looping over different values of k and threshold
-# for (i in seq_len(nrow(results) ) ) {
-# 
-#     # retrieving current kvalue and threshold
-#     kvalue <- results$kvalue[i]
-#     threshold <- results$threshold[i]
-# 
-#     # printing progress
-#     cat(
-#         "\nAssessing combination:", i, "out of", nrow(results), "combinations...\n",
-#         "k-value:", kvalue, "\n", "threshold:", threshold, "\n"
-#         )
-# 
-#     # when we assess a new k-value or new sim_id, we generate new data
-#     # and fit a new model...
-#     if (threshold == 1) {
-# 
-#         # generate some data
-#         raw_df <- generate_data(
-#             n_trials = n_trials, n_ppt = n_ppt, outvar = outvar,
-#             srate = srate, ronset = ronset
-#             )
-# 
-#         # sanity check
-#         # head(raw_df)
-# 
-#         # averaging across participants
-#         summary_df <- raw_df %>%
-#             summarise(
-#                 eeg_mean = mean(eeg),
-#                 eeg_sd = sd(eeg),
-#                 .by = c(participant, condition, time)
-#                 )
-# 
-#         # defining a contrast for condition
-#         contrasts(summary_df$condition) <- c(-0.5, 0.5)
-# 
-#         # fitting the GAM
-#         error_model <- brm(
-#             # using by-participant SD of EEG across trials
-#             eeg_mean | se(eeg_sd) ~
-#                 condition + s(time, bs = "cr", k = kvalue, by = condition) +
-#                 (1 | participant),
-#             data = summary_df,
-#             family = gaussian(),
-#             warmup = 2000,
-#             iter = 5000,
-#             chains = 8,
-#             cores = 8,
-#             ###################################
-#             # when running on the HPC cluster #
-#             ###################################
-#             backend = "cmdstanr",
-#             stan_model_args = list(stanc_options = list("O1") )
-#             )
-# 
-#         prob_y_above_0 <- error_model$data %>%
-#             # add_epred_draws(object = error_model, ndraws = n_post_samples) %>%
-#             add_epred_draws(object = error_model) %>%
-#             data.frame() %>%
-#             dplyr::select(participant, time, condition, .epred, .draw) %>%
-#             pivot_wider(names_from = condition, values_from = .epred) %>%
-#             mutate(epred_diff = cond2 - cond1) %>%
-#             # computing mean posterior probability at the group level
-#             group_by(time) %>%
-#             summarise(m = mean(epred_diff > 0) ) %>%
-#             mutate(prob_ratio = m / (1 - m) ) %>%
-#             ungroup() %>%
-#             # ensuring there is no 0 or +Inf values
-#             mutate(
-#                 prob_ratio = ifelse(
-#                     is.infinite(prob_ratio),
-#                     ndraws(error_model),
-#                     prob_ratio
-#                     )
-#                 ) %>%
-#             mutate(
-#                 prob_ratio = ifelse(
-#                     prob_ratio == 0,
-#                     1 / ndraws(error_model),
-#                     prob_ratio
-#                     )
-#                 )
-# 
-#     }
-# 
-#     # finding onset, offset, and peak for the current threshold
-#     exceeding_times <- prob_y_above_0 %>%
-#         dplyr::filter(prob_ratio > threshold) %>%
-#         summarise(
-#             cluster_onset = min(time, na.rm = TRUE),
-#             cluster_offset = max(time, na.rm = TRUE)
-#             ) %>%
-#         mutate(true_onset = true_onset, true_offset = true_offset)
-# 
-#     # storing the results in the dataframe
-#     if (nrow(exceeding_times) > 0) {
-# 
-#         results$estimated_onset[i] <- exceeding_times$cluster_onset
-#         results$estimated_offset[i] <- exceeding_times$cluster_offset
-# 
-#         # computing errors
-#         results$error_onset[i] <- abs(
-#             exceeding_times$cluster_onset - exceeding_times$true_onset
-#             )
-#         results$error_offset[i] <- abs(
-#             exceeding_times$cluster_offset - exceeding_times$true_offset
-#             )
-# 
-#     }
-# 
-# }
-
-# saving the simulation results
-# saveRDS(object = results, file = "results/meta_gam_error_properties.rds")
-
+This figure further reveals that the optimal `k` and `threshold` values may differ for the onset and offset values, and there there seems to exist a trade-off between these two parameters: lower `k` values lead to poorer estimations, but these poor estimations can be compensated (only to some extent) by higher `threshold` values (and reciprocally).
+
+```{r fig-onset-error, eval = TRUE, echo = FALSE, results = "hide", dev = "png", dpi = 300, fig.width = 9, fig.asp = 0.5, fig.cap = "Average estimation error (RMSE) for the onset (left) and offset (right) according to various basis dimension and threshold values for the BGAM (computed from 100 simulated datasets)."}
 # running the simulation on the HPC cluster
@@
 # results <- readRDS(file = "results/meta_gam_error_properties.rds")
+# results <- readRDS(file = "results/meta_gam_error_properties_100sims.rds")
 results <- readRDS(file = "results/meta_gam_error_properties_cluster.rds")
 
+# defining the true onset and offset in seconds
+true_onset <- 0.160
+true_peak <- 0.250
+true_offset <- 0.342
+
 # plotting the results
-results %>%
-    group_by(kvalue, threshold, sim_id) %>%
-    # mutate(error_onset = ) %>%
-    summarise(
-        # value = if (onset_offset[1] == "onset") min(value) else max(value),
-        cluster_onset = min(cluster_onset),
-        cluster_offset = max(cluster_offset)
-        # .groups = "drop"
-        ) %>%
-    ungroup() %>%
-    group_by(kvalue, threshold) %>%
-    # summarise(across(estimated_onset:error_offset, median) ) %>%
-    # summarise(across(estimated_onset:error_offset, mean) ) %>%
-    mutate(
-        error_onset = abs(cluster_onset - true_onset),
-        error_offset = abs(cluster_offset - true_offset)
-        ) %>%
-    summarise(across(error_onset:error_offset, mean) ) %>%
-    ungroup() %>%
-    pivot_longer(cols = error_onset:error_offset) %>%
-    mutate(name = factor(x = name, levels = c("error_onset", "error_offset") ) ) %>%
-    ggplot(
-        aes(
-            x = as.factor(kvalue),
-            y = as.factor(threshold),
-            fill = value
-            )
-        ) +
-    geom_raster(interpolate = FALSE) +
-    # geom_tile() +
-    # geom_point(
-    #     data = results %>%
+# results %>%
+#     group_by(kvalue, threshold, sim_id) %>%
+#     summarise(
+#         # value = if (onset_offset[1] == "onset") min(value) else max(value),
+#         cluster_onset = min(cluster_onset),
+#         cluster_offset = max(cluster_offset)
+#         # .groups = "drop"
+#         ) %>%
+#     ungroup() %>%
 #     group_by(kvalue, threshold) %>%
-    #         summarise(across(estimated_onset:error_offset, mean) ) %>%
+#     # summarise(across(estimated_onset:error_offset, median) ) %>%
+#     # summarise(across(estimated_onset:error_offset, mean) ) %>%
+#     summarise(
+#         # error_onset = mean(abs(cluster_onset - true_onset) ),
+#         # error_offset = mean(abs(cluster_offset - true_offset) )
+#         error_onset = sqrt(mean(cluster_onset - true_onset)^2),
+#         error_offset = sqrt(mean(cluster_offset - true_offset)^2)
+#         ) %>%
 #     ungroup() %>%
-    #         na.omit() %>%
-    #         filter(error_onset == min(error_onset) ),
-    #     # aes(x = kvalue, y = threshold),
-    #     aes(x = kvalue, y = value),
-    #     color = "orangered", size = 4, shape = 4,
-    #     # show.legend = FALSE
-    #     show.legend = TRUE
-    #     ) +
-    # scale_x_continuous(expand = c(0, 0) ) +
-    # scale_y_continuous(expand = c(0, 0) ) +
-    facet_wrap(~name) +
-    scale_x_discrete(expand = c(0, 0) ) +
-    scale_y_discrete(expand = c(0, 0), breaks = c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50) ) +
-    # scale_y_continuous(n.breaks=10) +
-    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-    # scale_fill_viridis_c(direction = 1) +
-    labs(
-        x = "GAM basis dimension (k)",
-        y = "Threshold",
-        fill = "Error (s)"
-        )
-
-# results %>%
-#     group_by(kvalue, threshold) %>%
-#     summarise(across(estimated_onset:error_offset, mean) ) %>%
-#     ungroup() %>%
-#     # head()
 #     pivot_longer(cols = error_onset:error_offset) %>%
@@
 #         aes(
-#             x = threshold,
-#             y = value,
-#             colour = kvalue,
-#             fill = kvalue
+#             x = as.factor(kvalue),
+#             y = as.factor(threshold),
+#             fill = value
 #             )
 #         ) +
-#     # geom_raster(interpolate = FALSE) +
+#     geom_raster(interpolate = FALSE) +
 #     # geom_tile() +
-#     # geom_line() +
-#     geom_point() +
 #     # geom_point(
-#     #     data = results %>% dplyr::filter(error_onset == min(error_onset, na.rm = TRUE) ),
+#     #     data = results %>%
+#     #         group_by(kvalue, threshold) %>%
+#     #         summarise(across(estimated_onset:error_offset, mean) ) %>%
+#     #         ungroup() %>%
+#     #         na.omit() %>%
+#     #         filter(error_onset == min(error_onset) ),
 #     #     # aes(x = kvalue, y = threshold),
-#     #     aes(x = as.factor(kvalue), y = threshold),
+#     #     aes(x = kvalue, y = value),
 #     #     color = "orangered", size = 4, shape = 4,
@@
 #     #     ) +
-#     # scale_x_continuous(expand = c(0, 0) ) +
-#     # scale_y_continuous(expand = c(0, 0) ) +
 #     facet_wrap(~name) +
-#     scale_colour_viridis_c() +
-#     # scale_x_discrete(expand = c(0, 0) ) +
-#     # scale_y_discrete(expand = c(0, 0) ) +
-#     # scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) )
+#     scale_x_discrete(expand = c(0, 0) ) +
+#     scale_y_discrete(expand = c(0, 0), breaks = c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50) ) +
+#     scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
 #     labs(
@@
 #         y = "Threshold",
-#         fill = "Error (s)"
+#         fill = "RMSE"
 #         )
 
-# plotting the results
-# onset_plot <- results %>%
-#     # ggplot(aes(x = eps, y = threshold, fill = error_onset) ) +
-#     # ggplot(aes(x = as.factor(kvalue), y = threshold, fill = error_onset) ) +
-#     ggplot(aes(x = as.factor(kvalue), y = as.factor(threshold), fill = error_onset) ) +
-#     geom_tile(show.legend = FALSE) +
-#     geom_point(
-#         data = results %>% dplyr::filter(error_onset == min(error_onset, na.rm = TRUE) ),
-#         # aes(x = kvalue, y = threshold),
-#         aes(x = as.factor(kvalue), y = threshold),
-#         color = "orangered", size = 4, shape = 4,
-#         # show.legend = FALSE
-#         show.legend = TRUE
+results %>%
+    # filter(kvalue > 5) %>%
+    group_by(kvalue, threshold, sim_id) %>%
+    summarise(
+        cluster_onset = min(cluster_onset),
+        cluster_offset = max(cluster_offset),
+        .groups = "drop"
+        ) %>%
+    ungroup() %>%
+    group_by(kvalue, threshold) %>%
+    summarise(
+        # error_onset = median(cluster_onset - true_onset),
+        # error_offset = median(cluster_offset - true_offset)
+        # error_onset = median(abs(cluster_onset - true_onset) ),
+        # error_offset = median(abs(cluster_offset - true_offset) )
+        error_onset = sqrt(mean(cluster_onset - true_onset)^2),
+        error_offset = sqrt(mean(cluster_offset - true_offset)^2)
+        ) %>%
+    ungroup() %>%
+    pivot_longer(cols = error_onset:error_offset) %>%
+    mutate(
+        name = factor(
+            x = name,
+            levels = c("error_onset", "error_offset"),
+            labels = c("onset", "offset")
+            )
+        ) %>%
+    ggplot(
+        aes(
+            x = threshold,
+            y = value,
+            colour = as.factor(kvalue),
+            fill = as.factor(kvalue),
+            group = as.factor(kvalue),
+            )
+        ) +
+    geom_line(linewidth = 0.5) +
+    geom_point(size = 0.8) +
+    # geom_segment(
+    #     data = . %>% filter(value == min(value), .by = name),
+    #     aes(
+    #         x = threshold,
+    #         xend = threshold,
+    #         y = 0.01,
+    #         yend = 0,
+    #         colour = as.factor(kvalue)
+    #         ),
+    #     arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
+    #     inherit.aes = FALSE,
+    #     linewidth = 1.0,
+    #     show.legend = FALSE
     #     ) +
-#     # scale_x_continuous(expand = c(0, 0) ) +
-#     # scale_y_continuous(expand = c(0, 0) ) +
-#     scale_x_discrete(expand = c(0, 0) ) +
-#     scale_y_discrete(expand = c(0, 0) ) +
-#     scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-#     labs(
-#         title = "Onset error",
-#         # x = "eps",
-#         x = "GAM basis dimension (k)",
-#         y = "Threshold",
-#         fill = "Error (s)"
-#         )
-# 
-# offset_plot <- results %>%
-#     # ggplot(aes(x = eps, y = threshold, fill = error_offset) ) +
-#     # ggplot(aes(x = kvalue, y = threshold, fill = error_offset) ) +
-#     ggplot(aes(x = as.factor(kvalue), y = as.factor(threshold), fill = error_offset) ) +
-#     geom_tile(show.legend = FALSE) +
-#     geom_point(
-#         data = results %>% dplyr::filter(error_offset == min(error_offset, na.rm = TRUE) ),
-#         # aes(x = kvalue, y = threshold),
-#         aes(x = as.factor(kvalue), y = threshold),
-#         color = "orangered", size = 4, shape = 4,
-#         # show.legend = FALSE
-#         show.legend = TRUE
-#         ) +
-#     # scale_x_continuous(expand = c(0, 0) ) +
-#     # scale_y_continuous(expand = c(0, 0) ) +
-#     scale_x_discrete(expand = c(0, 0) ) +
-#     scale_y_discrete(expand = c(0, 0) ) +
-#     scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-#     labs(
-#         title = "Offset error",
-#         # x = "eps",
-#         x = "GAM basis dimension (k)",
-#         y = "Threshold",
-#         fill = "Error (s)"
-#         )
-
-# combining the plots
-# onset_plot + offset_plot + plot_layout(guides = "collect")
-
-# displaying best parameters values (according to onset error)
-# results %>%
-#     dplyr::select(-estimated_peak, -error_peak) %>%
-#     arrange(error_onset) %>%
-#     data.frame() %>%
-#     head()
-
-# displaying best parameters values (according to offset error)
-# results %>%
-#     dplyr::select(-estimated_peak, -error_peak) %>%
-#     arrange(error_offset) %>%
-#     data.frame() %>%
-#     head()
-
-# displaying best parameters values (according to both onset and offset error)
-# results %>%
-#     group_by(kvalue, threshold) %>%
-#     summarise(across(estimated_onset:error_offset, mean) ) %>%
-#     # summarise(across(estimated_onset:error_offset, mode) ) %>%
-#     ungroup() %>%
-#     arrange(error_onset, error_offset) %>%
-#     data.frame() %>%
-#     head(40)
+    # geom_smooth(alpha = 0.1, linewidth = 0.8) +
+    facet_wrap(~name) +
+    scale_fill_viridis_d() +
+    scale_colour_viridis_d() +
+    scale_y_continuous(
+        labels = scales::label_number(),
+        transform = "sqrt"
+        ) +
+    labs(
+        x = "BGAM threshold",
+        y = "Estimation error (RMSE)",
+        colour = "BGAM basis\ndimension (k)",
+        fill = "BGAM basis\ndimension (k)"
+        )
 ```
@@
 
-We then compared the ability of the GAMM to correctly estimate the onset and offset of the ERP difference to other widely-used methods.
+We then compared the ability of the BGAMM to correctly estimate the onset and offset of the ERP difference to other widely-used methods.
 First, we conducted mass-univariate t-tests (thus treating each timestep independently) and identified the onset and offset of the ERP difference as the first and last values crossing an arbitrary significance threshold ($\alpha = 0.05$).
@@
 We also compared these estimates to the onset and offset as estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared t-values [as in @rousselet_using_2025].
-@fig-corrections illustrates the onsets and offsets estimated by each method on a single simulated dataset and shows that all methods over-estimate the true onset and under-estimate the true offset.
-
-```{=html}
-<!--
-
-For visualisation and interpretability purposes, we converted p-values to s-values, which can be interpreted as bits of surprising information, assuming a null effect [@greenland2019] (@fig-corrections).
-
--->
-```
+@fig-corrections illustrates the onsets and offsets estimated by each method on a single simulated dataset and shows that all methods systematically overestimate the true onset and underestimate the true offset.
 
@@
 
-clusters <- find_clusters(df = tests_results) %>%
+clusters <- find_clusters(tests_results) %>%
     rename(method = name, onset = cluster_onset, offset = cluster_offset) %>%
@@
     labs(x = "Time (s)", y = "Squared t-value", colour = NULL)
-
-# plotting -log2(p) (surprisal under H0) with thresholds (on p-values)
-# https://lesslikely.com/statistics/s-values/
-# a s-value of 4 (bits of information) is no more surprising than
-# getting all heads on 4 fair coin tosses
-# pvals_plot <- tests_results %>%
-#     pivot_longer(cols = pval:pval_holm) %>%
-#     mutate(sval = -log2(value) ) %>%
-#     ggplot(aes(x = time, y = sval, colour = name) ) +
-#     geom_density(
-#         aes(fill = name),
-#         stat = "identity",
-#         alpha = 0.8,
-#         show.legend = FALSE
-#         ) +
-#     geom_hline(yintercept = -log2(aath), linetype = 2) +
-#     geom_vline(xintercept = true_onset, linetype = 2) +
-#     geom_vline(xintercept = true_offset, linetype = 2) +
-#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 5) ) +
-#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 5) ) +
-#     labs(
-#         x = "Time (s)",
-#         y = "-log2(p-value)"
-#         )
-
-# combining the two plots
-# tvals_plot + pvals_plot +
-#     # plot_annotation(title = plot_title, tag_levels = "A") &
-#     # theme(plot.title = element_text(hjust = 0.5) )
-#     plot_annotation(tag_levels = "A") + plot_layout(guides = "collect")
 ```
@@
 
-The various methods were compared using the median error (i.e., true-estimated) and variance of onset/offset estimates computed on 10.000 simulated datasets.
+To assess the accuracy of group-level onset estimation, the various methods were compared using the bias (i.e., median(estimated-true)), median absolute error (MAE), root mean square error (RMSE), variance, and median absolute deviation (MAD) of onset/offset estimates computed on 1000 simulated datasets.
+As in @rousselet_using_2025, each participant was assigned a random onset betwee, 150 and 170ms.
 
@@
 
-To provide a more exhaustive assessment of these methods, we  also evaluated the performance of the proposed approach on actual MEG data [decoding results from @nalborczyk:inprep].
+To complement the simulation study, we evaluated the performance of the various methods on actual MEG data [decoding results from @nalborczyk:inprep].
 In this study, we conducted time-resolved multivariate pattern analysis (MVPA, also known as decoding) of MEG data during reading tasks.
 As a result, we obtain a timecourse of decoding performance (ROC AUC), bounded between 0 and 1, for each participant (for a total of 32 participants).
-Next, we want to *test* whether the group-level average decoding accuracy is above chance (i.e., 0.5) at each timestep (@fig-decoding-data).
-To achieve this, we fitted a similar GAM as discussed previously, but we replaced the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one to account for the bounded nature of AUC values (between 0 and 1) [for a tutorial on Beta regression, see @coretta2025].
+Next, we wanted to *test* whether the group-level average decoding accuracy is above chance (i.e., 0.5) at each timestep (@fig-decoding-data).
+To achieve this, we fitted a BGAM as introduced previously, but we replaced the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one to account for the bounded nature of AUC values (between 0 and 1) [for a tutorial on Beta regression, see @coretta2025].
 
@@
 
-```{=html}
-<!--
-
-OR, we could compute the onset/offset at the participant-level and correlate estimates across splits, BUT, individual-level estimates are not available for all methods...
-
-To test consistency between participants we calculated the split-half consistency across 1000 permutations.
-For each permutation, the participants were randomly split into two halves.
-Onset and offset were estimated at the participant level.
-Then, the Spearman rank correlation between the halves was calculated, corrected by the Spearman-Brown split-half reliability correction.
-A null distribution was calculated by randomly onset/offset values between words within one of the participant halves.
-
--->
-```
-
-```{r fig-decoding-data, cache = FALSE, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Group-level average decoding performance (N=32) superimposed with the GAM predictions (in blue) and the region of practical equivalence (ROPE, in orange) computed from the baseline period.
+```{r fig-decoding-data, cache = FALSE, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Group-level average decoding performance (N=32) superimposed with the GAM predictions (in blue) and the region of practical equivalence (ROPE, in orange) computed from the baseline period [data from @nalborczyk:inprep].
 The blue horizontal markers indicate the timesteps at which the posterior probability ratio exceeds 20."}
@@
 
-# finding onset and offset for the current threshold
-# exceeding_times <- prob_y_above_0 %>%
-#     dplyr::filter(prob_ratio > post_prob_ratio_threshold) %>%
-#     summarise(
-#         cluster_onset = min(time, na.rm = TRUE),
-#         cluster_offset = max(time, na.rm = TRUE)
-#         )
-# or finding all clusters
-# defining a function to find onset and offset in timeseries
-# find_clusters <- function (mask, timeseries) {
-#     
-#     # initialising onset and offset values
-#     onset <- NA
-#     offset <- NA
-#     
-#     # identifying the onset and offset
-#     masked_timesteps <- try(timeseries[which(mask)], silent = TRUE)
-#     onset <- head(x = masked_timesteps, n = 1)
-#     offset <- tail(x = masked_timesteps, n = 1)
-#     
-#     # returning the onset and offset
-#     return (c(onset, offset) )
-#     
-# }
-# 
-# find_onset_offset(
-#         tests_results$pval <= alpha_level,
-#         tests_results$time
-#         )
-# prob_y_above_0 %>%
-#     dplyr::filter(prob_ratio >= post_prob_ratio_threshold) %>%
-#     dplyr::pull(time)
+# defining a function to find all onset and offset in timeseries
 find_clusters <- function (df, threshold = 20) {
@@
 
+# defining a threshold
 post_prob_ratio_threshold <- 20
@@
 
-@fig-simulation-results shows a summary of the simulation results, revealing that the proposed approach (`BGAM`) has the lowest median error and variance for both the onset and offset estimates.
-The `Cluster mass`, `Holm`, and `Change point` also have good performance, but surprisingly, the `TFCE` method has relatively bad performance for estimating the effect offset.
-Unsurprisingly, the `Raw p-value`, `FDR BH95`, and `FDR BY01` methods show the worst performance.
-
-```{r table-simulation-results, eval = FALSE, echo = FALSE}
-# loading the simulation results
-# Here are a couple of example tables (@table-simulation-results)....
-sim_results <- readRDS(file = "results/errors_results_100sims.rds")
-
-# summarising the results
-sim_results_summary <- sim_results %>%
-    # keeping only threshold = 20 for BGAM and threshold = 0.05 for frequentist results
-    filter(threshold %in% c(0.05, 20) ) %>%
-    mutate(
-        error = case_when(
-            grepl("onset", onset_offset) ~ value - true_onset,
-            grepl("offset", onset_offset) ~ value - true_offset
-            )
-        ) %>%
-    group_by(method, onset_offset) %>%
-    summarise(
-        mean_error = mean(error, na.rm = TRUE),
-        median_error = median(error, na.rm = TRUE),
-        median_abs_error = median(abs(error), na.rm = TRUE),
-        rmse = sqrt(mean(error)^2),
-        variance = var(error, na.rm = TRUE),
-        mad = mad(error, na.rm = TRUE),
-        .groups = "drop"
-        ) %>%
-    ungroup() %>%
-    mutate(
-        method = factor(
-            x = method,
-            levels = c(
-                "raw_p", "pval_bh", "pval_by", "pval_holm",
-                "cluster_mass", "cluster_tfce", "cpt", "brms"
-                ),
-            labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-                "TFCE", "Change point", "BGAM (new method)"
-                ),
-            )
-        ) %>%
-    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
-    data.frame()
-
-# creating the table
-# library(tinytable)
-# tt(
-#     x = sim_results_summary,
-#     digits = 2
-#     )
-```
-
-```{r fig-simulation-results, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median error and variance of onset and offset estimates for each method.
-The vertical dashed line represents an unbiased estimate of the true onset/ofsset.
+@fig-simulation-results shows a summary of the simulation results, revealing that the proposed approach (`BGAM`) has the lowest median absolute error (MAE) and variance for both the onset and offset estimates.
+The `Cluster mass` and `Change point` also have good performance, but surprisingly, the `TFCE` method has relatively bad performance for estimating the effect offset (similar performance to the `Holm` and `FDR BY01` methods).
+Unsurprisingly, the `Raw p-value` and `FDR BH95` methods show the worst performance.
+
+```{r fig-simulation-results, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median absolute error and variance of onset and offset estimates for each method.
 Variance is plotted on a log10 scale for visual purposes."}
@@
 
+# plotting the distribution of results
+# sim_results %>%
+#     # keeping only threshold = 20 for BGAM and threshold = 0.05 for frequentist results
+#     filter(threshold %in% c(0.05, 20) ) %>%
+#     mutate(
+#         error = case_when(
+#             # grepl("onset", onset_offset) ~ abs(value - true_onset),
+#             # grepl("offset", onset_offset) ~ abs(value - true_offset)
+#             grepl("onset", onset_offset) ~ value - true_onset,
+#             grepl("offset", onset_offset) ~ value - true_offset
+#             )
+#         ) %>%
+#     mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
+#     mutate(
+#         method = factor(
+#             x = method,
+#             levels = c(
+#                 "raw_p", "pval_bh", "pval_by", "pval_holm",
+#                 "cluster_mass", "cluster_tfce", "cpt", "brms"
+#                 ),
+#             labels = c(
+#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+#                 "TFCE", "Change point", "BGAM"
+#                 ),
+#             )
+#         ) %>%
+#     mutate(
+#         plot_row = case_when(
+#             method %in% c("Raw p-value", "Holm") ~ 1,
+#             method %in% c("FDR BH95", "FDR BY01") ~ 2,
+#             method %in% c("Cluster mass", "TFCE") ~ 3,
+#             method %in% c("Change point", "BGAM") ~ 4
+#             )
+#         ) %>%
+#     ggplot(
+#         aes(
+#             x = error,
+#             colour = method,
+#             fill = method
+#             )
+#         ) +
+#     geom_vline(xintercept = 0, linetype = 2) +
+#     geom_density(aes(fill = NULL) ) +
+#     facet_grid(plot_row~onset_offset, scales = "free_x") +
+#     theme_light(base_size = 12, base_family = "Open Sans") +
+#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+#     labs(x = "Onset/Offset (s)", y = "Density")
+
 # computing and plotting bias and variance
-# bias := true_onset - median(sim_onset_distrib)
-# variance := variance(sim_onset_distrib)
+# bias := median(sim_onset_offset_distrib - true_onset_offset)
+# variance := var(sim_onset_offset_distrib - true_onset_offset)
 sim_results %>%
@@
         error = case_when(
-            # grepl("onset", measure) ~ abs(value - true_onset),
-            # grepl("offset", measure) ~ abs(value - true_offset)
+            # grepl("onset", onset_offset) ~ abs(value - true_onset),
+            # grepl("offset", onset_offset) ~ abs(value - true_offset)
             grepl("onset", onset_offset) ~ value - true_onset,
@@
         median_error = median(error, na.rm = TRUE),
-        variance = var(error, na.rm = TRUE),
-        # variance = mad(error, na.rm = TRUE),
+        median_abs_error = median(abs(error), na.rm = TRUE),
+        # variance = var(error, na.rm = TRUE),
+        variance = var(value, na.rm = TRUE),
+        mad = mad(value, na.rm = TRUE),
         .groups = "drop"
@@
         aes(
-            x = median_error,
+            # x = median_error,
+            # y = median_abs_error,
+            x = median_abs_error,
             y = variance,
@@
         ) +
-    geom_vline(xintercept = 0, linetype = 2) +
+    # geom_vline(xintercept = 0, linetype = 2) +
     geom_point(
@@
     scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    labs(x = "Median error (s)", y = "Variance (log-scale)")
+    # labs(x = "Median error (s)", y = "Variance (log-scale)")
+    # labs(x = "Median error (s)", y = "Median absolute error (log-scale)")
+    labs(x = "Median absolute error (s)", y = "Variance (log-scale)")
 
@@
 ggsave(
-    filename = "figures/simulation_results_median_error_variance.png",
+    # filename = "figures/simulation_results_median_error_variance.png",
+    # filename = "figures/simulation_results_bias_mae.png",
+    filename = "figures/simulation_results_mae_variance.png",
     width = 12, height = 6, dpi = 300,
@@
 
+Results are further summarised in @tbl-simulation-results, which shows that the `BGAM` is perfectly unbiased (i.e., it has a bias of 0s) for the onset and almost exactly unbiased for the onset (with a bias of approximately -2ms).
+The `Bias` column shows that all other methods tend to estimate the onset later than the true onset and to estimate the offet earlier than the true offset.
+As can be seen from this table, the `BGAM` has the best performance on all included metrics.
+
+```{r tbl-simulation-results, eval = TRUE, echo = FALSE, results = "asis"}
+#| tbl-cap: "Summary statistics of the onset and offset estimates for each method (ordered by MAE)."
+
+# loading the simulation results
+sim_results <- readRDS(file = "results/errors_results_100sims.rds")
+
+# summarising the results
+sim_results_summary <- sim_results %>%
+    # keeping only threshold = 20 for BGAM and threshold = 0.05 for frequentist results
+    filter(threshold %in% c(0.05, 20) ) %>%
+    mutate(
+        error = case_when(
+            grepl("onset", onset_offset) ~ value - true_onset,
+            grepl("offset", onset_offset) ~ value - true_offset
+            )
+        ) %>%
+    group_by(method, onset_offset) %>%
+    summarise(
+        median_error = median(error, na.rm = TRUE),
+        median_abs_error = median(abs(error), na.rm = TRUE),
+        rmse = sqrt(mean(error)^2),
+        variance = var(error, na.rm = TRUE),
+        mad = mad(error, na.rm = TRUE),
+        .groups = "drop"
+        ) %>%
+    ungroup() %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c(
+                "raw_p", "pval_bh", "pval_by", "pval_holm",
+                "cluster_mass", "cluster_tfce", "cpt", "brms"
+                ),
+            labels = c(
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+                "TFCE", "Change point", "BGAM (new method)"
+                ),
+            )
+        ) %>%
+    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
+    data.frame()
+
+# creating the table
+sim_results_summary %>%
+    mutate(
+        onset_offset = factor(
+            x = onset_offset,
+            levels = c("offset", "onset")
+            )
+        ) %>%
+    arrange(median_abs_error) %>%
+    group_by(onset_offset) %>%
+    gt() %>%
+    # gt(groupname_col = "onset_offset") %>%
+    fmt_number(
+        columns = everything(),
+        decimals = 4
+        ) %>%
+    cols_align(
+        align = "center",
+        columns = everything()
+        ) %>%
+    cols_align(
+        align = "left",
+        columns = method
+        ) %>%
+    cols_label(
+        method = "",
+        median_error = "Bias",
+        median_abs_error = "MAE",
+        rmse = "RMSE",
+        variance = "Variance",
+        mad = "MAD"
+        ) %>%
+    tab_options(
+        table.align = "center",
+        # table.width = "75%",
+        table.font.size = 12
+        ) %>%
+    as_latex()
+```
+
 \newpage
@@
 
-@fig-onset-offset shows the group-level average decoding performance through time with onset and offset estimates for each method.
+@fig-onset-offset shows the group-level average decoding performance through time with onset and offset estimates from each method.
 Overall, this figure shows that both the `Raw p-value` and `FDR BH95` methods are extremely lenient, considering that the decoding performance is above chance before the onset of the stimulus (false positive) and until the end of the trial.
@@
 
-[^1]: It should be noted that although each method can produce several "clusters" of "significant" timesteps, we only considered the first (onset) and last (offset) timesteps identified by each method to compute the estimation error.
+[^1]: It should be noted that although each method can produce several "clusters" of timesteps, we only considered the first (onset) and last (offset) timesteps identified by each method to compute the estimation error.
 
@@
 
+The TFCE performs worse than the cluster-sum approach, which was anticipated by @rousselet_using_2025 based on the initial results of @smith2009...
+
 ## Limitations and future directions
@@
 The error properties depend on the threshold parameter, a value of 10 or 20 seems to be a reasonable default, but the optimal threshold parameter can be adjusted using split-half reliability assessment...
-
-Can be applied to any 1D timeseries (e.g., pupillometry, electromyography)... Extending the approach to spatiotemporal data (i.e., time + sensors)...
+also depends on `k`...
+
+Can be applied to any 1D timeseries (e.g., pupillometry, electromyography)... Extending the approach to spatiotemporal data (i.e., time + sensors) or spatiotemporal time-frequency 4D data...
 
@@
 
-```{r fig-sim-timegen, echo = FALSE, dpi = 300, fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
+```{r fig-sim-timegen, echo = FALSE, dpi = 300, out.width = "75%", fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
 # number of participants
@@
 # generating participant-specific cross-temporal generalisation matrices
-timegen_data <- tibble()
+# timegen_data <- tibble()
+timegen_data <- data.frame()
 
@@
         offset_factor <- 1 /
-            (1 + exp((t1 - (decode_offset + offset_variability)) / sigma_smooth) )
+            (1 + exp((t1 - (decode_offset + offset_variability) ) / sigma_smooth) )
 
@@
     auc ~ t2(train_time, test_time, bs = "tp", k = 20),
+    # auc ~ t2(train_time, test_time, bs = "tp", k = 30),
     data = timegen_data,
@@
     file = "models/timegen_gam_t2.rds"
+    # file = "models/timegen_gam_t2_k30.rds"
     )
@@
 \newpage
-
-```{=html}
-<!--
-
-# Threshold-free cluster enhancement
-
-Cluster-based permutation approaches require defining a cluster-forming threshold (e.g., a t- or f-value) as the initial step of the algorithm.
-As different cluster-forming thresholds lead to clusters with different spatial or temporal extent, this initial threshold modulates the sensitivity of the subsequent permutation test.
-The threshold-free cluster enhancement method (TFCE) was introduced by @smith2009 to overcome this choice of an arbitrary threshold.
-
-In brief, the TFCE method works as follows.
-Instead of picking an arbitrary cluster-forming threshold (e.g., $t=2$), we try all (or many) possible thresholds in a given range and check whether a given timestep/voxel belongs to a significant cluster under any of the set of thresholds.
-Then, instead of using cluster mass (e.g., the sum of squared t-values within the cluster), we use a weighted average between the cluster extend ($e$, how broad is the cluster, that is, how many connected samples it contains) and the cluster height ($h$, how high is the cluster, that is, how large is the test statistic).
-The TFCE score at each timestep/voxel $t$ is given by:
-
-$$
-\text{TFCE(t)} = \int_{h=h_{0}}^{h_{t}} e(h)^{E} h^{H} \mathrm{d}h
-$$
-
-where $h_{0}$ is typically $0$ and parameters $E$ and $H$ are set a priori (typically to 0.5 and 2, respectively) and control the influence of the extend and height on the TFCE.
-Then, p-value for timestep/voxel $t$ is computed by comparing it TFCE with the null distribution of TFCE values.
-For each permuted signal, we keep the maximal value over the whole signal for the null distribution of the TFCE....
-But see @sassenhagen2019...
-
-\newpage
-
--->
-```
 
@@
 
-# Integration with `MNE-Python`
+# `R` package and integration with `MNE-Python`
 
 For users who are already familiar with `brms`, the recommended pipeline is to import ERPs or decoding results in `R` and analyse these data using the code provided in the main paper.
-However, it is also possible to simply call functions from the `neurogam R` package (available at <https://github.com/lnalborczyk/neurogam>), which come with sensible defaults.
+However, it is also possible to call functions from the `neurogam R` package (available at <https://github.com/lnalborczyk/neurogam>), which come with sensible defaults.
 
@@
 # using the testing_through_time() function from the neurogam package
+# this may take a few minutes (or hours depending the machine's performance
+# and data size)...
 gam_onset_offset <- testing_through_time(
+    # dataframe with M/EEG data in long format
     data = raw_df,
-    threshold = 10,
+    # threshold for defining clusters (20 by default)
+    threshold = 20,
+    # the *_id arguments are use to specify the relevant columns in data 
     participant_id = "participant", meeg_id = "eeg",
     time_id = "time", predictor_id = "condition",
-    warmup = 1000, iter = 5000, chains = 4, cores = 4
+    # number of warmup MCMC iterations
+    warmup = 1000,
+    # total number of MCMC iterations
+    iter = 5000,
+    # number of MCMCs
+    chains = 4,
+    # number of parallel cores to use for running the MCMCs
+    cores = 4
     )
@@
 The `neurogam` package can also be called from `Python` using the `rpy2` module, and can easily be integrated into `MNE-Python` pipelines.
-For example, we use it below to estimate the onset and offset of effects for one channel from a MNE evoked object.
+For example, we use it below to estimate the onset and offset of effects for one EEG channel from a MNE evoked object.
 Note that the code used to reshape the `sample MNE` dataset is provided later, and we refer to the [MNE documentation](https://mne.tools/stable/auto_tutorials/epochs/50_epochs_to_data_frame.html) about converting `MNE` epochs to `Pandas` dataframes in long format.

### 2025-04-10 brms_meeg.tex
updating the manuscript (intro)

@@
 conventional approach involves mass-univariate statistical testing
-across time, followed by corrections for multiple comparisons (e.g.,
-FDR, FWER) or cluster-based inference. While effective for controlling
-error rates, cluster-based methods come with a significant limitation:
-they shift the focus of inference from individual time points to
-clusters, making it difficult to draw precise conclusions about the
-onset or offset of observed effects. Here, we introduce a
-\emph{model-based} approch for analysing M/EEG timeseries such as
-event-related potentials (ERPs) or decoding performance over time. Our
-method leverages Bayesian generalised additive multilevel models
-(GAMMs), providing posterior probabilities that an effect is above zero
-(or above chance) at each time point, while naturally accounting for
-temporal dependencies and between-subject variability. Using both
-simulated and actual M/EEG datasets, we demonstrate that this approach
-substantially outperforms conventional methods in estimating the onset
-and offset of neural effects, yielding more precise and reliable
-results. We also provide an R package implementing the method and show
-how it can be seamlessly integrated into M/EEG analysis pipelines using
-MNE-Python.}
-
-\keywords{EEG, MEG, generalised additive models, mixed-effects
+across time and space, followed by corrections for multiple comparisons
+such as cluster-based inference. While effective for controlling error
+rates at the cluster-level, cluster-based inference comes with a
+significant limitation: by shifting the focus of inference from
+individual time points to clusters, it makes difficult to draw precise
+conclusions about the onset or offset of observed effects. Here, we
+introduce a \emph{model-based} approach for analysing M/EEG timeseries
+such as event-related potentials (ERPs) or decoding performance over
+time. Our method leverages Bayesian generalised additive multilevel
+models (GAMMs), providing posterior probabilities that an effect is
+above zero (or above chance) at each time point, while naturally
+accounting for temporal dependencies and between-subject variability.
+Using both simulated and actual M/EEG datasets, we demonstrate that this
+approach substantially outperforms conventional methods in estimating
+the onset and offset of neural effects, yielding more precise and
+reliable results. We provide an R package implementing the method and
+show how it can be seamlessly integrated into M/EEG analysis pipelines
+using MNE-Python.}
+
+\keywords{EEG, MEG, cluster-based inference, simulation, multiple
+comparisons, generalised additive models, mixed-effects
 models, multilevel models, Bayesian statistics, brms}
@@
 
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{longtable}
+\usepackage{colortbl}
+\usepackage{array}
+\usepackage{anyfontsize}
+\usepackage{multirow}
 \usepackage{mathtools}
+\usepackage{booktabs, caption, longtable, colortbl, array}
 \makeatletter
@@
 
+\subsection{Problem statement}\label{problem-statement}
+
 Understanding the temporal dynamics of cognitive processes requires
 methods that can capture fast-changing neural activity with high
-temporal resolution. Magnetoencephalography (MEG) and
-electroencephalography (EEG) are two such methods, widely used in
-cognitive neuroscience for their ability to track brain activity at the
-millisecond scale. These techniques provide rich time series data that
-reflect how neural responses unfold in response to stimuli or tasks. A
-central goal in many M/EEG studies is to determine whether and when
-neural responses differ across experimental conditions or participant
-groups.
-
-Here are some useful references to be discussed
-(\citeproc{ref-combrisson_exceeding_2015}{Combrisson \& Jerbi, 2015};
-\citeproc{ref-ehinger_unfold_2019}{Ehinger \& Dimigen, 2019};
-\citeproc{ref-frossard2021}{Frossard \& Renaud, 2021},
-\citeproc{ref-frossard2022}{2022}; \citeproc{ref-gramfort2013}{Gramfort,
-2013}; \citeproc{ref-hayasaka_validating_2003}{Hayasaka, 2003};
-\citeproc{ref-luck_how_2017}{Luck \& Gaspelin, 2017};
+temporal resolution. Magnetoencephalography and electroencephalography
+(M/EEG) are two such methods, widely used in cognitive neuroscience for
+their ability to track brain activity at the millisecond scale. These
+techniques provide rich time series data that reflect how neural
+responses unfold in response to stimuli or tasks. A central goal in many
+M/EEG studies is to determine whether, when, and where neural responses
+differ across experimental conditions or participant groups.
+
+The conventional approach involves mass-univariate statistical testing
+with a unique threshold, followed by corrections for multiple
+comparisons with the goal of maintaining the familywise error rate
+(FWER) or false discovery rate (FDR) at the nominal level. Cluster-based
+inference is the most common way of achieving this sort of error control
+in the M/EEG literature, being the default option in several software
+programs. While effective for controlling error rates, cluster-based
+inference comes with a significant limitation: they shift the focus of
+inference from individual time points to clusters, making it difficult
+to draw precise conclusions about the onset or offset of observed
+effects (\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
+\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}). As
+pointed by Maris \& Oostenveld (\citeproc{ref-maris2007}{2007}); ``there
+is a conflict between this interest in localized effects and our choice
+for a global null hypothesis: by controlling the FA {[}false alarm{]}
+rate under this global null hypothesis, one cannot quantify the
+uncertainty in the spatiotemporal localization of the effect''. Even
+worse, as noted by Rosenblatt et al.
+(\citeproc{ref-rosenblatt2018}{2018}), cluster-based inference suffers
+from low spatial resolution: ``Since discovering a cluster means that
+`there exists at least one voxel with an evoked response in the
+cluster', and not that `all the voxels in the cluster have an evoked
+response', it follows that the larger the detected cluster, the less
+information we have on the location of the activation.'' As a
+consequence, cluster-based inferenc has been shown\ldots{}
+
+To overcome the limitations of cluster-based inference, we introduce a
+novel model-based approach for precisely localising M/EEG effects in
+time (space, and other dimensions). The proposed approach, based on
+Bayesian generalised additive multilevel models, allows quantifying the
+posterior probability of effects being above chance at the level of
+timesteps, sensors, voxels, etc., while naturally taking into account
+spatiotemporal dependencies present in M/EEG timeseries. We compare the
+performance of the proposed approach to well-established alternative
+methods using both simulated and actual M/EEG data and show that it
+significantly outperforms alternative methods in estimating the onset
+and offset of M/EEG effects.
+
+\subsection{Cluster-based inference}\label{cluster-based-inference}
+
+Different methods exist to control the family-wiser error rate (FWER),
+defined as the type-1 error rate (false positive) over an ensemble
+(family) of tests\ldots{} for instance the Bonferroni correction
+(\citeproc{ref-dunn1961}{Dunn, 1961}), however this method is generally
+overconservative as it assumes statistical independence of tests, an
+assumptions that is clearly violated in the context of M/EEG timeseries
+with spatiotemporal dependencies\ldots{} or the false discovery rate
+(FDR), defined as the proportion of false positive among positive tests
+(e.g., \citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
+\citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001})\ldots{}
+
+On popular technique to account for spatiotemporal dependencies while
+controlling the FWER is cluster-based inference. Description of
+cluster-based approaches (see \citeproc{ref-sassenhagen2019}{Sassenhagen
+\& Draschkow, 2019})\ldots{} A typical cluster-based inference consists
+of two successive steps. First, clusters are defined as sets of
+contiguous voxels, channels, and/or timesteps, whose intensity/activity
+exceeds some predefined threshold. Clusters are then characterised by
+their height (i.e., maximal value), extent (number of constituent
+elements), or some combination of both, such at its ``mass'', for
+instance by summing the statistics within a cluster, an approach
+refereed to as ``cluster mass'' (\citeproc{ref-maris2007}{Maris \&
+Oostenveld, 2007}; \citeproc{ref-pernet2015}{Pernet et al., 2015}).
+Then, the null hypothesis is tested by assessing the probability.. As
+different cluster-forming thresholds lead to clusters with different
+spatial or temporal extent, this initial threshold modulates the
+sensitivity of the subsequent permutation test. The threshold-free
+cluster enhancement method (TFCE) was introduced by S. Smith \& Nichols
+(\citeproc{ref-smith2009}{2009}) to overcome this choice of an arbitrary
+threshold.
+
+In brief, the TFCE method works as follows. Instead of picking an
+arbitrary cluster-forming threshold (e.g., \(t=2\)), we try all (or
+many) possible thresholds in a given range and check whether a given
+timestep/voxel belongs to a significant cluster under any of the set of
+thresholds. Then, instead of using cluster mass (e.g., the sum of
+squared t-values within the cluster), we use a weighted average between
+the cluster extend (\(e\), how broad is the cluster, that is, how many
+connected samples it contains) and the cluster height (\(h\), how high
+is the cluster, that is, how large is the test statistic). The TFCE
+score at each timestep/voxel \(t\) is given by:
+
+\[
+\text{TFCE(t)} = \int_{h=h_{0}}^{h_{t}} e(h)^{E} h^{H} \mathrm{d}h
+\]
+
+where \(h_{0}\) is typically \(0\) and parameters \(E\) and \(H\) are
+set a priori (typically to 0.5 and 2, respectively) and control the
+influence of the extend and height on the TFCE. Then, p-value for
+timestep/voxel \(t\) is computed by comparing it TFCE with the null
+distribution of TFCE values. For each permuted signal, we keep the
+maximal value over the whole signal for the null distribution of the
+TFCE. The TFCE combined with permutation (assuming a large enough number
+of permutations) has been shown to provide accurate type 1 FWER (e.g.,
+\citeproc{ref-pernet2015}{Pernet et al., 2015}). However, previous
+simulation work showed that cluster-based methods have indeed poor
+performance in localising the onset of M/EEG effects (e.g.,
+\citeproc{ref-rousselet_using_2025}{Rousselet, 2025};
+\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}).
+
+To sum up this section, cluster-based inference main limitations is that
+it provides inference at the cluster level only, not allowing inference
+at the level of timesteps, sensors, etc. Thus, it does not allow
+inferring the localisation of effects (\citeproc{ref-maris2011}{Maris,
+2011}; for more details on cluster-based inference, see for instance
 \citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
-\citeproc{ref-pernet2015}{Pernet et al., 2015})\ldots{} See also Maris
-(\citeproc{ref-maris2011}{2011}) and Rosenblatt et al.
-(\citeproc{ref-rosenblatt2018}{2018}) (history of cluster-based
-approaches)\ldots{} Clusters failures (\citeproc{ref-eklund2016}{Eklund
-et al., 2016})\ldots{}
-
-\subsection{Problem statement}\label{problem-statement}
-
-Description of cluster-based approaches (see
-\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019})\ldots{}
-
-However, as pointed by the original authors themselves
-(\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007}), ``there is a
-conflict between this interest in localized effects and our choice for a
-global null hypothesis: by controlling the FA {[}false alarm{]} rate
-under this global null hypothesis, one cannot quantify the uncertainty
-in the spatiotemporal localization of the effect''
-(\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
-\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}). In
-contrast, the proposed approach, based on Bayesian generalised additive
-multilevel models, allows quantifying the posterior probability of
-effects being above chance at the level of timesteps, voxels, sensors
-(etc), while naturally taking into account spatiotemporal dependencies
-present in M/EEG timeseries.
-
-\subsection{Related work}\label{related-work}
+\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}). In the
+following, we briefly review previous modelling work. Then, we provide a
+brief introduction to Bayesian statistical modelling, before moving to
+generalised additive models (GAMs) and generalised additive multilevel
+models (GAMMs) to illustrate how they can be used to precisely localise
+the onset and offset of M/EEG effects.
+
+\subsection{Previous work on modelling M/EEG
+data}\label{previous-work-on-modelling-meeg-data}
 
@@
 \citeproc{ref-dunagan2024}{Dunagan et al., 2024})\ldots{}
+
+As put by Rousselet (\citeproc{ref-rousselet_using_2025}{2025}),
+concluding on the onset of effect based on a series of univariate
+tests\ldots{} commits to three fallacies\ldots{} here, we want to avoid
+these by introducing a model-based approach, which naturally take into
+account the temporal dependencies in the data to output a series of
+posterior probabilities\ldots{}
 
@@
 decoding performance is above chance level. However, this approach
-provides timeseries of BFs that ignores temporal dependencies..
-
-\subsection{Bayesian regression
-modelling}\label{bayesian-regression-modelling}
-
-Short intro/recap about Bayesian (linear and generalised) regression
-models\ldots{}
+provides timeseries of BFs that ignores temporal dependencies\ldots{}
 
@@
 terminology at this stage: \emph{splines} are functions composed of
-simpler functions. These simpler functions are basis functions (e.g., a
-cubic polynomial) and the set of basis functions is a \emph{basis}. Each
-basis function gets its coefficient and the resultant spline is the sum
-of these weighted basis functions. Splines coefficients are penalised
+simpler functions. These simpler functions are basis functions (e.g.,
+cubic polynomial, thin-plate) and the set of basis functions is a
+\emph{basis}. Each basis function gets its coefficient and the resultant
+spline is the sum of these weighted basis functions
+(Figure~\ref{fig-intro-gam}). Splines coefficients are penalised
 (usually through the squared of the smooth functions' second derivative)
@@
 
-Introduction to multilevel GAMs
+The Bayesian approach to statistical modelling is characterised by its
+reliance on probability theory to (\citeproc{ref-gelman2020}{Gelman et
+al., 2020})\ldots{} In this framework, all unknown entities are assigned
+probability distributions reflecting the uncertainty\ldots{} These
+probability distributions are commonly referred to as ``priors'' and
+represent some state of knowledge about unknown quantities before seeing
+any data. There are debates among Bayesian practitioners as to whether
+prior distributions should encore subjective (personal) beliefs
+or\ldots{} but these debates are outside the scope of the present paper
+and we therefore the interested reader to dedicated work (e.g., XX;
+YY)\ldots{} In practice, weakly informative priors are often used as
+default priors in situations in which subjective priors are difficult to
+define/elicit\ldots{} Bayesian models are then fitted on empirical
+(actual or simulated) data to update prior states of knowledge to
+posterior states of knowledge using Bayes theorem, or in practise,
+sampling-based approximations of the posterior distribution\ldots{}
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-intro-gam}Different types of GAM(M)s. \textbf{A}:
+GAMs predictions are computed as the weigthed sum (in black) of basis
+functions (here thin-plate basis functions, in grey). \textbf{B}:
+Constant-effect GAM, with 5 participants in colours and the group-level
+prediction in black. \textbf{C}: Varying-intercept + varying-slope GAMM
+(with common smoother). \textbf{D}: Varying-intercept + varying-slope +
+varying-smoother GAMM. In this model, each participant gets its own
+intercept, slope, and degree of `wiggliness' (smoother).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-intro-gam-1.png}
+
+}
+
+\end{figure}%
+
+See Figure~\ref{fig-intro-gam}\ldots{} Introduction to multilevel GAMs
 (\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al.,
-2019})\ldots{}
-
-Now describe the Bayesian GAMM (\citeproc{ref-miller2025}{Miller,
-2025})\ldots{} Proper inclusion of varying/random effects in the model
-specification protects against overly wiggly curves
-(\citeproc{ref-baayen2020}{Baayen \& Linke, 2020})\ldots{} Generalising
-to scale and shape or ``distributional GAMs''
-(\citeproc{ref-rigby2005}{Rigby \& Stasinopoulos, 2005};
+2019})\ldots{} Now describe the Bayesian GAMM
+(\citeproc{ref-miller2025}{Miller, 2025})\ldots{} Proper inclusion of
+varying/random effects in the model specification protects against
+overly wiggly curves (\citeproc{ref-baayen2020}{Baayen \& Linke,
+2020})\ldots{} Generalising to scale and shape or ``distributional
+GAMs'' (\citeproc{ref-rigby2005}{Rigby \& Stasinopoulos, 2005};
 \citeproc{ref-umlauf2018}{Umlauf et al., 2018}) and applied to
@@
 precisely identify the onset and offset of M/EEG effects (e.g., ERPs,
-decoding performance), we aimed to provide a model-based approach for
+decoding performance), we developed a model-based approach for
 estimating the onset and offset of such effects. To achieve this, we
 leveraged Bayesian generalised additive multilevel models (BGAMMs)
-fitted via the \texttt{brms\ R} package and compared the performance of
-this approach to conventional methods on both simulated and actual M/EEG
-data.
+fitted in \texttt{R} via the \texttt{brms} package and compared the
+performance of this approach to conventional methods on both simulated
+and actual M/EEG data.
 
@@
 Rousselet (\citeproc{ref-rousselet_using_2025}{2025}), the signal was
-generated from truncated Gaussian with an objective onset at 160 ms, a
-peak at 250 ms, and an offset at 342 ms. We simulated this signal for
-250 timesteps between 0 and 0.5s, akin to a 500 Hz sampling rate. We
-simulated data for a group of 20 participants with 50 trials per
-participant and condition (Figure~\ref{fig-eeg}).
+generated from a truncated Gaussian distribution with an objective onset
+at 160 ms, a peak at 250 ms, and an offset at 342 ms. We simulated this
+signal for 250 timesteps between 0 and 0.5s, akin to a 500 Hz sampling
+rate. We simulated data for a group of 20 participants (with variable
+true onset) with 50 trials per participant and condition
+(Figure~\ref{fig-eeg}).
 
@@
 
-\subsection{Model fitting}\label{model-fitting}
-
-We then fitted a Bayesian GAM using the \texttt{brms} package
+\subsection{Model description and model
+fitting}\label{model-description-and-model-fitting}
+
+We then fitted a Bayesian GAM (BGAM) using the \texttt{brms} package
 (\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018};
@@
 
-\CommentTok{\# timing the model fitting}
-\CommentTok{\# around 600s for 2000 iterations on 4 chains and 4 cores}
-\CommentTok{\# around 190s for 2000 iterations on 4 chains and 4 cores + cmdstanr and 01}
-\CommentTok{\# around 440s for 2000 iterations on 4 chains and 4 cores + threading(2), cmdstanr and 01}
-\CommentTok{\# library(tictoc)}
-\CommentTok{\# tic()}
-
-\CommentTok{\# fitting the GAM}
+\CommentTok{\# fitting the BGAM}
 \NormalTok{gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
@@
     \AttributeTok{file =} \StringTok{"models/gam.rds"}
-    \DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
-    \CommentTok{\# speeding up the model fitting \#}
-    \DocumentationTok{\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#}
-    \CommentTok{\# backend = "cmdstanr",}
-    \CommentTok{\# threads = threading(2),}
-    \CommentTok{\# stan\_model\_args = list(stanc\_options = list("O1") )}
 \NormalTok{    )}
-
-\CommentTok{\# stop timing}
-\CommentTok{\# toc()}
 \end{Highlighting}
@@
 
-We depict the posterior predictions together with the posterior estimate
-of the slope for \texttt{condition} at each timestep
-(Figure~\ref{fig-plot-post-slope}).
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-plot-post-slope}Posterior estimate of the EEG
-activity in each condition (left) and posterior estimate of the
-difference in EEG activity (right) according to the GAM.}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-plot-post-slope-1.png}
-
-}
-
-\end{figure}%
-
-We then compute the posterior probability of the slope for
-\texttt{condition} being above \(0\) (Figure~\ref{fig-post-prob-ratio},
-left). We can also express this as the ratio of posterior probabilities
-(i.e., \(p/(1-p)\)) and visualise the timecourse of this ratio
-superimposed with the conventional thresholds on evidence ratios
-(Figure~\ref{fig-post-prob-ratio}, right). A ratio of 10 means that the
-probability of the difference being above 0 is 10 times higher than the
-probability of the difference not being above 0, given the data, the
-priors, and other model's assumptions.
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-post-prob-ratio}Left: Posterior probability of the
-EEG difference (slope) being above 0 according to the GAM. Right: Ratio
-of posterior probability according to the GAM (on a log10 scale).
-Timesteps above threshold (10) are highlighted in green.}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.png}
-
-}
-
-\end{figure}%
-
-\newpage
-
-\subsection{Multilevel modelling using summary
-statistics}\label{multilevel-modelling-using-summary-statistics}
-
-The previous model only included constant (fixed) effects, thus not
-properly accounting for between-participant variability. We next fit a
-multilevel version of the GAM. Although it is possible to fit a GAMM at
-the single-trial level, we present a resource-efficient version of the
-model that is fitted directly on the by-participant summary statistics
-(mean and SD), similar to what is done in meta-analysis.
+However, the previous model only included constant (fixed) effects, thus
+not properly accounting for between-participant variability. We next fit
+a multilevel version of the BGAM (BGAMM). Although it is possible to fit
+a BGAMM at the single-trial level, we present a resource-efficient
+version of the model that is fitted directly on the by-participant
+summary statistics (mean and SD), similar to what is done in
+meta-analysis.
 
@@
 
-\CommentTok{\# fitting the GAM}
+\CommentTok{\# fitting the BGAMM}
 \NormalTok{meta\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
@@
 
+We depict the posterior predictions together with the posterior estimate
+of the slope for \texttt{condition} at each timestep
+(Figure~\ref{fig-plot-post-slope}). This figure suggests that the BGAMM
+provides an adequate description of the simulated data.
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-plot-post-slope}Posterior estimate of the EEG
+activity in each condition (left) and posterior estimate of the
+difference in EEG activity (right) according to the BGAMM.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-plot-post-slope-1.png}
+
+}
+
+\end{figure}%
+
+We then compute the posterior probability of the slope for
+\texttt{condition} being above \(0\) (Figure~\ref{fig-post-prob-ratio},
+left). We can also express this as the ratio of posterior probabilities
+(i.e., \(p/(1-p)\)) and visualise the timecourse of this ratio
+superimposed with the conventional thresholds on evidence ratios
+(Figure~\ref{fig-post-prob-ratio}, right). Note that a ratio of 10 means
+that the probability of the difference being above 0 is 10 times higher
+than the probability of the difference not being above 0, given the
+data, the priors, and other model's assumptions. Thresholding the
+posterior probability ratio thus provides a model-based approach for
+estimating the onset and offset of M/EEG effects.
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-post-prob-ratio}Left: Posterior probability of the
+EEG difference (slope) being above 0 according to the BGAMM. Right:
+Ratio of posterior probability according to the BGAMM (on a log10
+scale). Timesteps above threshold (10) are highlighted in green.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.png}
+
+}
+
+\end{figure}%
+
+\newpage
+
 \subsection{Error properties of the proposed
@@
 
-We then compute the difference between the true and estimated
-onset/offset of the EEG difference according to various \texttt{k} (GAM
-basis dimension) and \texttt{threshold} values. Remember that the EEG
-signal was generated from a truncated Gaussian with an objective onset
-at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
-Figure~\ref{fig-onset-error} shows that the multilevel GAM can
-\emph{almost exactly} recover the true onset and offset values, given
-some reasonable choice of \texttt{k} and \texttt{threshold} values
-(e.g., a threshold of 20). We provide more detailed recommendations on
-how to set \texttt{k} in Section~\ref{sec-basis}.
+We then assess the performance of the proposed approach by computing the
+difference between the true and estimated onset/offset of the EEG
+difference according to various \texttt{k} (BGAM basis dimension) and
+\texttt{threshold} values. Remember that the EEG signal was generated
+from a truncated Gaussian with an objective onset at 160 ms, a maximum
+at 250 ms, and an offset at 342 ms. Figure~\ref{fig-onset-error} shows
+that the multilevel GAM can almost exactly recover the true onset and
+offset values, given some reasonable choice of \texttt{k} and
+\texttt{threshold} values. We provide more detailed recommendations on
+how to set \texttt{k} in Section~\ref{sec-basis}. This figure further
+reveals that the optimal \texttt{k} and \texttt{threshold} values may
+differ for the onset and offset values, and there there seems to exist a
+trade-off between these two parameters: lower \texttt{k} values lead to
+poorer estimations, but these poor estimations can be compensated (only
+to some extent) by higher \texttt{threshold} values (and reciprocally).
 
@@
 
-\caption{\label{fig-onset-error}Average estimation error for the onset
-(left) and offset (right) according to various basis dimension and
-threshold values (according to the multilevel GAM).}
+\caption{\label{fig-onset-error}Average estimation error (RMSE) for the
+onset (left) and offset (right) according to various basis dimension and
+threshold values for the BGAM (computed from 100 simulated datasets).}
 
@@
 
-We then compared the ability of the GAMM to correctly estimate the onset
-and offset of the ERP difference to other widely-used methods. First, we
-conducted mass-univariate t-tests (thus treating each timestep
+We then compared the ability of the BGAMM to correctly estimate the
+onset and offset of the ERP difference to other widely-used methods.
+First, we conducted mass-univariate t-tests (thus treating each timestep
 independently) and identified the onset and offset of the ERP difference
@@
 estimated by each method on a single simulated dataset and shows that
-all methods over-estimate the true onset and under-estimate the true
-offset.
+all methods systematically overestimate the true onset and underestimate
+the true offset.
 
@@
 onsets/offsets identified using the raw p-values, the corrected p-values
-(BH95, BY01, Holm), the cluster-based methods (Cluster-mass, TFCE), or
+(BH95, BY01, Holm), the cluster-based methods (Cluster mass, TFCE), or
 using the binary segmentation method (Change point).}
@@
 
-The various methods were compared using the median error (i.e.,
-true-estimated) and variance of onset/offset estimates computed on
-10.000 simulated datasets.
+To assess the accuracy of group-level onset estimation, the various
+methods were compared using the bias (i.e., median(estimated-true)),
+median absolute error (MAE), root mean square error (RMSE), variance,
+and median absolute deviation (MAD) of onset/offset estimates computed
+on 1000 simulated datasets. As in Rousselet
+(\citeproc{ref-rousselet_using_2025}{2025}), each participant was
+assigned a random onset betwee, 150 and 170ms.
 
@@
 
-To provide a more exhaustive assessment of these methods, we also
-evaluated the performance of the proposed approach on actual MEG data
-(decoding results from \citeproc{ref-nalborczyk:inprep}{Nalborczyk et
-al., in preparation}). In this study, we conducted time-resolved
-multivariate pattern analysis (MVPA, also known as decoding) of MEG data
-during reading tasks. As a result, we obtain a timecourse of decoding
-performance (ROC AUC), bounded between 0 and 1, for each participant
-(for a total of 32 participants). Next, we want to \emph{test} whether
-the group-level average decoding accuracy is above chance (i.e., 0.5) at
-each timestep (Figure~\ref{fig-decoding-data}). To achieve this, we
-fitted a similar GAM as discussed previously, but we replaced the
-\(\mathrm{Normal}\) likelihood function by a \(\mathrm{Beta}\) one to
-account for the bounded nature of AUC values (between 0 and 1) (for a
-tutorial on Beta regression, see \citeproc{ref-coretta2025}{Coretta \&
-Bürkner, 2025}).
+To complement the simulation study, we evaluated the performance of the
+various methods on actual MEG data (decoding results from
+\citeproc{ref-nalborczyk:inprep}{Nalborczyk et al., in preparation}). In
+this study, we conducted time-resolved multivariate pattern analysis
+(MVPA, also known as decoding) of MEG data during reading tasks. As a
+result, we obtain a timecourse of decoding performance (ROC AUC),
+bounded between 0 and 1, for each participant (for a total of 32
+participants). Next, we wanted to \emph{test} whether the group-level
+average decoding accuracy is above chance (i.e., 0.5) at each timestep
+(Figure~\ref{fig-decoding-data}). To achieve this, we fitted a BGAM as
+introduced previously, but we replaced the \(\mathrm{Normal}\)
+likelihood function by a \(\mathrm{Beta}\) one to account for the
+bounded nature of AUC values (between 0 and 1) (for a tutorial on Beta
+regression, see \citeproc{ref-coretta2025}{Coretta \& Bürkner, 2025}).
 
@@
 the region of practical equivalence (ROPE, in orange) computed from the
-baseline period. The blue horizontal markers indicate the timesteps at
-which the posterior probability ratio exceeds 20.}
+baseline period (data from \citeproc{ref-nalborczyk:inprep}{Nalborczyk
+et al., in preparation}). The blue horizontal markers indicate the
+timesteps at which the posterior probability ratio exceeds 20.}
 
@@
 results, revealing that the proposed approach (\texttt{BGAM}) has the
-lowest median error and variance for both the onset and offset
-estimates. The \texttt{Cluster-mass}, \texttt{Holm}, and
-\texttt{Change\ point} also have good performance, but surprisingly, the
-\texttt{TFCE} method has relatively bad performance for estimating the
-effect offset. Unsurprisingly, the \texttt{Raw\ p-value},
-\texttt{FDR\ BH95}, and \texttt{FDR\ BY01} methods show the worst
-performance.
+lowest median absolute error (MAE) and variance for both the onset and
+offset estimates. The \texttt{Cluster\ mass} and \texttt{Change\ point}
+also have good performance, but surprisingly, the \texttt{TFCE} method
+has relatively bad performance for estimating the effect offset (similar
+performance to the \texttt{Holm} and \texttt{FDR\ BY01} methods).
+Unsurprisingly, the \texttt{Raw\ p-value} and \texttt{FDR\ BH95} methods
+show the worst performance.
 
@@
 
-\caption{\label{fig-simulation-results}Median error and variance of
-onset and offset estimates for each method. The vertical dashed line
-represents an unbiased estimate of the true onset/ofsset. Variance is
+\caption{\label{fig-simulation-results}Median absolute error and
+variance of onset and offset estimates for each method. Variance is
 plotted on a log10 scale for visual purposes.}
@@
 \end{figure}%
+
+Results are further summarised in Table~\ref{tbl-simulation-results},
+which shows that the \texttt{BGAM} is perfectly unbiased (i.e., it has a
+bias of 0s) for the onset and almost exactly unbiased for the onset
+(with a bias of approximately -2ms). The \texttt{Bias} column shows that
+all other methods tend to estimate the onset later than the true onset
+and to estimate the offet earlier than the true offset. As can be seen
+from this table, the \texttt{BGAM} has the best performance on all
+included metrics.
+
+\begin{table}
+
+{\caption{{Summary statistics of the onset and offset estimates for each
+method (ordered by MAE).}{\label{tbl-simulation-results}}}}
+
+\fontsize{9.0pt}{10.8pt}\selectfont
+\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lccccc}
+\toprule
+ & Bias & MAE & RMSE & Variance & MAD \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{6}{l}{onset} \\[2.5pt] 
+\midrule\addlinespace[2.5pt]
+BGAM (new method) & 0.0000 & 0.0020 & 0.0001 & 0.0000 & 0.0030 \\ 
+Cluster mass & 0.0060 & 0.0060 & 0.0057 & 0.0000 & 0.0044 \\ 
+Change point & 0.0060 & 0.0060 & 0.0051 & 0.0000 & 0.0059 \\ 
+FDR BY01 & 0.0120 & 0.0120 & 0.0443 & 0.0087 & 0.0059 \\ 
+TFCE & 0.0140 & 0.0140 & 0.0270 & 0.0019 & 0.0059 \\ 
+Holm & 0.0140 & 0.0140 & 0.0270 & 0.0019 & 0.0059 \\ 
+FDR BH95 & 0.0080 & 0.0460 & 0.0599 & 0.0192 & 0.0801 \\ 
+Raw p-value & 0.0060 & 0.1250 & 0.0711 & 0.0252 & 0.1942 \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{6}{l}{offset} \\[2.5pt] 
+\midrule\addlinespace[2.5pt]
+BGAM (new method) & -0.0020 & 0.0040 & 0.0021 & 0.0000 & 0.0030 \\ 
+Cluster mass & -0.0080 & 0.0080 & 0.0084 & 0.0000 & 0.0059 \\ 
+Change point & -0.0100 & 0.0100 & 0.0096 & 0.0000 & 0.0059 \\ 
+FDR BY01 & -0.0140 & 0.0150 & 0.0204 & 0.0056 & 0.0059 \\ 
+TFCE & -0.0180 & 0.0180 & 0.0261 & 0.0016 & 0.0030 \\ 
+Holm & -0.0180 & 0.0180 & 0.0262 & 0.0016 & 0.0030 \\ 
+FDR BH95 & -0.0100 & 0.0360 & 0.0578 & 0.0189 & 0.0682 \\ 
+Raw p-value & -0.0080 & 0.1180 & 0.0725 & 0.0252 & 0.1868 \\ 
+\bottomrule
+\end{tabular*}
+
+\end{table}
 
@@
 Figure~\ref{fig-onset-offset} shows the group-level average decoding
-performance through time with onset and offset estimates for each
+performance through time with onset and offset estimates from each
 method. Overall, this figure shows that both the \texttt{Raw\ p-value}
@@
 to result in fewer clusters.\footnote{It should be noted that although
-  each method can produce several ``clusters'' of ``significant''
-  timesteps, we only considered the first (onset) and last (offset)
-  timesteps identified by each method to compute the estimation error.}
+  each method can produce several ``clusters'' of timesteps, we only
+  considered the first (onset) and last (offset) timesteps identified by
+  each method to compute the estimation error.}
 
@@
 
+The TFCE performs worse than the cluster-sum approach, which was
+anticipated by Rousselet (\citeproc{ref-rousselet_using_2025}{2025})
+based on the initial results of S. Smith \& Nichols
+(\citeproc{ref-smith2009}{2009})\ldots{}
+
 \subsection{Limitations and future
@@
 20 seems to be a reasonable default, but the optimal threshold parameter
-can be adjusted using split-half reliability assessment\ldots{}
+can be adjusted using split-half reliability assessment\ldots{} also
+depends on \texttt{k}\ldots{}
 
@@
 electromyography)\ldots{} Extending the approach to spatiotemporal data
-(i.e., time + sensors)\ldots{}
+(i.e., time + sensors) or spatiotemporal time-frequency 4D data\ldots{}
 
@@
 (\citeproc{ref-glue}{Hester \& Bryan, 2024}), grateful v. 0.2.11
-(\citeproc{ref-grateful}{Rodriguez-Sanchez \& Jackson, 2024}), knitr v.
-1.50 (\citeproc{ref-knitr2014}{Xie, 2014},
-\citeproc{ref-knitr2015}{2015}, \citeproc{ref-knitr2025}{2025}),
-MetBrewer v. 0.2.0 (\citeproc{ref-MetBrewer}{Mills, 2022}), neurogam v.
-0.0.1 (\citeproc{ref-neurogam}{Nalborczyk, 2025}), pakret v. 0.2.2
+(\citeproc{ref-grateful}{Rodriguez-Sanchez \& Jackson, 2024}), gt v.
+1.0.0 (\citeproc{ref-gt}{Iannone et al., 2025}), knitr v. 1.50
+(\citeproc{ref-knitr2014}{Xie, 2014}, \citeproc{ref-knitr2015}{2015},
+\citeproc{ref-knitr2025}{2025}), MetBrewer v. 0.2.0
+(\citeproc{ref-MetBrewer}{Mills, 2022}), neurogam v. 0.0.1
+(\citeproc{ref-neurogam}{Nalborczyk, 2025}), pakret v. 0.2.2
 (\citeproc{ref-pakret}{Gallou, 2024}), patchwork v. 1.3.0
@@
 (\citeproc{ref-scales}{Wickham et al., 2023}), scico v. 1.5.0
-(\citeproc{ref-scico}{T. L. Pedersen \& Crameri, 2023}), tidybayes v.
-3.0.7 (\citeproc{ref-tidybayes}{Kay, 2024}), tidytext v. 0.4.2
+(\citeproc{ref-scico}{T. L. Pedersen \& Crameri, 2023}), tictoc v. 1.2.1
+(\citeproc{ref-tictoc}{Izrailev, 2024}), tidybayes v. 3.0.7
+(\citeproc{ref-tidybayes}{Kay, 2024}), tidytext v. 0.4.2
 (\citeproc{ref-tidytext}{Silge \& Robinson, 2016}), tidyverse v. 2.0.0
@@
 
-\bibitem[\citeproctext]{ref-combrisson_exceeding_2015}
-Combrisson, E., \& Jerbi, K. (2015). Exceeding chance level by chance:
-{The} caveat of theoretical chance levels in brain signal classification
-and statistical assessment of decoding accuracy. \emph{Journal of
-Neuroscience Methods}, \emph{250}, 126--136.
-\url{https://doi.org/10.1016/j.jneumeth.2015.01.010}
-
 \bibitem[\citeproctext]{ref-coretta2025}
@@
 
-\bibitem[\citeproctext]{ref-ehinger_unfold_2019}
-Ehinger, B. V., \& Dimigen, O. (2019). Unfold: An integrated toolbox for
-overlap correction, non-linear modeling, and regression-based {EEG}
-analysis. \emph{PeerJ}, \emph{7}, e7838.
-\url{https://doi.org/10.7717/peerj.7838}
-
-\bibitem[\citeproctext]{ref-eklund2016}
-Eklund, A., Nichols, T. E., \& Knutsson, H. (2016). Cluster failure: Why
-fMRI inferences for spatial extent have inflated false-positive rates.
-\emph{Proceedings of the National Academy of Sciences}, \emph{113}(28),
-7900--7905. \url{https://doi.org/10.1073/pnas.1602413113}
+\bibitem[\citeproctext]{ref-dunn1961}
+Dunn, O. J. (1961). Multiple Comparisons among Means. \emph{Journal of
+the American Statistical Association}, \emph{56}(293), 52--64.
+\url{https://doi.org/10.1080/01621459.1961.10482090}
 
@@
 
-\bibitem[\citeproctext]{ref-frossard2021}
-Frossard, J., \& Renaud, O. (2021). Permutation Tests for Regression,
-ANOVA, and Comparison of Signals: The {\textbf{permuco}} Package.
-\emph{Journal of Statistical Software}, \emph{99}(15).
-\url{https://doi.org/10.18637/jss.v099.i15}
-
-\bibitem[\citeproctext]{ref-frossard2022}
-Frossard, J., \& Renaud, O. (2022). The cluster depth tests: Toward
-point-wise strong control of the family-wise error rate in massively
-univariate tests with application to M/EEG. \emph{NeuroImage},
-\emph{247}, 118824.
-\url{https://doi.org/10.1016/j.neuroimage.2021.118824}
-
 \bibitem[\citeproctext]{ref-gabry2019}
@@
 1383--1400. \url{https://doi.org/10.1016/j.neuroimage.2005.11.048}
-
-\bibitem[\citeproctext]{ref-hayasaka_validating_2003}
-Hayasaka, S. (2003). Validating cluster size inference: Random field and
-permutation methods. \emph{NeuroImage}, \emph{20}(4), 2343--2356.
-\url{https://doi.org/10.1016/j.neuroimage.2003.08.003}
 
@@
 
+\bibitem[\citeproctext]{ref-gt}
+Iannone, R., Cheng, J., Schloerke, B., Hughes, E., Lauer, A., Seo, J.,
+Brevoort, K., \& Roy, O. (2025). \emph{{gt}: Easily create
+presentation-ready display tables}.
+\url{https://CRAN.R-project.org/package=gt}
+
+\bibitem[\citeproctext]{ref-tictoc}
+Izrailev, S. (2024). \emph{{tictoc}: Functions for timing r scripts, as
+well as implementations of {``{Stack}''} and {``{StackList}''}
+structures}. \url{https://CRAN.R-project.org/package=tictoc}
+
 \bibitem[\citeproctext]{ref-tidybayes}
@@
 
-\bibitem[\citeproctext]{ref-luck_how_2017}
-Luck, S. J., \& Gaspelin, N. (2017). How to get statistically
-significant effects in any {ERP} experiment (and why you shouldn't).
-\emph{Psychophysiology}, \emph{54}(1), 146--157.
-\url{https://doi.org/10.1111/psyp.12639}
-
 \bibitem[\citeproctext]{ref-easystats}
@@
 of m/EEG effects with bayesian generalised additive multilevel models}.
+\url{https://github.com/lnalborczyk/neurogam}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-timegen-1.png}
+\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-timegen-1.png}
 
@@
 \NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
+    \CommentTok{\# auc \textasciitilde{} t2(train\_time, test\_time, bs = "tp", k = 30),}
     \AttributeTok{data =}\NormalTok{ timegen\_data,}
@@
     \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"}
+    \CommentTok{\# file = "models/timegen\_gam\_t2\_k30.rds"}
 \NormalTok{    )}
@@
 
-\section{\texorpdfstring{Integration with
-\texttt{MNE-Python}}{Integration with MNE-Python}}\label{integration-with-mne-python}
+\section{\texorpdfstring{\texttt{R} package and integration with
+\texttt{MNE-Python}}{R package and integration with MNE-Python}}\label{r-package-and-integration-with-mne-python}
 
@@
 these data using the code provided in the main paper. However, it is
-also possible to simply call functions from the \texttt{neurogam\ R}
-package (available at \url{https://github.com/lnalborczyk/neurogam}),
-which come with sensible defaults.
+also possible to call functions from the \texttt{neurogam\ R} package
+(available at \url{https://github.com/lnalborczyk/neurogam}), which come
+with sensible defaults.
 
@@
 \CommentTok{\# using the testing\_through\_time() function from the neurogam package}
+\CommentTok{\# this may take a few minutes (or hours depending the machine\textquotesingle{}s performance}
+\CommentTok{\# and data size)...}
 \NormalTok{gam\_onset\_offset }\OtherTok{\textless{}{-}} \FunctionTok{testing\_through\_time}\NormalTok{(}
+    \CommentTok{\# dataframe with M/EEG data in long format}
     \AttributeTok{data =}\NormalTok{ raw\_df,}
-    \AttributeTok{threshold =} \DecValTok{10}\NormalTok{,}
+    \CommentTok{\# threshold for defining clusters (20 by default)}
+    \AttributeTok{threshold =} \DecValTok{20}\NormalTok{,}
+    \CommentTok{\# the *\_id arguments are use to specify the relevant columns in data }
     \AttributeTok{participant\_id =} \StringTok{"participant"}\NormalTok{, }\AttributeTok{meeg\_id =} \StringTok{"eeg"}\NormalTok{,}
     \AttributeTok{time\_id =} \StringTok{"time"}\NormalTok{, }\AttributeTok{predictor\_id =} \StringTok{"condition"}\NormalTok{,}
-    \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{, }\AttributeTok{iter =} \DecValTok{5000}\NormalTok{, }\AttributeTok{chains =} \DecValTok{4}\NormalTok{, }\AttributeTok{cores =} \DecValTok{4}
+    \CommentTok{\# number of warmup MCMC iterations}
+    \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
+    \CommentTok{\# total number of MCMC iterations}
+    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
+    \CommentTok{\# number of MCMCs}
+    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
+    \CommentTok{\# number of parallel cores to use for running the MCMCs}
+    \AttributeTok{cores =} \DecValTok{4}
 \NormalTok{    )}
@@
 \texttt{MNE-Python} pipelines. For example, we use it below to estimate
-the onset and offset of effects for one channel from a MNE evoked
+the onset and offset of effects for one EEG channel from a MNE evoked
 object. Note that the code used to reshape the \texttt{sample\ MNE}

### 2025-04-10 code/errors_simulation_cluster.R

@@ -2,7 +2,7 @@
 # Monte-Carlo simulation of onset/offset error properties #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on April 7, 2025                           #
+# Last updated on April 9, 2025                           #
 ###########################################################
 
 # importing R packages
@@ -16,6 +16,10 @@ library(brms)
 # for setting up the cluster
 library(doParallel)
 library(foreach)
+
+# timing the simulations
+library(tictoc)
+tic()
 
 # total number of total CPU cores available on the HPC node
 total_cores <- 12
@@ -36,9 +40,6 @@ source("code/eeg_noise.R")
 # importing the ERP template with true onset = 160 ms, F = 81, and max at F = 126
 source("code/erp_template.R")
 
-# importing helper functions from Guillaume Rouselet
-# source("code/functions.R")
-
 # importing home-made helper functions
 source("code/utils.R")
 
@@ -46,7 +47,7 @@ source("code/utils.R")
 meanpower <- unlist(read.table("code/meanpower.txt") )
 
 # defining the total number of simulations to be performed
-nsims <- 100
+nsims <- 1000
 
 # defining simulation parameters
 n_trials <- 50 # number of trials
@@ -337,7 +338,10 @@ sim_results <- bind_rows(sim_results_list)
 
 # saving the simulation results
 # saveRDS(object = sim_results, file = "./results/errors_results_cluster.rds")
-saveRDS(object = sim_results, file = "./results/errors_results_100sims.rds")
+saveRDS(object = sim_results, file = "./results/errors_results_1000sims.rds")
 
 # stopping the cluster
 stopCluster(cl)
+
+# timing the simulations
+toc()

### 2025-04-10 code/generate_eeg_data.R

@@ -4,14 +4,8 @@ source("code/eeg_noise.R")
 # importing the ERP template with true onset = 160 ms, F=81, and max at F=126
 source("code/erp_template.R")
 
-# importing helper functions
-# source("code/functions.R")
-
 # to use with the eeg_noise function
 meanpower <- unlist(read.table("code/meanpower.txt") )
-
-# defining the number of simulations to be performed
-# nsims <- 100
 
 # defining simulation parameters
 n_trials <- 50 # number of trials
@@ -23,9 +17,6 @@ ronset <- seq(from = 150, to = 170, by = 2) # random onset for each participant
 # defining the true onset and offset in seconds
 true_onset <- 0.160
 true_offset <- 0.342
-
-# defining significance (alpha) thresholds
-# significance_thresholds <- c(0.05, 0.01, 0.005, 0.001)
 
 # defining a function to generate a dataframe
 generate_data <- function (n_trials, n_ppt, outvar, srate, ronset) {
@@ -59,7 +50,7 @@ generate_data <- function (n_trials, n_ppt, outvar, srate, ronset) {
             trial = c(rep(1:n_trials, each = length(Xf) ), rep(1:n_trials, each = length(Xf) ) ),
             condition = factor(rep(c("cond1", "cond2"), each = Nf * n_trials) ),
             participant = paste0("participant_", sprintf("%02d", P) )
-        ) %>%
+            ) %>%
             select(participant, condition, trial, x, y)
         
         # and appending it to previous participants

### 2025-04-10 code/meta_gam_errors_cluster.R

@@ -2,7 +2,7 @@
 # Monte-Carlo simulation of onset/offset error properties #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on April 7, 2025                           #
+# Last updated on April 9, 2025                           #
 ###########################################################
 
 # importing R packages
@@ -14,6 +14,10 @@ library(brms)
 # setting up parallel workers
 library(purrr)
 library(furrr)
+
+# timing the simulations
+library(tictoc)
+tic()
 
 # total number of total CPU cores available on the HPC node
 total_cores <- 12
@@ -49,7 +53,7 @@ single_iteration <- function (i) {
     temp_results <- data.frame()
     
     # number of posterior samples to use to compute the posterior probability ratio
-    n_post_samples <- 1e4
+    # n_post_samples <- 1e4
     
     # defining the parameter grid
     param_grid <- crossing(
@@ -75,7 +79,7 @@ single_iteration <- function (i) {
                 n_trials = n_trials, n_ppt = n_ppt, outvar = outvar,
                 srate = srate, ronset = ronset
                 )
-    
+            
             # averaging across participants
             summary_df <- raw_df %>%
                 summarise(
@@ -116,8 +120,8 @@ single_iteration <- function (i) {
                 )
             
             prob_y_above_0 <- error_model$data %>%
-                add_epred_draws(object = error_model, ndraws = n_post_samples) %>%
-                # add_epred_draws(object = error_model) %>%
+                # add_epred_draws(object = error_model, ndraws = n_post_samples) %>%
+                add_epred_draws(object = error_model) %>%
                 data.frame() %>%
                 dplyr::select(participant, time, condition, .epred, .draw) %>%
                 pivot_wider(names_from = condition, values_from = .epred) %>%
@@ -131,16 +135,16 @@ single_iteration <- function (i) {
                 mutate(
                     prob_ratio = ifelse(
                         is.infinite(prob_ratio),
-                        # ndraws(error_model),
-                        n_post_samples,
+                        ndraws(error_model),
+                        # n_post_samples,
                         prob_ratio
                         )
                     ) %>%
                 mutate(
                     prob_ratio = ifelse(
                         prob_ratio == 0,
-                        # 1 / ndraws(error_model),
-                        1 / n_post_samples,
+                        1 / ndraws(error_model),
+                        # 1 / n_post_samples,
[…]

### 2025-04-10 code/split_half_reliability_cluster.R

@@ -2,7 +2,7 @@
 # Split-half reliability of onset/offset estimates        #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on April 7, 2025                           #
+# Last updated on April 9, 2025                           #
 ###########################################################
 
 # importing R packages
@@ -16,6 +16,10 @@ library(brms)
 # for setting up the cluster
 library(doParallel)
 library(foreach)
+
+# timing the simulations
+library(tictoc)
+tic()
 
 # total number of total CPU cores available on the HPC node
 total_cores <- 12
@@ -64,10 +68,10 @@ unique_participants <- unique(raw_df$participant)
 n_participants <- as.numeric(length(unique_participants) )
 
 # defining the number of splits
-n_splits <- 100
+n_splits <- 1000
 
 # number of posterior samples to use to compute the posterior probability ratio
-n_post_samples <- 1e4
+# n_post_samples <- 1e4
 
 # creating a dataframe where each participant appears in every split
 split_df <- crossing(participant = unique_participants, split_number = 1:n_splits)
@@ -141,9 +145,9 @@ reliability_results <- tryCatch({foreach(
         message_parallel("\nModel fitted...")
         prob_y_above_0 <- data.frame(time = unique(gam_split$data$time) )|>
             # using a subset of posterior samples
-            add_epred_draws(object = gam_split, ndraws = n_post_samples) |>
+            # add_epred_draws(object = gam_split, ndraws = n_post_samples) |>
             # or using all posterior samples
-            # add_epred_draws(object = gam_split)|>
+            add_epred_draws(object = gam_split)|>
             # converting to dataframe
             data.frame()|>
             group_by(time)|>
@@ -151,21 +155,21 @@ reliability_results <- tryCatch({foreach(
             mutate(prob_ratio = m / (1 - m) )|>
             ungroup()|>
             # ensuring there is no 0 or +Inf values
-            # mutate(prob_ratio = ifelse(is.infinite(prob_ratio), ndraws(gam_split), prob_ratio) )|>
-            # mutate(prob_ratio = ifelse(prob_ratio == 0, 1 / ndraws(gam_split), prob_ratio) )
-            mutate(prob_ratio = ifelse(is.infinite(prob_ratio), n_post_samples, prob_ratio) )|>
-            mutate(prob_ratio = ifelse(prob_ratio == 0, 1 / n_post_samples, prob_ratio) ) 
[…]

### 2025-04-10 brms_meeg.qmd
pushing a cleaner version of the intro

@@
 Here, we introduce a *model-based* approach for analysing M/EEG timeseries such as event-related potentials (ERPs) or decoding performance over time.
-Our method leverages Bayesian generalised additive multilevel models (GAMMs), providing posterior probabilities that an effect is above zero (or above chance) at each time point, while naturally accounting for temporal dependencies and between-subject variability.
+Our approach leverages Bayesian generalised additive multilevel models, providing posterior probabilities that an effect is above zero (or above chance) at each time point, while naturally accounting for temporal dependencies and between-subject variability.
 Using both simulated and actual M/EEG datasets, we demonstrate that this approach substantially outperforms conventional methods in estimating the onset and offset of neural effects, yielding more precise and reliable results.
-We provide an R package implementing the method and show how it can be seamlessly integrated into M/EEG analysis pipelines using MNE-Python."
+We provide an R package implementing the method and describe how it can be integrated into M/EEG analysis pipelines using MNE-Python."
 keywords: [EEG, MEG, cluster-based inference, simulation, multiple comparisons, generalised additive models, mixed-effects models, multilevel models, Bayesian statistics, brms]
@@
     latex-engine: xelatex
-    documentmode: doc # man, jou, doc, stu
+    documentmode: man # man, jou, doc, stu
     keep-tex: true
@@
 
-The conventional approach involves mass-univariate statistical testing with a unique threshold, followed by corrections for multiple comparisons with the goal of maintaining the familywise error rate (FWER) or false discovery rate (FDR) at the nominal level.
-Cluster-based inference is the most common way of achieving this sort of error control in the M/EEG literature, being the default option in several software programs.
-While effective for controlling error rates, cluster-based inference comes with a significant limitation: they shift the focus of inference from individual time points to clusters, making it difficult to draw precise conclusions about the onset or offset of observed effects [@maris2007; @sassenhagen2019].
-As pointed by @maris2007; "there is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA \[false alarm\] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect".
-Even worse, as noted by @rosenblatt2018, cluster-based inference suffers from low spatial resolution: "Since discovering a cluster means that 'there exists at least one voxel with an evoked response in the cluster', and not that 'all the voxels in the cluster have an evoked response', it follows that the larger the detected cluster, the less information we have on the location of the activation." As a consequence, cluster-based inferenc has been shown...
-
-To overcome the limitations of cluster-based inference, we introduce a novel model-based approach for precisely localising M/EEG effects in time (space, and other dimensions).
-The proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the posterior probability of effects being above chance at the level of timesteps, sensors, voxels, etc., while naturally taking into account spatiotemporal dependencies present in M/EEG timeseries.
+The conventional approach involves mass-univariate statistical testing through time and/or space followed by some form or correction for multiple comparisons with the goal of maintaining the familywise error rate (FWER) or false discovery rate (FDR) at the nominal level (e.g., 5%).
+Cluster-based inference is the most common way of achieving this sort of error control in the M/EEG literature, being the recommended approach in several software programs [e.g., `EEGlab` or `MNE-Python`, @delorme2004; @gramfort2013].
+While effective for controlling error rates, cluster-based inference comes with a significant limitation: by shifting the focus of inference from individual time points to clusters, it prevents the ability to draw precise conclusions about the onset or offset of observed effects [@maris2007; @sassenhagen2019].
+As pointed out by @maris2007; "there is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA [false alarm] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect".
+Even worse, @rosenblatt2018 note that cluster-based inference suffers from low spatial resolution: "Since discovering a cluster means that 'there exists at least one voxel with an evoked response in the cluster', and not that 'all the voxels in the cluster have an evoked response', it follows that the larger the detected cluster, the less information we have on the location of the activation." As a consequence, cluster-based inference is expected to perform poorly for localising the onset of M/EEG effects; a property that was later demonstrated in simulations studies [e.g., @rousselet_using_2025; @sassenhagen2019].
+
+To overcome the limitations of cluster-based inference, we introduce a novel *model-based* approach for precisely localising M/EEG effects in time, space, and other dimensions.
+The proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the posterior probability of effects being above chance at the level of timesteps, sensors, voxels, or other dimensions, while naturally taking into account spatiotemporal dependencies present in M/EEG timeseries.
 We compare the performance of the proposed approach to well-established alternative methods using both simulated and actual M/EEG data and show that it significantly outperforms alternative methods in estimating the onset and offset of M/EEG effects.
@@
 
+The issue with multiple comparisons...
 Different methods exist to control the family-wiser error rate (FWER), defined as the type-1 error rate (false positive) over an ensemble (family) of tests...
@@
 
-Here are some useful references to be discussed [@ehinger_unfold_2019; @luck_how_2017; @frossard2021; @frossard2022]...
+Some useful references to be discussed [@ehinger_unfold_2019; @frossard2021; @frossard2022]...
 
@@
 where $h_{0}$ is typically $0$ and parameters $E$ and $H$ are set a priori (typically to 0.5 and 2, respectively) and control the influence of the extend and height on the TFCE.
+Note that in practise, this intergral is approximated by a sum over small $h$ increments.
 Then, p-value for timestep/voxel $t$ is computed by comparing it TFCE with the null distribution of TFCE values.
@@
 The TFCE combined with permutation (assuming a large enough number of permutations) has been shown to provide accurate type 1 FWER [e.g., @pernet2015].
-However, previous simulation work showed that cluster-based methods have indeed poor performance in localising the onset of M/EEG effects [e.g., @rousselet_using_2025; @sassenhagen2019].
+However, previous simulation work showed that cluster-based methods (including the TFCE method) perform poorly in localising the onset of M/EEG effects [e.g., @rousselet_using_2025; @sassenhagen2019].
 
 To sum up this section, cluster-based inference main limitations is that it provides inference at the cluster level only, not allowing inference at the level of timesteps, sensors, etc.
-Thus, it does not allow inferring the localisation of effects [for more details on cluster-based inference, see for instance @maris2007; @maris2011; @sassenhagen2019].
-In the following, we briefly review previous modelling work.
-Then, we provide a brief introduction to Bayesian statistical modelling, before moving to generalised additive models (GAMs) and generalised additive multilevel models (GAMMs) to illustrate how they can be used to precisely localise the onset and offset of M/EEG effects.
+As a consequence, it does not allow inferring the localisation of effects [for more details on cluster-based inference, see for instance @maris2011; @maris2007; @sassenhagen2019].
+In the following, we briefly review previous M/EEG modelling work.
+Then, we provide a brief introduction to generalised additive models (GAMs) and Bayesian generalised additive multilevel models (BGAMMs) to illustrate how these models can be used to precisely localise the onset and offset of M/EEG effects.
 
@@
 
-Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]... Weighting single trials [@pernet2022]... The LIMO toolbox [@pernet2011]...
+Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]...
+Weighting single trials [@pernet2022]...
+The LIMO toolbox [@pernet2011] using linear and 2-stage regression (not a proper multilevel model)...
 
@@
 gam <- brm(
-    # cubic regression splines with k-1 basis functions
-    # eeg ~ condition + s(time, bs = "cr", k = 20, by = condition),
     # thin-plate regression splines with k-1 basis functions
@@
 However, the previous model only included constant (fixed) effects, thus not properly accounting for between-participant variability.
-We next fit a multilevel version of the BGAM (BGAMM).
-Although it is possible to fit a BGAMM at the single-trial level, we present a resource-efficient version of the model that is fitted directly on the by-participant summary statistics (mean and SD), similar to what is done in meta-analysis.
+We next fit a multilevel version of the BGAM [BGAMM, see @nalborczyk2019].
+Although it is possible to fit a BGAMM at the single-trial level, we present a computationally-lighter version of the model that is fitted directly on by-participant summary statistics (mean and SD), similar to what is done in meta-analysis.
 
@@
 
-```{r gamm, echo = FALSE, eval = FALSE, results = "hide"}
-# defining a contrast for condition
-contrasts(raw_df$condition) <- c(-0.5, 0.5)
-
-# fitting the GAMM with factor smooth (per participant)
-gamm <- brm(
-    # cubic regression splines with k-1 basis functions
-    eeg ~ condition +
-        s(time, bs = "cr", k = 20, by = condition) +
-        s(time, participant, bs = "fs"),
-    data = raw_df,
-    family = gaussian(),
-    warmup = 2000,
-    iter = 5000,
-    chains = 4,
-    cores = 4,
-    file = "models/gamm.rds"
-    )
-```
-
 We depict the posterior predictions together with the posterior estimate of the slope for `condition` at each timestep (@fig-plot-post-slope).
-This figure suggests that the BGAMM provides an adequate description of the simulated data.
+This figure suggests that the BGAMM provides an adequate description of the simulated data (see further posterior predictive checks in @sec-basis).
 
@@
 We then compute the posterior probability of the slope for `condition` being above $0$ (@fig-post-prob-ratio, left).
-We can also express this as the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio, right).
+From this quantity, we then compute the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio, right).
 Note that a ratio of 10 means that the probability of the difference being above 0 is 10 times higher than the probability of the difference not being above 0, given the data, the priors, and other model's assumptions.
@@
 Right: Ratio of posterior probability according to the BGAMM (on a log10 scale).
-Timesteps above threshold (10) are highlighted in green."}
+Timesteps above threshold (10) are highlighted in green.
+NB: the minimum and maximum possible ratio values are determined (bounded) by the number of available posterior samples in the model."}
 # defining a function to compute and plot the posterior probability ratio
@@
 
-## Comparing the identified onsets/offsets to other approaches
+## Comparing the onsets/offsets estimates from other approaches
 
@@
 
-To assess the accuracy of group-level onset estimation, the various methods were compared using the bias (i.e., median(estimated-true)), median absolute error (MAE), root mean square error (RMSE), variance, and median absolute deviation (MAD) of onset/offset estimates computed on 1000 simulated datasets.
-As in @rousselet_using_2025, each participant was assigned a random onset betwee, 150 and 170ms.
+To assess the accuracy of group-level onset estimation, the various methods were compared using the bias (i.e., median(estimated-true)), median absolute error (MAE), root mean square error (RMSE), variance, and median absolute deviation (MAD) of onset/offset estimates computed on 1,000 simulated datasets.
+As in @rousselet_using_2025, each participant was assigned a random onset between 150 and 170ms.
 
@@
 We assessed the reliability of the proposed approach using a form of permutation-based split-half reliability [as for instance in @rosenblatt2018], which consisted of the following steps.
-First, we created 1000 split halves of the data (i.e., with half the participants in the original data, that is, 16 participants).
+First, we created 1,000 split halves of the data (i.e., with half the participants in the original data, that is, 16 participants).
 For each split, we estimated the onset/offset using all methods described previously.
 Third, we summarised the distribution of onset/offset estimates using the median "error" (i.e., difference between the split estimate and the estimate obtained using the full dataset) and the variance across splits.
-This approach allows assessing how similar the estimate of each half split is to the full dataset (thus acting as a proxy for the population) and how variable the estimates are (across split halves).
+This approach allows assessing how similar the estimate of each half split is to the full dataset (thus acting as a proxy for the population) and how variable the estimates are across split halves.
 
@@
     )
-
-# fitting the Beta GAMM
-# meg_decoding_gam <- brm(
-#     # auc_mean ~ s(time, bs = "cr", k = 50),
-#     auc_mean | se(auc_sd) ~ s(time, bs = "cr", k = 10) +
-#         s(time, participant, bs = "fs", xt = "cr", k = 10, m = 1),
-#     data = decoding_summary_df,
-#     # family = Beta(),
-#     # warmup = 2000,
-#     # iter = 5000,
-#     chains = 4,
-#     cores = 4,
-#     file = "models/meg_decoding_gamm.rds"
-#     )
-
-# some posterior predictive checks
-# pp_check(gam_split, prefix = "ppd")
-# pp_check(gam_split)
-# pp_check(gam_split, type = "error_hist")
-# pp_check(gam_split, type = "stat_2d")
-# pp_check(gam_split, type = "ecdf_overlay")
 
@@
 @fig-onset-offset).[^2] Overall, the figure reveals that for all other methods, split datasets produce later onset estimates and earlier offset estimates (as compared to the estimates from the model fitted on the full dataset).
-These results highlight that the results of any method should be assessed using both i) good asymptotic properties on simulated data, ii) sensible identified clusters in actual data, and iii) reliable/stable estimates on actual data.
+These results highlight some desirable properties for a method aiming to precisely and reliably estimate the onset and offset of M/EEG effects, namely, it should i) have good asymptotic properties on simulated data, ii) provide sensible identified clusters in actual data, and iii) provide reliable/stable estimates on actual data.
 
@@
 # number of timepoints
-# n_timepoints <- 240 # (1.2s at 200Hz)
 n_timepoints <- 60 # (1.2s at 50Hz)
@@
 # generating participant-specific cross-temporal generalisation matrices
-# timegen_data <- tibble()
 timegen_data <- data.frame()
@@
     auc ~ t2(train_time, test_time, bs = "tp", k = 20),
-    # auc ~ t2(train_time, test_time, bs = "tp", k = 30),
     data = timegen_data,
     family = Beta(),
-    warmup = 1000,
-    iter = 2000,
+    warmup = 2000,
+    iter = 5000,
     chains = 8,
     cores = 8,
-    file = "models/timegen_gam_t2.rds"
-    # file = "models/timegen_gam_t2_k30.rds"
+    # file = "models/timegen_gam_t2.rds" # k = 10
+    file = "models/timegen_gam_t2_k20.rds" # k = 20
     )

### 2025-04-10 brms_meeg.tex
pushing a cleaner version of the intro

@@
 \documentclass[
-  doc,
+  man,
   floatsintext,
@@
 \usepackage{amssymb}
-
-\geometry{inner=1in, outer=1in}
-\fancyhfoffset[LE,RO]{0cm}
 
@@
 such as event-related potentials (ERPs) or decoding performance over
-time. Our method leverages Bayesian generalised additive multilevel
-models (GAMMs), providing posterior probabilities that an effect is
-above zero (or above chance) at each time point, while naturally
-accounting for temporal dependencies and between-subject variability.
-Using both simulated and actual M/EEG datasets, we demonstrate that this
-approach substantially outperforms conventional methods in estimating
-the onset and offset of neural effects, yielding more precise and
-reliable results. We provide an R package implementing the method and
-show how it can be seamlessly integrated into M/EEG analysis pipelines
-using MNE-Python.}
+time. Our approach leverages Bayesian generalised additive multilevel
+models, providing posterior probabilities that an effect is above zero
+(or above chance) at each time point, while naturally accounting for
+temporal dependencies and between-subject variability. Using both
+simulated and actual M/EEG datasets, we demonstrate that this approach
+substantially outperforms conventional methods in estimating the onset
+and offset of neural effects, yielding more precise and reliable
+results. We provide an R package implementing the method and describe
+how it can be integrated into M/EEG analysis pipelines using
+MNE-Python.}
 
@@
 The conventional approach involves mass-univariate statistical testing
-with a unique threshold, followed by corrections for multiple
-comparisons with the goal of maintaining the familywise error rate
-(FWER) or false discovery rate (FDR) at the nominal level. Cluster-based
-inference is the most common way of achieving this sort of error control
-in the M/EEG literature, being the default option in several software
-programs. While effective for controlling error rates, cluster-based
-inference comes with a significant limitation: they shift the focus of
-inference from individual time points to clusters, making it difficult
-to draw precise conclusions about the onset or offset of observed
-effects (\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
+through time and/or space followed by some form or correction for
+multiple comparisons with the goal of maintaining the familywise error
+rate (FWER) or false discovery rate (FDR) at the nominal level (e.g.,
+5\%). Cluster-based inference is the most common way of achieving this
+sort of error control in the M/EEG literature, being the recommended
+approach in several software programs (e.g., \texttt{EEGlab} or
+\texttt{MNE-Python}, \citeproc{ref-delorme2004}{Delorme \& Makeig,
+2004}; \citeproc{ref-gramfort2013}{Gramfort, 2013}). While effective for
+controlling error rates, cluster-based inference comes with a
+significant limitation: by shifting the focus of inference from
+individual time points to clusters, it prevents the ability to draw
+precise conclusions about the onset or offset of observed effects
+(\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
 \citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}). As
-pointed by Maris \& Oostenveld (\citeproc{ref-maris2007}{2007}); ``there
-is a conflict between this interest in localized effects and our choice
-for a global null hypothesis: by controlling the FA {[}false alarm{]}
-rate under this global null hypothesis, one cannot quantify the
+pointed out by Maris \& Oostenveld (\citeproc{ref-maris2007}{2007});
+``there is a conflict between this interest in localized effects and our
+choice for a global null hypothesis: by controlling the FA {[}false
+alarm{]} rate under this global null hypothesis, one cannot quantify the
 uncertainty in the spatiotemporal localization of the effect''. Even
-worse, as noted by Rosenblatt et al.
-(\citeproc{ref-rosenblatt2018}{2018}), cluster-based inference suffers
-from low spatial resolution: ``Since discovering a cluster means that
-`there exists at least one voxel with an evoked response in the
-cluster', and not that `all the voxels in the cluster have an evoked
-response', it follows that the larger the detected cluster, the less
-information we have on the location of the activation.'' As a
-consequence, cluster-based inferenc has been shown\ldots{}
+worse, Rosenblatt et al. (\citeproc{ref-rosenblatt2018}{2018}) note that
+cluster-based inference suffers from low spatial resolution: ``Since
+discovering a cluster means that `there exists at least one voxel with
+an evoked response in the cluster', and not that `all the voxels in the
+cluster have an evoked response', it follows that the larger the
+detected cluster, the less information we have on the location of the
+activation.'' As a consequence, cluster-based inference is expected to
+perform poorly for localising the onset of M/EEG effects; a property
+that was later demonstrated in simulations studies (e.g.,
+\citeproc{ref-rousselet_using_2025}{Rousselet, 2025};
+\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}).
 
 To overcome the limitations of cluster-based inference, we introduce a
-novel model-based approach for precisely localising M/EEG effects in
-time (space, and other dimensions). The proposed approach, based on
+novel \emph{model-based} approach for precisely localising M/EEG effects
+in time, space, and other dimensions. The proposed approach, based on
 Bayesian generalised additive multilevel models, allows quantifying the
 posterior probability of effects being above chance at the level of
-timesteps, sensors, voxels, etc., while naturally taking into account
-spatiotemporal dependencies present in M/EEG timeseries. We compare the
-performance of the proposed approach to well-established alternative
-methods using both simulated and actual M/EEG data and show that it
-significantly outperforms alternative methods in estimating the onset
-and offset of M/EEG effects.
+timesteps, sensors, voxels, or other dimensions, while naturally taking
+into account spatiotemporal dependencies present in M/EEG timeseries. We
+compare the performance of the proposed approach to well-established
+alternative methods using both simulated and actual M/EEG data and show
+that it significantly outperforms alternative methods in estimating the
+onset and offset of M/EEG effects.
 
@@
 
-Different methods exist to control the family-wiser error rate (FWER),
-defined as the type-1 error rate (false positive) over an ensemble
-(family) of tests\ldots{} for instance the Bonferroni correction
-(\citeproc{ref-dunn1961}{Dunn, 1961}), however this method is generally
-overconservative as it assumes statistical independence of tests, an
-assumptions that is clearly violated in the context of M/EEG timeseries
-with spatiotemporal dependencies\ldots{} or the false discovery rate
-(FDR), defined as the proportion of false positive among positive tests
-(e.g., \citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
+The issue with multiple comparisons\ldots{} Different methods exist to
+control the family-wiser error rate (FWER), defined as the type-1 error
+rate (false positive) over an ensemble (family) of tests\ldots{} for
+instance the Bonferroni correction (\citeproc{ref-dunn1961}{Dunn,
+1961}), however this method is generally overconservative as it assumes
+statistical independence of tests, an assumptions that is clearly
+violated in the context of M/EEG timeseries with spatiotemporal
+dependencies\ldots{} or the false discovery rate (FDR), defined as the
+proportion of false positive among positive tests (e.g.,
+\citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
 \citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001})\ldots{}
@@
 set a priori (typically to 0.5 and 2, respectively) and control the
-influence of the extend and height on the TFCE. Then, p-value for
-timestep/voxel \(t\) is computed by comparing it TFCE with the null
-distribution of TFCE values. For each permuted signal, we keep the
-maximal value over the whole signal for the null distribution of the
-TFCE. The TFCE combined with permutation (assuming a large enough number
-of permutations) has been shown to provide accurate type 1 FWER (e.g.,
-\citeproc{ref-pernet2015}{Pernet et al., 2015}). However, previous
-simulation work showed that cluster-based methods have indeed poor
-performance in localising the onset of M/EEG effects (e.g.,
-\citeproc{ref-rousselet_using_2025}{Rousselet, 2025};
+influence of the extend and height on the TFCE. Note that in practise,
+this intergral is approximated by a sum over small \(h\) increments.
+Then, p-value for timestep/voxel \(t\) is computed by comparing it TFCE
+with the null distribution of TFCE values. For each permuted signal, we
+keep the maximal value over the whole signal for the null distribution
+of the TFCE. The TFCE combined with permutation (assuming a large enough
+number of permutations) has been shown to provide accurate type 1 FWER
+(e.g., \citeproc{ref-pernet2015}{Pernet et al., 2015}). However,
+previous simulation work showed that cluster-based methods (including
+the TFCE method) perform poorly in localising the onset of M/EEG effects
+(e.g., \citeproc{ref-rousselet_using_2025}{Rousselet, 2025};
 \citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}).
@@
 it provides inference at the cluster level only, not allowing inference
-at the level of timesteps, sensors, etc. Thus, it does not allow
-inferring the localisation of effects (\citeproc{ref-maris2011}{Maris,
-2011}; for more details on cluster-based inference, see for instance
-\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
-\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}). In the
-following, we briefly review previous modelling work. Then, we provide a
-brief introduction to Bayesian statistical modelling, before moving to
-generalised additive models (GAMs) and generalised additive multilevel
-models (GAMMs) to illustrate how they can be used to precisely localise
-the onset and offset of M/EEG effects.
+at the level of timesteps, sensors, etc. As a consequence, it does not
+allow inferring the localisation of effects (for more details on
+cluster-based inference, see for instance
+\citeproc{ref-maris2011}{Maris, 2011}; \citeproc{ref-maris2007}{Maris \&
+Oostenveld, 2007}; \citeproc{ref-sassenhagen2019}{Sassenhagen \&
+Draschkow, 2019}). In the following, we briefly review previous M/EEG
+modelling work. Then, we provide a brief introduction to generalised
+additive models (GAMs) and Bayesian generalised additive multilevel
+models (BGAMMs) to illustrate how these models can be used to precisely
+localise the onset and offset of M/EEG effects.
 
@@
 2022})\ldots{} The LIMO toolbox (\citeproc{ref-pernet2011}{Pernet et
-al., 2011})\ldots{}
+al., 2011}) using linear and 2-stage regression (not a proper multilevel
+model)\ldots{}
 
@@
 \NormalTok{gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-    \CommentTok{\# cubic regression splines with k{-}1 basis functions}
-    \CommentTok{\# eeg \textasciitilde{} condition + s(time, bs = "cr", k = 20, by = condition),}
     \CommentTok{\# thin{-}plate regression splines with k{-}1 basis functions}
@@
 not properly accounting for between-participant variability. We next fit
-a multilevel version of the BGAM (BGAMM). Although it is possible to fit
-a BGAMM at the single-trial level, we present a resource-efficient
-version of the model that is fitted directly on the by-participant
-summary statistics (mean and SD), similar to what is done in
-meta-analysis.
+a multilevel version of the BGAM (BGAMM, see
+\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}). Although it is
+possible to fit a BGAMM at the single-trial level, we present a
+computationally-lighter version of the model that is fitted directly on
+by-participant summary statistics (mean and SD), similar to what is done
+in meta-analysis.
 
@@
 (Figure~\ref{fig-plot-post-slope}). This figure suggests that the BGAMM
-provides an adequate description of the simulated data.
+provides an adequate description of the simulated data (see further
+posterior predictive checks in Section~\ref{sec-basis}).
 
@@
 \texttt{condition} being above \(0\) (Figure~\ref{fig-post-prob-ratio},
-left). We can also express this as the ratio of posterior probabilities
-(i.e., \(p/(1-p)\)) and visualise the timecourse of this ratio
-superimposed with the conventional thresholds on evidence ratios
+left). From this quantity, we then compute the ratio of posterior
+probabilities (i.e., \(p/(1-p)\)) and visualise the timecourse of this
+ratio superimposed with the conventional thresholds on evidence ratios
 (Figure~\ref{fig-post-prob-ratio}, right). Note that a ratio of 10 means
@@
 Ratio of posterior probability according to the BGAMM (on a log10
-scale). Timesteps above threshold (10) are highlighted in green.}
+scale). Timesteps above threshold (10) are highlighted in green. NB: the
+minimum and maximum possible ratio values are determined (bounded) by
+the number of available posterior samples in the model.}
 
@@
 
-\subsection{Comparing the identified onsets/offsets to other
-approaches}\label{comparing-the-identified-onsetsoffsets-to-other-approaches}
+\subsection{Comparing the onsets/offsets estimates from other
+approaches}\label{comparing-the-onsetsoffsets-estimates-from-other-approaches}
 
@@
 and median absolute deviation (MAD) of onset/offset estimates computed
-on 1000 simulated datasets. As in Rousselet
+on 1,000 simulated datasets. As in Rousselet
 (\citeproc{ref-rousselet_using_2025}{2025}), each participant was
-assigned a random onset betwee, 150 and 170ms.
+assigned a random onset between 150 and 170ms.
 
@@
 \citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}), which consisted
-of the following steps. First, we created 1000 split halves of the data
+of the following steps. First, we created 1,000 split halves of the data
 (i.e., with half the participants in the original data, that is, 16
@@
 acting as a proxy for the population) and how variable the estimates are
-(across split halves).
+across split halves.
 
@@
 {\caption{{Summary statistics of the onset and offset estimates for each
-method (ordered by MAE).}{\label{tbl-simulation-results}}}}
+method (ordered by MAE).}{\label{tbl-simulation-results}}}
+\vspace{-20pt}}
 
@@
 earlier offset estimates (as compared to the estimates from the model
-fitted on the full dataset). These results highlight that the results of
-any method should be assessed using both i) good asymptotic properties
-on simulated data, ii) sensible identified clusters in actual data, and
-iii) reliable/stable estimates on actual data.
+fitted on the full dataset). These results highlight some desirable
+properties for a method aiming to precisely and reliably estimate the
+onset and offset of M/EEG effects, namely, it should i) have good
+asymptotic properties on simulated data, ii) provide sensible identified
+clusters in actual data, and iii) provide reliable/stable estimates on
+actual data.
 
@@
 
+\bibitem[\citeproctext]{ref-delorme2004}
+Delorme, A., \& Makeig, S. (2004). EEGLAB: an open source toolbox for
+analysis of single-trial EEG dynamics including independent component
+analysis. \emph{Journal of Neuroscience Methods}, \emph{134}(1), 9--21.
+\url{https://doi.org/10.1016/j.jneumeth.2003.10.009}
+
 \bibitem[\citeproctext]{ref-dimigen2021}
@@
 \NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
-    \CommentTok{\# auc \textasciitilde{} t2(train\_time, test\_time, bs = "tp", k = 30),}
     \AttributeTok{data =}\NormalTok{ timegen\_data,}
@@
     \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"}
-    \CommentTok{\# file = "models/timegen\_gam\_t2\_k30.rds"}
+    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"} \CommentTok{\# k = 10}
+    \CommentTok{\# file = "models/timegen\_gam\_t2\_k20.rds" \# k = 20}
 \NormalTok{    )}

### 2025-04-11 brms_meeg.qmd
new sections in the intro

@@
 # see https://www.overleaf.com/learn/latex/Font_typefaces
-# and systemfonts::system_fonts()
-# pdf-engine: xelatex
-# monofont: Roboto
+# see available font via systemfonts::system_fonts()
+# monofont: Monaco
+# monofont: Courier
 mainfont: CMU Serif
@@
 donotrepeattitle: true
+suppress-title-introduction: true
 lang: en
@@
     latex-engine: xelatex
-    documentmode: man # man, jou, doc, stu
+    documentmode: doc # man, jou, doc, stu
     keep-tex: true
@@
           \usepackage{booktabs, caption, longtable, colortbl, array}
+          \usepackage{setspace}
+          \onehalfspacing
 ---
@@
 
-The conventional approach involves mass-univariate statistical testing through time and/or space followed by some form or correction for multiple comparisons with the goal of maintaining the familywise error rate (FWER) or false discovery rate (FDR) at the nominal level (e.g., 5%).
-Cluster-based inference is the most common way of achieving this sort of error control in the M/EEG literature, being the recommended approach in several software programs [e.g., `EEGlab` or `MNE-Python`, @delorme2004; @gramfort2013].
-While effective for controlling error rates, cluster-based inference comes with a significant limitation: by shifting the focus of inference from individual time points to clusters, it prevents the ability to draw precise conclusions about the onset or offset of observed effects [@maris2007; @sassenhagen2019].
-As pointed out by @maris2007; "there is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA [false alarm] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect".
+The conventional approach involves mass-univariate statistical testing through time and/or space followed by some form or correction for multiple comparisons with the goal of maintaining the familywise error rate (FWER) or the false discovery rate (FDR) at the nominal level (e.g., 5%).
+Cluster-based inference is the most common way of achieving this sort of error control in the M/EEG literature, being the recommended approach in several software programs [e.g., `EEGlab`, @delorme2004; `MNE-Python`,  @gramfort2013].
+While effective for controlling error rates, cluster-based inference comes with a significant limitation: by shifting the focus of inference from individual datapoints (e.g., timesteps, sensors, voxels) to clusters, it prevents the ability to draw precise conclusions about the spatiotemporal localisation of such effects [@maris2007; @sassenhagen2019].
+As pointed out by @maris2007: "there is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA [false alarm] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect".
 Even worse, @rosenblatt2018 note that cluster-based inference suffers from low spatial resolution: "Since discovering a cluster means that 'there exists at least one voxel with an evoked response in the cluster', and not that 'all the voxels in the cluster have an evoked response', it follows that the larger the detected cluster, the less information we have on the location of the activation." As a consequence, cluster-based inference is expected to perform poorly for localising the onset of M/EEG effects; a property that was later demonstrated in simulations studies [e.g., @rousselet_using_2025; @sassenhagen2019].
@@
 To overcome the limitations of cluster-based inference, we introduce a novel *model-based* approach for precisely localising M/EEG effects in time, space, and other dimensions.
-The proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the posterior probability of effects being above chance at the level of timesteps, sensors, voxels, or other dimensions, while naturally taking into account spatiotemporal dependencies present in M/EEG timeseries.
+The proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the posterior probability of effects being above chance at the level of timesteps, sensors, voxels, etc, while naturally taking into account spatiotemporal dependencies present in M/EEG data.
 We compare the performance of the proposed approach to well-established alternative methods using both simulated and actual M/EEG data and show that it significantly outperforms alternative methods in estimating the onset and offset of M/EEG effects.
 
-## Cluster-based inference
-
-The issue with multiple comparisons...
-Different methods exist to control the family-wiser error rate (FWER), defined as the type-1 error rate (false positive) over an ensemble (family) of tests...
-for instance the Bonferroni correction [@dunn1961], however this method is generally overconservative as it assumes statistical independence of tests, an assumptions that is clearly violated in the context of M/EEG timeseries with spatiotemporal dependencies...
-or the false discovery rate (FDR), defined as the proportion of false positive among positive tests [e.g., @benjamini1995; @benjamini2001]...
-
-<!--
-
-Some useful references to be discussed [@ehinger_unfold_2019; @frossard2021; @frossard2022]...
-
--->
-
-On popular technique to account for spatiotemporal dependencies while controlling the FWER is cluster-based inference.
-Description of cluster-based approaches [see @sassenhagen2019]...
-A typical cluster-based inference consists of two successive steps.
-First, clusters are defined as sets of contiguous voxels, channels, and/or timesteps, whose intensity/activity exceeds some predefined threshold.
-Clusters are then characterised by their height (i.e., maximal value), extent (number of constituent elements), or some combination of both, such at its "mass", for instance by summing the statistics within a cluster, an approach refereed to as "cluster mass" [@maris2007; @pernet2015].
-Then, the null hypothesis is tested by assessing the probability..
+## Statistical errors and cluster-based inference
+
+The issues with multiple comparisons represent a common and well-recognised danger in neuroimaging and M/EEG research where the collected data allows for a multitude of potential hypothesis tests and is characterised by complex structures of spatiotemporal dependencies.
+The probability of obtaining at least one false positive in an ensemble (family) of $m$ tests (i.e., the FWER) is computed as $1-\left(1-\alpha\right)^{m}$ (for $m=10$ tests and $\alpha=0.05$, it is approximately equal to $0.4$).
+Different methods exist to control the FWER (i.e., bring it back to $\alpha$).
+Most methods apply a simple correction to series of "raw" $p$-values issued from univariate statistical tests (e.g., t-tests).
+For instance, the Bonferroni correction [@dunn1961] consists in setting the significance threshold to $\alpha/m$, or equivalently, multiplying the "raw" $p$-values by $m$ and using the standard $\alpha$ significance threshold.
+This method is generally overconservative (i.e., under-powered) as it assumes statistical independence of the tests, an assumption that is clearly violated in the context of M/EEG timeseries characterised by massive spatiotemporal dependencies.
+Some alternative methods aims at controlling the FDR, defined as the proportion of false positive *among positive tests* [e.g., @benjamini1995; @benjamini2001].
+However, a major limitation of both types of corrections is that they do not take into account the spatial and temporal information contained in M/EEG data.
+
+A popular technique to account for spatiotemporal dependencies while controlling the FWER is cluster-based inference [@bullmore1999; @maris2007].
+A typical cluster-based inference consists of two successive steps [for more details on cluster-based inference, see for instance @frossard2022; @maris2011; @maris2007; @sassenhagen2019].
+First, clusters are defined as sets of contiguous timesteps, sensors, voxels, etc, whose activity, summarised by some test statistic (e.g., a $t$-value), exceeds a predefined threshold (e.g., the 95th percentile of the parametric null distribution).
+Clusters are then characterised by their height (i.e., maximal value), extent (number of constituent elements), or some combination of both, for instance by summing the statistics within a cluster, an approach referred to as "cluster mass" [@maris2007; @pernet2015].
+Then, the null hypothesis is tested by computing a $p$-value for each identified cluster by comparing its mass with the null distribution of cluster masses (obtained via permutation).
+As alluded previously, a significant cluster is a cluster which contains *at least one* significant time-point.
+As such, it would be incorrect to conclude, for instance, that the timestep of a significant cluster is the first moment at which some conditions differ [@frossard2022; @sassenhagen2019].
+In other words, because the inference is performed at the second step (i.e., once clusters have been formed), it prevents any conclusion to be made about individual datapoints (e.g., timesteps, sensors, etc).
+
 As different cluster-forming thresholds lead to clusters with different spatial or temporal extent, this initial threshold modulates the sensitivity of the subsequent permutation test.
-The threshold-free cluster enhancement method (TFCE) was introduced by @smith2009 to overcome this choice of an arbitrary threshold.
-
+The threshold-free cluster enhancement (TFCE) method was introduced by @smith2009 to overcome this choice of an arbitrary threshold.
 In brief, the TFCE method works as follows.
-Instead of picking an arbitrary cluster-forming threshold (e.g., $t=2$), we try all (or many) possible thresholds in a given range and check whether a given timestep/voxel belongs to a significant cluster under any of the set of thresholds.
-Then, instead of using cluster mass (e.g., the sum of squared t-values within the cluster), we use a weighted average between the cluster extend ($e$, how broad is the cluster, that is, how many connected samples it contains) and the cluster height ($h$, how high is the cluster, that is, how large is the test statistic).
-The TFCE score at each timestep/voxel $t$ is given by:
+Instead of picking an arbitrary cluster-forming threshold (e.g., $t=2$), the methods consist in trying all (or many) possible thresholds in a given range and checking whether a given datapoint (e.g., timestep, sensor, voxel) belongs to a significant cluster under any of the set of thresholds.
+Then, instead of using cluster mass, one uses a weighted average between the cluster extend ($e$, how broad is the cluster, that is, how many connected samples it contains) and the cluster height ($h$, how high is the cluster, that is, how large is the test statistic).
+The TFCE score at each timestep $t$ is given by:
 
 $$
-\text{TFCE(t)} = \int_{h=h_{0}}^{h_{t}} e(h)^{E} h^{H} \mathrm{d}h
+\text{TFCE(t)} = \int_{h=h_{0}}^{h=h_{t}} e(h)^{E} h^{H} \mathrm{d}h
 $$
@@
 where $h_{0}$ is typically $0$ and parameters $E$ and $H$ are set a priori (typically to 0.5 and 2, respectively) and control the influence of the extend and height on the TFCE.
-Note that in practise, this intergral is approximated by a sum over small $h$ increments.
-Then, p-value for timestep/voxel $t$ is computed by comparing it TFCE with the null distribution of TFCE values.
+Note that in practise, this integral is approximated by a sum over small $h$ increments.
+Then, a p-value for each timestep $t$ is computed by comparing its TFCE with the null distribution of TFCE values (obtained via permutation).
 For each permuted signal, we keep the maximal value over the whole signal for the null distribution of the TFCE.
-The TFCE combined with permutation (assuming a large enough number of permutations) has been shown to provide accurate type 1 FWER [e.g., @pernet2015].
-However, previous simulation work showed that cluster-based methods (including the TFCE method) perform poorly in localising the onset of M/EEG effects [e.g., @rousselet_using_2025; @sassenhagen2019].
-
-To sum up this section, cluster-based inference main limitations is that it provides inference at the cluster level only, not allowing inference at the level of timesteps, sensors, etc.
-As a consequence, it does not allow inferring the localisation of effects [for more details on cluster-based inference, see for instance @maris2011; @maris2007; @sassenhagen2019].
+The TFCE combined with permutation (assuming a large enough number of permutations) has been shown to provide accurate FWER control [e.g., @pernet2015].
+However, further simulation work showed that cluster-based methods (including TFCE) perform poorly in localising the onset of M/EEG effects [e.g., @rousselet_using_2025; @sassenhagen2019].
+
+To sum up, the main limitation of cluster-based inference is that it allows for inference at the cluster level only, not allowing inference at the level of timesteps, sensors, etc.
+As a consequence, it does not allow inferring the precise spatial and temporal localisation of effects.
 In the following, we briefly review previous M/EEG modelling work.
-Then, we provide a brief introduction to generalised additive models (GAMs) and Bayesian generalised additive multilevel models (BGAMMs) to illustrate how these models can be used to precisely localise the onset and offset of M/EEG effects.
+Then, we provide a short introduction to generalised additive models (GAMs) and Bayesian generalised additive multilevel models (BGAMMs) to illustrate how these models can be used to precisely localise the onset and offset of M/EEG effects.
 
@@
 
-From @dimigen2021: Recently, spline regression has been applied to ERPs [e.g., @hendrix2017; @kryuchkova2012]... GAMMs for EEG data [@abugaber2023; @meulman2015; @meulman2023]...
+From @dimigen2021: Recently, spline regression has been applied to ERPs [e.g., @hendrix2017; @kryuchkova2012]...
+GAMMs for EEG data [@abugaber2023; @meulman2015; @meulman2023]...
+See also the UNFOLD toolbox [@ehinger_unfold_2019]...
 
@@
 Splines coefficients are penalised (usually through the squared of the smooth functions' second derivative) in a way that can be interpreted, in Bayesian terms, as a prior on the "wiggliness" of the function.
-In other words, more complex (wiggly) basis function are penalised.
+In other words, more complex (wiggly) basis functions are penalised.
 
@@
 However, the previous model only included constant (fixed) effects, thus not properly accounting for between-participant variability.
-We next fit a multilevel version of the BGAM [BGAMM, see @nalborczyk2019].
+We next fit a multilevel version of the BGAM [BGAMM, for an introduction to Bayesian multilevel models in `brms`, see @nalborczyk2019].
 Although it is possible to fit a BGAMM at the single-trial level, we present a computationally-lighter version of the model that is fitted directly on by-participant summary statistics (mean and SD), similar to what is done in meta-analysis.
@@
     )
+
+# fitting the BGAMM
+# meta_gamm <- brm(
+#     # using by-participant SD of ERPs across trials
+#     eeg_mean | se(eeg_sd) ~ condition +
+#         s(participant, bs = "re") +
+#         s(time, participant, bs = "fs", by = condition, m = 1),
+#     data = summary_df,
+#     family = gaussian(),
+#     warmup = 2000,
+#     iter = 5000,
+#     chains = 8,
+#     cores = 8,
+#     control = list(adapt_delta = 0.95),
+#     file = "models/meta_gamm.rds"
+#     )
 ```
@@
 Thresholding the posterior probability ratio thus provides a model-based approach for estimating the onset and offset of M/EEG effects.
+An important advantage is that the proposed approach can be extended to virtually any model structure.
+Moreover, the approach provides both group-level and individual-level onset and offset estimates of M/EEG effects.
 
@@
 
+```{r fig-bgamm-varying-effects, eval = FALSE, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior estimate of the EEG activity in each condition (left) and posterior estimate of the difference in EEG activity (right) according to the BGAMM."}
+####################################################################
+# TO-DO: plotting the onset/offset estimate per participant?
+#################################################################
+
+model <- meta_gamm
+plot_post_preds(model = meta_gamm) + plot_post_slope(model = meta_gamm)
+
+plot_post_test_ratio_individual <- function (model, sesoi = 0, threshold = 10) {
+
+    # retrieving the number of posterior samples
+    n_posterior_samples <- ndraws(model)
+
+    # retrieving posterior draws and compute the difference per participant & time
+    prob_y_above_0_indiv <- model$data %>%
+        add_epred_draws(object = model) %>%
+        data.frame() %>%
+        dplyr::select(participant, time, condition, .epred, .draw) %>%
+        pivot_wider(names_from = condition, values_from = .epred) %>%
+        mutate(epred_diff = cond2 - cond1) %>%
+        group_by(participant, time) %>%
+        summarise(m = mean(epred_diff > 0 + sesoi), .groups = "drop") %>%
+        mutate(
+            prob_ratio = m / (1 - m),
+            prob_ratio = ifelse(is.infinite(prob_ratio), n_posterior_samples, prob_ratio),
+            prob_ratio = ifelse(prob_ratio == 0, 1 / n_posterior_samples, prob_ratio),
+            above_thres = ifelse(prob_ratio >= threshold, 1, NA)
+            )
+
+    # plotting posterior probabilities per participant
+    post_plot_indiv <- prob_y_above_0_indiv %>%
+        ggplot(aes(x = time, y = m, group = participant, color = participant) ) +
+        geom_hline(yintercept = 0.5, linetype = 2) +
+        geom_vline(xintercept = true_onset, linetype = 2) +
+        geom_vline(xintercept = true_offset, linetype = 2) +
+        geom_line(alpha = 0.5) +
+        labs(
+            x = "Time (s)",
+            y = expression(Pr(beta>0~"|"~data, model) ),
+            title = "Posterior probabilities per participant"
+            ) +
+        ylim(c(0, 1) ) +
+        theme(legend.position = "none")
+
+    # plotting posterior probability ratio per participant (log scale)
+    ratio_plot_indiv <- prob_y_above_0_indiv %>%
+        ggplot(aes(x = time, y = prob_ratio, group = participant, color = participant) ) +
+        geom_hline(yintercept = 1, linetype = "dashed") +
+        geom_hline(yintercept = 10, linetype = "dashed", color = "darkred") +
+        geom_hline(yintercept = 1 / 10, linetype = "dashed", color = "darkred") +
+        geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
+        geom_hline(yintercept = 1 / 100, linetype = "dashed", color = "orangered") +
+        geom_vline(xintercept = true_onset, linetype = 2) +
+        geom_vline(xintercept = true_offset, linetype = 2) +
+        geom_line(alpha = 0.5) +
+        geom_point(
+          data = . %>% dplyr::filter(!is.na(above_thres) ),
+          aes(y = threshold),
+          colour = "darkgreen",
+          shape = 16,
+          size = 1.5,
+          na.rm = TRUE
+        ) +
+        scale_y_log10(
+          labels = label_log(digits = 2),
+          limits = c(1 / n_posterior_samples, n_posterior_samples),
+          breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
+        ) +
+        labs(
+          x = "Time (s)",
+          y = expression(Pr(beta>0~"|"~data) / (1 - Pr(beta>0~"|"~data) ) ),
+          title = "Posterior probability ratios per participant"
+        ) +
+        theme(legend.position = "none")
+
+    # combining plots
+    post_plot_indiv + ratio_plot_indiv
+
+}
+```
+
 \newpage
@@
 ## Error properties of the proposed approach
+
+<!--
+
+Note: Given the focus on the precision of onset/offset estimates, is it correct/relevant to mention that we are interested here in quantifying type-S and type-M (rather than type-1 or type-2) errors?
+
+-->
 
@@
 
-We then compared the ability of the BGAMM to correctly estimate the onset and offset of the ERP difference to other widely-used methods.
+We then compared the ability of the BGAMM to accurately estimate the onset and offset of the ERP difference to other widely-used methods.
 First, we conducted mass-univariate t-tests (thus treating each timestep independently) and identified the onset and offset of the ERP difference as the first and last values crossing an arbitrary significance threshold ($\alpha = 0.05$).
-We then followed the same approach but after applying different forms of multiplicity correction to the p-values.
-We compared two methods that control the false discovery rate (FDR) [i.e., BH95, @benjamini1995; and BY01, @benjamini2001], one method that controls the family-wise error rate (FWER) [i.e., Holm–Bonferroni method, @holm1979], and two cluster-based permutation methods [permutation with a single cluster-forming threshold and threshold-free cluster enhancement, TFCE, @smith2009].
-The BH95, BY01, and Holm corrections were applied to the p-values using the `p.adjust()` function in `R`.
-The cluster-based inference was implemented using a cluster-sum statistic of squared t-values, as implemented in `MNE-Python` [@gramfort2013], called via `r pkrt("reticulate")`.
-We also compared these estimates to the onset and offset as estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared t-values [as in @rousselet_using_2025].
+We then followed the same approach but after applying different forms of multiplicity correction to the $p$-values.
+We compared two methods that control the FDR [i.e., `BH95`, @benjamini1995; and `BY01`, @benjamini2001], one method that controls the FWER [i.e., Holm–Bonferroni method, @holm1979], and two cluster-based permutation methods [permutation with a single cluster-forming threshold and threshold-free cluster enhancement, `TFCE`, @smith2009].
+The `BH95`, `BY01`, and `Holm` corrections were applied to the p-values using the `p.adjust()` function in `R`.
+The cluster-based inference was implemented using a cluster-sum statistic of squared $t$-values, as implemented in `MNE-Python` [@gramfort2013], called via `r pkrt("reticulate")`.
+We also compared these estimates to the onset and offset as estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared $t$-values [as in @rousselet_using_2025].
 @fig-corrections illustrates the onsets and offsets estimated by each method on a single simulated dataset and shows that all methods systematically overestimate the true onset and underestimate the true offset.
@@
 As in @rousselet_using_2025, each participant was assigned a random onset between 150 and 170ms.
+Whereas the present article focuses on one-dimensional signas (e.g., one M/EEG channel), we provice a 2D application in @sec-2D.
 
@@
 In this study, we conducted time-resolved multivariate pattern analysis (MVPA, also known as decoding) of MEG data during reading tasks.
-As a result, we obtain a timecourse of decoding performance (ROC AUC), bounded between 0 and 1, for each participant (for a total of 32 participants).
+As a result, we obtain a timecourse of decoding performance (ROC AUC), bounded between 0 and 1, for each participant ($N=32$).
 Next, we wanted to *test* whether the group-level average decoding accuracy is above chance (i.e., 0.5) at each timestep (@fig-decoding-data).
@@
 
-Note that although we chose a basis dimension of $k=50$, which seems appropriate for the present data, this choice should be adapted according to the properties of the modelled data (e.g., signal-to-noise ratio, prior low-pass filtering, sampling rate, etc) and should be assessed by the usual model checking tools (e.g., posterior predictive checks, see also @sec-basis).
+Note that although we chose a basis dimension of $k=50$, which seems appropriate for the present data, this choice should be adapted according to the properties of the modelled data (e.g., signal-to-noise ratio, prior low-pass filtering, sampling rate) and should be assessed by the usual model checking tools (e.g., posterior predictive checks, see also @sec-basis).
 To better distinguish signal from noise, we also defined a region of practical equivalence [ROPE, @kruschke2017], defined as the chance level plus the standard deviation of the (group-level average) decoding performance during the baseline period.
@@
 ```{r tbl-simulation-results, eval = TRUE, echo = FALSE, results = "asis"}
-#| tbl-cap: "Summary statistics of the onset and offset estimates for each method (ordered by MAE)."
+#| tbl-cap: "Summary statistics of the onset and offset estimates for each method (ordered by the absolute value of the bias)."
 
@@
         ) %>%
-    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
+    mutate(
+        onset_offset = factor(x = onset_offset, levels = c("onset", "offset") )
+        ) %>%
     data.frame()
@@
         ) %>%
-    arrange(median_abs_error) %>%
+    arrange(abs(median_error) ) %>%
     group_by(onset_offset) %>%
     gt() %>%
-    # gt(groupname_col = "onset_offset") %>%
     fmt_number(
@@
 
+```{r tbl-reliability-results, eval = TRUE, echo = FALSE, results = "asis"}
+#| tbl-cap: "Summary statistics of the onset and offset estimates for each method (ordered by the absolute value of the bias)."
+
+# retrieving the reliability results
+reliability_results <- readRDS(file = "results/reliability_results_100splits.rds")
+
+# extracting the full data estimates
+reliability_errors <- reliability_results %>%
+    data.frame() %>%
+    # keeping only the first and last timestep
+    group_by(split_id, method, onset_offset) %>%
+    summarise(
+        value = if (onset_offset[1] == "onset") min(value) else max(value),
+        .groups = "drop"
+        ) %>%
+    ungroup() %>%
+    select(-split_id) %>%
+    left_join(
+        full_reliability_results,
+        by = c("method", "onset_offset"),
+        suffix = c("_split", "_full")
+        ) %>%
+    group_by(method, onset_offset) %>%
+    mutate(error = value_split - value_full) %>%
+    ungroup()
+
+# summarising the results
+reliability_summary <- reliability_errors %>%
+    # mutate(
+    #     error = case_when(
+    #         grepl("onset", onset_offset) ~ value - true_onset,
+    #         grepl("offset", onset_offset) ~ value - true_offset
+    #         )
+    #     ) %>%
+    group_by(method, onset_offset) %>%
+    summarise(
+        median_error = median(error, na.rm = TRUE),
+        median_abs_error = median(abs(error), na.rm = TRUE),
+        rmse = sqrt(mean(error)^2),
+        variance = var(error, na.rm = TRUE),
+        mad = mad(error, na.rm = TRUE),
+        .groups = "drop"
+        ) %>%
+    ungroup() %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c(
+                "raw_p", "pval_bh", "pval_by", "pval_holm",
+                "cluster_mass", "cluster_tfce", "cpt", "brms"
+                ),
+            labels = c(
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+                "TFCE", "Change point", "BGAM (new method)"
+                ),
+            )
+        ) %>%
+    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
+    data.frame()
+
+# creating the table
+reliability_summary %>%
+    mutate(
+        onset_offset = factor(
+            x = onset_offset,
+            levels = c("offset", "onset")
+            )
+        ) %>%
+    # arrange(median_abs_error) %>%
+    arrange(abs(median_error) ) %>%
+    group_by(onset_offset) %>%
+    gt() %>%
+    # gt(groupname_col = "onset_offset") %>%
+    fmt_number(
+        columns = everything(),
+        decimals = 4
+        ) %>%
+    cols_align(
+        align = "center",
+        columns = everything()
+        ) %>%
+    cols_align(
+        align = "left",
+        columns = method
+        ) %>%
+    cols_label(
+        method = "",
+        median_error = "Bias",
+        median_abs_error = "MAE",
+        rmse = "RMSE",
+        variance = "Variance",
+        mad = "MAD"
+        ) %>%
+    tab_options(
+        table.align = "center",
+        # table.width = "75%",
+        table.font.size = 12
+        ) %>%
+    as_latex()
+```
+
 \newpage
@@
 The TFCE performs worse than the cluster-sum approach, which was anticipated by @rousselet_using_2025 based on the initial results of @smith2009...
+we did not include the cluster-depth algorithm [@frossard2022], as @rousselet_using_2025 already showed its performance were worse than the cluster mass algorithm...
 
@@
 note however, that the same caveats apply to all methods...
+
+Discussion about the nature of the model: we modelled the surface M/EEG signals, however, the true interests (probably) lie in the brain, that is, in the source space...
+we could build a "full" Bayesian model of the generated EEG signal (i.e., including hypotheses about the source and a forward model), but this model would become computationally heavier...
 
@@
 
-# Application to 2D time-resolved decoding results (cross-temporal generalisation)
+# Application to 2D time-resolved decoding results (cross-temporal generalisation) {#sec-2D}
 
@@
 
-```{r gam-gp-timegen, message = FALSE}
+```{r gam-timegen, eval = TRUE, echo = TRUE, results = "hide"}
 # fitting a GAM with two temporal dimensions
@@
     # 2D thin-plate spline (tp)
+    # auc ~ t2(train_time, test_time, bs = "tp", k = 10),
     auc ~ t2(train_time, test_time, bs = "tp", k = 20),
@@
     )
-
+```
+
+```{r gp-timegen, eval = FALSE, echo = FALSE, results = "hide"}
 # fitting a GP with two temporal dimensions
-# timegen_gp <- brm(
-#     auc ~ gp(train_time, test_time, k = 20),
-#     data = timegen_data,
-#     family = Beta(),
-#     control = list(adapt_delta = 0.95),
-#     iter = 2000,
-#     chains = 4,
-#     cores = 4,
-#     file = "models/timegen_gp.rds"
-#     )
-```
-
-```{r gam-timegen-preds, eval = FALSE, echo = FALSE, fig.cap = "Posterior predictions of the GAM fitted on decoding accuracy over time."}
+timegen_gp <- brm(
+    auc ~ gp(train_time, test_time, k = 20),
+    data = timegen_data,
+    family = Beta(),
+    control = list(adapt_delta = 0.95),
+    iter = 2000,
+    chains = 4,
+    cores = 4,
+    file = "models/timegen_gp.rds"
+    )
+```
+
+```{r gam-timegen-preds, eval = FALSE, echo = FALSE}
 # plotting the posterior predictions
@@
     surface = TRUE,
-    theme = theme_light(),
+    # theme = theme_light(),
     plot = TRUE
@@
 
-```{r gam-timegen-post-preds, echo = FALSE, fig.width = 8, fig.asp = 0.6, fig.cap = "Predicted AUC values (left) and posterior probabilies of decoding accuracy being above chance level (right) according to the bivariate GAM."}
-# defining a function to plot posterior prob of slope above 0
-plot_post_timegen <- function (model, data, chance_level = 0.5, eps = 0.1) {
+```{r gam-timegen-post-preds, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Simulated data with thresholds derived from the BGAM (left), predicted AUC values (middle), and ratios of posterior probabilies of decoding accuracy being above chance level (right) according to the bivariate BGAM (for k=10, top row, and k=20, bottom row)."}
+# defining a function to plot posterior predictions
+plot_post_timegen <- function (model, data = NULL, chance_level = 0.5, sesoi = 0, thresholds = c(10, 100) ) {
+
+    # if no data is specified, use model$data
+    if (is.null(data) ) data <- model$data
 
@@
         mutate(post_prob = colMeans(posterior_samples) ) %>%
-        mutate(m = colMeans(posterior_samples > (chance_level + eps) ) ) %>%
-        mutate(prob_ratio = m / (1 - m) )
+        mutate(m = colMeans(posterior_samples > (chance_level + sesoi) ) ) %>%
+        mutate(prob_ratio = m / (1 - m) ) %>%
+        mutate(prob_ratio = ifelse(prob_ratio == Inf, ndraws(model), prob_ratio) ) %>%
+        mutate(prob_ratio = ifelse(prob_ratio == 0, 1 / ndraws(model), prob_ratio) )
 
     # plotting it
-    # prob_y_above_0 %>% filter(prob_ratio != Inf) %>% pull(prob_ratio) %>% range()
     prob_above_chance_plot <- prob_y_above_0 %>%
-        mutate(prob_ratio = ifelse(prob_ratio == Inf, 2000, prob_ratio) ) %>%
-        mutate(prob_ratio = ifelse(prob_ratio == 0, 0.0005, prob_ratio) ) %>%
         ggplot(aes(x = train_time, y = test_time, fill = prob_ratio) ) +
@@
         coord_fixed() +
-        scale_fill_scico(palette = "vik", midpoint = 0, name = "p/(1-p)", trans = "log10") +
-        # scale_fill_scico(palette = "vik", midpoint = 1, name = "p/(1-p)") +
-        # theme_light() +
+        scale_fill_scico(
+            palette = "vik",
+            midpoint = 0,
+            name = "Posterior ratio",
+            trans = "log10"
+            ) +
         labs(x = "Testing time (s)", y = "Training time (s)")
 
+    # plotting cross-temporal generalization matrix
+    threshold_data_plot <- data %>%
+        summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>%
+        ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
+        geom_tile() +
+        geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
+        geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
+        geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
+        scale_x_continuous(expand = c(0, 0) ) +
+        scale_y_continuous(expand = c(0, 0) ) + 
+        coord_fixed() +
+        scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
+        labs(x = "Testing time (s)", y = "Training time (s)") +
+        geom_contour(
+            data = prob_y_above_0,
+            aes(x = train_time, y = test_time, z = prob_ratio),
+            breaks = thresholds,
+            color = "black",
+            linewidth = 1,
+            inherit.aes = FALSE
+            )
+
         # combining the two plots
-    post_preds_plot + prob_above_chance_plot
-
-    # some tests
-    # post_preds_plot +
-    #     geom_contour(
-    #         data = prob_y_above_0,
-    #         aes(x = train_time, y = test_time, z = prob_ratio),
-    #         breaks = 10,
-    #         color = "red",
-    #         linewidth = 1,
-    #         inherit.aes = FALSE
-    #         )
+        threshold_data_plot + post_preds_plot + prob_above_chance_plot
 
@@
 
-# using the function with the decoding GAM
-plot_post_timegen(
-    model = timegen_gam, data = timegen_data,
-    chance_level = 0.5, eps = 0.01
-    )
+# importing the models
+bgam_k10 <- readRDS(file = "models/timegen_gam_t2.rds")
+bgam_k20 <- readRDS(file = "models/timegen_gam_t2_k20.rds")
+baseline_sd <- timegen_data %>% filter(train_time < 0 & test_time < 0.2) %>% summarise(sd(auc) ) %>% pull()
+
+# plotting predictions for each model
+k10_plot <- plot_post_timegen(
+    model = bgam_k10,
+    chance_level = 0.5, sesoi = baseline_sd,
+    # thresholds = c(10, 100)
+    thresholds = 10
+    )
+
+k20_plot <- plot_post_timegen(
+    model = bgam_k20,
+    chance_level = 0.5, sesoi = baseline_sd,
+    # thresholds = c(10, 100),
+    thresholds = 10
+    )
+
+# combining the plots
+k10_plot / k20_plot
 ```
@@
 # using the testing_through_time() function from the neurogam package
-# this may take a few minutes (or hours depending the machine's performance
-# and data size)...
+# this may take a few minutes (or hours depending the machine's
+# performance and data size)...
 gam_onset_offset <- testing_through_time(
@@
     threshold = 20,
-    # the *_id arguments are use to specify the relevant columns in data 
+    # the *_id arguments are used to specify the relevant columns in data 
     participant_id = "participant", meeg_id = "eeg",
@@
 For example, we use it below to estimate the onset and offset of effects for one EEG channel from a MNE evoked object.
-Note that the code used to reshape the `sample MNE` dataset is provided later, and we refer to the [MNE documentation](https://mne.tools/stable/auto_tutorials/epochs/50_epochs_to_data_frame.html) about converting `MNE` epochs to `Pandas` dataframes in long format.
+Note that the code used to reshape the `sample MNE` dataset is available in the online supplementary materials, and we refer to the [MNE documentation](https://mne.tools/stable/auto_tutorials/epochs/50_epochs_to_data_frame.html) about converting `MNE` epochs to `Pandas` dataframes in long format (i.e., with one observation per row).
 
@@
 
-```{python, eval = FALSE, echo = TRUE}
+```{python, eval = FALSE, echo = FALSE}
 # loading the Python modules

### 2025-04-11 brms_meeg.tex
new sections in the intro

@@
 \documentclass[
-  man,
+  doc,
   floatsintext,
@@
   notimes,
+  donotrepeattitle,
   colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}
@@
 \usepackage{amssymb}
+
+\geometry{inner=1in, outer=1in}
+\fancyhfoffset[LE,RO]{0cm}
 
@@
 \usepackage{booktabs, caption, longtable, colortbl, array}
+\usepackage{setspace}
+\onehalfspacing
 \makeatletter
@@
 multiple comparisons with the goal of maintaining the familywise error
-rate (FWER) or false discovery rate (FDR) at the nominal level (e.g.,
-5\%). Cluster-based inference is the most common way of achieving this
-sort of error control in the M/EEG literature, being the recommended
-approach in several software programs (e.g., \texttt{EEGlab} or
-\texttt{MNE-Python}, \citeproc{ref-delorme2004}{Delorme \& Makeig,
-2004}; \citeproc{ref-gramfort2013}{Gramfort, 2013}). While effective for
-controlling error rates, cluster-based inference comes with a
-significant limitation: by shifting the focus of inference from
-individual time points to clusters, it prevents the ability to draw
-precise conclusions about the onset or offset of observed effects
+rate (FWER) or the false discovery rate (FDR) at the nominal level
+(e.g., 5\%). Cluster-based inference is the most common way of achieving
+this sort of error control in the M/EEG literature, being the
+recommended approach in several software programs (e.g.,
+\texttt{EEGlab}, \citeproc{ref-delorme2004}{Delorme \& Makeig, 2004};
+\texttt{MNE-Python}, \citeproc{ref-gramfort2013}{Gramfort, 2013}). While
+effective for controlling error rates, cluster-based inference comes
+with a significant limitation: by shifting the focus of inference from
+individual datapoints (e.g., timesteps, sensors, voxels) to clusters, it
+prevents the ability to draw precise conclusions about the
+spatiotemporal localisation of such effects
 (\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
 \citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}). As
-pointed out by Maris \& Oostenveld (\citeproc{ref-maris2007}{2007});
+pointed out by Maris \& Oostenveld (\citeproc{ref-maris2007}{2007}):
 ``there is a conflict between this interest in localized effects and our
@@
 posterior probability of effects being above chance at the level of
-timesteps, sensors, voxels, or other dimensions, while naturally taking
-into account spatiotemporal dependencies present in M/EEG timeseries. We
-compare the performance of the proposed approach to well-established
-alternative methods using both simulated and actual M/EEG data and show
-that it significantly outperforms alternative methods in estimating the
-onset and offset of M/EEG effects.
-
-\subsection{Cluster-based inference}\label{cluster-based-inference}
-
-The issue with multiple comparisons\ldots{} Different methods exist to
-control the family-wiser error rate (FWER), defined as the type-1 error
-rate (false positive) over an ensemble (family) of tests\ldots{} for
-instance the Bonferroni correction (\citeproc{ref-dunn1961}{Dunn,
-1961}), however this method is generally overconservative as it assumes
-statistical independence of tests, an assumptions that is clearly
-violated in the context of M/EEG timeseries with spatiotemporal
-dependencies\ldots{} or the false discovery rate (FDR), defined as the
-proportion of false positive among positive tests (e.g.,
-\citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
-\citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001})\ldots{}
-
-On popular technique to account for spatiotemporal dependencies while
-controlling the FWER is cluster-based inference. Description of
-cluster-based approaches (see \citeproc{ref-sassenhagen2019}{Sassenhagen
-\& Draschkow, 2019})\ldots{} A typical cluster-based inference consists
-of two successive steps. First, clusters are defined as sets of
-contiguous voxels, channels, and/or timesteps, whose intensity/activity
-exceeds some predefined threshold. Clusters are then characterised by
-their height (i.e., maximal value), extent (number of constituent
-elements), or some combination of both, such at its ``mass'', for
-instance by summing the statistics within a cluster, an approach
-refereed to as ``cluster mass'' (\citeproc{ref-maris2007}{Maris \&
-Oostenveld, 2007}; \citeproc{ref-pernet2015}{Pernet et al., 2015}).
-Then, the null hypothesis is tested by assessing the probability.. As
-different cluster-forming thresholds lead to clusters with different
+timesteps, sensors, voxels, etc, while naturally taking into account
+spatiotemporal dependencies present in M/EEG data. We compare the
+performance of the proposed approach to well-established alternative
+methods using both simulated and actual M/EEG data and show that it
+significantly outperforms alternative methods in estimating the onset
+and offset of M/EEG effects.
+
+\subsection{Statistical errors and cluster-based
+inference}\label{statistical-errors-and-cluster-based-inference}
+
+The issues with multiple comparisons represent a common and
+well-recognised danger in neuroimaging and M/EEG research where the
+collected data allows for a multitude of potential hypothesis tests and
+is characterised by complex structures of spatiotemporal dependencies.
+The probability of obtaining at least one false positive in an ensemble
+(family) of \(m\) tests (i.e., the FWER) is computed as
+\(1-\left(1-\alpha\right)^{m}\) (for \(m=10\) tests and \(\alpha=0.05\),
+it is approximately equal to \(0.4\)). Different methods exist to
+control the FWER (i.e., bring it back to \(\alpha\)). Most methods apply
+a simple correction to series of ``raw'' \(p\)-values issued from
+univariate statistical tests (e.g., t-tests). For instance, the
+Bonferroni correction (\citeproc{ref-dunn1961}{Dunn, 1961}) consists in
+setting the significance threshold to \(\alpha/m\), or equivalently,
+multiplying the ``raw'' \(p\)-values by \(m\) and using the standard
+\(\alpha\) significance threshold. This method is generally
+overconservative (i.e., under-powered) as it assumes statistical
+independence of the tests, an assumption that is clearly violated in the
+context of M/EEG timeseries characterised by massive spatiotemporal
+dependencies. Some alternative methods aims at controlling the FDR,
+defined as the proportion of false positive \emph{among positive tests}
+(e.g., \citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
+\citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001}). However, a
+major limitation of both types of corrections is that they do not take
+into account the spatial and temporal information contained in M/EEG
+data.
+
+A popular technique to account for spatiotemporal dependencies while
+controlling the FWER is cluster-based inference
+(\citeproc{ref-bullmore1999}{Bullmore et al., 1999};
+\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007}). A typical
+cluster-based inference consists of two successive steps (for more
+details on cluster-based inference, see for instance
+\citeproc{ref-frossard2022}{Frossard \& Renaud, 2022};
+\citeproc{ref-maris2011}{Maris, 2011}; \citeproc{ref-maris2007}{Maris \&
+Oostenveld, 2007}; \citeproc{ref-sassenhagen2019}{Sassenhagen \&
+Draschkow, 2019}). First, clusters are defined as sets of contiguous
+timesteps, sensors, voxels, etc, whose activity, summarised by some test
+statistic (e.g., a \(t\)-value), exceeds a predefined threshold (e.g.,
+the 95th percentile of the parametric null distribution). Clusters are
+then characterised by their height (i.e., maximal value), extent (number
+of constituent elements), or some combination of both, for instance by
+summing the statistics within a cluster, an approach referred to as
+``cluster mass'' (\citeproc{ref-maris2007}{Maris \& Oostenveld, 2007};
+\citeproc{ref-pernet2015}{Pernet et al., 2015}). Then, the null
+hypothesis is tested by computing a \(p\)-value for each identified
+cluster by comparing its mass with the null distribution of cluster
+masses (obtained via permutation). As alluded previously, a significant
+cluster is a cluster which contains \emph{at least one} significant
+time-point. As such, it would be incorrect to conclude, for instance,
+that the timestep of a significant cluster is the first moment at which
+some conditions differ (\citeproc{ref-frossard2022}{Frossard \& Renaud,
+2022}; \citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}).
+In other words, because the inference is performed at the second step
+(i.e., once clusters have been formed), it prevents any conclusion to be
+made about individual datapoints (e.g., timesteps, sensors, etc).
+
+As different cluster-forming thresholds lead to clusters with different
 spatial or temporal extent, this initial threshold modulates the
 sensitivity of the subsequent permutation test. The threshold-free
-cluster enhancement method (TFCE) was introduced by S. Smith \& Nichols
+cluster enhancement (TFCE) method was introduced by S. Smith \& Nichols
 (\citeproc{ref-smith2009}{2009}) to overcome this choice of an arbitrary
-threshold.
-
-In brief, the TFCE method works as follows. Instead of picking an
-arbitrary cluster-forming threshold (e.g., \(t=2\)), we try all (or
-many) possible thresholds in a given range and check whether a given
-timestep/voxel belongs to a significant cluster under any of the set of
-thresholds. Then, instead of using cluster mass (e.g., the sum of
-squared t-values within the cluster), we use a weighted average between
-the cluster extend (\(e\), how broad is the cluster, that is, how many
-connected samples it contains) and the cluster height (\(h\), how high
-is the cluster, that is, how large is the test statistic). The TFCE
-score at each timestep/voxel \(t\) is given by:
+threshold. In brief, the TFCE method works as follows. Instead of
+picking an arbitrary cluster-forming threshold (e.g., \(t=2\)), the
+methods consist in trying all (or many) possible thresholds in a given
+range and checking whether a given datapoint (e.g., timestep, sensor,
+voxel) belongs to a significant cluster under any of the set of
+thresholds. Then, instead of using cluster mass, one uses a weighted
+average between the cluster extend (\(e\), how broad is the cluster,
+that is, how many connected samples it contains) and the cluster height
+(\(h\), how high is the cluster, that is, how large is the test
+statistic). The TFCE score at each timestep \(t\) is given by:
 
 \[
-\text{TFCE(t)} = \int_{h=h_{0}}^{h_{t}} e(h)^{E} h^{H} \mathrm{d}h
+\text{TFCE(t)} = \int_{h=h_{0}}^{h=h_{t}} e(h)^{E} h^{H} \mathrm{d}h
 \]
@@
 influence of the extend and height on the TFCE. Note that in practise,
-this intergral is approximated by a sum over small \(h\) increments.
-Then, p-value for timestep/voxel \(t\) is computed by comparing it TFCE
-with the null distribution of TFCE values. For each permuted signal, we
-keep the maximal value over the whole signal for the null distribution
-of the TFCE. The TFCE combined with permutation (assuming a large enough
-number of permutations) has been shown to provide accurate type 1 FWER
-(e.g., \citeproc{ref-pernet2015}{Pernet et al., 2015}). However,
-previous simulation work showed that cluster-based methods (including
-the TFCE method) perform poorly in localising the onset of M/EEG effects
-(e.g., \citeproc{ref-rousselet_using_2025}{Rousselet, 2025};
+this integral is approximated by a sum over small \(h\) increments.
+Then, a p-value for each timestep \(t\) is computed by comparing its
+TFCE with the null distribution of TFCE values (obtained via
+permutation). For each permuted signal, we keep the maximal value over
+the whole signal for the null distribution of the TFCE. The TFCE
+combined with permutation (assuming a large enough number of
+permutations) has been shown to provide accurate FWER control (e.g.,
+\citeproc{ref-pernet2015}{Pernet et al., 2015}). However, further
+simulation work showed that cluster-based methods (including TFCE)
+perform poorly in localising the onset of M/EEG effects (e.g.,
+\citeproc{ref-rousselet_using_2025}{Rousselet, 2025};
 \citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}).
 
-To sum up this section, cluster-based inference main limitations is that
-it provides inference at the cluster level only, not allowing inference
+To sum up, the main limitation of cluster-based inference is that it
+allows for inference at the cluster level only, not allowing inference
 at the level of timesteps, sensors, etc. As a consequence, it does not
-allow inferring the localisation of effects (for more details on
-cluster-based inference, see for instance
-\citeproc{ref-maris2011}{Maris, 2011}; \citeproc{ref-maris2007}{Maris \&
-Oostenveld, 2007}; \citeproc{ref-sassenhagen2019}{Sassenhagen \&
-Draschkow, 2019}). In the following, we briefly review previous M/EEG
-modelling work. Then, we provide a brief introduction to generalised
-additive models (GAMs) and Bayesian generalised additive multilevel
-models (BGAMMs) to illustrate how these models can be used to precisely
+allow inferring the precise spatial and temporal localisation of
+effects. In the following, we briefly review previous M/EEG modelling
+work. Then, we provide a short introduction to generalised additive
+models (GAMs) and Bayesian generalised additive multilevel models
+(BGAMMs) to illustrate how these models can be used to precisely
 localise the onset and offset of M/EEG effects.
@@
 \citeproc{ref-meulman2015}{Meulman et al., 2015},
-\citeproc{ref-meulman2023}{2023})\ldots{}
+\citeproc{ref-meulman2023}{2023})\ldots{} See also the UNFOLD toolbox
+(\citeproc{ref-ehinger_unfold_2019}{Ehinger \& Dimigen, 2019})\ldots{}
 
@@
 ``wiggliness'' of the function. In other words, more complex (wiggly)
-basis function are penalised.
+basis functions are penalised.
 
@@
 not properly accounting for between-participant variability. We next fit
-a multilevel version of the BGAM (BGAMM, see
+a multilevel version of the BGAM (BGAMM, for an introduction to Bayesian
+multilevel models in \texttt{brms}, see
 \citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}). Although it is
@@
 \NormalTok{    )}
+
+\CommentTok{\# fitting the BGAMM}
+\CommentTok{\# meta\_gamm \textless{}{-} brm(}
+\CommentTok{\#     \# using by{-}participant SD of ERPs across trials}
+\CommentTok{\#     eeg\_mean | se(eeg\_sd) \textasciitilde{} condition +}
+\CommentTok{\#         s(participant, bs = "re") +}
+\CommentTok{\#         s(time, participant, bs = "fs", by = condition, m = 1),}
+\CommentTok{\#     data = summary\_df,}
+\CommentTok{\#     family = gaussian(),}
+\CommentTok{\#     warmup = 2000,}
+\CommentTok{\#     iter = 5000,}
+\CommentTok{\#     chains = 8,}
+\CommentTok{\#     cores = 8,}
+\CommentTok{\#     control = list(adapt\_delta = 0.95),}
+\CommentTok{\#     file = "models/meta\_gamm.rds"}
+\CommentTok{\#     )}
 \end{Highlighting}
@@
 posterior probability ratio thus provides a model-based approach for
-estimating the onset and offset of M/EEG effects.
+estimating the onset and offset of M/EEG effects. An important advantage
+is that the proposed approach can be extended to virtually any model
+structure. Moreover, the approach provides both group-level and
+individual-level onset and offset estimates of M/EEG effects.
 
@@
 
-We then compared the ability of the BGAMM to correctly estimate the
+We then compared the ability of the BGAMM to accurately estimate the
 onset and offset of the ERP difference to other widely-used methods.
@@
 after applying different forms of multiplicity correction to the
-p-values. We compared two methods that control the false discovery rate
-(FDR) (i.e., BH95, \citeproc{ref-benjamini1995}{Benjamini \& Hochberg,
-1995}; and BY01, \citeproc{ref-benjamini2001}{Benjamini \& Yekutieli,
-2001}), one method that controls the family-wise error rate (FWER)
-(i.e., Holm--Bonferroni method, \citeproc{ref-holm1979}{Holm, 1979}),
-and two cluster-based permutation methods (permutation with a single
-cluster-forming threshold and threshold-free cluster enhancement, TFCE,
-\citeproc{ref-smith2009}{S. Smith \& Nichols, 2009}). The BH95, BY01,
-and Holm corrections were applied to the p-values using the
-\texttt{p.adjust()} function in \texttt{R}. The cluster-based inference
-was implemented using a cluster-sum statistic of squared t-values, as
-implemented in \texttt{MNE-Python}
+\(p\)-values. We compared two methods that control the FDR (i.e.,
+\texttt{BH95}, \citeproc{ref-benjamini1995}{Benjamini \& Hochberg,
+1995}; and \texttt{BY01}, \citeproc{ref-benjamini2001}{Benjamini \&
+Yekutieli, 2001}), one method that controls the FWER (i.e.,
+Holm--Bonferroni method, \citeproc{ref-holm1979}{Holm, 1979}), and two
+cluster-based permutation methods (permutation with a single
+cluster-forming threshold and threshold-free cluster enhancement,
+\texttt{TFCE}, \citeproc{ref-smith2009}{S. Smith \& Nichols, 2009}). The
+\texttt{BH95}, \texttt{BY01}, and \texttt{Holm} corrections were applied
+to the p-values using the \texttt{p.adjust()} function in \texttt{R}.
+The cluster-based inference was implemented using a cluster-sum
+statistic of squared \(t\)-values, as implemented in \texttt{MNE-Python}
 (\citeproc{ref-gramfort2013}{Gramfort, 2013}), called via the \texttt{R}
@@
 (\citeproc{ref-changepoint}{Killick et al., 2022}), and applied directly
-to the squared t-values (as in
+to the squared \(t\)-values (as in
 \citeproc{ref-rousselet_using_2025}{Rousselet, 2025}).
@@
 (\citeproc{ref-rousselet_using_2025}{2025}), each participant was
-assigned a random onset between 150 and 170ms.
+assigned a random onset between 150 and 170ms. Whereas the present
+article focuses on one-dimensional signas (e.g., one M/EEG channel), we
+provice a 2D application in Section~\ref{sec-2D}.
 
@@
 result, we obtain a timecourse of decoding performance (ROC AUC),
-bounded between 0 and 1, for each participant (for a total of 32
-participants). Next, we wanted to \emph{test} whether the group-level
-average decoding accuracy is above chance (i.e., 0.5) at each timestep
+bounded between 0 and 1, for each participant (\(N=32\)). Next, we
+wanted to \emph{test} whether the group-level average decoding accuracy
+is above chance (i.e., 0.5) at each timestep
 (Figure~\ref{fig-decoding-data}). To achieve this, we fitted a BGAM as
@@
 according to the properties of the modelled data (e.g., signal-to-noise
-ratio, prior low-pass filtering, sampling rate, etc) and should be
-assessed by the usual model checking tools (e.g., posterior predictive
-checks, see also Section~\ref{sec-basis}). To better distinguish signal
-from noise, we also defined a region of practical equivalence (ROPE,
+ratio, prior low-pass filtering, sampling rate) and should be assessed
+by the usual model checking tools (e.g., posterior predictive checks,
+see also Section~\ref{sec-basis}). To better distinguish signal from
+noise, we also defined a region of practical equivalence (ROPE,
 \citeproc{ref-kruschke2017}{Kruschke \& Liddell, 2017}), defined as the
@@
 {\caption{{Summary statistics of the onset and offset estimates for each
-method (ordered by MAE).}{\label{tbl-simulation-results}}}
-\vspace{-20pt}}
+method (ordered by the absolute value of the
+bias).}{\label{tbl-simulation-results}}}}
 
@@
 Change point & 0.0060 & 0.0060 & 0.0051 & 0.0000 & 0.0059 \\ 
+Raw p-value & 0.0060 & 0.1250 & 0.0711 & 0.0252 & 0.1942 \\ 
+FDR BH95 & 0.0080 & 0.0460 & 0.0599 & 0.0192 & 0.0801 \\ 
 FDR BY01 & 0.0120 & 0.0120 & 0.0443 & 0.0087 & 0.0059 \\ 
@@
 Holm & 0.0140 & 0.0140 & 0.0270 & 0.0019 & 0.0059 \\ 
-FDR BH95 & 0.0080 & 0.0460 & 0.0599 & 0.0192 & 0.0801 \\ 
-Raw p-value & 0.0060 & 0.1250 & 0.0711 & 0.0252 & 0.1942 \\ 
 \midrule\addlinespace[2.5pt]
@@
 Cluster mass & -0.0080 & 0.0080 & 0.0084 & 0.0000 & 0.0059 \\ 
+Raw p-value & -0.0080 & 0.1180 & 0.0725 & 0.0252 & 0.1868 \\ 
 Change point & -0.0100 & 0.0100 & 0.0096 & 0.0000 & 0.0059 \\ 
+FDR BH95 & -0.0100 & 0.0360 & 0.0578 & 0.0189 & 0.0682 \\ 
 FDR BY01 & -0.0140 & 0.0150 & 0.0204 & 0.0056 & 0.0059 \\ 
@@
 Holm & -0.0180 & 0.0180 & 0.0262 & 0.0016 & 0.0030 \\ 
-FDR BH95 & -0.0100 & 0.0360 & 0.0578 & 0.0189 & 0.0682 \\ 
-Raw p-value & -0.0080 & 0.1180 & 0.0725 & 0.0252 & 0.1868 \\ 
 \bottomrule
@@
 
+\begin{table}
+
+{\caption{{Summary statistics of the onset and offset estimates for each
+method (ordered by the absolute value of the
+bias).}{\label{tbl-reliability-results}}}}
+
+\fontsize{9.0pt}{10.8pt}\selectfont
+\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lccccc}
+\toprule
+ & Bias & MAE & RMSE & Variance & MAD \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{6}{l}{onset} \\[2.5pt] 
+\midrule\addlinespace[2.5pt]
+BGAM (new method) & 0.0000 & 0.0000 & 0.0106 & 0.0008 & 0.0000 \\ 
+Cluster mass & 0.0000 & 0.0000 & 0.0019 & 0.0000 & 0.0000 \\ 
+Change point & 0.0050 & 0.0050 & 0.0011 & 0.0001 & 0.0074 \\ 
+TFCE & 0.0100 & 0.0100 & 0.0120 & 0.0001 & 0.0074 \\ 
+FDR BY01 & 0.0100 & 0.0100 & 0.0065 & 0.0023 & 0.0000 \\ 
+Holm & 0.0100 & 0.0100 & 0.0118 & 0.0001 & 0.0074 \\ 
+FDR BH95 & 0.0250 & 0.0900 & 0.0101 & 0.0076 & 0.0964 \\ 
+Raw p-value & -0.0300 & 0.0650 & 0.0201 & 0.0062 & 0.0815 \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{6}{l}{offset} \\[2.5pt] 
+\midrule\addlinespace[2.5pt]
+Raw p-value & 0.0000 & 0.0300 & 0.0256 & 0.0047 & 0.0445 \\ 
+BGAM (new method) & -0.0050 & 0.0750 & 0.0716 & 0.0274 & 0.1038 \\ 
+Change point & 0.0250 & 0.0650 & 0.0566 & 0.0090 & 0.1186 \\ 
+FDR BH95 & -0.0750 & 0.0750 & 0.1053 & 0.0153 & 0.1483 \\ 
+TFCE & -0.1150 & 0.1450 & 0.1094 & 0.0146 & 0.1260 \\ 
+Holm & -0.1150 & 0.1450 & 0.1091 & 0.0133 & 0.1186 \\ 
+FDR BY01 & -0.1450 & 0.1450 & 0.1197 & 0.0154 & 0.0890 \\ 
+Cluster mass & -0.1650 & 0.1650 & 0.1448 & 0.0051 & 0.0297 \\ 
+\bottomrule
+\end{tabular*}
+
+\end{table}
+
 \newpage
@@
 based on the initial results of S. Smith \& Nichols
-(\citeproc{ref-smith2009}{2009})\ldots{}
+(\citeproc{ref-smith2009}{2009})\ldots{} we did not include the
+cluster-depth algorithm (\citeproc{ref-frossard2022}{Frossard \& Renaud,
+2022}), as Rousselet (\citeproc{ref-rousselet_using_2025}{2025}) already
+showed its performance were worse than the cluster mass
+algorithm\ldots{}
 
@@
 all methods\ldots{}
+
+Discussion about the nature of the model: we modelled the surface M/EEG
+signals, however, the true interests (probably) lie in the brain, that
+is, in the source space\ldots{} we could build a ``full'' Bayesian model
+of the generated EEG signal (i.e., including hypotheses about the source
+and a forward model), but this model would become computationally
+heavier\ldots{}
 
@@
 
+\bibitem[\citeproctext]{ref-bullmore1999}
+Bullmore, E. T., Suckling, J., Overmeyer, S., Rabe-Hesketh, S., Taylor,
+E., \& Brammer, M. J. (1999). Global, voxel, and cluster tests, by
+theory and permutation, for a difference between two groups of
+structural MR images of the brain. \emph{IEEE Transactions on Medical
+Imaging}, \emph{18}(1), 32--42. \url{https://doi.org/10.1109/42.750253}
+
 \bibitem[\citeproctext]{ref-brms2017}
@@
 
+\bibitem[\citeproctext]{ref-ehinger_unfold_2019}
+Ehinger, B. V., \& Dimigen, O. (2019). Unfold: An integrated toolbox for
+overlap correction, non-linear modeling, and regression-based {EEG}
+analysis. \emph{PeerJ}, \emph{7}, e7838.
+\url{https://doi.org/10.7717/peerj.7838}
+
 \bibitem[\citeproctext]{ref-fischer2013}
@@
 \url{https://doi.org/10.1016/j.neuron.2013.07.006}
+
+\bibitem[\citeproctext]{ref-frossard2022}
+Frossard, J., \& Renaud, O. (2022). The cluster depth tests: Toward
+point-wise strong control of the family-wise error rate in massively
+univariate tests with application to M/EEG. \emph{NeuroImage},
+\emph{247}, 118824.
+\url{https://doi.org/10.1016/j.neuroimage.2021.118824}
 
@@
 \section{Application to 2D time-resolved decoding results
-(cross-temporal
-generalisation)}\label{application-to-2d-time-resolved-decoding-results-cross-temporal-generalisation}
+(cross-temporal generalisation)}\label{sec-2D}
 
@@
     \CommentTok{\# 2D thin{-}plate spline (tp)}
+    \CommentTok{\# auc \textasciitilde{} t2(train\_time, test\_time, bs = "tp", k = 10),}
 \NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
@@
     \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
-    \AttributeTok{iter =} \DecValTok{2000}\NormalTok{,}
+    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
+    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
     \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
     \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2.rds"} \CommentTok{\# k = 10}
-    \CommentTok{\# file = "models/timegen\_gam\_t2\_k20.rds" \# k = 20}
+    \CommentTok{\# file = "models/timegen\_gam\_t2.rds" \# k = 10}
+    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2\_k20.rds"} \CommentTok{\# k = 20}
 \NormalTok{    )}
-
-\CommentTok{\# fitting a GP with two temporal dimensions}
-\CommentTok{\# timegen\_gp \textless{}{-} brm(}
-\CommentTok{\#     auc \textasciitilde{} gp(train\_time, test\_time, k = 20),}
-\CommentTok{\#     data = timegen\_data,}
-\CommentTok{\#     family = Beta(),}
-\CommentTok{\#     control = list(adapt\_delta = 0.95),}
-\CommentTok{\#     iter = 2000,}
-\CommentTok{\#     chains = 4,}
-\CommentTok{\#     cores = 4,}
-\CommentTok{\#     file = "models/timegen\_gp.rds"}
-\CommentTok{\#     )}
 \end{Highlighting}
@@
 
-\caption{Predicted AUC values (left) and posterior probabilies of
+\caption{Simulated data with thresholds derived from the BGAM (left),
+predicted AUC values (middle), and ratios of posterior probabilies of
 decoding accuracy being above chance level (right) according to the
-bivariate GAM.}
+bivariate BGAM (for k=10, top row, and k=20, bottom row).}
 
@@
 \CommentTok{\# using the testing\_through\_time() function from the neurogam package}
-\CommentTok{\# this may take a few minutes (or hours depending the machine\textquotesingle{}s performance}
-\CommentTok{\# and data size)...}
+\CommentTok{\# this may take a few minutes (or hours depending the machine\textquotesingle{}s}
+\CommentTok{\# performance and data size)...}
 \NormalTok{gam\_onset\_offset }\OtherTok{\textless{}{-}} \FunctionTok{testing\_through\_time}\NormalTok{(}
@@
     \AttributeTok{threshold =} \DecValTok{20}\NormalTok{,}
-    \CommentTok{\# the *\_id arguments are use to specify the relevant columns in data }
+    \CommentTok{\# the *\_id arguments are used to specify the relevant columns in data }
     \AttributeTok{participant\_id =} \StringTok{"participant"}\NormalTok{, }\AttributeTok{meeg\_id =} \StringTok{"eeg"}\NormalTok{,}
@@
 object. Note that the code used to reshape the \texttt{sample\ MNE}
-dataset is provided later, and we refer to the
+dataset is available in the online supplementary materials, and we refer
+to the
 \href{https://mne.tools/stable/auto_tutorials/epochs/50_epochs_to_data_frame.html}{MNE
 documentation} about converting \texttt{MNE} epochs to \texttt{Pandas}
-dataframes in long format.
+dataframes in long format (i.e., with one observation per row).
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# loading the Python modules}
-\ImportTok{import}\NormalTok{ mne}
-\ImportTok{from}\NormalTok{ mne.datasets }\ImportTok{import}\NormalTok{ sample}
-\ImportTok{import}\NormalTok{ numpy }\ImportTok{as}\NormalTok{ np}
-\ImportTok{import}\NormalTok{ pandas }\ImportTok{as}\NormalTok{ pd}
-
-\CommentTok{\# defining the path to the "sample" dataset}
-\NormalTok{path }\OperatorTok{=}\NormalTok{ sample.data\_path()}
-
-\CommentTok{\# loading the evoked data}
-\NormalTok{evokeds }\OperatorTok{=}\NormalTok{ mne.read\_evokeds(path }\OperatorTok{/} \StringTok{"MEG"} \OperatorTok{/} \StringTok{"sample"} \OperatorTok{/} \StringTok{"sample\_audvis{-}ave.fif"}\NormalTok{)}
-
-\CommentTok{\# defining a function to reshape the data}
-\KeywordTok{def}\NormalTok{ reshape\_eeg\_channels\_as\_participants(evokeds, condition\_labels):}
-
-\NormalTok{    all\_dfs }\OperatorTok{=}\NormalTok{ []}
-
-    \ControlFlowTok{for}\NormalTok{ evoked, condition }\KeywordTok{in} \BuiltInTok{zip}\NormalTok{(evokeds, condition\_labels):}
-
-        \CommentTok{\# selecting only EEG channels}
-\NormalTok{        picks }\OperatorTok{=}\NormalTok{ mne.pick\_types(evoked.info, meg}\OperatorTok{=}\VariableTok{False}\NormalTok{, eeg}\OperatorTok{=}\VariableTok{True}\NormalTok{)}
-\NormalTok{        data }\OperatorTok{=}\NormalTok{ evoked.data[picks, :]}
-\NormalTok{        channel\_names }\OperatorTok{=}\NormalTok{ np.array(evoked.ch\_names)[picks]}
-
-\NormalTok{        n\_channels, n\_times }\OperatorTok{=}\NormalTok{ data.shape}
-        \CommentTok{\# repeating time for each channel}
-\NormalTok{        times }\OperatorTok{=}\NormalTok{ np.tile(evoked.times, n\_channels)}
-\NormalTok{        meeg\_values }\OperatorTok{=}\NormalTok{ data.flatten()}
-\NormalTok{        pseudo\_participant\_ids }\OperatorTok{=}\NormalTok{ np.repeat(channel\_names, n\_times)}
-
-        \CommentTok{\# converting to dataframe}
-\NormalTok{        df }\OperatorTok{=}\NormalTok{ pd.DataFrame(\{}
-            \StringTok{"participant"}\NormalTok{: pseudo\_participant\_ids,}
-            \StringTok{"time"}\NormalTok{: times,}
-            \StringTok{"eeg"}\NormalTok{: meeg\_values,}
-            \StringTok{"condition"}\NormalTok{: condition}
-\NormalTok{            \})}
-
-\NormalTok{        all\_dfs.append(df)}
-
-    \ControlFlowTok{return}\NormalTok{ pd.concat(all\_dfs, ignore\_index}\OperatorTok{=}\VariableTok{True}\NormalTok{)}
-
-\CommentTok{\# picking two evoked conditions}
-\NormalTok{faces }\OperatorTok{=}\NormalTok{ evokeds[}\DecValTok{0}\NormalTok{]}
-\NormalTok{scrambled }\OperatorTok{=}\NormalTok{ evokeds[}\DecValTok{1}\NormalTok{]}
-
-\CommentTok{\# reshaping data by pretending each EEG channel is a participant}
-\NormalTok{reshaped\_df }\OperatorTok{=}\NormalTok{ reshape\_eeg\_channels\_as\_participants(}
-\NormalTok{    evokeds}\OperatorTok{=}\NormalTok{[faces, scrambled],}
-\NormalTok{    condition\_labels}\OperatorTok{=}\NormalTok{[}\StringTok{"faces"}\NormalTok{, }\StringTok{"scrambled"}\NormalTok{]}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
 \end{document}

### 2025-04-11 code/split_half_reliability_cluster.R

@@ -2,7 +2,7 @@
 # Split-half reliability of onset/offset estimates        #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on April 9, 2025                           #
+# Last updated on April 11, 2025                          #
 ###########################################################
 
 # importing R packages

### 2025-05-22 manuscript.qmd
pushing a first version of intro+methods+results

--- 2025-04-11 brms_meeg.qmd
@@
     conflict-of-interest: "The authors have no conflicts of interest to disclose."
-abstract: "Time-resolved electrophysiological measurements such as those obtained through magneto- or electro-encephalography (M/EEG) offer a unique window into the neural activity underlying cognitive processes.
+abstract: "Time-resolved electrophysiological measurements such as those obtained through magneto- or electro-encephalography (M/EEG) offer a unique window onto the neural activity underlying cognitive processes.
 Researchers are often interested in determining whether and when these signals differ across experimental conditions or participant groups.
 The conventional approach involves mass-univariate statistical testing across time and space, followed by corrections for multiple comparisons such as cluster-based inference.
-While effective for controlling error rates at the cluster-level, cluster-based inference comes with a significant limitation: by shifting the focus of inference from individual time points to clusters, it makes difficult to draw precise conclusions about the onset or offset of observed effects.
+While effective for controlling error rates at the cluster-level, cluster-based inference comes with a significant limitation: by shifting the focus of inference from individual time points to clusters, it prevents drawing conclusions about the precise onset or offset of observed effects.
 Here, we introduce a *model-based* approach for analysing M/EEG timeseries such as event-related potentials (ERPs) or decoding performance over time.
@@
 
+# disabling hyperlinks
+# options(cli.hyperlink = FALSE)
+
 # setting a default ggplot2 theme
@@
 These techniques provide rich time series data that reflect how neural responses unfold in response to stimuli or tasks.
-A central goal in many M/EEG studies is to determine whether, when, and where neural responses differ across experimental conditions or participant groups.
-
-The conventional approach involves mass-univariate statistical testing through time and/or space followed by some form or correction for multiple comparisons with the goal of maintaining the familywise error rate (FWER) or the false discovery rate (FDR) at the nominal level (e.g., 5%).
+A central goal in many M/EEG studies is to determine whether, when, and where neural responses differ across experimental conditions or groups.
+
+The conventional approach involves mass-univariate statistical testing through time and/or space followed by some form of correction for multiple comparisons with the goal of maintaining the familywise error rate (FWER) or the false discovery rate (FDR) at the nominal level (e.g., 5%).
 Cluster-based inference is the most common way of achieving this sort of error control in the M/EEG literature, being the recommended approach in several software programs [e.g., `EEGlab`, @delorme2004; `MNE-Python`, @gramfort2013].
@@
 As pointed out by @maris2007: "there is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA [false alarm] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect".
-Even worse, @rosenblatt2018 note that cluster-based inference suffers from low spatial resolution: "Since discovering a cluster means that 'there exists at least one voxel with an evoked response in the cluster', and not that 'all the voxels in the cluster have an evoked response', it follows that the larger the detected cluster, the less information we have on the location of the activation." As a consequence, cluster-based inference is expected to perform poorly for localising the onset of M/EEG effects; a property that was later demonstrated in simulations studies [e.g., @rousselet_using_2025; @sassenhagen2019].
+Even worse, @rosenblatt2018 note that cluster-based inference suffers from low spatial resolution: "Since discovering a cluster means that 'there exists at least one voxel with an evoked response in the cluster', and not that 'all the voxels in the cluster have an evoked response', it follows that the larger the detected cluster, the less information we have on the location of the activation." As a consequence, cluster-based inference is expected to perform poorly for localising the onset of M/EEG effects; a property that was later demonstrated in simulation studies [e.g., @rousselet_using_2025; @sassenhagen2019].
 
@@
 
-The issues with multiple comparisons represent a common and well-recognised danger in neuroimaging and M/EEG research where the collected data allows for a multitude of potential hypothesis tests and is characterised by complex structures of spatiotemporal dependencies.
+The issues with multiple comparisons represent a common and well-recognised danger in neuroimaging and M/EEG research, where the collected data allows for a multitude of potential hypothesis tests and is characterised by complex structures of spatiotemporal dependencies.
 The probability of obtaining at least one false positive in an ensemble (family) of $m$ tests (i.e., the FWER) is computed as $1-\left(1-\alpha\right)^{m}$ (for $m=10$ tests and $\alpha=0.05$, it is approximately equal to $0.4$).
 Different methods exist to control the FWER (i.e., bring it back to $\alpha$).
-Most methods apply a simple correction to series of "raw" $p$-values issued from univariate statistical tests (e.g., t-tests).
-For instance, the Bonferroni correction [@dunn1961] consists in setting the significance threshold to $\alpha/m$, or equivalently, multiplying the "raw" $p$-values by $m$ and using the standard $\alpha$ significance threshold.
+Most methods apply a simple correction to series of $p$-values issued from univariate statistical tests (e.g., t-tests).
+For instance, the Bonferroni correction [@dunn1961] consists in setting the significance threshold to $\alpha/m$, or equivalently, multiplying the $p$-values by $m$ and using the standard $\alpha$ significance threshold.
 This method is generally overconservative (i.e., under-powered) as it assumes statistical independence of the tests, an assumption that is clearly violated in the context of M/EEG timeseries characterised by massive spatiotemporal dependencies.
@@
 As such, it would be incorrect to conclude, for instance, that the timestep of a significant cluster is the first moment at which some conditions differ [@frossard2022; @sassenhagen2019].
-In other words, because the inference is performed at the second step (i.e., once clusters have been formed), it prevents any conclusion to be made about individual datapoints (e.g., timesteps, sensors, etc).
+In other words, because the inference is performed at the second step (i.e., once clusters have been formed), no conclusion can be made about individual datapoints (e.g., timesteps, sensors, etc).
 
@@
 where $h_{0}$ is typically $0$ and parameters $E$ and $H$ are set a priori (typically to 0.5 and 2, respectively) and control the influence of the extend and height on the TFCE.
-Note that in practise, this integral is approximated by a sum over small $h$ increments.
-Then, a p-value for each timestep $t$ is computed by comparing its TFCE with the null distribution of TFCE values (obtained via permutation).
+Note that in practice, this integral is approximated by a sum over small $h$ increments.
+Then, a *p*-value for each timestep $t$ is computed by comparing its TFCE with the null distribution of TFCE values (obtained via permutation).
 For each permuted signal, we keep the maximal value over the whole signal for the null distribution of the TFCE.
@@
 As a consequence, it does not allow inferring the precise spatial and temporal localisation of effects.
-In the following, we briefly review previous M/EEG modelling work.
-Then, we provide a short introduction to generalised additive models (GAMs) and Bayesian generalised additive multilevel models (BGAMMs) to illustrate how these models can be used to precisely localise the onset and offset of M/EEG effects.
+In the following, we briefly review previous modelling work of M/EEG data.
+Then, we provide a short introduction to generalised additive models (GAMs) to illustrate how these models can be used to precisely estimate the onset and offset of M/EEG effects.
 
@@
 
-Recent example of GLM for EEG [@wüllhorst2025; @fischer2013]...
-See also [@rousselet2008; @hauk2006]...
-Example of two-stage regression analysis [i.e., individual-level then group-level, @dunagan2024]...
-
-As put by @rousselet_using_2025, concluding on the onset of effect based on a series of univariate tests...
-commits to three fallacies...
-here, we want to avoid these by introducing a model-based approach, which naturally take into account the temporal dependencies in the data to output a series of posterior probabilities...
-
-See also the rERP framework [@smith2014a; @smith2014b] and @tremblay2014...
-
-From @dimigen2021: Recently, spline regression has been applied to ERPs [e.g., @hendrix2017; @kryuchkova2012]...
-GAMMs for EEG data [@abugaber2023; @meulman2015; @meulman2023]...
-See also the UNFOLD toolbox [@ehinger_unfold_2019]...
-
-Disentangling overlapping processes [@skukies_brain_2024; @skukies_modelling_2021]...
-Weighting single trials [@pernet2022]...
-The LIMO toolbox [@pernet2011] using linear and 2-stage regression (not a proper multilevel model)...
-
-Recently, @teichmann2022 provided a detailed tutorial on using Bayes factors (BFs) to analyse the 1D or 2D output from MVPA, that is, for testing, at every timestep, whether decoding performance is above chance level.
-However, this approach provides timeseries of BFs that ignores temporal dependencies...
+<!--
+
+Weighting single trials [@pernet2022]... From @dimigen2021: Recently, spline regression has been applied to ERPs [e.g., @hendrix2017; @kryuchkova2012]...
+
+M/EEG recordings at the scalp measure activity stemming from various parts of the brain and contain many artefacts not related to the cognitive and neural processes of interest.
+Therefore, analysing M/EEG data requires methods for separating the parts of the signals that are related to cognitive and neural processes and the parts of the signals that contain task-unrelated "noise".
+The most widely-used method for achieving this separation consists in estimating the event-related potential (ERP), defined as the brain response elicited by some specific time-locked sensory, cognitive, or motor event.
+Most often, the ERP is computed as the average response over trials computed from epochs time-locked on the event of interest (e.g., stimulus onset).
+However, this simple averaging approach comes with several limitations, such as being limited to simple categorical designs.
+
+-->
+
+Scalp-recorded M/EEG signals capture neural activity originating from various brain regions and are often contaminated by artifacts unrelated to the cognitive processes under investigation.
+Consequently, analysing M/EEG data necessitates methods that can disentangle task-relevant neural signals from extraneous "noise." A widely adopted technique for this purpose is the estimation of event-related potentials (ERPs), which are stereotyped electrophysiological responses time-locked to specific sensory, cognitive, or motor events.
+Typically, ERPs are derived by averaging EEG or MEG epochs across multiple trials aligned to the event of interest (e.g., stimulus onset), thereby enhancing the signal-to-noise ratio by attenuating non-time-locked activity.
+However, this averaging approach has notable limitations: it assumes consistent latency and amplitude across trials and is primarily suited for simple categorical designs.
+Such assumptions may not hold in more complex experimental paradigms, potentially leading to suboptimal ERP estimations.
+
+<!--
+
+To overcome the limitations of averaging, a number of model-based options for estimating ERPs have been proposed.
+These approaches are justified by the observation that the averaging approach to ERPs is equivalent to an intercept-only linear regression in a simple categorical design with no overlapping factors [@smith2014a].
+The regression approach to ERPs is much more flexible than simpl averaging, as it allows modelling linear and non-linear effects of continuous predictors such as word frequency or age [e.g., @smith2014a; @smith2014b; @tremblay2014] and allow modelling and disentangling overlapping processes [e.g., @skukies_brain_2024; @skukies_modelling_2021] as implemented in the UNFOLD toolbox [@ehinger_unfold_2019].
+The regression approach, as implemented for instance in the LIMO toolbox [@pernet2011], is a multi-stage procedure in which a participant-level regression model is fitted independently to each datapoint (e.g., channel, timestep, etc.) to estimate the ERP, followed by group-level statistical analyses of the regression coefficients, itself followed by multiplicity corrections or cluster-based inference [for recent applied examples, see for instance @dunagan2025; @wüllhorst2025].
+Whereas this approach allows including linear and non-linear discrete and continuous predictors when estimating the ERP, it has several limitations.
+First, the participant-level modelling ignores spatiotemporal dependencies as it treats all datapoints independently.
+Second, this approach does not take into account the participant (or stimulus) variability when testing group-level hypotheses (as could be done for instance via multilevel modelling).
+Finally, because the output of this procedure is summarised by cluster-based inference, its conclusions suffer from the same limitations discussed in the previous section.
+
+-->
+
+To overcome the limitations of simple averaging, several model-based approaches for estimating ERPs have been proposed.
+These methods are motivated by the observation that traditional ERP averaging is mathematically equivalent to fitting an intercept-only linear regression model in a simple categorical design without overlapping events [@smith2014a].
+In contrast to simple averaging, regression-based approaches to ERP estimation offer substantially greater flexibility.
+Notably, they allow for the modelling of both linear and nonlinear effects of continuous predictors--such as word frequency or age [e.g., @smith2014a; @smith2014b; @tremblay2014]--and enable the disentangling of overlapping cognitive processes [e.g., @ehinger_unfold_2019; @skukies_brain_2024; @skukies_modelling_2021].
+One widely used implementation of this approach is provided by the LIMO EEG toolbox [@pernet2011], which follows a multi-stage analysis pipeline.
+First, a separate regression model is fit for each datapoint (e.g., each time point and electrode) at the individual participant level to estimate ERP responses.
+This is followed by group-level statistical analyses of the resulting regression coefficients, often accompanied by multiple comparison corrections or cluster-based inference [for recent applied examples, see @dunagan2025; @wüllhorst2025].
+
+Although this framework allows for the inclusion of a wide range of predictors--both continuous and categorical, linear and nonlinear--it still has important limitations.
+First, fitting separate models for each datapoint ignores the spatiotemporal dependencies inherent in M/EEG data, potentially reducing statistical power and interpretability.
+Second, the subsequent group-level analyses typically do not account for hierarchical dependencies, between-participant or between-stimulus variability, which could otherwise be addressed through multilevel modelling.
+Finally, because the output of this procedure is summarised by cluster-based inference, its conclusions remain subject to the limitations discussed in the previous section.
+
+<!--
+
+Building upon the regression framework, recent advances allow modelling and disentangling overlapping processes [e.g., @skukies_brain_2024; @skukies_modelling_2021] as implemented in the UNFOLD toolbox [@ehinger_unfold_2019]...
+
+In addition to modelling the non-linear effects of continuous predictors (e.g., word frequency) on the ERP amplitude, GAMs have been used for modelling the effect of time, that is, the shape of the ERP itself [@abugaber2023; @baayen2018; @meulman2015; @meulman2023]...
+In the next section, we turn to a brief introduction to generalised additive models and discuss their use for modelling M/EEG timeseries...
+
+-->
+
+Beyond modelling nonlinear effects of continuous predictors on ERP amplitudes, generalised additive models (GAMs) have been employed to capture the temporal dynamics of ERPs themselves, effectively modelling the shape of the waveform over time [@abugaber2023; @baayen2018; @meulman2015; @meulman2023].
+This approach allows for the estimation of smooth, data-driven functions that characterise how neural responses evolve over time, offering a flexible alternative to traditional linear models.
+In the following section, we provide a brief introduction to GAMs, highlighting their applicability to M/EEG time series analysis and the advantages they offer over conventional methods.
 
@@
 
-See for instance these tutorials [@sóskuthy2017; @winter2016] or application to phonetic data [@wieling2018; @sóskuthy2021] or this introduction [@baayen2020] or these reference books [@hastie2017; @wood2017]...
-application to pupillometry [@vanrij2019]...
-GAMLSS for neuroimaging data [@dinga2021]...
-Modelling auto-correlation in GAMMs + EEG example [@baayen2018]...
-
-In generalised additive models (GAMs), the functional relationship between predictors and response variable is decomposed into a sum of low-dimensional non-parametric functions.
+In generalised additive models (GAMs), the functional relationship between the predictors and the response variable is decomposed into a sum of low-dimensional non-parametric functions.
 A typical GAM has the following form:
@@
 where $y_{i} \sim \mathrm{EF}\left(\mu_{i}, \phi\right)$ denotes that the observations $y_{i}$ are distributed as some member of the exponential family of distributions (e.g., Gaussian, Gamma, Beta, Poisson) with mean $\mu_{i}$ and scale parameter $\phi$; $g(\cdot)$ is the link function, $\mathbf{A}_{i}$ is the $i$th row of a known parametric model matrix, $\gamma$ is a vector of parameters for the parametric terms (to be estimated), $f_{j}$ is a smooth function of covariate $x_{j}$ (to be estimated as well).
-The smooth functions $f_{j}$ are represented in the model via penalised splines basis expansions of the covariates, that are a weighted sum of $K$ simpler, basis functions:
+The smooth functions $f_{j}$ are represented in the model as a weighted sum of $K$ simpler, basis functions:
 
@@
 where $\beta_{jk}$ is the weight (coefficient) associated with the $k$th basis function $b_{jk}()$ evaluated at the covariate value $x_{ij}$ for the $j$th smooth function $f_{j}$.
-To clarify the terminology at this stage: *splines* are functions composed of simpler functions.
-These simpler functions are basis functions (e.g., cubic polynomial, thin-plate) and the set of basis functions is a *basis*.
-Each basis function gets its coefficient and the resultant spline is the sum of these weighted basis functions (@fig-intro-gam).
-Splines coefficients are penalised (usually through the squared of the smooth functions' second derivative) in a way that can be interpreted, in Bayesian terms, as a prior on the "wiggliness" of the function.
-In other words, more complex (wiggly) basis functions are penalised.
+To clarify the terminology at this point: *splines* are functions composed of simpler functions.
+These simpler functions are called *basis functions* (e.g., cubic polynomial, thin-plate) and the set of basis functions is called a *basis*.
+Each basis function gets its coefficient and the resultant spline is the sum of these weighted basis functions (@fig-intro-gam\ignorespaces A).
+Splines coefficients are penalised (usually through the squared of the smooth functions' second derivative) in a way that can be interpreted, in Bayesian terms, as a prior on the "wiggliness" of the function [@miller2025; @wood2017].
+In other words, more complex (wiggly) basis functions are automatically penalised.
+
+<!--
+
+A detailed discussion of the technical aspects of smooth functions is beyond the scope of this article (see Wood, 2017a); or (Baayen, Vasishth, Kliegl, & Bates, 2017; van Rij, Vaci, Wurm, & Feldman, in press; Wieling, 2018, for an introduction), but in the context of this work, it can be thought of as a continuous, potentially wiggly but not abruptly changing line that is expressed over time.
+
+Generalised additive models can be understood in a Bayesian framework [@miller2025; @wood2017]...
 
@@
 
-The Bayesian approach to statistical modelling is characterised by its reliance on probability theory to [@gelman2020]...
-In this framework, all unknown entities are assigned probability distributions reflecting the uncertainty...
-These probability distributions are commonly referred to as "priors" and represent some state of knowledge about unknown quantities before seeing any data.
 There are debates among Bayesian practitioners as to whether prior distributions should encore subjective (personal) beliefs or...
@@
 In practice, weakly informative priors are often used as default priors in situations in which subjective priors are difficult to define/elicit...
+
+The Bayesian approach to statistical modelling is characterised by its reliance on probability theory to quantify uncertainty about parameter estimates [e.g., @gelman2020].
+In this framework, all unknown entities are assigned probability distributions reflecting the uncertainty...
+These probability distributions are commonly referred to as "priors" and represent some state of knowledge about unknown quantities before seeing any data.
 Bayesian models are then fitted on empirical (actual or simulated) data to update prior states of knowledge to posterior states of knowledge using Bayes theorem, or in practise, sampling-based approximations of the posterior distribution...
 
-```{r fig-intro-gam, echo = FALSE, fig.width = 9, fig.cap = "Different types of GAM(M)s.
+-->
+
+```{r fig-intro-gam, echo = FALSE, fig.width = 9, fig.cap = "Different types of GAMs.
 **A**: GAMs predictions are computed as the weigthed sum (in black) of basis functions (here thin-plate basis functions, in grey).
 **B**: Constant-effect GAM, with 5 participants in colours and the group-level prediction in black.
-**C**: Varying-intercept + varying-slope GAMM (with common smoother).
+**C**: Varying-intercept + varying-slope GAMM (with constant smoother).
 **D**: Varying-intercept + varying-slope + varying-smoother GAMM.
@@
 ########################
-# simulating som edata #
+# simulating some data #
 ########################
@@
         ylim(-2, +2) +
-        # scale_colour_manual(values = met.brewer(name = "Johnson", n = 10) ) +
         theme(legend.position = "none")
@@
         ylim(-5, 5) +
-        scale_colour_manual(values = met.brewer(name = "Johnson", n = 5) ) +
+        scale_colour_manual(values = met.brewer(name = "Johnson", n = n_subjects) ) +
         labs(x = "Predictor", y = "Response")
@@
 # plotting, note that the basis plot uses the fixed-effect model for simplicity
+# df2 <- df %>% filter(subject %in% c(unique(df$subject)[1:4]) )
 p1 <- plot_basis_functions(df, gam_fixed)
@@
 
-See @fig-intro-gam...
+<!--
+
+Or (Baayen, Vasishth, Kliegl, & Bates, 2017; van Rij, Vaci, Wurm, & Feldman, in press; Wieling, 2018, for an introduction), but in the context of this work, it can be thought of as a continuous, potentially wiggly but not abruptly changing line that is expressed over time.
+
+This penalises complexity automatically [@meulman2023; @wieling2018]...
+ERP itself [@abugaber2023; @baayen2018; @meulman2015; @meulman2023]...
+See for instance these tutorials [@sóskuthy2017; @winter2016] or application to phonetic data [@wieling2018; @sóskuthy2021]...
+or this introduction [@baayen2020] or these reference books [@hastie2017; @wood2017]...
+
+A detailed discussion of the technical aspects of GAMs is beyond the scope of this article [see reference books such as @hastie2017; @wood2017] but it should be noted that GAMs have been successfully been applied to various types of timeseries data throughout the cognitive sciences.
+For instance, it has been applied to pupillometry [e.g., @vanrij2019], articulography [e.g., @wieling2018], formant contours [e.g., @sóskuthy2021], neuroimaging data [e.g., @dinga2021] and ERPs [e.g., @abugaber2023; @baayen2018; @meulman2015; @meulman2023].
+The appeal of GAMs for modelling surface (sensor-level) M/EEG data comes from their ability to deal with the complex shape of the ERP without overfitting it.
+Generalising to scale and shape or "distributional GAMs" [@rigby2005; @umlauf2018], allowing for the specification of linear predictors for both the shape and scale parameters of many distributions [for an application to neuroimaging data, see @dinga2021]...
+
 Introduction to multilevel GAMs [@pedersen_hierarchical_2019]...
-Now describe the Bayesian GAMM [@miller2025]...
 Proper inclusion of varying/random effects in the model specification protects against overly wiggly curves [@baayen2020]...
-Generalising to scale and shape or "distributional GAMs" [@rigby2005; @umlauf2018] and applied to neuroimaging data [@dinga2021]...
-
-Instead of averaging, obtain the smooth ERP signal from multilevel GAM... less susceptible to outliers [@meulman2023]...
+@fig-intro-gam\ignorespaces D...
+Instead of averaging, we obtain the smooth ERP signal from multilevel GAM...
+which has been shown to be less susceptible to outliers [@meulman2023]...
+
+-->
+
+A detailed treatment of the technical underpinnings of GAMs is beyond the scope of this article [see reference books such as @hastie2017; @wood2017].
+However, it is worth emphasising that GAMs have been successfully applied to a wide range of time series data across the cognitive sciences, including pupillometry [e.g., @vanrij2019], articulography [e.g., @wieling2018], speech formant dynamics [e.g., @sóskuthy2021], neuroimaging data [e.g., @dinga2021], and event-related potentials [e.g., @abugaber2023; @baayen2018; @meulman2015; @meulman2023].
+Their appeal for modelling M/EEG data lies in their ability to flexibly capture the complex shape of ERP waveforms without overfitting, through the use of smooth functions constrained by penalisation.
+Recent extensions, such as distributional GAMs [@rigby2005; @umlauf2018], allow researchers to model not only the mean structure but also the variance (or scale) and other distributional properties as functions of predictors--a feature that has proven useful in modelling neuroimaging data [e.g., @dinga2021].
+Moreover, hierarchical or multilevel GAMs [@pedersen_hierarchical_2019] provide a principled way to account for the nested structure of M/EEG data (e.g., trials within participants), enabling the inclusion of varying intercepts, slopes, and smoothers (as illustrated in @fig-intro-gam\ignorespaces C-D).
+This approach mitigates the risk of overfitting and reduces the influence of outliers on smooth estimates [@baayen2020; @meulman2023].
+
+<!--
+
+In the following, we introduce our proposed approach based on Bayesian generalized additive multilevel models (Bayesian GAMMs), which builds on these strengths while providing fully probabilistic estimates of ERP dynamics.
+We illustrate how this framework enables direct inference on the presence and timing of neural effects at the trial level, across conditions and participants, and show that it outperforms conventional methods in both simulated and real M/EEG datasets.
+
+-->
 
@@
 
-Given the previously reported limitations of conventional methods to precisely identify the onset and offset of M/EEG effects (e.g., ERPs, decoding performance), we developed a model-based approach for estimating the onset and offset of such effects.
-To achieve this, we leveraged Bayesian generalised additive multilevel models (BGAMMs) fitted in `R` via the `brms` package and compared the performance of this approach to conventional methods on both simulated and actual M/EEG data.
+<!--
+
+Recently, @teichmann2022 provided a detailed tutorial on using Bayes factors (BFs) to analyse the 1D or 2D output from MVPA, that is, for testing, at every timestep, whether decoding performance is above chance level.
+However, this approach provides timeseries of BFs that ignores temporal dependencies...
+In the following...
+
+As put by @rousselet_using_2025, concluding on the onset of effect based on a series of univariate tests...
+commits to three fallacies...
+here, we want to avoid these by introducing a model-based approach, which naturally take into account the temporal dependencies in the data to output a series of posterior probabilities...
+
+Given the previously reported limitations of cluster-based inference to precisely identify the onset and offset of M/EEG effects (e.g., ERPs, decoding performance), we developed a model-based approach for estimating the onset and offset of M/EEG effects.
+To achieve this, we leveraged Bayesian generalised additive multilevel models fitted in `R` via the `brms` package and compared the performance of this approach to conventional methods on both simulated and actual M/EEG data.
+
+-->
+
+Cluster-based permutation tests are widely used in M/EEG research to identify statistically significant effects across time and space.
+However, these methods have notable limitations, particularly in accurately determining the precise onset and offset of neural effects.
+To address these limitations, we developed a model-based approach relying on Bayesian generalised additive multilevel models implemented in `R` via the `brms` package [@brms2017; @brms2018].
+We evaluated the performance of this approach against conventional methods using both simulated and actual M/EEG data.
+Our findings demonstrate that this method provides more precise and reliable estimates of effect onsets and offsets, offering a valuable tool for researchers seeking to enhance the temporal resolution of their analyses.
 
@@
 We simulated data for a group of 20 participants (with variable true onset) with 50 trials per participant and condition (@fig-eeg).
+All figures and simulation results can be reproduced using the `R` code available on GitHub: <https://github.com/lnalborczyk/brms_meeg>.
 
@@
 The error band represents the mean +/- 1 standard error of the mean."}
+# setting the seed for reproducible results
+set.seed(111)
+
 # importing R version of Matlab code from Yeung et al. (2004)
@@
 
-We then fitted a Bayesian GAM (BGAM) using the `brms` package [@brms2017; @brms2018; @nalborczyk2019].
-We used the default priors in `brms` (i.e., weakly informative priors).
+We then fitted a Bayesian GAM (BGAM) using the `brms` package [@brms2017; @brms2018; @nalborczyk2019] and default priors (i.e., weakly informative priors).
 We ran eight Markov Chain Monte-Carlo (MCMC) to approximate the posterior distribution, including each 5000 iterations and a warmup of 2000 iterations, yielding a total of $8 \times (5000-2000) = 24000$ posterior samples to use for inference.
@@
 ```{r gam, results = "hide"}
-# averaging across participants
+# computing the average EEG signals for each participant
 ppt_df <- raw_df %>%
     group_by(participant, condition, time) %>%
-    summarise(eeg = mean(eeg) ) %>%
+    summarise(eeg_mean = mean(eeg) ) %>%
     ungroup()
@@
     # thin-plate regression splines with k-1 basis functions
-    eeg ~ condition + s(time, bs = "tp", k = 20, by = condition),
+    eeg_mean ~ 1 + condition + s(time, bs = "tp", k = 20, by = condition),
     data = ppt_df,
@@
 However, the previous model only included constant (fixed) effects, thus not properly accounting for between-participant variability.
-We next fit a multilevel version of the BGAM [BGAMM, for an introduction to Bayesian multilevel models in `brms`, see @nalborczyk2019].
-Although it is possible to fit a BGAMM at the single-trial level, we present a computationally-lighter version of the model that is fitted directly on by-participant summary statistics (mean and SD), similar to what is done in meta-analysis.
+We next fitted a multilevel version of the BGAM [BGAMM, for an introduction to Bayesian multilevel models in `brms`, see @nalborczyk2019] including a varying intercept and slope for participant (but only with a constant smoother).
+Although it is possible to fit a BGAMM using data at the single-trial level, we present a computationally lighter version of the model that is fitted directly on by-participant summary statistics (mean and SD), similar to what is done in meta-analysis.
 
 ```{r meta-gam, results = "hide"}
-# averaging across participants
+# computing the mean and SD of EEG signals for each participant
 summary_df <- raw_df %>%
-    summarise(
-        eeg_mean = mean(eeg),
-        eeg_sd = sd(eeg),
-        .by = c(participant, condition, time)
-        )
+    group_by(participant, condition, time) %>%
+    summarise(eeg_mean = mean(eeg), eeg_sd = sd(eeg) ) %>%
+    ungroup()
 
@@
 meta_gam <- brm(
-    # using by-participant SD of ERPs across trials
+    # using the by-participant mean and SD of ERPs
     eeg_mean | se(eeg_sd) ~
-        condition + s(time, bs = "cr", k = 20, by = condition) +
-        (1 | participant),
+        1 + condition + (1 + condition | participant) +
+        s(time, bs = "tp", k = 20, by = condition),
     data = summary_df,
@@
     )
-
+```
+
+```{r meta-gamm, echo = FALSE, eval = FALSE}
 # fitting the BGAMM
-# meta_gamm <- brm(
-#     # using by-participant SD of ERPs across trials
-#     eeg_mean | se(eeg_sd) ~ condition +
-#         s(participant, bs = "re") +
-#         s(time, participant, bs = "fs", by = condition, m = 1),
+# gamm <- brm(
+#     # thin-plate regression splines with k-1 basis functions
+#     eeg_mean ~ 1 + condition + (1 + condition | participant) + s(time, bs = "tp", k = 20, by = condition),
+#     data = ppt_df,
+#     family = gaussian(),
+#     warmup = 2000,
+#     iter = 5000,
+#     chains = 8,
+#     cores = 8,
+#     file = "models/gamm.rds"
+#     )
+
+# fitting the BGAMM
+# meta_gam_k10 <- brm(
+#     # using the by-participant SD of ERPs
+#     eeg_mean | se(eeg_sd) ~
+#         1 + condition + (1 + condition | participant) +
+#         s(time, bs = "tp", k = 10, by = condition),
 #     data = summary_df,
@@
 #     cores = 8,
-#     control = list(adapt_delta = 0.95),
-#     file = "models/meta_gamm.rds"
+#     file = "models/meta_gam_k10.rds"
 #     )
+
+# comparing the two models
+# loo(meta_gam, meta_gam_k10)
+# library(easystats)
+# model_comp <- compare_performance(meta_gam, meta_gam_k10)
+# model_comp
+# plot(model_comp)
 ```
@@
         re_formula = NULL,
-        prob = 0.99
+        prob = 0.95
         )[[1]]
@@
     # computing the posterior probability for the slope over time
-    post_prob_slope <- model$data |>
-        add_epred_draws(object = model) |>
-        data.frame() |>
-        dplyr::select(participant, time, condition, .epred, .draw) |>
-        pivot_wider(names_from = condition, values_from = .epred) |>
-        mutate(epred_diff = cond2 - cond1) |>
+    post_prob_slope <- model$data %>%
+        add_epred_draws(object = model) %>%
+        data.frame() %>%
+        dplyr::select(participant, time, condition, .epred, .draw) %>%
+        pivot_wider(names_from = condition, values_from = .epred) %>%
+        mutate(epred_diff = cond2 - cond1) %>%
         # computing mean posterior probability at the group level
-        group_by(time) |>
+        group_by(time) %>%
         summarise(
@@
 We then compute the posterior probability of the slope for `condition` being above $0$ (@fig-post-prob-ratio, left).
-From this quantity, we then compute the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio, right).
-Note that a ratio of 10 means that the probability of the difference being above 0 is 10 times higher than the probability of the difference not being above 0, given the data, the priors, and other model's assumptions.
-Thresholding the posterior probability ratio thus provides a model-based approach for estimating the onset and offset of M/EEG effects.
+From this quantity, we compute the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio, right).
+Note that a ratio of 10 means that the probability of the difference being above 0 is 10 times higher than the probability of the difference not being above 0, given the data, the priors, and other model's assumptions.[^1] Thresholding the posterior probability ratio thus provides a model-based approach for estimating the onset and offset of M/EEG effects, whose properties will be assessed in the simulation study.
 An important advantage is that the proposed approach can be extended to virtually any model structure.
-Moreover, the approach provides both group-level and individual-level onset and offset estimates of M/EEG effects.
-
-```{r tests, echo = FALSE, eval = FALSE}
+
+[^1]: This posterior probability ratio, or posterior odds, is equivalent to a Bayes factor, assuming 1:1 prior odds.
+
+```{r tests2, echo = FALSE, eval = FALSE}
 # compute percentage in ROPE
@@
 # retrieving posterior draws for EEG over time
-gam <- full_model
-post_samples <- posterior_epred(object = gam, newdata = ppt_df, allow_new_levels = TRUE)
+# gam <- full_model
+# post_samples <- posterior_epred(object = gam, newdata = ppt_df, allow_new_levels = TRUE)
+gam <- meta_gam
+post_samples <- posterior_epred(object = gam, newdata = summary_df, allow_new_levels = TRUE)
 
@@
 # rope_range <- c(-0.1, 0.1)
-baseline_sd <- ppt_df %>%
-    pivot_wider(names_from = condition, values_from = eeg) %>%
+baseline_sd <- summary_df %>%
+    pivot_wider(names_from = condition, values_from = eeg_mean) %>%
     mutate(cond_diff = abs(cond1 - cond2) ) %>%
@@
 # storing the results in a dataframe
-rope_results <- data.frame(time = ppt_df$time, prop_in_rope = prop_in_rope)
+# rope_results <- data.frame(time = ppt_df$time, prop_in_rope = prop_in_rope)
+rope_results <- data.frame(time = summary_df$time, prop_in_rope = prop_in_rope)
 
@@
 # full_model <- gam
-full_model <- readRDS("models/gam.rds")
+# full_model <- readRDS("models/gam.rds")
+full_model <- meta_gam
 
 # fitting the null model (no 'condition' effect)
+# null_model <- brm(
+#     eeg ~ s(time, bs = "tp", k = 20),
+#     data = ppt_df,
+#     family = gaussian(),
+#     warmup = 2000,
+#     iter = 5000,
+#     chains = 8,
+#     cores = 8,
+#     file = "models/null_gam.rds"
+#     )
+
+# fitting the BGAMM
 null_model <- brm(
-    eeg ~ s(time, bs = "tp", k = 20),
-    data = ppt_df,
+    # using by-participant SD of ERPs across trials
+    eeg_mean | se(eeg_sd) ~
+        1 + (1 | participant) +
+        s(time, bs = "tp", k = 20),
+    data = summary_df,
     family = gaussian(),
@@
     chains = 8,
-    cores = 8,
-    file = "models/null_gam.rds"
+    cores = 8
+    # file = "models/meta_gamm.rds"
     )
@@
 
-# computing pointwise Bayes Factors as exp(lpd_diff)
+# computing pointwise Bayes factors as exp(lpd_diff)
 BF_10_pointwise <- exp(lpd_diff)
@@
     summarise(BF_10 = mean(LPD_diff), .by = time) %>%
+    # summarise(BF_10 = mean(BF_10), .by = time) %>%
     # ggplot(aes(x = time, y = value, colour = name) ) +
@@
 Timesteps above threshold (10) are highlighted in green.
-NB: the minimum and maximum possible ratio values are determined (bounded) by the number of available posterior samples in the model."}
+NB: the minimum and maximum possible ratio values are determined (bounded) by the number of posterior samples in the model."}
 # defining a function to compute and plot the posterior probability ratio
@@
     n_posterior_samples <- ndraws(model)
+
     prob_y_above_0 <- model$data %>%
@@
         # ensuring there is no 0 or +Inf values
-        mutate(prob_ratio = ifelse(is.infinite(prob_ratio), n_posterior_samples, prob_ratio) ) %>%
+        # mutate(prob_ratio = ifelse(is.infinite(prob_ratio), n_posterior_samples, prob_ratio) ) %>%
+        mutate(
+            prob_ratio = ifelse(
+                prob_ratio > n_posterior_samples, n_posterior_samples, prob_ratio
+                )
+            ) %>%
         mutate(prob_ratio = ifelse(prob_ratio == 0, 1 / n_posterior_samples, prob_ratio) )
@@
 # TO-DO: plotting the onset/offset estimate per participant?
+# see https://discourse.mc-stan.org/t/specifying-varying-effect-structures-with-hierarchical-generalised-additive-models-gams/35596
 #################################################################
 
-model <- meta_gamm
-plot_post_preds(model = meta_gamm) + plot_post_slope(model = meta_gamm)
+# fixed-effect GAM
+gam_fixed <- gam(
+    y ~ s(x),
+    data = df
+    )
+
+# random intercept + slope
+gam_int_slope <- gam(
+    y ~ s(x) +
+        s(subject, bs = "re") +
+        s(x, subject, bs = "re"),
+    data = df
+    )
+
+# random smooth per subject + global smooth
+gam_rand_smooth <- gam(
+    y ~ s(x) +
+        s(subject, bs = "re") +
+        s(x, subject, bs = "fs", m = 1),
+    data = df
+    )
+
+# fitting the BGAMM
+# see https://discourse.mc-stan.org/t/specifying-varying-effect-structures-with-hierarchical-generalised-additive-models-gams/35596/2
+# tic()
+# 5519.743 sec elapsed
+full_meta_gamm <- brm(
+    # using by-participant SD of ERPs across trials
+    eeg_mean | se(eeg_sd) ~ 1 + condition +
+        s(time, by = condition, bs = "tp", k = 20) +
+        s(time, participant, by = condition, bs = "fs", xt = list(bs = "tp") ),
+    data = summary_df,
+    family = gaussian(),
+    warmup = 2000,
+    iter = 5000,
+    chains = 8,
+    cores = 8,
+    control = list(adapt_delta = 0.95)
+    # file = "models/full_meta_gamm.rds"
+    )
+# toc()
+
+# model <- meta_gamm
+# plot_post_preds(model = meta_gamm) + plot_post_slope(model = meta_gamm)
+# model <- full_meta_gamm
+# plot_post_preds(model = full_meta_gamm) + plot_post_slope(model = full_meta_gamm)
+# model <- meta_gam
+# model <- gamm
 
@@
         add_epred_draws(object = model) %>%
+        # add_epred_draws(object = model, ndraws = 1e2) %>%
         data.frame() %>%
@@
         pivot_wider(names_from = condition, values_from = .epred) %>%
+        # pivot_wider(id_cols = c(participant, time), names_from = condition, values_from = .epred) %>%
         mutate(epred_diff = cond2 - cond1) %>%
         group_by(participant, time) %>%
-        summarise(m = mean(epred_diff > 0 + sesoi), .groups = "drop") %>%
+        # summarise(m = mean(epred_diff > 0 + sesoi), .groups = "drop") %>%
+        summarise(m = mean(epred_diff > 0 + sesoi) ) %>%
         mutate(
@@
             above_thres = ifelse(prob_ratio >= threshold, 1, NA)
-            )
+            ) %>%
+        ungroup()
+
+    # some tests
+    # model$data %>%
+    #     add_epred_draws(object = model, ndraws = 1e2) %>%
+    #     data.frame() %>%
+    #     dplyr::select(participant, time, condition, .epred, .draw) %>%
+    #     filter(participant == "participant_01" & time == 0 & .draw == 1)
+    # 
+    # prob_y_above_0_indiv %>%
+    #     filter(participant == "participant_02" & time == 0 & .draw == 2)
+    # 
+    # prob_y_above_0_indiv %>%
+    #     pivot_wider(names_from = condition, values_from = .epred) %>%
+    #     filter(participant == "participant_02" & time == 0 & .draw == 2)
 
@@
 }
+
+# plotting the predictions
+# plot_post_test_ratio_individual(model = full_meta_gamm)
+# plot_post_test_ratio_individual(model = meta_gam)
+plot_post_test_ratio_individual(model = gamm)
+
+# plotting the predictions at the participant level
+plot_post_preds_ppt <- function (model, data = NULL, true_onset = NULL, true_offset = NULL) {
+
+    if (is.null(data) ) data <- model$data
+
+    # retrieving predictions at the individual level
+    # cond_effects <- conditional_effects(
+    #     # defining the model
+    #     x = model,
+    #     # defining the interaction of interest
+    #     effects = "time:condition",
+    #     # defining the "conditions" (participant) on which computing predictions
+    #     conditions = make_conditions(x = model$data, vars = "participant"),
+    #     # getting predictions
+    #     method = "posterior_epred",
+    #     # including all random/varying effects
+    #     re_formula = NULL
+    #     )[[1]]
+
+    # retrieving posterior draws and computing the difference per participant & time
+    post_preds <- model$data %>%
+        add_epred_draws(object = model, re_formula = NULL, ndraws = 500) %>%
+        data.frame() %>%
+        dplyr::select(participant, time, condition, .epred, .draw) %>%
+        group_by(participant, time, condition) %>%
+        summarise(
+            .epred_mean = mean(.epred),
+            .epred_lower = quantile(.epred, probs = 0.025),
+            .epred_upper = quantile(.epred, probs = 0.975),
+            .groups = "drop"
+            ) %>%
+        ungroup()
+
+    # retrieving observed means (optional)
+    obs_means <- data %>%
+        group_by(participant, time, condition) %>%
+        summarise(eeg_mean = mean(eeg_mean), .groups = "drop")
+
+    # plotting
+    post_preds %>%
+        ggplot(aes(x = time, y = .epred_mean, colour = condition, fill = condition) ) +
+        geom_vline(xintercept = true_onset, linetype = 2) +
+        geom_vline(xintercept = true_offset, linetype = 2) +
+        geom_hline(yintercept = 0, linetype = 2) +
+        geom_point(
+            data = obs_means, aes(y = eeg_mean),
+            pch = 21, size = 1, alpha = 0.2,
+            show.legend = FALSE
+            ) +
+        geom_line() +
+        geom_ribbon(aes(ymin = .epred_lower, ymax = .epred_upper, colour = NULL), alpha = 0.2) +
+        facet_wrap(~participant, scales = "free_y") +
+        scale_colour_met_d(name = "Johnson") +
+        scale_fill_met_d(name = "Johnson") +
+        labs(x = "Time (s)", y = "EEG signal (a.u.)", title = "Posterior predictions by participant")
+
+}
+
+plot_post_preds_ppt(model = meta_gam)
 ```
@@
 
-Note: Given the focus on the precision of onset/offset estimates, is it correct/relevant to mention that we are interested here in quantifying type-S and type-M (rather than type-1 or type-2) errors?
+Note: Given the focus on the precision of onset/offset estimates, is it correct/relevant/sensible to mention that we are interested here in quantifying type-S and type-M (rather than type-1 or type-2) errors?
 
@@
 
-We then assess the performance of the proposed approach by computing the difference between the true and estimated onset/offset of the EEG difference according to various `k` (BGAM basis dimension) and `threshold` values.
+We then assessed the performance of the proposed approach by computing the difference between the true and estimated onset/offset of the EEG difference according to various `k` (basis dimension) and `threshold` values.
 Remember that the EEG signal was generated from a truncated Gaussian with an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
@@
 # loading the simulation results
-# results <- readRDS(file = "results/meta_gam_error_properties.rds")
-# results <- readRDS(file = "results/meta_gam_error_properties_100sims.rds")
 results <- readRDS(file = "results/meta_gam_error_properties_cluster.rds")
@@
 #     pivot_longer(cols = error_onset:error_offset) %>%
-#     mutate(name = factor(x = name, levels = c("error_onset", "error_offset") ) ) %>%
+#     mutate(
+#         name = factor(
+#             x = name,
+#             levels = c("error_onset", "error_offset"),
+#             labels = c("onset", "offset")
+#             )
+#         ) %>%
 #     ggplot(
@@
 #     scale_y_discrete(expand = c(0, 0), breaks = c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50) ) +
-#     scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
+#     scale_fill_gradientn(
+#         colors = rev(met.brewer("Hokusai1") ),
+#         transform = "sqrt"
+#         ) +
 #     labs(
@@
 
-## Comparing the onsets/offsets estimates from other approaches
-
-We then compared the ability of the BGAMM to accurately estimate the onset and offset of the ERP difference to other widely-used methods.
+## Comparing the onsets/offsets estimates across approaches
+
+We then compared the ability of the BGAM to accurately estimate the onset and offset of the ERP difference to other widely-used methods.
 First, we conducted mass-univariate t-tests (thus treating each timestep independently) and identified the onset and offset of the ERP difference as the first and last values crossing an arbitrary significance threshold ($\alpha = 0.05$).
@@
 
-To assess the accuracy of group-level onset estimation, the various methods were compared using the bias (i.e., median(estimated-true)), median absolute error (MAE), root mean square error (RMSE), variance, and median absolute deviation (MAD) of onset/offset estimates computed on 1,000 simulated datasets.
+To assess the accuracy of group-level onset estimation, the various methods were compared using the bias (defined as the median difference between the estimated and true value of the onset/offset), median absolute error (MAE), root mean square error (RMSE), variance, and median absolute deviation (MAD) of onset/offset estimates computed on 10,000 simulated datasets.
 As in @rousselet_using_2025, each participant was assigned a random onset between 150 and 170ms.
-Whereas the present article focuses on one-dimensional signas (e.g., one M/EEG channel), we provice a 2D application in @sec-2D.
+Whereas the focus of the present article focuses is on one-dimensional signals (e.g., one M/EEG channel), we provide a 2D application in @sec-2D.
 
@@
 
-To complement the simulation study, we evaluated the performance of the various methods on actual MEG data [decoding results from @nalborczyk:inprep].
+To complement the simulation study, we evaluated the performance of all methods on actual MEG data [decoding results from @nalborczyk:inprep].
 In this study, we conducted time-resolved multivariate pattern analysis (MVPA, also known as decoding) of MEG data during reading tasks.
@@
 
+# computing the smallest effect size of interest (SESOI)
+# as the 95% confidence interval during the baseline
+# sesoi <- decoding_summary_df %>%
+#     filter(time < 0) %>%
+#     summarise(auc_mean = mean(auc_mean), .by = time) %>%
+#     summarise(sesoi = 1.96 * (sd(auc_mean) / sqrt(n() ) ) ) %>%
+#     pull(sesoi)
+
 # fitting the Beta GAM
@@
     auc_mean ~ s(time, bs = "cr", k = 50),
+    # auc_mean ~ 1 + baseline + s(time, bs = "tp", k = 30, by = baseline),
     data = decoding_summary_df,
     family = Beta(),
-    warmup = 2000,
-    iter = 5000,
+    warmup = 1000,
+    iter = 4000,
     chains = 4,
@@
     file = "models/meg_decoding_gam.rds"
+    # file = "models/meg_decoding_gam_baseline.rds"
     )
@@
     # ensuring there is no 0 or +Inf values
-    mutate(prob_ratio = ifelse(is.infinite(prob_ratio), ndraws(meg_decoding_gam), prob_ratio) ) %>%
-    mutate(prob_ratio = ifelse(prob_ratio == 0, 1 / ndraws(meg_decoding_gam), prob_ratio) ) %>%
+    mutate(prob_ratio = pmin(prob_ratio, ndraws(meg_decoding_gam) ) ) %>%
+    mutate(prob_ratio = pmax(prob_ratio, 1 / ndraws(meg_decoding_gam) ) ) %>%
     ungroup()
@@
         filter(above_threshold) %>%
-        mutate(cluster_id = cumsum(cluster_change)) %>%
+        mutate(cluster_id = cumsum(cluster_change) ) %>%
         group_by(cluster_id) %>%
-        summarise(cluster_onset = first(time), cluster_offset = last(time), .groups = "drop") %>%
+        summarise(
+            cluster_onset = first(time), cluster_offset = last(time),
+            .groups = "drop"
+            ) %>%
         select(cluster_onset, cluster_offset)
@@
     conditional_effects(x = meg_decoding_gam),
-    # line_args = list(colour = "steelblue", fill = "steelblue", alpha = 0.3),
     line_args = list(colour = brms_color, fill = brms_color, alpha = 0.3),
@@
         fill ="orangered",
-        # fill ="darkred",
         alpha = 0.2
@@
             ),
-        # colour = "steelblue",
         colour = brms_color,
@@
 
+```{r fig-post-prob-ratio-decoding, eval = FALSE, echo = FALSE, out.width = "100%", fig.width = 9, fig.asp = 0.5, fig.cap = "Left: Posterior probability of the EEG difference (slope) being above 0 according to the BGAMM.
+Right: Ratio of posterior probability according to the BGAMM (on a log10 scale).
+Timesteps above threshold (10) are highlighted in green.
+NB: the minimum and maximum possible ratio values are determined (bounded) by the number of posterior samples in the model."}
+# Computing the EDF?
+# From https://discourse.mc-stan.org/t/standard-operating-procedures-for-anova/3964/2
+# If Kruschke has code that does what you want, then you can probably call his code on some columns of as.matrix(fit).
+If it really has to be in coda format, there is an As.mcmc.list function in the rstan package that will convert a stanfit object to a mcmc.list, but you probably have to call it on the $stanfit element of the brmsfit object.
+# See the bamlss::samplestat() function at https://github.com/cran/bamlss/blob/0a2a4154e0c0099ba8ac0b372ae7460e57972741/R/BAMLSS.R#L2397
+# Maybe export brms posterior samples to mcmc.list and then use this function?
+# See https://mc-stan.org/rstan/reference/As.mcmc.list.html
+library(bamlss)
+# mcmc_samples <- as.mcmc(meg_decoding_gam)
+mcmc_samples <- as.mcmc(meg_decoding_gam$fit)
+mcmc_samples_stats <- samplestats(
+    samples = mcmc_samples,
+    x = bamlss.frame(formula = auc_mean ~ s(time, bs = "tp", k = 50), data = decoding_summary_df)$x,
+    # y = bamlss.frame(formula = auc_mean ~ s(time, bs = "tp", k = 50), data = decoding_summary_df)$y,
+    # family = beta_bamlss,
+    # logLik = FALSE
+    )
+mcmc_samples_stats
+
+# or using bamlss directly
+b_model <- bamlss(
+    formula = auc_mean ~ s(time, bs = "tp", k = 50),
+    family = beta_bamlss,
+    data = decoding_summary_df
+    )
+
+# computing the EDF (around 44.6)
+samplestats(samples = b_model)
+
+# comparing to p_WAIC (around 48.4)
+waic(meg_decoding_gam)
+
+# computing the posterior probability ratio (PPR)
+model <- meg_decoding_gam
+threshold <- 10
+
+# defining a function to compute and plot the posterior odds
+plot_posterior_odds <- function (model, threshold = 10) {
+
+    # retrieving the number of posterior samples
+    n_posterior_samples <- ndraws(model)
+
+    # defiing the full time sequence
+    # time_seq <- seq(min(model$data$time), max(model$data$time), length.out = 300)
+    time_seg <- unique(model$data$time)
+
+    # generating new data
+    newdata <- data.frame(time = time_seq)
+
+    # getting posterior expected values (marginal)
+    posterior_preds <- posterior_epred(object = model, newdata = newdata, re_formula = NA)
+
+    # computing posterior mean and 95% CI at each time
+    posterior_preds_df <- as.data.frame(posterior_preds) %>%
+        mutate(draw = row_number() ) %>%
+        pivot_longer(-draw, names_to = "time_index", values_to = "auc") %>%
+        mutate(time = rep(time_seq, each = nrow(posterior_preds) ) ) %>%
+        group_by(time) %>%
+        summarise(
+            auc_mean = mean(auc),
+            auc_lower = quantile(auc, 0.05),
+            auc_upper = quantile(auc, 0.95)
+            ) %>%
+        ungroup()
+
+    # computing the average upper 96% credible interval during baseline (i.e., before 0s)
+    upper_baseline_ci <- posterior_preds_df %>%
+        filter(time < 0) %>%
+        summarise(upper_baseline_ci = mean(auc_upper) ) %>%
+        pull(upper_baseline_ci)
+
+    # computing the posterior probability of AUC > upper CrI of baseline
+    prob_y_above_baseline <- model$data %>%
+        add_epred_draws(object = model, ndraws = 100) %>%
+        data.frame() %>%
+        # head()
+        # dplyr::select(participant, time, condition, .epred, .draw) %>%
+        # pivot_wider(names_from = condition, values_from = .epred) %>%
+        # mutate(epred_diff = cond2 - cond1) %>%
+        # dplyr::select(time, baseline, .epred, .draw) %>%
+        # pivot_wider(names_from = baseline, values_from = .epred) %>%
+        # head()
+        # mutate(epred_diff = `c - cond1) %>%
+        # posteriors_preds %>%
+        # computing the mean posterior probability at the group level
+        group_by(time) %>%
+        summarise(m = mean(.epred > upper_baseline_ci) ) %>%
+        mutate(prob_ratio = m / (1 - m) ) %>%
+        ungroup() %>%
+        # bounding posterior probabilities by the number of posterior samples
+        mutate(prob_ratio = pmin(prob_ratio, n_posterior_samples) ) %>%
+        mutate(prob_ratio = pmax(prob_ratio, 1 / n_posterior_samples) ) %>%
+        ungroup()
+
+    # plotting it
+    post_plot <- prob_y_above_baseline %>%
+        ggplot(aes(x = time, y = m) ) +
+        geom_hline(yintercept = 0.5, linetype = 2) +
+        # geom_vline(xintercept = true_onset, linetype = 2) +
+        # geom_vline(xintercept = true_offset, linetype = 2) +
+        geom_line() +
+        ylim(c(0, 1) ) +
+        labs(x = "Time (s)", y = expression(Pr(beta>0~"|"~data, model) ) )
+
+    # plotting it
+    ratio_plot <- prob_y_above_baseline %>%
+        mutate(above_thres = ifelse(prob_ratio >= threshold, 1, NA) ) %>%
+        ggplot(aes(x = time, y = prob_ratio) ) +
+        geom_hline(yintercept = 1, linetype = "dashed") +
+        geom_hline(yintercept = 10, linetype = "dashed", color = "darkred") +
+        geom_hline(yintercept = 1/10, linetype = "dashed", color = "darkred") +
+        geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
+        geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
+        # geom_vline(xintercept = true_onset, linetype = 2) +
+        # geom_vline(xintercept = true_offset, linetype = 2) +
+        geom_line(linewidth = 0.8) +
+        geom_point(
+            data = . %>% dplyr::filter(!is.na(above_thres) ),
+            aes(y = threshold),
+            colour = "darkgreen",
+            shape = 16,
+            size = 2,
+            na.rm = TRUE
+            ) +
+        scale_y_log10(
+            labels = label_log(digits = 2),
+            limits = c(1 / n_posterior_samples, n_posterior_samples),
+            breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
+            ) +
+        labs(
+            x = "Time (s)",
+            y = expression(Pr(beta>0~"|"~data) / (1 - Pr(beta>0~"|"~data) ) )
+            )
+
+    # plotting both the posterior probability and the ratio...
+    post_plot + ratio_plot
+
+}
+
+# plotting the posterior odds
+plot_posterior_odds(model = meg_decoding_gam, threshold = 10)
+
+####################################
+# or testing an alternative method #
+####################################
+
+# defining the full time sequence
+time_seq <- seq(min(decoding_summary_df$time), max(decoding_summary_df$time), length.out = 300)
+
+# generating new data
+newdata <- data.frame(time = time_seq)
+
+# retrieving the posterior expected values (marginal)
+posterior <- posterior_epred(object = meg_decoding_gam, newdata = newdata, re_formula = NA)
+
+# computing posterior mean and 95% CI at each time
+pred_df <- as.data.frame(posterior) %>%
+    mutate(draw = row_number() ) %>%
+    pivot_longer(-draw, names_to = "time_index", values_to = "auc") %>%
+    mutate(time = rep(time_seq, each = nrow(posterior) ) ) %>%
+    group_by(time) %>%
+    summarise(
+        auc_mean = mean(auc),
+        auc_lower = quantile(auc, 0.025),
+        auc_upper = quantile(auc, 0.975)
+        )
+
+# subsetting newdata to time < 0
+baseline_indices <- which(time_seq < 0)
+baseline_draws <- posterior[, baseline_indices]
+
+# computing the average baseline AUC per draw
+baseline_per_draw <- rowMeans(baseline_draws)
+
+# for timepoints ≥ 0
+post_indices <- which(time_seq >= 0)
+time_post <- time_seq[post_indices]
+
+# subsetting post-stimulus predictions
+posterior_post <- posterior[, post_indices]
+
+# computing posterior probability that AUC > pre-stim baseline
+post_prob <- sapply(1:length(post_indices), function (i) {
+
+    mean(posterior_post[, i] > baseline_per_draw)
+
+})
+
+# assembling into dataframe
+post_df <- data.frame(
+    time = time_post,
+    prob_auc_gt_baseline = post_prob
+    )
+
+# plotting the results
+n_posterior_samples <- ndraws(meg_decoding_gam)
+post_df %>%
+    ggplot(aes(x = time, y = prob_auc_gt_baseline / (1-prob_auc_gt_baseline) ) ) +
+    geom_line(size = 1, color = "darkgreen") +
+    geom_hline(yintercept = 0.95, linetype = "dashed", color = "gray") +
+    scale_y_log10(
+        labels = label_log(digits = 2),
+        limits = c(1 / n_posterior_samples, n_posterior_samples),
+        breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
+        ) +
+    labs(
+        title = "Posterior Probability AUC > Pre-Stimulus Baseline",
+        x = "Time (s)", y = "P(AUC > baseline)"
+        ) +
+    theme_bw()
+
+# or using yet another approach
+estimate_edf_brms <- function (
+    model, 
+    term,           # e.g., "s(time)"
+    time_seq = NULL,
+    n = 200,        # number of time points
+    re_formula = NA # exclude group-level effects
+    ) {
+
+  # Extract time variable name from term, e.g., "s(time)" → "time"
+  time_var <- gsub("s\\(([^,]+).*\\)", "\\1", term)
+
+  # Create prediction grid
+  if (is.null(time_seq)) {
+    time_range <- range(model$data[[time_var]], na.rm = TRUE)
+    time_seq <- seq(time_range[1], time_range[2], length.out = n)
+  }
+
+  # newdata <- data.frame(!!time_var := time_seq)
+  newdata <- data.frame(time_seq)
+  names(newdata) <- time_var
+
+  # Extract posterior smooth draws only (excludes intercept & fixed effects)
+  smooth_draws <- posterior_smooths(
+    model,
+    newdata = newdata,
+    smooth = term,
+    re_formula = re_formula,
+    resp = NULL
+  )
+
+  # Compute posterior mean at each timepoint
+  posterior_mean <- colMeans(smooth_draws)
+
+  # Compute variance of the posterior mean across time
+  var_mean <- var(posterior_mean)
+
+  # Compute variance at each timepoint across draws
+  pointwise_var <- apply(smooth_draws, 2, var)
+
+  # Mean pointwise variance
+  mean_pointwise_var <- mean(pointwise_var)
+
+  # EDF estimate
+  edf_estimate <- var_mean / mean_pointwise_var
+
+  return(list(
+    edf = edf_estimate,
+    var_mean = var_mean,
+    mean_pointwise_var = mean_pointwise_var,
+    time = time_seq,
+    smooth_mean = posterior_mean,
+    smooth_sd = sqrt(pointwise_var)
+  ))
+}
+
+edf_result <- estimate_edf_brms(
+  model = meg_decoding_gam,
+  #term = "s(time)"
+  term = 's(time,bs="tp",k=50)'
+)
+
+print(paste("Estimated EDF:", round(edf_result$edf, 2)))
+
+plot_df <- data.frame(
+  time = edf_result$time,
+  mean = edf_result$smooth_mean,
+  sd = edf_result$smooth_sd
+)
+
+ggplot(plot_df, aes(x = time, y = mean)) +
+  geom_line(color = "blue") +
+  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd), alpha = 0.2) +
+  labs(
+    title = paste("Smooth with Posterior SD | Estimated EDF ≈", round(edf_result$edf, 2)),
+    x = "Time", y = "Smooth Estimate"
+  ) +
+  theme_bw()
+```
+
 \newpage
@@
 
-@fig-simulation-results shows a summary of the simulation results, revealing that the proposed approach (`BGAM`) has the lowest median absolute error (MAE) and variance for both the onset and offset estimates.
+@fig-simulation-results shows a summary of the simulation results, revealing that the proposed approach (`BGAM`) has the lowest median error for both the onset and offset estimates.
 The `Cluster mass` and `Change point` also have good performance, but surprisingly, the `TFCE` method has relatively bad performance for estimating the effect offset (similar performance to the `Holm` and `FDR BY01` methods).
-Unsurprisingly, the `Raw p-value` and `FDR BH95` methods show the worst performance.
-
-```{r fig-simulation-results, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median absolute error and variance of onset and offset estimates for each method.
-Variance is plotted on a log10 scale for visual purposes."}
-# loading the simulation results
-sim_results <- readRDS(file = "results/errors_results_100sims.rds")
+Unsurprisingly, the `FDR BH95` and `Raw p-value` methods show the worst performance.
+
+```{r fig-simulation-results, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median error and median absolute deviation of the error for the onset (left) and offset (right) estimates according to each method.
+Methods are ordered from lowest (top) to highest (bottom) median absolute error (separately for the onset and offset estimates)."}
+# loading all RDS files
+files <- list.files(
+    path = "results/errors_cluster_results", 
+    pattern = ".rds", 
+    full.names = TRUE
+    )
+
+# sort files numerically by the number in the filename
+files <- files[order(readr::parse_number(files) )]
+
+# read and bind all files, adding a sim_id
+sim_results <- map2_dfr(
+    .x = files,
+    .y = seq_along(files),
+    .f = ~ readRDS(.x)
+    # .f = ~ readRDS(.x) |> mutate(sim_id = .y)
+    # .f = function (file, index) {
+    #     offset <- (index - 1) * splits_per_file
+    #     readRDS(file) %>% mutate(split_id = as.numeric(split_id) ) %>%
+    #         mutate(split_id = split_id + offset) %>%
+    #         mutate(split_id = formatC(x = split_id, width = 3, flag = 0) ) 
+    #     }
+    )
+
+# sanity check
+# n_distinct(sim_results$simulation_id)
 
@@
 #     summarise(
+#         bias_onset = median(onset - true_onset),
 #         MAE_onset = median(abs(onset - true_onset) ),
 #         variance_onset = var(onset),
+#         bias_offset = median(offset - true_offset),
 #         MAE_offset = median(abs(offset - true_offset) ),
@@
 #         ) %>%
-#     arrange(MAE_onset, MAE_offset)
+#     # arrange(MAE_onset, MAE_offset)
+#     arrange(bias_onset, bias_offset)
 
@@
 #     geom_density(aes(fill = NULL) ) +
+#     # geom_histogram(breaks = "FD") +
 #     facet_grid(plot_row~onset_offset, scales = "free_x") +
@@
 # variance := var(sim_onset_offset_distrib - true_onset_offset)
+# sim_results %>%
+#     # keeping only threshold = 20 for BGAM and threshold = 0.05 for frequentist results
+#     filter(threshold %in% c(0.05, 20) ) %>%
+#     mutate(
+#         error = case_when(
+#             # grepl("onset", onset_offset) ~ abs(value - true_onset),
+#             # grepl("offset", onset_offset) ~ abs(value - true_offset)
+#             grepl("onset", onset_offset) ~ value - true_onset,
+#             grepl("offset", onset_offset) ~ value - true_offset
+#             )
+#         ) %>%
+#     group_by(method, onset_offset) %>%
+#     summarise(
+#         median_error = median(error, na.rm = TRUE),
+#         median_abs_error = median(abs(error), na.rm = TRUE),
+#         # variance = var(error, na.rm = TRUE),
+#         variance = var(value, na.rm = TRUE),
+#         mad = mad(value, na.rm = TRUE),
+#         .groups = "drop"
+#         ) %>%
+#     ungroup() %>%
+#     mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
+#     mutate(
+#         method = factor(
+#             x = method,
+#             levels = c(
+#                 "raw_p", "pval_bh", "pval_by", "pval_holm",
+#                 "cluster_mass", "cluster_tfce", "cpt", "brms"
+#                 ),
+#             labels = c(
+#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+#                 "TFCE", "Change point", "BGAM (new method)"
+#                 ),
+#             )
+#         ) %>%
+#     ggplot(
+#         aes(
+#             x = median_error,
+#             # y = mad,
+#             # y = median_abs_error,
+#             # y = median_abs_error,
+#             # x = median_abs_error,
+#             y = variance,
+#             colour = method,
+#             fill = method
+#             )
+#         ) +
+#     geom_vline(xintercept = 0, linetype = 2) +
+#     # geom_errorbar(
+#     #     aes(
+#     #         xmin = median_error - mad, xmax = median_error + mad,
+#     #         #ymin = median_abs_error - mad, ymax = median_abs_error + mad
+#     #         )
+#     #     ) +
+#     #     geom_errorbar(
+#     #     aes(
+#     #         #xmin = median_error - mad, xmax = median_error + mad,
+#     #         ymin = median_abs_error - mad, ymax = median_abs_error + mad
+#     #         )
+#     #     ) +
+#     geom_point(
+#         size = 2,
+#         pch = 21,
+#         colour = "white",
+#         show.legend = FALSE
+#         ) +
+#     geom_label_repel(
+#         aes(label = method),
+#         colour = "white",
+#         size = 3,
+#         segment.color = NA,
+#         show.legend = FALSE
+#         ) +
+#     # scale_y_log10() +
+#     facet_wrap(~onset_offset) +
+#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+#     # labs(x = "Median error (s)", y = "Variance (log-scale)")
+#     # labs(x = "Median error (s)", y = "Median absolute error (log-scale)")
+#     labs(x = "Median absolute error (s)", y = "Variance (log-scale)")
+
+# or plotting only the error in descending order
 sim_results %>%
-    # keeping only threshold = 20 for BGAM and threshold = 0.05 for frequentist results
+    # keeping only threshold = 10 for BGAM and threshold = 0.05 for frequentist results
     filter(threshold %in% c(0.05, 20) ) %>%
+    # filter(threshold %in% c(0.05, 10) ) %>%
     mutate(
@@
         ) %>%
+    # converting to ms
+    # mutate(error = error * 1e3) %>%
     group_by(method, onset_offset) %>%
     summarise(
+        mean_error = mean(error, na.rm = TRUE),
         median_error = median(error, na.rm = TRUE),
         median_abs_error = median(abs(error), na.rm = TRUE),
-        # variance = var(error, na.rm = TRUE),
-        variance = var(value, na.rm = TRUE),
+        variance = var(error, na.rm = TRUE),
+        # variance = var(value, na.rm = TRUE),
+        sem = sd(value, na.rm = TRUE) / sqrt(n() ),
         mad = mad(value, na.rm = TRUE),
@@
         ) %>%
+    mutate(
+        method_reordered = tidytext::reorder_within(
+            x = method,
+            # by = 1 / error,
+            # fun = median,
+            # by = 1 / abs(median_error),
+            by = 1 / median_abs_error,
+            # by = 1 / abs(mean_error),
+            within = onset_offset
+            )
+        ) %>%
+    # mutate(
+    #     method_reordered = fct_reorder(
+    #         method, error, .fun = ~abs(median(.x) ), .na_rm = TRUE
+    #         )
+    #     ) %>%
+    # arrange(desc(error) ) %>%
+    # arrange(error) %>%
+    # head()
     ggplot(
         aes(
-            # x = median_error,
-            # y = median_abs_error,
-            x = median_abs_error,
-            y = variance,
+            x = median_error,
+            # x = error,
+            xmin = median_error - mad,
+            xmax = median_error + mad,
+            # x = mean_error,
+            # xmin = mean_error - variance,
+            # xmax = mean_error + variance,
+            y = method_reordered,
             colour = method,
@@
         ) +
-    # geom_vline(xintercept = 0, linetype = 2) +
-    geom_point(
-        size = 2,
-        pch = 21,
-        colour = "white",
+    geom_vline(xintercept = 0, linetype = 2) +
+    geom_pointrange(
+        size = 0.25,
+        linewidth = 0.75,
         show.legend = FALSE
         ) +
-    geom_label_repel(
-        aes(label = method),
-        colour = "white",
-        size = 3,
-        segment.color = NA,
-        show.legend = FALSE
-        ) +
-    scale_y_log10() +
-    facet_wrap(~onset_offset) +
+    # stat_histinterval(
+    #     # breaks = "FD",
+    #     # n = 1e3,
+    #     # point_interval = "median_qi",
+    #     # .width = c(0.5, 0.95),
+    #     .width = 0.1,
+    #     # colour = "black",
+    #     # slab_colour = "white",
+    #     # slab_linewidth = 0,
+    #     slab_alpha = 0.8,
+    #     # fill = "steelblue"
+    #     show.legend = FALSE
+    #     ) +
+    facet_wrap(~onset_offset, scales = "free") +
+    scale_y_reordered() +
     scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
     scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    # labs(x = "Median error (s)", y = "Variance (log-scale)")
-    # labs(x = "Median error (s)", y = "Median absolute error (log-scale)")
-    labs(x = "Median absolute error (s)", y = "Variance (log-scale)")
+    labs(
+        x = "Estimation error (s)",
+        # x = "Mean error (s)",
+        y = "Method"
+        )
 
 # saving the plot
-ggsave(
-    # filename = "figures/simulation_results_median_error_variance.png",
-    # filename = "figures/simulation_results_bias_mae.png",
-    filename = "figures/simulation_results_mae_variance.png",
-    width = 12, height = 6, dpi = 300,
-    device = "png"
-    )
+# ggsave(
+#     # filename = "figures/simulation_results_median_error_variance.png",
+#     # filename = "figures/simulation_results_bias_mae.png",
+#     filename = "figures/simulation_results_mae_variance.png",
+#     width = 12, height = 6, dpi = 300,
+#     device = "png"
+#     )
+
+# plotting the distribution of results
+# sim_results %>%
+#     # keeping only threshold = 10 for BGAM and threshold = 0.05 for frequentist results
+#     filter(threshold %in% c(0.05, 20) ) %>%
+#     mutate(
+#         error = case_when(
+#             # grepl("onset", onset_offset) ~ abs(value - true_onset),
+#             # grepl("offset", onset_offset) ~ abs(value - true_offset)
+#             grepl("onset", onset_offset) ~ value - true_onset,
+#             grepl("offset", onset_offset) ~ value - true_offset
+#             )
+#         ) %>%
+#     # converting to ms
+#     # mutate(error = error * 1e3) %>%
+#     # group_by(method, onset_offset) %>%
+#     # summarise(
+#     #     mean_error = mean(error, na.rm = TRUE),
+#     #     median_error = median(error, na.rm = TRUE),
+#     #     median_abs_error = median(abs(error), na.rm = TRUE),
+#     #     variance = var(error, na.rm = TRUE),
+#     #     # variance = var(value, na.rm = TRUE),
+#     #     sem = sd(value, na.rm = TRUE) / sqrt(n() ),
+#     #     mad = mad(value, na.rm = TRUE),
+#     #     .groups = "drop"
+#     #     ) %>%
+#     # ungroup() %>%
+#     mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
+#     mutate(
+#         method = factor(
+#             x = method,
+#             levels = c(
+#                 "raw_p", "pval_bh", "pval_by", "pval_holm",
+#                 "cluster_mass", "cluster_tfce", "cpt", "brms"
+#                 ),
+#             labels = c(
+#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+#                 "TFCE", "Change point", "BGAM (new method)"
+#                 ),
+#             )
+#         ) %>%
+#     mutate(
+#         method_reordered = tidytext::reorder_within(
+#             x = method,
+#             by = abs(error),
+#             fun = median,
+#             # by = 1 / abs(median_error),
+#             # by = 1 / abs(mean_error),
+#             within = onset_offset
+#             )
+#         ) %>%
+#     # mutate(
+#     #     method_reordered = fct_reorder(
+#     #         method, error, .fun = ~abs(median(.x) ), .na_rm = TRUE
+#     #         )
+#     #     ) %>%
+#     # arrange(desc(error) ) %>%
+#     # arrange(error) %>%
+#     # head()
+#     ggplot(
+#         aes(
+#             # x = median_error,
+#             x = error,
+#             # xmin = median_error - mad,
+#             # xmax = median_error + mad,
+#             # x = mean_error,
+#             # xmin = mean_error - variance,
+#             # xmax = mean_error + variance,
+#             y = method_reordered,
+#             colour = method,
+#             fill = method
+#             )
+#         ) +
+#     geom_vline(xintercept = 0, linetype = 2) +
+#     # geom_pointrange(
+#     #     size = 0.25,
+#     #     linewidth = 0.75,
+#     #     show.legend = FALSE
+#     #     ) +
+#     stat_slabinterval(
+#         n = 200,
+#         normalize = "groups",
+#         point_interval = "median_qi",
+#         # .width = c(0.1, 0.9),
+#         # .width = 0.2,
+#         .width = 0,
+#         interval_size_range = c(1, 1),
+#         fatten_point = 1.25,
+#         # slab_colour = "white",
+#         # slab_linewidth = 0,
+#         # density = "unbounded",
+#         slab_alpha = 0.5,
+#         show.legend = FALSE
+#         ) +
+#     facet_wrap(~onset_offset, scales = "free") +
+#     scale_y_reordered() +
+#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+#     labs(
+#         x = "Estimation error (s)",
+#         y = "Method"
+#         )
 ```
 
-Results are further summarised in @tbl-simulation-results, which shows that the `BGAM` is perfectly unbiased (i.e., it has a bias of 0s) for the onset and almost exactly unbiased for the onset (with a bias of approximately -2ms).
-The `Bias` column shows that all other methods tend to estimate the onset later than the true onset and to estimate the offet earlier than the true offset.
+Results are further summarised in @tbl-simulation-results, which shows that the `BGAM` is almost perfectly unbiased (i.e., it has a bias of approximately 2ms for the onset and 4ms for the offset).
+The `Bias` column shows that all methods tend to estimate the onset later than the true onset and to estimate the offset earlier than the true offset.
 As can be seen from this table, the `BGAM` has the best performance on all included metrics.
@@
 ```{r tbl-simulation-results, eval = TRUE, echo = FALSE, results = "asis"}
-#| tbl-cap: "Summary statistics of the onset and offset estimates for each method (ordered by the absolute value of the bias)."
-
-# loading the simulation results
-sim_results <- readRDS(file = "results/errors_results_100sims.rds")
-
+#| tbl-cap: "Summary statistics of the onset and offset estimates for each method (in ms, ordered by the MAE) according to the main simulation study."
 # summarising the results
@@
     filter(threshold %in% c(0.05, 20) ) %>%
+    # filter(threshold %in% c(0.05, 10) ) %>%
     mutate(
         error = case_when(
-            grepl("onset", onset_offset) ~ value - true_onset,
-            grepl("offset", onset_offset) ~ value - true_offset
-            )
-        ) %>%
+            # grepl("onset", onset_offset) ~ value - true_onset,
+            # grepl("offset", onset_offset) ~ value - true_offset
+            grepl("onset", onset_offset) ~ value * 1e3 - true_onset * 1e3,
+            grepl("offset", onset_offset) ~ value * 1e3 - true_offset * 1e3
+            )
+        ) %>%
+    # converting to ms
+    # mutate(error = error * 1e3) %>%
     group_by(method, onset_offset) %>%
     summarise(
+        # mean_error = mean(error, na.rm = TRUE),
         median_error = median(error, na.rm = TRUE),
+        # mean_abs_error = mean(abs(error), na.rm = TRUE),
         median_abs_error = median(abs(error), na.rm = TRUE),
@@
         variance = var(error, na.rm = TRUE),
+        # variance = var(value, na.rm = TRUE),
         mad = mad(error, na.rm = TRUE),
@@
         ) %>%
-    arrange(abs(median_error) ) %>%
+    # arrange(abs(median_error) ) %>%
+    # arrange(abs(mean_error) ) %>%
+    arrange(median_abs_error) %>%
     group_by(onset_offset) %>%
@@
         columns = everything(),
-        decimals = 4
+        decimals = 2
         ) %>%
@@
         method = "",
+        # mean_error = "Bias_mean",
         median_error = "Bias",
         median_abs_error = "MAE",
+        # mean_abs_error = "MAE",
         rmse = "RMSE",
@@
 The `Change point` and `Cluster mass` methods seem the most conservative methods, identifying a time window from approximately +60ms to +500ms.
-The `Holm`, `TFCE`, and `BGAM` methods produce similar estimates of onset and offset, ranging from approximately +60ms to +650ms, although the `BGAM` method seems to result in fewer clusters.[^1]
-
-[^1]: It should be noted that although each method can produce several "clusters" of timesteps, we only considered the first (onset) and last (offset) timesteps identified by each method to compute the estimation error.
+The `Holm`, `TFCE`, and `BGAM` methods produce similar estimates of onset and offset, ranging from approximately +60ms to +650ms, although the `BGAM` method seems to result in fewer clusters.[^2]
+
+[^2]: It should be noted that although each method can produce several "clusters" of timesteps, we only considered the first (onset) and last (offset) timesteps identified by each method to compute the estimation error.
 
 ```{r fig-onset-offset, echo = FALSE, fig.width = 8, fig.asp = 0.6, fig.cap = "Group-level average decoding performance through time with onset and offset estimates for each method [data from @nalborczyk:inprep]."}
-# loading the results
-# reliability_results <- readRDS(file = "results/reliability_results.rds")
-# reliability_results <- readRDS(file = "results/reliability_results_cluster.rds")
-reliability_results <- readRDS(file = "results/reliability_results_100splits.rds")
+# importing results
+# reliability_results <- readRDS(file = "results/reliability_results_100splits.rds")
+
+# loading all RDS files
+files <- list.files(
+    path = "results/reliability_cluster_results",
+    pattern = ".rds", 
+    full.names = TRUE
+    )
+
+# sorting files numerically by the number in the filename
+files <- files[order(readr::parse_number(files) )]
+
+# reading and binding all files, adding a sim_id
+reliability_results <- map2_dfr(
+    .x = files,
+    .y = seq_along(files),
+    .f = ~ readRDS(.x)
+    # .f = ~ readRDS(.x) |> mutate(sim_id = .y)
+    # .f = function (file, index) {
+    #     offset <- (index - 1) * splits_per_file
+    #     readRDS(file) %>% mutate(split_id = as.numeric(split_id) ) %>%
+    #         mutate(split_id = split_id + offset) %>%
+    #         mutate(split_id = formatC(x = split_id, width = 3, flag = 0) ) 
+    #     }
+    )
+
+# sanity checks
+# unique(reliability_results$split_id)
+# n_distinct(reliability_results$split_id)
 
@@
 full_reliability_results <- reliability_results %>%
-    filter(split_id == max(split_id) ) %>%
+    # filter(split_id == max(split_id) ) %>%
+    # filter(str_detect(string = split_id, pattern = "_s051") ) %>%
+    filter(split_id == "a001_s051") %>%
     # keeping only the first and last timestep
@@
 # saving the plot
-ggsave(
-    filename = "figures/meg_decoding_results_all_methods.png",
-    width = 12, height = 6, dpi = 300,
-    device = "png"
-    )
+# ggsave(
+#     filename = "figures/meg_decoding_results_all_methods.png",
+#     width = 12, height = 6, dpi = 300,
+#     device = "png"
+#     )
 ```
@@
 This figure reveals that the `BGAM` *onset and offset* estimates on each split are the closest to the estimates from the full dataset on average (0ms difference for the onset estimate and 5ms difference for the offset estimate).
-The `Raw p-value` method has similar performance, but given the aberrant estimates it produces (cf.
+The `Raw p-value` method has similar performance for the offset, but given the aberrant estimates it produces (cf.
 @fig-onset-offset), the fact that it is consistent between data splits and the full dataset is not convincing on its own.
 The `Change point` method also has a very good performance (i.e., very low difference between split estimates and full estimates), but produces too short cluster of significant decoding performance (cf.
-@fig-onset-offset).[^2] Overall, the figure reveals that for all other methods, split datasets produce later onset estimates and earlier offset estimates (as compared to the estimates from the model fitted on the full dataset).
-These results highlight some desirable properties for a method aiming to precisely and reliably estimate the onset and offset of M/EEG effects, namely, it should i) have good asymptotic properties on simulated data, ii) provide sensible identified clusters in actual data, and iii) provide reliable/stable estimates on actual data.
-
-[^2]: As in @rousselet_using_2025, we fixed the number of expected change points to two in the binary segmentation algorithm, thus producing always one cluster.
+@fig-onset-offset).[^3] Overall, the figure reveals that for all other methods, split datasets produce later onset estimates and earlier offset estimates (compared to the estimates from the model fitted on the full dataset).
+These results highlight some desirable properties for a method aiming to precisely and reliably estimate the onset and offset of M/EEG effects, namely, it should i) have good asymptotic properties on simulated data, ii) provide sensible identified clusters in actual data, and iii) provide reliable/stable estimates across various partitions of the data.
+All these desiderata seem to be fulfilled by the BGAM approach.
+
+[^3]: As in @rousselet_using_2025, we fixed the number of expected change points to two in the binary segmentation algorithm, thus producing always one cluster.
 
 ```{r fig-reliability, echo = FALSE, cache = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median error and median absolute deviation of the error for the onset (left) and offset (right) estimates according to each method.
-Methods are ordered from lowest (top) to highest (bottom) median absolute error (separately for the onset and offset estimates)."}
+Methods are ordered from smallest (top) to largest (bottom) median absolute error separately for the onset and offset estimates."}
 # extracting the full data estimates
 full_reliability_results <- reliability_results %>%
-    filter(split_id == max(split_id) ) %>%
+    # filter(split_id == max(split_id) ) %>%
+    filter(split_id == "a001_s051") %>%
     # keeping only the first and last timestep
@@
 
-# or plotting only the error in descending order
+# sanity check
+# reliability_results %>%
+#     filter(str_detect(string = split_id, pattern = "s051", negate = TRUE) ) %>%
+#     pull(split_id) %>%
+#     n_distinct()
+
+# plotting the error in descending order
 reliability_results %>%
-    filter(split_id < max(split_id) ) %>%
+    filter(str_detect(string = split_id, pattern = "s051", negate = TRUE) ) %>%
     data.frame() %>%
@@
         # variance = sum(error, na.rm = TRUE)
+        mean_error = mean(error, na.rm = TRUE),
         median_error = median(error, na.rm = TRUE),
         # variance = IQR(error, na.rm = TRUE)
-        variance = mad(error, na.rm = TRUE)
+        mad = mad(error, na.rm = TRUE),
+        variance = var(error, na.rm = TRUE)
         # .groups = "drop"
@@
             by = 1 / abs(median_error),
+            # by = 1 / abs(mean_error),
             within = onset_offset
@@
             x = median_error,
-            xmin = median_error - variance,
-            xmax = median_error + variance,
+            xmin = median_error - mad,
+            xmax = median_error + mad,
+            # x = mean_error,
+            # xmin = mean_error - variance,
+            # xmax = mean_error + variance,
             y = method_reordered,
@@
     labs(
-        x = "Median difference to full dataset (s)",
+        x = "Difference to full dataset (s)",
         y = "Method"
         )
+
+# plotting the distribution of results
+# reliability_results %>%
+#     # filter(split_id < max(split_id) ) %>%
+#     filter(str_detect(string = split_id, pattern = "s051", negate = TRUE) ) %>%
+#     data.frame() %>%
+#     # keeping only the first and last timestep
+#     group_by(split_id, method, onset_offset) %>%
+#     summarise(
+#         value = if (onset_offset[1] == "onset") min(value) else max(value),
+#         .groups = "drop"
+#         ) %>%
+#     ungroup() %>%
+#     select(-split_id) %>%
+#     left_join(
+#         full_reliability_results,
+#         # by = "measure",
+#         by = c("method", "onset_offset"),
+#         suffix = c("_split", "_full")
+#         ) %>%
+#     group_by(method, onset_offset) %>%
+#     # mutate(error = abs(value_split - value_full) ) %>%
+#     mutate(error = value_split - value_full) %>%
+#     mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
+#     mutate(
+#         method = factor(
+#             x = method,
+#             levels = c(
+#                 "raw_p", "pval_bh", "pval_by", "pval_holm",
+#                 "cluster_mass", "cluster_tfce", "cpt", "brms"
+#                 ),
+#             labels = c(
+#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+#                 "TFCE", "Change point", "BGAM (new method)"
+#                 ),
+#             )
+#         ) %>%
+#     mutate(
+#         method_reordered = tidytext::reorder_within(
+#             x = method,
+#             by = abs(error),
+#             fun = median,
+#             # by = 1 / abs(median_error),
+#             # by = 1 / abs(mean_error),
+#             within = onset_offset
+#             )
+#         ) %>%
+#     # mutate(
+#     #     method_reordered = fct_reorder(
+#     #         method, error, .fun = ~abs(median(.x) ), .na_rm = TRUE
+#     #         )
+#     #     ) %>%
+#     # arrange(desc(error) ) %>%
+#     # arrange(error) %>%
+#     # head()
+#     ggplot(
+#         aes(
+#             # x = median_error,
+#             x = error,
+#             # xmin = median_error - mad,
+#             # xmax = median_error + mad,
+#             # x = mean_error,
+#             # xmin = mean_error - variance,
+#             # xmax = mean_error + variance,
+#             y = method_reordered,
+#             colour = method,
+#             fill = method
+#             )
+#         ) +
+#     geom_vline(xintercept = 0, linetype = 2) +
+#     # geom_pointrange(
+#     #     size = 0.25,
+#     #     linewidth = 0.75,
+#     #     show.legend = FALSE
+#     #     ) +
+#     stat_slabinterval(
+#         n = 200,
+#         normalize = "groups",
+#         point_interval = "median_qi",
+#         # .width = c(0.1, 0.9),
+#         # .width = 0.2,
+#         .width = 0,
+#         interval_size_range = c(1, 1),
+#         fatten_point = 1.25,
+#         # slab_colour = "white",
+#         # slab_linewidth = 0,
+#         # density = "unbounded",
+#         slab_alpha = 0.5,
+#         show.legend = FALSE
+#         ) +
+#     facet_wrap(~onset_offset, scales = "free") +
+#     scale_y_reordered() +
+#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+#     labs(
+#         x = "Estimation error (s)",
+#         y = "Method"
+#         )
 ```
 
-```{r fig-reliability-distributions, echo = FALSE, eval = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Median absolute error and variance of onset (left) and offset (right) estimates for each method."}
-# plotting distributions
-reliability_results %>%
-    slice(1:nrow(.)-1) %>%
-    pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
-    data.frame() %>%
-    left_join(
-        full_reliability_results,
-        by = "measure",
-        suffix = c("_split", "_full")
-        ) %>%
-    # mutate(error = abs(value_split - value_full) ) %>%
-    mutate(error = value_split - value_full) %>%
-    # mutate(error = (value_split - value_full)^2) %>%
-    separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
-    # now plotting
-    mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
-    filter(method != "brms_full") %>%
-    mutate(
-        method = factor(
-            x = method,
-            levels = c(
-                "p", "bh", "by", "holm", "cluster_mass",
-                "cluster_tfce", "cpt", "brms"
-                ),
-            labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-                "TFCE", "Change point", "BGAM"
-                ),
-            )
-        ) %>%
-    mutate(
-        plot_row = case_when(
-            method %in% c("Raw p-value", "Holm") ~ 1,
-            method %in% c("FDR BH95", "FDR BY01") ~ 2,
-            method %in% c("Cluster mass", "TFCE") ~ 3,
-            method %in% c("Change point", "BGAM") ~ 4
-            )
-        ) %>%
-    ggplot(
-        aes(
-            x = error,
-            colour = method,
-            fill = method
-            )
-        ) +
-    geom_vline(xintercept = 0, linetype = 2) +
-    geom_density(aes(fill = NULL) ) +
-    facet_grid(plot_row~type, scales = "free_x") +
-    theme_light(base_size = 12, base_family = "Open Sans") +
-    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    labs(x = "Onset/Offset (s)", y = "Density")
-```
-
 ```{r tbl-reliability-results, eval = TRUE, echo = FALSE, results = "asis"}
-#| tbl-cap: "Summary statistics of the onset and offset estimates for each method (ordered by the absolute value of the bias)."
-
-# retrieving the reliability results
-reliability_results <- readRDS(file = "results/reliability_results_100splits.rds")
+#| tbl-cap: "Summary statistics of the onset and offset estimates for each method (in ms, ordered by the MAE) according to the reliability simulation study."
 
@@
     group_by(method, onset_offset) %>%
+    mutate(error = 1e3 * error) %>%
     summarise(
+        # mean_error = mean(error, na.rm = TRUE),
         median_error = median(error, na.rm = TRUE),
@@
     # arrange(median_abs_error) %>%
+    # arrange(abs(median_error) ) %>%
+    # arrange(median_abs_error) %>%
     arrange(abs(median_error) ) %>%
@@
         method = "",
+        # mean_error = "Bias_mean",
         median_error = "Bias",
@@
 
-We kept the exemplary models simple, but can be extended by adding varying/random effects (intercept and slope) for item (e.g., word)... but also continuous predictors at the trial level?
+We kept the exemplary models simple, but can be extended by adding varying/random effects (intercept and slope) for item (e.g., word)...
+but also continuous predictors at the trial level, overlapping factors, etc...
 
@@
 
-```{r, echo = FALSE}
+```{r, eval = TRUE, echo = FALSE, cache = FALSE}
+# checking the included packages
+# grateful::scan_packages(pkgs = "Session")
+
 # citing the R packages we have used
@@
     omit = c("permuco", "reticulate"),
-    out.dir = "."
+    out.dir = getwd(),
+    dependencies = FALSE
     )
@@
 
-Centre de Calcul Intensif d’Aix-Marseille is acknowledged for granting access to its high performance computing resources.
+Centre de Calcul Intensif d'Aix-Marseille is acknowledged for granting access to its high performance computing resources.
 
@@
 
-```{r fig-sim-timegen, echo = FALSE, dpi = 300, out.width = "75%", fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
-# number of participants
-n_participants <- 10
-
-# number of timepoints
-n_timepoints <- 60 # (1.2s at 50Hz)
-timepoints <- seq(from = -0.2, to = 1, length.out = n_timepoints)
-
-# smooth onset and offset parameters
-decode_onset <- 0.2
-decode_offset <- 0.3
-
-# smoothness of onset/offset
-sigma_smooth <- 0.05
-
-# generating participant-specific cross-temporal generalisation matrices
-timegen_data <- data.frame()
-
-# for each participant
-for (p in 1:n_participants) {
-
-    # adding variability in peak amplitude and width per participant
-    peak_variability <- runif(1, 30, 50) # random peak strength (higher = stronger decoding)
-    smooth_variability <- runif(1, 0.04, 0.06) # random smoothness variation
-    onset_variability <- runif(1, -0.05, 0.05) # small jitter in onset time
-    offset_variability <- runif(1, -0.05, 0.05) # small jitter in offset time
-
-    # computing smooth transition function (soft onset and offset with participant variation)
-    decode_window <- outer(timepoints, timepoints, function (t1, t2) {
-
-        # smooth onset
-        onset_factor <- 1 /
-            (1 + exp(-(t1 - (decode_onset + onset_variability) ) / sigma_smooth) )
-
-        # smooth offset
-        offset_factor <- 1 /
-            (1 + exp((t1 - (decode_offset + offset_variability) ) / sigma_smooth) )
-
-        # diagonal Gaussian
-        onset_factor * offset_factor * exp(-((t1 - t2)^2) / (2 * smooth_variability^2) )
-
-    })
-
-    # defining smooth alpha and beta parameters for Beta distribution
-    base_alpha <- 10
-    peak_alpha <- peak_variability
-    alpha_matrix <- base_alpha + peak_alpha * decode_window
-    beta_matrix <- base_alpha
-
-    # generating AUC values using the Beta distribution
-    auc_matrix_beta <- matrix(
-        rbeta(n_timepoints^2, shape1 = alpha_matrix, shape2 = beta_matrix),
-        nrow = n_timepoints, ncol = n_timepoints
-        )
-
-    # converting to long format for plotting
-    temp_data <- expand.grid(
-        train_time = timepoints, test_time = timepoints
-        ) %>%
-        mutate(auc = as.vector(auc_matrix_beta), participant = p)
-
-    # storing results
-    timegen_data <- bind_rows(timegen_data, temp_data)
-
-}
-
-# plotting cross-temporal generalization matrix
+```{r fig-timegen-tickertape, eval = FALSE, echo = FALSE, dpi = 300, out.width = "75%", fig.cap = "Exemplary group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
+# importing the MEG decoding data
+library(reticulate)
+use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
+np <- import("numpy")
+data <- np$load("data/group_level_decoding_4l_2s_vs_7l_2s_visual_blocks_temporal_generalisation_all.npy")
+# data <- np$load("data/group_level_decoding_7l_2s_vs_7l_3s_visual_blocks_temporal_generalisation_all.npy")
+
+# converting to R array
+data_r <- py_to_r(data)
+str(data_r)
+dim(data_r)
+
+# original time vector (before decimation)
+time_points <- seq(from = -0.5, to = 1.0, length.out = dim(data_r)[2])
+
+# removing the first 300ms before stimulus onset
+start_idx <- which.min(abs(time_points - (-0.2) ) )
+end_idx <- which.min(abs(time_points - 1.0) )
+
+# cutting data_r to keep only time points between -0.2s and 1.0s
+data_cut <- data_r[, start_idx:end_idx, start_idx:end_idx]
+
+# updating time vector accordingly
+time_points_cut <- time_points[start_idx:end_idx]
+
+# checking new dimensions
+dim(data_cut)
+length(time_points_cut)
+
+# defining a decimation factor
+decim <- 4
+
+# selecting every 4th time point
+data_cut_decim <- data_cut[, seq(1, dim(data_cut)[2], by = decim), seq(1, dim(data_cut)[3], by = decim)]
+
+# retrieving the number of participants, time points
+n_participants <- dim(data_cut_decim)[1]
+n_times <- dim(data_cut_decim)[2]
+
+# defining a time vector (if not already defined)
+time_points <- seq(from = -0.2, to = 1.0, length.out = n_times)
+
+# converting array to long format using `as.data.frame.table`
+timegen_data <- as.data.frame.table(data_cut_decim)
+
+# renaming and converting indices
+colnames(timegen_data) <- c("participant", "train_idx", "test_idx", "auc")
+
+timegen_data <- timegen_data %>%
+    mutate(
+    participant = as.integer(participant),
+    train_idx = as.integer(train_idx),
+    test_idx = as.integer(test_idx),
+    train_time = time_points[train_idx],
+    test_time = time_points[test_idx]
+    ) %>%
+    select(train_time, test_time, auc, participant) %>%
+    filter(participant < 11)
+
+# plotting the cross-temporal generalisation matrix
 timegen_data %>%
@@
 
+```{r fig-sim-timegen, echo = FALSE, dpi = 300, out.width = "75%", fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
+# number of participants
+n_participants <- 10
+
+# number of timepoints
+n_timepoints <- 60 # (1.2s at 50Hz)
+timepoints <- seq(from = -0.2, to = 1, length.out = n_timepoints)
+
+# smooth onset and offset parameters
+decode_onset <- 0.2
+decode_offset <- 0.3
+
+# smoothness of onset/offset
+sigma_smooth <- 0.05
+
+# generating participant-specific cross-temporal generalisation matrices
+timegen_data <- data.frame()
+
+# for each participant
+for (p in 1:n_participants) {
+
+    # adding variability in peak amplitude and width per participant
+    peak_variability <- runif(1, 30, 50) # random peak strength (higher = stronger decoding)
+    smooth_variability <- runif(1, 0.04, 0.06) # random smoothness variation
+    onset_variability <- runif(1, -0.05, 0.05) # small jitter in onset time
+    offset_variability <- runif(1, -0.05, 0.05) # small jitter in offset time
+
+    # computing smooth transition function (soft onset and offset with participant variation)
+    decode_window <- outer(timepoints, timepoints, function (t1, t2) {
+
+        # smooth onset
+        onset_factor <- 1 /
+            (1 + exp(-(t1 - (decode_onset + onset_variability) ) / sigma_smooth) )
+
+        # smooth offset
+        offset_factor <- 1 /
+            (1 + exp((t1 - (decode_offset + offset_variability) ) / sigma_smooth) )
+
+        # diagonal Gaussian
+        onset_factor * offset_factor * exp(-((t1 - t2)^2) / (2 * smooth_variability^2) )
+
+    })
+
+    # defining smooth alpha and beta parameters for Beta distribution
+    base_alpha <- 10
+    peak_alpha <- peak_variability
+    alpha_matrix <- base_alpha + peak_alpha * decode_window
+    beta_matrix <- base_alpha
+
+    # generating AUC values using the Beta distribution
+    auc_matrix_beta <- matrix(
+        rbeta(n_timepoints^2, shape1 = alpha_matrix, shape2 = beta_matrix),
+        nrow = n_timepoints, ncol = n_timepoints
+        )
+
+    # converting to long format for plotting
+    temp_data <- expand.grid(
+        train_time = timepoints, test_time = timepoints
+        ) %>%
+        mutate(auc = as.vector(auc_matrix_beta), participant = p)
+
+    # storing results
+    timegen_data <- bind_rows(timegen_data, temp_data)
+
+}
+
+# plotting the cross-temporal generalisation matrix
+timegen_data %>%
+    summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>%
+    ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
+    geom_tile() +
+    geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
+    geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
+    geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
+    scale_x_continuous(expand = c(0, 0) ) +
+    scale_y_continuous(expand = c(0, 0) ) + 
+    coord_fixed() +
+    scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
+    labs(x = "Testing time (s)", y = "Training time (s)")
+```
+
 ```{r mne-timegen, eval = FALSE, echo = FALSE}
@@
 
-```{=html}
-<!--
-
-Now, we want to test whether and when decoding performance is above chance level (0.5 for a binary decoding task).
-These two models are computationally heavier to fit (more observations and 2D smooth functions)...
-
-See https://bookdown.org/epeterson_2010/bios526_book/Module_5_3.html#bivariate-splines
-
--->
-```
-
 To model cross-temporal generalisation matrices of decoding performance (ROC AUC), we extended the initial (decoding) GAM to take into account the bivariate temporal distribution of AUC values, thus producing naturally smoothed estimates (timecourses) of AUC values and posterior probabilities.
@@
 
-where $\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|$ is the Euclidean distance between the covariate $\mathbf{s}_{i}$ and the knot location $\mathbf{k}_{m}$. We fitted this model using `brms`...
-
-```{r gam-timegen, eval = TRUE, echo = TRUE, results = "hide"}
+where $\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|$ is the Euclidean distance between the covariate $\mathbf{s}_{i}$ and the knot location $\mathbf{k}_{m}$.
+We fitted this model using `brms`...
+using the `t2()` smooth constructor with full penalties [@pedersen_hierarchical_2019]...
+
+```{r gam-timegen, eval = FALSE, echo = TRUE, results = "hide"}
 # fitting a GAM with two temporal dimensions
@@
     )
+
+# fitting a GAM with two temporal dimensions to empirical data
+timegen_gam <- brm(
+    # 2D thin-plate spline (tp)
+    # auc ~ t2(train_time, test_time, bs = "tp", k = 30),
+    auc ~ t2(train_time, test_time, bs = "tp", k = 20, full = TRUE),
+    data = timegen_data,
+    family = Beta(),
+    warmup = 1000,
+    iter = 3000,
+    chains = 8,
+    cores = 8,
+    # file = "models/timegen_gam_meg_t2_k30.rds"
+    file = "models/timegen_gam_meg_t2_k20_full.rds"
+    )
+
+# retrieving the data
+# timegen_data <- timegen_gam$data
+
+# PPCs
+# pp_check(timegen_gam)
 ```
@@
 
-```{r gam-timegen-preds, eval = FALSE, echo = FALSE}
-# plotting the posterior predictions
-plot(
-    conditional_effects(x = timegen_gam, prob = 0.95),
-    points = FALSE,
-    # point_args = list(size = 1, pch = 21, colour = "grey"),
-    # line_args = list(colour = "black"),
-    method = "posterior_epred",
-    surface = TRUE,
-    # theme = theme_light(),
-    plot = TRUE
-    )#[[1]] +
-    # geom_hline(yintercept = 0.5, linetype = 2) +
-    # geom_vline(xintercept = 0.0, linetype = 2) +
-    # ylim(0, 1) +
-    # labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
-
-# posterior predictive checks
-pp_check(timegen_gam)
-```
-
-```{r gam-timegen-post-preds, echo = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Simulated data with thresholds derived from the BGAM (left), predicted AUC values (middle), and ratios of posterior probabilies of decoding accuracy being above chance level (right) according to the bivariate BGAM (for k=10, top row, and k=20, bottom row)."}
+```{r gam-timegen-post-preds, eval = FALSE, echo = FALSE, cache = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Simulated data with thresholds derived from the BGAM (left), predicted AUC values (middle), and posterior probabily ratios of decoding accuracy being above chance level (right) according to the bivariate BGAM (for k=10, top row, and k=20, bottom row)."}
 # defining a function to plot posterior predictions
@@
     # defining a sequence of x values to make predictions
+    # preds_conds <- crossing(
+    #     train_time = unique(data$train_time)[c(TRUE, FALSE)],
+    #     test_time = unique(data$test_time)[c(TRUE, FALSE)]
+    #     )
     preds_conds <- crossing(
-        train_time = unique(data$train_time)[c(TRUE, FALSE)],
-        test_time = unique(data$test_time)[c(TRUE, FALSE)]
+        train_time = unique(data$train_time),
+        test_time = unique(data$test_time)
         )
@@
         prob = 0.95,
-        ndraws = 100
+        ndraws = 1e2,
+        # ndraws = 1e3,
+        resolution = n_distinct(data$train_time)
         )[[1]]
+
+    # sanity checks
+    # nrow(preds_conds)
+    # unique(post_preds$train_time)
+    # unique(post_preds$test_time)
+    # post_preds %>%
+    #     summarise(auc = mean(estimate__), .by = c(train_time, test_time) ) %>%
+    #     # head()
+    #     nrow()
+
+    # combining AUC ranges from both plots
+    all_aucs <- c(
+        post_preds %>% summarise(auc = mean(estimate__), .by = c(train_time, test_time) ) %>% pull(auc),
+        data %>% summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>% pull(auc)
+        )
+
+    # determining common limits
+    fill_limits <- range(all_aucs, na.rm = TRUE)
 
@@
         coord_fixed() +
-        scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
+        # scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
+        scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC", limits = fill_limits) +
         labs(x = "Testing time (s)", y = "Training time (s)")
@@
             midpoint = 0,
-            name = "Posterior ratio",
+            name = "Post. prob. ratio",
             trans = "log10"
@@
 
-    # plotting cross-temporal generalization matrix
+    # plotting cross-temporal generalisation matrix
     threshold_data_plot <- data %>%
@@
         coord_fixed() +
-        scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
+        # scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
+        scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC", limits = fill_limits) +
         labs(x = "Testing time (s)", y = "Training time (s)") +
@@
 
-        # combining the two plots
+    # combining the plots
     threshold_data_plot + post_preds_plot + prob_above_chance_plot
@@
 bgam_k20 <- readRDS(file = "models/timegen_gam_t2_k20.rds")
-baseline_sd <- timegen_data %>% filter(train_time < 0 & test_time < 0.2) %>% summarise(sd(auc) ) %>% pull()
+
+# computing the SESOI as the baseline SD
+baseline_sd <- timegen_data %>%
+    filter(train_time < 0 & test_time < 0.2) %>%
+    summarise(sd(auc) ) %>% pull()
+
+# posterior predictive checks
+# pp_check(bgam_k10)
+# pp_check(bgam_k20)
 
@@
     model = bgam_k10,
-    chance_level = 0.5, sesoi = baseline_sd,
+    # model = timegen_gam,
+    chance_level = 0.5,
+    sesoi = baseline_sd,
     # thresholds = c(10, 100)
-    thresholds = 10
+    # thresholds = 10
+    thresholds = 20
     )
@@
     model = bgam_k20,
-    chance_level = 0.5, sesoi = baseline_sd,
+    chance_level = 0.5,
+    sesoi = baseline_sd,
     # thresholds = c(10, 100),
-    thresholds = 10
+    # thresholds = 10
+    thresholds = 20
     )
@@
 
-Here provide recommendation about how to define `k`.
+Here we should provide recommendation about how to define `k`...
 An option is to vary `k` and examine the predictions and posterior predictive checks (PPCs) of each model...
@@
 However, it is not possible to provide general recommendations, as the optimal `k` depends on the sampling rate, the preprocessing steps (e.g., signal-to-noise ratio, low-pass filtering, etc), and the neural dynamics of the phenomenon under study.
+Choose `k` as large a possible (given limits on computation time etc)?
+Or use LOOIC/WAIC (seem to select too simple models)?
 
@@
 For users who are already familiar with `brms`, the recommended pipeline is to import ERPs or decoding results in `R` and analyse these data using the code provided in the main paper.
-However, it is also possible to call functions from the `neurogam R` package (available at <https://github.com/lnalborczyk/neurogam>), which come with sensible defaults.
+It is also possible to call functions from the `neurogam R` package (available at <https://github.com/lnalborczyk/neurogam>) which come with sensible defaults.
 

### 2025-05-22 manuscript.tex
pushing a first version of intro+methods+results

--- 2025-04-11 brms_meeg.tex
@@
 obtained through magneto- or electro-encephalography (M/EEG) offer a
-unique window into the neural activity underlying cognitive processes.
+unique window onto the neural activity underlying cognitive processes.
 Researchers are often interested in determining whether and when these
@@
 significant limitation: by shifting the focus of inference from
-individual time points to clusters, it makes difficult to draw precise
-conclusions about the onset or offset of observed effects. Here, we
+individual time points to clusters, it prevents drawing conclusions
+about the precise onset or offset of observed effects. Here, we
 introduce a \emph{model-based} approach for analysing M/EEG timeseries
@@
 M/EEG studies is to determine whether, when, and where neural responses
-differ across experimental conditions or participant groups.
+differ across experimental conditions or groups.
 
 The conventional approach involves mass-univariate statistical testing
-through time and/or space followed by some form or correction for
+through time and/or space followed by some form of correction for
 multiple comparisons with the goal of maintaining the familywise error
@@
 perform poorly for localising the onset of M/EEG effects; a property
-that was later demonstrated in simulations studies (e.g.,
+that was later demonstrated in simulation studies (e.g.,
 \citeproc{ref-rousselet_using_2025}{Rousselet, 2025};
@@
 The issues with multiple comparisons represent a common and
-well-recognised danger in neuroimaging and M/EEG research where the
+well-recognised danger in neuroimaging and M/EEG research, where the
 collected data allows for a multitude of potential hypothesis tests and
@@
 control the FWER (i.e., bring it back to \(\alpha\)). Most methods apply
-a simple correction to series of ``raw'' \(p\)-values issued from
-univariate statistical tests (e.g., t-tests). For instance, the
-Bonferroni correction (\citeproc{ref-dunn1961}{Dunn, 1961}) consists in
-setting the significance threshold to \(\alpha/m\), or equivalently,
-multiplying the ``raw'' \(p\)-values by \(m\) and using the standard
-\(\alpha\) significance threshold. This method is generally
-overconservative (i.e., under-powered) as it assumes statistical
-independence of the tests, an assumption that is clearly violated in the
-context of M/EEG timeseries characterised by massive spatiotemporal
-dependencies. Some alternative methods aims at controlling the FDR,
-defined as the proportion of false positive \emph{among positive tests}
-(e.g., \citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
+a simple correction to series of \(p\)-values issued from univariate
+statistical tests (e.g., t-tests). For instance, the Bonferroni
+correction (\citeproc{ref-dunn1961}{Dunn, 1961}) consists in setting the
+significance threshold to \(\alpha/m\), or equivalently, multiplying the
+\(p\)-values by \(m\) and using the standard \(\alpha\) significance
+threshold. This method is generally overconservative (i.e.,
+under-powered) as it assumes statistical independence of the tests, an
+assumption that is clearly violated in the context of M/EEG timeseries
+characterised by massive spatiotemporal dependencies. Some alternative
+methods aims at controlling the FDR, defined as the proportion of false
+positive \emph{among positive tests} (e.g.,
+\citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
 \citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001}). However, a
@@
 In other words, because the inference is performed at the second step
-(i.e., once clusters have been formed), it prevents any conclusion to be
-made about individual datapoints (e.g., timesteps, sensors, etc).
+(i.e., once clusters have been formed), no conclusion can be made about
+individual datapoints (e.g., timesteps, sensors, etc).
 
@@
 set a priori (typically to 0.5 and 2, respectively) and control the
-influence of the extend and height on the TFCE. Note that in practise,
+influence of the extend and height on the TFCE. Note that in practice,
 this integral is approximated by a sum over small \(h\) increments.
-Then, a p-value for each timestep \(t\) is computed by comparing its
-TFCE with the null distribution of TFCE values (obtained via
+Then, a \emph{p}-value for each timestep \(t\) is computed by comparing
+its TFCE with the null distribution of TFCE values (obtained via
 permutation). For each permuted signal, we keep the maximal value over
@@
 allow inferring the precise spatial and temporal localisation of
-effects. In the following, we briefly review previous M/EEG modelling
-work. Then, we provide a short introduction to generalised additive
-models (GAMs) and Bayesian generalised additive multilevel models
-(BGAMMs) to illustrate how these models can be used to precisely
-localise the onset and offset of M/EEG effects.
+effects. In the following, we briefly review previous modelling work of
+M/EEG data. Then, we provide a short introduction to generalised
+additive models (GAMs) to illustrate how these models can be used to
+precisely estimate the onset and offset of M/EEG effects.
 
@@
 
-Recent example of GLM for EEG (\citeproc{ref-fischer2013}{Fischer \&
-Ullsperger, 2013}; \citeproc{ref-wuxfcllhorst2025}{Wüllhorst et al.,
-2025})\ldots{} See also (\citeproc{ref-hauk2006}{Hauk et al., 2006};
-\citeproc{ref-rousselet2008}{Rousselet et al., 2008})\ldots{} Example of
-two-stage regression analysis (i.e., individual-level then group-level,
-\citeproc{ref-dunagan2024}{Dunagan et al., 2024})\ldots{}
-
-As put by Rousselet (\citeproc{ref-rousselet_using_2025}{2025}),
-concluding on the onset of effect based on a series of univariate
-tests\ldots{} commits to three fallacies\ldots{} here, we want to avoid
-these by introducing a model-based approach, which naturally take into
-account the temporal dependencies in the data to output a series of
-posterior probabilities\ldots{}
-
-See also the rERP framework (\citeproc{ref-smith2014a}{N. J. Smith \&
-Kutas, 2014a}, \citeproc{ref-smith2014b}{2014b}) and Tremblay \& Newman
-(\citeproc{ref-tremblay2014}{2014})\ldots{}
-
-From Dimigen \& Ehinger (\citeproc{ref-dimigen2021}{2021}): Recently,
-spline regression has been applied to ERPs (e.g.,
-\citeproc{ref-hendrix2017}{Hendrix et al., 2017};
-\citeproc{ref-kryuchkova2012}{Kryuchkova et al., 2012})\ldots{} GAMMs
-for EEG data (\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
+Scalp-recorded M/EEG signals capture neural activity originating from
+various brain regions and are often contaminated by artifacts unrelated
+to the cognitive processes under investigation. Consequently, analysing
+M/EEG data necessitates methods that can disentangle task-relevant
+neural signals from extraneous ``noise.'' A widely adopted technique for
+this purpose is the estimation of event-related potentials (ERPs), which
+are stereotyped electrophysiological responses time-locked to specific
+sensory, cognitive, or motor events. Typically, ERPs are derived by
+averaging EEG or MEG epochs across multiple trials aligned to the event
+of interest (e.g., stimulus onset), thereby enhancing the
+signal-to-noise ratio by attenuating non-time-locked activity. However,
+this averaging approach has notable limitations: it assumes consistent
+latency and amplitude across trials and is primarily suited for simple
+categorical designs. Such assumptions may not hold in more complex
+experimental paradigms, potentially leading to suboptimal ERP
+estimations.
+
+To overcome the limitations of simple averaging, several model-based
+approaches for estimating ERPs have been proposed. These methods are
+motivated by the observation that traditional ERP averaging is
+mathematically equivalent to fitting an intercept-only linear regression
+model in a simple categorical design without overlapping events
+(\citeproc{ref-smith2014a}{N. J. Smith \& Kutas, 2014a}). In contrast to
+simple averaging, regression-based approaches to ERP estimation offer
+substantially greater flexibility. Notably, they allow for the modelling
+of both linear and nonlinear effects of continuous predictors--such as
+word frequency or age (e.g., \citeproc{ref-smith2014a}{N. J. Smith \&
+Kutas, 2014a}, \citeproc{ref-smith2014b}{2014b};
+\citeproc{ref-tremblay2014}{Tremblay \& Newman, 2014})--and enable the
+disentangling of overlapping cognitive processes (e.g.,
+\citeproc{ref-ehinger_unfold_2019}{Ehinger \& Dimigen, 2019};
+\citeproc{ref-skukies_brain_2024}{Skukies et al., 2024};
+\citeproc{ref-skukies_modelling_2021}{Skukies \& Ehinger, 2021}). One
+widely used implementation of this approach is provided by the LIMO EEG
+toolbox (\citeproc{ref-pernet2011}{Pernet et al., 2011}), which follows
+a multi-stage analysis pipeline. First, a separate regression model is
+fit for each datapoint (e.g., each time point and electrode) at the
+individual participant level to estimate ERP responses. This is followed
+by group-level statistical analyses of the resulting regression
+coefficients, often accompanied by multiple comparison corrections or
+cluster-based inference (for recent applied examples, see
+\citeproc{ref-dunagan2025}{Dunagan et al., 2025};
+\citeproc{ref-wuxfcllhorst2025}{Wüllhorst et al., 2025}).
+
+Although this framework allows for the inclusion of a wide range of
+predictors--both continuous and categorical, linear and nonlinear--it
+still has important limitations. First, fitting separate models for each
+datapoint ignores the spatiotemporal dependencies inherent in M/EEG
+data, potentially reducing statistical power and interpretability.
+Second, the subsequent group-level analyses typically do not account for
+hierarchical dependencies, between-participant or between-stimulus
+variability, which could otherwise be addressed through multilevel
+modelling. Finally, because the output of this procedure is summarised
+by cluster-based inference, its conclusions remain subject to the
+limitations discussed in the previous section.
+
+Beyond modelling nonlinear effects of continuous predictors on ERP
+amplitudes, generalised additive models (GAMs) have been employed to
+capture the temporal dynamics of ERPs themselves, effectively modelling
+the shape of the waveform over time
+(\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
+\citeproc{ref-baayen2018}{Baayen et al., 2018};
 \citeproc{ref-meulman2015}{Meulman et al., 2015},
-\citeproc{ref-meulman2023}{2023})\ldots{} See also the UNFOLD toolbox
-(\citeproc{ref-ehinger_unfold_2019}{Ehinger \& Dimigen, 2019})\ldots{}
-
-Disentangling overlapping processes
-(\citeproc{ref-skukies_brain_2024}{Skukies et al., 2024};
-\citeproc{ref-skukies_modelling_2021}{Skukies \& Ehinger, 2021})\ldots{}
-Weighting single trials (\citeproc{ref-pernet2022}{Pernet,
-2022})\ldots{} The LIMO toolbox (\citeproc{ref-pernet2011}{Pernet et
-al., 2011}) using linear and 2-stage regression (not a proper multilevel
-model)\ldots{}
-
-Recently, Teichmann (\citeproc{ref-teichmann2022}{2022}) provided a
-detailed tutorial on using Bayes factors (BFs) to analyse the 1D or 2D
-output from MVPA, that is, for testing, at every timestep, whether
-decoding performance is above chance level. However, this approach
-provides timeseries of BFs that ignores temporal dependencies\ldots{}
+\citeproc{ref-meulman2023}{2023}). This approach allows for the
+estimation of smooth, data-driven functions that characterise how neural
+responses evolve over time, offering a flexible alternative to
+traditional linear models. In the following section, we provide a brief
+introduction to GAMs, highlighting their applicability to M/EEG time
+series analysis and the advantages they offer over conventional methods.
 
@@
 
-See for instance these tutorials
-(\citeproc{ref-suxf3skuthy2017}{Sóskuthy, 2017};
-\citeproc{ref-winter2016}{Winter \& Wieling, 2016}) or application to
-phonetic data (\citeproc{ref-suxf3skuthy2021}{Sóskuthy, 2021};
-\citeproc{ref-wieling2018}{Wieling, 2018}) or this introduction
-(\citeproc{ref-baayen2020}{Baayen \& Linke, 2020}) or these reference
-books (\citeproc{ref-hastie2017}{Hastie \& Tibshirani, 2017};
-\citeproc{ref-wood2017}{Wood, 2017a})\ldots{} application to
-pupillometry (\citeproc{ref-vanrij2019}{Rij et al., 2019})\ldots{}
-GAMLSS for neuroimaging data (\citeproc{ref-dinga2021}{Dinga et al.,
-2021})\ldots{} Modelling auto-correlation in GAMMs + EEG example
-(\citeproc{ref-baayen2018}{Baayen et al., 2018})\ldots{}
-
 In generalised additive models (GAMs), the functional relationship
-between predictors and response variable is decomposed into a sum of
-low-dimensional non-parametric functions. A typical GAM has the
+between the predictors and the response variable is decomposed into a
+sum of low-dimensional non-parametric functions. A typical GAM has the
 following form:
@@
 function of covariate \(x_{j}\) (to be estimated as well). The smooth
-functions \(f_{j}\) are represented in the model via penalised splines
-basis expansions of the covariates, that are a weighted sum of \(K\)
-simpler, basis functions:
+functions \(f_{j}\) are represented in the model as a weighted sum of
+\(K\) simpler, basis functions:
 
@@
 \(x_{ij}\) for the \(j\)th smooth function \(f_{j}\). To clarify the
-terminology at this stage: \emph{splines} are functions composed of
-simpler functions. These simpler functions are basis functions (e.g.,
-cubic polynomial, thin-plate) and the set of basis functions is a
-\emph{basis}. Each basis function gets its coefficient and the resultant
-spline is the sum of these weighted basis functions
-(Figure~\ref{fig-intro-gam}). Splines coefficients are penalised
-(usually through the squared of the smooth functions' second derivative)
-in a way that can be interpreted, in Bayesian terms, as a prior on the
-``wiggliness'' of the function. In other words, more complex (wiggly)
-basis functions are penalised.
-
-\subsection{Bayesian generalised additive multilevel
-models}\label{bayesian-generalised-additive-multilevel-models}
-
-The Bayesian approach to statistical modelling is characterised by its
-reliance on probability theory to (\citeproc{ref-gelman2020}{Gelman et
-al., 2020})\ldots{} In this framework, all unknown entities are assigned
-probability distributions reflecting the uncertainty\ldots{} These
-probability distributions are commonly referred to as ``priors'' and
-represent some state of knowledge about unknown quantities before seeing
-any data. There are debates among Bayesian practitioners as to whether
-prior distributions should encore subjective (personal) beliefs
-or\ldots{} but these debates are outside the scope of the present paper
-and we therefore the interested reader to dedicated work (e.g., XX;
-YY)\ldots{} In practice, weakly informative priors are often used as
-default priors in situations in which subjective priors are difficult to
-define/elicit\ldots{} Bayesian models are then fitted on empirical
-(actual or simulated) data to update prior states of knowledge to
-posterior states of knowledge using Bayes theorem, or in practise,
-sampling-based approximations of the posterior distribution\ldots{}
+terminology at this point: \emph{splines} are functions composed of
+simpler functions. These simpler functions are called \emph{basis
+functions} (e.g., cubic polynomial, thin-plate) and the set of basis
+functions is called a \emph{basis}. Each basis function gets its
+coefficient and the resultant spline is the sum of these weighted basis
+functions (Figure~\ref{fig-intro-gam}\ignorespaces A). Splines
+coefficients are penalised (usually through the squared of the smooth
+functions' second derivative) in a way that can be interpreted, in
+Bayesian terms, as a prior on the ``wiggliness'' of the function
+(\citeproc{ref-miller2025}{Miller, 2025}; \citeproc{ref-wood2017}{Wood,
+2017a}). In other words, more complex (wiggly) basis functions are
+automatically penalised.
 
@@
 
-\caption{\label{fig-intro-gam}Different types of GAM(M)s. \textbf{A}:
-GAMs predictions are computed as the weigthed sum (in black) of basis
+\caption{\label{fig-intro-gam}Different types of GAMs. \textbf{A}: GAMs
+predictions are computed as the weigthed sum (in black) of basis
 functions (here thin-plate basis functions, in grey). \textbf{B}:
@@
 prediction in black. \textbf{C}: Varying-intercept + varying-slope GAMM
-(with common smoother). \textbf{D}: Varying-intercept + varying-slope +
-varying-smoother GAMM. In this model, each participant gets its own
+(with constant smoother). \textbf{D}: Varying-intercept + varying-slope
++ varying-smoother GAMM. In this model, each participant gets its own
 intercept, slope, and degree of `wiggliness' (smoother).}
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-intro-gam-1.png}
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-intro-gam-1.png}
 
@@
 
-See Figure~\ref{fig-intro-gam}\ldots{} Introduction to multilevel GAMs
-(\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al.,
-2019})\ldots{} Now describe the Bayesian GAMM
-(\citeproc{ref-miller2025}{Miller, 2025})\ldots{} Proper inclusion of
-varying/random effects in the model specification protects against
-overly wiggly curves (\citeproc{ref-baayen2020}{Baayen \& Linke,
-2020})\ldots{} Generalising to scale and shape or ``distributional
-GAMs'' (\citeproc{ref-rigby2005}{Rigby \& Stasinopoulos, 2005};
-\citeproc{ref-umlauf2018}{Umlauf et al., 2018}) and applied to
-neuroimaging data (\citeproc{ref-dinga2021}{Dinga et al., 2021})\ldots{}
-
-Instead of averaging, obtain the smooth ERP signal from multilevel
-GAM\ldots{} less susceptible to outliers
-(\citeproc{ref-meulman2023}{Meulman et al., 2023})\ldots{}
+A detailed treatment of the technical underpinnings of GAMs is beyond
+the scope of this article (see reference books such as
+\citeproc{ref-hastie2017}{Hastie \& Tibshirani, 2017};
+\citeproc{ref-wood2017}{Wood, 2017a}). However, it is worth emphasising
+that GAMs have been successfully applied to a wide range of time series
+data across the cognitive sciences, including pupillometry (e.g.,
+\citeproc{ref-vanrij2019}{Rij et al., 2019}), articulography (e.g.,
+\citeproc{ref-wieling2018}{Wieling, 2018}), speech formant dynamics
+(e.g., \citeproc{ref-suxf3skuthy2021}{Sóskuthy, 2021}), neuroimaging
+data (e.g., \citeproc{ref-dinga2021}{Dinga et al., 2021}), and
+event-related potentials (e.g., \citeproc{ref-abugaber2023}{Abugaber et
+al., 2023}; \citeproc{ref-baayen2018}{Baayen et al., 2018};
+\citeproc{ref-meulman2015}{Meulman et al., 2015},
+\citeproc{ref-meulman2023}{2023}). Their appeal for modelling M/EEG data
+lies in their ability to flexibly capture the complex shape of ERP
+waveforms without overfitting, through the use of smooth functions
+constrained by penalisation. Recent extensions, such as distributional
+GAMs (\citeproc{ref-rigby2005}{Rigby \& Stasinopoulos, 2005};
+\citeproc{ref-umlauf2018}{Umlauf et al., 2018a}), allow researchers to
+model not only the mean structure but also the variance (or scale) and
+other distributional properties as functions of predictors--a feature
+that has proven useful in modelling neuroimaging data (e.g.,
+\citeproc{ref-dinga2021}{Dinga et al., 2021}). Moreover, hierarchical or
+multilevel GAMs (\citeproc{ref-pedersen_hierarchical_2019}{E. J.
+Pedersen et al., 2019}) provide a principled way to account for the
+nested structure of M/EEG data (e.g., trials within participants),
+enabling the inclusion of varying intercepts, slopes, and smoothers (as
+illustrated in Figure~\ref{fig-intro-gam}\ignorespaces C-D). This
+approach mitigates the risk of overfitting and reduces the influence of
+outliers on smooth estimates (\citeproc{ref-baayen2020}{Baayen \& Linke,
+2020}; \citeproc{ref-meulman2023}{Meulman et al., 2023}).
 
@@
 
-Given the previously reported limitations of conventional methods to
-precisely identify the onset and offset of M/EEG effects (e.g., ERPs,
-decoding performance), we developed a model-based approach for
-estimating the onset and offset of such effects. To achieve this, we
-leveraged Bayesian generalised additive multilevel models (BGAMMs)
-fitted in \texttt{R} via the \texttt{brms} package and compared the
-performance of this approach to conventional methods on both simulated
-and actual M/EEG data.
+Cluster-based permutation tests are widely used in M/EEG research to
+identify statistically significant effects across time and space.
+However, these methods have notable limitations, particularly in
+accurately determining the precise onset and offset of neural effects.
+To address these limitations, we developed a model-based approach
+relying on Bayesian generalised additive multilevel models implemented
+in \texttt{R} via the \texttt{brms} package
+(\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018}).
+We evaluated the performance of this approach against conventional
+methods using both simulated and actual M/EEG data. Our findings
+demonstrate that this method provides more precise and reliable
+estimates of effect onsets and offsets, offering a valuable tool for
+researchers seeking to enhance the temporal resolution of their
+analyses.
 
@@
 true onset) with 50 trials per participant and condition
-(Figure~\ref{fig-eeg}).
+(Figure~\ref{fig-eeg}). All figures and simulation results can be
+reproduced using the \texttt{R} code available on GitHub:
+\url{https://github.com/lnalborczyk/brms_meeg}.
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-eeg-1.png}
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-eeg-1.png}
 
@@
 (\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018};
-\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}). We used the
-default priors in \texttt{brms} (i.e., weakly informative priors). We
-ran eight Markov Chain Monte-Carlo (MCMC) to approximate the posterior
-distribution, including each 5000 iterations and a warmup of 2000
-iterations, yielding a total of \(8 \times (5000-2000) = 24000\)
-posterior samples to use for inference. Posterior convergence was
-assessed examining trace plots as well as the Gelman--Rubin statistic
-\(\hat{R}\) (\citeproc{ref-gabry2019}{Gabry et al., 2019};
+\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}) and default
+priors (i.e., weakly informative priors). We ran eight Markov Chain
+Monte-Carlo (MCMC) to approximate the posterior distribution, including
+each 5000 iterations and a warmup of 2000 iterations, yielding a total
+of \(8 \times (5000-2000) = 24000\) posterior samples to use for
+inference. Posterior convergence was assessed examining trace plots as
+well as the Gelman--Rubin statistic \(\hat{R}\)
+(\citeproc{ref-gabry2019}{Gabry et al., 2019};
 \citeproc{ref-gelman2020}{Gelman et al., 2020}). The \texttt{brms}
 package uses the same syntax as the \texttt{R} package \texttt{mgcv} v
-1.9-1 (\citeproc{ref-mgcv}{Wood, 2017b}) for specifying smooth effects.
+1.9-3 (\citeproc{ref-mgcv}{Wood, 2017b}) for specifying smooth effects.
 Figure~\ref{fig-plot-post-slope} shows the predictions of this model
@@
 \begin{Highlighting}[]
-\CommentTok{\# averaging across participants}
+\CommentTok{\# computing the average EEG signals for each participant}
 \NormalTok{ppt\_df }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
     \FunctionTok{group\_by}\NormalTok{(participant, condition, time) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{summarise}\NormalTok{(}\AttributeTok{eeg =} \FunctionTok{mean}\NormalTok{(eeg) ) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{summarise}\NormalTok{(}\AttributeTok{eeg\_mean =} \FunctionTok{mean}\NormalTok{(eeg) ) }\SpecialCharTok{\%\textgreater{}\%}
     \FunctionTok{ungroup}\NormalTok{()}
@@
     \CommentTok{\# thin{-}plate regression splines with k{-}1 basis functions}
-\NormalTok{    eeg }\SpecialCharTok{\textasciitilde{}}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
+\NormalTok{    eeg\_mean }\SpecialCharTok{\textasciitilde{}} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
     \AttributeTok{data =}\NormalTok{ ppt\_df,}
@@
 However, the previous model only included constant (fixed) effects, thus
-not properly accounting for between-participant variability. We next fit
-a multilevel version of the BGAM (BGAMM, for an introduction to Bayesian
-multilevel models in \texttt{brms}, see
-\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}). Although it is
-possible to fit a BGAMM at the single-trial level, we present a
-computationally-lighter version of the model that is fitted directly on
-by-participant summary statistics (mean and SD), similar to what is done
-in meta-analysis.
+not properly accounting for between-participant variability. We next
+fitted a multilevel version of the BGAM (BGAMM, for an introduction to
+Bayesian multilevel models in \texttt{brms}, see
+\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}) including a
+varying intercept and slope for participant (but only with a constant
+smoother). Although it is possible to fit a BGAMM using data at the
+single-trial level, we present a computationally lighter version of the
+model that is fitted directly on by-participant summary statistics (mean
+and SD), similar to what is done in meta-analysis.
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# averaging across participants}
+\CommentTok{\# computing the mean and SD of EEG signals for each participant}
 \NormalTok{summary\_df }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{summarise}\NormalTok{(}
-        \AttributeTok{eeg\_mean =} \FunctionTok{mean}\NormalTok{(eeg),}
-        \AttributeTok{eeg\_sd =} \FunctionTok{sd}\NormalTok{(eeg),}
-        \AttributeTok{.by =} \FunctionTok{c}\NormalTok{(participant, condition, time)}
-\NormalTok{        )}
+    \FunctionTok{group\_by}\NormalTok{(participant, condition, time) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{summarise}\NormalTok{(}\AttributeTok{eeg\_mean =} \FunctionTok{mean}\NormalTok{(eeg), }\AttributeTok{eeg\_sd =} \FunctionTok{sd}\NormalTok{(eeg) ) }\SpecialCharTok{\%\textgreater{}\%}
+    \FunctionTok{ungroup}\NormalTok{()}
 
@@
 \NormalTok{meta\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-    \CommentTok{\# using by{-}participant SD of ERPs across trials}
+    \CommentTok{\# using the by{-}participant mean and SD of ERPs}
 \NormalTok{    eeg\_mean }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
-\NormalTok{        condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition) }\SpecialCharTok{+}
-\NormalTok{        (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ participant),}
+        \DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition }\SpecialCharTok{|}\NormalTok{ participant) }\SpecialCharTok{+}
+        \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
     \AttributeTok{data =}\NormalTok{ summary\_df,}
@@
 \NormalTok{    )}
-
-\CommentTok{\# fitting the BGAMM}
-\CommentTok{\# meta\_gamm \textless{}{-} brm(}
-\CommentTok{\#     \# using by{-}participant SD of ERPs across trials}
-\CommentTok{\#     eeg\_mean | se(eeg\_sd) \textasciitilde{} condition +}
-\CommentTok{\#         s(participant, bs = "re") +}
-\CommentTok{\#         s(time, participant, bs = "fs", by = condition, m = 1),}
-\CommentTok{\#     data = summary\_df,}
-\CommentTok{\#     family = gaussian(),}
-\CommentTok{\#     warmup = 2000,}
-\CommentTok{\#     iter = 5000,}
-\CommentTok{\#     chains = 8,}
-\CommentTok{\#     cores = 8,}
-\CommentTok{\#     control = list(adapt\_delta = 0.95),}
-\CommentTok{\#     file = "models/meta\_gamm.rds"}
-\CommentTok{\#     )}
 \end{Highlighting}
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-plot-post-slope-1.png}
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-plot-post-slope-1.png}
 
@@
 \texttt{condition} being above \(0\) (Figure~\ref{fig-post-prob-ratio},
-left). From this quantity, we then compute the ratio of posterior
+left). From this quantity, we compute the ratio of posterior
 probabilities (i.e., \(p/(1-p)\)) and visualise the timecourse of this
@@
 than the probability of the difference not being above 0, given the
-data, the priors, and other model's assumptions. Thresholding the
-posterior probability ratio thus provides a model-based approach for
-estimating the onset and offset of M/EEG effects. An important advantage
-is that the proposed approach can be extended to virtually any model
-structure. Moreover, the approach provides both group-level and
-individual-level onset and offset estimates of M/EEG effects.
+data, the priors, and other model's assumptions.\footnote{This posterior
+  probability ratio, or posterior odds, is equivalent to a Bayes factor,
+  assuming 1:1 prior odds.} Thresholding the posterior probability ratio
+thus provides a model-based approach for estimating the onset and offset
+of M/EEG effects, whose properties will be assessed in the simulation
+study. An important advantage is that the proposed approach can be
+extended to virtually any model structure.
 
@@
 minimum and maximum possible ratio values are determined (bounded) by
-the number of available posterior samples in the model.}
+the number of posterior samples in the model.}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-post-prob-ratio-1.png}
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-post-prob-ratio-1.png}
 
@@
 
-We then assess the performance of the proposed approach by computing the
-difference between the true and estimated onset/offset of the EEG
-difference according to various \texttt{k} (BGAM basis dimension) and
+We then assessed the performance of the proposed approach by computing
+the difference between the true and estimated onset/offset of the EEG
+difference according to various \texttt{k} (basis dimension) and
 \texttt{threshold} values. Remember that the EEG signal was generated
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-error-1.png}
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-onset-error-1.png}
 
@@
 
-\subsection{Comparing the onsets/offsets estimates from other
-approaches}\label{comparing-the-onsetsoffsets-estimates-from-other-approaches}
-
-We then compared the ability of the BGAMM to accurately estimate the
+\subsection{Comparing the onsets/offsets estimates across
+approaches}\label{comparing-the-onsetsoffsets-estimates-across-approaches}
+
+We then compared the ability of the BGAM to accurately estimate the
 onset and offset of the ERP difference to other widely-used methods.
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-corrections-1.png}
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-corrections-1.png}
 
@@
 To assess the accuracy of group-level onset estimation, the various
-methods were compared using the bias (i.e., median(estimated-true)),
-median absolute error (MAE), root mean square error (RMSE), variance,
-and median absolute deviation (MAD) of onset/offset estimates computed
-on 1,000 simulated datasets. As in Rousselet
+methods were compared using the bias (defined as the median difference
+between the estimated and true value of the onset/offset), median
+absolute error (MAE), root mean square error (RMSE), variance, and
+median absolute deviation (MAD) of onset/offset estimates computed on
+10,000 simulated datasets. As in Rousselet
 (\citeproc{ref-rousselet_using_2025}{2025}), each participant was
-assigned a random onset between 150 and 170ms. Whereas the present
-article focuses on one-dimensional signas (e.g., one M/EEG channel), we
-provice a 2D application in Section~\ref{sec-2D}.
+assigned a random onset between 150 and 170ms. Whereas the focus of the
+present article focuses is on one-dimensional signals (e.g., one M/EEG
+channel), we provide a 2D application in Section~\ref{sec-2D}.
 
@@
 
-To complement the simulation study, we evaluated the performance of the
-various methods on actual MEG data (decoding results from
+To complement the simulation study, we evaluated the performance of all
+methods on actual MEG data (decoding results from
 \citeproc{ref-nalborczyk:inprep}{Nalborczyk et al., in preparation}). In
@@
 
-\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-decoding-data-1.png}
+\includegraphics[width=0.75\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-decoding-data-1.png}
 
@@
 results, revealing that the proposed approach (\texttt{BGAM}) has the
-lowest median absolute error (MAE) and variance for both the onset and
-offset estimates. The \texttt{Cluster\ mass} and \texttt{Change\ point}
-also have good performance, but surprisingly, the \texttt{TFCE} method
-has relatively bad performance for estimating the effect offset (similar
-performance to the \texttt{Holm} and \texttt{FDR\ BY01} methods).
-Unsurprisingly, the \texttt{Raw\ p-value} and \texttt{FDR\ BH95} methods
-show the worst performance.
+lowest median error for both the onset and offset estimates. The
+\texttt{Cluster\ mass} and \texttt{Change\ point} also have good
+performance, but surprisingly, the \texttt{TFCE} method has relatively
+bad performance for estimating the effect offset (similar performance to
+the \texttt{Holm} and \texttt{FDR\ BY01} methods). Unsurprisingly, the
+\texttt{FDR\ BH95} and \texttt{Raw\ p-value} methods show the worst
+performance.
 
@@
 
-\caption{\label{fig-simulation-results}Median absolute error and
-variance of onset and offset estimates for each method. Variance is
-plotted on a log10 scale for visual purposes.}
+\caption{\label{fig-simulation-results}Median error and median absolute
+deviation of the error for the onset (left) and offset (right) estimates
+according to each method. Methods are ordered from lowest (top) to
+highest (bottom) median absolute error (separately for the onset and
+offset estimates).}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-simulation-results-1.png}
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-simulation-results-1.png}
 
@@
 Results are further summarised in Table~\ref{tbl-simulation-results},
-which shows that the \texttt{BGAM} is perfectly unbiased (i.e., it has a
-bias of 0s) for the onset and almost exactly unbiased for the onset
-(with a bias of approximately -2ms). The \texttt{Bias} column shows that
-all other methods tend to estimate the onset later than the true onset
-and to estimate the offet earlier than the true offset. As can be seen
-from this table, the \texttt{BGAM} has the best performance on all
-included metrics.
+which shows that the \texttt{BGAM} is almost perfectly unbiased (i.e.,
+it has a bias of approximately 2ms for the onset and 4ms for the
+offset). The \texttt{Bias} column shows that all methods tend to
+estimate the onset later than the true onset and to estimate the offset
+earlier than the true offset. As can be seen from this table, the
+\texttt{BGAM} has the best performance on all included metrics.
 
@@
 {\caption{{Summary statistics of the onset and offset estimates for each
-method (ordered by the absolute value of the
-bias).}{\label{tbl-simulation-results}}}}
+method (in ms, ordered by the MAE) according to the main simulation
+study.}{\label{tbl-simulation-results}}}}
 
@@
 \midrule\addlinespace[2.5pt]
-BGAM (new method) & 0.0000 & 0.0020 & 0.0001 & 0.0000 & 0.0030 \\ 
-Cluster mass & 0.0060 & 0.0060 & 0.0057 & 0.0000 & 0.0044 \\ 
-Change point & 0.0060 & 0.0060 & 0.0051 & 0.0000 & 0.0059 \\ 
-Raw p-value & 0.0060 & 0.1250 & 0.0711 & 0.0252 & 0.1942 \\ 
-FDR BH95 & 0.0080 & 0.0460 & 0.0599 & 0.0192 & 0.0801 \\ 
-FDR BY01 & 0.0120 & 0.0120 & 0.0443 & 0.0087 & 0.0059 \\ 
-TFCE & 0.0140 & 0.0140 & 0.0270 & 0.0019 & 0.0059 \\ 
-Holm & 0.0140 & 0.0140 & 0.0270 & 0.0019 & 0.0059 \\ 
+BGAM (new method) & 2.00 & 2.00 & 1.03 & 43.83 & 2.97 \\ 
+Cluster mass & 6.00 & 6.00 & 6.78 & 228.16 & 5.93 \\ 
+Change point & 6.00 & 6.00 & 5.21 & 34.54 & 5.93 \\ 
+FDR BY01 & 12.00 & 14.00 & 35.44 & 7,659.08 & 5.93 \\ 
+TFCE & 16.00 & 16.00 & 29.88 & 2,785.39 & 5.93 \\ 
+Holm & 16.00 & 16.00 & 29.91 & 2,710.69 & 5.93 \\ 
+FDR BH95 & 8.00 & 48.00 & 57.50 & 18,986.98 & 83.03 \\ 
+Raw p-value & 6.00 & 118.00 & 72.39 & 25,072.78 & 183.84 \\ 
 \midrule\addlinespace[2.5pt]
@@
 \midrule\addlinespace[2.5pt]
-BGAM (new method) & -0.0020 & 0.0040 & 0.0021 & 0.0000 & 0.0030 \\ 
-Cluster mass & -0.0080 & 0.0080 & 0.0084 & 0.0000 & 0.0059 \\ 
-Raw p-value & -0.0080 & 0.1180 & 0.0725 & 0.0252 & 0.1868 \\ 
-Change point & -0.0100 & 0.0100 & 0.0096 & 0.0000 & 0.0059 \\ 
-FDR BH95 & -0.0100 & 0.0360 & 0.0578 & 0.0189 & 0.0682 \\ 
-FDR BY01 & -0.0140 & 0.0150 & 0.0204 & 0.0056 & 0.0059 \\ 
-TFCE & -0.0180 & 0.0180 & 0.0261 & 0.0016 & 0.0030 \\ 
-Holm & -0.0180 & 0.0180 & 0.0262 & 0.0016 & 0.0030 \\ 
+BGAM (new method) & -4.00 & 4.00 & 3.00 & 45.58 & 2.97 \\ 
+Cluster mass & -8.00 & 8.00 & 8.60 & 200.72 & 5.93 \\ 
+Change point & -10.00 & 10.00 & 9.22 & 34.58 & 5.93 \\ 
+FDR BY01 & -14.00 & 16.00 & 37.88 & 7,708.00 & 5.93 \\ 
+TFCE & -18.00 & 18.00 & 32.15 & 2,822.35 & 5.93 \\ 
+Holm & -18.00 & 18.00 & 31.96 & 2,720.13 & 5.93 \\ 
+FDR BH95 & -10.00 & 46.00 & 60.10 & 19,025.38 & 83.03 \\ 
+Raw p-value & -8.00 & 112.00 & 71.59 & 24,969.31 & 177.91 \\ 
 \bottomrule
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-onset-offset-1.png}
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-onset-offset-1.png}
 
@@
 (0ms difference for the onset estimate and 5ms difference for the offset
-estimate). The \texttt{Raw\ p-value} method has similar performance, but
-given the aberrant estimates it produces (cf.
+estimate). The \texttt{Raw\ p-value} method has similar performance for
+the offset, but given the aberrant estimates it produces (cf.
 Figure~\ref{fig-onset-offset}), the fact that it is consistent between
@@
 for all other methods, split datasets produce later onset estimates and
-earlier offset estimates (as compared to the estimates from the model
+earlier offset estimates (compared to the estimates from the model
 fitted on the full dataset). These results highlight some desirable
@@
 asymptotic properties on simulated data, ii) provide sensible identified
-clusters in actual data, and iii) provide reliable/stable estimates on
-actual data.
+clusters in actual data, and iii) provide reliable/stable estimates
+across various partitions of the data. All these desiderata seem to be
+fulfilled by the BGAM approach.
 
@@
 deviation of the error for the onset (left) and offset (right) estimates
-according to each method. Methods are ordered from lowest (top) to
-highest (bottom) median absolute error (separately for the onset and
-offset estimates).}
+according to each method. Methods are ordered from smallest (top) to
+largest (bottom) median absolute error separately for the onset and
+offset estimates.}
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-reliability-1.png}
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-reliability-1.png}
 
@@
 {\caption{{Summary statistics of the onset and offset estimates for each
-method (ordered by the absolute value of the
-bias).}{\label{tbl-reliability-results}}}}
+method (in ms, ordered by the MAE) according to the reliability
+simulation study.}{\label{tbl-reliability-results}}}}
 
@@
 \midrule\addlinespace[2.5pt]
-BGAM (new method) & 0.0000 & 0.0000 & 0.0106 & 0.0008 & 0.0000 \\ 
-Cluster mass & 0.0000 & 0.0000 & 0.0019 & 0.0000 & 0.0000 \\ 
-Change point & 0.0050 & 0.0050 & 0.0011 & 0.0001 & 0.0074 \\ 
-TFCE & 0.0100 & 0.0100 & 0.0120 & 0.0001 & 0.0074 \\ 
-FDR BY01 & 0.0100 & 0.0100 & 0.0065 & 0.0023 & 0.0000 \\ 
-Holm & 0.0100 & 0.0100 & 0.0118 & 0.0001 & 0.0074 \\ 
-FDR BH95 & 0.0250 & 0.0900 & 0.0101 & 0.0076 & 0.0964 \\ 
-Raw p-value & -0.0300 & 0.0650 & 0.0201 & 0.0062 & 0.0815 \\ 
+BGAM (new method) & 0.00 & 0.00 & 6.91 & 829.66 & 0.00 \\ 
+Cluster mass & 0.00 & 0.00 & 2.07 & 8.33 & 0.00 \\ 
+TFCE & 5.00 & 5.00 & 10.67 & 101.05 & 7.41 \\ 
+Holm & 5.00 & 5.00 & 10.04 & 96.64 & 7.41 \\ 
+Change point & 5.00 & 5.00 & 1.74 & 96.94 & 7.41 \\ 
+FDR BY01 & 10.00 & 10.00 & 3.71 & 1,581.49 & 0.00 \\ 
+FDR BH95 & 10.00 & 90.00 & 6.30 & 6,970.82 & 118.61 \\ 
+Raw p-value & -40.00 & 60.00 & 27.50 & 5,545.80 & 74.13 \\ 
 \midrule\addlinespace[2.5pt]
@@
 \midrule\addlinespace[2.5pt]
-Raw p-value & 0.0000 & 0.0300 & 0.0256 & 0.0047 & 0.0445 \\ 
-BGAM (new method) & -0.0050 & 0.0750 & 0.0716 & 0.0274 & 0.1038 \\ 
-Change point & 0.0250 & 0.0650 & 0.0566 & 0.0090 & 0.1186 \\ 
-FDR BH95 & -0.0750 & 0.0750 & 0.1053 & 0.0153 & 0.1483 \\ 
-TFCE & -0.1150 & 0.1450 & 0.1094 & 0.0146 & 0.1260 \\ 
-Holm & -0.1150 & 0.1450 & 0.1091 & 0.0133 & 0.1186 \\ 
-FDR BY01 & -0.1450 & 0.1450 & 0.1197 & 0.0154 & 0.0890 \\ 
-Cluster mass & -0.1650 & 0.1650 & 0.1448 & 0.0051 & 0.0297 \\ 
+Raw p-value & 0.00 & 35.00 & 39.65 & 7,386.26 & 51.89 \\ 
+BGAM (new method) & -15.00 & 60.00 & 38.07 & 23,511.47 & 66.72 \\ 
+Change point & 20.00 & 60.00 & 54.11 & 8,354.85 & 81.54 \\ 
+FDR BH95 & -75.00 & 75.00 & 111.46 & 15,361.69 & 140.85 \\ 
+FDR BY01 & -145.00 & 150.00 & 122.62 & 17,421.16 & 88.96 \\ 
+Holm & -160.00 & 160.00 & 123.00 & 13,560.13 & 126.02 \\ 
+Cluster mass & -165.00 & 165.00 & 150.79 & 4,272.92 & 29.65 \\ 
+TFCE & -180.00 & 180.00 & 131.70 & 13,720.28 & 96.37 \\ 
 \bottomrule
@@
 varying/random effects (intercept and slope) for item (e.g.,
-word)\ldots{} but also continuous predictors at the trial level?
+word)\ldots{} but also continuous predictors at the trial level,
+overlapping factors, etc\ldots{}
 
@@
 following R packages: assertthat v. 0.2.1
-(\citeproc{ref-assertthat}{Wickham, 2019}), brms v. 2.22.0
+(\citeproc{ref-assertthat}{Wickham, 2019}), bamlss v. 1.2.5
+(\citeproc{ref-bamlss2024}{Muschinski et al., 2024};
+\citeproc{ref-bamlss2018}{Umlauf et al., 2018b},
+\citeproc{ref-bamlss2021}{2021}), brms v. 2.22.0
 (\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018},
@@
 furrr v. 0.3.1 (\citeproc{ref-furrr}{Vaughan \& Dancho, 2022}), future
-v. 1.34.0 (\citeproc{ref-RJ-2021-048}{Bengtsson, 2021}), ggrepel v.
-0.9.6 (\citeproc{ref-ggrepel}{Slowikowski, 2024}), glue v. 1.8.0
-(\citeproc{ref-glue}{Hester \& Bryan, 2024}), grateful v. 0.2.11
+v. 1.49.0 (\citeproc{ref-future}{Bengtsson, 2021}), ggrepel v. 0.9.6
+(\citeproc{ref-ggrepel}{Slowikowski, 2024}), glue v. 1.8.0
+(\citeproc{ref-glue}{Hester \& Bryan, 2024}), grateful v. 0.2.12
 (\citeproc{ref-grateful}{Rodriguez-Sanchez \& Jackson, 2024}), gt v.
@@
 \citeproc{ref-knitr2025}{2025}), MetBrewer v. 0.2.0
-(\citeproc{ref-MetBrewer}{Mills, 2022}), neurogam v. 0.0.1
+(\citeproc{ref-MetBrewer}{Mills, 2022}), mgcv v. 1.9.3
+(\citeproc{ref-mgcv2003}{Wood, 2003b}, \citeproc{ref-mgcv2004}{2004},
+\citeproc{ref-mgcv2011}{2011}, \citeproc{ref-mgcv2017}{2017c};
+\citeproc{ref-mgcv2016}{Wood et al., 2016}), neurogam v. 0.0.1
 (\citeproc{ref-neurogam}{Nalborczyk, 2025}), pakret v. 0.2.2
@@
 \citeproc{ref-rmarkdown2018}{Xie et al., 2018},
-\citeproc{ref-rmarkdown2020}{2020}), scales v. 1.3.0
-(\citeproc{ref-scales}{Wickham et al., 2023}), scico v. 1.5.0
+\citeproc{ref-rmarkdown2020}{2020}), scales v. 1.4.0
+(\citeproc{ref-scales}{Wickham et al., 2025}), scico v. 1.5.0
 (\citeproc{ref-scico}{T. L. Pedersen \& Crameri, 2023}), tictoc v. 1.2.1
@@
 
-\bibitem[\citeproctext]{ref-RJ-2021-048}
+\bibitem[\citeproctext]{ref-future}
 Bengtsson, H. (2021). A unifying framework for parallel and distributed
@@
 
-\bibitem[\citeproctext]{ref-dimigen2021}
-Dimigen, O., \& Ehinger, B. V. (2021). Regression-based analysis of
-combined EEG and eye-tracking data: Theory and applications.
-\emph{Journal of Vision}, \emph{21}(1), 3.
-\url{https://doi.org/10.1167/jov.21.1.3}
-
 \bibitem[\citeproctext]{ref-dinga2021}
@@
 
-\bibitem[\citeproctext]{ref-dunagan2024}
+\bibitem[\citeproctext]{ref-dunagan2025}
 Dunagan, D., Jordan, T., Hale, J. T., Pylkkänen, L., \& Chacón, D. A.
-(2024). \emph{Evaluating the timecourses of morpho-orthographic,
-lexical, and grammatical processing following rapid parallel visual
-presentation: An EEG investigation in english}.
-\url{http://dx.doi.org/10.1101/2024.04.10.588861}
+(2025). Evaluating the timecourses of morpho-orthographic, lexical, and
+grammatical processing following rapid parallel visual presentation: An
+EEG investigation in English. \emph{Cognition}, \emph{257}, 106080.
+\url{https://doi.org/10.1016/j.cognition.2025.106080}
 
@@
 \url{https://doi.org/10.7717/peerj.7838}
-
-\bibitem[\citeproctext]{ref-fischer2013}
-Fischer, Adrian~G., \& Ullsperger, M. (2013). Real and Fictive Outcomes
-Are Processed Differently but Converge on a Common Adaptive Mechanism.
-\emph{Neuron}, \emph{79}(6), 1243--1255.
-\url{https://doi.org/10.1016/j.neuron.2013.07.006}
 
@@
 
-\bibitem[\citeproctext]{ref-hauk2006}
-Hauk, O., Davis, M. H., Ford, M., Pulvermüller, F., \& Marslen-Wilson,
-W. D. (2006). The time course of visual word recognition as revealed by
-linear regression analysis of ERP data. \emph{NeuroImage}, \emph{30}(4),
-1383--1400. \url{https://doi.org/10.1016/j.neuroimage.2005.11.048}
-
-\bibitem[\citeproctext]{ref-hendrix2017}
-Hendrix, P., Bolger, P., \& Baayen, H. (2017). Distinct ERP signatures
-of word frequency, phrase frequency, and prototypicality in speech
-production. \emph{Journal of Experimental Psychology: Learning, Memory,
-and Cognition}, \emph{43}(1), 128--149.
-\url{https://doi.org/10.1037/a0040332}
-
 \bibitem[\citeproctext]{ref-glue}
@@
 
-\bibitem[\citeproctext]{ref-kryuchkova2012}
-Kryuchkova, T., Tucker, B. V., Wurm, L. H., \& Baayen, R. H. (2012).
-Danger and usefulness are detected early in auditory lexical processing:
-Evidence from electroencephalography. \emph{Brain and Language},
-\emph{122}(2), 81--91. \url{https://doi.org/10.1016/j.bandl.2012.05.005}
-
 \bibitem[\citeproctext]{ref-easystats}
@@
 \url{https://CRAN.R-project.org/package=MetBrewer}
+
+\bibitem[\citeproctext]{ref-bamlss2024}
+Muschinski, T., Mayr, G. J., Simon, T., Umlauf, N., \& Zeileis, A.
+(2024). Cholesky-based multivariate {G}aussian regression.
+\emph{Econometrics and Statistics}, \emph{29}, 261--281.
+\url{https://doi.org/10.1016/j.ecosta.2022.03.001}
 
@@
 
-\bibitem[\citeproctext]{ref-pernet2022}
-Pernet, C. R. (2022). Electroencephalography robust statistical linear
-modelling using a single weight per trial. \emph{Aperture Neuro},
-\emph{2}, 1--22.
-\url{https://doi.org/10.52294/apertureneuro.2022.2.seoo9435}
-
 \bibitem[\citeproctext]{ref-pernet2011}
@@
 
-\bibitem[\citeproctext]{ref-suxf3skuthy2017}
-Sóskuthy, M. (2017). \emph{Generalised additive mixed models for dynamic
-analysis in linguistics: A practical introduction}.
-\url{https://doi.org/10.48550/ARXIV.1703.05339}
-
 \bibitem[\citeproctext]{ref-suxf3skuthy2021}
@@
 \emph{84}, 101017. \url{https://doi.org/10.1016/j.wocn.2020.101017}
-
-\bibitem[\citeproctext]{ref-teichmann2022}
-Teichmann, L. (2022). An empirically driven guide on using bayes factors
-for m/EEG decoding. \emph{Aperture Neuro}, \emph{2}, 1--10.
-\url{https://doi.org/10.52294/apertureneuro.2022.2.maoc6465}
 
@@
 
+\bibitem[\citeproctext]{ref-bamlss2021}
+Umlauf, N., Klein, N., Simon, T., \& Zeileis, A. (2021). {bamlss}: A
+{L}ego toolbox for flexible {B}ayesian regression (and beyond).
+\emph{Journal of Statistical Software}, \emph{100}(4), 1--53.
+\url{https://doi.org/10.18637/jss.v100.i04}
+
 \bibitem[\citeproctext]{ref-umlauf2018}
-Umlauf, N., Klein, N., \& Zeileis, A. (2018). BAMLSS: Bayesian Additive
+Umlauf, N., Klein, N., \& Zeileis, A. (2018a). BAMLSS: Bayesian Additive
 Models for Location, Scale, and Shape (and Beyond). \emph{Journal of
@@
 \url{https://doi.org/10.1080/10618600.2017.1407325}
+
+\bibitem[\citeproctext]{ref-bamlss2018}
+Umlauf, N., Klein, N., \& Zeileis, A. (2018b). {BAMLSS}: {B}ayesian
+additive models for location, scale and shape (and beyond).
+\emph{Journal of Computational and Graphical Statistics}, \emph{27}(3),
+612--627. \url{https://doi.org/10.1080/10618600.2017.1407325}
 
@@
 \bibitem[\citeproctext]{ref-scales}
-Wickham, H., Pedersen, T. L., \& Seidel, D. (2023). \emph{{scales}:
+Wickham, H., Pedersen, T. L., \& Seidel, D. (2025). \emph{{scales}:
 Scale functions for visualization}.
@@
 
-\bibitem[\citeproctext]{ref-winter2016}
-Winter, B., \& Wieling, M. (2016). How to analyze linguistic change
-using mixed models, Growth Curve Analysis and Generalized Additive
-Modeling. \emph{Journal of Language Evolution}, \emph{1}(1), 7--18.
-\url{https://doi.org/10.1093/jole/lzv003}
-
 \bibitem[\citeproctext]{ref-wood2003}
-Wood, S. N. (2003). Thin Plate Regression Splines. \emph{Journal of the
+Wood, S. N. (2003a). Thin Plate Regression Splines. \emph{Journal of the
 Royal Statistical Society Series B: Statistical Methodology},
 \emph{65}(1), 95--114. \url{https://doi.org/10.1111/1467-9868.00374}
+
+\bibitem[\citeproctext]{ref-mgcv2003}
+Wood, S. N. (2003b). Thin-plate regression splines. \emph{Journal of the
+Royal Statistical Society (B)}, \emph{65}(1), 95--114.
+\url{https://doi.org/10.1111/1467-9868.00374}
+
+\bibitem[\citeproctext]{ref-mgcv2004}
+Wood, S. N. (2004). Stable and efficient multiple smoothing parameter
+estimation for generalized additive models. \emph{Journal of the
+American Statistical Association}, \emph{99}(467), 673--686.
+\url{https://doi.org/10.1198/016214504000000980}
+
+\bibitem[\citeproctext]{ref-mgcv2011}
+Wood, S. N. (2011). Fast stable restricted maximum likelihood and
+marginal likelihood estimation of semiparametric generalized linear
+models. \emph{Journal of the Royal Statistical Society (B)},
+\emph{73}(1), 3--36.
+\url{https://doi.org/10.1111/j.1467-9868.2010.00749.x}
 
@@
 with r} (2nd ed.). Chapman; Hall/CRC.
+
+\bibitem[\citeproctext]{ref-mgcv2017}
+Wood, S. N. (2017c). \emph{Generalized {A}dditive {M}odels: An
+introduction with {R}} (2nd ed.). Chapman; Hall/CRC.
+
+\bibitem[\citeproctext]{ref-mgcv2016}
+Wood, S. N., Pya, N., \& Säfken, B. (2016). Smoothing parameter and
+model selection for general smooth models (with discussion).
+\emph{Journal of the American Statistical Association}, \emph{111},
+1548--1575. \url{https://doi.org/10.1080/01621459.2016.1180986}
 
@@
 
-\includegraphics[width=0.75\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-sim-timegen-1.png}
+\includegraphics[width=0.75\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-sim-timegen-1.png}
 
@@
 \(R \times R \rightarrow R\). A popular bivariate basis function uses
-\emph{thin-plate splines} (\citeproc{ref-wood2003}{Wood, 2003}), which
+\emph{thin-plate splines} (\citeproc{ref-wood2003}{Wood, 2003a}), which
 extend to \(\mathbf{s}_{i} \in \mathbb{R}^{d}\) and \(\partial l_{g}\)
@@
 \(\mathbf{k}_{m}\). We fitted this model using \texttt{brms}\ldots{}
+using the \texttt{t2()} smooth constructor with full penalties
+(\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al.,
+2019})\ldots{}
 
@@
 \NormalTok{    )}
+
+\CommentTok{\# fitting a GAM with two temporal dimensions to empirical data}
+\NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
+    \CommentTok{\# 2D thin{-}plate spline (tp)}
+    \CommentTok{\# auc \textasciitilde{} t2(train\_time, test\_time, bs = "tp", k = 30),}
+\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{full =} \ConstantTok{TRUE}\NormalTok{),}
+    \AttributeTok{data =}\NormalTok{ timegen\_data,}
+    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
+    \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
+    \AttributeTok{iter =} \DecValTok{3000}\NormalTok{,}
+    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
+    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
+    \CommentTok{\# file = "models/timegen\_gam\_meg\_t2\_k30.rds"}
+    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_meg\_t2\_k20\_full.rds"}
+\NormalTok{    )}
+
+\CommentTok{\# retrieving the data}
+\CommentTok{\# timegen\_data \textless{}{-} timegen\_gam$data}
+
+\CommentTok{\# PPCs}
+\CommentTok{\# pp\_check(timegen\_gam)}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{figure}[H]
-
-\caption{Simulated data with thresholds derived from the BGAM (left),
-predicted AUC values (middle), and ratios of posterior probabilies of
-decoding accuracy being above chance level (right) according to the
-bivariate BGAM (for k=10, top row, and k=20, bottom row).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/gam-timegen-post-preds-1.png}
-
-}
-
-\end{figure}%
 
@@
 
-Here provide recommendation about how to define \texttt{k}. An option is
-to vary \texttt{k} and examine the predictions and posterior predictive
-checks (PPCs) of each model\ldots{} In this example
-(Figure~\ref{fig-choose-k})\ldots{} However, it is not possible to
-provide general recommendations, as the optimal \texttt{k} depends on
-the sampling rate, the preprocessing steps (e.g., signal-to-noise ratio,
-low-pass filtering, etc), and the neural dynamics of the phenomenon
-under study.
+Here we should provide recommendation about how to define
+\texttt{k}\ldots{} An option is to vary \texttt{k} and examine the
+predictions and posterior predictive checks (PPCs) of each model\ldots{}
+In this example (Figure~\ref{fig-choose-k})\ldots{} However, it is not
+possible to provide general recommendations, as the optimal \texttt{k}
+depends on the sampling rate, the preprocessing steps (e.g.,
+signal-to-noise ratio, low-pass filtering, etc), and the neural dynamics
+of the phenomenon under study. Choose \texttt{k} as large a possible
+(given limits on computation time etc)? Or use LOOIC/WAIC (seem to
+select too simple models)?
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{brms_meeg_files/figure-pdf/fig-choose-k-1.png}
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-choose-k-1.png}
 
@@
 pipeline is to import ERPs or decoding results in \texttt{R} and analyse
-these data using the code provided in the main paper. However, it is
-also possible to call functions from the \texttt{neurogam\ R} package
-(available at \url{https://github.com/lnalborczyk/neurogam}), which come
+these data using the code provided in the main paper. It is also
+possible to call functions from the \texttt{neurogam\ R} package
+(available at \url{https://github.com/lnalborczyk/neurogam}) which come
 with sensible defaults.

### 2025-05-22 code/edf_estimation.R

@@ -1,0 +1,113 @@
+###########################################################
+# Comparing various EDF estimates                         #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on May 22, 2025                            #
+###########################################################
+
+library(tidyverse)
+library(bamlss)
+library(brms)
+
+# importing the reshaped MEG decoding results
+decoding_df <- read.csv(file = "data/decoding_results_reshaped.csv") %>%
+    # removing some participants
+    mutate(participant_id = cur_group_id(), .by = participant) %>%
+    filter(participant_id < 33) %>%
+    select(-participant_id)
+
+# averaging across participants
+decoding_summary_df <- decoding_df %>%
+    summarise(
+        auc_mean = mean(auc),
+        auc_sd = sd(auc),
+        .by = c(participant, time)
+        )
+
+# setting possible k values
+k_values <- seq(from = 10, to = 50, by = 5)
+
+# initialising en empty dataframe to store the results
+results <- data.frame()
+
+# fitting the model for each k
+for (k_value in k_values) {
+    
+    # printing progress
+    cat("\nProcessing k =", k_value, "\n")
+    
+    # fitting the model with bamlss
+    b_model <- bamlss(
+        formula = auc_mean ~ s(time, bs = "tp", k = k_value),
+        family = beta_bamlss,
+        data = decoding_summary_df
+        )
+    
+    # computing the EDF
+    ss <- samplestats(samples = b_model)
+    
+    # constructing the smooth term dynamically
+    smooth_term <- glue::glue("s(time, bs = 'tp', k = {k_value})")
+    
+    # full formula
+    formula_str <- glue::glue("auc_mean ~ {smooth_term}")
+    
+    # convert to formula
+    formula_obj <- brms::bf(formula_str)
+    
+    # fitting the Beta GAM
+    meg_decoding_gam <- brm(
+        formula = formula_obj,
+        data = decoding_summary_df,
+        family = brms::Beta(),
+        warmup = 1000,
+        iter = 4000,
+        chains = 4,
+        cores = 4
+        )
+    
+    # computing the p_WAIC
+    pwaic <- waic(meg_decoding_gam)
+    looic <- loo(meg_decoding_gam)
+    
+    # combining the estimates in a dataframe
[…]

### 2025-05-22 code/errors_simulation_cluster.R

@@ -2,8 +2,10 @@
 # Monte-Carlo simulation of onset/offset error properties #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on April 9, 2025                           #
+# Last updated on May 9, 2025                             #
 ###########################################################
+
+options(error = function() {traceback(2); quit("no", status = 1, runLast = FALSE)})
 
 # importing R packages
 library(changepoint)
@@ -11,18 +13,20 @@ library(reticulate)
 library(tidybayes)
 library(dplyr)
 library(tidyr)
+library(glue)
 library(brms)
 
 # for setting up the cluster
 library(doParallel)
 library(foreach)
 
-# timing the simulations
-library(tictoc)
-tic()
+# retrieving the array ID
+args <- commandArgs(trailingOnly = TRUE)
+array_id <- as.numeric(args[1])
+cat("Running job for array_id =", array_id, "\n")
 
 # total number of total CPU cores available on the HPC node
-total_cores <- 12
+total_cores <- 16
 
 # number of cores per brms model
 cores_per_model <- 4
@@ -34,20 +38,20 @@ n_parallel <- total_cores / cores_per_model
 cl <- makeCluster(n_parallel)
 registerDoParallel(cl)
 
+# importing home-made helper functions
+source("code/utils.R")
+
 # importing R version of Matlab code from Yeung et al. (2004)
 source("code/eeg_noise.R")
 
 # importing the ERP template with true onset = 160 ms, F = 81, and max at F = 126
 source("code/erp_template.R")
 
-# importing home-made helper functions
-source("code/utils.R")
-
 # to use with the eeg_noise function
-meanpower <- unlist(read.table("code/meanpower.txt") )
+# meanpower <- unlist(read.table("code/meanpower.txt") )
 
 # defining the total number of simulations to be performed
-nsims <- 1000
+n_sims <- 8
 
 # defining simulation parameters
 n_trials <- 50 # number of trials
@@ -67,13 +71,16 @@ significance_thresholds <- c(0.05, 0.01, 0.005, 0.001)
 threshold_values <- c(1, 3, 6, 10, 20, 50, 100)
 
 # initialising empty simulation results
-sim_results <- data.frame()
+# sim_results <- data.frame()
 
 # for each simulation
-sim_results_list <- foreach(
-    i = 1:nsims,
+sim_results <- foreach(
[…]

### 2025-05-22 code/old/model_examples.R

@@ -1,0 +1,97 @@
+# averaging across participants
+# ppt_df <- raw_df %>%
+#     group_by(participant, condition, time) %>%
+#     summarise(eeg = mean(eeg) ) %>%
+#     ungroup()
+
+# defining a contrast for condition
+# contrasts(ppt_df$condition) <- c(-0.5, 0.5)
+
+# fitting the BGAM
+# gam <- brm(
+#     # thin-plate regression splines with k-1 basis functions
+#     eeg ~ condition + s(time, bs = "tp", k = 20, by = condition),
+#     data = ppt_df,
+#     family = gaussian(),
+#     warmup = 2000,
+#     iter = 5000,
+#     chains = 8,
+#     cores = 8,
+#     file = "models/gam.rds"
+#     )
+# 
+# meta_gam <- brm(
+#     # using by-participant SD of ERPs across trials
+#     eeg_mean | se(eeg_sd) ~
+#         condition + s(time, bs = "tp", k = 20, by = condition) +
+#         (1 + condition | participant),
+#     data = summary_df,
+#     family = gaussian(),
+#     warmup = 2000,
+#     iter = 5000,
+#     chains = 8,
+#     cores = 8,
+#     file = "models/meta_gam.rds"
+#     )
+# 
+# full_meta_gamm <- brm(
+#     # using by-participant SD of ERPs across trials
+#     eeg_mean | se(eeg_sd) ~ 1 + condition +
+#         (1 + condition | participant) +
+#         s(time, by = condition, bs = "tp", k = 20) +
+#         # s(participant, by = condition, bs = "re", xt = list(bs = "tp") ),
+#         s(time, participant, by = condition, bs = "fs", xt = list(bs = "tp") ),
+#     data = summary_df,
+#     family = gaussian(),
+#     warmup = 2000,
+#     iter = 5000,
+#     chains = 8,
+#     cores = 8,
+#     control = list(adapt_delta = 0.95),
+#     file = "models/full_meta_gamm.rds"
+#     )
+
+# errors on the HPC cluster
+# Error in unserialize(socklist[[n]]) : error reading from connection
+# Calls: %dopar% ... recvOneData -> recvOneData.SOCKcluster -> unserialize
+# Execution halted
[…]

### 2025-05-22 code/old/split_half_reliability_cluster.R

@@ -1,0 +1,502 @@
+###########################################################
+# Split-half reliability of onset/offset estimates        #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on April 14, 2025                          #
+###########################################################
+
+# importing R packages
+library(changepoint)
+library(reticulate)
+library(tidybayes)
+library(dplyr)
+library(tidyr)
+library(brms)
+
+# for setting up the cluster
+library(doParallel)
+library(foreach)
+
+# timing the simulations
+# library(tictoc)
+# tic()
+
+# total number of total CPU cores available on the HPC node
+total_cores <- 64
+
+# number of cores per brms model
+cores_per_model <- 4
+
+# number of parallel simulations to run at once
+n_parallel <- total_cores / cores_per_model
+
+# registering parallel backend
+cl <- makeCluster(n_parallel)
+registerDoParallel(cl)
+
+# importing home-made helper functions
+source("code/utils.R")
+
+# importing the reshaped MEG decoding results
+raw_df <- read.csv(file = "data/decoding_results_reshaped.csv") |>
+    # removing the last participant (to get a round number of participants)
+    mutate(participant_id = cur_group_id(), .by = participant) |>
+    filter(participant_id < 33) |>
+    select(-participant_id)
+
+# averaging across participants
+summary_df <- raw_df |>
+    summarise(
+        auc_mean = mean(auc),
+        auc_sd = sd(auc),
+        auc_se = sd(auc) / sqrt(n() ),
+        .by = c(participant, time)
+        )
+
+# computing the smallest effect size of interest (SESOI)
+# as the SD of decoding performance during the baseline
+full_sesoi <- summary_df %>%
+    filter(time < 0) %>%
+    summarise(auc_mean = mean(auc_mean), .by = time) %>%
+    summarise(sesoi = sd(auc_mean) ) %>%
+    pull(sesoi)
+
+# getting unique participants
+unique_participants <- unique(raw_df$participant)
+
+# getting the number of participants
+n_participants <- as.numeric(length(unique_participants) )
+
+# defining the number of splits
+# n_splits <- 1000
+n_splits <- 32
+
+# number of posterior samples to use to compute the posterior probability ratio
+# n_post_samples <- 1e4
+
+# creating a dataframe where each participant appears in every split
+split_df <- crossing(participant = unique_participants, split_number = 1:n_splits)
+
[…]

### 2025-05-22 code/split_half_reliability_cluster.R

@@ -2,8 +2,10 @@
 # Split-half reliability of onset/offset estimates        #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on April 11, 2025                          #
+# Last updated on May 8, 2025                             #
 ###########################################################
+
+options(error = function() {traceback(2); quit("no", status = 1, runLast = FALSE)})
 
 # importing R packages
 library(changepoint)
@@ -11,18 +13,20 @@ library(reticulate)
 library(tidybayes)
 library(dplyr)
 library(tidyr)
+library(glue)
 library(brms)
 
 # for setting up the cluster
 library(doParallel)
 library(foreach)
 
-# timing the simulations
-library(tictoc)
-tic()
+# retrieving the array ID
+args <- commandArgs(trailingOnly = TRUE)
+array_id <- as.numeric(args[1])
+cat("Running job for array_id =", array_id, "\n")
 
 # total number of total CPU cores available on the HPC node
-total_cores <- 12
+total_cores <- 16
 
 # number of cores per brms model
 cores_per_model <- 4
@@ -67,8 +71,8 @@ unique_participants <- unique(raw_df$participant)
 # getting the number of participants
 n_participants <- as.numeric(length(unique_participants) )
 
-# defining the number of splits
-n_splits <- 1000
+# defining the total number of splits (per node)
+n_splits <- 50
 
 # number of posterior samples to use to compute the posterior probability ratio
 # n_post_samples <- 1e4
@@ -93,20 +97,25 @@ post_prob_ratio_threshold <- 20
 alpha_level <- 0.05
 
 # for each simulation
-reliability_results <- tryCatch({foreach(
+reliability_results <- foreach(
     i = seq_len(n_splits), .combine = bind_rows,
     .packages = c("reticulate", "brms", "dplyr", "tidyr", "changepoint", "tidybayes")
     ) %dopar% {
+
+	tryCatch({
         
+	    # formatting the current split ID
+	    this_split_id <- sprintf("a%03d_s%03d", array_id, i)
+	    
         # importing Python modules locally
-        message_parallel("\nImporting Python modules...")
-        use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
[…]

### 2025-05-22 code/utils.R

@@ -104,6 +104,14 @@ find_clusters <- function (data, threshold = 10, above_threshold = TRUE) {
             ) |>
         data.frame()
     
+    if (nrow(clusters) == 0) {
+        clusters <- data.frame(
+            cluster_id = NA_integer_,
+            cluster_onset = NA_real_,
+            cluster_offset = NA_real_
+            )
+        }
+    
     return (clusters)
     
 }

### 2025-08-22 manuscript.qmd
pushing revised manuscript

@@
     conflict-of-interest: "The authors have no conflicts of interest to disclose."
-abstract: "Time-resolved electrophysiological measurements such as those obtained through magneto- or electro-encephalography (M/EEG) offer a unique window onto the neural activity underlying cognitive processes.
+abstract: "Time-resolved electrophysiological measurements such as those obtained through magneto- and electroencephalography (M/EEG) offer a unique window onto the neural activity underlying cognitive processes.
 Researchers are often interested in determining whether and when these signals differ across experimental conditions or participant groups.
-The conventional approach involves mass-univariate statistical testing across time and space, followed by corrections for multiple comparisons such as cluster-based inference.
+The conventional approach involves mass-univariate statistical testing across time and space followed by corrections for multiple comparisons or some form of cluster-based inference.
 While effective for controlling error rates at the cluster-level, cluster-based inference comes with a significant limitation: by shifting the focus of inference from individual time points to clusters, it prevents drawing conclusions about the precise onset or offset of observed effects.
-Here, we introduce a *model-based* approach for analysing M/EEG timeseries such as event-related potentials (ERPs) or decoding performance over time.
-Our approach leverages Bayesian generalised additive multilevel models, providing posterior probabilities that an effect is above zero (or above chance) at each time point, while naturally accounting for temporal dependencies and between-subject variability.
-Using both simulated and actual M/EEG datasets, we demonstrate that this approach substantially outperforms conventional methods in estimating the onset and offset of neural effects, yielding more precise and reliable results.
-We provide an R package implementing the method and describe how it can be integrated into M/EEG analysis pipelines using MNE-Python."
-keywords: [EEG, MEG, cluster-based inference, simulation, multiple comparisons, generalised additive models, mixed-effects models, multilevel models, Bayesian statistics, brms]
+Here, we present a *model-based* alternative for analysing M/EEG timeseries, such as event-related potentials or time-resolved decoding accuracy.
+Our approach leverages Bayesian generalised additive multilevel models, providing posterior odds that an effect exceeds zero (or chance) at each time point, while naturally accounting for temporal dependencies and between-subject variability.
+Using both simulated and empirical M/EEG datasets, we show that this approach substantially outperforms conventional methods in estimating the onset and offset of neural effects, yielding more precise and reliable estimates.
+We provide an open-source R package implementing the method and describe how it can be integrated into M/EEG analysis pipelines using MNE-Python."
+keywords: [EEG, MEG, cluster-based inference, multiple comparisons, generalised additive models, mixed-effects models, multilevel models, Bayesian statistics, brms]
 floatsintext: true
 numbered-lines: true
+word-count: true
 bibliography:
@@
 donotrepeattitle: true
-suppress-title-introduction: true
+number-sections: true
+number-depth: 3
 lang: en
@@
     documentmode: doc # man, jou, doc, stu
+    # number_sections: true
     keep-tex: true
@@
 library(ggrepel)
+library(ggpubr)
 library(scales)
@@
     eval = TRUE,
-    echo = TRUE,
+    echo = FALSE,
     warning = FALSE,
@@
 
-# disabling hyperlinks
-# options(cli.hyperlink = FALSE)
-
 # setting a default ggplot2 theme
@@
 ```
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
 
@@
 As pointed out by @maris2007: "there is a conflict between this interest in localized effects and our choice for a global null hypothesis: by controlling the FA [false alarm] rate under this global null hypothesis, one cannot quantify the uncertainty in the spatiotemporal localization of the effect".
-Even worse, @rosenblatt2018 note that cluster-based inference suffers from low spatial resolution: "Since discovering a cluster means that 'there exists at least one voxel with an evoked response in the cluster', and not that 'all the voxels in the cluster have an evoked response', it follows that the larger the detected cluster, the less information we have on the location of the activation." As a consequence, cluster-based inference is expected to perform poorly for localising the onset of M/EEG effects; a property that was later demonstrated in simulation studies [e.g., @rousselet_using_2025; @sassenhagen2019].
+Even worse, @rosenblatt2018 note that cluster-based inference suffers from low spatial resolution: "Since discovering a cluster means that 'there exists at least one voxel with an evoked response in the cluster', and not that 'all the voxels in the cluster have an evoked response', it follows that the larger the detected cluster, the less information we have on the location of the activation." As a consequence, cluster-based inference is expected to perform poorly for identifying the onset of M/EEG effects; a property that was later demonstrated in simulation studies [e.g., @rousselet_using_2025; @sassenhagen2019].
 
 To overcome the limitations of cluster-based inference, we introduce a novel *model-based* approach for precisely localising M/EEG effects in time, space, and other dimensions.
-The proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the posterior probability of effects being above chance at the level of timesteps, sensors, voxels, etc, while naturally taking into account spatiotemporal dependencies present in M/EEG data.
+The proposed approach, based on Bayesian generalised additive multilevel models, allows quantifying the posterior odds of effects being above chance at the level of timesteps, sensors, voxels, etc, while naturally taking into account spatiotemporal dependencies present in M/EEG data.
 We compare the performance of the proposed approach to well-established alternative methods using both simulated and actual M/EEG data and show that it significantly outperforms alternative methods in estimating the onset and offset of M/EEG effects.
@@
 The issues with multiple comparisons represent a common and well-recognised danger in neuroimaging and M/EEG research, where the collected data allows for a multitude of potential hypothesis tests and is characterised by complex structures of spatiotemporal dependencies.
-The probability of obtaining at least one false positive in an ensemble (family) of $m$ tests (i.e., the FWER) is computed as $1-\left(1-\alpha\right)^{m}$ (for $m=10$ tests and $\alpha=0.05$, it is approximately equal to $0.4$).
-Different methods exist to control the FWER (i.e., bring it back to $\alpha$).
+The probability of obtaining at least one false positive in an ensemble (family) of $m$ tests (i.e., the FWER) is computed as $1-\left(1-\alpha\right)^{m}$ (for $m=10$ independent tests and $\alpha=0.05$, it is approximately equal to $0.4$).
+Different methods exist to control the FWER, that is, to bring it back to $\alpha$.
 Most methods apply a simple correction to series of $p$-values issued from univariate statistical tests (e.g., t-tests).
@@
 As such, it would be incorrect to conclude, for instance, that the timestep of a significant cluster is the first moment at which some conditions differ [@frossard2022; @sassenhagen2019].
-In other words, because the inference is performed at the second step (i.e., once clusters have been formed), no conclusion can be made about individual datapoints (e.g., timesteps, sensors, etc).
+Because the inference is performed at the second step (i.e., once clusters have been formed), no conclusion can be made about individual datapoints (e.g., timesteps, sensors, etc).
 
@@
 where $h_{0}$ is typically $0$ and parameters $E$ and $H$ are set a priori (typically to 0.5 and 2, respectively) and control the influence of the extend and height on the TFCE.
-Note that in practice, this integral is approximated by a sum over small $h$ increments.
+In practice, this integral is approximated by a sum over small $h$ increments.
 Then, a *p*-value for each timestep $t$ is computed by comparing its TFCE with the null distribution of TFCE values (obtained via permutation).
@@
 However, this averaging approach has notable limitations: it assumes consistent latency and amplitude across trials and is primarily suited for simple categorical designs.
-Such assumptions may not hold in more complex experimental paradigms, potentially leading to suboptimal ERP estimations.
-
-<!--
-
-To overcome the limitations of averaging, a number of model-based options for estimating ERPs have been proposed.
-These approaches are justified by the observation that the averaging approach to ERPs is equivalent to an intercept-only linear regression in a simple categorical design with no overlapping factors [@smith2014a].
-The regression approach to ERPs is much more flexible than simpl averaging, as it allows modelling linear and non-linear effects of continuous predictors such as word frequency or age [e.g., @smith2014a; @smith2014b; @tremblay2014] and allow modelling and disentangling overlapping processes [e.g., @skukies_brain_2024; @skukies_modelling_2021] as implemented in the UNFOLD toolbox [@ehinger_unfold_2019].
-The regression approach, as implemented for instance in the LIMO toolbox [@pernet2011], is a multi-stage procedure in which a participant-level regression model is fitted independently to each datapoint (e.g., channel, timestep, etc.) to estimate the ERP, followed by group-level statistical analyses of the regression coefficients, itself followed by multiplicity corrections or cluster-based inference [for recent applied examples, see for instance @dunagan2025; @wüllhorst2025].
-Whereas this approach allows including linear and non-linear discrete and continuous predictors when estimating the ERP, it has several limitations.
-First, the participant-level modelling ignores spatiotemporal dependencies as it treats all datapoints independently.
-Second, this approach does not take into account the participant (or stimulus) variability when testing group-level hypotheses (as could be done for instance via multilevel modelling).
-Finally, because the output of this procedure is summarised by cluster-based inference, its conclusions suffer from the same limitations discussed in the previous section.
-
--->
+Such assumptions may not hold in more complex experimental paradigms, potentially leading to suboptimal ERP estimations [e.g., @smith2014a].
 
@@
 In contrast to simple averaging, regression-based approaches to ERP estimation offer substantially greater flexibility.
-Notably, they allow for the modelling of both linear and nonlinear effects of continuous predictors--such as word frequency or age [e.g., @smith2014a; @smith2014b; @tremblay2014]--and enable the disentangling of overlapping cognitive processes [e.g., @ehinger_unfold_2019; @skukies_brain_2024; @skukies_modelling_2021].
+Notably, they allow for the modelling of both linear and nonlinear effects of continuous predictors, such as word frequency or age [e.g., @smith2014a; @smith2014b; @tremblay2014], and enable the disentangling of overlapping cognitive processes [e.g., @ehinger_unfold_2019; @skukies_brain_2024; @skukies_modelling_2021].
 One widely used implementation of this approach is provided by the LIMO EEG toolbox [@pernet2011], which follows a multi-stage analysis pipeline.
 First, a separate regression model is fit for each datapoint (e.g., each time point and electrode) at the individual participant level to estimate ERP responses.
-This is followed by group-level statistical analyses of the resulting regression coefficients, often accompanied by multiple comparison corrections or cluster-based inference [for recent applied examples, see @dunagan2025; @wüllhorst2025].
+This is followed by group-level statistical analyses of the resulting regression coefficients, often accompanied by corrections for multiple comparisons or cluster-based inference [for recent applied examples, see @dunagan2025; @wüllhorst2025].
 
@@
 First, fitting separate models for each datapoint ignores the spatiotemporal dependencies inherent in M/EEG data, potentially reducing statistical power and interpretability.
-Second, the subsequent group-level analyses typically do not account for hierarchical dependencies, between-participant or between-stimulus variability, which could otherwise be addressed through multilevel modelling.
+Second, the subsequent group-level analyses typically do not account for hierarchical dependencies which could otherwise be addressed through multilevel modelling.
 Finally, because the output of this procedure is summarised by cluster-based inference, its conclusions remain subject to the limitations discussed in the previous section.
 
-<!--
-
-Building upon the regression framework, recent advances allow modelling and disentangling overlapping processes [e.g., @skukies_brain_2024; @skukies_modelling_2021] as implemented in the UNFOLD toolbox [@ehinger_unfold_2019]...
-
-In addition to modelling the non-linear effects of continuous predictors (e.g., word frequency) on the ERP amplitude, GAMs have been used for modelling the effect of time, that is, the shape of the ERP itself [@abugaber2023; @baayen2018; @meulman2015; @meulman2023]...
-In the next section, we turn to a brief introduction to generalised additive models and discuss their use for modelling M/EEG timeseries...
-
--->
-
-Beyond modelling nonlinear effects of continuous predictors on ERP amplitudes, generalised additive models (GAMs) have been employed to capture the temporal dynamics of ERPs themselves, effectively modelling the shape of the waveform over time [@abugaber2023; @baayen2018; @meulman2015; @meulman2023].
+Beyond modelling nonlinear effects of continuous predictors on ERP amplitudes, GAMs have been employed to capture the temporal dynamics of ERPs themselves, effectively modelling the shape of the waveform over time [@abugaber2023; @baayen2018; @meulman2015; @meulman2023].
 This approach allows for the estimation of smooth, data-driven functions that characterise how neural responses evolve over time, offering a flexible alternative to traditional linear models.
@@
 
-In generalised additive models (GAMs), the functional relationship between the predictors and the response variable is decomposed into a sum of low-dimensional non-parametric functions.
+In generalised additive models, the functional relationship between the predictors and the response variable is decomposed into a sum of low-dimensional non-parametric functions.
 A typical GAM has the following form:
@@
 These simpler functions are called *basis functions* (e.g., cubic polynomial, thin-plate) and the set of basis functions is called a *basis*.
-Each basis function gets its coefficient and the resultant spline is the sum of these weighted basis functions (@fig-intro-gam\ignorespaces A).
-Splines coefficients are penalised (usually through the squared of the smooth functions' second derivative) in a way that can be interpreted, in Bayesian terms, as a prior on the "wiggliness" of the function [@miller2025; @wood2017].
+Each basis function is weighted by its coefficient and the resultant spline is the sum of these weighted basis functions (@fig-intro-gam\ignorespaces A).
+Splines coefficients are penalised (usually through the square of the smooth functions' second derivative) in a way that can be interpreted, in Bayesian terms, as a prior on the "wiggliness" of the function [@miller2025; @wood2017].
 In other words, more complex (wiggly) basis functions are automatically penalised.
@@
 
-```{r fig-intro-gam, echo = FALSE, fig.width = 9, fig.cap = "Different types of GAMs.
+```{r fig-intro-gam, fig.width = 9, fig.cap = "Different types of GAMs.
 **A**: GAMs predictions are computed as the weigthed sum (in black) of basis functions (here thin-plate basis functions, in grey).
@@
 Their appeal for modelling M/EEG data lies in their ability to flexibly capture the complex shape of ERP waveforms without overfitting, through the use of smooth functions constrained by penalisation.
-Recent extensions, such as distributional GAMs [@rigby2005; @umlauf2018], allow researchers to model not only the mean structure but also the variance (or scale) and other distributional properties as functions of predictors--a feature that has proven useful in modelling neuroimaging data [e.g., @dinga2021].
+Recent extensions, such as distributional GAMs [@rigby2005; @umlauf2018], allow researchers to model not only the mean structure but also the variance (or scale) and other distributional properties as functions of predictors, a feature that has proven useful in modelling neuroimaging data [e.g., @dinga2021].
 Moreover, hierarchical or multilevel GAMs [@pedersen_hierarchical_2019] provide a principled way to account for the nested structure of M/EEG data (e.g., trials within participants), enabling the inclusion of varying intercepts, slopes, and smoothers (as illustrated in @fig-intro-gam\ignorespaces C-D).
 This approach mitigates the risk of overfitting and reduces the influence of outliers on smooth estimates [@baayen2020; @meulman2023].
-
-<!--
-
-In the following, we introduce our proposed approach based on Bayesian generalized additive multilevel models (Bayesian GAMMs), which builds on these strengths while providing fully probabilistic estimates of ERP dynamics.
-We illustrate how this framework enables direct inference on the presence and timing of neural effects at the trial level, across conditions and participants, and show that it outperforms conventional methods in both simulated and real M/EEG datasets.
-
--->
 
@@
 We evaluated the performance of this approach against conventional methods using both simulated and actual M/EEG data.
-Our findings demonstrate that this method provides more precise and reliable estimates of effect onsets and offsets, offering a valuable tool for researchers seeking to enhance the temporal resolution of their analyses.
-
-# Methods
-
-## M/EEG data simulation
-
+Our findings demonstrate that this method provides more precise and reliable estimates of effects' onset and offset than conventional approaches such as cluster-based inference.
+
+# Benchmarking with known ground truth
+
+## Methods
+
+### M/EEG data simulation
+
+To assess the accuracy of group-level onset and offset estimation of our proposed method, we simulated EEG with known onset and offset values.
 Following the approach of @sassenhagen2019 and @rousselet_using_2025, we simulated EEG data stemming from two conditions, one with noise only, and the other with noise + signal.
@@
 We simulated data for a group of 20 participants (with variable true onset) with 50 trials per participant and condition (@fig-eeg).
-All figures and simulation results can be reproduced using the `R` code available on GitHub: <https://github.com/lnalborczyk/brms_meeg>.
+All figures and simulation results can be reproduced using the `R` code available online at: <https://github.com/lnalborczyk/brms_meeg>.
 
@@
 
-## Model description and model fitting
-
-We then fitted a Bayesian GAM (BGAM) using the `brms` package [@brms2017; @brms2018; @nalborczyk2019] and default priors (i.e., weakly informative priors).
+### Model description and model fitting
+
+We then fitted a Bayesian GAM (BGAM) using the `brms` package [@brms2017; @brms2018] and default priors (i.e., weakly informative priors).
 We ran eight Markov Chain Monte-Carlo (MCMC) to approximate the posterior distribution, including each 5000 iterations and a warmup of 2000 iterations, yielding a total of $8 \times (5000-2000) = 24000$ posterior samples to use for inference.
@@
 However, the previous model only included constant (fixed) effects, thus not properly accounting for between-participant variability.
-We next fitted a multilevel version of the BGAM [BGAMM, for an introduction to Bayesian multilevel models in `brms`, see @nalborczyk2019] including a varying intercept and slope for participant (but only with a constant smoother).
+We next fitted a multilevel version of the BGAM [BGAMM, for an introduction to Bayesian multilevel models in `brms`, see @nalborczyk2019] including a varying intercept and slope for participant (but with a constant smoother).
 Although it is possible to fit a BGAMM using data at the single-trial level, we present a computationally lighter version of the model that is fitted directly on by-participant summary statistics (mean and SD), similar to what is done in meta-analysis.
@@
 We depict the posterior predictions together with the posterior estimate of the slope for `condition` at each timestep (@fig-plot-post-slope).
-This figure suggests that the BGAMM provides an adequate description of the simulated data (see further posterior predictive checks in @sec-basis).
+This figure suggests that the BGAMM provides an adequate description of the simulated data (see further posterior predictive checks in @apx-basis).
 
@@
 We then compute the posterior probability of the slope for `condition` being above $0$ (@fig-post-prob-ratio, left).
-From this quantity, we compute the ratio of posterior probabilities (i.e., $p/(1-p)$) and visualise the timecourse of this ratio superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio, right).
-Note that a ratio of 10 means that the probability of the difference being above 0 is 10 times higher than the probability of the difference not being above 0, given the data, the priors, and other model's assumptions.[^1] Thresholding the posterior probability ratio thus provides a model-based approach for estimating the onset and offset of M/EEG effects, whose properties will be assessed in the simulation study.
+From this quantity, we compute the ratio of posterior probabilities (i.e., $p/(1-p)$), or posterior odds, and visualise the timecourse of these odds superimposed with the conventional thresholds on evidence ratios (@fig-post-prob-ratio, right).
+A ratio of 10 means that the probability of the difference being above 0 is 10 times higher than the probability of the difference not being above 0, given the data, the priors, and other model's assumptions.[^1] Thresholding the posterior odds thus provides a model-based approach for estimating the onset and offset of M/EEG effects, whose properties will be assessed in the simulation study.
 An important advantage is that the proposed approach can be extended to virtually any model structure.
 
-[^1]: This posterior probability ratio, or posterior odds, is equivalent to a Bayes factor, assuming 1:1 prior odds.
+[^1]: These posterior odds are equivalent to a Bayes factor, assuming 1:1 prior odds.
 
@@
 ```{r fig-post-prob-ratio, echo = FALSE, out.width = "100%", fig.width = 9, fig.asp = 0.5, fig.cap = "Left: Posterior probability of the EEG difference (slope) being above 0 according to the BGAMM.
-Right: Ratio of posterior probability according to the BGAMM (on a log10 scale).
+Right: Posterior odds according to the BGAMM (on a log10 scale).
 Timesteps above threshold (10) are highlighted in green.
-NB: the minimum and maximum possible ratio values are determined (bounded) by the number of posterior samples in the model."}
+NB: the minimum and maximum possible posterior odds are determined (bounded) by the number of posterior samples in the model."}
 # defining a function to compute and plot the posterior probability ratio
@@
         ylim(c(0, 1) ) +
-        labs(x = "Time (s)", y = expression(Pr(beta>0~"|"~data, model) ) )
+        # labs(x = "Time (s)", y = expression(Pr(beta>0~"|"~data, model) ) )
+        labs(x = "Time (s)", y = expression(p(beta>0~"|"~data) ) )
 
@@
             x = "Time (s)",
-            y = expression(Pr(beta>0~"|"~data) / (1 - Pr(beta>0~"|"~data) ) )
+            y = expression(p(beta>0~"|"~data) / (1 - p(beta>0~"|"~data) ) )
             )
@@
             x = "Time (s)",
-            y = expression(Pr(beta>0~"|"~data, model) ),
+            # y = expression(Pr(beta>0~"|"~data, model) ),
+            y = expression(p(beta>0~"|"~data) ),
             title = "Posterior probabilities per participant"
@@
             x = "Time (s)",
-            y = expression(Pr(beta>0~"|"~data) / (1 - Pr(beta>0~"|"~data) ) ),
+            y = expression(p(beta>0~"|"~data) / (1 - p(beta>0~"|"~data) ) ),
             title = "Posterior probability ratios per participant"
@@
 
+<!--
+
 \newpage
 
-## Error properties of the proposed approach
-
-<!--
-
-Note: Given the focus on the precision of onset/offset estimates, is it correct/relevant/sensible to mention that we are interested here in quantifying type-S and type-M (rather than type-1 or type-2) errors?
-
--->
+### Error properties of the proposed approach
 
@@
 @fig-onset-error shows that the multilevel GAM can almost exactly recover the true onset and offset values, given some reasonable choice of `k` and `threshold` values.
-We provide more detailed recommendations on how to set `k` in @sec-basis.
-This figure further reveals that the optimal `k` and `threshold` values may differ for the onset and offset values, and there there seems to exist a trade-off between these two parameters: lower `k` values lead to poorer estimations, but these poor estimations can be compensated (only to some extent) by higher `threshold` values (and reciprocally).
+We provide more detailed recommendations on how to set `k` in @apx-basis.
+This figure further reveals that the optimal `k` and `threshold` values may differ for the onset and offset values, and there seems to exist a trade-off between these two parameters: lower `k` values lead to poorer estimations, but these poor estimations can be compensated (only to some extent) by higher `threshold` values (and reciprocally).
 
@@
 
-## Comparing the onsets/offsets estimates across approaches
+-->
+
+### Comparing the onset/offset estimates across approaches
 
@@
 The cluster-based inference was implemented using a cluster-sum statistic of squared $t$-values, as implemented in `MNE-Python` [@gramfort2013], called via `r pkrt("reticulate")`.
-We also compared these estimates to the onset and offset as estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared $t$-values [as in @rousselet_using_2025].
-@fig-corrections illustrates the onsets and offsets estimated by each method on a single simulated dataset and shows that all methods systematically overestimate the true onset and underestimate the true offset.
+We also compared these estimates to the onset and offset as estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared $t$-values [as in @rousselet_using_2025].^[As in @rousselet_using_2025, we fixed the number of expected change points to two in the binary segmentation algorithm, thus producing always one cluster.] @fig-corrections illustrates the onsets and offsets estimated by each method on a single simulated dataset and shows that all methods systematically overestimate the true onset and underestimate the true offset.
+In addition, the `Raw p-value`, `FDR BH95`, and `FDR BY01` methods identify clusters well before the true onset and after the true offset.
 
@@
 # estimating onset and offset using the binary segmentation method
-res <- cpt.meanvar(data = tests_results$tval, method = "BinSeg", Q = 2)
+res <- cpt.meanvar(data = tests_results$tval, method = "BinSeg", Q = 2, penalty = "MBIC", minseglen = 2)
 onset_cpt <- tests_results$time[res@cpts[1]]
 offset_cpt <- tests_results$time[res@cpts[2]]
+
+# sanity checks
+# ts.plot(tests_results$tval)
+# for (i in 1:length(res@cpts) ) {
+#     
+#     abline(v = res@cpts[i])
+#     
+# }
 
@@
 
-\newpage
-
-## Simulation study
-
-To assess the accuracy of group-level onset estimation, the various methods were compared using the bias (defined as the median difference between the estimated and true value of the onset/offset), median absolute error (MAE), root mean square error (RMSE), variance, and median absolute deviation (MAD) of onset/offset estimates computed on 10,000 simulated datasets.
-As in @rousselet_using_2025, each participant was assigned a random onset between 150 and 170ms.
-Whereas the focus of the present article focuses is on one-dimensional signals (e.g., one M/EEG channel), we provide a 2D application in @sec-2D.
-
-## Application to actual MEG data
-
-To complement the simulation study, we evaluated the performance of all methods on actual MEG data [decoding results from @nalborczyk:inprep].
-In this study, we conducted time-resolved multivariate pattern analysis (MVPA, also known as decoding) of MEG data during reading tasks.
-As a result, we obtain a timecourse of decoding performance (ROC AUC), bounded between 0 and 1, for each participant ($N=32$).
-Next, we wanted to *test* whether the group-level average decoding accuracy is above chance (i.e., 0.5) at each timestep (@fig-decoding-data).
-To achieve this, we fitted a BGAM as introduced previously, but we replaced the $\mathrm{Normal}$ likelihood function by a $\mathrm{Beta}$ one to account for the bounded nature of AUC values (between 0 and 1) [for a tutorial on Beta regression, see @coretta2025].
-
-Note that although we chose a basis dimension of $k=50$, which seems appropriate for the present data, this choice should be adapted according to the properties of the modelled data (e.g., signal-to-noise ratio, prior low-pass filtering, sampling rate) and should be assessed by the usual model checking tools (e.g., posterior predictive checks, see also @sec-basis).
-To better distinguish signal from noise, we also defined a region of practical equivalence [ROPE, @kruschke2017], defined as the chance level plus the standard deviation of the (group-level average) decoding performance during the baseline period.
-
-```{r decoding-model, eval = FALSE}
-# fitting the Beta GAM
-meg_decoding_gam <- brm(
-    auc ~ s(time, bs = "cr", k = 50),
-    data = decoding_df,
-    family = Beta(),
-    warmup = 2000,
-    iter = 5000,
-    chains = 4,
-    cores = 4
-    )
-```
-
-We assessed the reliability of the proposed approach using a form of permutation-based split-half reliability [as for instance in @rosenblatt2018], which consisted of the following steps.
-First, we created 1,000 split halves of the data (i.e., with half the participants in the original data, that is, 16 participants).
-For each split, we estimated the onset/offset using all methods described previously.
-Third, we summarised the distribution of onset/offset estimates using the median "error" (i.e., difference between the split estimate and the estimate obtained using the full dataset) and the variance across splits.
-This approach allows assessing how similar the estimate of each half split is to the full dataset (thus acting as a proxy for the population) and how variable the estimates are across split halves.
-
-```{r fig-decoding-data, cache = FALSE, echo = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Group-level average decoding performance (N=32) superimposed with the GAM predictions (in blue) and the region of practical equivalence (ROPE, in orange) computed from the baseline period [data from @nalborczyk:inprep].
-The blue horizontal markers indicate the timesteps at which the posterior probability ratio exceeds 20."}
-# importing the reshaped MEG decoding results
-decoding_df <- read.csv(file = "data/decoding_results_reshaped.csv") %>%
-    # removing some participants
-    mutate(participant_id = cur_group_id(), .by = participant) %>%
-    filter(participant_id < 33) %>%
-    select(-participant_id)
-
-# averaging across participants
-decoding_summary_df <- decoding_df %>%
-    summarise(
-        auc_mean = mean(auc),
-        auc_sd = sd(auc),
-        .by = c(participant, time)
-        )
-
-# defining the chance level
-chance_level <- 0.5
-
-# computing the smallest effect size of interest (SESOI)
-# as the SD of decoding performance during the baseline
-sesoi <- decoding_summary_df %>%
-    filter(time < 0) %>%
-    summarise(auc_mean = mean(auc_mean), .by = time) %>%
-    summarise(sesoi = sd(auc_mean) ) %>%
-    pull(sesoi)
-
-# computing the smallest effect size of interest (SESOI)
-# as the 95% confidence interval during the baseline
-# sesoi <- decoding_summary_df %>%
-#     filter(time < 0) %>%
-#     summarise(auc_mean = mean(auc_mean), .by = time) %>%
-#     summarise(sesoi = 1.96 * (sd(auc_mean) / sqrt(n() ) ) ) %>%
-#     pull(sesoi)
-
-# fitting the Beta GAM
-meg_decoding_gam <- brm(
-    auc_mean ~ s(time, bs = "cr", k = 50),
-    # auc_mean ~ 1 + baseline + s(time, bs = "tp", k = 30, by = baseline),
-    data = decoding_summary_df,
-    family = Beta(),
-    warmup = 1000,
-    iter = 4000,
-    chains = 4,
-    cores = 4,
-    file = "models/meg_decoding_gam.rds"
-    # file = "models/meg_decoding_gam_baseline.rds"
-    )
-
-# computing the posterior probability
-prob_y_above_0 <- data.frame(time = unique(meg_decoding_gam$data$time) ) %>%
-    # retrieving the posterior samples
-    add_epred_draws(object = meg_decoding_gam) %>%
-    # converting to dataframe
-    data.frame() %>%
-    # computing mean posterior probability at the group level
-    group_by(time) %>%
-    summarise(m = mean(.epred > (0 + chance_level + sesoi) ) ) %>%
-    mutate(prob_ratio = m / (1 - m) ) %>%
-    # ensuring there is no 0 or +Inf values
-    mutate(prob_ratio = pmin(prob_ratio, ndraws(meg_decoding_gam) ) ) %>%
-    mutate(prob_ratio = pmax(prob_ratio, 1 / ndraws(meg_decoding_gam) ) ) %>%
-    ungroup()
-
-# defining a function to find all onset and offset in timeseries
-find_clusters <- function (df, threshold = 20) {
-
-    df %>%
-        mutate(above_threshold = prob_ratio >= threshold) %>%
-        mutate(cluster_change = c(TRUE, diff(above_threshold) != 0) ) %>%
-        filter(above_threshold) %>%
-        mutate(cluster_id = cumsum(cluster_change) ) %>%
-        group_by(cluster_id) %>%
-        summarise(
-            cluster_onset = first(time), cluster_offset = last(time),
-            .groups = "drop"
-            ) %>%
-        select(cluster_onset, cluster_offset)
-
-}
-
-# defining a threshold
-post_prob_ratio_threshold <- 20
-clusters <- find_clusters(prob_y_above_0, threshold = post_prob_ratio_threshold)
-
-# checking model's predictions against raw data
-brms_color <- met.brewer(name = "Johnson", n = 8)[8]
-plot(
-    conditional_effects(x = meg_decoding_gam),
-    line_args = list(colour = brms_color, fill = brms_color, alpha = 0.3),
-    points = FALSE, plot = FALSE
-    )[[1]] +
-    geom_hline(yintercept = 0.5, linetype = 2) +
-    geom_vline(xintercept = 0.0, linetype = 2) +
-    annotate(
-        geom = "rect",
-        xmin = -Inf, xmax = Inf,
-        ymin = chance_level - sesoi, ymax = chance_level + sesoi,
-        fill ="orangered",
-        alpha = 0.2
-        ) +
-    geom_segment(
-        data = clusters,
-        aes(
-            x = cluster_onset,
-            xend = cluster_offset,
-            y = 0.48, yend = 0.48
-            ),
-        colour = brms_color,
-        inherit.aes = FALSE,
-        lineend = "round",
-        linewidth = 2
-        ) +
-    geom_line(
-        data = decoding_summary_df %>% summarise(auc_mean = mean(auc_mean), .by = time),
-        aes(x = time, y = auc_mean), inherit.aes = FALSE
-        ) +
-    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
-```
-
-```{r fig-post-prob-ratio-decoding, eval = FALSE, echo = FALSE, out.width = "100%", fig.width = 9, fig.asp = 0.5, fig.cap = "Left: Posterior probability of the EEG difference (slope) being above 0 according to the BGAMM.
-Right: Ratio of posterior probability according to the BGAMM (on a log10 scale).
-Timesteps above threshold (10) are highlighted in green.
-NB: the minimum and maximum possible ratio values are determined (bounded) by the number of posterior samples in the model."}
-# Computing the EDF?
-# From https://discourse.mc-stan.org/t/standard-operating-procedures-for-anova/3964/2
-# If Kruschke has code that does what you want, then you can probably call his code on some columns of as.matrix(fit).
-If it really has to be in coda format, there is an As.mcmc.list function in the rstan package that will convert a stanfit object to a mcmc.list, but you probably have to call it on the $stanfit element of the brmsfit object.
-# See the bamlss::samplestat() function at https://github.com/cran/bamlss/blob/0a2a4154e0c0099ba8ac0b372ae7460e57972741/R/BAMLSS.R#L2397
-# Maybe export brms posterior samples to mcmc.list and then use this function?
-# See https://mc-stan.org/rstan/reference/As.mcmc.list.html
-library(bamlss)
-# mcmc_samples <- as.mcmc(meg_decoding_gam)
-mcmc_samples <- as.mcmc(meg_decoding_gam$fit)
-mcmc_samples_stats <- samplestats(
-    samples = mcmc_samples,
-    x = bamlss.frame(formula = auc_mean ~ s(time, bs = "tp", k = 50), data = decoding_summary_df)$x,
-    # y = bamlss.frame(formula = auc_mean ~ s(time, bs = "tp", k = 50), data = decoding_summary_df)$y,
-    # family = beta_bamlss,
-    # logLik = FALSE
-    )
-mcmc_samples_stats
-
-# or using bamlss directly
-b_model <- bamlss(
-    formula = auc_mean ~ s(time, bs = "tp", k = 50),
-    family = beta_bamlss,
-    data = decoding_summary_df
-    )
-
-# computing the EDF (around 44.6)
-samplestats(samples = b_model)
-
-# comparing to p_WAIC (around 48.4)
-waic(meg_decoding_gam)
-
-# computing the posterior probability ratio (PPR)
-model <- meg_decoding_gam
-threshold <- 10
-
-# defining a function to compute and plot the posterior odds
-plot_posterior_odds <- function (model, threshold = 10) {
-
-    # retrieving the number of posterior samples
-    n_posterior_samples <- ndraws(model)
-
-    # defiing the full time sequence
-    # time_seq <- seq(min(model$data$time), max(model$data$time), length.out = 300)
-    time_seg <- unique(model$data$time)
-
-    # generating new data
-    newdata <- data.frame(time = time_seq)
-
-    # getting posterior expected values (marginal)
-    posterior_preds <- posterior_epred(object = model, newdata = newdata, re_formula = NA)
-
-    # computing posterior mean and 95% CI at each time
-    posterior_preds_df <- as.data.frame(posterior_preds) %>%
-        mutate(draw = row_number() ) %>%
-        pivot_longer(-draw, names_to = "time_index", values_to = "auc") %>%
-        mutate(time = rep(time_seq, each = nrow(posterior_preds) ) ) %>%
-        group_by(time) %>%
-        summarise(
-            auc_mean = mean(auc),
-            auc_lower = quantile(auc, 0.05),
-            auc_upper = quantile(auc, 0.95)
-            ) %>%
-        ungroup()
-
-    # computing the average upper 96% credible interval during baseline (i.e., before 0s)
-    upper_baseline_ci <- posterior_preds_df %>%
-        filter(time < 0) %>%
-        summarise(upper_baseline_ci = mean(auc_upper) ) %>%
-        pull(upper_baseline_ci)
-
-    # computing the posterior probability of AUC > upper CrI of baseline
-    prob_y_above_baseline <- model$data %>%
-        add_epred_draws(object = model, ndraws = 100) %>%
-        data.frame() %>%
-        # head()
-        # dplyr::select(participant, time, condition, .epred, .draw) %>%
-        # pivot_wider(names_from = condition, values_from = .epred) %>%
-        # mutate(epred_diff = cond2 - cond1) %>%
-        # dplyr::select(time, baseline, .epred, .draw) %>%
-        # pivot_wider(names_from = baseline, values_from = .epred) %>%
-        # head()
-        # mutate(epred_diff = `c - cond1) %>%
-        # posteriors_preds %>%
-        # computing the mean posterior probability at the group level
-        group_by(time) %>%
-        summarise(m = mean(.epred > upper_baseline_ci) ) %>%
-        mutate(prob_ratio = m / (1 - m) ) %>%
-        ungroup() %>%
-        # bounding posterior probabilities by the number of posterior samples
-        mutate(prob_ratio = pmin(prob_ratio, n_posterior_samples) ) %>%
-        mutate(prob_ratio = pmax(prob_ratio, 1 / n_posterior_samples) ) %>%
-        ungroup()
-
-    # plotting it
-    post_plot <- prob_y_above_baseline %>%
-        ggplot(aes(x = time, y = m) ) +
-        geom_hline(yintercept = 0.5, linetype = 2) +
-        # geom_vline(xintercept = true_onset, linetype = 2) +
-        # geom_vline(xintercept = true_offset, linetype = 2) +
-        geom_line() +
-        ylim(c(0, 1) ) +
-        labs(x = "Time (s)", y = expression(Pr(beta>0~"|"~data, model) ) )
-
-    # plotting it
-    ratio_plot <- prob_y_above_baseline %>%
-        mutate(above_thres = ifelse(prob_ratio >= threshold, 1, NA) ) %>%
-        ggplot(aes(x = time, y = prob_ratio) ) +
-        geom_hline(yintercept = 1, linetype = "dashed") +
-        geom_hline(yintercept = 10, linetype = "dashed", color = "darkred") +
-        geom_hline(yintercept = 1/10, linetype = "dashed", color = "darkred") +
-        geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
-        geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
-        # geom_vline(xintercept = true_onset, linetype = 2) +
-        # geom_vline(xintercept = true_offset, linetype = 2) +
-        geom_line(linewidth = 0.8) +
-        geom_point(
-            data = . %>% dplyr::filter(!is.na(above_thres) ),
-            aes(y = threshold),
-            colour = "darkgreen",
-            shape = 16,
-            size = 2,
-            na.rm = TRUE
-            ) +
-        scale_y_log10(
-            labels = label_log(digits = 2),
-            limits = c(1 / n_posterior_samples, n_posterior_samples),
-            breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
-            ) +
-        labs(
-            x = "Time (s)",
-            y = expression(Pr(beta>0~"|"~data) / (1 - Pr(beta>0~"|"~data) ) )
-            )
-
-    # plotting both the posterior probability and the ratio...
-    post_plot + ratio_plot
-
-}
-
-# plotting the posterior odds
-plot_posterior_odds(model = meg_decoding_gam, threshold = 10)
-
-####################################
-# or testing an alternative method #
-####################################
-
-# defining the full time sequence
-time_seq <- seq(min(decoding_summary_df$time), max(decoding_summary_df$time), length.out = 300)
-
-# generating new data
-newdata <- data.frame(time = time_seq)
-
-# retrieving the posterior expected values (marginal)
-posterior <- posterior_epred(object = meg_decoding_gam, newdata = newdata, re_formula = NA)
-
-# computing posterior mean and 95% CI at each time
-pred_df <- as.data.frame(posterior) %>%
-    mutate(draw = row_number() ) %>%
-    pivot_longer(-draw, names_to = "time_index", values_to = "auc") %>%
-    mutate(time = rep(time_seq, each = nrow(posterior) ) ) %>%
-    group_by(time) %>%
-    summarise(
-        auc_mean = mean(auc),
-        auc_lower = quantile(auc, 0.025),
-        auc_upper = quantile(auc, 0.975)
-        )
-
-# subsetting newdata to time < 0
-baseline_indices <- which(time_seq < 0)
-baseline_draws <- posterior[, baseline_indices]
-
-# computing the average baseline AUC per draw
-baseline_per_draw <- rowMeans(baseline_draws)
-
-# for timepoints ≥ 0
-post_indices <- which(time_seq >= 0)
-time_post <- time_seq[post_indices]
-
-# subsetting post-stimulus predictions
-posterior_post <- posterior[, post_indices]
-
-# computing posterior probability that AUC > pre-stim baseline
-post_prob <- sapply(1:length(post_indices), function (i) {
-
-    mean(posterior_post[, i] > baseline_per_draw)
-
-})
-
-# assembling into dataframe
-post_df <- data.frame(
-    time = time_post,
-    prob_auc_gt_baseline = post_prob
-    )
-
-# plotting the results
-n_posterior_samples <- ndraws(meg_decoding_gam)
-post_df %>%
-    ggplot(aes(x = time, y = prob_auc_gt_baseline / (1-prob_auc_gt_baseline) ) ) +
-    geom_line(size = 1, color = "darkgreen") +
-    geom_hline(yintercept = 0.95, linetype = "dashed", color = "gray") +
-    scale_y_log10(
-        labels = label_log(digits = 2),
-        limits = c(1 / n_posterior_samples, n_posterior_samples),
-        breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
-        ) +
-    labs(
-        title = "Posterior Probability AUC > Pre-Stimulus Baseline",
-        x = "Time (s)", y = "P(AUC > baseline)"
-        ) +
-    theme_bw()
-
-# or using yet another approach
-estimate_edf_brms <- function (
-    model, 
-    term,           # e.g., "s(time)"
-    time_seq = NULL,
-    n = 200,        # number of time points
-    re_formula = NA # exclude group-level effects
-    ) {
-
-  # Extract time variable name from term, e.g., "s(time)" → "time"
-  time_var <- gsub("s\\(([^,]+).*\\)", "\\1", term)
-
-  # Create prediction grid
-  if (is.null(time_seq)) {
-    time_range <- range(model$data[[time_var]], na.rm = TRUE)
-    time_seq <- seq(time_range[1], time_range[2], length.out = n)
-  }
-
-  # newdata <- data.frame(!!time_var := time_seq)
-  newdata <- data.frame(time_seq)
-  names(newdata) <- time_var
-
-  # Extract posterior smooth draws only (excludes intercept & fixed effects)
-  smooth_draws <- posterior_smooths(
-    model,
-    newdata = newdata,
-    smooth = term,
-    re_formula = re_formula,
-    resp = NULL
-  )
-
-  # Compute posterior mean at each timepoint
-  posterior_mean <- colMeans(smooth_draws)
-
-  # Compute variance of the posterior mean across time
-  var_mean <- var(posterior_mean)
-
-  # Compute variance at each timepoint across draws
-  pointwise_var <- apply(smooth_draws, 2, var)
-
-  # Mean pointwise variance
-  mean_pointwise_var <- mean(pointwise_var)
-
-  # EDF estimate
-  edf_estimate <- var_mean / mean_pointwise_var
-
-  return(list(
-    edf = edf_estimate,
-    var_mean = var_mean,
-    mean_pointwise_var = mean_pointwise_var,
-    time = time_seq,
-    smooth_mean = posterior_mean,
-    smooth_sd = sqrt(pointwise_var)
-  ))
-}
-
-edf_result <- estimate_edf_brms(
-  model = meg_decoding_gam,
-  #term = "s(time)"
-  term = 's(time,bs="tp",k=50)'
-)
-
-print(paste("Estimated EDF:", round(edf_result$edf, 2)))
-
-plot_df <- data.frame(
-  time = edf_result$time,
-  mean = edf_result$smooth_mean,
-  sd = edf_result$smooth_sd
-)
-
-ggplot(plot_df, aes(x = time, y = mean)) +
-  geom_line(color = "blue") +
-  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd), alpha = 0.2) +
-  labs(
-    title = paste("Smooth with Posterior SD | Estimated EDF ≈", round(edf_result$edf, 2)),
-    x = "Time", y = "Smooth Estimate"
-  ) +
-  theme_bw()
-```
-
-\newpage
-
-# Results
+### Simulation study
+
+To assess the accuracy of group-level onset and offset estimation, all methods were compared by computing the bias (defined as the mean difference between the estimated and true value of the onset/offset), mean absolute error (MAE), root mean square error (RMSE), and variance of onset/offset estimates from 10,000 simulated datasets.
+Following @rousselet_using_2025, each participant was assigned a random onset between 150 and 170ms.
+Whereas the present article focuses on one-dimensional signals (e.g., one M/EEG channel), we provide a 2D application in @apx-2D.
+
+<!--
+
+## Results
 
@@
 
-@fig-simulation-results shows a summary of the simulation results, revealing that the proposed approach (`BGAM`) has the lowest median error for both the onset and offset estimates.
-The `Cluster mass` and `Change point` also have good performance, but surprisingly, the `TFCE` method has relatively bad performance for estimating the effect offset (similar performance to the `Holm` and `FDR BY01` methods).
+-->
+
+## Results
+
+@fig-simulation-results shows a summary of the simulation results, revealing that the proposed approach (`BGAM`) has the lowest error for both the onset and offset estimates.
+The `Cluster mass` and `Change point` methods also have good performance, but perhaps surprisingly, the `TFCE` method performs poorly for estimating the offset of the effect (with performance similar to the `Holm` method).
 Unsurprisingly, the `FDR BH95` and `Raw p-value` methods show the worst performance.
 
-```{r fig-simulation-results, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median error and median absolute deviation of the error for the onset (left) and offset (right) estimates according to each method.
-Methods are ordered from lowest (top) to highest (bottom) median absolute error (separately for the onset and offset estimates)."}
+```{r fig-simulation-results, echo = FALSE, fig.width = 9, fig.asp = 0.6, fig.cap = "Mean error and variance of onset and offset estimates according to each method.
+Variance is plotted on a log10 scale for visual purposes."}
 # loading all RDS files
@@
 
-# sort files numerically by the number in the filename
+# sorting files numerically by the number in the filename
 files <- files[order(readr::parse_number(files) )]
 
-# read and bind all files, adding a sim_id
+# reading and binding all files
 sim_results <- map2_dfr(
@@
     .f = ~ readRDS(.x)
-    # .f = ~ readRDS(.x) |> mutate(sim_id = .y)
-    # .f = function (file, index) {
-    #     offset <- (index - 1) * splits_per_file
-    #     readRDS(file) %>% mutate(split_id = as.numeric(split_id) ) %>%
-    #         mutate(split_id = split_id + offset) %>%
-    #         mutate(split_id = formatC(x = split_id, width = 3, flag = 0) ) 
-    #     }
     )
@@
 #     summarise(
-#         bias_onset = median(onset - true_onset),
-#         MAE_onset = median(abs(onset - true_onset) ),
+#         bias_onset = mean(onset - true_onset),
+#         MAE_onset = mean(abs(onset - true_onset) ),
 #         variance_onset = var(onset),
-#         bias_offset = median(offset - true_offset),
-#         MAE_offset = median(abs(offset - true_offset) ),
+#         bias_offset = mean(offset - true_offset),
+#         MAE_offset = mean(abs(offset - true_offset) ),
 #         variance_offset = var(offset),
@@
 #         ) %>%
-#     # arrange(MAE_onset, MAE_offset)
-#     arrange(bias_onset, bias_offset)
+#     arrange(MAE_onset, MAE_offset)
+#     # arrange(bias_onset, bias_offset)
+#     # arrange(variance_onset, variance_offset)
 
@@
 #     geom_density(aes(fill = NULL) ) +
-#     # geom_histogram(breaks = "FD") +
 #     facet_grid(plot_row~onset_offset, scales = "free_x") +
@@
 # computing and plotting bias and variance
-# bias := median(sim_onset_offset_distrib - true_onset_offset)
+# bias := mean(sim_onset_offset_distrib - true_onset_offset)
 # variance := var(sim_onset_offset_distrib - true_onset_offset)
+sim_results %>%
+    # keeping only threshold = 20 for BGAM and threshold = 0.05 for frequentist results
+    filter(threshold %in% c(0.05, 20) ) %>%
+    mutate(
+        error = case_when(
+            # grepl("onset", onset_offset) ~ abs(value - true_onset),
+            # grepl("offset", onset_offset) ~ abs(value - true_offset)
+            grepl("onset", onset_offset) ~ value - true_onset,
+            grepl("offset", onset_offset) ~ value - true_offset
+            )
+        ) %>%
+    group_by(method, onset_offset) %>%
+    summarise(
+        mean_error = mean(error, na.rm = TRUE),
+        median_error = median(error, na.rm = TRUE),
+        mean_abs_error = mean(abs(error), na.rm = TRUE),
+        median_abs_error = median(abs(error), na.rm = TRUE),
+        variance = var(error, na.rm = TRUE),
+        mad = mad(value, na.rm = TRUE),
+        .groups = "drop"
+        ) %>%
+    ungroup() %>%
+    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c(
+                "raw_p", "pval_bh", "pval_by", "pval_holm",
+                "cluster_mass", "cluster_tfce", "cpt", "brms"
+                ),
+            labels = c(
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+                "TFCE", "Change point", "BGAM (our method)"
+                ),
+            )
+        ) %>%
+    ggplot(
+        aes(
+            x = mean_error,
+            # y = mad,
+            # y = median_abs_error,
+            # y = median_abs_error,
+            # x = median_abs_error,
+            y = variance,
+            # y = mean_abs_error,
+            colour = method,
+            fill = method
+            )
+        ) +
+    geom_vline(xintercept = 0, linetype = 2) +
+    # geom_errorbar(
+    #     aes(
+    #         xmin = median_error - mad, xmax = median_error + mad,
+    #         #ymin = median_abs_error - mad, ymax = median_abs_error + mad
+    #         )
+    #     ) +
+    #     geom_errorbar(
+    #     aes(
+    #         #xmin = median_error - mad, xmax = median_error + mad,
+    #         ymin = median_abs_error - mad, ymax = median_abs_error + mad
+    #         )
+    #     ) +
+    geom_point(
+        size = 2,
+        pch = 21,
+        colour = "white",
+        show.legend = FALSE
+        ) +
+    geom_label_repel(
+        aes(label = method),
+        colour = "white",
+        size = 3,
+        segment.color = NA,
+        show.legend = FALSE
+        ) +
+    # scale_x_log10(guide = "axis_logticks") +
+    scale_y_log10(
+        guide = "axis_logticks",
+        # breaks = trans_breaks("log10", function(x) 10^x),
+        # labels = trans_format("log10", math_format(10^.x) )
+        labels = label_log(digits = 2)
+        ) +
+    facet_wrap(~onset_offset) +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
+    labs(x = "Mean error (s)", y = "Variance (log-scale)")
+
+# or plotting only the error in descending order
 # sim_results %>%
-#     # keeping only threshold = 20 for BGAM and threshold = 0.05 for frequentist results
+#     # keeping only threshold = 10 for BGAM and threshold = 0.05 for frequentist results
 #     filter(threshold %in% c(0.05, 20) ) %>%
+#     # filter(threshold %in% c(0.05, 10) ) %>%
 #     mutate(
@@
 #         ) %>%
+#     # converting to ms
+#     # mutate(error = error * 1e3) %>%
 #     group_by(method, onset_offset) %>%
 #     summarise(
+#         mean_error = mean(error, na.rm = TRUE),
 #         median_error = median(error, na.rm = TRUE),
 #         median_abs_error = median(abs(error), na.rm = TRUE),
-#         # variance = var(error, na.rm = TRUE),
-#         variance = var(value, na.rm = TRUE),
+#         variance = var(error, na.rm = TRUE),
+#         # variance = var(value, na.rm = TRUE),
+#         sem = sd(value, na.rm = TRUE) / sqrt(n() ),
 #         mad = mad(value, na.rm = TRUE),
@@
 #                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-#                 "TFCE", "Change point", "BGAM (new method)"
+#                 "TFCE", "Change point", "BGAM (our method)"
 #                 ),
@@
 #         ) %>%
+#     mutate(
+#         method_reordered = tidytext::reorder_within(
+#             x = method,
+#             # by = 1 / error,
+#             # fun = median,
+#             # by = 1 / abs(median_error),
+#             by = 1 / median_abs_error,
+#             # by = 1 / abs(mean_error),
+#             within = onset_offset
+#             )
+#         ) %>%
+#     # mutate(
+#     #     method_reordered = fct_reorder(
+#     #         method, error, .fun = ~abs(median(.x) ), .na_rm = TRUE
+#     #         )
+#     #     ) %>%
+#     # arrange(desc(error) ) %>%
+#     # arrange(error) %>%
+#     # head()
 #     ggplot(
@@
 #             x = median_error,
-#             # y = mad,
-#             # y = median_abs_error,
-#             # y = median_abs_error,
-#             # x = median_abs_error,
-#             y = variance,
+#             # x = error,
+#             xmin = median_error - mad,
+#             xmax = median_error + mad,
+#             # x = mean_error,
+#             # xmin = mean_error - variance,
+#             # xmax = mean_error + variance,
+#             y = method_reordered,
 #             colour = method,
@@
 #     geom_vline(xintercept = 0, linetype = 2) +
-#     # geom_errorbar(
-#     #     aes(
-#     #         xmin = median_error - mad, xmax = median_error + mad,
-#     #         #ymin = median_abs_error - mad, ymax = median_abs_error + mad
-#     #         )
-#     #     ) +
-#     #     geom_errorbar(
-#     #     aes(
-#     #         #xmin = median_error - mad, xmax = median_error + mad,
-#     #         ymin = median_abs_error - mad, ymax = median_abs_error + mad
-#     #         )
-#     #     ) +
-#     geom_point(
-#         size = 2,
-#         pch = 21,
-#         colour = "white",
+#     geom_pointrange(
+#         size = 0.25,
+#         linewidth = 0.75,
 #         show.legend = FALSE
 #         ) +
-#     geom_label_repel(
-#         aes(label = method),
-#         colour = "white",
-#         size = 3,
-#         segment.color = NA,
-#         show.legend = FALSE
-#         ) +
-#     # scale_y_log10() +
-#     facet_wrap(~onset_offset) +
+#     # stat_histinterval(
+#     #     # breaks = "FD",
+#     #     # n = 1e3,
+#     #     # point_interval = "median_qi",
+#     #     # .width = c(0.5, 0.95),
+#     #     .width = 0.1,
+#     #     # colour = "black",
+#     #     # slab_colour = "white",
+#     #     # slab_linewidth = 0,
+#     #     slab_alpha = 0.8,
+#     #     # fill = "steelblue"
+#     #     show.legend = FALSE
+#     #     ) +
+#     facet_wrap(~onset_offset, scales = "free") +
+#     scale_y_reordered() +
 #     scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
 #     scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-#     # labs(x = "Median error (s)", y = "Variance (log-scale)")
-#     # labs(x = "Median error (s)", y = "Median absolute error (log-scale)")
-#     labs(x = "Median absolute error (s)", y = "Variance (log-scale)")
-
-# or plotting only the error in descending order
-sim_results %>%
-    # keeping only threshold = 10 for BGAM and threshold = 0.05 for frequentist results
-    filter(threshold %in% c(0.05, 20) ) %>%
-    # filter(threshold %in% c(0.05, 10) ) %>%
-    mutate(
-        error = case_when(
-            # grepl("onset", onset_offset) ~ abs(value - true_onset),
-            # grepl("offset", onset_offset) ~ abs(value - true_offset)
-            grepl("onset", onset_offset) ~ value - true_onset,
-            grepl("offset", onset_offset) ~ value - true_offset
-            )
-        ) %>%
-    # converting to ms
-    # mutate(error = error * 1e3) %>%
-    group_by(method, onset_offset) %>%
-    summarise(
-        mean_error = mean(error, na.rm = TRUE),
-        median_error = median(error, na.rm = TRUE),
-        median_abs_error = median(abs(error), na.rm = TRUE),
-        variance = var(error, na.rm = TRUE),
-        # variance = var(value, na.rm = TRUE),
-        sem = sd(value, na.rm = TRUE) / sqrt(n() ),
-        mad = mad(value, na.rm = TRUE),
-        .groups = "drop"
-        ) %>%
-    ungroup() %>%
-    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
-    mutate(
-        method = factor(
-            x = method,
-            levels = c(
-                "raw_p", "pval_bh", "pval_by", "pval_holm",
-                "cluster_mass", "cluster_tfce", "cpt", "brms"
-                ),
-            labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-                "TFCE", "Change point", "BGAM (new method)"
-                ),
-            )
-        ) %>%
-    mutate(
-        method_reordered = tidytext::reorder_within(
-            x = method,
-            # by = 1 / error,
-            # fun = median,
-            # by = 1 / abs(median_error),
-            by = 1 / median_abs_error,
-            # by = 1 / abs(mean_error),
-            within = onset_offset
-            )
-        ) %>%
-    # mutate(
-    #     method_reordered = fct_reorder(
-    #         method, error, .fun = ~abs(median(.x) ), .na_rm = TRUE
-    #         )
-    #     ) %>%
-    # arrange(desc(error) ) %>%
-    # arrange(error) %>%
-    # head()
-    ggplot(
-        aes(
-            x = median_error,
-            # x = error,
-            xmin = median_error - mad,
-            xmax = median_error + mad,
-            # x = mean_error,
-            # xmin = mean_error - variance,
-            # xmax = mean_error + variance,
-            y = method_reordered,
-            colour = method,
-            fill = method
-            )
-        ) +
-    geom_vline(xintercept = 0, linetype = 2) +
-    geom_pointrange(
-        size = 0.25,
-        linewidth = 0.75,
-        show.legend = FALSE
-        ) +
-    # stat_histinterval(
-    #     # breaks = "FD",
-    #     # n = 1e3,
-    #     # point_interval = "median_qi",
-    #     # .width = c(0.5, 0.95),
-    #     .width = 0.1,
-    #     # colour = "black",
-    #     # slab_colour = "white",
-    #     # slab_linewidth = 0,
-    #     slab_alpha = 0.8,
-    #     # fill = "steelblue"
-    #     show.legend = FALSE
-    #     ) +
-    facet_wrap(~onset_offset, scales = "free") +
-    scale_y_reordered() +
-    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    labs(
-        x = "Estimation error (s)",
-        # x = "Mean error (s)",
-        y = "Method"
-        )
+#     labs(
+#         x = "Estimation error (s)",
+#         # x = "Mean error (s)",
+#         y = "Method"
+#         )
 
@@
 
-Results are further summarised in @tbl-simulation-results, which shows that the `BGAM` is almost perfectly unbiased (i.e., it has a bias of approximately 2ms for the onset and 4ms for the offset).
+These results are further summarised in @tbl-simulation-results, which shows that the `BGAM` method is almost perfectly unbiased (i.e., it has a bias of approximately 0.1ms for the onset and 2.4ms for the offset).
 The `Bias` column shows that all methods tend to estimate the onset later than the true onset and to estimate the offset earlier than the true offset.
-As can be seen from this table, the `BGAM` has the best performance on all included metrics.
+As can be seen from this table, the `BGAM` method has the best performance on all included metrics (except for the `Variance` of the offset estimate, where the `Change point` method performs better, presumably because it was constrained to identifying a single cluster).
 
 ```{r tbl-simulation-results, eval = TRUE, echo = FALSE, results = "asis"}
-#| tbl-cap: "Summary statistics of the onset and offset estimates for each method (in ms, ordered by the MAE) according to the main simulation study."
+#| tbl-cap: "Summary statistics of onset/offset estimates for each method (in ms, ordered by the MAE)."
 # summarising the results
@@
     summarise(
-        # mean_error = mean(error, na.rm = TRUE),
-        median_error = median(error, na.rm = TRUE),
-        # mean_abs_error = mean(abs(error), na.rm = TRUE),
-        median_abs_error = median(abs(error), na.rm = TRUE),
+        mean_error = mean(error, na.rm = TRUE),
+        # median_error = median(error, na.rm = TRUE),
+        mean_abs_error = mean(abs(error), na.rm = TRUE),
+        # median_abs_error = median(abs(error), na.rm = TRUE),
         rmse = sqrt(mean(error)^2),
         variance = var(error, na.rm = TRUE),
-        # variance = var(value, na.rm = TRUE),
-        mad = mad(error, na.rm = TRUE),
+        # mad = mad(error, na.rm = TRUE),
         .groups = "drop"
@@
                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-                "TFCE", "Change point", "BGAM (new method)"
+                "TFCE", "Change point", "BGAM (our method)"
                 ),
@@
     # arrange(abs(mean_error) ) %>%
-    arrange(median_abs_error) %>%
+    # arrange(median_abs_error) %>%
+    arrange(mean_abs_error) %>%
     group_by(onset_offset) %>%
@@
         # mean_error = "Bias_mean",
-        median_error = "Bias",
-        median_abs_error = "MAE",
-        # mean_abs_error = "MAE",
+        # median_error = "Bias",
+        mean_error = "Bias",
+        # median_abs_error = "MAE",
+        mean_abs_error = "MAE",
         rmse = "RMSE",
-        variance = "Variance",
-        mad = "MAD"
+        variance = "Variance"
+        # mad = "MAD"
         ) %>%
@@
     as_latex()
+
 ```
 
-\newpage
-
-## Application to actual MEG data (reliability)
-
-@fig-onset-offset shows the group-level average decoding performance through time with onset and offset estimates from each method.
-Overall, this figure shows that both the `Raw p-value` and `FDR BH95` methods are extremely lenient, considering that the decoding performance is above chance before the onset of the stimulus (false positive) and until the end of the trial.
-The `Change point` and `Cluster mass` methods seem the most conservative methods, identifying a time window from approximately +60ms to +500ms.
-The `Holm`, `TFCE`, and `BGAM` methods produce similar estimates of onset and offset, ranging from approximately +60ms to +650ms, although the `BGAM` method seems to result in fewer clusters.[^2]
-
-[^2]: It should be noted that although each method can produce several "clusters" of timesteps, we only considered the first (onset) and last (offset) timesteps identified by each method to compute the estimation error.
-
-```{r fig-onset-offset, echo = FALSE, fig.width = 8, fig.asp = 0.6, fig.cap = "Group-level average decoding performance through time with onset and offset estimates for each method [data from @nalborczyk:inprep]."}
-# importing results
-# reliability_results <- readRDS(file = "results/reliability_results_100splits.rds")
-
-# loading all RDS files
-files <- list.files(
-    path = "results/reliability_cluster_results",
-    pattern = ".rds", 
-    full.names = TRUE
-    )
-
-# sorting files numerically by the number in the filename
-files <- files[order(readr::parse_number(files) )]
-
-# reading and binding all files, adding a sim_id
-reliability_results <- map2_dfr(
-    .x = files,
-    .y = seq_along(files),
-    .f = ~ readRDS(.x)
-    # .f = ~ readRDS(.x) |> mutate(sim_id = .y)
-    # .f = function (file, index) {
-    #     offset <- (index - 1) * splits_per_file
-    #     readRDS(file) %>% mutate(split_id = as.numeric(split_id) ) %>%
-    #         mutate(split_id = split_id + offset) %>%
-    #         mutate(split_id = formatC(x = split_id, width = 3, flag = 0) ) 
-    #     }
-    )
-
-# sanity checks
-# unique(reliability_results$split_id)
-# n_distinct(reliability_results$split_id)
-
-# extracting the full data estimates
-full_reliability_results <- reliability_results %>%
-    # filter(split_id == max(split_id) ) %>%
-    # filter(str_detect(string = split_id, pattern = "_s051") ) %>%
-    filter(split_id == "a001_s051") %>%
-    # keeping only the first and last timestep
-    # group_by(split_id, method, onset_offset) %>%
-    # summarise(
-    #     value = if (onset_offset[1] == "onset") min(value) else max(value),
-    #     .groups = "drop"
-    #     ) %>%
-    # ungroup() %>%
-    # pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
-    # data.frame() %>%
-    # select(-split_id) %>%
-    # separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
+# Application to actual MEG data
+
+## Methods
+
+To complement the simulation study, we evaluated the performance of all methods on actual MEG data [@nalborczyk:inprep].
+In this study, the authors conducted time-resolved multivariate pattern analysis (MVPA, also known as decoding) of MEG data recorded in 32 human participants during a reading task.
+As a result, the authors obtained a timecourse of decoding accuray (ROC AUC), bounded between 0 and 1, for each participant.
+To test whether the group-level average decoding accuracy was above chance (i.e., 0.5) at each timestep, we fitted a BGAM as introduced previously with a basis dimension $k=50$ and retained all timesteps exceeding a posterior odds of 20.
+To better distinguish signal from noise, we defined a region of practical equivalence [ROPE, @kruschke2017] as the upper 90% quantile of decoding performance during the baseline period (i.e., before stimulus onset).
+Although we chose a basis dimension of $k=50$, which seemed appropriate for the present data, this choice should be adapted according to the properties of the modelled data (e.g., signal-to-noise ratio, prior low-pass filtering, sampling rate) and should be assessed by the usual model checking tools (e.g., models comparison, posterior predictive checks, see @apx-basis).
+
+```{r fig-decoding-data, cache = FALSE, echo = FALSE, eval = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Group-level average decoding performance (N=32) superimposed with the BGAM predictions (in blue) and the region of practical equivalence (ROPE, in orange) computed from the baseline period [data from @nalborczyk:inprep].
+The blue horizontal markers indicate the timesteps at which the posterior probability ratio exceeds 20."}
+# importing the reshaped MEG decoding results
+decoding_df <- read.csv(file = "data/decoding_results_reshaped.csv") %>%
+    # removing some participants
+    mutate(participant_id = cur_group_id(), .by = participant) %>%
+    dplyr::filter(participant_id < 33) %>%
+    select(-participant_id)
+
+# averaging across participants
+decoding_summary_df <- decoding_df %>%
+    summarise(
+        auc_mean = mean(auc),
+        auc_sd = sd(auc),
+        .by = c(participant, time)
+        )
+
+# defining the chance level
+chance_level <- 0.5
+
+# computing the smallest effect size of interest (SESOI)
+# as the upper 90% quantile of decoding performance during the baseline
+sesoi <- decoding_summary_df %>%
+    dplyr::filter(time < 0) %>%
+    summarise(auc_mean = mean(auc_mean), .by = time) %>%
+    summarise(sesoi = quantile(x = auc_mean, probs = 0.90) ) %>%
+    # summarise(sesoi = quantile(x = auc_mean, probs = 0.95) ) %>%
+    pull(sesoi) - chance_level
+
+# computing the smallest effect size of interest (SESOI)
+# as the SD of decoding performance during the baseline
+# sesoi <- decoding_summary_df %>%
+#     dplyr::filter(time < 0) %>%
+#     summarise(auc_mean = mean(auc_mean), .by = time) %>%
+#     summarise(sesoi = sd(auc_mean) ) %>%
+#     pull(sesoi)
+
+# computing the smallest effect size of interest (SESOI)
+# as the 95% confidence interval during the baseline
+# sesoi <- decoding_summary_df %>%
+#     filter(time < 0) %>%
+#     summarise(auc_mean = mean(auc_mean), .by = time) %>%
+#     summarise(sesoi = 1.96 * (sd(auc_mean) / sqrt(n() ) ) ) %>%
+#     pull(sesoi)
+
+# fitting the GAMs
+meg_decoding_gam_gaussian_k40 <- brm(
+    auc_mean ~ s(time, bs = "cr", k = 40),
+    data = decoding_summary_df,
+    family = gaussian(),
+    warmup = 1000,
+    iter = 5000,
+    chains = 4,
+    cores = 4
+    )
+
+meg_decoding_gam_gaussian_k50 <- brm(
+    auc_mean ~ s(time, bs = "cr", k = 50),
+    data = decoding_summary_df,
+    family = gaussian(),
+    # warmup = 1000,
+    # iter = 5000,
+    warmup = 2000,
+    iter = 10000,
+    chains = 4,
+    cores = 4
+    )
+
+meg_decoding_gam_gaussian_k60 <- brm(
+    auc_mean ~ s(time, bs = "cr", k = 60),
+    data = decoding_summary_df,
+    family = gaussian(),
+    warmup = 1000,
+    iter = 5000,
+    chains = 4,
+    cores = 4
+    )
+
+meg_decoding_gam_beta_k40 <- brm(
+    auc_mean ~ s(time, bs = "cr", k = 40),
+    data = decoding_summary_df,
+    family = Beta(),
+    warmup = 1000,
+    iter = 5000,
+    chains = 4,
+    cores = 4
+    )
+
+meg_decoding_gam_beta_k50 <- brm(
+    auc_mean ~ s(time, bs = "cr", k = 50),
+    data = decoding_summary_df,
+    family = Beta(),
+    warmup = 1000,
+    iter = 5000,
+    chains = 4,
+    cores = 4
+    )
+
+meg_decoding_gam_beta_k60 <- brm(
+    auc_mean ~ s(time, bs = "cr", k = 60),
+    data = decoding_summary_df,
+    family = Beta(),
+    warmup = 1000,
+    iter = 5000,
+    chains = 4,
+    cores = 4
+    )
+
+# PPCs
+pp_check(meg_decoding_gam_gaussian_k40)
+pp_check(meg_decoding_gam_gaussian_k50)
+pp_check(meg_decoding_gam_gaussian_k60)
+pp_check(meg_decoding_gam_beta_k40)
+pp_check(meg_decoding_gam_beta_k50)
+pp_check(meg_decoding_gam_beta_k60)
+
+# models comparison
+loo(
+    meg_decoding_gam_gaussian_k40, meg_decoding_gam_gaussian_k50,
+    meg_decoding_gam_gaussian_k60,
+    meg_decoding_gam_beta_k40, meg_decoding_gam_beta_k50,
+    meg_decoding_gam_beta_k60
+    )
+
+meg_decoding_gam_gaussian_k40 <- add_criterion(x = meg_decoding_gam_gaussian_k40, criterio = "loo")
+meg_decoding_gam_gaussian_k50 <- add_criterion(x = meg_decoding_gam_gaussian_k50, criterio = "loo")
+meg_decoding_gam_gaussian_k60 <- add_criterion(x = meg_decoding_gam_gaussian_k60, criterio = "loo")
+meg_decoding_gam_beta_k40 <- add_criterion(x = meg_decoding_gam_beta_k30, criterio = "loo")
+meg_decoding_gam_beta_k50 <- add_criterion(x = meg_decoding_gam_beta_k50, criterio = "loo")
+meg_decoding_gam_beta_k60 <- add_criterion(x = meg_decoding_gam_beta_k60, criterio = "loo")
+
+models_comp <- loo_compare(
+    meg_decoding_gam_gaussian_k40, meg_decoding_gam_gaussian_k50,
+    meg_decoding_gam_gaussian_k60,
+    meg_decoding_gam_beta_k40, meg_decoding_gam_beta_k50,
+    meg_decoding_gam_beta_k60
+    )
+
+# summary
+print(x = models_comp, simplify = FALSE, digits = 3)
+
+# computing the posterior probability
+# prob_y_above_0 <- data.frame(time = unique(meg_decoding_gam$data$time) ) %>%
+prob_y_above_0 <- data.frame(time = unique(meg_decoding_gam_gaussian_k50$data$time) ) %>%
+    # retrieving the posterior samples
+    # add_epred_draws(object = meg_decoding_gam) %>%
+    add_epred_draws(object = meg_decoding_gam_gaussian_k50) %>%
+    # converting to dataframe
+    data.frame() %>%
+    # computing mean posterior probability at the group level
+    group_by(time) %>%
+    summarise(m = mean(.epred > (0 + chance_level + sesoi) ) ) %>%
+    # summarise(m = mean(.epred > (sesoi) ) ) %>%
+    mutate(prob_ratio = m / (1 - m) ) %>%
+    # ensuring there is no 0 or +Inf values
+    mutate(prob_ratio = pmin(prob_ratio, ndraws(meg_decoding_gam_gaussian_k50) ) ) %>%
+    mutate(prob_ratio = pmax(prob_ratio, 1 / ndraws(meg_decoding_gam_gaussian_k50) ) ) %>%
+    ungroup()
+
+# defining a function to find all onset and offset in timeseries
+find_clusters <- function (df, threshold) {
+
+    df %>%
+        mutate(above_threshold = prob_ratio >= threshold) %>%
+        mutate(cluster_change = c(TRUE, diff(above_threshold) != 0) ) %>%
+        dplyr::filter(above_threshold) %>%
+        mutate(cluster_id = cumsum(cluster_change) ) %>%
+        group_by(cluster_id) %>%
+        summarise(
+            cluster_onset = first(time), cluster_offset = last(time),
+            .groups = "drop"
+            ) %>%
+        select(cluster_onset, cluster_offset)
+
+}
+
+# defining a threshold
+post_prob_ratio_threshold <- 20
+clusters <- find_clusters(prob_y_above_0, threshold = post_prob_ratio_threshold)
+
+# defining a colour for the BGAM method
+brms_color <- met.brewer(name = "Johnson", n = 8)[8]
+
+# checking model's predictions against raw data
+plot(
+    # conditional_effects(x = meg_decoding_gam),
+    conditional_effects(x = meg_decoding_gam_gaussian_k50),
+    line_args = list(colour = brms_color, fill = brms_color, alpha = 0.3),
+    points = FALSE, plot = FALSE
+    )[[1]] +
+    geom_hline(yintercept = 0.5, linetype = 2) +
+    geom_vline(xintercept = 0.0, linetype = 2) +
+    annotate(
+        geom = "rect",
+        xmin = -Inf,
+        xmax = Inf,
+        ymin = chance_level - sesoi,
+        ymax = chance_level + sesoi,
+        # ymin = chance_level,
+        # ymax = sesoi,
+        fill ="orangered",
+        alpha = 0.2
+        ) +
+    geom_segment(
+        data = clusters,
+        aes(
+            x = cluster_onset,
+            xend = cluster_offset,
+            y = 0.48, yend = 0.48
+            ),
+        colour = brms_color,
+        inherit.aes = FALSE,
+        lineend = "round",
+        linewidth = 2
+        ) +
+    geom_line(
+        data = decoding_summary_df %>% summarise(auc_mean = mean(auc_mean), .by = time),
+        aes(x = time, y = auc_mean), inherit.aes = FALSE
+        ) +
+    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
+```
+
+```{r fig-post-prob-ratio-decoding, eval = FALSE, echo = FALSE, out.width = "100%", fig.width = 9, fig.asp = 0.5, fig.cap = "Left: Posterior probability of the EEG difference (slope) being above 0 according to the BGAMM.
+Right: Ratio of posterior probability according to the BGAMM (on a log10 scale).
+Timesteps above threshold (10) are highlighted in green.
+NB: the minimum and maximum possible ratio values are determined (bounded) by the number of posterior samples in the model."}
+# Computing the EDF?
+# From https://discourse.mc-stan.org/t/standard-operating-procedures-for-anova/3964/2
+# If Kruschke has code that does what you want, then you can probably call his code on some columns of as.matrix(fit).
+If it really has to be in coda format, there is an As.mcmc.list function in the rstan package that will convert a stanfit object to a mcmc.list, but you probably have to call it on the $stanfit element of the brmsfit object.
+# See the bamlss::samplestat() function at https://github.com/cran/bamlss/blob/0a2a4154e0c0099ba8ac0b372ae7460e57972741/R/BAMLSS.R#L2397
+# Maybe export brms posterior samples to mcmc.list and then use this function?
+# See https://mc-stan.org/rstan/reference/As.mcmc.list.html
+library(bamlss)
+# mcmc_samples <- as.mcmc(meg_decoding_gam)
+mcmc_samples <- as.mcmc(meg_decoding_gam$fit)
+mcmc_samples_stats <- samplestats(
+    samples = mcmc_samples,
+    x = bamlss.frame(formula = auc_mean ~ s(time, bs = "tp", k = 50), data = decoding_summary_df)$x,
+    # y = bamlss.frame(formula = auc_mean ~ s(time, bs = "tp", k = 50), data = decoding_summary_df)$y,
+    # family = beta_bamlss,
+    # logLik = FALSE
+    )
+mcmc_samples_stats
+
+# or using bamlss directly
+b_model <- bamlss(
+    formula = auc_mean ~ s(time, bs = "tp", k = 50),
+    family = beta_bamlss,
+    data = decoding_summary_df
+    )
+
+# computing the EDF (around 44.6)
+samplestats(samples = b_model)
+
+# comparing to p_WAIC (around 48.4)
+waic(meg_decoding_gam)
+
+# computing the posterior probability ratio (PPR)
+model <- meg_decoding_gam
+threshold <- 10
+
+# defining a function to compute and plot the posterior odds
+plot_posterior_odds <- function (model, threshold = 10) {
+
+    # retrieving the number of posterior samples
+    n_posterior_samples <- ndraws(model)
+
+    # defiing the full time sequence
+    # time_seq <- seq(min(model$data$time), max(model$data$time), length.out = 300)
+    time_seg <- unique(model$data$time)
+
+    # generating new data
+    newdata <- data.frame(time = time_seq)
+
+    # getting posterior expected values (marginal)
+    posterior_preds <- posterior_epred(object = model, newdata = newdata, re_formula = NA)
+
+    # computing posterior mean and 95% CI at each time
+    posterior_preds_df <- as.data.frame(posterior_preds) %>%
+        mutate(draw = row_number() ) %>%
+        pivot_longer(-draw, names_to = "time_index", values_to = "auc") %>%
+        mutate(time = rep(time_seq, each = nrow(posterior_preds) ) ) %>%
+        group_by(time) %>%
+        summarise(
+            auc_mean = mean(auc),
+            auc_lower = quantile(auc, 0.05),
+            auc_upper = quantile(auc, 0.95)
+            ) %>%
+        ungroup()
+
+    # computing the average upper 96% credible interval during baseline (i.e., before 0s)
+    upper_baseline_ci <- posterior_preds_df %>%
+        filter(time < 0) %>%
+        summarise(upper_baseline_ci = mean(auc_upper) ) %>%
+        pull(upper_baseline_ci)
+
+    # computing the posterior probability of AUC > upper CrI of baseline
+    prob_y_above_baseline <- model$data %>%
+        add_epred_draws(object = model, ndraws = 100) %>%
+        data.frame() %>%
+        # head()
+        # dplyr::select(participant, time, condition, .epred, .draw) %>%
+        # pivot_wider(names_from = condition, values_from = .epred) %>%
+        # mutate(epred_diff = cond2 - cond1) %>%
+        # dplyr::select(time, baseline, .epred, .draw) %>%
+        # pivot_wider(names_from = baseline, values_from = .epred) %>%
+        # head()
+        # mutate(epred_diff = `c - cond1) %>%
+        # posteriors_preds %>%
+        # computing the mean posterior probability at the group level
+        group_by(time) %>%
+        summarise(m = mean(.epred > upper_baseline_ci) ) %>%
+        mutate(prob_ratio = m / (1 - m) ) %>%
+        ungroup() %>%
+        # bounding posterior probabilities by the number of posterior samples
+        mutate(prob_ratio = pmin(prob_ratio, n_posterior_samples) ) %>%
+        mutate(prob_ratio = pmax(prob_ratio, 1 / n_posterior_samples) ) %>%
+        ungroup()
+
+    # plotting it
+    post_plot <- prob_y_above_baseline %>%
+        ggplot(aes(x = time, y = m) ) +
+        geom_hline(yintercept = 0.5, linetype = 2) +
+        # geom_vline(xintercept = true_onset, linetype = 2) +
+        # geom_vline(xintercept = true_offset, linetype = 2) +
+        geom_line() +
+        ylim(c(0, 1) ) +
+        labs(x = "Time (s)", y = expression(p(beta>0~"|"~data, model) ) )
+
+    # plotting it
+    ratio_plot <- prob_y_above_baseline %>%
+        mutate(above_thres = ifelse(prob_ratio >= threshold, 1, NA) ) %>%
+        ggplot(aes(x = time, y = prob_ratio) ) +
+        geom_hline(yintercept = 1, linetype = "dashed") +
+        geom_hline(yintercept = 10, linetype = "dashed", color = "darkred") +
+        geom_hline(yintercept = 1/10, linetype = "dashed", color = "darkred") +
+        geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
+        geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
+        # geom_vline(xintercept = true_onset, linetype = 2) +
+        # geom_vline(xintercept = true_offset, linetype = 2) +
+        geom_line(linewidth = 0.8) +
+        geom_point(
+            data = . %>% dplyr::filter(!is.na(above_thres) ),
+            aes(y = threshold),
+            colour = "darkgreen",
+            shape = 16,
+            size = 2,
+            na.rm = TRUE
+            ) +
+        scale_y_log10(
+            labels = label_log(digits = 2),
+            limits = c(1 / n_posterior_samples, n_posterior_samples),
+            breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
+            ) +
+        labs(
+            x = "Time (s)",
+            y = expression(p(beta>0~"|"~data) / (1 - p(beta>0~"|"~data) ) )
+            )
+
+    # plotting both the posterior probability and the ratio...
+    post_plot + ratio_plot
+
+}
+
+# plotting the posterior odds
+plot_posterior_odds(model = meg_decoding_gam, threshold = 10)
+
+####################################
+# or testing an alternative method #
+####################################
+
+# defining the full time sequence
+time_seq <- seq(min(decoding_summary_df$time), max(decoding_summary_df$time), length.out = 300)
+
+# generating new data
+newdata <- data.frame(time = time_seq)
+
+# retrieving the posterior expected values (marginal)
+posterior <- posterior_epred(object = meg_decoding_gam, newdata = newdata, re_formula = NA)
+
+# computing posterior mean and 95% CI at each time
+pred_df <- as.data.frame(posterior) %>%
+    mutate(draw = row_number() ) %>%
+    pivot_longer(-draw, names_to = "time_index", values_to = "auc") %>%
+    mutate(time = rep(time_seq, each = nrow(posterior) ) ) %>%
+    group_by(time) %>%
+    summarise(
+        auc_mean = mean(auc),
+        auc_lower = quantile(auc, 0.025),
+        auc_upper = quantile(auc, 0.975)
+        )
+
+# subsetting newdata to time < 0
+baseline_indices <- which(time_seq < 0)
+baseline_draws <- posterior[, baseline_indices]
+
+# computing the average baseline AUC per draw
+baseline_per_draw <- rowMeans(baseline_draws)
+
+# for timepoints ≥ 0
+post_indices <- which(time_seq >= 0)
+time_post <- time_seq[post_indices]
+
+# subsetting post-stimulus predictions
+posterior_post <- posterior[, post_indices]
+
+# computing posterior probability that AUC > pre-stim baseline
+post_prob <- sapply(1:length(post_indices), function (i) {
+
+    mean(posterior_post[, i] > baseline_per_draw)
+
+})
+
+# assembling into dataframe
+post_df <- data.frame(
+    time = time_post,
+    prob_auc_gt_baseline = post_prob
+    )
+
+# plotting the results
+n_posterior_samples <- ndraws(meg_decoding_gam)
+post_df %>%
+    ggplot(aes(x = time, y = prob_auc_gt_baseline / (1-prob_auc_gt_baseline) ) ) +
+    geom_line(size = 1, color = "darkgreen") +
+    geom_hline(yintercept = 0.95, linetype = "dashed", color = "gray") +
+    scale_y_log10(
+        labels = label_log(digits = 2),
+        limits = c(1 / n_posterior_samples, n_posterior_samples),
+        breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
+        ) +
+    labs(
+        title = "Posterior Probability AUC > Pre-Stimulus Baseline",
+        x = "Time (s)", y = "P(AUC > baseline)"
+        ) +
+    theme_bw()
+
+# or using yet another approach
+estimate_edf_brms <- function (
+    model, 
+    term,           # e.g., "s(time)"
+    time_seq = NULL,
+    n = 200,        # number of time points
+    re_formula = NA # exclude group-level effects
+    ) {
+
+  # Extract time variable name from term, e.g., "s(time)" → "time"
+  time_var <- gsub("s\\(([^,]+).*\\)", "\\1", term)
+
+  # Create prediction grid
+  if (is.null(time_seq)) {
+    time_range <- range(model$data[[time_var]], na.rm = TRUE)
+    time_seq <- seq(time_range[1], time_range[2], length.out = n)
+  }
+
+  # newdata <- data.frame(!!time_var := time_seq)
+  newdata <- data.frame(time_seq)
+  names(newdata) <- time_var
+
+  # Extract posterior smooth draws only (excludes intercept & fixed effects)
+  smooth_draws <- posterior_smooths(
+    model,
+    newdata = newdata,
+    smooth = term,
+    re_formula = re_formula,
+    resp = NULL
+  )
+
+  # Compute posterior mean at each timepoint
+  posterior_mean <- colMeans(smooth_draws)
+
+  # Compute variance of the posterior mean across time
+  var_mean <- var(posterior_mean)
+
+  # Compute variance at each timepoint across draws
+  pointwise_var <- apply(smooth_draws, 2, var)
+
+  # Mean pointwise variance
+  mean_pointwise_var <- mean(pointwise_var)
+
+  # EDF estimate
+  edf_estimate <- var_mean / mean_pointwise_var
+
+  return(list(
+    edf = edf_estimate,
+    var_mean = var_mean,
+    mean_pointwise_var = mean_pointwise_var,
+    time = time_seq,
+    smooth_mean = posterior_mean,
+    smooth_sd = sqrt(pointwise_var)
+  ))
+}
+
+edf_result <- estimate_edf_brms(
+  model = meg_decoding_gam,
+  #term = "s(time)"
+  term = 's(time,bs="tp",k=50)'
+)
+
+print(paste("Estimated EDF:", round(edf_result$edf, 2)))
+
+plot_df <- data.frame(
+  time = edf_result$time,
+  mean = edf_result$smooth_mean,
+  sd = edf_result$smooth_sd
+)
+
+ggplot(plot_df, aes(x = time, y = mean)) +
+  geom_line(color = "blue") +
+  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd), alpha = 0.2) +
+  labs(
+    title = paste("Smooth with Posterior SD | Estimated EDF ≈", round(edf_result$edf, 2)),
+    x = "Time", y = "Smooth Estimate"
+  ) +
+  theme_bw()
+```
+
+## Results
+
+@fig-onset-offset shows the group-level average decoding performance through time superimposed with onset and offset estimates from each method.
+Overall, this figure shows that both the `Raw p-value` and `FDR BH95` methods are extremely lenient, identifying clusters of above-chance decoding accuracy before the onset of the stimulus (false positive) and until the end of the trial.
+The `Change point` method seems to be the most conservative one, identifying a single cluster spanning from approximately +60ms to +450ms.
+The `Holm`, `Cluster mass`, `TFCE`, and `BGAM` methods produce roughly similar estimates of onset and offset, ranging from approximately +60ms to +650ms (considering only the first and last identified timesteps), although the `BGAM` method seems to result in fewer clusters.
+
+```{r fig-onset-offset, echo = FALSE, results = "hide", fig.width = 8, fig.asp = 0.6, fig.cap = "Group-level average decoding performance through time with clusters of higher-than-chance decoding performance as identified by each method [data from @nalborczyk:inprep]."}
+# importing the reshaped MEG decoding results
+decoding_df <- read.csv(file = "data/decoding_results_reshaped.csv") %>%
+    # removing the last participant (to get a round number of participants)
+    mutate(participant_id = cur_group_id(), .by = participant) %>%
+    dplyr::filter(participant_id < 33) %>%
+    select(-participant_id)
+
+# averaging across participants
+decoding_summary_df <- decoding_df %>%
+    summarise(
+        auc_mean = mean(auc),
+        auc_sd = sd(auc),
+        .by = c(participant, time)
+        )
+
+# defining the chance level
+chance_level <- 0.5
+
+# computing the SESOI as the baseline upper 95% quantile
+full_sesoi <- decoding_summary_df %>%
+    filter(time < 0) %>%
+    summarise(auc_mean = mean(auc_mean), .by = time) %>%
+    summarise(quantile(x = auc_mean, probs = 0.90) ) %>%
+    pull() - chance_level
+
+# importing Python modules
+use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
+
+# fitting the final GAM
+full_gam <- brm(
+    auc_mean ~ s(time, bs = "cr", k = 50),
+    data = decoding_summary_df,
+    family = gaussian(),
+    warmup = 2000,
+    iter = 10000,
+    chains = 8,
+    cores = 8,
+    file = "models/full_decoding_gam.rds"
+    )
+
+# sanity check
+cat("\nNumber of posterior samples in the full model:", ndraws(full_gam) )
+
+# computing the posterior probability
+prob_y_above_0 <- data.frame(time = unique(full_gam$data$time) ) %>%
+    # retrieving the posterior samples
+    add_epred_draws(object = full_gam) %>%
+    # using a subset of posterior samples
+    # add_epred_draws(object = full_gam, ndraws = n_post_samples) |>
+    # converting to dataframe
+    data.frame() %>%
+    # computing mean posterior probability at the group level
+    group_by(time) %>%
+    summarise(m = mean(.epred > (0 + chance_level + full_sesoi) ) ) %>%
+    # summarise(m = mean(.epred >= full_sesoi) ) |>
+    mutate(prob_ratio = m / (1 - m) ) %>%
+    # ensuring there is no 0 or +Inf values
+    mutate(prob_ratio = pmin(prob_ratio, ndraws(full_gam) ) ) %>%
+    mutate(prob_ratio = pmax(prob_ratio, 1 / ndraws(full_gam) ) ) %>%
+    ungroup()
+
+# importing home-made helper functions
+source("code/utils.R")
+
+# importing Python modules
+np <- import("numpy")
+mne <- import("mne")
+
+# defining alpha and posterior odds thresholds        
+alpha_level <- 0.05
+post_prob_ratio_threshold <- 20
+
+# finding clusters
+onset_offset_brms <- find_clusters(
+    data = prob_y_above_0 %>% select(time, value = prob_ratio),
+    threshold = post_prob_ratio_threshold
+    ) %>%
+    mutate(split_id = "full") %>%
+    select(split_id, cluster_id, onset = cluster_onset, offset = cluster_offset) %>%
+    mutate(method = "brms") %>%
+    pivot_longer(cols = onset:offset) %>%
+    select(split_id, method, cluster_id, onset_offset = name, value) %>%
+    data.frame()
+
+# converting data to a matrix (for later use in MNE functions)
+data_matrix <- matrix(
+    data = decoding_summary_df$auc_mean,
+    ncol = length(unique(decoding_summary_df$time) ),
+    byrow = TRUE
+    )
+
+# massive univariate t-tests
+tests_results <- decoding_summary_df %>%
+    group_by(time) %>%
+    summarise(
+        tval = t.test(x = auc_mean, mu = chance_level)$statistic^2,
+        pval = t.test(x = auc_mean, mu = chance_level)$p.value
+        ) %>%
+    mutate(
+        pval_bh = p.adjust(p = pval, method = "BH"),
+        pval_by = p.adjust(p = pval, method = "BY"),
+        pval_holm = p.adjust(p = pval, method = "holm")
+        ) %>%
+    ungroup()
+
+# using the binary segmentation method to identify onsets and offsets
+res <- cpt.meanvar(data = tests_results$tval, method = "BinSeg", Q = 2)
+
+onset_offset_cpt <- data.frame(
+    split_id = "full",
+    cluster_id = 1,
+    onset = tests_results$time[res@cpts[1]],
+    offset = tests_results$time[res@cpts[2]]
+    ) %>%
+    mutate(method = "cpt") %>%
+    pivot_longer(cols = onset:offset) %>%
+    select(split_id, method, cluster_id, onset_offset = name, value) %>%
+    data.frame()
+
+# finding clusters
+onset_offset_raw_p <- find_clusters(
+    data = tests_results %>% mutate(pval = pval * (-1) ) |> select(time, value = pval),
+    threshold = -alpha_level
+    ) %>%
+    mutate(split_id = "full") %>%
+    select(split_id, cluster_id, onset = cluster_onset, offset = cluster_offset) %>%
+    mutate(method = "raw_p") %>%
+    pivot_longer(cols = onset:offset) %>%
+    select(split_id, method, cluster_id, onset_offset = name, value) %>%
+    data.frame()
+
+onset_offset_pval_bh <- find_clusters(
+    data = tests_results %>% mutate(pval = pval_bh * (-1) ) |> select(time, value = pval),
+    threshold = -alpha_level
+    ) %>%
+    mutate(split_id = "full") %>%
+    select(split_id, cluster_id, onset = cluster_onset, offset = cluster_offset) %>%
+    mutate(method = "pval_bh") %>%
+    pivot_longer(cols = onset:offset) %>%
+    select(split_id, method, cluster_id, onset_offset = name, value) %>%
+    data.frame()
+
+onset_offset_pval_by <- find_clusters(
+    data = tests_results %>% mutate(pval = pval_by * (-1) ) |> select(time, value = pval),
+    threshold = -alpha_level
+    ) %>%
+    mutate(split_id = "full") %>%
+    select(split_id, cluster_id, onset = cluster_onset, offset = cluster_offset) %>%
+    mutate(method = "pval_by") %>%
+    pivot_longer(cols = onset:offset) %>%
+    select(split_id, method, cluster_id, onset_offset = name, value) %>%
+    data.frame()
+
+onset_offset_pval_holm <- find_clusters(
+    data = tests_results %>% mutate(pval = pval_holm * (-1) ) |> select(time, value = pval),
+    threshold = -alpha_level
+    ) %>%
+    mutate(split_id = "full") %>%
+    select(split_id, cluster_id, onset = cluster_onset, offset = cluster_offset) %>%
+    mutate(method = "pval_holm") %>%
+    pivot_longer(cols = onset:offset) %>%
+    select(split_id, method, cluster_id, onset_offset = name, value) %>%
+    data.frame()
+
+# running the MNE cluster-based permutation
+p_values_cluster_mass <- freq_stats_cluster_matrix(
+    X = data_matrix - chance_level,
+    timesteps = unique(decoding_summary_df$time),
+    cluster_type = "mass",
+    alpha_level = alpha_level,
+    verbose = FALSE
+    )
+p_values_tfce <- freq_stats_cluster_matrix(
+    X = data_matrix - chance_level,
+    timesteps = unique(decoding_summary_df$time),
+    cluster_type = "tfce",
+    alpha_level = alpha_level,
+    verbose = FALSE
+    )
+
+onset_offset_cluster_mass <- find_clusters(
+    data = p_values_cluster_mass %>% mutate(pval = pval*(-1) ) %>% select(time, value = pval),
+    threshold = -alpha_level
+    ) %>%
+    mutate(split_id = "full") %>%
+    select(split_id, cluster_id, onset = cluster_onset, offset = cluster_offset) %>%
+    mutate(method = "cluster_mass") %>%
+    pivot_longer(cols = onset:offset) %>%
+    select(split_id, method, cluster_id, onset_offset = name, value) %>%
+    data.frame()
+
+onset_offset_cluster_tfce <- find_clusters(
+    data = p_values_tfce %>% mutate(pval = pval*(-1) ) %>% select(time, value = pval),
+    threshold = -alpha_level
+    ) %>%
+    mutate(split_id = "full") %>%
+    select(split_id, cluster_id, onset = cluster_onset, offset = cluster_offset) %>%
+    mutate(method = "cluster_tfce") %>%
+    pivot_longer(cols = onset:offset) %>%
+    select(split_id, method, cluster_id, onset_offset = name, value) %>%
+    data.frame()
+
+# binding all results together
+full_reliability_results <- bind_rows(
+    onset_offset_brms, onset_offset_raw_p, onset_offset_pval_bh,
+    onset_offset_pval_by, onset_offset_pval_holm, onset_offset_cpt,
+    onset_offset_cluster_mass, onset_offset_cluster_tfce
+    ) %>%
+    mutate(sesoi = full_sesoi) %>%
     pivot_wider(names_from = onset_offset, values_from = value) %>%
@@
                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-                "TFCE", "Change point", "BGAM (new method)"
+                "TFCE", "Change point", "BGAM (our method)"
                 )
@@
     mutate(y_offset = 0.5 - as.numeric(factor(method) ) * 0.01 - 0.01)
+
+# importing results
+# reliability_results <- readRDS(file = "results/reliability_results_100splits.rds")
+
+# loading all RDS files
+# files <- list.files(
+#     path = "results/reliability_cluster_results",
+#     pattern = ".rds", 
+#     full.names = TRUE
+#     )
+# files <- list.files(
+#     # path = "reliability_cluster_results",
+#     path = "reliability_jaccard_cluster_results",
+#     pattern = ".rds", 
+#     full.names = TRUE
+#     )
+
+# importing the data splits
+# data_splits <- list.files(
+#     path = "reliability_data_splits",
+#     pattern = ".csv",
+#     full.names = TRUE
+#     )
+
+# sorting files numerically by the number in the filename
+# files <- files[order(readr::parse_number(files) )]
+# data_splits <- data_splits[order(readr::parse_number(data_splits) )]
+
+# reading and binding all files
+# reliability_results <- map2_dfr(
+#     .x = files,
+#     .y = seq_along(files),
+#     .f = ~ readRDS(.x)
+#     # .f = ~ readRDS(.x) |> mutate(sim_id = .y)
+#     # .f = function (file, index) {
+#     #     offset <- (index - 1) * splits_per_file
+#     #     readRDS(file) %>% mutate(split_id = as.numeric(split_id) ) %>%
+#     #         mutate(split_id = split_id + offset) %>%
+#     #         mutate(split_id = formatC(x = split_id, width = 3, flag = 0) ) 
+#     #     }
+#     )
+
+# checking errors
+# reliability_results %>%
+#     filter(!is.na(error) )
+
+# removing NAs
+# reliability_results <- reliability_results %>%
+#     # select(-error, -stringAsFactors) %>%
+#     na.omit()
+
+# sanity checks (N=1122)
+# unique(reliability_results$split_id)
+# n_distinct(reliability_results$split_id)
+
+# reading and binding all files, adding a split_id
+# splits_results <- map2_dfr(
+#     .x = data_splits,
+#     .y = seq_along(data_splits),
+#     .f = ~ read.csv(.x)
+#     ) %>%
+#     # mutate(array_id = as.numeric(array_id) ) %>%
+#     mutate(split_id = paste(array_id, sprintf("s%03d", split_number), sep = "_") ) %>%
+#     # removing unused columns
+#     select(-split_number, -array_id)
+
+# or reading splits list directly
+# splits_results <- read.csv(file = "uniform_jaccard_data_splits.csv")
+# splits_results <- read.csv(file = "uniform_jaccard_data_splits_100pairs.csv")
+# splits_results <- read.csv(file = "data/uniform_jaccard_data_splits_1000pairs.csv")
+
+# extracting the full data estimates
+# full_reliability_results <- reliability_results %>%
+#     # filter(split_id == max(split_id) ) %>%
+#     # filter(str_detect(string = split_id, pattern = "_s051") ) %>%
+#     # dplyr::filter(split_id == "a001_s051") %>%
+#     dplyr::filter(split_id == "1_full") %>%
+#     # keeping only the first and last timestep
+#     # group_by(split_id, method, onset_offset) %>%
+#     # summarise(
+#     #     value = if (onset_offset[1] == "onset") min(value) else max(value),
+#     #     .groups = "drop"
+#     #     ) %>%
+#     # ungroup() %>%
+#     # pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
+#     # data.frame() %>%
+#     # select(-split_id) %>%
+#     # separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
+#     pivot_wider(names_from = onset_offset, values_from = value) %>%
+#     mutate(
+#         method = factor(
+#             x = method,
+#             levels = c(
+#                 "raw_p", "pval_bh", "pval_by", "pval_holm", "cluster_mass",
+#                 "cluster_tfce", "cpt", "brms"
+#                 ),
+#             labels = c(
+#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+#                 "TFCE", "Change point", "BGAM (our method)"
+#                 )
+#             )
+#         ) %>%
+#     mutate(y_offset = 0.5 - as.numeric(factor(method) ) * 0.01 - 0.01)
 
@@
 
+We then assessed the sensitivity of the various methods using a form of permutation-based sensitivity study, which consisted of the following steps.
+First, we created a large number of split halves of the data, that is, subsets of the dataset containing only 16 out of 32 participants.
+For each possible pair of subsets, we have 16 possible levels of overlap/similarity that can be quantified using the Jaccard index, ranging from 0 (perfectly disjoints subsets) to $\approx0.88$ (identical subsets except one participant).
+For each of these 16 levels of Jaccard similarity, we created 1,000 pairs of subsets, resulting in 16,000 pairs of subsets in total.
+For each of these pairs, we estimated the onset and offset according to each method and computed the absolute difference in onset/offset estimates.
+Finally, we estimated the Spearman's rank correlation coefficient (which quantifies the strength of a monotonic relation between two variables) between the Jaccard similarity and the absolute difference in onset/offset estimates.
+The rational for this procedure is that sensitive methods should produce similar onset/offset estimates for similar subsets and dissimilar onset/offset estimates for dissimilar subsets.
+The results of this procedure are summarised in @fig-splits-similarity.
+
+<!--
+
+Finally, we fitted a polynomial (cubic) regression to predict the absolute difference in onset/offset estimates according to the Jaccard similarity.
+This analysis allowed us to disentangle two properties of each method: i) whether it produces similar onset/offset estimates for similar subsets (in the limit case where subsets are identical, this is reflected by the intercept parameter), and ii) how sensitive each method is to variation in subsets' similarity (which is reflected by the two slope parameters, for the linear and cubic terms, respectively).
+In simple terms, sensitive methods should produce similar onset/offset estimates for similar subsets and dissimilar onset/offset estimates for dissimilar subsets.
+The results of this procedure are summarised in @fig-splits-similarity.
+
+This figure reveals that the `BGAM` *onset and offset* estimates on each split are the closest to the estimates from the full dataset on average (0ms difference for the onset estimate and 5ms difference for the offset estimate).
+The `Raw p-value` method has similar performance for the offset, but given the aberrant estimates it produces (cf.
+@fig-onset-offset), the fact that it is consistent between data splits and the full dataset is not convincing on its own.
+The `Change point` method also has a very good performance (i.e., very low difference between split estimates and full estimates), but produces too short cluster of significant decoding performance (cf.
+@fig-onset-offset).[^3] Overall, the figure reveals that for all other methods, split datasets produce later onset estimates and earlier offset estimates (compared to the estimates from the model fitted on the full dataset).
+These results highlight some desirable properties for a method aiming to precisely and reliably estimate the onset and offset of M/EEG effects, namely, it should i) have good asymptotic properties on simulated data, ii) provide sensible identified clusters in actual data, and iii) provide reliable/stable estimates across various partitions of the data.
+All these desiderata seem to be fulfilled by the BGAM approach.
+
+-->
+
+```{r fig-splits-similarity, eval = TRUE, echo = FALSE, fig.width = 12, fig.asp = 0.6, fig.cap = "Relation between data subsets' dissimilarity (x-axis) and difference in onset (blue) and offset (orange) estimates (y-axis) according to each method."}
+# reading splits list directly
+# splits_results <- read.csv(file = "uniform_jaccard_data_splits.csv")
+# splits_results <- read.csv(file = "uniform_jaccard_data_splits_100pairs.csv")
+splits_results <- read.csv(file = "data/uniform_jaccard_data_splits_1000pairs.csv")
+
+# listing all RDS files
+files <- list.files(
+    path = "results/reliability_jaccard_cluster_results",
+    pattern = ".rds", 
+    full.names = TRUE
+    )
+
+# sorting files numerically by the number in the filename
+files <- files[order(readr::parse_number(files) )]
+
+# reading and binding all files
+reliability_results <- map2_dfr(
+    .x = files,
+    .y = seq_along(files),
+    .f = ~ readRDS(.x)
+    )
+
+# checking errors
+# reliability_results %>%
+#     filter(!is.na(error) )
+
+# removing NAs
+reliability_results <- reliability_results %>%
+    # select(-error, -stringAsFactors) %>%
+    na.omit()
+
+# This analysis assesses whether the method is "self-consistent" across data splits...
+# Step 1: creating a list of participants per split_id
+# split_participants <- splits_results %>%
+#     rename(split_id = split_number) %>%
+#     group_by(split_id) %>%
+#     dplyr::filter(group == "A") %>%
+#     summarise(participants = list(participant) ) %>%
+#     ungroup()
+
+# Step 2: creating all pairwise combinations of split_ids
+# split_pairs <- expand.grid(
+#     split_id1 = split_participants$split_id,
+#     split_id2 = split_participants$split_id,
+#     stringsAsFactors = FALSE
+#     ) %>%
+#     # avoiding duplicates and self-comparison
+#     dplyr::filter(split_id1 < split_id2)
+
+# Step 3: joining the participant lists for each pair
+# split_pairs <- split_pairs %>%
+#     left_join(split_participants, by = c("split_id1" = "split_id") ) %>%
+#     rename(participants1 = participants) %>%
+#     left_join(split_participants, by = c("split_id2" = "split_id") ) %>%
+#     rename(participants2 = participants)
+
+# Step 4: computing overlap as size of intersection / union
+# see also https://www.rdocumentation.org/packages/genieclust/versions/1.1.6/topics/compare_partitions
+# split_pairs <- split_pairs %>%
+#     mutate(
+#         intersection = map2_int(participants1, participants2, ~length(intersect(.x, .y) ) ),
+#         union = map2_int(participants1, participants2, ~length(union(.x, .y) ) ),
+#         jaccard_similarity = intersection / union,
+#         # jaccard_similarity = map2_dbl(participants1, participants2, ~jaccard(.x, .y) )
+#         # ari = map2(participants1, participants2, ~ARI(.x[[1]], .y[[1]]) ),
+#         # nmi = map2_dbl(participants1, participants2, ~NMI(.x[[1]], .y[[1]]) ),
+#         # ami = map2_dbl(participants1, participants2, ~AMI(.x[[1]], .y[[1]]), .progress = TRUE)
+#         )
+
+# checking the results
+# sort(unique(split_pairs$intersection) )
+# hist(split_pairs$intersection)
+
+# plotting it
+# split_pairs %>%
+#     ggplot(aes(x = intersection, y = jaccard_similarity) ) +
+#     geom_point() +
+#     geom_smooth(method = "lm") +
+#     theme_bw()
+
+##########################################################
+# now computing similarity of onset and offset estimates #
+##########################################################
+
+# Step 1: creating a list of participants per split_id
+# reliability_splits <- reliability_results %>%
+#     # filter(str_detect(string = split_id, pattern = "full", negate = TRUE) ) %>%
+#     # mutate(split_id = str_sub(string = split_id, start = 6, end = -1) )
+#     dplyr::filter(split_id != "a001_s051") %>%
+#     # keeping only the first and last timestep
+#     group_by(split_id, method, onset_offset) %>%
+#     summarise(
+#         value = if (onset_offset[1] == "onset") min(value) else max(value),
+#         .groups = "drop"
+#         ) %>%
+#     ungroup()
+
+# Step 2: creating all pairwise combinations of split_ids
+# and computing the absolute difference of estimated onset/offset
+# reliability_splits_results <- reliability_splits %>%
+#   # joining the data frame with itself by method and onset_offset
+#   inner_join(
+#       reliability_splits,
+#       by = c("method", "onset_offset"),
+#       suffix = c("_1", "_2")
+#       ) %>%
+#   # keeping only unique pairs (avoid self-pairs and duplicates)
+#   dplyr::filter(split_id_1 < split_id_2) %>%
+#   # computing the absolute difference in value
+#   mutate(value_diff = abs(value_1 - value_2) ) %>%
+#   select(method, onset_offset, split_id_1, split_id_2, value_diff)
+
+# joining these results with splits similarity
+# head(reliability_splits_results)
+# full_df <- split_pairs %>%
+#     select(split_id_1 = split_id1, split_id_2 = split_id2, jaccard_similarity) %>%
+#     # head()
+#     inner_join(reliability_splits_results, by = c("split_id_1", "split_id_2") )
+
+# checking similarity values and effectives
+# full_df %>%
+#     summarise(
+#         effective_sim = n(),
+#         .by = c(onset_offset, method, jaccard_similarity)
+#         ) %>%
+#     arrange(jaccard_similarity)
+
+########################################################################
+# or working directly from pre-existing jaccard similarity csv file
+######################################################################
+
+reliability_results2 <- reliability_results %>%
+    # dplyr::filter(str_detect(string = split_id, pattern = "full", negate = TRUE) ) %>%
+    # keeping only the first and last timestep
+    group_by(split_id, method, onset_offset) %>%
+    summarise(
+        value = if (onset_offset[1] == "onset") min(value) else max(value),
+        .groups = "drop"
+        ) %>%
+    ungroup() %>%
+    mutate(split_id = str_sub(string = split_id, start = 6, end = -3) ) %>%
+    # removing split with error
+    # filter(split_id != "s00171") %>%
+    group_by(split_id, method, onset_offset) %>%
+    mutate(value_diff = abs(diff(value) ) ) %>%
+    # mutate(value_diff = diff(value) ) %>%
+    # mutate(value_diff = diff(value)^2) %>%
+    # rmse = sqrt(mean(error)^2),
+    ungroup() %>%
+    select(-value) %>%
+    distinct()
+
+splits_results2 <- splits_results %>%
+    mutate(split_id = formatC(x = split_id, width = 5, flag = 0) ) %>%
+    mutate(split_id = paste0("s", split_id) ) %>%
+    select(-group, -participant) %>%
+    distinct()
+
+full_df <- full_join(reliability_results2, splits_results2, by = "split_id")
+
+# sanity checks
+# table(full_df$jaccard)
+# unique(full_df$jaccard)
+# n_distinct(full_df$jaccard)
+# unique(splits_results2$jaccard)
+# unique(reliability_results2$split_id)
+
+# find rows with NAs
+# full_df[!complete.cases(full_df), ] %>% data.frame()
+# reliability_results2 %>% filter(split_id == "s00088")
+# reliability_results %>% filter(str_detect(string = split_id, pattern = "s00088") )
+
+# custom function to implement min-max scaling (i.e., between 0 and 1)
+# minMax <- function (x) {
+#     
+#     std_score <- (x - min(x) ) / (max(x) - min(x) )
+#     
+#     return (std_score)
+#     
+# }
+
+# polynomial regression degree
+# poly_reg_degree <- 2
+
+# or fitting a GAM for each combination
+# library(ggpmisc)
+# library(broom)
+# library(glue)
+# library(mgcv)
+# 
+# # 1. Nest data by method & onset_offset
+# model_df <- full_df %>%
+#     na.omit() %>%
+#     group_by(method, onset_offset, jaccard) %>%
+#     summarise(
+#         mean_value_diff = mean(x = value_diff, na.rm = TRUE),
+#         sem = sd(x = value_diff, na.rm = TRUE) / sqrt(n() )
+#         ) %>%
+#     ungroup() %>%
+#     mutate(
+#         method = factor(
+#             x = method,
+#             levels = c(
+#                 "raw_p", "pval_bh", "pval_by", "pval_holm", "cluster_mass",
+#                 "cluster_tfce", "cpt", "brms"
+#                 ),
+#             labels = c(
+#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+#                 "TFCE", "Change point", "BGAM (our method)"
+#                 )
+#             )
+#         ) %>%
+#     # mutate(jaccard = 1-jaccard) %>%
+#     filter(!is.na(mean_value_diff), !is.na(jaccard) ) %>%
+#     group_by(method, onset_offset) %>%
+#     nest()
+# 
+# # 2. Fit a GAM to each subset
+# model_df <- model_df %>%
+#     mutate(
+#         gam_model = map(
+#             data, ~ gam(mean_value_diff ~ s(jaccard),
+#                         # family = Gamma(link = "log"),
+#                         data = .x)
+#             )
+#         )
+# 
+# # 3. Generate pretty label with edf and R² for plotting
+# model_df <- model_df %>%
+#     mutate(
+#         eq_label = map2_chr(
+#             gam_model,
+#             data,
+#             ~ {
+#                 r2 <- summary(.x)$r.sq
+#                 edf <- summary(.x)$s.table[, "edf"] %>% round(2)
+#                 glue("edf = {edf}, R² = {round(r2, 2)}")
+#                 })
+#         )
+# 
+# # 4. Create predictions for plotting
+# model_df <- model_df %>%
+#   mutate(
+#     preds = map2(data, gam_model, ~ {
+#       tibble(
+#         jaccard = seq(min(.x$jaccard), max(.x$jaccard), length.out = 100)
+#       ) %>%
+#         mutate(mean_value_diff = predict(.y, newdata = ., type = "response"))
+#     })
+#   )
+# 
+# # 5. Unnest everything for ggplot
+# plot_data <- model_df %>%
+#   select(method, onset_offset, eq_label, preds) %>%
+#   unnest(preds)
+# 
+# # 6. Create base plot with data points and fitted GAM lines
+# full_df %>%
+#     na.omit() %>%
+#     # nrow()
+#     # dplyr::filter(jaccard < 1 & jaccard > 0) %>%
+#     group_by(method, onset_offset, jaccard) %>%
+#     summarise(
+#         mean_value_diff = mean(x = value_diff, na.rm = TRUE),
+#         # mean_value_diff = sqrt(mean(x = value_diff, na.rm = TRUE) ),
+#         sem = sd(x = value_diff, na.rm = TRUE) / sqrt(n() )
+#         ) %>%
+#     ungroup() %>%
+#     mutate(
+#         method = factor(
+#             x = method,
+#             levels = c(
+#                 "raw_p", "pval_bh", "pval_by", "pval_holm", "cluster_mass",
+#                 "cluster_tfce", "cpt", "brms"
+#                 ),
+#             labels = c(
+#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+#                 "TFCE", "Change point", "BGAM (our method)"
+#                 )
+#             )
+#         ) %>%
+#     ggplot(aes(x = 1-jaccard, y = mean_value_diff, colour = onset_offset) ) +
+#     geom_point(alpha = 0.6) +
+#     geom_line(data = plot_data, aes(y = mean_value_diff), size = 1.2) +
+#     geom_label(
+#         data = plot_data %>% group_by(method, onset_offset) %>% slice(1),
+#         aes(label = eq_label, x = 0.05, y = Inf),
+#         vjust = 1.2, hjust = 0,
+#         fill = "white", colour = "black",
+#         size = 3, inherit.aes = FALSE
+#         ) +
+#     facet_wrap(~method, nrow = 2) +
+#     theme_bw() +
+#         labs(
+#         x = "Jaccard dissimilarity",
+#         y = "Average absolute difference in onset/offset estimate (s)",
+#         colour = ""
+#         )
+
+# plotting the results
+full_df %>%
+    na.omit() %>%
+    mutate(jaccard_dis = 1-jaccard) %>%
+    # dplyr::filter(jaccard < 1 & jaccard > 0) %>%
+    # dplyr::filter(1-jaccard > 0 & 1-jaccard < 1) %>%
+    dplyr::filter(jaccard_dis > 0) %>%
+    group_by(method, onset_offset, jaccard_dis) %>%
+    summarise(
+        mean_value_diff = mean(x = value_diff, na.rm = TRUE),
+        sem = sd(x = value_diff, na.rm = TRUE) / sqrt(n() )
+        ) %>%
+    ungroup() %>%
+    mutate(
+        method = factor(
+            x = method,
+            levels = c(
+                "raw_p", "pval_bh", "pval_by", "pval_holm", "cluster_mass",
+                "cluster_tfce", "cpt", "brms"
+                ),
+            labels = c(
+                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
+                "TFCE", "Change point", "BGAM (our method)"
+                )
+            )
+        ) %>%
+    ggplot(
+        aes(
+            # x = minMax(1-jaccard_similarity),
+            # x = 1-jaccard_similarity,
+            # x = max(jaccard)-jaccard,
+            x = jaccard_dis,
+            y = mean_value_diff,
+            fill = onset_offset,
+            colour = onset_offset,
+            group = onset_offset
+            )
+        ) +
+    # geom_point(size = 0.5, alpha = 0.05) +
+    geom_smooth(
+        # method = "glm",
+        # method.args = list(family = Gamma(link = "inverse") ),
+        # method.args = list(family = Gamma(link = "log") ),
+        # method.args = list(family = inverse.gaussian() ),
+        method = "lm",
+        # method = "gam",
+        # formula = y ~ stats::poly(x, degree = 3, raw = TRUE),
+        # formula = y ~ stats::poly(x, degree = poly_reg_degree, raw = TRUE),
+        # formula = rank(y) ~ rank(x),
+        # formula = y ~ logit_scaled(x),
+        # method.args = list(family = "binomial"),
+        alpha = 0.2,
+        fullrange = TRUE
+        ) +
+    # geom_point() +
+    geom_pointrange(
+        aes(ymin = mean_value_diff - sem, ymax = mean_value_diff + sem),
+        fatten = 1
+        ) +
+    stat_cor(
+        # showing the R and R^2
+        # aes(label = paste(..r.label.., ..rr.label.., sep = "~`,`~") ),
+        aes(label = paste(after_stat(r.label), after_stat(rr.label), sep = "~`,`~") ),
+        method = "spearman",
+        # method = "pearson",
+        # method = "kendall",
+        digits = 3,
+        # label.x = 0,
+        # label.y = 1,
+        label.x.npc = "left",
+        label.y.npc = "top",
+        geom = "label",
+        colour = "white",
+        size = 3.5
+        ) +
+    # stat_regline_equation(
+    #     aes(label = paste(..eq.label.., ..rr.label.., sep = "~`,`~") ),
+    #     # formula = rank(y) ~rank(x),
+    #     # formula = y ~ logit_scaled(x),
+    #     formula = y ~ stats::poly(x = x, degree = poly_reg_degree, raw = TRUE),
+    #     # method = "glm",
+    #     # method.args = list(family = Gamma(link = "log")),
+    #     label.x = 0,
+    #     label.y.npc = "top",
+    #     geom = "label",
+    #     colour = "white",
+    #     size = 3.5
+    #     ) +
+    # scale_y_log10(guide = "axis_logticks", limits = c(NA, 0.5) ) +
+    scale_y_log10(guide = "axis_logticks", limits = c(NA, 1) ) +
+    facet_wrap(~method, nrow = 2) +
+    theme_bw(base_size = 12) +
+    scale_fill_manual(values = met.brewer(name = "Johnson", n = 2) ) +
+    scale_colour_manual(values = met.brewer(name = "Johnson", n = 2) ) +
+    labs(
+        x = "Jaccard dissimilarity",
+        y = "Average absolute difference in onset/offset estimate (s)",
+        fill = "",
+        colour = ""
+        )
+
+# saving the plot
+# ggsave(
+#     filename = "figures/splits_summary_similarity_100pairs.png",
+#     width = 16, height = 8, dpi = 200,
+#     device = "png"
+#     )
+```
+
+This figure shows that, among the methods that performed best in the simulation study (i.e., `Cluster mass`, `Change point`, and `BGAM`), onset estimates remain highly stable across subsets of participants with varying Jaccard similarity, it varies from around 1ms for most similar subsets to around 10ms for most dissimilar subsets.
+Additionally, for both onset and offset estimates, the average pairwise difference increases monotonically with Jaccard dissimilarity, as indicated by the Spearman's rank correlation coefficient.
+For the onset estimates, the `Holm`, `Cluster mass`, `TFCE`, and `BGAM` methods exhibit the strongest monotonic relation with subset similarity (all $\rho\text{s} > 0.9$), whereas for offset estimates, all methods demonstrate excellent performance (all $\rho\text{s} > 0.9$) with the `BGAM` method showing the highest sensitivity ($\rho\approx0.994$).
+However, given the aberrant clusters identified by the `Raw p-value`, `FDR BH95`, and `FDR BY01` methods (see @fig-onset-offset), their sensitivity to variation in subset similarity is not meaningful.
+
+<!--
+
 @fig-reliability shows the median difference between the onset and offset estimates from each data split and the onset and offset estimates from the full dataset (x-axis) along with the variance of its onset and offset estimates across data splits (error bar).
@@
 
-```{r fig-reliability, echo = FALSE, cache = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median error and median absolute deviation of the error for the onset (left) and offset (right) estimates according to each method.
+```{r fig-reliability, eval = FALSE, echo = FALSE, cache = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median error and median absolute deviation of the error for the onset (left) and offset (right) estimates according to each method.
 Methods are ordered from smallest (top) to largest (bottom) median absolute error separately for the onset and offset estimates."}
@@
     "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-    "TFCE", "Change point", "BGAM (new method)"
+    "TFCE", "Change point", "BGAM (our method)"
     )
@@
 
+-->
+
 # Discussion
 
-## Summary of the proposed approach
-
-Overall, before concluding on the onset/offset of effect based on the model, we need to ensure that the model provides a faithful description of the data-generating process (e.g., via posterior predictive checks etc)...
-
-The TFCE performs worse than the cluster-sum approach, which was anticipated by @rousselet_using_2025 based on the initial results of @smith2009...
-we did not include the cluster-depth algorithm [@frossard2022], as @rousselet_using_2025 already showed its performance were worse than the cluster mass algorithm...
-
-## Limitations and future directions
-
-As in previous simulation work [e.g., @rousselet2008; @sassenhagen2019], the present simulation results depend on various choices such as the specific cluster-forming algorithm and threshold, signal-to-noise ratio, negative impact of preprocessing steps (e.g., low-pass filter) on temporal resolution...
-note however, that the same caveats apply to all methods...
-
-Discussion about the nature of the model: we modelled the surface M/EEG signals, however, the true interests (probably) lie in the brain, that is, in the source space...
-we could build a "full" Bayesian model of the generated EEG signal (i.e., including hypotheses about the source and a forward model), but this model would become computationally heavier...
-
-The error properties depend on the threshold parameter, a value of 10 or 20 seems to be a reasonable default, but the optimal threshold parameter can be adjusted using split-half reliability assessment...
-also depends on `k`...
-
-Can be applied to any 1D timeseries (e.g., pupillometry, electromyography)... Extending the approach to spatiotemporal data (i.e., time + sensors) or spatiotemporal time-frequency 4D data...
-
-We kept the exemplary models simple, but can be extended by adding varying/random effects (intercept and slope) for item (e.g., word)...
-but also continuous predictors at the trial level, overlapping factors, etc...
+In brief, our results show that the model-based approach we introduced outperforms conventional methods such as cluster-based inference in identifying the onset and offset of M/EEG effects.
+Performance was assessed both on simulated data--allowing us to evaluate the method's ability to recover ground-truth onset and offset values--and on actual MEG data--allowing us to assess its sensitivity to realistic data properties (subset similarity).
+Together, these results highlight desirable properties for any method aiming to precisely and reliably estimate the onset and offset of M/EEG effects: it should i) recover true onsets and offsets in simulation (good asymptotic behaviour), ii) identify clusters that are interpretable and consistent in empirical data, and iii) show sensitivity to subtle changes in the data.
+Our approach meets all three of these desiderata.
+
+As with previous simulation studies [e.g., @rousselet2008; @sassenhagen2019], results inevitably depend on design choices, including the specific cluster-forming algorithm and threshold (for cluster-based methods), the signal-to-noise ratio, and the potential degradation of temporal resolution introduced by preprocessing steps such as low-pass filtering.
+However, these constraints apply equally to all methods tested, so relative differences in performance remain meaningful.
+
+**I am not sure what to do with this paragraph (not sure how to discuss these results/observations):** Interestingly, the TFCE method performed worse than the traditional cluster-sum approach, consistent with the predictions of @rousselet_using_2025 based on the original findings of @smith2009.
+We also found a striking overlap in the clusters identified by the `Holm` and `TFCE` procedures (cf.
+@fig-onset-offset and @fig-splits-similarity).
+Whereas these two approaches are conceptually distinct; `Holm` controlling the family-wise error rate through sequential *p*-value adjustment, `TFCE` enhancing signal based on spatiotemporal support; their similarity in practice may arise because both effectively prioritise extended, moderately strong effects over isolated high-intensity points.
+This convergence warrants further methodological work.
+We chose not to include the cluster-depth algorithm [@frossard2022], as prior work already shows its performance to be inferior to cluster-mass approaches in similar contexts [@rousselet_using_2025].
+
+A critical consideration for any model-based approach is that the model must adequately capture the underlying data-generating process.
+Misspecified models are likely to produce biased or unreliable onset/offset estimates.
+This underscores the importance of thorough model diagnostics, including posterior predictive checks, fit assessments, and model comparison [@gelman2020].
+As with any inferential modelling framework, transparency in assumptions and rigorous validation procedures are indispensable to ensure replicability and generalisability of the results.
+
+Another important methodological consideration concerns the selection of model hyperparameters, such as the threshold for posterior odds and the number of basis functions.
+Although our simulations suggest that these parameters influence the precision and reliability of onset and offset estimates, optimal values may vary depending on the signal's temporal dynamics and signal-to-noise characteristics.
+Future work could explore principled approaches to hyperparameter tuning, including cross-validation or fully Bayesian model selection using tools such as leave-one-out cross-validation (LOO-CV) or Bayes factors [@gelman2020].
+We provide initial guidance in @apx-basis and advocate for future development of adaptive heuristics to support flexible yet parsimonious model specification.
+
+Currently, our approach estimates temporal effects independently at each sensor (1D temporal data).
+Extending the current framework to incorporate additional temporal (see @apx-2D) or spatial dimensions would improve both sensitivity and interpretability.
+Such extensions could draw on methods from spatial epidemiology and geostatistics using either GAMMs or approximate Gaussian process regression [e.g., @rasmussen2005; @riutort-mayol_practical_2023], depending on computational feasibility.
+
+To facilitate adoption, we developed the `neurogram` open-source `R` package [@neurogam], which implements the proposed method using `brms`.
+The package integrates seamlessly with `MNE-Python` [@gramfort2013], enabling researchers to process M/EEG data in `Python` and import them directly into `R` for model-based inference without cumbersome data export.
+This interoperability, described in @apx-package, is designed to encourage broader use of model-based approaches in cognitive neuroscience.
+
+In conclusion, we introduced a model-based approach for estimating the onset and offset of M/EEG effects.
+Across simulated and empirical datasets, we showed that the method yields more precise and sensitive estimates than conventional cluster-based approaches.
+These results highlight the potential of flexible, model-based alternatives for characterising time-resolved neural dynamics, particularly in applications where accurate temporal localisation is critical.
 
@@
 
+\setcounter{secnumdepth}{0}
+
 # Data and code availability
 
-The simulation results as well as the `R` code to reproduce the simulations are available on GitHub: <https://github.com/lnalborczyk/brms_meeg>.
+The simulation results as well as the `R` code to reproduce the simulations are available at: <https://github.com/lnalborczyk/brms_meeg>.
 The `neurogam R` package is available at <https://github.com/lnalborczyk/neurogam>.
@@
     output = "paragraph",
-    omit = c("permuco", "reticulate"),
+    omit = c("bamlss", "changepoint", "permuco", "reticulate"),
     out.dir = getwd(),
-    dependencies = FALSE
+    dependencies = FALSE,
+    quiet = TRUE
+    # progress = FALSE,
+    # errors = "ignored"
     )
@@
 
-# Appendix
-
-# Application to 2D time-resolved decoding results (cross-temporal generalisation) {#sec-2D}
-
-Assume we have M/EEG data and we have conducted cross-temporal generalisation analyses [@king2014].
-As a result, we have a 2D matrix where each element contains the decoding accuracy (e.g., ROC AUC) of a classifier trained at timestep $\text{training}_{i}$ and tested at timestep $\text{testing}_{j}$ (@fig-sim-timegen).
-
-```{r fig-timegen-tickertape, eval = FALSE, echo = FALSE, dpi = 300, out.width = "75%", fig.cap = "Exemplary group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
+# Application to 2D time-resolved decoding results (cross-temporal generalisation) {#apx-2D}
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+We conducted a cross-temporal generalisation analysis of the decoding data from @nalborczyk:inprep, in which we assessed the performance of classifiers trained and tested at various timesteps of the trial [@king2014].
+This analysis was performed at the participant level, resulting in a 2D matrix where each element contains the decoding accuracy (ROC AUC) of a classifier trained at timestep $\text{training}_{i}$ and tested at timestep $\text{testing}_{j}$ for each participant (@fig-timegen).
+
+```{r fig-timegen, eval = TRUE, echo = FALSE, dpi = 300, out.width = "75%", fig.cap = "Group-level average cross-temporal generalisation matrix of decoding performance [data from @nalborczyk:inprep]."}
 # importing the MEG decoding data
@@
 data <- np$load("data/group_level_decoding_4l_2s_vs_7l_2s_visual_blocks_temporal_generalisation_all.npy")
-# data <- np$load("data/group_level_decoding_7l_2s_vs_7l_3s_visual_blocks_temporal_generalisation_all.npy")
 
@@
 data_r <- py_to_r(data)
-str(data_r)
-dim(data_r)
+
+# sanity checks
+# str(data_r)
+# dim(data_r)
 
@@
 # checking new dimensions
-dim(data_cut)
-length(time_points_cut)
+# dim(data_cut)
+# length(time_points_cut)
 
@@
 
-```{r fig-sim-timegen, echo = FALSE, dpi = 300, out.width = "75%", fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
+```{r fig-sim-timegen, eval = FALSE, echo = FALSE, dpi = 300, out.width = "75%", fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
 # number of participants
@@
 
-To model cross-temporal generalisation matrices of decoding performance (ROC AUC), we extended the initial (decoding) GAM to take into account the bivariate temporal distribution of AUC values, thus producing naturally smoothed estimates (timecourses) of AUC values and posterior probabilities.
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+To model cross-temporal generalisation matrices of decoding performance (ROC AUC), we extended our initial BGAM to take into account the bivariate temporal distribution of AUC values, thus producing naturally smoothed estimates (timecourses) of AUC values and posterior probabilities.
 This model can be written as follows:
@@
 
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
 where we assume that AUC values come from a $\mathrm{Beta}$ distribution with two parameters $\mu$ and $\phi$.
@@
 
-Note that $b_{m}(,)$ is a basis function that maps $R \times R \rightarrow R$.
-A popular bivariate basis function uses *thin-plate splines* [@wood2003], which extend to $\mathbf{s}_{i} \in \mathbb{R}^{d}$ and $\partial l_{g}$ penalties.
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+A popular bivariate basis function uses *thin-plate splines* [@wood2003].
 These splines are designed to interpolate and approximate smooth surfaces over two dimensions (hence the "bivariate" term).
@@
 
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
 using the the radial basis function given by:
@@
 
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
 where $\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|$ is the Euclidean distance between the covariate $\mathbf{s}_{i}$ and the knot location $\mathbf{k}_{m}$.
-We fitted this model using `brms`...
-using the `t2()` smooth constructor with full penalties [@pedersen_hierarchical_2019]...
-
-```{r gam-timegen, eval = FALSE, echo = TRUE, results = "hide"}
+We fitted this model using `brms` and the `t2()` tensor product smooth constructor with full penalties [@pedersen_hierarchical_2019].
+We ran eight MCMCs to approximate the posterior distribution, including each 5000 iterations and a warmup of 2000 iterations, yielding a total of $8 \times (5000-2000) = 24000$ posterior samples to be used for inference.
+
+```{r gam-timegen, eval = TRUE, echo = TRUE, results = "hide"}
 # fitting a GAM with two temporal dimensions
 timegen_gam <- brm(
-    # 2D thin-plate spline (tp)
-    # auc ~ t2(train_time, test_time, bs = "tp", k = 10),
-    auc ~ t2(train_time, test_time, bs = "tp", k = 20),
+    # 2D thin-plate spline (tp) with full penalties
+    auc ~ t2(train_time, test_time, bs = "tp", k = 30, full = TRUE),
     data = timegen_data,
-    family = Beta(),
-    warmup = 2000,
+    family = brms::Beta(),
+    warmup = 1000,
     iter = 5000,
@@
     cores = 8,
-    # file = "models/timegen_gam_t2.rds" # k = 10
-    file = "models/timegen_gam_t2_k20.rds" # k = 20
-    )
-
-# fitting a GAM with two temporal dimensions to empirical data
-timegen_gam <- brm(
-    # 2D thin-plate spline (tp)
-    # auc ~ t2(train_time, test_time, bs = "tp", k = 30),
-    auc ~ t2(train_time, test_time, bs = "tp", k = 20, full = TRUE),
-    data = timegen_data,
-    family = Beta(),
-    warmup = 1000,
-    iter = 3000,
-    chains = 8,
-    cores = 8,
-    # file = "models/timegen_gam_meg_t2_k30.rds"
-    file = "models/timegen_gam_meg_t2_k20_full.rds"
-    )
-
-# retrieving the data
-# timegen_data <- timegen_gam$data
-
-# PPCs
-# pp_check(timegen_gam)
+    control = list(adapt_delta = 0.95, max_treedepth = 15),
+    backend = "cmdstanr",
+    stan_model_args = list(stanc_options = list("O1") ),
+    file = "models/timegen_gam_meg_t2_k30_full.rds"
+    )
 ```
 
-```{r gp-timegen, eval = FALSE, echo = FALSE, results = "hide"}
-# fitting a GP with two temporal dimensions
-timegen_gp <- brm(
-    auc ~ gp(train_time, test_time, k = 20),
-    data = timegen_data,
-    family = Beta(),
-    control = list(adapt_delta = 0.95),
-    iter = 2000,
-    chains = 4,
-    cores = 4,
-    file = "models/timegen_gp.rds"
-    )
-```
-
-```{r gam-timegen-post-preds, eval = FALSE, echo = FALSE, cache = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Simulated data with thresholds derived from the BGAM (left), predicted AUC values (middle), and posterior probabily ratios of decoding accuracy being above chance level (right) according to the bivariate BGAM (for k=10, top row, and k=20, bottom row)."}
+```{r fig-gam-timegen-post-preds, eval = TRUE, echo = FALSE, cache = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Predicted AUC values with threshold (left) and posterior odds of decoding accuracy being above chance (right) according to the bivariate BGAM."}
 # defining a function to plot posterior predictions
-plot_post_timegen <- function (model, data = NULL, chance_level = 0.5, sesoi = 0, thresholds = c(10, 100) ) {
+plot_post_timegen <- function (
+        model, data = NULL, chance_level = 0.5,
+        sesoi = 0, thresholds = c(10, 100)
+        ) {
 
@@
     # defining a sequence of x values to make predictions
-    # preds_conds <- crossing(
-    #     train_time = unique(data$train_time)[c(TRUE, FALSE)],
-    #     test_time = unique(data$test_time)[c(TRUE, FALSE)]
-    #     )
     preds_conds <- crossing(
@@
 
+    # retrieving posterior samples
+    posterior_samples <- posterior_epred(object = model, newdata = preds_conds)
+
+    # computing the probability that y is above 0
+    prob_y_above_0 <- preds_conds %>%
+        mutate(post_prob = colMeans(posterior_samples) ) %>%
+        mutate(m = colMeans(posterior_samples > (chance_level + sesoi) ) ) %>%
+        mutate(prob_ratio = m / (1 - m) ) %>%
+        mutate(prob_ratio = pmin(prob_ratio, ndraws(model) ) ) %>%
+        mutate(prob_ratio = pmax(prob_ratio, 1 / ndraws(model) ) )
+
+    # plotting cross-temporal generalisation matrix
+    # data_plot <- data %>%
+    #     summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>%
+    #     ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
+    #     geom_tile() +
+    #     geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
+    #     geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
+    #     geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
+    #     scale_x_continuous(expand = c(0, 0) ) +
+    #     scale_y_continuous(expand = c(0, 0) ) + 
+    #     coord_fixed() +
+    #     scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC", limits = fill_limits) +
+    #     labs(x = "Testing time (s)", y = "Training time (s)")
+
     # plotting cross-temporal generalization matrix
@@
         geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
+        geom_contour(
+            data = prob_y_above_0,
+            aes(x = train_time, y = test_time, z = prob_ratio),
+            breaks = thresholds,
+            color = "black",
+            linewidth = 1,
+            inherit.aes = FALSE
+            ) +
         scale_x_continuous(expand = c(0, 0) ) +
@@
         coord_fixed() +
-        # scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
         scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC", limits = fill_limits) +
         labs(x = "Testing time (s)", y = "Training time (s)")
-
-    # retrieving posterior samples
-    posterior_samples <- posterior_epred(object = model, newdata = preds_conds)
-
-    # computing the probability that y is above 0
-    prob_y_above_0 <- preds_conds %>%
-        mutate(post_prob = colMeans(posterior_samples) ) %>%
-        mutate(m = colMeans(posterior_samples > (chance_level + sesoi) ) ) %>%
-        mutate(prob_ratio = m / (1 - m) ) %>%
-        mutate(prob_ratio = ifelse(prob_ratio == Inf, ndraws(model), prob_ratio) ) %>%
-        mutate(prob_ratio = ifelse(prob_ratio == 0, 1 / ndraws(model), prob_ratio) )
 
@@
             midpoint = 0,
-            name = "Post. prob. ratio",
+            name = "Posterior odds",
             trans = "log10"
@@
 
-    # plotting cross-temporal generalisation matrix
-    threshold_data_plot <- data %>%
-        summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>%
-        ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
-        geom_tile() +
-        geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
-        geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
-        geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
-        scale_x_continuous(expand = c(0, 0) ) +
-        scale_y_continuous(expand = c(0, 0) ) + 
-        coord_fixed() +
-        # scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
-        scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC", limits = fill_limits) +
-        labs(x = "Testing time (s)", y = "Training time (s)") +
-        geom_contour(
-            data = prob_y_above_0,
-            aes(x = train_time, y = test_time, z = prob_ratio),
-            breaks = thresholds,
-            color = "black",
-            linewidth = 1,
-            inherit.aes = FALSE
-            )
-
     # combining the plots
-    threshold_data_plot + post_preds_plot + prob_above_chance_plot
+    post_preds_plot + prob_above_chance_plot
 
@@
 
-# importing the models
-bgam_k10 <- readRDS(file = "models/timegen_gam_t2.rds")
-bgam_k20 <- readRDS(file = "models/timegen_gam_t2_k20.rds")
-
 # computing the SESOI as the baseline SD
-baseline_sd <- timegen_data %>%
-    filter(train_time < 0 & test_time < 0.2) %>%
-    summarise(sd(auc) ) %>% pull()
-
-# posterior predictive checks
-# pp_check(bgam_k10)
-# pp_check(bgam_k20)
-
-# plotting predictions for each model
-k10_plot <- plot_post_timegen(
-    model = bgam_k10,
-    # model = timegen_gam,
-    chance_level = 0.5,
-    sesoi = baseline_sd,
-    # thresholds = c(10, 100)
-    # thresholds = 10
-    thresholds = 20
-    )
-
-k20_plot <- plot_post_timegen(
-    model = bgam_k20,
-    chance_level = 0.5,
-    sesoi = baseline_sd,
-    # thresholds = c(10, 100),
-    # thresholds = 10
-    thresholds = 20
-    )
-
-# combining the plots
-k10_plot / k20_plot
+# baseline_sd <- timegen_data %>%
+#     filter(train_time < 0 & test_time < 0) %>%
+#     summarise(sd(auc) ) %>%
+#     pull()
+
+# computing the SESOI as the baseline upper 99% quantile
+baseline_upper_quantile <- timegen_data %>%
+    filter(train_time < 0 & test_time < 0) %>%
+    summarise(quantile(x = auc, probs = 0.99) ) %>%
+    pull()
+
+# computing predictions and posterior odds contours
+post_preds_plot <- plot_post_timegen(
+    model = timegen_gam,
+    data = timegen_data,
+    # chance_level = 0.5,
+    # sesoi = baseline_sd,
+    chance_level = baseline_upper_quantile,
+    sesoi = 0,
+    # thresholds = c(20, 50, 100)
+    thresholds = 10
+    )
+
+# showing the plot
+post_preds_plot
 ```
 
-Could be extended to spatial and temporal dimensions with formulas such as `te(x, y, Time, d = c(2, 1) )`...
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+@fig-gam-timegen-post-preds shows the predictions from the model (left) superimposed with the identified cluster as defined by thresholding the posterior odds (right).
+Notably, this model could be extended to a multilevel bivariate GAM via `t2(train_time, test_time, participant, bs = c("tp", "tp", "re"), m = 2, full = TRUE)` and could be generalised to account for both spatial (`x1` and `x2`) and temporal (`time`) dimensions with formulas such as `te(x, y, time, d = c(2, 1) )`.
+The performance of such models should be assessed in future work.
+
+<!--
 
@@
 Whereas Bayesian linear regression outputs a distribution over the parameters of some predefined parametric model, the GP approach, in contrast, is a non-parametric approach, in that it finds a distribution over the possible functions that are consistent with the observed data.
-However, note that nonparametric does not mean there aren't parameters, it means that there are infinitely many parameters.
+However, "nonparametric" does not mean there aren't parameters, it means that there are infinitely many parameters.
 
@@
 
+-->
+
 \newpage
 
-# How to choose the GAM basis dimension? {#sec-basis}
-
-Here we should provide recommendation about how to define `k`...
-An option is to vary `k` and examine the predictions and posterior predictive checks (PPCs) of each model...
-In this example (@fig-choose-k)...
-However, it is not possible to provide general recommendations, as the optimal `k` depends on the sampling rate, the preprocessing steps (e.g., signal-to-noise ratio, low-pass filtering, etc), and the neural dynamics of the phenomenon under study.
-Choose `k` as large a possible (given limits on computation time etc)?
-Or use LOOIC/WAIC (seem to select too simple models)?
-
-```{r fig-choose-k, results = "hide", echo = FALSE, fig.width = 9, fig.asp = 1, fig.cap = "Posterior predictions and posterior predictive checks for the GAM with varying k (in rows)."}
+# How to choose the GAM basis dimension? {#apx-basis}
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+There is no universal recommendation for choosing the optimal value of `k`, as it depends on several factors, including the sampling rate, preprocessing steps (e.g., signal-to-noise ratio, low-pass filtering), and the underlying neural dynamics of the phenomenon under investigation.
+One strategy is to set `k` as high as computational constraints allow, as suggested by previous authors [e.g., @pedersen_hierarchical_2019].
+Alternatively, one can fit a series of models with different `k` values and compare them using information criteria such as LOOIC or WAIC, alongside posterior predictive checks (PPCs), to select the model that best captures the structure of the data.
+We illustrate this approach below.
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+@fig-choose-k presents the posterior predictions and two forms of posterior predictive checks (PPCs) for each GAM fit using different numbers of basis functions ($k \in \{5, 10, 20, 40, 80\}$).
+With the exception of the $k=5$ model, all other fits yield satisfactory PPCs, indicating that the predicted data closely resemble the empirical observations.
+However, model comparison using the leave-one-out information crtierio (LOOIC), as summarised in @tbl-simulation-results, identifies the $k=40$ model as the best-performing one in terms of LOOIC, closely followed by the $k=80$ model.
+This suggests that the optimal number of basis functions likely lies between these two values.
+Future simulation studies could further investigate how such model selection criteria relate to the precision of onset and offset estimates.
+
+```{r fig-choose-k1, results = "hide", echo = FALSE, fig.width = 10, fig.asp = 1.25, fig.cap = "Posterior predictions and posterior predictive checks for the GAM with varying k (in rows)."}
 # defining a list of possible k values
-# k_values <- c(5, 10, 20, 50)
+# k_values <- c(5, 10, 20, 40, 80)
 
@@
 # ppc_stat2d_plots <- list()
-# 
-# # looping over k values
+
+# looping over k values
 # for (k_value in k_values) {
@@
 #     gam <- brm(
-#         eeg ~ condition + s(time, bs = "tp", k = k_value, by = condition),
+#         eeg_mean ~ condition + s(time, bs = "tp", k = k_value, by = condition),
 #         data = ppt_df,
 #         family = gaussian(),
-#         warmup = 2000,
+#         warmup = 1000,
 #         iter = 5000,
 #         chains = 4,
-#         cores = 4
+#         cores = 4,
+#         file = paste0("results/gam_k", as.character(k_value), ".rds")
+#         )
+#     
+#     # computing model comparison indices
+#     gam <- add_criterion(
+#         x = gam,
+#         # criterion = c("loo", "waic", "bayes_R2", "loo_R2", "marglik"),
+#         criterion = c("loo", "waic", "bayes_R2"),
+#         model_name = paste0("gam_k", as.character(k_value) )
 #         )
@@
 
+# saving the models
+# saveRDS(object = gam_models, file = "results/gam_k_models.rds")
+
 # saving the results
@@
 
+# importing the models
+gam_models <- readRDS(file = "results/gam_k_models.rds")
+
 # importing the results
@@
 
-# importing the results
-
 # pp_check(gam, ndraws = 100, type = "ecdf_overlay") + 
@@
 
-# arranging all plots in a 4x4 grid using patchwork
-# final_plot <- (
-#     post_pred_plots[["5"]] + ppc_dens_plots[["5"]] + ppc_ecdf_plots[["5"]] + ppc_stat2d_plots[["5"]] +
-#     post_pred_plots[["10"]] + ppc_dens_plots[["10"]] + ppc_ecdf_plots[["10"]] + ppc_stat2d_plots[["10"]] +
-#     post_pred_plots[["20"]] + ppc_dens_plots[["20"]] + ppc_ecdf_plots[["20"]] + ppc_stat2d_plots[["20"]] +
-#     post_pred_plots[["50"]] + ppc_dens_plots[["50"]] + ppc_ecdf_plots[["50"]] + ppc_stat2d_plots[["50"]]
-#     ) + plot_layout(ncol = 4, nrow = 4)
+# arranging plots using patchwork
 final_plot <- (
@@
     post_pred_plots[["20"]] + ppc_dens_plots[["20"]] + ppc_stat2d_plots[["20"]] +
-    post_pred_plots[["50"]] + ppc_dens_plots[["50"]] + ppc_stat2d_plots[["50"]]
-    ) + plot_layout(ncol = 3, nrow = 4)
+    post_pred_plots[["40"]] + ppc_dens_plots[["40"]] + ppc_stat2d_plots[["40"]] +
+    post_pred_plots[["80"]] + ppc_dens_plots[["80"]] + ppc_stat2d_plots[["80"]]
+    ) + plot_layout(ncol = 3, nrow = 5)
 
 # displaying the final plot
-print(final_plot)
+# print(final_plot)
 
@@
 
+```{r tbl-k-results, eval = TRUE, echo = FALSE, results = "asis", fig.pos = "h", out.extra = ''}
+#| tbl-cap: "Models comparison with LOOIC. Models are arranged by the diffence in expected log-pointwise density (ELPD) to the best model (i.e., k=40)."
+#| # comparing the models
+models_comp <- brms::loo_compare(
+    gam_models[["5"]],
+    gam_models[["10"]],
+    gam_models[["20"]],
+    gam_models[["40"]],
+    gam_models[["80"]],
+    criterion = "loo"
+    )
+
+data.frame(models_comp) %>%
+    rownames_to_column(var = "k") %>%
+    mutate(k = str_extract(string = k, pattern = '"[^"]*"') ) %>%
+    mutate(k = str_remove_all(string = k, pattern = '^"|"$') ) %>%
+    # mutate(across(.cols = where(is.numeric), .fns = ~round(.x, 2) ) ) %>%
+    gt() %>%
+    fmt_number(
+        columns = everything(),
+        decimals = 2
+        ) %>%
+    cols_align(
+        align = "center",
+        columns = everything()
+        ) %>%
+    cols_align(
+        align = "left",
+        columns = k
+        ) %>%
+    # cols_label(
+    #     # elpd_diff = "\\$Delta$ diff"
+    #     ) %>%
+    tab_options(
+        table.align = "center",
+        table.font.size = 10
+        ) %>%
+    as_latex()
+```
+
+```{r fig-choose-k, results = "hide", echo = FALSE, fig.width = 10, fig.asp = 1.25, fig.cap = "Posterior predictions and posterior predictive checks for the GAM with varying k (in rows)."}
+# displaying the final plot
+print(final_plot)
+```
+
 \newpage
 
-# `R` package and integration with `MNE-Python`
-
-For users who are already familiar with `brms`, the recommended pipeline is to import ERPs or decoding results in `R` and analyse these data using the code provided in the main paper.
+# `R` package and integration with `MNE-Python` {#apx-package}
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+For readers who are already familiar with `brms`, the recommended pipeline is to use the code provided in the main paper (available at <https://github.com/lnalborczyk/brms_meeg>).
 It is also possible to call functions from the `neurogam R` package (available at <https://github.com/lnalborczyk/neurogam>) which come with sensible defaults.
@@
 # using the testing_through_time() function from the neurogam package
-# this may take a few minutes (or hours depending the machine's
-# performance and data size)...
+# this may take a few minutes (or hours depending on the machine's
+# performance and the size of the dataset)...
 gam_onset_offset <- testing_through_time(
@@
     data = raw_df,
-    # threshold for defining clusters (20 by default)
-    threshold = 20,
     # the *_id arguments are used to specify the relevant columns in data 
@@
     time_id = "time", predictor_id = "condition",
+    # posterior odds threshold for defining clusters (20 by default)
+    threshold = 20,
     # number of warmup MCMC iterations
@@
 
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
 The `neurogam` package can also be called from `Python` using the `rpy2` module, and can easily be integrated into `MNE-Python` pipelines.
-For example, we use it below to estimate the onset and offset of effects for one EEG channel from a MNE evoked object.
-Note that the code used to reshape the `sample MNE` dataset is available in the online supplementary materials, and we refer to the [MNE documentation](https://mne.tools/stable/auto_tutorials/epochs/50_epochs_to_data_frame.html) about converting `MNE` epochs to `Pandas` dataframes in long format (i.e., with one observation per row).
+For example, we use it below to estimate the onset and offset of effects for one EEG channel from an MNE evoked object.
+The code used to reshape the `sample MNE` dataset is available in the online supplementary materials, and we further refer to the [MNE documentation](https://mne.tools/stable/auto_tutorials/epochs/50_epochs_to_data_frame.html) about converting `MNE` epochs to `Pandas` dataframes in long format (i.e., with one observation per row).
 
@@
     data=reshaped_df_r,
-    threshold=10,
+    threshold=20,
     multilevel=False

### 2025-08-22 manuscript.tex
pushing revised manuscript

@@
   notimes,
-  donotrepeattitle,
   colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}
@@
 \usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
+\usepackage{subcaption}
+
+\renewcommand\thesubfigure{\Alph{subfigure}}
 \setcounter{topnumber}{2}
@@
 \abstract{Time-resolved electrophysiological measurements such as those
-obtained through magneto- or electro-encephalography (M/EEG) offer a
+obtained through magneto- and electroencephalography (M/EEG) offer a
 unique window onto the neural activity underlying cognitive processes.
@@
 conventional approach involves mass-univariate statistical testing
-across time and space, followed by corrections for multiple comparisons
-such as cluster-based inference. While effective for controlling error
-rates at the cluster-level, cluster-based inference comes with a
+across time and space followed by corrections for multiple comparisons
+or some form of cluster-based inference. While effective for controlling
+error rates at the cluster-level, cluster-based inference comes with a
 significant limitation: by shifting the focus of inference from
 individual time points to clusters, it prevents drawing conclusions
-about the precise onset or offset of observed effects. Here, we
-introduce a \emph{model-based} approach for analysing M/EEG timeseries
-such as event-related potentials (ERPs) or decoding performance over
-time. Our approach leverages Bayesian generalised additive multilevel
-models, providing posterior probabilities that an effect is above zero
-(or above chance) at each time point, while naturally accounting for
-temporal dependencies and between-subject variability. Using both
-simulated and actual M/EEG datasets, we demonstrate that this approach
-substantially outperforms conventional methods in estimating the onset
-and offset of neural effects, yielding more precise and reliable
-results. We provide an R package implementing the method and describe
-how it can be integrated into M/EEG analysis pipelines using
-MNE-Python.}
-
-\keywords{EEG, MEG, cluster-based inference, simulation, multiple
+about the precise onset or offset of observed effects. Here, we present
+a \emph{model-based} alternative for analysing M/EEG timeseries, such as
+event-related potentials or time-resolved decoding accuracy. Our
+approach leverages Bayesian generalised additive multilevel models,
+providing posterior odds that an effect exceeds zero (or chance) at each
+time point, while naturally accounting for temporal dependencies and
+between-subject variability. Using both simulated and empirical M/EEG
+datasets, we show that this approach substantially outperforms
+conventional methods in estimating the onset and offset of neural
+effects, yielding more precise and reliable estimates. We provide an
+open-source R package implementing the method and describe how it can be
+integrated into M/EEG analysis pipelines using MNE-Python. }
+
+\keywords{EEG, MEG, cluster-based inference, multiple
 comparisons, generalised additive models, mixed-effects
@@
 Pasteur, 13100
-Aix-en-Provence, France, email: ladislas.nalborczyk@cnrs.fr}
+Aix-en-Provence, France, email: \href{mailto:ladislas.nalborczyk@cnrs.fr}{ladislas.nalborczyk@cnrs.fr}}
 }
@@
 
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{3}
 
@@
 \resetlinenumber[1]
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
 
@@
 activation.'' As a consequence, cluster-based inference is expected to
-perform poorly for localising the onset of M/EEG effects; a property
+perform poorly for identifying the onset of M/EEG effects; a property
 that was later demonstrated in simulation studies (e.g.,
@@
 Bayesian generalised additive multilevel models, allows quantifying the
-posterior probability of effects being above chance at the level of
-timesteps, sensors, voxels, etc, while naturally taking into account
-spatiotemporal dependencies present in M/EEG data. We compare the
-performance of the proposed approach to well-established alternative
-methods using both simulated and actual M/EEG data and show that it
-significantly outperforms alternative methods in estimating the onset
-and offset of M/EEG effects.
+posterior odds of effects being above chance at the level of timesteps,
+sensors, voxels, etc, while naturally taking into account spatiotemporal
+dependencies present in M/EEG data. We compare the performance of the
+proposed approach to well-established alternative methods using both
+simulated and actual M/EEG data and show that it significantly
+outperforms alternative methods in estimating the onset and offset of
+M/EEG effects.
 
@@
 (family) of \(m\) tests (i.e., the FWER) is computed as
-\(1-\left(1-\alpha\right)^{m}\) (for \(m=10\) tests and \(\alpha=0.05\),
-it is approximately equal to \(0.4\)). Different methods exist to
-control the FWER (i.e., bring it back to \(\alpha\)). Most methods apply
-a simple correction to series of \(p\)-values issued from univariate
-statistical tests (e.g., t-tests). For instance, the Bonferroni
-correction (\citeproc{ref-dunn1961}{Dunn, 1961}) consists in setting the
-significance threshold to \(\alpha/m\), or equivalently, multiplying the
-\(p\)-values by \(m\) and using the standard \(\alpha\) significance
-threshold. This method is generally overconservative (i.e.,
-under-powered) as it assumes statistical independence of the tests, an
-assumption that is clearly violated in the context of M/EEG timeseries
-characterised by massive spatiotemporal dependencies. Some alternative
-methods aims at controlling the FDR, defined as the proportion of false
-positive \emph{among positive tests} (e.g.,
-\citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
+\(1-\left(1-\alpha\right)^{m}\) (for \(m=10\) independent tests and
+\(\alpha=0.05\), it is approximately equal to \(0.4\)). Different
+methods exist to control the FWER, that is, to bring it back to
+\(\alpha\). Most methods apply a simple correction to series of
+\(p\)-values issued from univariate statistical tests (e.g., t-tests).
+For instance, the Bonferroni correction (\citeproc{ref-dunn1961}{Dunn,
+1961}) consists in setting the significance threshold to \(\alpha/m\),
+or equivalently, multiplying the \(p\)-values by \(m\) and using the
+standard \(\alpha\) significance threshold. This method is generally
+overconservative (i.e., under-powered) as it assumes statistical
+independence of the tests, an assumption that is clearly violated in the
+context of M/EEG timeseries characterised by massive spatiotemporal
+dependencies. Some alternative methods aims at controlling the FDR,
+defined as the proportion of false positive \emph{among positive tests}
+(e.g., \citeproc{ref-benjamini1995}{Benjamini \& Hochberg, 1995};
 \citeproc{ref-benjamini2001}{Benjamini \& Yekutieli, 2001}). However, a
@@
 2022}; \citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}).
-In other words, because the inference is performed at the second step
-(i.e., once clusters have been formed), no conclusion can be made about
-individual datapoints (e.g., timesteps, sensors, etc).
+Because the inference is performed at the second step (i.e., once
+clusters have been formed), no conclusion can be made about individual
+datapoints (e.g., timesteps, sensors, etc).
 
@@
 set a priori (typically to 0.5 and 2, respectively) and control the
-influence of the extend and height on the TFCE. Note that in practice,
-this integral is approximated by a sum over small \(h\) increments.
-Then, a \emph{p}-value for each timestep \(t\) is computed by comparing
-its TFCE with the null distribution of TFCE values (obtained via
-permutation). For each permuted signal, we keep the maximal value over
-the whole signal for the null distribution of the TFCE. The TFCE
-combined with permutation (assuming a large enough number of
-permutations) has been shown to provide accurate FWER control (e.g.,
+influence of the extend and height on the TFCE. In practice, this
+integral is approximated by a sum over small \(h\) increments. Then, a
+\emph{p}-value for each timestep \(t\) is computed by comparing its TFCE
+with the null distribution of TFCE values (obtained via permutation).
+For each permuted signal, we keep the maximal value over the whole
+signal for the null distribution of the TFCE. The TFCE combined with
+permutation (assuming a large enough number of permutations) has been
+shown to provide accurate FWER control (e.g.,
 \citeproc{ref-pernet2015}{Pernet et al., 2015}). However, further
@@
 experimental paradigms, potentially leading to suboptimal ERP
-estimations.
+estimations (e.g., \citeproc{ref-smith2014a}{N. J. Smith \& Kutas,
+2014a}).
 
@@
 substantially greater flexibility. Notably, they allow for the modelling
-of both linear and nonlinear effects of continuous predictors--such as
+of both linear and nonlinear effects of continuous predictors, such as
 word frequency or age (e.g., \citeproc{ref-smith2014a}{N. J. Smith \&
 Kutas, 2014a}, \citeproc{ref-smith2014b}{2014b};
-\citeproc{ref-tremblay2014}{Tremblay \& Newman, 2014})--and enable the
+\citeproc{ref-tremblay2014}{Tremblay \& Newman, 2014}), and enable the
 disentangling of overlapping cognitive processes (e.g.,
@@
 by group-level statistical analyses of the resulting regression
-coefficients, often accompanied by multiple comparison corrections or
-cluster-based inference (for recent applied examples, see
+coefficients, often accompanied by corrections for multiple comparisons
+or cluster-based inference (for recent applied examples, see
 \citeproc{ref-dunagan2025}{Dunagan et al., 2025};
@@
 Second, the subsequent group-level analyses typically do not account for
-hierarchical dependencies, between-participant or between-stimulus
-variability, which could otherwise be addressed through multilevel
-modelling. Finally, because the output of this procedure is summarised
-by cluster-based inference, its conclusions remain subject to the
-limitations discussed in the previous section.
+hierarchical dependencies which could otherwise be addressed through
+multilevel modelling. Finally, because the output of this procedure is
+summarised by cluster-based inference, its conclusions remain subject to
+the limitations discussed in the previous section.
 
 Beyond modelling nonlinear effects of continuous predictors on ERP
-amplitudes, generalised additive models (GAMs) have been employed to
-capture the temporal dynamics of ERPs themselves, effectively modelling
-the shape of the waveform over time
-(\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
+amplitudes, GAMs have been employed to capture the temporal dynamics of
+ERPs themselves, effectively modelling the shape of the waveform over
+time (\citeproc{ref-abugaber2023}{Abugaber et al., 2023};
 \citeproc{ref-baayen2018}{Baayen et al., 2018};
@@
 
-In generalised additive models (GAMs), the functional relationship
-between the predictors and the response variable is decomposed into a
-sum of low-dimensional non-parametric functions. A typical GAM has the
+In generalised additive models, the functional relationship between the
+predictors and the response variable is decomposed into a sum of
+low-dimensional non-parametric functions. A typical GAM has the
 following form:
@@
 functions} (e.g., cubic polynomial, thin-plate) and the set of basis
-functions is called a \emph{basis}. Each basis function gets its
-coefficient and the resultant spline is the sum of these weighted basis
-functions (Figure~\ref{fig-intro-gam}\ignorespaces A). Splines
-coefficients are penalised (usually through the squared of the smooth
+functions is called a \emph{basis}. Each basis function is weighted by
+its coefficient and the resultant spline is the sum of these weighted
+basis functions (Figure~\ref{fig-intro-gam}\ignorespaces A). Splines
+coefficients are penalised (usually through the square of the smooth
 functions' second derivative) in a way that can be interpreted, in
@@
 GAMs (\citeproc{ref-rigby2005}{Rigby \& Stasinopoulos, 2005};
-\citeproc{ref-umlauf2018}{Umlauf et al., 2018a}), allow researchers to
+\citeproc{ref-umlauf2018}{Umlauf et al., 2018}), allow researchers to
 model not only the mean structure but also the variance (or scale) and
-other distributional properties as functions of predictors--a feature
+other distributional properties as functions of predictors, a feature
 that has proven useful in modelling neuroimaging data (e.g.,
@@
 demonstrate that this method provides more precise and reliable
-estimates of effect onsets and offsets, offering a valuable tool for
-researchers seeking to enhance the temporal resolution of their
-analyses.
-
-\section{Methods}\label{methods}
-
-\subsection{M/EEG data simulation}\label{meeg-data-simulation}
-
+estimates of effects' onset and offset than conventional approaches such
+as cluster-based inference.
+
+\section{Benchmarking with known ground
+truth}\label{benchmarking-with-known-ground-truth}
+
+\subsection{Methods}\label{methods}
+
+\subsubsection{M/EEG data simulation}\label{meeg-data-simulation}
+
+To assess the accuracy of group-level onset and offset estimation of our
+proposed method, we simulated EEG with known onset and offset values.
 Following the approach of Sassenhagen \& Draschkow
@@
 (Figure~\ref{fig-eeg}). All figures and simulation results can be
-reproduced using the \texttt{R} code available on GitHub:
+reproduced using the \texttt{R} code available online at:
 \url{https://github.com/lnalborczyk/brms_meeg}.
@@
 
-\subsection{Model description and model
+\subsubsection{Model description and model
 fitting}\label{model-description-and-model-fitting}
@@
 We then fitted a Bayesian GAM (BGAM) using the \texttt{brms} package
-(\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018};
-\citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}) and default
-priors (i.e., weakly informative priors). We ran eight Markov Chain
-Monte-Carlo (MCMC) to approximate the posterior distribution, including
-each 5000 iterations and a warmup of 2000 iterations, yielding a total
-of \(8 \times (5000-2000) = 24000\) posterior samples to use for
-inference. Posterior convergence was assessed examining trace plots as
-well as the Gelman--Rubin statistic \(\hat{R}\)
-(\citeproc{ref-gabry2019}{Gabry et al., 2019};
+(\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018})
+and default priors (i.e., weakly informative priors). We ran eight
+Markov Chain Monte-Carlo (MCMC) to approximate the posterior
+distribution, including each 5000 iterations and a warmup of 2000
+iterations, yielding a total of \(8 \times (5000-2000) = 24000\)
+posterior samples to use for inference. Posterior convergence was
+assessed examining trace plots as well as the Gelman--Rubin statistic
+\(\hat{R}\) (\citeproc{ref-gabry2019}{Gabry et al., 2019};
 \citeproc{ref-gelman2020}{Gelman et al., 2020}). The \texttt{brms}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# computing the average EEG signals for each participant}
-\NormalTok{ppt\_df }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{group\_by}\NormalTok{(participant, condition, time) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{summarise}\NormalTok{(}\AttributeTok{eeg\_mean =} \FunctionTok{mean}\NormalTok{(eeg) ) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{ungroup}\NormalTok{()}
-
-\CommentTok{\# defining a contrast for condition}
-\FunctionTok{contrasts}\NormalTok{(ppt\_df}\SpecialCharTok{$}\NormalTok{condition) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{)}
-
-\CommentTok{\# fitting the BGAM}
-\NormalTok{gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-    \CommentTok{\# thin{-}plate regression splines with k{-}1 basis functions}
-\NormalTok{    eeg\_mean }\SpecialCharTok{\textasciitilde{}} \DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition }\SpecialCharTok{+} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
-    \AttributeTok{data =}\NormalTok{ ppt\_df,}
-    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
-    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/gam.rds"}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
 However, the previous model only included constant (fixed) effects, thus
@@
 \citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}) including a
-varying intercept and slope for participant (but only with a constant
+varying intercept and slope for participant (but with a constant
 smoother). Although it is possible to fit a BGAMM using data at the
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# computing the mean and SD of EEG signals for each participant}
-\NormalTok{summary\_df }\OtherTok{\textless{}{-}}\NormalTok{ raw\_df }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{group\_by}\NormalTok{(participant, condition, time) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{summarise}\NormalTok{(}\AttributeTok{eeg\_mean =} \FunctionTok{mean}\NormalTok{(eeg), }\AttributeTok{eeg\_sd =} \FunctionTok{sd}\NormalTok{(eeg) ) }\SpecialCharTok{\%\textgreater{}\%}
-    \FunctionTok{ungroup}\NormalTok{()}
-
-\CommentTok{\# defining a contrast for condition}
-\FunctionTok{contrasts}\NormalTok{(summary\_df}\SpecialCharTok{$}\NormalTok{condition) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{)}
-
-\CommentTok{\# fitting the BGAMM}
-\NormalTok{meta\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-    \CommentTok{\# using the by{-}participant mean and SD of ERPs}
-\NormalTok{    eeg\_mean }\SpecialCharTok{|} \FunctionTok{se}\NormalTok{(eeg\_sd) }\SpecialCharTok{\textasciitilde{}}
-        \DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ condition }\SpecialCharTok{|}\NormalTok{ participant) }\SpecialCharTok{+}
-        \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{by =}\NormalTok{ condition),}
-    \AttributeTok{data =}\NormalTok{ summary\_df,}
-    \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
-    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
-    \AttributeTok{file =} \StringTok{"models/meta\_gam.rds"}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
 We depict the posterior predictions together with the posterior estimate
@@
 provides an adequate description of the simulated data (see further
-posterior predictive checks in Section~\ref{sec-basis}).
+posterior predictive checks in \hyperref[apx-basis]{Appendix~B}).
 
@@
 left). From this quantity, we compute the ratio of posterior
-probabilities (i.e., \(p/(1-p)\)) and visualise the timecourse of this
-ratio superimposed with the conventional thresholds on evidence ratios
-(Figure~\ref{fig-post-prob-ratio}, right). Note that a ratio of 10 means
-that the probability of the difference being above 0 is 10 times higher
-than the probability of the difference not being above 0, given the
-data, the priors, and other model's assumptions.\footnote{This posterior
-  probability ratio, or posterior odds, is equivalent to a Bayes factor,
-  assuming 1:1 prior odds.} Thresholding the posterior probability ratio
-thus provides a model-based approach for estimating the onset and offset
-of M/EEG effects, whose properties will be assessed in the simulation
-study. An important advantage is that the proposed approach can be
-extended to virtually any model structure.
+probabilities (i.e., \(p/(1-p)\)), or posterior odds, and visualise the
+timecourse of these odds superimposed with the conventional thresholds
+on evidence ratios (Figure~\ref{fig-post-prob-ratio}, right). A ratio of
+10 means that the probability of the difference being above 0 is 10
+times higher than the probability of the difference not being above 0,
+given the data, the priors, and other model's assumptions.\footnote{These
+  posterior odds are equivalent to a Bayes factor, assuming 1:1 prior
+  odds.} Thresholding the posterior odds thus provides a model-based
+approach for estimating the onset and offset of M/EEG effects, whose
+properties will be assessed in the simulation study. An important
+advantage is that the proposed approach can be extended to virtually any
+model structure.
 
@@
 EEG difference (slope) being above 0 according to the BGAMM. Right:
-Ratio of posterior probability according to the BGAMM (on a log10
-scale). Timesteps above threshold (10) are highlighted in green. NB: the
-minimum and maximum possible ratio values are determined (bounded) by
-the number of posterior samples in the model.}
+Posterior odds according to the BGAMM (on a log10 scale). Timesteps
+above threshold (10) are highlighted in green. NB: the minimum and
+maximum possible posterior odds are determined (bounded) by the number
+of posterior samples in the model.}
 
@@
 
-\newpage
-
-\subsection{Error properties of the proposed
-approach}\label{error-properties-of-the-proposed-approach}
-
-We then assessed the performance of the proposed approach by computing
-the difference between the true and estimated onset/offset of the EEG
-difference according to various \texttt{k} (basis dimension) and
-\texttt{threshold} values. Remember that the EEG signal was generated
-from a truncated Gaussian with an objective onset at 160 ms, a maximum
-at 250 ms, and an offset at 342 ms. Figure~\ref{fig-onset-error} shows
-that the multilevel GAM can almost exactly recover the true onset and
-offset values, given some reasonable choice of \texttt{k} and
-\texttt{threshold} values. We provide more detailed recommendations on
-how to set \texttt{k} in Section~\ref{sec-basis}. This figure further
-reveals that the optimal \texttt{k} and \texttt{threshold} values may
-differ for the onset and offset values, and there there seems to exist a
-trade-off between these two parameters: lower \texttt{k} values lead to
-poorer estimations, but these poor estimations can be compensated (only
-to some extent) by higher \texttt{threshold} values (and reciprocally).
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-onset-error}Average estimation error (RMSE) for the
-onset (left) and offset (right) according to various basis dimension and
-threshold values for the BGAM (computed from 100 simulated datasets).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-onset-error-1.png}
-
-}
-
-\end{figure}%
-
-\newpage
-
-\subsection{Comparing the onsets/offsets estimates across
-approaches}\label{comparing-the-onsetsoffsets-estimates-across-approaches}
+\subsubsection{Comparing the onset/offset estimates across
+approaches}\label{comparing-the-onsetoffset-estimates-across-approaches}
 
@@
 to the squared \(t\)-values (as in
-\citeproc{ref-rousselet_using_2025}{Rousselet, 2025}).
+\citeproc{ref-rousselet_using_2025}{Rousselet, 2025}).\footnote{As in
+  Rousselet (\citeproc{ref-rousselet_using_2025}{2025}), we fixed the
+  number of expected change points to two in the binary segmentation
+  algorithm, thus producing always one cluster.}
 Figure~\ref{fig-corrections} illustrates the onsets and offsets
@@
 all methods systematically overestimate the true onset and underestimate
-the true offset.
+the true offset. In addition, the \texttt{Raw\ p-value},
+\texttt{FDR\ BH95}, and \texttt{FDR\ BY01} methods identify clusters
+well before the true onset and after the true offset.
 
@@
 
-\newpage
-
-\subsection{Simulation study}\label{simulation-study}
-
-To assess the accuracy of group-level onset estimation, the various
-methods were compared using the bias (defined as the median difference
-between the estimated and true value of the onset/offset), median
-absolute error (MAE), root mean square error (RMSE), variance, and
-median absolute deviation (MAD) of onset/offset estimates computed on
-10,000 simulated datasets. As in Rousselet
-(\citeproc{ref-rousselet_using_2025}{2025}), each participant was
-assigned a random onset between 150 and 170ms. Whereas the focus of the
-present article focuses is on one-dimensional signals (e.g., one M/EEG
-channel), we provide a 2D application in Section~\ref{sec-2D}.
-
-\subsection{Application to actual MEG
-data}\label{application-to-actual-meg-data}
-
-To complement the simulation study, we evaluated the performance of all
-methods on actual MEG data (decoding results from
-\citeproc{ref-nalborczyk:inprep}{Nalborczyk et al., in preparation}). In
-this study, we conducted time-resolved multivariate pattern analysis
-(MVPA, also known as decoding) of MEG data during reading tasks. As a
-result, we obtain a timecourse of decoding performance (ROC AUC),
-bounded between 0 and 1, for each participant (\(N=32\)). Next, we
-wanted to \emph{test} whether the group-level average decoding accuracy
-is above chance (i.e., 0.5) at each timestep
-(Figure~\ref{fig-decoding-data}). To achieve this, we fitted a BGAM as
-introduced previously, but we replaced the \(\mathrm{Normal}\)
-likelihood function by a \(\mathrm{Beta}\) one to account for the
-bounded nature of AUC values (between 0 and 1) (for a tutorial on Beta
-regression, see \citeproc{ref-coretta2025}{Coretta \& Bürkner, 2025}).
-
-Note that although we chose a basis dimension of \(k=50\), which seems
-appropriate for the present data, this choice should be adapted
-according to the properties of the modelled data (e.g., signal-to-noise
-ratio, prior low-pass filtering, sampling rate) and should be assessed
-by the usual model checking tools (e.g., posterior predictive checks,
-see also Section~\ref{sec-basis}). To better distinguish signal from
-noise, we also defined a region of practical equivalence (ROPE,
-\citeproc{ref-kruschke2017}{Kruschke \& Liddell, 2017}), defined as the
-chance level plus the standard deviation of the (group-level average)
-decoding performance during the baseline period.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# fitting the Beta GAM}
-\NormalTok{meg\_decoding\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{s}\NormalTok{(time, }\AttributeTok{bs =} \StringTok{"cr"}\NormalTok{, }\AttributeTok{k =} \DecValTok{50}\NormalTok{),}
-    \AttributeTok{data =}\NormalTok{ decoding\_df,}
-    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
-    \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{4}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{4}
-\NormalTok{    )}
-\end{Highlighting}
-\end{Shaded}
-
-We assessed the reliability of the proposed approach using a form of
-permutation-based split-half reliability (as for instance in
-\citeproc{ref-rosenblatt2018}{Rosenblatt et al., 2018}), which consisted
-of the following steps. First, we created 1,000 split halves of the data
-(i.e., with half the participants in the original data, that is, 16
-participants). For each split, we estimated the onset/offset using all
-methods described previously. Third, we summarised the distribution of
-onset/offset estimates using the median ``error'' (i.e., difference
-between the split estimate and the estimate obtained using the full
-dataset) and the variance across splits. This approach allows assessing
-how similar the estimate of each half split is to the full dataset (thus
-acting as a proxy for the population) and how variable the estimates are
-across split halves.
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-decoding-data}Group-level average decoding
-performance (N=32) superimposed with the GAM predictions (in blue) and
-the region of practical equivalence (ROPE, in orange) computed from the
-baseline period (data from \citeproc{ref-nalborczyk:inprep}{Nalborczyk
-et al., in preparation}). The blue horizontal markers indicate the
-timesteps at which the posterior probability ratio exceeds 20.}
-
-\centering{
-
-\includegraphics[width=0.75\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-decoding-data-1.png}
-
-}
-
-\end{figure}%
-
-\newpage
-
-\section{Results}\label{results}
-
-This section is divided in two parts. First, we present the results from
-the simulation study, assessing the bias and variance of each method
-when applied to simulated data in which the ground truth is known.
-Second, we present the results obtained when applying the different
-methods to actual MEG data (decoding performance through time),
-assessing the reliability of the estimates provided by each method.
-
-\subsection{Simulation study (bias and
-variance)}\label{simulation-study-bias-and-variance}
+\subsubsection{Simulation study}\label{simulation-study}
+
+To assess the accuracy of group-level onset and offset estimation, all
+methods were compared by computing the bias (defined as the mean
+difference between the estimated and true value of the onset/offset),
+mean absolute error (MAE), root mean square error (RMSE), and variance
+of onset/offset estimates from 10,000 simulated datasets. Following
+Rousselet (\citeproc{ref-rousselet_using_2025}{2025}), each participant
+was assigned a random onset between 150 and 170ms. Whereas the present
+article focuses on one-dimensional signals (e.g., one M/EEG channel), we
+provide a 2D application in \hyperref[apx-2D]{Appendix~A}.
+
+\subsection{Results}\label{results}
 
@@
 results, revealing that the proposed approach (\texttt{BGAM}) has the
-lowest median error for both the onset and offset estimates. The
-\texttt{Cluster\ mass} and \texttt{Change\ point} also have good
-performance, but surprisingly, the \texttt{TFCE} method has relatively
-bad performance for estimating the effect offset (similar performance to
-the \texttt{Holm} and \texttt{FDR\ BY01} methods). Unsurprisingly, the
-\texttt{FDR\ BH95} and \texttt{Raw\ p-value} methods show the worst
-performance.
+lowest error for both the onset and offset estimates. The
+\texttt{Cluster\ mass} and \texttt{Change\ point} methods also have good
+performance, but perhaps surprisingly, the \texttt{TFCE} method performs
+poorly for estimating the offset of the effect (with performance similar
+to the \texttt{Holm} method). Unsurprisingly, the \texttt{FDR\ BH95} and
+\texttt{Raw\ p-value} methods show the worst performance.
 
@@
 
-\caption{\label{fig-simulation-results}Median error and median absolute
-deviation of the error for the onset (left) and offset (right) estimates
-according to each method. Methods are ordered from lowest (top) to
-highest (bottom) median absolute error (separately for the onset and
-offset estimates).}
+\caption{\label{fig-simulation-results}Mean error and variance of onset
+and offset estimates according to each method. Variance is plotted on a
+log10 scale for visual purposes.}
 
@@
 
-Results are further summarised in Table~\ref{tbl-simulation-results},
-which shows that the \texttt{BGAM} is almost perfectly unbiased (i.e.,
-it has a bias of approximately 2ms for the onset and 4ms for the
-offset). The \texttt{Bias} column shows that all methods tend to
-estimate the onset later than the true onset and to estimate the offset
-earlier than the true offset. As can be seen from this table, the
-\texttt{BGAM} has the best performance on all included metrics.
+These results are further summarised in
+Table~\ref{tbl-simulation-results}, which shows that the \texttt{BGAM}
+method is almost perfectly unbiased (i.e., it has a bias of
+approximately 0.1ms for the onset and 2.4ms for the offset). The
+\texttt{Bias} column shows that all methods tend to estimate the onset
+later than the true onset and to estimate the offset earlier than the
+true offset. As can be seen from this table, the \texttt{BGAM} method
+has the best performance on all included metrics (except for the
+\texttt{Variance} of the offset estimate, where the
+\texttt{Change\ point} method performs better, presumably because it was
+constrained to identifying a single cluster).
 
@@
 
-{\caption{{Summary statistics of the onset and offset estimates for each
-method (in ms, ordered by the MAE) according to the main simulation
-study.}{\label{tbl-simulation-results}}}}
+{\caption{{Summary statistics of onset/offset estimates for each method
+(in ms, ordered by the MAE).}{\label{tbl-simulation-results}}}}
 
 \fontsize{9.0pt}{10.8pt}\selectfont
-\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lccccc}
+\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lcccc}
 \toprule
- & Bias & MAE & RMSE & Variance & MAD \\ 
+ & Bias & MAE & RMSE & Variance \\ 
 \midrule\addlinespace[2.5pt]
-\multicolumn{6}{l}{onset} \\[2.5pt] 
+\multicolumn{5}{l}{onset} \\[2.5pt] 
 \midrule\addlinespace[2.5pt]
-BGAM (new method) & 2.00 & 2.00 & 1.03 & 43.83 & 2.97 \\ 
-Cluster mass & 6.00 & 6.00 & 6.78 & 228.16 & 5.93 \\ 
-Change point & 6.00 & 6.00 & 5.21 & 34.54 & 5.93 \\ 
-FDR BY01 & 12.00 & 14.00 & 35.44 & 7,659.08 & 5.93 \\ 
-TFCE & 16.00 & 16.00 & 29.88 & 2,785.39 & 5.93 \\ 
-Holm & 16.00 & 16.00 & 29.91 & 2,710.69 & 5.93 \\ 
-FDR BH95 & 8.00 & 48.00 & 57.50 & 18,986.98 & 83.03 \\ 
-Raw p-value & 6.00 & 118.00 & 72.39 & 25,072.78 & 183.84 \\ 
+BGAM (our method) & 0.11 & 2.76 & 0.11 & 28.78 \\ 
+Change point & 5.17 & 6.51 & 5.17 & 33.58 \\ 
+Cluster mass & 6.64 & 7.62 & 6.64 & 207.86 \\ 
+Holm & 29.83 & 32.01 & 29.83 & 2,763.68 \\ 
+TFCE & 29.73 & 32.07 & 29.73 & 2,823.11 \\ 
+FDR BY01 & 36.67 & 50.80 & 36.67 & 7,872.30 \\ 
+FDR BH95 & 58.64 & 102.49 & 58.64 & 19,054.58 \\ 
+Raw p-value & 70.72 & 132.86 & 70.72 & 25,004.65 \\ 
 \midrule\addlinespace[2.5pt]
-\multicolumn{6}{l}{offset} \\[2.5pt] 
+\multicolumn{5}{l}{offset} \\[2.5pt] 
 \midrule\addlinespace[2.5pt]
-BGAM (new method) & -4.00 & 4.00 & 3.00 & 45.58 & 2.97 \\ 
-Cluster mass & -8.00 & 8.00 & 8.60 & 200.72 & 5.93 \\ 
-Change point & -10.00 & 10.00 & 9.22 & 34.58 & 5.93 \\ 
-FDR BY01 & -14.00 & 16.00 & 37.88 & 7,708.00 & 5.93 \\ 
-TFCE & -18.00 & 18.00 & 32.15 & 2,822.35 & 5.93 \\ 
-Holm & -18.00 & 18.00 & 31.96 & 2,720.13 & 5.93 \\ 
-FDR BH95 & -10.00 & 46.00 & 60.10 & 19,025.38 & 83.03 \\ 
-Raw p-value & -8.00 & 112.00 & 71.59 & 24,969.31 & 177.91 \\ 
+BGAM (our method) & -2.35 & 3.46 & 2.35 & 53.63 \\ 
+Cluster mass & -8.64 & 9.44 & 8.64 & 193.63 \\ 
+Change point & -9.28 & 9.82 & 9.28 & 33.03 \\ 
+Holm & -31.86 & 34.03 & 31.86 & 2,764.12 \\ 
+TFCE & -31.84 & 34.19 & 31.84 & 2,834.11 \\ 
+FDR BY01 & -36.68 & 51.37 & 36.68 & 7,648.53 \\ 
+FDR BH95 & -58.87 & 102.63 & 58.87 & 18,939.58 \\ 
+Raw p-value & -72.93 & 133.33 & 72.93 & 25,012.72 \\ 
 \bottomrule
@@
 
+\section{Application to actual MEG
+data}\label{application-to-actual-meg-data}
+
+\subsection{Methods}\label{methods-1}
+
+To complement the simulation study, we evaluated the performance of all
+methods on actual MEG data (\citeproc{ref-nalborczyk:inprep}{Nalborczyk
+et al., in preparation}). In this study, the authors conducted
+time-resolved multivariate pattern analysis (MVPA, also known as
+decoding) of MEG data recorded in 32 human participants during a reading
+task. As a result, the authors obtained a timecourse of decoding accuray
+(ROC AUC), bounded between 0 and 1, for each participant. To test
+whether the group-level average decoding accuracy was above chance
+(i.e., 0.5) at each timestep, we fitted a BGAM as introduced previously
+with a basis dimension \(k=50\) and retained all timesteps exceeding a
+posterior odds of 20. To better distinguish signal from noise, we
+defined a region of practical equivalence (ROPE,
+\citeproc{ref-kruschke2017}{Kruschke \& Liddell, 2017}) as the upper
+90\% quantile of decoding performance during the baseline period (i.e.,
+before stimulus onset). Although we chose a basis dimension of \(k=50\),
+which seemed appropriate for the present data, this choice should be
+adapted according to the properties of the modelled data (e.g.,
+signal-to-noise ratio, prior low-pass filtering, sampling rate) and
+should be assessed by the usual model checking tools (e.g., models
+comparison, posterior predictive checks, see
+\hyperref[apx-basis]{Appendix~B}).
+
+\subsection{Results}\label{results-1}
+
+Figure~\ref{fig-onset-offset} shows the group-level average decoding
+performance through time superimposed with onset and offset estimates
+from each method. Overall, this figure shows that both the
+\texttt{Raw\ p-value} and \texttt{FDR\ BH95} methods are extremely
+lenient, identifying clusters of above-chance decoding accuracy before
+the onset of the stimulus (false positive) and until the end of the
+trial. The \texttt{Change\ point} method seems to be the most
+conservative one, identifying a single cluster spanning from
+approximately +60ms to +450ms. The \texttt{Holm},
+\texttt{Cluster\ mass}, \texttt{TFCE}, and \texttt{BGAM} methods produce
+roughly similar estimates of onset and offset, ranging from
+approximately +60ms to +650ms (considering only the first and last
+identified timesteps), although the \texttt{BGAM} method seems to result
+in fewer clusters.
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-onset-offset}Group-level average decoding
+performance through time with clusters of higher-than-chance decoding
+performance as identified by each method (data from
+\citeproc{ref-nalborczyk:inprep}{Nalborczyk et al., in preparation}).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-onset-offset-1.png}
+
+}
+
+\end{figure}%
+
+We then assessed the sensitivity of the various methods using a form of
+permutation-based sensitivity study, which consisted of the following
+steps. First, we created a large number of split halves of the data,
+that is, subsets of the dataset containing only 16 out of 32
+participants. For each possible pair of subsets, we have 16 possible
+levels of overlap/similarity that can be quantified using the Jaccard
+index, ranging from 0 (perfectly disjoints subsets) to \(\approx0.88\)
+(identical subsets except one participant). For each of these 16 levels
+of Jaccard similarity, we created 1,000 pairs of subsets, resulting in
+16,000 pairs of subsets in total. For each of these pairs, we estimated
+the onset and offset according to each method and computed the absolute
+difference in onset/offset estimates. Finally, we estimated the
+Spearman's rank correlation coefficient (which quantifies the strength
+of a monotonic relation between two variables) between the Jaccard
+similarity and the absolute difference in onset/offset estimates. The
+rational for this procedure is that sensitive methods should produce
+similar onset/offset estimates for similar subsets and dissimilar
+onset/offset estimates for dissimilar subsets. The results of this
+procedure are summarised in Figure~\ref{fig-splits-similarity}.
+
+\begin{figure}[!htb]
+
+\caption{\label{fig-splits-similarity}Relation between data subsets'
+dissimilarity (x-axis) and difference in onset (blue) and offset
+(orange) estimates (y-axis) according to each method.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-splits-similarity-1.png}
+
+}
+
+\end{figure}%
+
+This figure shows that, among the methods that performed best in the
+simulation study (i.e., \texttt{Cluster\ mass}, \texttt{Change\ point},
+and \texttt{BGAM}), onset estimates remain highly stable across subsets
+of participants with varying Jaccard similarity, it varies from around
+1ms for most similar subsets to around 10ms for most dissimilar subsets.
+Additionally, for both onset and offset estimates, the average pairwise
+difference increases monotonically with Jaccard dissimilarity, as
+indicated by the Spearman's rank correlation coefficient. For the onset
+estimates, the \texttt{Holm}, \texttt{Cluster\ mass}, \texttt{TFCE}, and
+\texttt{BGAM} methods exhibit the strongest monotonic relation with
+subset similarity (all \(\rho\text{s} > 0.9\)), whereas for offset
+estimates, all methods demonstrate excellent performance (all
+\(\rho\text{s} > 0.9\)) with the \texttt{BGAM} method showing the
+highest sensitivity (\(\rho\approx0.994\)). However, given the aberrant
+clusters identified by the \texttt{Raw\ p-value}, \texttt{FDR\ BH95},
+and \texttt{FDR\ BY01} methods (see Figure~\ref{fig-onset-offset}),
+their sensitivity to variation in subset similarity is not meaningful.
+
+\section{Discussion}\label{discussion}
+
+In brief, our results show that the model-based approach we introduced
+outperforms conventional methods such as cluster-based inference in
+identifying the onset and offset of M/EEG effects. Performance was
+assessed both on simulated data--allowing us to evaluate the method's
+ability to recover ground-truth onset and offset values--and on actual
+MEG data--allowing us to assess its sensitivity to realistic data
+properties (subset similarity). Together, these results highlight
+desirable properties for any method aiming to precisely and reliably
+estimate the onset and offset of M/EEG effects: it should i) recover
+true onsets and offsets in simulation (good asymptotic behaviour), ii)
+identify clusters that are interpretable and consistent in empirical
+data, and iii) show sensitivity to subtle changes in the data. Our
+approach meets all three of these desiderata.
+
+As with previous simulation studies (e.g.,
+\citeproc{ref-rousselet2008}{Rousselet et al., 2008};
+\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}), results
+inevitably depend on design choices, including the specific
+cluster-forming algorithm and threshold (for cluster-based methods), the
+signal-to-noise ratio, and the potential degradation of temporal
+resolution introduced by preprocessing steps such as low-pass filtering.
+However, these constraints apply equally to all methods tested, so
+relative differences in performance remain meaningful.
+
+\textbf{I am not sure what to do with this paragraph (not sure how to
+discuss these results/observations):} Interestingly, the TFCE method
+performed worse than the traditional cluster-sum approach, consistent
+with the predictions of Rousselet
+(\citeproc{ref-rousselet_using_2025}{2025}) based on the original
+findings of S. Smith \& Nichols (\citeproc{ref-smith2009}{2009}). We
+also found a striking overlap in the clusters identified by the
+\texttt{Holm} and \texttt{TFCE} procedures (cf.
+Figure~\ref{fig-onset-offset} and Figure~\ref{fig-splits-similarity}).
+Whereas these two approaches are conceptually distinct; \texttt{Holm}
+controlling the family-wise error rate through sequential \emph{p}-value
+adjustment, \texttt{TFCE} enhancing signal based on spatiotemporal
+support; their similarity in practice may arise because both effectively
+prioritise extended, moderately strong effects over isolated
+high-intensity points. This convergence warrants further methodological
+work. We chose not to include the cluster-depth algorithm
+(\citeproc{ref-frossard2022}{Frossard \& Renaud, 2022}), as prior work
+already shows its performance to be inferior to cluster-mass approaches
+in similar contexts (\citeproc{ref-rousselet_using_2025}{Rousselet,
+2025}).
+
+A critical consideration for any model-based approach is that the model
+must adequately capture the underlying data-generating process.
+Misspecified models are likely to produce biased or unreliable
+onset/offset estimates. This underscores the importance of thorough
+model diagnostics, including posterior predictive checks, fit
+assessments, and model comparison (\citeproc{ref-gelman2020}{Gelman et
+al., 2020}). As with any inferential modelling framework, transparency
+in assumptions and rigorous validation procedures are indispensable to
+ensure replicability and generalisability of the results.
+
+Another important methodological consideration concerns the selection of
+model hyperparameters, such as the threshold for posterior odds and the
+number of basis functions. Although our simulations suggest that these
+parameters influence the precision and reliability of onset and offset
+estimates, optimal values may vary depending on the signal's temporal
+dynamics and signal-to-noise characteristics. Future work could explore
+principled approaches to hyperparameter tuning, including
+cross-validation or fully Bayesian model selection using tools such as
+leave-one-out cross-validation (LOO-CV) or Bayes factors
+(\citeproc{ref-gelman2020}{Gelman et al., 2020}). We provide initial
+guidance in \hyperref[apx-basis]{Appendix~B} and advocate for future
+development of adaptive heuristics to support flexible yet parsimonious
+model specification.
+
+Currently, our approach estimates temporal effects independently at each
+sensor (1D temporal data). Extending the current framework to
+incorporate additional temporal (see \hyperref[apx-2D]{Appendix~A}) or
+spatial dimensions would improve both sensitivity and interpretability.
+Such extensions could draw on methods from spatial epidemiology and
+geostatistics using either GAMMs or approximate Gaussian process
+regression (e.g., \citeproc{ref-rasmussen2005}{Rasmussen \& Williams,
+2005}; \citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
+2023}), depending on computational feasibility.
+
+To facilitate adoption, we developed the \texttt{neurogram} open-source
+\texttt{R} package (\citeproc{ref-neurogam}{Nalborczyk, 2025}), which
+implements the proposed method using \texttt{brms}. The package
+integrates seamlessly with \texttt{MNE-Python}
+(\citeproc{ref-gramfort2013}{Gramfort, 2013}), enabling researchers to
+process M/EEG data in \texttt{Python} and import them directly into
+\texttt{R} for model-based inference without cumbersome data export.
+This interoperability, described in \hyperref[apx-package]{Appendix~C},
+is designed to encourage broader use of model-based approaches in
+cognitive neuroscience.
+
+In conclusion, we introduced a model-based approach for estimating the
+onset and offset of M/EEG effects. Across simulated and empirical
+datasets, we showed that the method yields more precise and sensitive
+estimates than conventional cluster-based approaches. These results
+highlight the potential of flexible, model-based alternatives for
+characterising time-resolved neural dynamics, particularly in
+applications where accurate temporal localisation is critical.
+
 \newpage
 
-\subsection{Application to actual MEG data
-(reliability)}\label{application-to-actual-meg-data-reliability}
-
-Figure~\ref{fig-onset-offset} shows the group-level average decoding
-performance through time with onset and offset estimates from each
-method. Overall, this figure shows that both the \texttt{Raw\ p-value}
-and \texttt{FDR\ BH95} methods are extremely lenient, considering that
-the decoding performance is above chance before the onset of the
-stimulus (false positive) and until the end of the trial. The
-\texttt{Change\ point} and \texttt{Cluster\ mass} methods seem the most
-conservative methods, identifying a time window from approximately +60ms
-to +500ms. The \texttt{Holm}, \texttt{TFCE}, and \texttt{BGAM} methods
-produce similar estimates of onset and offset, ranging from
-approximately +60ms to +650ms, although the \texttt{BGAM} method seems
-to result in fewer clusters.\footnote{It should be noted that although
-  each method can produce several ``clusters'' of timesteps, we only
-  considered the first (onset) and last (offset) timesteps identified by
-  each method to compute the estimation error.}
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-onset-offset}Group-level average decoding
-performance through time with onset and offset estimates for each method
-(data from \citeproc{ref-nalborczyk:inprep}{Nalborczyk et al., in
-preparation}).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-onset-offset-1.png}
-
-}
-
-\end{figure}%
-
-Figure~\ref{fig-reliability} shows the median difference between the
-onset and offset estimates from each data split and the onset and offset
-estimates from the full dataset (x-axis) along with the variance of its
-onset and offset estimates across data splits (error bar). This figure
-reveals that the \texttt{BGAM} \emph{onset and offset} estimates on each
-split are the closest to the estimates from the full dataset on average
-(0ms difference for the onset estimate and 5ms difference for the offset
-estimate). The \texttt{Raw\ p-value} method has similar performance for
-the offset, but given the aberrant estimates it produces (cf.
-Figure~\ref{fig-onset-offset}), the fact that it is consistent between
-data splits and the full dataset is not convincing on its own. The
-\texttt{Change\ point} method also has a very good performance (i.e.,
-very low difference between split estimates and full estimates), but
-produces too short cluster of significant decoding performance (cf.
-Figure~\ref{fig-onset-offset}).\footnote{As in Rousselet
-  (\citeproc{ref-rousselet_using_2025}{2025}), we fixed the number of
-  expected change points to two in the binary segmentation algorithm,
-  thus producing always one cluster.} Overall, the figure reveals that
-for all other methods, split datasets produce later onset estimates and
-earlier offset estimates (compared to the estimates from the model
-fitted on the full dataset). These results highlight some desirable
-properties for a method aiming to precisely and reliably estimate the
-onset and offset of M/EEG effects, namely, it should i) have good
-asymptotic properties on simulated data, ii) provide sensible identified
-clusters in actual data, and iii) provide reliable/stable estimates
-across various partitions of the data. All these desiderata seem to be
-fulfilled by the BGAM approach.
-
-\begin{figure}[!htb]
-
-\caption{\label{fig-reliability}Median error and median absolute
-deviation of the error for the onset (left) and offset (right) estimates
-according to each method. Methods are ordered from smallest (top) to
-largest (bottom) median absolute error separately for the onset and
-offset estimates.}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-reliability-1.png}
-
-}
-
-\end{figure}%
-
-\begin{table}
-
-{\caption{{Summary statistics of the onset and offset estimates for each
-method (in ms, ordered by the MAE) according to the reliability
-simulation study.}{\label{tbl-reliability-results}}}}
-
-\fontsize{9.0pt}{10.8pt}\selectfont
-\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lccccc}
-\toprule
- & Bias & MAE & RMSE & Variance & MAD \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{6}{l}{onset} \\[2.5pt] 
-\midrule\addlinespace[2.5pt]
-BGAM (new method) & 0.00 & 0.00 & 6.91 & 829.66 & 0.00 \\ 
-Cluster mass & 0.00 & 0.00 & 2.07 & 8.33 & 0.00 \\ 
-TFCE & 5.00 & 5.00 & 10.67 & 101.05 & 7.41 \\ 
-Holm & 5.00 & 5.00 & 10.04 & 96.64 & 7.41 \\ 
-Change point & 5.00 & 5.00 & 1.74 & 96.94 & 7.41 \\ 
-FDR BY01 & 10.00 & 10.00 & 3.71 & 1,581.49 & 0.00 \\ 
-FDR BH95 & 10.00 & 90.00 & 6.30 & 6,970.82 & 118.61 \\ 
-Raw p-value & -40.00 & 60.00 & 27.50 & 5,545.80 & 74.13 \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{6}{l}{offset} \\[2.5pt] 
-\midrule\addlinespace[2.5pt]
-Raw p-value & 0.00 & 35.00 & 39.65 & 7,386.26 & 51.89 \\ 
-BGAM (new method) & -15.00 & 60.00 & 38.07 & 23,511.47 & 66.72 \\ 
-Change point & 20.00 & 60.00 & 54.11 & 8,354.85 & 81.54 \\ 
-FDR BH95 & -75.00 & 75.00 & 111.46 & 15,361.69 & 140.85 \\ 
-FDR BY01 & -145.00 & 150.00 & 122.62 & 17,421.16 & 88.96 \\ 
-Holm & -160.00 & 160.00 & 123.00 & 13,560.13 & 126.02 \\ 
-Cluster mass & -165.00 & 165.00 & 150.79 & 4,272.92 & 29.65 \\ 
-TFCE & -180.00 & 180.00 & 131.70 & 13,720.28 & 96.37 \\ 
-\bottomrule
-\end{tabular*}
-
-\end{table}
-
-\newpage
-
-\section{Discussion}\label{discussion}
-
-\subsection{Summary of the proposed
-approach}\label{summary-of-the-proposed-approach}
-
-Overall, before concluding on the onset/offset of effect based on the
-model, we need to ensure that the model provides a faithful description
-of the data-generating process (e.g., via posterior predictive checks
-etc)\ldots{}
-
-The TFCE performs worse than the cluster-sum approach, which was
-anticipated by Rousselet (\citeproc{ref-rousselet_using_2025}{2025})
-based on the initial results of S. Smith \& Nichols
-(\citeproc{ref-smith2009}{2009})\ldots{} we did not include the
-cluster-depth algorithm (\citeproc{ref-frossard2022}{Frossard \& Renaud,
-2022}), as Rousselet (\citeproc{ref-rousselet_using_2025}{2025}) already
-showed its performance were worse than the cluster mass
-algorithm\ldots{}
-
-\subsection{Limitations and future
-directions}\label{limitations-and-future-directions}
-
-As in previous simulation work (e.g.,
-\citeproc{ref-rousselet2008}{Rousselet et al., 2008};
-\citeproc{ref-sassenhagen2019}{Sassenhagen \& Draschkow, 2019}), the
-present simulation results depend on various choices such as the
-specific cluster-forming algorithm and threshold, signal-to-noise ratio,
-negative impact of preprocessing steps (e.g., low-pass filter) on
-temporal resolution\ldots{} note however, that the same caveats apply to
-all methods\ldots{}
-
-Discussion about the nature of the model: we modelled the surface M/EEG
-signals, however, the true interests (probably) lie in the brain, that
-is, in the source space\ldots{} we could build a ``full'' Bayesian model
-of the generated EEG signal (i.e., including hypotheses about the source
-and a forward model), but this model would become computationally
-heavier\ldots{}
-
-The error properties depend on the threshold parameter, a value of 10 or
-20 seems to be a reasonable default, but the optimal threshold parameter
-can be adjusted using split-half reliability assessment\ldots{} also
-depends on \texttt{k}\ldots{}
-
-Can be applied to any 1D timeseries (e.g., pupillometry,
-electromyography)\ldots{} Extending the approach to spatiotemporal data
-(i.e., time + sensors) or spatiotemporal time-frequency 4D data\ldots{}
-
-We kept the exemplary models simple, but can be extended by adding
-varying/random effects (intercept and slope) for item (e.g.,
-word)\ldots{} but also continuous predictors at the trial level,
-overlapping factors, etc\ldots{}
-
-\newpage
+\setcounter{secnumdepth}{0}
 
@@
 The simulation results as well as the \texttt{R} code to reproduce the
-simulations are available on GitHub:
+simulations are available at:
 \url{https://github.com/lnalborczyk/brms_meeg}. The \texttt{neurogam\ R}
@@
 following R packages: assertthat v. 0.2.1
-(\citeproc{ref-assertthat}{Wickham, 2019}), bamlss v. 1.2.5
-(\citeproc{ref-bamlss2024}{Muschinski et al., 2024};
-\citeproc{ref-bamlss2018}{Umlauf et al., 2018b},
-\citeproc{ref-bamlss2021}{2021}), brms v. 2.22.0
+(\citeproc{ref-assertthat}{Wickham, 2019}), brms v. 2.22.0
 (\citeproc{ref-brms2017}{Bürkner, 2017}, \citeproc{ref-brms2018}{2018},
-\citeproc{ref-brms2021}{2021}), changepoint v. 2.3
-(\citeproc{ref-changepoint2024}{Killick et al., 2024};
-\citeproc{ref-changepoint2014}{Killick \& Eckley, 2014}), doParallel v.
-1.0.17 (\citeproc{ref-doParallel}{Corporation \& Weston, 2022}),
-easystats v. 0.7.4 (\citeproc{ref-easystats}{Lüdecke et al., 2022}),
-foreach v. 1.5.2 (\citeproc{ref-foreach}{Microsoft \& Weston, 2022}),
-furrr v. 0.3.1 (\citeproc{ref-furrr}{Vaughan \& Dancho, 2022}), future
-v. 1.49.0 (\citeproc{ref-future}{Bengtsson, 2021}), ggrepel v. 0.9.6
+\citeproc{ref-brms2021}{2021}), doParallel v. 1.0.17
+(\citeproc{ref-doParallel}{Corporation \& Weston, 2022}), easystats v.
+0.7.4 (\citeproc{ref-easystats}{Lüdecke et al., 2022}), foreach v. 1.5.2
+(\citeproc{ref-foreach}{Microsoft \& Weston, 2022}), furrr v. 0.3.1
+(\citeproc{ref-furrr}{Vaughan \& Dancho, 2022}), future v. 1.58.0
+(\citeproc{ref-future}{Bengtsson, 2021}), ggpubr v. 0.6.0
+(\citeproc{ref-ggpubr}{Kassambara, 2023}), ggrepel v. 0.9.6
 (\citeproc{ref-ggrepel}{Slowikowski, 2024}), glue v. 1.8.0
@@
 (\citeproc{ref-scales}{Wickham et al., 2025}), scico v. 1.5.0
-(\citeproc{ref-scico}{T. L. Pedersen \& Crameri, 2023}), tictoc v. 1.2.1
+(\citeproc{ref-scico}{T. L. Pedersen \& Crameri, 2023}), signal v. 1.8.1
+(\citeproc{ref-signal}{signal developers, 2023}), tictoc v. 1.2.1
 (\citeproc{ref-tictoc}{Izrailev, 2024}), tidybayes v. 3.0.7
@@
 
-\bibitem[\citeproctext]{ref-coretta2025}
-Coretta, S., \& Bürkner, P.-. C. (2025). \emph{Bayesian beta regressions
-with brms in r: A tutorial for phoneticians}.
-\url{http://dx.doi.org/10.31219/osf.io/f9rqg_v1}
-
 \bibitem[\citeproctext]{ref-doParallel}
@@
 
+\bibitem[\citeproctext]{ref-ggpubr}
+Kassambara, A. (2023). \emph{{ggpubr}: {``{ggplot2}''} based publication
+ready plots}. \url{https://CRAN.R-project.org/package=ggpubr}
+
 \bibitem[\citeproctext]{ref-tidybayes}
@@
 
-\bibitem[\citeproctext]{ref-changepoint2014}
-Killick, R., \& Eckley, I. A. (2014). {changepoint}: An {R} package for
-changepoint analysis. \emph{Journal of Statistical Software},
-\emph{58}(3), 1--19.
-\url{https://www.jstatsoft.org/article/view/v058i03}
-
 \bibitem[\citeproctext]{ref-changepoint}
 Killick, R., Haynes, K., \& Eckley, I. A. (2022). \emph{{changepoint}:
-An {R} package for changepoint analysis}.
-\url{https://CRAN.R-project.org/package=changepoint}
-
-\bibitem[\citeproctext]{ref-changepoint2024}
-Killick, R., Haynes, K., \& Eckley, I. A. (2024). \emph{{changepoint}:
 An {R} package for changepoint analysis}.
@@
 \url{https://CRAN.R-project.org/package=MetBrewer}
-
-\bibitem[\citeproctext]{ref-bamlss2024}
-Muschinski, T., Mayr, G. J., Simon, T., Umlauf, N., \& Zeileis, A.
-(2024). Cholesky-based multivariate {G}aussian regression.
-\emph{Econometrics and Statistics}, \emph{29}, 261--281.
-\url{https://doi.org/10.1016/j.ecosta.2022.03.001}
 
@@
 
+\bibitem[\citeproctext]{ref-signal}
+signal developers. (2023). \emph{{signal}: Signal processing}.
+\url{https://r-forge.r-project.org/projects/signal/}
+
 \bibitem[\citeproctext]{ref-tidytext}
@@
 
-\bibitem[\citeproctext]{ref-bamlss2021}
-Umlauf, N., Klein, N., Simon, T., \& Zeileis, A. (2021). {bamlss}: A
-{L}ego toolbox for flexible {B}ayesian regression (and beyond).
-\emph{Journal of Statistical Software}, \emph{100}(4), 1--53.
-\url{https://doi.org/10.18637/jss.v100.i04}
-
 \bibitem[\citeproctext]{ref-umlauf2018}
-Umlauf, N., Klein, N., \& Zeileis, A. (2018a). BAMLSS: Bayesian Additive
+Umlauf, N., Klein, N., \& Zeileis, A. (2018). BAMLSS: Bayesian Additive
 Models for Location, Scale, and Shape (and Beyond). \emph{Journal of
@@
 \url{https://doi.org/10.1080/10618600.2017.1407325}
-
-\bibitem[\citeproctext]{ref-bamlss2018}
-Umlauf, N., Klein, N., \& Zeileis, A. (2018b). {BAMLSS}: {B}ayesian
-additive models for location, scale and shape (and beyond).
-\emph{Journal of Computational and Graphical Statistics}, \emph{27}(3),
-612--627. \url{https://doi.org/10.1080/10618600.2017.1407325}
 
@@
 \section{Application to 2D time-resolved decoding results
-(cross-temporal generalisation)}\label{sec-2D}
-
-Assume we have M/EEG data and we have conducted cross-temporal
-generalisation analyses (\citeproc{ref-king2014}{King \& Dehaene,
-2014}). As a result, we have a 2D matrix where each element contains the
-decoding accuracy (e.g., ROC AUC) of a classifier trained at timestep
-\(\text{training}_{i}\) and tested at timestep \(\text{testing}_{j}\)
-(Figure~\ref{fig-sim-timegen}).
+(cross-temporal generalisation)}\label{apx-2D}
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+We conducted a cross-temporal generalisation analysis of the decoding
+data from Nalborczyk et al. (\citeproc{ref-nalborczyk:inprep}{in
+preparation}), in which we assessed the performance of classifiers
+trained and tested at various timesteps of the trial
+(\citeproc{ref-king2014}{King \& Dehaene, 2014}). This analysis was
+performed at the participant level, resulting in a 2D matrix where each
+element contains the decoding accuracy (ROC AUC) of a classifier trained
+at timestep \(\text{training}_{i}\) and tested at timestep
+\(\text{testing}_{j}\) for each participant (Figure~\ref{fig-timegen}).
 
@@
 
-\caption{\label{fig-sim-timegen}Exemplary (simulated) group-level
-average cross-temporal generalisation matrix of decoding performance
-(ROC AUC).}
+\caption{\label{fig-timegen}Group-level average cross-temporal
+generalisation matrix of decoding performance (data from
+\citeproc{ref-nalborczyk:inprep}{Nalborczyk et al., in preparation}).}
 
@@
 
-\includegraphics[width=0.75\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-sim-timegen-1.png}
+\includegraphics[width=0.75\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-timegen-1.png}
 
@@
 
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
 To model cross-temporal generalisation matrices of decoding performance
-(ROC AUC), we extended the initial (decoding) GAM to take into account
-the bivariate temporal distribution of AUC values, thus producing
-naturally smoothed estimates (timecourses) of AUC values and posterior
+(ROC AUC), we extended our initial BGAM to take into account the
+bivariate temporal distribution of AUC values, thus producing naturally
+smoothed estimates (timecourses) of AUC values and posterior
 probabilities. This model can be written as follows:
@@
 \]
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
 
@@
 
-Note that \(b_{m}(,)\) is a basis function that maps
-\(R \times R \rightarrow R\). A popular bivariate basis function uses
-\emph{thin-plate splines} (\citeproc{ref-wood2003}{Wood, 2003a}), which
-extend to \(\mathbf{s}_{i} \in \mathbb{R}^{d}\) and \(\partial l_{g}\)
-penalties. These splines are designed to interpolate and approximate
-smooth surfaces over two dimensions (hence the ``bivariate'' term). For
-\(d=2\) dimensions and \(l=2\) (smoothness penalty involving second
-order derivative):
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+A popular bivariate basis function uses \emph{thin-plate splines}
+(\citeproc{ref-wood2003}{Wood, 2003a}). These splines are designed to
+interpolate and approximate smooth surfaces over two dimensions (hence
+the ``bivariate'' term). For \(d=2\) dimensions and \(l=2\) (smoothness
+penalty involving second order derivative):
 
@@
 
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
 where \(\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|\) is the Euclidean
 distance between the covariate \(\mathbf{s}_{i}\) and the knot location
-\(\mathbf{k}_{m}\). We fitted this model using \texttt{brms}\ldots{}
-using the \texttt{t2()} smooth constructor with full penalties
+\(\mathbf{k}_{m}\). We fitted this model using \texttt{brms} and the
+\texttt{t2()} tensor product smooth constructor with full penalties
 (\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al.,
-2019})\ldots{}
+2019}). We ran eight MCMCs to approximate the posterior distribution,
+including each 5000 iterations and a warmup of 2000 iterations, yielding
+a total of \(8 \times (5000-2000) = 24000\) posterior samples to be used
+for inference.
 
@@
 \NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-    \CommentTok{\# 2D thin{-}plate spline (tp)}
-    \CommentTok{\# auc \textasciitilde{} t2(train\_time, test\_time, bs = "tp", k = 10),}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{),}
+    \CommentTok{\# 2D thin{-}plate spline (tp) with full penalties}
+\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{30}\NormalTok{, }\AttributeTok{full =} \ConstantTok{TRUE}\NormalTok{),}
     \AttributeTok{data =}\NormalTok{ timegen\_data,}
-    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{warmup =} \DecValTok{2000}\NormalTok{,}
+    \AttributeTok{family =}\NormalTok{ brms}\SpecialCharTok{::}\FunctionTok{Beta}\NormalTok{(),}
+    \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
     \AttributeTok{iter =} \DecValTok{5000}\NormalTok{,}
@@
     \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
-    \CommentTok{\# file = "models/timegen\_gam\_t2.rds" \# k = 10}
-    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_t2\_k20.rds"} \CommentTok{\# k = 20}
+    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{15}\NormalTok{),}
+    \AttributeTok{backend =} \StringTok{"cmdstanr"}\NormalTok{,}
+    \AttributeTok{stan\_model\_args =} \FunctionTok{list}\NormalTok{(}\AttributeTok{stanc\_options =} \FunctionTok{list}\NormalTok{(}\StringTok{"O1"}\NormalTok{) ),}
+    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_meg\_t2\_k30\_full.rds"}
 \NormalTok{    )}
-
-\CommentTok{\# fitting a GAM with two temporal dimensions to empirical data}
-\NormalTok{timegen\_gam }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
-    \CommentTok{\# 2D thin{-}plate spline (tp)}
-    \CommentTok{\# auc \textasciitilde{} t2(train\_time, test\_time, bs = "tp", k = 30),}
-\NormalTok{    auc }\SpecialCharTok{\textasciitilde{}} \FunctionTok{t2}\NormalTok{(train\_time, test\_time, }\AttributeTok{bs =} \StringTok{"tp"}\NormalTok{, }\AttributeTok{k =} \DecValTok{20}\NormalTok{, }\AttributeTok{full =} \ConstantTok{TRUE}\NormalTok{),}
-    \AttributeTok{data =}\NormalTok{ timegen\_data,}
-    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
-    \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
-    \AttributeTok{iter =} \DecValTok{3000}\NormalTok{,}
-    \AttributeTok{chains =} \DecValTok{8}\NormalTok{,}
-    \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
-    \CommentTok{\# file = "models/timegen\_gam\_meg\_t2\_k30.rds"}
-    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_meg\_t2\_k20\_full.rds"}
-\NormalTok{    )}
-
-\CommentTok{\# retrieving the data}
-\CommentTok{\# timegen\_data \textless{}{-} timegen\_gam$data}
-
-\CommentTok{\# PPCs}
-\CommentTok{\# pp\_check(timegen\_gam)}
 \end{Highlighting}
@@
 
-Could be extended to spatial and temporal dimensions with formulas such
-as \texttt{te(x,\ y,\ Time,\ d\ =\ c(2,\ 1)\ )}\ldots{}
+\begin{figure}[!htb]
+
+\caption{\label{fig-gam-timegen-post-preds}Predicted AUC values with
+threshold (left) and posterior odds of decoding accuracy being above
+chance (right) according to the bivariate BGAM.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{manuscript_files/figure-pdf/fig-gam-timegen-post-preds-1.png}
+
+}
+
+\end{figure}%
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+Figure~\ref{fig-gam-timegen-post-preds} shows the predictions from the
+model (left) superimposed with the identified cluster as defined by
+thresholding the posterior odds (right). Notably, this model could be
+extended to a multilevel bivariate GAM via
+\texttt{t2(train\_time,\ test\_time,\ participant,\ bs\ =\ c("tp",\ "tp",\ "re"),\ m\ =\ 2,\ full\ =\ TRUE)}
+and could be generalised to account for both spatial (\texttt{x1} and
+\texttt{x2}) and temporal (\texttt{time}) dimensions with formulas such
+as \texttt{te(x,\ y,\ time,\ d\ =\ c(2,\ 1)\ )}. The performance of such
+models should be assessed in future work.
 
@@
 
-\section{Alternative to GAMs: Approximate Gaussian Process
-regression}\label{alternative-to-gams-approximate-gaussian-process-regression}
-
-A Gaussian process (GP) is a stochastic process that defines the
-distribution over a collection of random variables indexed by a
-continuous variable, that is \(\{f(t): t \in \mathcal{T}\}\) for some
-index set \(\mathcal{T}\) (\citeproc{ref-rasmussen2005}{Rasmussen \&
-Williams, 2005};
-\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
-2023}). Whereas Bayesian linear regression outputs a distribution over
-the parameters of some predefined parametric model, the GP approach, in
-contrast, is a non-parametric approach, in that it finds a distribution
-over the possible functions that are consistent with the observed data.
-However, note that nonparametric does not mean there aren't parameters,
-it means that there are infinitely many parameters.
-
-From
-\href{https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp}{brms
-documentation}: A GP is a stochastic process, which describes the
-relation between one or more predictors
-\(x=\left(x_{1}, \ldots, x_{d}\right)\) and a response \(f(x)\), where
-\(d\) is the number of predictors. A GP is the generalization of the
-multivariate normal distribution to an infinite number of dimensions.
-Thus, it can be interpreted as a prior over functions. The values of
-\(f()\) at any finite set of locations are jointly multivariate normal,
-with a covariance matrix defined by the covariance kernel
-\(k_p\left(x_i, x_j\right)\), where \(p\) is the vector of parameters of
-the GP:
-
-\[
-\left(f\left(x_{1}\right), \ldots f\left(x_{n}\right) \sim \operatorname{MVN}\left(0, \left(k_p\left(x_{i}, x_{j}\right)\right)_{i, j=1}^{n}\right)\right.
-\]
-
-The smoothness and general behaviour of the function \(f\) depends only
-on the choice of covariance kernel, which ensures that values that are
-close together in the input space will be mapped to similar output
-values\ldots{}
-
-From this perspective, \(f\) is a realisation of an infinite dimensional
-normal distribution:
-
-\[
-f \sim \mathrm{Normal}(0, C(\lambda))
-\]
-
-where \(C\) is a covariance kernel with hyperparameters \(\lambda\) that
-defines the covariance between two function values \(f\left(t_1\right)\)
-and \(f\left(t_2\right)\) for two time points \(t_1\) and \(t_2\)
-(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). Similar to
-the different choices of the basis function for splines, different
-choices of the covariance kernel lead to different GPs. In this article,
-we consider the squared-exponential (a.k.a. radial basis function)
-kernel, which computes the squared distance between points and converts
-it into a measure of similarity. It is defined as:
-
-\[
-C(\lambda) := C\left(t_1, t_2, \sigma, \gamma\right) := \sigma^2 \exp \left(-\frac{||t_1-t_2||^{2}}{2 \gamma^2}\right)
-\]
-
-with hyperparameters \(\lambda = (\sigma, \gamma)\), expressing the
-overall scale of GP and the length-scale, respectively
-(\citeproc{ref-rasmussen2005}{Rasmussen \& Williams, 2005}). The
-advantages of this kernel are that it is computationally efficient and
-(infinitely) smooth making it a reasonable choice for the purposes of
-the present article. Here again, \(\lambda\) hyperparameters are
-estimated from the data, along with all other model parameters.
-
-Taken from
-\url{https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10c-Gaussian-processes.html}:
-For a given vector \(\mathbf{x}\), we can use the kernel to construct
-finite multi-variate normal distribution associated with it like so:
-
-\[
-\mathbf{x} \mapsto_{G P} \operatorname{MVNormal}(m(\mathbf{x}), k(\mathbf{x}, \mathbf{x}))
-\]
-
-where \(m\) is a function that specifies the mean for the distribution
-associated with \(\mathbf{x}\). This mapping is essentially the Gaussian
-process: a systematic association of vectors of arbitrary length with a
-suitable multi-variate normal distribution.
-
-Low-rank approximate Gaussian processes are of main interest in machine
-learning and statistics due to the high computational demands of exact
-Gaussian process models
-(\citeproc{ref-riutort-mayol_practical_2023}{Riutort-Mayol et al.,
-2023})\ldots{}
-
-\newpage
-
-\section{How to choose the GAM basis dimension?}\label{sec-basis}
-
-Here we should provide recommendation about how to define
-\texttt{k}\ldots{} An option is to vary \texttt{k} and examine the
-predictions and posterior predictive checks (PPCs) of each model\ldots{}
-In this example (Figure~\ref{fig-choose-k})\ldots{} However, it is not
-possible to provide general recommendations, as the optimal \texttt{k}
-depends on the sampling rate, the preprocessing steps (e.g.,
-signal-to-noise ratio, low-pass filtering, etc), and the neural dynamics
-of the phenomenon under study. Choose \texttt{k} as large a possible
-(given limits on computation time etc)? Or use LOOIC/WAIC (seem to
-select too simple models)?
+\section{How to choose the GAM basis dimension?}\label{apx-basis}
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+There is no universal recommendation for choosing the optimal value of
+\texttt{k}, as it depends on several factors, including the sampling
+rate, preprocessing steps (e.g., signal-to-noise ratio, low-pass
+filtering), and the underlying neural dynamics of the phenomenon under
+investigation. One strategy is to set \texttt{k} as high as
+computational constraints allow, as suggested by previous authors (e.g.,
+\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al., 2019}).
+Alternatively, one can fit a series of models with different \texttt{k}
+values and compare them using information criteria such as LOOIC or
+WAIC, alongside posterior predictive checks (PPCs), to select the model
+that best captures the structure of the data. We illustrate this
+approach below.
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+Figure~\ref{fig-choose-k} presents the posterior predictions and two
+forms of posterior predictive checks (PPCs) for each GAM fit using
+different numbers of basis functions (\(k \in \{5, 10, 20, 40, 80\}\)).
+With the exception of the \(k=5\) model, all other fits yield
+satisfactory PPCs, indicating that the predicted data closely resemble
+the empirical observations. However, model comparison using the
+leave-one-out information crtierio (LOOIC), as summarised in
+Table~\ref{tbl-simulation-results}, identifies the \(k=40\) model as the
+best-performing one in terms of LOOIC, closely followed by the \(k=80\)
+model. This suggests that the optimal number of basis functions likely
+lies between these two values. Future simulation studies could further
+investigate how such model selection criteria relate to the precision of
+onset and offset estimates.
+
+\begin{table}
+
+{\caption{{Models comparison with LOOIC. Models are arranged by the
+diffence in expected log-pointwise density (ELPD) to the best model
+(i.e., k=40).}{\label{tbl-k-results}}}}
+
+\fontsize{7.5pt}{9.0pt}\selectfont
+\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lcccccccc}
+\toprule
+k & elpd\_diff & se\_diff & elpd\_loo & se\_elpd\_loo & p\_loo & se\_p\_loo & looic & se\_looic \\ 
+\midrule\addlinespace[2.5pt]
+40 & 0.00 & 0.00 & 5,579.11 & 69.47 & 47.79 & 0.74 & -11,158.21 & 138.94 \\ 
+80 & -0.63 & 0.38 & 5,578.47 & 69.48 & 50.30 & 0.78 & -11,156.94 & 138.97 \\ 
+20 & -6.43 & 3.38 & 5,572.68 & 69.59 & 36.99 & 0.58 & -11,145.35 & 139.17 \\ 
+10 & -36.12 & 8.85 & 5,542.98 & 69.88 & 19.29 & 0.32 & -11,085.97 & 139.76 \\ 
+5 & -1,605.12 & 53.40 & 3,973.99 & 71.54 & 11.13 & 0.19 & -7,947.98 & 143.09 \\ 
+\bottomrule
+\end{tabular*}
+
+\end{table}
 
@@
 \section{\texorpdfstring{\texttt{R} package and integration with
-\texttt{MNE-Python}}{R package and integration with MNE-Python}}\label{r-package-and-integration-with-mne-python}
-
-For users who are already familiar with \texttt{brms}, the recommended
-pipeline is to import ERPs or decoding results in \texttt{R} and analyse
-these data using the code provided in the main paper. It is also
-possible to call functions from the \texttt{neurogam\ R} package
-(available at \url{https://github.com/lnalborczyk/neurogam}) which come
-with sensible defaults.
+\texttt{MNE-Python}}{R package and integration with MNE-Python}}\label{apx-package}
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+For readers who are already familiar with \texttt{brms}, the recommended
+pipeline is to use the code provided in the main paper (available at
+\url{https://github.com/lnalborczyk/brms_meeg}). It is also possible to
+call functions from the \texttt{neurogam\ R} package (available at
+\url{https://github.com/lnalborczyk/neurogam}) which come with sensible
+defaults.
 
@@
 \CommentTok{\# using the testing\_through\_time() function from the neurogam package}
-\CommentTok{\# this may take a few minutes (or hours depending the machine\textquotesingle{}s}
-\CommentTok{\# performance and data size)...}
+\CommentTok{\# this may take a few minutes (or hours depending on the machine\textquotesingle{}s}
+\CommentTok{\# performance and the size of the dataset)...}
 \NormalTok{gam\_onset\_offset }\OtherTok{\textless{}{-}} \FunctionTok{testing\_through\_time}\NormalTok{(}
@@
     \AttributeTok{data =}\NormalTok{ raw\_df,}
-    \CommentTok{\# threshold for defining clusters (20 by default)}
-    \AttributeTok{threshold =} \DecValTok{20}\NormalTok{,}
     \CommentTok{\# the *\_id arguments are used to specify the relevant columns in data }
@@
     \AttributeTok{time\_id =} \StringTok{"time"}\NormalTok{, }\AttributeTok{predictor\_id =} \StringTok{"condition"}\NormalTok{,}
+    \CommentTok{\# posterior odds threshold for defining clusters (20 by default)}
+    \AttributeTok{threshold =} \DecValTok{20}\NormalTok{,}
     \CommentTok{\# number of warmup MCMC iterations}
@@
 
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
 The \texttt{neurogam} package can also be called from \texttt{Python}
@@
 \texttt{MNE-Python} pipelines. For example, we use it below to estimate
-the onset and offset of effects for one EEG channel from a MNE evoked
-object. Note that the code used to reshape the \texttt{sample\ MNE}
-dataset is available in the online supplementary materials, and we refer
-to the
+the onset and offset of effects for one EEG channel from an MNE evoked
+object. The code used to reshape the \texttt{sample\ MNE} dataset is
+available in the online supplementary materials, and we further refer to
+the
 \href{https://mne.tools/stable/auto_tutorials/epochs/50_epochs_to_data_frame.html}{MNE
@@
 \NormalTok{    data}\OperatorTok{=}\NormalTok{reshaped\_df\_r,}
-\NormalTok{    threshold}\OperatorTok{=}\DecValTok{10}\NormalTok{,}
+\NormalTok{    threshold}\OperatorTok{=}\DecValTok{20}\NormalTok{,}
 \NormalTok{    multilevel}\OperatorTok{=}\VariableTok{False}

### 2025-08-22 code/errors_simulation_cluster.R

@@ -2,10 +2,11 @@
 # Monte-Carlo simulation of onset/offset error properties #
 # Written by Ladislas Nalborczyk                          #
 # Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on May 9, 2025                             #
+# Last updated on August 7, 2025                          #
 ###########################################################
 
-options(error = function() {traceback(2); quit("no", status = 1, runLast = FALSE)})
+# logging error messages on the HPC cluster
+options(error = function() { traceback(2); quit("no", status = 1, runLast = FALSE) })
 
 # importing R packages
 library(changepoint)
@@ -48,10 +49,10 @@ source("code/eeg_noise.R")
 source("code/erp_template.R")
 
 # to use with the eeg_noise function
-# meanpower <- unlist(read.table("code/meanpower.txt") )
+meanpower <- unlist(read.table("code/meanpower.txt") )
 
 # defining the total number of simulations to be performed
-n_sims <- 8
+n_sims <- 500
 
 # defining simulation parameters
 n_trials <- 50 # number of trials
@@ -70,13 +71,12 @@ significance_thresholds <- c(0.05, 0.01, 0.005, 0.001)
 # defining possible (posterior probability ratio) threshold values
 threshold_values <- c(1, 3, 6, 10, 20, 50, 100)
 
-# initialising empty simulation results
-# sim_results <- data.frame()
-
 # for each simulation
 sim_results <- foreach(
-    i = seq_len(n_sims), .combine = bind_rows,
-    .packages = c("brms", "dplyr", "tidyr", "tidybayes", "changepoint", "reticulate")
+    i = seq_len(n_sims),
+    .combine = bind_rows,
+    .errorhandling = "pass",
+    .packages = c("reticulate", "brms", "dplyr", "tidyr", "tidybayes", "changepoint")
     ) %dopar% {
         
         # formatting the current sim ID
@@ -84,13 +84,14 @@ sim_results <- foreach(
         
         # importing Python modules locally
         message_parallel("\nImporting Python modules...")
-        use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
+        # use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
         
         # importing Python modules on the HPC cluster
         # py_require(packages = c("numpy", "mne"), python_version = "3.12.0")
-        # np <- import("numpy")
[…]

### 2025-08-22 code/peaks_across_splits.R

@@ -1,0 +1,120 @@
+###########################################################
+# Defining subsets of participants with uniform Jaccard   #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on July 15, 2025                            #
+###########################################################
+
+# importing R packages
+library(tidyverse)
+library(signal)
+
+# for setting up the cluster
+library(doParallel)
+library(foreach)
+
+# total number of total CPU cores available on my laptop
+total_cores <- 12
+
+# registering parallel backend
+cl <- makeCluster(total_cores)
+registerDoParallel(cl)
+
+# importing home-made helper functions
+source("code/utils.R")
+
+# importing the data splits list
+# split_df <- read.csv(file = "uniform_jaccard_data_splits.csv")
+split_df <- read.csv(file = "uniform_jaccard_data_splits_100pairs.csv")
+
+# importing the reshaped MEG decoding results
+raw_df <- read.csv(file = "data/decoding_results_reshaped.csv") |>
+    # removing the last participant (to get a round number of participants)
+    mutate(participant_id = cur_group_id(), .by = participant) |>
+    dplyr::filter(participant_id < 33) |>
+    select(-participant_id)
+
+# averaging across participants
+summary_df <- raw_df |>
+    summarise(
+        auc_mean = mean(auc),
+        auc_sd = sd(auc),
+        auc_se = sd(auc) / sqrt(n() ),
+        .by = c(participant, time)
+        )
+
+# creating a dataframe with all (split_number, group) combinations to iterate over
+split_group_combinations <- crossing(
+    split_number = split_df$split_id,
+    group = split_df$group
+    ) |>
+    # creating a custom simulation ID per group
+    mutate(split_id = sprintf("s%05d_%s", split_number, group) )
+
+# for each simulation
+reliability_results <- foreach(
+    # i = seq_len(n_splits), .combine = bind_rows,
+    i = 1:nrow(split_group_combinations), .combine = bind_rows,
+    .packages = c("dplyr", "tidyr", "signal")
+    ) %dopar% {
+    
+        # extracting split_number and group for this simulation
+        current_split <- split_group_combinations$split_number[i]
+        current_group <- split_group_combinations$group[i]
+        current_split_id <- split_group_combinations$split_id[i]
+        
+        # printing progress
+        message_parallel(sprintf("\nRunning simulation: %s (%d out of %d)...", current_split_id, i, nrow(split_group_combinations) ) )
+        
[…]

### 2025-08-22 code/remaining_reliability_simulations.R

@@ -1,0 +1,57 @@
+###########################################################
+# Finding simulations IDs that did not finish             #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on August 7, 2025                          #
+###########################################################
+
+library(tidyverse)
+
+# importing original splits
+splits <- read.csv(file = "data/uniform_jaccard_data_splits_1000pairs.csv")
+
+# sanity check
+range(splits$split_id)
+
+# listing results files
+files <- list.files(
+    path = "reliability_jaccard_cluster_results",
+    pattern = ".rds", 
+    full.names = TRUE
+    )
+
+# sorting files numerically by the number in the filename
+files <- files[order(readr::parse_number(files) )]
+
+# importing these files
+reliability_results <- map2_dfr(
+    .x = files,
+    .y = seq_along(files),
+    .f = ~ readRDS(.x)
+    )
+
+# finding splits IDs that are in splits$split_id but not in reliability_results$split_id
+splits_in_results <- reliability_results %>%
+    mutate(split_id = str_sub(string = split_id, start = 7, end = 11) ) %>%
+    pull(split_id) %>%
+    unique()
+    
+all_splits <- sprintf("%05d", splits$split_id) %>% unique()
+
+# splits IDs that are both in splits$split_id and reliability_results$split_id
+intersect(splits_in_results, all_splits) %>% length()
+intersect(splits_in_results, all_splits)
+
+# splits IDs that are in splits$split_id but not in reliability_results$split_id
+setdiff(all_splits, splits_in_results) %>% length()
+setdiff(all_splits, splits_in_results)
+
+# exporting a new csv file with remaining simulation IDs
+remaining_splits <- splits %>%
+    filter(sprintf("%05d", splits$split_id) %in% setdiff(all_splits, splits_in_results) )
+
+# sanity check
+remaining_splits %>% pull(split_id) %>% n_distinct()
+
+# exporting it
+write.csv(x = remaining_splits, file = "data/uniform_jaccard_data_splits_1000pairs_remaining2.csv", row.names = FALSE)

### 2025-08-22 code/split_half_reliability_cluster.R

@@ -1,525 +1,0 @@
-###########################################################
-# Split-half reliability of onset/offset estimates        #
-# Written by Ladislas Nalborczyk                          #
-# Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on May 8, 2025                             #
-###########################################################
-
-options(error = function() {traceback(2); quit("no", status = 1, runLast = FALSE)})
-
-# importing R packages
-library(changepoint)
-library(reticulate)
-library(tidybayes)
-library(dplyr)
-library(tidyr)
-library(glue)
-library(brms)
-
-# for setting up the cluster
-library(doParallel)
-library(foreach)
-
-# retrieving the array ID
-args <- commandArgs(trailingOnly = TRUE)
-array_id <- as.numeric(args[1])
-cat("Running job for array_id =", array_id, "\n")
-
-# total number of total CPU cores available on the HPC node
-total_cores <- 16
-
-# number of cores per brms model
-cores_per_model <- 4
-
-# number of parallel simulations to run at once
-n_parallel <- total_cores / cores_per_model
-
-# registering parallel backend
-cl <- makeCluster(n_parallel)
-registerDoParallel(cl)
-
-# importing home-made helper functions
-source("code/utils.R")
-
-# importing the reshaped MEG decoding results
-raw_df <- read.csv(file = "data/decoding_results_reshaped.csv") |>
-    # removing the last participant (to get a round number of participants)
-    mutate(participant_id = cur_group_id(), .by = participant) |>
-    filter(participant_id < 33) |>
-    select(-participant_id)
-
-# averaging across participants
-summary_df <- raw_df |>
-    summarise(
-        auc_mean = mean(auc),
-        auc_sd = sd(auc),
-        auc_se = sd(auc) / sqrt(n() ),
-        .by = c(participant, time)
-        )
-
-# computing the smallest effect size of interest (SESOI)
-# as the SD of decoding performance during the baseline
-full_sesoi <- summary_df %>%
-    filter(time < 0) %>%
-    summarise(auc_mean = mean(auc_mean), .by = time) %>%
-    summarise(sesoi = sd(auc_mean) ) %>%
-    pull(sesoi)
-
-# getting unique participants
-unique_participants <- unique(raw_df$participant)
-
-# getting the number of participants
-n_participants <- as.numeric(length(unique_participants) )
-
-# defining the total number of splits (per node)
-n_splits <- 50
-
[…]

### 2025-08-22 code/split_half_reliability_cluster_uniform_jaccard.R

@@ -1,0 +1,396 @@
+###########################################################
+# Split-half reliability of onset/offset estimates        #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on August 7, 2025                          #
+###########################################################
+
+# logging error messages on the HPC cluster
+# options(error = function () {traceback(2); quit("no", status = 1, runLast = FALSE)})
+
+# importing R packages
+library(changepoint)
+library(reticulate)
+library(tidybayes)
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(glue)
+library(brms)
+
+# for setting up the cluster
+library(doParallel)
+library(foreach)
+
+# retrieving the array ID
+# ensuring this is passed in the SLURM job array script: #SBATCH --array=1-50
+args <- commandArgs(trailingOnly = TRUE)
+# array_id <- as.numeric(args[1])
+array_id_increment <- 200 # id increment for remaining simulations
+array_id <- array_id_increment + as.numeric(args[1])
+cat("Running job for array_id =", array_id, "\n")
+
+# total number of total CPU cores available on the HPC node
+total_cores <- 16
+
+# number of cores per brms model
+cores_per_model <- 4
+
+# number of parallel simulations to run at once
+n_parallel <- total_cores / cores_per_model
+
+# registering parallel backend
+cl <- makeCluster(n_parallel)
+registerDoParallel(cl)
+
+# importing home-made helper functions
+source("code/utils.R")
+
+# importing the reshaped MEG decoding results
+raw_df <- read.csv(file = "data/decoding_results_reshaped.csv") |>
+    # removing the last participant (to get a round number of participants)
+    mutate(participant_id = cur_group_id(), .by = participant) |>
+    dplyr::filter(participant_id < 33) |>
+    select(-participant_id)
+
+# averaging across participants
+summary_df <- raw_df |>
+    summarise(
+        auc_mean = mean(auc),
+        auc_sd = sd(auc),
+        auc_se = sd(auc) / sqrt(n() ),
+        .by = c(participant, time)
+        )
+
+# defining the decoding chance level
+chance_level <- 0.5
+
+# importing the data splits list
+# split_df <- read.csv(file = "data/uniform_jaccard_data_splits_100pairs.csv")
+# split_df <- read.csv(file = "data/uniform_jaccard_data_splits_1000pairs.csv")
[…]

### 2025-08-22 code/uniform_jaccard.R

@@ -1,0 +1,255 @@
+###########################################################
+# Defining subsets of participants with uniform Jaccard   #
+# Written by Ladislas Nalborczyk                          #
+# Contact: ladislas.nalborczyk@gmail.com                  #
+# Last updated on July 30, 2025                           #
+###########################################################
+
+# importing R packages
+library(tidyverse)
+library(glue)
+
+# importing the reshaped MEG decoding results
+raw_df <- read.csv(file = "data/decoding_results_reshaped.csv") |>
+    # removing the last participant (to get a round number of participants)
+    mutate(participant_id = cur_group_id(), .by = participant) |>
+    dplyr::filter(participant_id < 33) |>
+    select(-participant_id)
+
+# getting unique participants
+unique_participants <- unique(raw_df$participant)
+
+# getting the number of participants
+n_participants <- as.numeric(length(unique_participants) )
+
+# defining the total number of splits
+# n_splits <- 50
+
+# creating a dataframe where each participant appears in every split
+# split_df <- crossing(participant = unique_participants, split_number = 1:n_splits)
+
+# assigning alternating groups (A or B) within each split for each participant
+# split_df <- split_df |>
+#     group_by(split_number) |>
+#     # shuffling groups for each split
+#     mutate(group = sample(rep(c("A", "B"), length.out = n() ) ) ) |>
+#     ungroup()
+
+# computing all possible Jaccard values for overlaps 0 to 16
+overlaps <- 0:16
+jaccard_values <- overlaps / (n_participants - overlaps)
+jaccard_values <- round(jaccard_values, 5) # for numerical clarity
+
+jaccard_table <- data.frame(
+    overlap = overlaps,
+    jaccard = jaccard_values
+    )
+
+# checking the results
+print(jaccard_table)
+
+# defining some variables
+n_total <- n_participants
+subset_size <- n_participants / 2
+participants <- as.character(1:n_total)
+
+# storing pairs
+jaccard_exact_pairs <- vector("list", length(overlaps) )
+names(jaccard_exact_pairs) <- paste0("J_", jaccard_values)
+
+# desired number of pairs
+n_pairs <- 1000
+
+# getting 1000 pairs of subsets for each level of Jaccard similarity
+for (i in seq_along(overlaps) ) {
+    
[…]

### 2025-08-22 code/utils.R

@@ -120,12 +120,15 @@ find_clusters <- function (data, threshold = 10, above_threshold = TRUE) {
 freq_stats_cluster_matrix <- function (
         X, timesteps,
         cluster_type = c("mass", "tfce"),
-        alpha_level = 0.05
+        alpha_level = 0.05,
+        verbose = FALSE
         ) {
     
     # converting R matrix to NumPy array
     X_np <- np$array(X, dtype = "float64")
-    message(print(X_np) )
+    
+    # sanity check
+    if (verbose) {message(print(X_np) )}
     
     # defining the type of cluster-based method to use
     if (cluster_type == "mass") {
@@ -154,20 +157,21 @@ freq_stats_cluster_matrix <- function (
             threshold = threshold,
             n_permutations = as.integer(2^12),
             n_jobs = as.integer(1),
-            verbose = TRUE
+            verbose = verbose
             )
-    }, error = function(e) {
+    }, error = function (e) {
         message("Python error: ", e$message)
         message(reticulate::py_last_error() )
         return(NULL)
     })
     
-    if (is.null(results)) {
+    if (is.null(results) ) {
         message(reticulate::py_last_error() )
         stop("Python function failed.")
     }
     
-    message("Length of results: ", length(results) )
+    # sanity check
+    if (verbose) {message("Length of results: ", length(results) )}
     
     if (length(results) < 3) {
         stop("Python function did not return 3 results as expected.")
@@ -187,6 +191,8 @@ freq_stats_cluster_matrix <- function (
         # defining the (main) cluster onset and offset
         main_cluster_onset <- clusters[which(p_values < alpha_level)][[1]][[1]]$start+1
         main_cluster_offset <- clusters[which(p_values < alpha_level)][[1]][[1]]$stop
+        
+        # old
         # main_cluster_offset <- clusters[which(p_values < alpha_level)][[1]][[1]]$stop+1
         
         # assigning p-values to clusters
@@ -196,6 +202,8 @@ freq_stats_cluster_matrix <- function (
             cluster_mask <- clusters[[i]][[1]]
             idx_start <- cluster_mask$start+1
             idx_stop <- cluster_mask$stop
+            
+            # old
             # idx_stop <- cluster_mask$stop+1
             
             # retrieving the p-value for this cluster

### 2025-08-27 manuscript.qmd
pushing final version of manuscript

@@
 # see available font via systemfonts::system_fonts()
-# monofont: Monaco
-# monofont: Courier
 mainfont: CMU Serif
@@
 lang: en
-toc: true
+toc: false
 language:
@@
     documentmode: doc # man, jou, doc, stu
-    # number_sections: true
     keep-tex: true
@@
 The issues with multiple comparisons represent a common and well-recognised danger in neuroimaging and M/EEG research, where the collected data allows for a multitude of potential hypothesis tests and is characterised by complex structures of spatiotemporal dependencies.
-The probability of obtaining at least one false positive in an ensemble (family) of $m$ tests (i.e., the FWER) is computed as $1-\left(1-\alpha\right)^{m}$ (for $m=10$ independent tests and $\alpha=0.05$, it is approximately equal to $0.4$).
+The probability of obtaining at least one false positive in an ensemble (family) of $m$ independent tests (i.e., the FWER) is computed as $1-\left(1-\alpha\right)^{m}$ (for $m=10$ independent tests and $\alpha=0.05$, it is approximately equal to $0.4$).
 Different methods exist to control the FWER, that is, to bring it back to $\alpha$.
@@
 
-<!--
-
-Weighting single trials [@pernet2022]... From @dimigen2021: Recently, spline regression has been applied to ERPs [e.g., @hendrix2017; @kryuchkova2012]...
-
-M/EEG recordings at the scalp measure activity stemming from various parts of the brain and contain many artefacts not related to the cognitive and neural processes of interest.
-Therefore, analysing M/EEG data requires methods for separating the parts of the signals that are related to cognitive and neural processes and the parts of the signals that contain task-unrelated "noise".
-The most widely-used method for achieving this separation consists in estimating the event-related potential (ERP), defined as the brain response elicited by some specific time-locked sensory, cognitive, or motor event.
-Most often, the ERP is computed as the average response over trials computed from epochs time-locked on the event of interest (e.g., stimulus onset).
-However, this simple averaging approach comes with several limitations, such as being limited to simple categorical designs.
-
--->
-
 Scalp-recorded M/EEG signals capture neural activity originating from various brain regions and are often contaminated by artifacts unrelated to the cognitive processes under investigation.
@@
 In other words, more complex (wiggly) basis functions are automatically penalised.
-
-<!--
-
-A detailed discussion of the technical aspects of smooth functions is beyond the scope of this article (see Wood, 2017a); or (Baayen, Vasishth, Kliegl, & Bates, 2017; van Rij, Vaci, Wurm, & Feldman, in press; Wieling, 2018, for an introduction), but in the context of this work, it can be thought of as a continuous, potentially wiggly but not abruptly changing line that is expressed over time.
-
-Generalised additive models can be understood in a Bayesian framework [@miller2025; @wood2017]...
-
-## Bayesian generalised additive multilevel models
-
-There are debates among Bayesian practitioners as to whether prior distributions should encore subjective (personal) beliefs or...
-but these debates are outside the scope of the present paper and we therefore the interested reader to dedicated work (e.g., XX; YY)...
-In practice, weakly informative priors are often used as default priors in situations in which subjective priors are difficult to define/elicit...
-
-The Bayesian approach to statistical modelling is characterised by its reliance on probability theory to quantify uncertainty about parameter estimates [e.g., @gelman2020].
-In this framework, all unknown entities are assigned probability distributions reflecting the uncertainty...
-These probability distributions are commonly referred to as "priors" and represent some state of knowledge about unknown quantities before seeing any data.
-Bayesian models are then fitted on empirical (actual or simulated) data to update prior states of knowledge to posterior states of knowledge using Bayes theorem, or in practise, sampling-based approximations of the posterior distribution...
-
--->
 
@@
         geom_line(
-            # data = basis_long, aes(x = x, y = value, color = basis),
             data = basis_long, aes(x = x, y = value, group = basis),
@@
             ) +
-        # geom_text(
-        #     data = label_df,
-        #     aes(x = x, y = value, label = label, color = basis),
-        #     hjust = 0, nudge_x = 0.1, size = 2
-        #     ) +
-        # ggrepel::geom_text_repel(
-        #     data = label_df,
-        #     aes(x = x, y = value, label = label, color = basis),
-        #     hjust = 0, nudge_x = 1,
-        #     size = 2, direction = "y"
-        #     ) +
         labs(x = "Predictor", y = "Response") +
@@
 # plotting, note that the basis plot uses the fixed-effect model for simplicity
-# df2 <- df %>% filter(subject %in% c(unique(df$subject)[1:4]) )
 p1 <- plot_basis_functions(df, gam_fixed)
@@
 ```
-
-<!--
-
-Or (Baayen, Vasishth, Kliegl, & Bates, 2017; van Rij, Vaci, Wurm, & Feldman, in press; Wieling, 2018, for an introduction), but in the context of this work, it can be thought of as a continuous, potentially wiggly but not abruptly changing line that is expressed over time.
-
-This penalises complexity automatically [@meulman2023; @wieling2018]...
-ERP itself [@abugaber2023; @baayen2018; @meulman2015; @meulman2023]...
-See for instance these tutorials [@sóskuthy2017; @winter2016] or application to phonetic data [@wieling2018; @sóskuthy2021]...
-or this introduction [@baayen2020] or these reference books [@hastie2017; @wood2017]...
-
-A detailed discussion of the technical aspects of GAMs is beyond the scope of this article [see reference books such as @hastie2017; @wood2017] but it should be noted that GAMs have been successfully been applied to various types of timeseries data throughout the cognitive sciences.
-For instance, it has been applied to pupillometry [e.g., @vanrij2019], articulography [e.g., @wieling2018], formant contours [e.g., @sóskuthy2021], neuroimaging data [e.g., @dinga2021] and ERPs [e.g., @abugaber2023; @baayen2018; @meulman2015; @meulman2023].
-The appeal of GAMs for modelling surface (sensor-level) M/EEG data comes from their ability to deal with the complex shape of the ERP without overfitting it.
-Generalising to scale and shape or "distributional GAMs" [@rigby2005; @umlauf2018], allowing for the specification of linear predictors for both the shape and scale parameters of many distributions [for an application to neuroimaging data, see @dinga2021]...
-
-Introduction to multilevel GAMs [@pedersen_hierarchical_2019]...
-Proper inclusion of varying/random effects in the model specification protects against overly wiggly curves [@baayen2020]...
-@fig-intro-gam\ignorespaces D...
-Instead of averaging, we obtain the smooth ERP signal from multilevel GAM...
-which has been shown to be less susceptible to outliers [@meulman2023]...
-
--->
 
@@
 ## Objectives
-
-<!--
-
-Recently, @teichmann2022 provided a detailed tutorial on using Bayes factors (BFs) to analyse the 1D or 2D output from MVPA, that is, for testing, at every timestep, whether decoding performance is above chance level.
-However, this approach provides timeseries of BFs that ignores temporal dependencies...
-In the following...
-
-As put by @rousselet_using_2025, concluding on the onset of effect based on a series of univariate tests...
-commits to three fallacies...
-here, we want to avoid these by introducing a model-based approach, which naturally take into account the temporal dependencies in the data to output a series of posterior probabilities...
-
-Given the previously reported limitations of cluster-based inference to precisely identify the onset and offset of M/EEG effects (e.g., ERPs, decoding performance), we developed a model-based approach for estimating the onset and offset of M/EEG effects.
-To achieve this, we leveraged Bayesian generalised additive multilevel models fitted in `R` via the `brms` package and compared the performance of this approach to conventional methods on both simulated and actual M/EEG data.
-
--->
 
@@
 We ran eight Markov Chain Monte-Carlo (MCMC) to approximate the posterior distribution, including each 5000 iterations and a warmup of 2000 iterations, yielding a total of $8 \times (5000-2000) = 24000$ posterior samples to use for inference.
-Posterior convergence was assessed examining trace plots as well as the Gelman–Rubin statistic $\hat{R}$ [@gabry2019; @gelman2020].
+Posterior convergence was assessed examining trace plots as well as the Gelman–Rubin statistic $\hat{R}$ [@gabry2019; @gelman2020; @vehtari2021].
 The `brms` package uses the same syntax as `r pkrt("mgcv")` for specifying smooth effects.
@@
 
-However, the previous model only included constant (fixed) effects, thus not properly accounting for between-participant variability.
+However, the model whose predictions are depicted in @fig-plot-post-slope only included constant (fixed) effects, thus not properly accounting for between-participant variability.
 We next fitted a multilevel version of the BGAM [BGAMM, for an introduction to Bayesian multilevel models in `brms`, see @nalborczyk2019] including a varying intercept and slope for participant (but with a constant smoother).
@@
 
-```{r meta-gamm, echo = FALSE, eval = FALSE}
-# fitting the BGAMM
-# gamm <- brm(
-#     # thin-plate regression splines with k-1 basis functions
-#     eeg_mean ~ 1 + condition + (1 + condition | participant) + s(time, bs = "tp", k = 20, by = condition),
-#     data = ppt_df,
-#     family = gaussian(),
-#     warmup = 2000,
-#     iter = 5000,
-#     chains = 8,
-#     cores = 8,
-#     file = "models/gamm.rds"
-#     )
-
-# fitting the BGAMM
-# meta_gam_k10 <- brm(
-#     # using the by-participant SD of ERPs
-#     eeg_mean | se(eeg_sd) ~
-#         1 + condition + (1 + condition | participant) +
-#         s(time, bs = "tp", k = 10, by = condition),
-#     data = summary_df,
-#     family = gaussian(),
-#     warmup = 2000,
-#     iter = 5000,
-#     chains = 8,
-#     cores = 8,
-#     file = "models/meta_gam_k10.rds"
-#     )
-
-# comparing the two models
-# loo(meta_gam, meta_gam_k10)
-# library(easystats)
-# model_comp <- compare_performance(meta_gam, meta_gam_k10)
-# model_comp
-# plot(model_comp)
-```
-
 We depict the posterior predictions together with the posterior estimate of the slope for `condition` at each timestep (@fig-plot-post-slope).
@@
 [^1]: These posterior odds are equivalent to a Bayes factor, assuming 1:1 prior odds.
-
-```{r tests2, echo = FALSE, eval = FALSE}
-# compute percentage in ROPE
-library(easystats)
-# rope(
-#     x = gam,
-#     range = "default",
-#     ci = 0.95,
-#     ci_method = "ETI",
-#     effects = "fixed",
-#     # component = c("conditional", "zi", "zero_inflated", "all"),
-#     # parameters = NULL,
-#     verbose = TRUE
-#     )
-# rope_range(x = gam)
-
-# retrieving posterior draws for EEG over time
-# gam <- full_model
-# post_samples <- posterior_epred(object = gam, newdata = ppt_df, allow_new_levels = TRUE)
-gam <- meta_gam
-post_samples <- posterior_epred(object = gam, newdata = summary_df, allow_new_levels = TRUE)
-
-# defining the ROPE as baseline SD
-# rope_range <- c(-0.1, 0.1)
-baseline_sd <- summary_df %>%
-    pivot_wider(names_from = condition, values_from = eeg_mean) %>%
-    mutate(cond_diff = abs(cond1 - cond2) ) %>%
-    filter(time < 0.1) %>%
-    # head()
-    summarise(baseline_sd = sd(cond_diff) ) %>%
-    data.frame()
-
-rope_range <- c(-baseline_sd, +baseline_sd)
-
-# computing the HDI for each time step
-hdi_values <- apply(post_samples, 2, function (x) hdi(x, ci = 0.95) )
-
-# computing the  proportion of the HDI that falls within the ROPE
-prop_in_rope <- sapply(1:ncol(post_samples), function (i) {
-    lower_hdi <- hdi_values[[i]][1]
-    upper_hdi <- hdi_values[[i]][2]
-    # computing the  proportion inside ROPE
-    mean(lower_hdi >= rope_range[1] & upper_hdi <= rope_range[2])
-    })
-
-# storing the results in a dataframe
-# rope_results <- data.frame(time = ppt_df$time, prop_in_rope = prop_in_rope)
-rope_results <- data.frame(time = summary_df$time, prop_in_rope = prop_in_rope)
-
-# plotting the results
-rope_results %>%
-    summarise(prop_in_rope = mean(prop_in_rope), .by = time) %>%
-    ggplot(aes(x = time, y = prop_in_rope) ) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = true_offset, linetype = 2) +
-    geom_point()
-
-# loading or fitting the full model
-# full_model <- gam
-# full_model <- readRDS("models/gam.rds")
-full_model <- meta_gam
-
-# fitting the null model (no 'condition' effect)
-# null_model <- brm(
-#     eeg ~ s(time, bs = "tp", k = 20),
-#     data = ppt_df,
-#     family = gaussian(),
-#     warmup = 2000,
-#     iter = 5000,
-#     chains = 8,
-#     cores = 8,
-#     file = "models/null_gam.rds"
-#     )
-
-# fitting the BGAMM
-null_model <- brm(
-    # using by-participant SD of ERPs across trials
-    eeg_mean | se(eeg_sd) ~
-        1 + (1 | participant) +
-        s(time, bs = "tp", k = 20),
-    data = summary_df,
-    family = gaussian(),
-    warmup = 2000,
-    iter = 5000,
-    chains = 8,
-    cores = 8
-    # file = "models/meta_gamm.rds"
-    )
-
-# computing the marginal likelihoods using bridge sampling
-# library(bridgesampling)
-# full_bridge <- bridge_sampler(full_model)
-# null_bridge <- bridge_sampler(null_model)
-
-# computing the Bayes Factor in favor of full model
-# BF_10 <- bf(full_bridge, null_bridge)
-# print(BF_10)
-
-# extracting log-likelihoods for each model
-log_lik_full <- log_lik(full_model)
-log_lik_null <- log_lik(null_model)
-
-# computing the pointwise log pointwise predictive density (LPD)
-# averaging the log-likelihood across posterior draws
-lpd_full <- colMeans(log_lik_full)
-lpd_null <- colMeans(log_lik_null)
-
-# computing the LPD difference (full model - null model)
-lpd_diff <- lpd_full - lpd_null
-
-# computing pointwise Bayes factors as exp(lpd_diff)
-BF_10_pointwise <- exp(lpd_diff)
-
-# storing the results in a dataframe
-ppt_df <- full_model$data
-lpd_results <- data.frame(time = ppt_df$time, LPD_diff = lpd_diff, BF_10 = BF_10_pointwise)
-
-# viewing the results
-head(lpd_results)
-
-# plotting it
-lpd_results %>%
-    # pivot_longer(cols = LPD_diff:BF_10) %>%
-    summarise(BF_10 = mean(LPD_diff), .by = time) %>%
-    # summarise(BF_10 = mean(BF_10), .by = time) %>%
-    # ggplot(aes(x = time, y = value, colour = name) ) +
-    ggplot(aes(x = time, y = BF_10) ) +
-    geom_vline(xintercept = true_onset, linetype = 2) +
-    geom_vline(xintercept = true_offset, linetype = 2) +
-    geom_line() +
-    geom_point()
-```
 
@@
 
-```{r fig-bgamm-varying-effects, eval = FALSE, echo = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Posterior estimate of the EEG activity in each condition (left) and posterior estimate of the difference in EEG activity (right) according to the BGAMM."}
-####################################################################
-# TO-DO: plotting the onset/offset estimate per participant?
-# see https://discourse.mc-stan.org/t/specifying-varying-effect-structures-with-hierarchical-generalised-additive-models-gams/35596
-#################################################################
-
-# fixed-effect GAM
-gam_fixed <- gam(
-    y ~ s(x),
-    data = df
-    )
-
-# random intercept + slope
-gam_int_slope <- gam(
-    y ~ s(x) +
-        s(subject, bs = "re") +
-        s(x, subject, bs = "re"),
-    data = df
-    )
-
-# random smooth per subject + global smooth
-gam_rand_smooth <- gam(
-    y ~ s(x) +
-        s(subject, bs = "re") +
-        s(x, subject, bs = "fs", m = 1),
-    data = df
-    )
-
-# fitting the BGAMM
-# see https://discourse.mc-stan.org/t/specifying-varying-effect-structures-with-hierarchical-generalised-additive-models-gams/35596/2
-# tic()
-# 5519.743 sec elapsed
-full_meta_gamm <- brm(
-    # using by-participant SD of ERPs across trials
-    eeg_mean | se(eeg_sd) ~ 1 + condition +
-        s(time, by = condition, bs = "tp", k = 20) +
-        s(time, participant, by = condition, bs = "fs", xt = list(bs = "tp") ),
-    data = summary_df,
-    family = gaussian(),
-    warmup = 2000,
-    iter = 5000,
-    chains = 8,
-    cores = 8,
-    control = list(adapt_delta = 0.95)
-    # file = "models/full_meta_gamm.rds"
-    )
-# toc()
-
-# model <- meta_gamm
-# plot_post_preds(model = meta_gamm) + plot_post_slope(model = meta_gamm)
-# model <- full_meta_gamm
-# plot_post_preds(model = full_meta_gamm) + plot_post_slope(model = full_meta_gamm)
-# model <- meta_gam
-# model <- gamm
-
-plot_post_test_ratio_individual <- function (model, sesoi = 0, threshold = 10) {
-
-    # retrieving the number of posterior samples
-    n_posterior_samples <- ndraws(model)
-
-    # retrieving posterior draws and compute the difference per participant & time
-    prob_y_above_0_indiv <- model$data %>%
-        add_epred_draws(object = model) %>%
-        # add_epred_draws(object = model, ndraws = 1e2) %>%
-        data.frame() %>%
-        dplyr::select(participant, time, condition, .epred, .draw) %>%
-        pivot_wider(names_from = condition, values_from = .epred) %>%
-        # pivot_wider(id_cols = c(participant, time), names_from = condition, values_from = .epred) %>%
-        mutate(epred_diff = cond2 - cond1) %>%
-        group_by(participant, time) %>%
-        # summarise(m = mean(epred_diff > 0 + sesoi), .groups = "drop") %>%
-        summarise(m = mean(epred_diff > 0 + sesoi) ) %>%
-        mutate(
-            prob_ratio = m / (1 - m),
-            prob_ratio = ifelse(is.infinite(prob_ratio), n_posterior_samples, prob_ratio),
-            prob_ratio = ifelse(prob_ratio == 0, 1 / n_posterior_samples, prob_ratio),
-            above_thres = ifelse(prob_ratio >= threshold, 1, NA)
-            ) %>%
-        ungroup()
-
-    # some tests
-    # model$data %>%
-    #     add_epred_draws(object = model, ndraws = 1e2) %>%
-    #     data.frame() %>%
-    #     dplyr::select(participant, time, condition, .epred, .draw) %>%
-    #     filter(participant == "participant_01" & time == 0 & .draw == 1)
-    # 
-    # prob_y_above_0_indiv %>%
-    #     filter(participant == "participant_02" & time == 0 & .draw == 2)
-    # 
-    # prob_y_above_0_indiv %>%
-    #     pivot_wider(names_from = condition, values_from = .epred) %>%
-    #     filter(participant == "participant_02" & time == 0 & .draw == 2)
-
-    # plotting posterior probabilities per participant
-    post_plot_indiv <- prob_y_above_0_indiv %>%
-        ggplot(aes(x = time, y = m, group = participant, color = participant) ) +
-        geom_hline(yintercept = 0.5, linetype = 2) +
-        geom_vline(xintercept = true_onset, linetype = 2) +
-        geom_vline(xintercept = true_offset, linetype = 2) +
-        geom_line(alpha = 0.5) +
-        labs(
-            x = "Time (s)",
-            # y = expression(Pr(beta>0~"|"~data, model) ),
-            y = expression(p(beta>0~"|"~data) ),
-            title = "Posterior probabilities per participant"
-            ) +
-        ylim(c(0, 1) ) +
-        theme(legend.position = "none")
-
-    # plotting posterior probability ratio per participant (log scale)
-    ratio_plot_indiv <- prob_y_above_0_indiv %>%
-        ggplot(aes(x = time, y = prob_ratio, group = participant, color = participant) ) +
-        geom_hline(yintercept = 1, linetype = "dashed") +
-        geom_hline(yintercept = 10, linetype = "dashed", color = "darkred") +
-        geom_hline(yintercept = 1 / 10, linetype = "dashed", color = "darkred") +
-        geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
-        geom_hline(yintercept = 1 / 100, linetype = "dashed", color = "orangered") +
-        geom_vline(xintercept = true_onset, linetype = 2) +
-        geom_vline(xintercept = true_offset, linetype = 2) +
-        geom_line(alpha = 0.5) +
-        geom_point(
-            data = . %>% dplyr::filter(!is.na(above_thres) ),
-            aes(y = threshold),
-            colour = "darkgreen",
-            shape = 16,
-            size = 1.5,
-            na.rm = TRUE
-            ) +
-        scale_y_log10(
-            labels = label_log(digits = 2),
-            limits = c(1 / n_posterior_samples, n_posterior_samples),
-            breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
-            ) +
-        labs(
-            x = "Time (s)",
-            y = expression(p(beta>0~"|"~data) / (1 - p(beta>0~"|"~data) ) ),
-            title = "Posterior probability ratios per participant"
-            ) +
-        theme(legend.position = "none")
-
-    # combining plots
-    post_plot_indiv + ratio_plot_indiv
-
-}
-
-# plotting the predictions
-# plot_post_test_ratio_individual(model = full_meta_gamm)
-# plot_post_test_ratio_individual(model = meta_gam)
-plot_post_test_ratio_individual(model = gamm)
-
-# plotting the predictions at the participant level
-plot_post_preds_ppt <- function (model, data = NULL, true_onset = NULL, true_offset = NULL) {
-
-    if (is.null(data) ) data <- model$data
-
-    # retrieving predictions at the individual level
-    # cond_effects <- conditional_effects(
-    #     # defining the model
-    #     x = model,
-    #     # defining the interaction of interest
-    #     effects = "time:condition",
-    #     # defining the "conditions" (participant) on which computing predictions
-    #     conditions = make_conditions(x = model$data, vars = "participant"),
-    #     # getting predictions
-    #     method = "posterior_epred",
-    #     # including all random/varying effects
-    #     re_formula = NULL
-    #     )[[1]]
-
-    # retrieving posterior draws and computing the difference per participant & time
-    post_preds <- model$data %>%
-        add_epred_draws(object = model, re_formula = NULL, ndraws = 500) %>%
-        data.frame() %>%
-        dplyr::select(participant, time, condition, .epred, .draw) %>%
-        group_by(participant, time, condition) %>%
-        summarise(
-            .epred_mean = mean(.epred),
-            .epred_lower = quantile(.epred, probs = 0.025),
-            .epred_upper = quantile(.epred, probs = 0.975),
-            .groups = "drop"
-            ) %>%
-        ungroup()
-
-    # retrieving observed means (optional)
-    obs_means <- data %>%
-        group_by(participant, time, condition) %>%
-        summarise(eeg_mean = mean(eeg_mean), .groups = "drop")
-
-    # plotting
-    post_preds %>%
-        ggplot(aes(x = time, y = .epred_mean, colour = condition, fill = condition) ) +
-        geom_vline(xintercept = true_onset, linetype = 2) +
-        geom_vline(xintercept = true_offset, linetype = 2) +
-        geom_hline(yintercept = 0, linetype = 2) +
-        geom_point(
-            data = obs_means, aes(y = eeg_mean),
-            pch = 21, size = 1, alpha = 0.2,
-            show.legend = FALSE
-            ) +
-        geom_line() +
-        geom_ribbon(aes(ymin = .epred_lower, ymax = .epred_upper, colour = NULL), alpha = 0.2) +
-        facet_wrap(~participant, scales = "free_y") +
-        scale_colour_met_d(name = "Johnson") +
-        scale_fill_met_d(name = "Johnson") +
-        labs(x = "Time (s)", y = "EEG signal (a.u.)", title = "Posterior predictions by participant")
-
-}
-
-plot_post_preds_ppt(model = meta_gam)
-```
-
-<!--
-
-\newpage
-
-### Error properties of the proposed approach
-
-We then assessed the performance of the proposed approach by computing the difference between the true and estimated onset/offset of the EEG difference according to various `k` (basis dimension) and `threshold` values.
-Remember that the EEG signal was generated from a truncated Gaussian with an objective onset at 160 ms, a maximum at 250 ms, and an offset at 342 ms.
-@fig-onset-error shows that the multilevel GAM can almost exactly recover the true onset and offset values, given some reasonable choice of `k` and `threshold` values.
-We provide more detailed recommendations on how to set `k` in @apx-basis.
-This figure further reveals that the optimal `k` and `threshold` values may differ for the onset and offset values, and there seems to exist a trade-off between these two parameters: lower `k` values lead to poorer estimations, but these poor estimations can be compensated (only to some extent) by higher `threshold` values (and reciprocally).
-
-```{r fig-onset-error, eval = TRUE, echo = FALSE, results = "hide", dev = "png", dpi = 300, fig.width = 9, fig.asp = 0.5, fig.cap = "Average estimation error (RMSE) for the onset (left) and offset (right) according to various basis dimension and threshold values for the BGAM (computed from 100 simulated datasets)."}
-# running the simulation on the HPC cluster
-# source("code/meta_gam_errors_cluster.R")
-
-# loading the simulation results
-results <- readRDS(file = "results/meta_gam_error_properties_cluster.rds")
-
-# defining the true onset and offset in seconds
-true_onset <- 0.160
-true_peak <- 0.250
-true_offset <- 0.342
-
-# plotting the results
-# results %>%
-#     group_by(kvalue, threshold, sim_id) %>%
-#     summarise(
-#         # value = if (onset_offset[1] == "onset") min(value) else max(value),
-#         cluster_onset = min(cluster_onset),
-#         cluster_offset = max(cluster_offset)
-#         # .groups = "drop"
-#         ) %>%
-#     ungroup() %>%
-#     group_by(kvalue, threshold) %>%
-#     # summarise(across(estimated_onset:error_offset, median) ) %>%
-#     # summarise(across(estimated_onset:error_offset, mean) ) %>%
-#     summarise(
-#         # error_onset = mean(abs(cluster_onset - true_onset) ),
-#         # error_offset = mean(abs(cluster_offset - true_offset) )
-#         error_onset = sqrt(mean(cluster_onset - true_onset)^2),
-#         error_offset = sqrt(mean(cluster_offset - true_offset)^2)
-#         ) %>%
-#     ungroup() %>%
-#     pivot_longer(cols = error_onset:error_offset) %>%
-#     mutate(
-#         name = factor(
-#             x = name,
-#             levels = c("error_onset", "error_offset"),
-#             labels = c("onset", "offset")
-#             )
-#         ) %>%
-#     ggplot(
-#         aes(
-#             x = as.factor(kvalue),
-#             y = as.factor(threshold),
-#             fill = value
-#             )
-#         ) +
-#     geom_raster(interpolate = FALSE) +
-#     # geom_tile() +
-#     # geom_point(
-#     #     data = results %>%
-#     #         group_by(kvalue, threshold) %>%
-#     #         summarise(across(estimated_onset:error_offset, mean) ) %>%
-#     #         ungroup() %>%
-#     #         na.omit() %>%
-#     #         filter(error_onset == min(error_onset) ),
-#     #     # aes(x = kvalue, y = threshold),
-#     #     aes(x = kvalue, y = value),
-#     #     color = "orangered", size = 4, shape = 4,
-#     #     # show.legend = FALSE
-#     #     show.legend = TRUE
-#     #     ) +
-#     facet_wrap(~name) +
-#     scale_x_discrete(expand = c(0, 0) ) +
-#     scale_y_discrete(expand = c(0, 0), breaks = c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50) ) +
-#     scale_fill_gradientn(
-#         colors = rev(met.brewer("Hokusai1") ),
-#         transform = "sqrt"
-#         ) +
-#     labs(
-#         x = "GAM basis dimension (k)",
-#         y = "Threshold",
-#         fill = "RMSE"
-#         )
-
-results %>%
-    # filter(kvalue > 5) %>%
-    group_by(kvalue, threshold, sim_id) %>%
-    summarise(
-        cluster_onset = min(cluster_onset),
-        cluster_offset = max(cluster_offset),
-        .groups = "drop"
-        ) %>%
-    ungroup() %>%
-    group_by(kvalue, threshold) %>%
-    summarise(
-        # error_onset = median(cluster_onset - true_onset),
-        # error_offset = median(cluster_offset - true_offset)
-        # error_onset = median(abs(cluster_onset - true_onset) ),
-        # error_offset = median(abs(cluster_offset - true_offset) )
-        error_onset = sqrt(mean(cluster_onset - true_onset)^2),
-        error_offset = sqrt(mean(cluster_offset - true_offset)^2)
-        ) %>%
-    ungroup() %>%
-    pivot_longer(cols = error_onset:error_offset) %>%
-    mutate(
-        name = factor(
-            x = name,
-            levels = c("error_onset", "error_offset"),
-            labels = c("onset", "offset")
-            )
-        ) %>%
-    ggplot(
-        aes(
-            x = threshold,
-            y = value,
-            colour = as.factor(kvalue),
-            fill = as.factor(kvalue),
-            group = as.factor(kvalue),
-            )
-        ) +
-    geom_line(linewidth = 0.5) +
-    geom_point(size = 0.8) +
-    # geom_segment(
-    #     data = . %>% filter(value == min(value), .by = name),
-    #     aes(
-    #         x = threshold,
-    #         xend = threshold,
-    #         y = 0.01,
-    #         yend = 0,
-    #         colour = as.factor(kvalue)
-    #         ),
-    #     arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
-    #     inherit.aes = FALSE,
-    #     linewidth = 1.0,
-    #     show.legend = FALSE
-    #     ) +
-    # geom_smooth(alpha = 0.1, linewidth = 0.8) +
-    facet_wrap(~name) +
-    scale_fill_viridis_d() +
-    scale_colour_viridis_d() +
-    scale_y_continuous(
-        labels = scales::label_number(),
-        transform = "sqrt"
-        ) +
-    labs(
-        x = "BGAM threshold",
-        y = "Estimation error (RMSE)",
-        colour = "BGAM basis\ndimension (k)",
-        fill = "BGAM basis\ndimension (k)"
-        )
-```
-
-\newpage
-
--->
-
 ### Comparing the onset/offset estimates across approaches
@@
 The cluster-based inference was implemented using a cluster-sum statistic of squared $t$-values, as implemented in `MNE-Python` [@gramfort2013], called via `r pkrt("reticulate")`.
-We also compared these estimates to the onset and offset as estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared $t$-values [as in @rousselet_using_2025].^[As in @rousselet_using_2025, we fixed the number of expected change points to two in the binary segmentation algorithm, thus producing always one cluster.] @fig-corrections illustrates the onsets and offsets estimated by each method on a single simulated dataset and shows that all methods systematically overestimate the true onset and underestimate the true offset.
+We also compared these estimates to the onset and offset as estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared $t$-values [as in @rousselet_using_2025].[^2] @fig-corrections illustrates the onsets and offsets estimated by each method on a single simulated dataset and shows that all methods systematically overestimate the true onset and underestimate the true offset.
 In addition, the `Raw p-value`, `FDR BH95`, and `FDR BY01` methods identify clusters well before the true onset and after the true offset.
+
+[^2]: As in @rousselet_using_2025, we fixed the number of expected change points to two in the binary segmentation algorithm, thus producing always one cluster.
 
@@
 # estimating onset and offset using the binary segmentation method
-res <- cpt.meanvar(data = tests_results$tval, method = "BinSeg", Q = 2, penalty = "MBIC", minseglen = 2)
+res <- cpt.meanvar(
+    data = tests_results$tval, method = "BinSeg",
+    Q = 2, penalty = "MBIC", minseglen = 2
+    )
 onset_cpt <- tests_results$time[res@cpts[1]]
 offset_cpt <- tests_results$time[res@cpts[2]]
-
-# sanity checks
-# ts.plot(tests_results$tval)
-# for (i in 1:length(res@cpts) ) {
-#     
-#     abline(v = res@cpts[i])
-#     
-# }
 
@@
     geom_segment(
-        # data = all_methods_onset_offset,
         data = clusters,
@@
 
-<!--
-
-## Results
-
-This section is divided in two parts.
-First, we present the results from the simulation study, assessing the bias and variance of each method when applied to simulated data in which the ground truth is known.
-Second, we present the results obtained when applying the different methods to actual MEG data (decoding performance through time), assessing the reliability of the estimates provided by each method.
-
-## Simulation study (bias and variance)
-
--->
-
 ## Results
@@
     )
-
-# sanity check
-# n_distinct(sim_results$simulation_id)
-
-# identifying the best brms results
-# sim_results %>%
-#     filter(method == "brms") %>%
-#     pivot_wider(names_from = onset_offset, values_from = value) %>%
-#     summarise(
-#         bias_onset = mean(onset - true_onset),
-#         MAE_onset = mean(abs(onset - true_onset) ),
-#         variance_onset = var(onset),
-#         bias_offset = mean(offset - true_offset),
-#         MAE_offset = mean(abs(offset - true_offset) ),
-#         variance_offset = var(offset),
-#         .by = threshold
-#         ) %>%
-#     arrange(MAE_onset, MAE_offset)
-#     # arrange(bias_onset, bias_offset)
-#     # arrange(variance_onset, variance_offset)
 
@@
         error = case_when(
-            # grepl("onset", onset_offset) ~ abs(value - true_onset),
-            # grepl("offset", onset_offset) ~ abs(value - true_offset)
             grepl("onset", onset_offset) ~ value - true_onset,
@@
             x = mean_error,
-            # y = mad,
-            # y = median_abs_error,
-            # y = median_abs_error,
-            # x = median_abs_error,
             y = variance,
-            # y = mean_abs_error,
             colour = method,
@@
     geom_vline(xintercept = 0, linetype = 2) +
-    # geom_errorbar(
-    #     aes(
-    #         xmin = median_error - mad, xmax = median_error + mad,
-    #         #ymin = median_abs_error - mad, ymax = median_abs_error + mad
-    #         )
-    #     ) +
-    #     geom_errorbar(
-    #     aes(
-    #         #xmin = median_error - mad, xmax = median_error + mad,
-    #         ymin = median_abs_error - mad, ymax = median_abs_error + mad
-    #         )
-    #     ) +
     geom_point(
@@
         ) +
-    # scale_x_log10(guide = "axis_logticks") +
     scale_y_log10(
         guide = "axis_logticks",
-        # breaks = trans_breaks("log10", function(x) 10^x),
-        # labels = trans_format("log10", math_format(10^.x) )
         labels = label_log(digits = 2)
@@
     labs(x = "Mean error (s)", y = "Variance (log-scale)")
-
-# or plotting only the error in descending order
-# sim_results %>%
-#     # keeping only threshold = 10 for BGAM and threshold = 0.05 for frequentist results
-#     filter(threshold %in% c(0.05, 20) ) %>%
-#     # filter(threshold %in% c(0.05, 10) ) %>%
-#     mutate(
-#         error = case_when(
-#             # grepl("onset", onset_offset) ~ abs(value - true_onset),
-#             # grepl("offset", onset_offset) ~ abs(value - true_offset)
-#             grepl("onset", onset_offset) ~ value - true_onset,
-#             grepl("offset", onset_offset) ~ value - true_offset
-#             )
-#         ) %>%
-#     # converting to ms
-#     # mutate(error = error * 1e3) %>%
-#     group_by(method, onset_offset) %>%
-#     summarise(
-#         mean_error = mean(error, na.rm = TRUE),
-#         median_error = median(error, na.rm = TRUE),
-#         median_abs_error = median(abs(error), na.rm = TRUE),
-#         variance = var(error, na.rm = TRUE),
-#         # variance = var(value, na.rm = TRUE),
-#         sem = sd(value, na.rm = TRUE) / sqrt(n() ),
-#         mad = mad(value, na.rm = TRUE),
-#         .groups = "drop"
-#         ) %>%
-#     ungroup() %>%
-#     mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
-#     mutate(
-#         method = factor(
-#             x = method,
-#             levels = c(
-#                 "raw_p", "pval_bh", "pval_by", "pval_holm",
-#                 "cluster_mass", "cluster_tfce", "cpt", "brms"
-#                 ),
-#             labels = c(
-#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-#                 "TFCE", "Change point", "BGAM (our method)"
-#                 ),
-#             )
-#         ) %>%
-#     mutate(
-#         method_reordered = tidytext::reorder_within(
-#             x = method,
-#             # by = 1 / error,
-#             # fun = median,
-#             # by = 1 / abs(median_error),
-#             by = 1 / median_abs_error,
-#             # by = 1 / abs(mean_error),
-#             within = onset_offset
-#             )
-#         ) %>%
-#     # mutate(
-#     #     method_reordered = fct_reorder(
-#     #         method, error, .fun = ~abs(median(.x) ), .na_rm = TRUE
-#     #         )
-#     #     ) %>%
-#     # arrange(desc(error) ) %>%
-#     # arrange(error) %>%
-#     # head()
-#     ggplot(
-#         aes(
-#             x = median_error,
-#             # x = error,
-#             xmin = median_error - mad,
-#             xmax = median_error + mad,
-#             # x = mean_error,
-#             # xmin = mean_error - variance,
-#             # xmax = mean_error + variance,
-#             y = method_reordered,
-#             colour = method,
-#             fill = method
-#             )
-#         ) +
-#     geom_vline(xintercept = 0, linetype = 2) +
-#     geom_pointrange(
-#         size = 0.25,
-#         linewidth = 0.75,
-#         show.legend = FALSE
-#         ) +
-#     # stat_histinterval(
-#     #     # breaks = "FD",
-#     #     # n = 1e3,
-#     #     # point_interval = "median_qi",
-#     #     # .width = c(0.5, 0.95),
-#     #     .width = 0.1,
-#     #     # colour = "black",
-#     #     # slab_colour = "white",
-#     #     # slab_linewidth = 0,
-#     #     slab_alpha = 0.8,
-#     #     # fill = "steelblue"
-#     #     show.legend = FALSE
-#     #     ) +
-#     facet_wrap(~onset_offset, scales = "free") +
-#     scale_y_reordered() +
-#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-#     labs(
-#         x = "Estimation error (s)",
-#         # x = "Mean error (s)",
-#         y = "Method"
-#         )
-
-# saving the plot
-# ggsave(
-#     # filename = "figures/simulation_results_median_error_variance.png",
-#     # filename = "figures/simulation_results_bias_mae.png",
-#     filename = "figures/simulation_results_mae_variance.png",
-#     width = 12, height = 6, dpi = 300,
-#     device = "png"
-#     )
-
-# plotting the distribution of results
-# sim_results %>%
-#     # keeping only threshold = 10 for BGAM and threshold = 0.05 for frequentist results
-#     filter(threshold %in% c(0.05, 20) ) %>%
-#     mutate(
-#         error = case_when(
-#             # grepl("onset", onset_offset) ~ abs(value - true_onset),
-#             # grepl("offset", onset_offset) ~ abs(value - true_offset)
-#             grepl("onset", onset_offset) ~ value - true_onset,
-#             grepl("offset", onset_offset) ~ value - true_offset
-#             )
-#         ) %>%
-#     # converting to ms
-#     # mutate(error = error * 1e3) %>%
-#     # group_by(method, onset_offset) %>%
-#     # summarise(
-#     #     mean_error = mean(error, na.rm = TRUE),
-#     #     median_error = median(error, na.rm = TRUE),
-#     #     median_abs_error = median(abs(error), na.rm = TRUE),
-#     #     variance = var(error, na.rm = TRUE),
-#     #     # variance = var(value, na.rm = TRUE),
-#     #     sem = sd(value, na.rm = TRUE) / sqrt(n() ),
-#     #     mad = mad(value, na.rm = TRUE),
-#     #     .groups = "drop"
-#     #     ) %>%
-#     # ungroup() %>%
-#     mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
-#     mutate(
-#         method = factor(
-#             x = method,
-#             levels = c(
-#                 "raw_p", "pval_bh", "pval_by", "pval_holm",
-#                 "cluster_mass", "cluster_tfce", "cpt", "brms"
-#                 ),
-#             labels = c(
-#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-#                 "TFCE", "Change point", "BGAM (our method)"
-#                 ),
-#             )
-#         ) %>%
-#     mutate(
-#         method_reordered = tidytext::reorder_within(
-#             x = method,
-#             by = abs(error),
-#             fun = median,
-#             # by = 1 / abs(median_error),
-#             # by = 1 / abs(mean_error),
-#             within = onset_offset
-#             )
-#         ) %>%
-#     # mutate(
-#     #     method_reordered = fct_reorder(
-#     #         method, error, .fun = ~abs(median(.x) ), .na_rm = TRUE
-#     #         )
-#     #     ) %>%
-#     # arrange(desc(error) ) %>%
-#     # arrange(error) %>%
-#     # head()
-#     ggplot(
-#         aes(
-#             # x = median_error,
-#             x = error,
-#             # xmin = median_error - mad,
-#             # xmax = median_error + mad,
-#             # x = mean_error,
-#             # xmin = mean_error - variance,
-#             # xmax = mean_error + variance,
-#             y = method_reordered,
-#             colour = method,
-#             fill = method
-#             )
-#         ) +
-#     geom_vline(xintercept = 0, linetype = 2) +
-#     # geom_pointrange(
-#     #     size = 0.25,
-#     #     linewidth = 0.75,
-#     #     show.legend = FALSE
-#     #     ) +
-#     stat_slabinterval(
-#         n = 200,
-#         normalize = "groups",
-#         point_interval = "median_qi",
-#         # .width = c(0.1, 0.9),
-#         # .width = 0.2,
-#         .width = 0,
-#         interval_size_range = c(1, 1),
-#         fatten_point = 1.25,
-#         # slab_colour = "white",
-#         # slab_linewidth = 0,
-#         # density = "unbounded",
-#         slab_alpha = 0.5,
-#         show.legend = FALSE
-#         ) +
-#     facet_wrap(~onset_offset, scales = "free") +
-#     scale_y_reordered() +
-#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-#     labs(
-#         x = "Estimation error (s)",
-#         y = "Method"
-#         )
 ```
@@
     filter(threshold %in% c(0.05, 20) ) %>%
-    # filter(threshold %in% c(0.05, 10) ) %>%
     mutate(
         error = case_when(
-            # grepl("onset", onset_offset) ~ value - true_onset,
-            # grepl("offset", onset_offset) ~ value - true_offset
+            # coputing error and converting to ms
             grepl("onset", onset_offset) ~ value * 1e3 - true_onset * 1e3,
@@
         ) %>%
-    # converting to ms
-    # mutate(error = error * 1e3) %>%
     group_by(method, onset_offset) %>%
@@
         mean_error = mean(error, na.rm = TRUE),
-        # median_error = median(error, na.rm = TRUE),
         mean_abs_error = mean(abs(error), na.rm = TRUE),
-        # median_abs_error = median(abs(error), na.rm = TRUE),
         rmse = sqrt(mean(error)^2),
         variance = var(error, na.rm = TRUE),
-        # mad = mad(error, na.rm = TRUE),
         .groups = "drop"
@@
 
+# compute the best values per onset_offset and metric
+best_vals <- sim_results_summary %>%
+    group_by(onset_offset) %>%
+    summarise(
+        best_bias = mean_error[which.min(abs(mean_error) )],
+        best_mae = min(mean_abs_error),
+        best_rmse = min(rmse),
+        best_variance = min(variance),
+        .groups = "drop"
+        )
+
+# join back so we know which rows are "winners"
+sim_results_summary <- sim_results_summary %>%
+    left_join(best_vals, by = "onset_offset") %>%
+        mutate(
+        bold_bias = mean_error == best_bias,
+        bold_mae = mean_abs_error == best_mae,
+        bold_rmse = rmse == best_rmse,
+        bold_variance = variance == best_variance
+        )
+
 # creating the table
@@
         ) %>%
-    # arrange(abs(median_error) ) %>%
-    # arrange(abs(mean_error) ) %>%
-    # arrange(median_abs_error) %>%
     arrange(mean_abs_error) %>%
@@
         method = "",
-        # mean_error = "Bias_mean",
-        # median_error = "Bias",
         mean_error = "Bias",
-        # median_abs_error = "MAE",
         mean_abs_error = "MAE",
@@
         variance = "Variance"
-        # mad = "MAD"
         ) %>%
@@
         table.align = "center",
-        # table.width = "75%",
         table.font.size = 12
         ) %>%
+    # apply bold formatting
+    tab_style(
+        style = cell_text(weight = "bold"),
+        locations = cells_body(
+            columns = mean_error,
+            rows = bold_bias
+            )
+        ) %>%
+    tab_style(
+        style = cell_text(weight = "bold"),
+        locations = cells_body(
+            columns = mean_abs_error,
+            rows = bold_mae
+            )
+        ) %>%
+    tab_style(
+        style = cell_text(weight = "bold"),
+        locations = cells_body(
+            columns = rmse,
+            rows = bold_rmse
+            )
+        ) %>%
+    tab_style(
+        style = cell_text(weight = "bold"),
+        locations = cells_body(
+            columns = variance,
+            rows = bold_variance
+            )
+        ) %>%
+    # removing the extra columns
+    cols_hide(columns = starts_with(c("best", "bold") ) ) %>%
+    # converting to latex table
     as_latex()
-
 ```
@@
 Although we chose a basis dimension of $k=50$, which seemed appropriate for the present data, this choice should be adapted according to the properties of the modelled data (e.g., signal-to-noise ratio, prior low-pass filtering, sampling rate) and should be assessed by the usual model checking tools (e.g., models comparison, posterior predictive checks, see @apx-basis).
-
-```{r fig-decoding-data, cache = FALSE, echo = FALSE, eval = FALSE, out.width = "75%", fig.asp = 0.75, fig.cap = "Group-level average decoding performance (N=32) superimposed with the BGAM predictions (in blue) and the region of practical equivalence (ROPE, in orange) computed from the baseline period [data from @nalborczyk:inprep].
-The blue horizontal markers indicate the timesteps at which the posterior probability ratio exceeds 20."}
-# importing the reshaped MEG decoding results
-decoding_df <- read.csv(file = "data/decoding_results_reshaped.csv") %>%
-    # removing some participants
-    mutate(participant_id = cur_group_id(), .by = participant) %>%
-    dplyr::filter(participant_id < 33) %>%
-    select(-participant_id)
-
-# averaging across participants
-decoding_summary_df <- decoding_df %>%
-    summarise(
-        auc_mean = mean(auc),
-        auc_sd = sd(auc),
-        .by = c(participant, time)
-        )
-
-# defining the chance level
-chance_level <- 0.5
-
-# computing the smallest effect size of interest (SESOI)
-# as the upper 90% quantile of decoding performance during the baseline
-sesoi <- decoding_summary_df %>%
-    dplyr::filter(time < 0) %>%
-    summarise(auc_mean = mean(auc_mean), .by = time) %>%
-    summarise(sesoi = quantile(x = auc_mean, probs = 0.90) ) %>%
-    # summarise(sesoi = quantile(x = auc_mean, probs = 0.95) ) %>%
-    pull(sesoi) - chance_level
-
-# computing the smallest effect size of interest (SESOI)
-# as the SD of decoding performance during the baseline
-# sesoi <- decoding_summary_df %>%
-#     dplyr::filter(time < 0) %>%
-#     summarise(auc_mean = mean(auc_mean), .by = time) %>%
-#     summarise(sesoi = sd(auc_mean) ) %>%
-#     pull(sesoi)
-
-# computing the smallest effect size of interest (SESOI)
-# as the 95% confidence interval during the baseline
-# sesoi <- decoding_summary_df %>%
-#     filter(time < 0) %>%
-#     summarise(auc_mean = mean(auc_mean), .by = time) %>%
-#     summarise(sesoi = 1.96 * (sd(auc_mean) / sqrt(n() ) ) ) %>%
-#     pull(sesoi)
-
-# fitting the GAMs
-meg_decoding_gam_gaussian_k40 <- brm(
-    auc_mean ~ s(time, bs = "cr", k = 40),
-    data = decoding_summary_df,
-    family = gaussian(),
-    warmup = 1000,
-    iter = 5000,
-    chains = 4,
-    cores = 4
-    )
-
-meg_decoding_gam_gaussian_k50 <- brm(
-    auc_mean ~ s(time, bs = "cr", k = 50),
-    data = decoding_summary_df,
-    family = gaussian(),
-    # warmup = 1000,
-    # iter = 5000,
-    warmup = 2000,
-    iter = 10000,
-    chains = 4,
-    cores = 4
-    )
-
-meg_decoding_gam_gaussian_k60 <- brm(
-    auc_mean ~ s(time, bs = "cr", k = 60),
-    data = decoding_summary_df,
-    family = gaussian(),
-    warmup = 1000,
-    iter = 5000,
-    chains = 4,
-    cores = 4
-    )
-
-meg_decoding_gam_beta_k40 <- brm(
-    auc_mean ~ s(time, bs = "cr", k = 40),
-    data = decoding_summary_df,
-    family = Beta(),
-    warmup = 1000,
-    iter = 5000,
-    chains = 4,
-    cores = 4
-    )
-
-meg_decoding_gam_beta_k50 <- brm(
-    auc_mean ~ s(time, bs = "cr", k = 50),
-    data = decoding_summary_df,
-    family = Beta(),
-    warmup = 1000,
-    iter = 5000,
-    chains = 4,
-    cores = 4
-    )
-
-meg_decoding_gam_beta_k60 <- brm(
-    auc_mean ~ s(time, bs = "cr", k = 60),
-    data = decoding_summary_df,
-    family = Beta(),
-    warmup = 1000,
-    iter = 5000,
-    chains = 4,
-    cores = 4
-    )
-
-# PPCs
-pp_check(meg_decoding_gam_gaussian_k40)
-pp_check(meg_decoding_gam_gaussian_k50)
-pp_check(meg_decoding_gam_gaussian_k60)
-pp_check(meg_decoding_gam_beta_k40)
-pp_check(meg_decoding_gam_beta_k50)
-pp_check(meg_decoding_gam_beta_k60)
-
-# models comparison
-loo(
-    meg_decoding_gam_gaussian_k40, meg_decoding_gam_gaussian_k50,
-    meg_decoding_gam_gaussian_k60,
-    meg_decoding_gam_beta_k40, meg_decoding_gam_beta_k50,
-    meg_decoding_gam_beta_k60
-    )
-
-meg_decoding_gam_gaussian_k40 <- add_criterion(x = meg_decoding_gam_gaussian_k40, criterio = "loo")
-meg_decoding_gam_gaussian_k50 <- add_criterion(x = meg_decoding_gam_gaussian_k50, criterio = "loo")
-meg_decoding_gam_gaussian_k60 <- add_criterion(x = meg_decoding_gam_gaussian_k60, criterio = "loo")
-meg_decoding_gam_beta_k40 <- add_criterion(x = meg_decoding_gam_beta_k30, criterio = "loo")
-meg_decoding_gam_beta_k50 <- add_criterion(x = meg_decoding_gam_beta_k50, criterio = "loo")
-meg_decoding_gam_beta_k60 <- add_criterion(x = meg_decoding_gam_beta_k60, criterio = "loo")
-
-models_comp <- loo_compare(
-    meg_decoding_gam_gaussian_k40, meg_decoding_gam_gaussian_k50,
-    meg_decoding_gam_gaussian_k60,
-    meg_decoding_gam_beta_k40, meg_decoding_gam_beta_k50,
-    meg_decoding_gam_beta_k60
-    )
-
-# summary
-print(x = models_comp, simplify = FALSE, digits = 3)
-
-# computing the posterior probability
-# prob_y_above_0 <- data.frame(time = unique(meg_decoding_gam$data$time) ) %>%
-prob_y_above_0 <- data.frame(time = unique(meg_decoding_gam_gaussian_k50$data$time) ) %>%
-    # retrieving the posterior samples
-    # add_epred_draws(object = meg_decoding_gam) %>%
-    add_epred_draws(object = meg_decoding_gam_gaussian_k50) %>%
-    # converting to dataframe
-    data.frame() %>%
-    # computing mean posterior probability at the group level
-    group_by(time) %>%
-    summarise(m = mean(.epred > (0 + chance_level + sesoi) ) ) %>%
-    # summarise(m = mean(.epred > (sesoi) ) ) %>%
-    mutate(prob_ratio = m / (1 - m) ) %>%
-    # ensuring there is no 0 or +Inf values
-    mutate(prob_ratio = pmin(prob_ratio, ndraws(meg_decoding_gam_gaussian_k50) ) ) %>%
-    mutate(prob_ratio = pmax(prob_ratio, 1 / ndraws(meg_decoding_gam_gaussian_k50) ) ) %>%
-    ungroup()
-
-# defining a function to find all onset and offset in timeseries
-find_clusters <- function (df, threshold) {
-
-    df %>%
-        mutate(above_threshold = prob_ratio >= threshold) %>%
-        mutate(cluster_change = c(TRUE, diff(above_threshold) != 0) ) %>%
-        dplyr::filter(above_threshold) %>%
-        mutate(cluster_id = cumsum(cluster_change) ) %>%
-        group_by(cluster_id) %>%
-        summarise(
-            cluster_onset = first(time), cluster_offset = last(time),
-            .groups = "drop"
-            ) %>%
-        select(cluster_onset, cluster_offset)
-
-}
-
-# defining a threshold
-post_prob_ratio_threshold <- 20
-clusters <- find_clusters(prob_y_above_0, threshold = post_prob_ratio_threshold)
-
-# defining a colour for the BGAM method
-brms_color <- met.brewer(name = "Johnson", n = 8)[8]
-
-# checking model's predictions against raw data
-plot(
-    # conditional_effects(x = meg_decoding_gam),
-    conditional_effects(x = meg_decoding_gam_gaussian_k50),
-    line_args = list(colour = brms_color, fill = brms_color, alpha = 0.3),
-    points = FALSE, plot = FALSE
-    )[[1]] +
-    geom_hline(yintercept = 0.5, linetype = 2) +
-    geom_vline(xintercept = 0.0, linetype = 2) +
-    annotate(
-        geom = "rect",
-        xmin = -Inf,
-        xmax = Inf,
-        ymin = chance_level - sesoi,
-        ymax = chance_level + sesoi,
-        # ymin = chance_level,
-        # ymax = sesoi,
-        fill ="orangered",
-        alpha = 0.2
-        ) +
-    geom_segment(
-        data = clusters,
-        aes(
-            x = cluster_onset,
-            xend = cluster_offset,
-            y = 0.48, yend = 0.48
-            ),
-        colour = brms_color,
-        inherit.aes = FALSE,
-        lineend = "round",
-        linewidth = 2
-        ) +
-    geom_line(
-        data = decoding_summary_df %>% summarise(auc_mean = mean(auc_mean), .by = time),
-        aes(x = time, y = auc_mean), inherit.aes = FALSE
-        ) +
-    labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)")
-```
-
-```{r fig-post-prob-ratio-decoding, eval = FALSE, echo = FALSE, out.width = "100%", fig.width = 9, fig.asp = 0.5, fig.cap = "Left: Posterior probability of the EEG difference (slope) being above 0 according to the BGAMM.
-Right: Ratio of posterior probability according to the BGAMM (on a log10 scale).
-Timesteps above threshold (10) are highlighted in green.
-NB: the minimum and maximum possible ratio values are determined (bounded) by the number of posterior samples in the model."}
-# Computing the EDF?
-# From https://discourse.mc-stan.org/t/standard-operating-procedures-for-anova/3964/2
-# If Kruschke has code that does what you want, then you can probably call his code on some columns of as.matrix(fit).
-If it really has to be in coda format, there is an As.mcmc.list function in the rstan package that will convert a stanfit object to a mcmc.list, but you probably have to call it on the $stanfit element of the brmsfit object.
-# See the bamlss::samplestat() function at https://github.com/cran/bamlss/blob/0a2a4154e0c0099ba8ac0b372ae7460e57972741/R/BAMLSS.R#L2397
-# Maybe export brms posterior samples to mcmc.list and then use this function?
-# See https://mc-stan.org/rstan/reference/As.mcmc.list.html
-library(bamlss)
-# mcmc_samples <- as.mcmc(meg_decoding_gam)
-mcmc_samples <- as.mcmc(meg_decoding_gam$fit)
-mcmc_samples_stats <- samplestats(
-    samples = mcmc_samples,
-    x = bamlss.frame(formula = auc_mean ~ s(time, bs = "tp", k = 50), data = decoding_summary_df)$x,
-    # y = bamlss.frame(formula = auc_mean ~ s(time, bs = "tp", k = 50), data = decoding_summary_df)$y,
-    # family = beta_bamlss,
-    # logLik = FALSE
-    )
-mcmc_samples_stats
-
-# or using bamlss directly
-b_model <- bamlss(
-    formula = auc_mean ~ s(time, bs = "tp", k = 50),
-    family = beta_bamlss,
-    data = decoding_summary_df
-    )
-
-# computing the EDF (around 44.6)
-samplestats(samples = b_model)
-
-# comparing to p_WAIC (around 48.4)
-waic(meg_decoding_gam)
-
-# computing the posterior probability ratio (PPR)
-model <- meg_decoding_gam
-threshold <- 10
-
-# defining a function to compute and plot the posterior odds
-plot_posterior_odds <- function (model, threshold = 10) {
-
-    # retrieving the number of posterior samples
-    n_posterior_samples <- ndraws(model)
-
-    # defiing the full time sequence
-    # time_seq <- seq(min(model$data$time), max(model$data$time), length.out = 300)
-    time_seg <- unique(model$data$time)
-
-    # generating new data
-    newdata <- data.frame(time = time_seq)
-
-    # getting posterior expected values (marginal)
-    posterior_preds <- posterior_epred(object = model, newdata = newdata, re_formula = NA)
-
-    # computing posterior mean and 95% CI at each time
-    posterior_preds_df <- as.data.frame(posterior_preds) %>%
-        mutate(draw = row_number() ) %>%
-        pivot_longer(-draw, names_to = "time_index", values_to = "auc") %>%
-        mutate(time = rep(time_seq, each = nrow(posterior_preds) ) ) %>%
-        group_by(time) %>%
-        summarise(
-            auc_mean = mean(auc),
-            auc_lower = quantile(auc, 0.05),
-            auc_upper = quantile(auc, 0.95)
-            ) %>%
-        ungroup()
-
-    # computing the average upper 96% credible interval during baseline (i.e., before 0s)
-    upper_baseline_ci <- posterior_preds_df %>%
-        filter(time < 0) %>%
-        summarise(upper_baseline_ci = mean(auc_upper) ) %>%
-        pull(upper_baseline_ci)
-
-    # computing the posterior probability of AUC > upper CrI of baseline
-    prob_y_above_baseline <- model$data %>%
-        add_epred_draws(object = model, ndraws = 100) %>%
-        data.frame() %>%
-        # head()
-        # dplyr::select(participant, time, condition, .epred, .draw) %>%
-        # pivot_wider(names_from = condition, values_from = .epred) %>%
-        # mutate(epred_diff = cond2 - cond1) %>%
-        # dplyr::select(time, baseline, .epred, .draw) %>%
-        # pivot_wider(names_from = baseline, values_from = .epred) %>%
-        # head()
-        # mutate(epred_diff = `c - cond1) %>%
-        # posteriors_preds %>%
-        # computing the mean posterior probability at the group level
-        group_by(time) %>%
-        summarise(m = mean(.epred > upper_baseline_ci) ) %>%
-        mutate(prob_ratio = m / (1 - m) ) %>%
-        ungroup() %>%
-        # bounding posterior probabilities by the number of posterior samples
-        mutate(prob_ratio = pmin(prob_ratio, n_posterior_samples) ) %>%
-        mutate(prob_ratio = pmax(prob_ratio, 1 / n_posterior_samples) ) %>%
-        ungroup()
-
-    # plotting it
-    post_plot <- prob_y_above_baseline %>%
-        ggplot(aes(x = time, y = m) ) +
-        geom_hline(yintercept = 0.5, linetype = 2) +
-        # geom_vline(xintercept = true_onset, linetype = 2) +
-        # geom_vline(xintercept = true_offset, linetype = 2) +
-        geom_line() +
-        ylim(c(0, 1) ) +
-        labs(x = "Time (s)", y = expression(p(beta>0~"|"~data, model) ) )
-
-    # plotting it
-    ratio_plot <- prob_y_above_baseline %>%
-        mutate(above_thres = ifelse(prob_ratio >= threshold, 1, NA) ) %>%
-        ggplot(aes(x = time, y = prob_ratio) ) +
-        geom_hline(yintercept = 1, linetype = "dashed") +
-        geom_hline(yintercept = 10, linetype = "dashed", color = "darkred") +
-        geom_hline(yintercept = 1/10, linetype = "dashed", color = "darkred") +
-        geom_hline(yintercept = 100, linetype = "dashed", color = "orangered") +
-        geom_hline(yintercept = 1/100, linetype = "dashed", color = "orangered") +
-        # geom_vline(xintercept = true_onset, linetype = 2) +
-        # geom_vline(xintercept = true_offset, linetype = 2) +
-        geom_line(linewidth = 0.8) +
-        geom_point(
-            data = . %>% dplyr::filter(!is.na(above_thres) ),
-            aes(y = threshold),
-            colour = "darkgreen",
-            shape = 16,
-            size = 2,
-            na.rm = TRUE
-            ) +
-        scale_y_log10(
-            labels = label_log(digits = 2),
-            limits = c(1 / n_posterior_samples, n_posterior_samples),
-            breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
-            ) +
-        labs(
-            x = "Time (s)",
-            y = expression(p(beta>0~"|"~data) / (1 - p(beta>0~"|"~data) ) )
-            )
-
-    # plotting both the posterior probability and the ratio...
-    post_plot + ratio_plot
-
-}
-
-# plotting the posterior odds
-plot_posterior_odds(model = meg_decoding_gam, threshold = 10)
-
-####################################
-# or testing an alternative method #
-####################################
-
-# defining the full time sequence
-time_seq <- seq(min(decoding_summary_df$time), max(decoding_summary_df$time), length.out = 300)
-
-# generating new data
-newdata <- data.frame(time = time_seq)
-
-# retrieving the posterior expected values (marginal)
-posterior <- posterior_epred(object = meg_decoding_gam, newdata = newdata, re_formula = NA)
-
-# computing posterior mean and 95% CI at each time
-pred_df <- as.data.frame(posterior) %>%
-    mutate(draw = row_number() ) %>%
-    pivot_longer(-draw, names_to = "time_index", values_to = "auc") %>%
-    mutate(time = rep(time_seq, each = nrow(posterior) ) ) %>%
-    group_by(time) %>%
-    summarise(
-        auc_mean = mean(auc),
-        auc_lower = quantile(auc, 0.025),
-        auc_upper = quantile(auc, 0.975)
-        )
-
-# subsetting newdata to time < 0
-baseline_indices <- which(time_seq < 0)
-baseline_draws <- posterior[, baseline_indices]
-
-# computing the average baseline AUC per draw
-baseline_per_draw <- rowMeans(baseline_draws)
-
-# for timepoints ≥ 0
-post_indices <- which(time_seq >= 0)
-time_post <- time_seq[post_indices]
-
-# subsetting post-stimulus predictions
-posterior_post <- posterior[, post_indices]
-
-# computing posterior probability that AUC > pre-stim baseline
-post_prob <- sapply(1:length(post_indices), function (i) {
-
-    mean(posterior_post[, i] > baseline_per_draw)
-
-})
-
-# assembling into dataframe
-post_df <- data.frame(
-    time = time_post,
-    prob_auc_gt_baseline = post_prob
-    )
-
-# plotting the results
-n_posterior_samples <- ndraws(meg_decoding_gam)
-post_df %>%
-    ggplot(aes(x = time, y = prob_auc_gt_baseline / (1-prob_auc_gt_baseline) ) ) +
-    geom_line(size = 1, color = "darkgreen") +
-    geom_hline(yintercept = 0.95, linetype = "dashed", color = "gray") +
-    scale_y_log10(
-        labels = label_log(digits = 2),
-        limits = c(1 / n_posterior_samples, n_posterior_samples),
-        breaks = c(1/1e4, 1/1e3, 1/1e2, 1/10, 1, 10, 1e2, 1e3, 1e4)
-        ) +
-    labs(
-        title = "Posterior Probability AUC > Pre-Stimulus Baseline",
-        x = "Time (s)", y = "P(AUC > baseline)"
-        ) +
-    theme_bw()
-
-# or using yet another approach
-estimate_edf_brms <- function (
-    model, 
-    term,           # e.g., "s(time)"
-    time_seq = NULL,
-    n = 200,        # number of time points
-    re_formula = NA # exclude group-level effects
-    ) {
-
-  # Extract time variable name from term, e.g., "s(time)" → "time"
-  time_var <- gsub("s\\(([^,]+).*\\)", "\\1", term)
-
-  # Create prediction grid
-  if (is.null(time_seq)) {
-    time_range <- range(model$data[[time_var]], na.rm = TRUE)
-    time_seq <- seq(time_range[1], time_range[2], length.out = n)
-  }
-
-  # newdata <- data.frame(!!time_var := time_seq)
-  newdata <- data.frame(time_seq)
-  names(newdata) <- time_var
-
-  # Extract posterior smooth draws only (excludes intercept & fixed effects)
-  smooth_draws <- posterior_smooths(
-    model,
-    newdata = newdata,
-    smooth = term,
-    re_formula = re_formula,
-    resp = NULL
-  )
-
-  # Compute posterior mean at each timepoint
-  posterior_mean <- colMeans(smooth_draws)
-
-  # Compute variance of the posterior mean across time
-  var_mean <- var(posterior_mean)
-
-  # Compute variance at each timepoint across draws
-  pointwise_var <- apply(smooth_draws, 2, var)
-
-  # Mean pointwise variance
-  mean_pointwise_var <- mean(pointwise_var)
-
-  # EDF estimate
-  edf_estimate <- var_mean / mean_pointwise_var
-
-  return(list(
-    edf = edf_estimate,
-    var_mean = var_mean,
-    mean_pointwise_var = mean_pointwise_var,
-    time = time_seq,
-    smooth_mean = posterior_mean,
-    smooth_sd = sqrt(pointwise_var)
-  ))
-}
-
-edf_result <- estimate_edf_brms(
-  model = meg_decoding_gam,
-  #term = "s(time)"
-  term = 's(time,bs="tp",k=50)'
-)
-
-print(paste("Estimated EDF:", round(edf_result$edf, 2)))
-
-plot_df <- data.frame(
-  time = edf_result$time,
-  mean = edf_result$smooth_mean,
-  sd = edf_result$smooth_sd
-)
-
-ggplot(plot_df, aes(x = time, y = mean)) +
-  geom_line(color = "blue") +
-  geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd), alpha = 0.2) +
-  labs(
-    title = paste("Smooth with Posterior SD | Estimated EDF ≈", round(edf_result$edf, 2)),
-    x = "Time", y = "Smooth Estimate"
-  ) +
-  theme_bw()
-```
 
@@
 
-# importing results
-# reliability_results <- readRDS(file = "results/reliability_results_100splits.rds")
-
-# loading all RDS files
-# files <- list.files(
-#     path = "results/reliability_cluster_results",
-#     pattern = ".rds", 
-#     full.names = TRUE
-#     )
-# files <- list.files(
-#     # path = "reliability_cluster_results",
-#     path = "reliability_jaccard_cluster_results",
-#     pattern = ".rds", 
-#     full.names = TRUE
-#     )
-
-# importing the data splits
-# data_splits <- list.files(
-#     path = "reliability_data_splits",
-#     pattern = ".csv",
-#     full.names = TRUE
-#     )
-
-# sorting files numerically by the number in the filename
-# files <- files[order(readr::parse_number(files) )]
-# data_splits <- data_splits[order(readr::parse_number(data_splits) )]
-
-# reading and binding all files
-# reliability_results <- map2_dfr(
-#     .x = files,
-#     .y = seq_along(files),
-#     .f = ~ readRDS(.x)
-#     # .f = ~ readRDS(.x) |> mutate(sim_id = .y)
-#     # .f = function (file, index) {
-#     #     offset <- (index - 1) * splits_per_file
-#     #     readRDS(file) %>% mutate(split_id = as.numeric(split_id) ) %>%
-#     #         mutate(split_id = split_id + offset) %>%
-#     #         mutate(split_id = formatC(x = split_id, width = 3, flag = 0) ) 
-#     #     }
-#     )
-
-# checking errors
-# reliability_results %>%
-#     filter(!is.na(error) )
-
-# removing NAs
-# reliability_results <- reliability_results %>%
-#     # select(-error, -stringAsFactors) %>%
-#     na.omit()
-
-# sanity checks (N=1122)
-# unique(reliability_results$split_id)
-# n_distinct(reliability_results$split_id)
-
-# reading and binding all files, adding a split_id
-# splits_results <- map2_dfr(
-#     .x = data_splits,
-#     .y = seq_along(data_splits),
-#     .f = ~ read.csv(.x)
-#     ) %>%
-#     # mutate(array_id = as.numeric(array_id) ) %>%
-#     mutate(split_id = paste(array_id, sprintf("s%03d", split_number), sep = "_") ) %>%
-#     # removing unused columns
-#     select(-split_number, -array_id)
-
-# or reading splits list directly
-# splits_results <- read.csv(file = "uniform_jaccard_data_splits.csv")
-# splits_results <- read.csv(file = "uniform_jaccard_data_splits_100pairs.csv")
-# splits_results <- read.csv(file = "data/uniform_jaccard_data_splits_1000pairs.csv")
-
-# extracting the full data estimates
-# full_reliability_results <- reliability_results %>%
-#     # filter(split_id == max(split_id) ) %>%
-#     # filter(str_detect(string = split_id, pattern = "_s051") ) %>%
-#     # dplyr::filter(split_id == "a001_s051") %>%
-#     dplyr::filter(split_id == "1_full") %>%
-#     # keeping only the first and last timestep
-#     # group_by(split_id, method, onset_offset) %>%
-#     # summarise(
-#     #     value = if (onset_offset[1] == "onset") min(value) else max(value),
-#     #     .groups = "drop"
-#     #     ) %>%
-#     # ungroup() %>%
-#     # pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
-#     # data.frame() %>%
-#     # select(-split_id) %>%
-#     # separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
-#     pivot_wider(names_from = onset_offset, values_from = value) %>%
-#     mutate(
-#         method = factor(
-#             x = method,
-#             levels = c(
-#                 "raw_p", "pval_bh", "pval_by", "pval_holm", "cluster_mass",
-#                 "cluster_tfce", "cpt", "brms"
-#                 ),
-#             labels = c(
-#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-#                 "TFCE", "Change point", "BGAM (our method)"
-#                 )
-#             )
-#         ) %>%
-#     mutate(y_offset = 0.5 - as.numeric(factor(method) ) * 0.01 - 0.01)
-
 # checking model's predictions against raw data
@@
     geom_vline(xintercept = 0.0, linetype = 2) +
-    # annotate(
-    #     geom = "rect",
-    #     xmin = -Inf, xmax = Inf,
-    #     ymin = chance_level - sesoi, ymax = chance_level + sesoi,
-    #     fill ="orangered",
-    #     alpha = 0.2
-    #     ) +
     geom_segment(
@@
     labs(x = "Time (s)", y = "Decoding accuracy (ROC AUC)", colour = NULL)
-
-# saving the plot
-# ggsave(
-#     filename = "figures/meg_decoding_results_all_methods.png",
-#     width = 12, height = 6, dpi = 300,
-#     device = "png"
-#     )
 ```
@@
 
-<!--
-
-Finally, we fitted a polynomial (cubic) regression to predict the absolute difference in onset/offset estimates according to the Jaccard similarity.
-This analysis allowed us to disentangle two properties of each method: i) whether it produces similar onset/offset estimates for similar subsets (in the limit case where subsets are identical, this is reflected by the intercept parameter), and ii) how sensitive each method is to variation in subsets' similarity (which is reflected by the two slope parameters, for the linear and cubic terms, respectively).
-In simple terms, sensitive methods should produce similar onset/offset estimates for similar subsets and dissimilar onset/offset estimates for dissimilar subsets.
-The results of this procedure are summarised in @fig-splits-similarity.
-
-This figure reveals that the `BGAM` *onset and offset* estimates on each split are the closest to the estimates from the full dataset on average (0ms difference for the onset estimate and 5ms difference for the offset estimate).
-The `Raw p-value` method has similar performance for the offset, but given the aberrant estimates it produces (cf.
-@fig-onset-offset), the fact that it is consistent between data splits and the full dataset is not convincing on its own.
-The `Change point` method also has a very good performance (i.e., very low difference between split estimates and full estimates), but produces too short cluster of significant decoding performance (cf.
-@fig-onset-offset).[^3] Overall, the figure reveals that for all other methods, split datasets produce later onset estimates and earlier offset estimates (compared to the estimates from the model fitted on the full dataset).
-These results highlight some desirable properties for a method aiming to precisely and reliably estimate the onset and offset of M/EEG effects, namely, it should i) have good asymptotic properties on simulated data, ii) provide sensible identified clusters in actual data, and iii) provide reliable/stable estimates across various partitions of the data.
-All these desiderata seem to be fulfilled by the BGAM approach.
-
--->
-
 ```{r fig-splits-similarity, eval = TRUE, echo = FALSE, fig.width = 12, fig.asp = 0.6, fig.cap = "Relation between data subsets' dissimilarity (x-axis) and difference in onset (blue) and offset (orange) estimates (y-axis) according to each method."}
 # reading splits list directly
-# splits_results <- read.csv(file = "uniform_jaccard_data_splits.csv")
-# splits_results <- read.csv(file = "uniform_jaccard_data_splits_100pairs.csv")
 splits_results <- read.csv(file = "data/uniform_jaccard_data_splits_1000pairs.csv")
@@
 
-# checking errors
+# checking potential errors
 # reliability_results %>%
@@
 
-# This analysis assesses whether the method is "self-consistent" across data splits...
-# Step 1: creating a list of participants per split_id
-# split_participants <- splits_results %>%
-#     rename(split_id = split_number) %>%
-#     group_by(split_id) %>%
-#     dplyr::filter(group == "A") %>%
-#     summarise(participants = list(participant) ) %>%
-#     ungroup()
-
-# Step 2: creating all pairwise combinations of split_ids
-# split_pairs <- expand.grid(
-#     split_id1 = split_participants$split_id,
-#     split_id2 = split_participants$split_id,
-#     stringsAsFactors = FALSE
-#     ) %>%
-#     # avoiding duplicates and self-comparison
-#     dplyr::filter(split_id1 < split_id2)
-
-# Step 3: joining the participant lists for each pair
-# split_pairs <- split_pairs %>%
-#     left_join(split_participants, by = c("split_id1" = "split_id") ) %>%
-#     rename(participants1 = participants) %>%
-#     left_join(split_participants, by = c("split_id2" = "split_id") ) %>%
-#     rename(participants2 = participants)
-
-# Step 4: computing overlap as size of intersection / union
-# see also https://www.rdocumentation.org/packages/genieclust/versions/1.1.6/topics/compare_partitions
-# split_pairs <- split_pairs %>%
-#     mutate(
-#         intersection = map2_int(participants1, participants2, ~length(intersect(.x, .y) ) ),
-#         union = map2_int(participants1, participants2, ~length(union(.x, .y) ) ),
-#         jaccard_similarity = intersection / union,
-#         # jaccard_similarity = map2_dbl(participants1, participants2, ~jaccard(.x, .y) )
-#         # ari = map2(participants1, participants2, ~ARI(.x[[1]], .y[[1]]) ),
-#         # nmi = map2_dbl(participants1, participants2, ~NMI(.x[[1]], .y[[1]]) ),
-#         # ami = map2_dbl(participants1, participants2, ~AMI(.x[[1]], .y[[1]]), .progress = TRUE)
-#         )
-
-# checking the results
-# sort(unique(split_pairs$intersection) )
-# hist(split_pairs$intersection)
-
-# plotting it
-# split_pairs %>%
-#     ggplot(aes(x = intersection, y = jaccard_similarity) ) +
-#     geom_point() +
-#     geom_smooth(method = "lm") +
-#     theme_bw()
-
-##########################################################
-# now computing similarity of onset and offset estimates #
-##########################################################
-
-# Step 1: creating a list of participants per split_id
-# reliability_splits <- reliability_results %>%
-#     # filter(str_detect(string = split_id, pattern = "full", negate = TRUE) ) %>%
-#     # mutate(split_id = str_sub(string = split_id, start = 6, end = -1) )
-#     dplyr::filter(split_id != "a001_s051") %>%
-#     # keeping only the first and last timestep
-#     group_by(split_id, method, onset_offset) %>%
-#     summarise(
-#         value = if (onset_offset[1] == "onset") min(value) else max(value),
-#         .groups = "drop"
-#         ) %>%
-#     ungroup()
-
-# Step 2: creating all pairwise combinations of split_ids
-# and computing the absolute difference of estimated onset/offset
-# reliability_splits_results <- reliability_splits %>%
-#   # joining the data frame with itself by method and onset_offset
-#   inner_join(
-#       reliability_splits,
-#       by = c("method", "onset_offset"),
-#       suffix = c("_1", "_2")
-#       ) %>%
-#   # keeping only unique pairs (avoid self-pairs and duplicates)
-#   dplyr::filter(split_id_1 < split_id_2) %>%
-#   # computing the absolute difference in value
-#   mutate(value_diff = abs(value_1 - value_2) ) %>%
-#   select(method, onset_offset, split_id_1, split_id_2, value_diff)
-
-# joining these results with splits similarity
-# head(reliability_splits_results)
-# full_df <- split_pairs %>%
-#     select(split_id_1 = split_id1, split_id_2 = split_id2, jaccard_similarity) %>%
-#     # head()
-#     inner_join(reliability_splits_results, by = c("split_id_1", "split_id_2") )
-
-# checking similarity values and effectives
-# full_df %>%
-#     summarise(
-#         effective_sim = n(),
-#         .by = c(onset_offset, method, jaccard_similarity)
-#         ) %>%
-#     arrange(jaccard_similarity)
-
-########################################################################
-# or working directly from pre-existing jaccard similarity csv file
-######################################################################
-
+# importing the reliability results
 reliability_results2 <- reliability_results %>%
-    # dplyr::filter(str_detect(string = split_id, pattern = "full", negate = TRUE) ) %>%
     # keeping only the first and last timestep
@@
     mutate(value_diff = abs(diff(value) ) ) %>%
-    # mutate(value_diff = diff(value) ) %>%
-    # mutate(value_diff = diff(value)^2) %>%
-    # rmse = sqrt(mean(error)^2),
     ungroup() %>%
@@
 # full_df[!complete.cases(full_df), ] %>% data.frame()
-# reliability_results2 %>% filter(split_id == "s00088")
-# reliability_results %>% filter(str_detect(string = split_id, pattern = "s00088") )
-
-# custom function to implement min-max scaling (i.e., between 0 and 1)
-# minMax <- function (x) {
-#     
-#     std_score <- (x - min(x) ) / (max(x) - min(x) )
-#     
-#     return (std_score)
-#     
-# }
-
-# polynomial regression degree
-# poly_reg_degree <- 2
-
-# or fitting a GAM for each combination
-# library(ggpmisc)
-# library(broom)
-# library(glue)
-# library(mgcv)
-# 
-# # 1. Nest data by method & onset_offset
-# model_df <- full_df %>%
-#     na.omit() %>%
-#     group_by(method, onset_offset, jaccard) %>%
-#     summarise(
-#         mean_value_diff = mean(x = value_diff, na.rm = TRUE),
-#         sem = sd(x = value_diff, na.rm = TRUE) / sqrt(n() )
-#         ) %>%
-#     ungroup() %>%
-#     mutate(
-#         method = factor(
-#             x = method,
-#             levels = c(
-#                 "raw_p", "pval_bh", "pval_by", "pval_holm", "cluster_mass",
-#                 "cluster_tfce", "cpt", "brms"
-#                 ),
-#             labels = c(
-#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-#                 "TFCE", "Change point", "BGAM (our method)"
-#                 )
-#             )
-#         ) %>%
-#     # mutate(jaccard = 1-jaccard) %>%
-#     filter(!is.na(mean_value_diff), !is.na(jaccard) ) %>%
-#     group_by(method, onset_offset) %>%
-#     nest()
-# 
-# # 2. Fit a GAM to each subset
-# model_df <- model_df %>%
-#     mutate(
-#         gam_model = map(
-#             data, ~ gam(mean_value_diff ~ s(jaccard),
-#                         # family = Gamma(link = "log"),
-#                         data = .x)
-#             )
-#         )
-# 
-# # 3. Generate pretty label with edf and R² for plotting
-# model_df <- model_df %>%
-#     mutate(
-#         eq_label = map2_chr(
-#             gam_model,
-#             data,
-#             ~ {
-#                 r2 <- summary(.x)$r.sq
-#                 edf <- summary(.x)$s.table[, "edf"] %>% round(2)
-#                 glue("edf = {edf}, R² = {round(r2, 2)}")
-#                 })
-#         )
-# 
-# # 4. Create predictions for plotting
-# model_df <- model_df %>%
-#   mutate(
-#     preds = map2(data, gam_model, ~ {
-#       tibble(
-#         jaccard = seq(min(.x$jaccard), max(.x$jaccard), length.out = 100)
-#       ) %>%
-#         mutate(mean_value_diff = predict(.y, newdata = ., type = "response"))
-#     })
-#   )
-# 
-# # 5. Unnest everything for ggplot
-# plot_data <- model_df %>%
-#   select(method, onset_offset, eq_label, preds) %>%
-#   unnest(preds)
-# 
-# # 6. Create base plot with data points and fitted GAM lines
-# full_df %>%
-#     na.omit() %>%
-#     # nrow()
-#     # dplyr::filter(jaccard < 1 & jaccard > 0) %>%
-#     group_by(method, onset_offset, jaccard) %>%
-#     summarise(
-#         mean_value_diff = mean(x = value_diff, na.rm = TRUE),
-#         # mean_value_diff = sqrt(mean(x = value_diff, na.rm = TRUE) ),
-#         sem = sd(x = value_diff, na.rm = TRUE) / sqrt(n() )
-#         ) %>%
-#     ungroup() %>%
-#     mutate(
-#         method = factor(
-#             x = method,
-#             levels = c(
-#                 "raw_p", "pval_bh", "pval_by", "pval_holm", "cluster_mass",
-#                 "cluster_tfce", "cpt", "brms"
-#                 ),
-#             labels = c(
-#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-#                 "TFCE", "Change point", "BGAM (our method)"
-#                 )
-#             )
-#         ) %>%
-#     ggplot(aes(x = 1-jaccard, y = mean_value_diff, colour = onset_offset) ) +
-#     geom_point(alpha = 0.6) +
-#     geom_line(data = plot_data, aes(y = mean_value_diff), size = 1.2) +
-#     geom_label(
-#         data = plot_data %>% group_by(method, onset_offset) %>% slice(1),
-#         aes(label = eq_label, x = 0.05, y = Inf),
-#         vjust = 1.2, hjust = 0,
-#         fill = "white", colour = "black",
-#         size = 3, inherit.aes = FALSE
-#         ) +
-#     facet_wrap(~method, nrow = 2) +
-#     theme_bw() +
-#         labs(
-#         x = "Jaccard dissimilarity",
-#         y = "Average absolute difference in onset/offset estimate (s)",
-#         colour = ""
-#         )
 
@@
     mutate(jaccard_dis = 1-jaccard) %>%
-    # dplyr::filter(jaccard < 1 & jaccard > 0) %>%
-    # dplyr::filter(1-jaccard > 0 & 1-jaccard < 1) %>%
     dplyr::filter(jaccard_dis > 0) %>%
@@
         aes(
-            # x = minMax(1-jaccard_similarity),
-            # x = 1-jaccard_similarity,
-            # x = max(jaccard)-jaccard,
             x = jaccard_dis,
@@
         ) +
-    # geom_point(size = 0.5, alpha = 0.05) +
     geom_smooth(
-        # method = "glm",
-        # method.args = list(family = Gamma(link = "inverse") ),
-        # method.args = list(family = Gamma(link = "log") ),
-        # method.args = list(family = inverse.gaussian() ),
         method = "lm",
-        # method = "gam",
-        # formula = y ~ stats::poly(x, degree = 3, raw = TRUE),
-        # formula = y ~ stats::poly(x, degree = poly_reg_degree, raw = TRUE),
-        # formula = rank(y) ~ rank(x),
-        # formula = y ~ logit_scaled(x),
-        # method.args = list(family = "binomial"),
         alpha = 0.2,
@@
         ) +
-    # geom_point() +
     geom_pointrange(
@@
         # showing the R and R^2
-        # aes(label = paste(..r.label.., ..rr.label.., sep = "~`,`~") ),
         aes(label = paste(after_stat(r.label), after_stat(rr.label), sep = "~`,`~") ),
         method = "spearman",
-        # method = "pearson",
-        # method = "kendall",
         digits = 3,
-        # label.x = 0,
-        # label.y = 1,
         label.x.npc = "left",
@@
         ) +
-    # stat_regline_equation(
-    #     aes(label = paste(..eq.label.., ..rr.label.., sep = "~`,`~") ),
-    #     # formula = rank(y) ~rank(x),
-    #     # formula = y ~ logit_scaled(x),
-    #     formula = y ~ stats::poly(x = x, degree = poly_reg_degree, raw = TRUE),
-    #     # method = "glm",
-    #     # method.args = list(family = Gamma(link = "log")),
-    #     label.x = 0,
-    #     label.y.npc = "top",
-    #     geom = "label",
-    #     colour = "white",
-    #     size = 3.5
-    #     ) +
-    # scale_y_log10(guide = "axis_logticks", limits = c(NA, 0.5) ) +
     scale_y_log10(guide = "axis_logticks", limits = c(NA, 1) ) +
@@
         )
-
-# saving the plot
-# ggsave(
-#     filename = "figures/splits_summary_similarity_100pairs.png",
-#     width = 16, height = 8, dpi = 200,
-#     device = "png"
-#     )
 ```
@@
 
-<!--
-
-@fig-reliability shows the median difference between the onset and offset estimates from each data split and the onset and offset estimates from the full dataset (x-axis) along with the variance of its onset and offset estimates across data splits (error bar).
-This figure reveals that the `BGAM` *onset and offset* estimates on each split are the closest to the estimates from the full dataset on average (0ms difference for the onset estimate and 5ms difference for the offset estimate).
-The `Raw p-value` method has similar performance for the offset, but given the aberrant estimates it produces (cf.
-@fig-onset-offset), the fact that it is consistent between data splits and the full dataset is not convincing on its own.
-The `Change point` method also has a very good performance (i.e., very low difference between split estimates and full estimates), but produces too short cluster of significant decoding performance (cf.
-@fig-onset-offset).[^3] Overall, the figure reveals that for all other methods, split datasets produce later onset estimates and earlier offset estimates (compared to the estimates from the model fitted on the full dataset).
-These results highlight some desirable properties for a method aiming to precisely and reliably estimate the onset and offset of M/EEG effects, namely, it should i) have good asymptotic properties on simulated data, ii) provide sensible identified clusters in actual data, and iii) provide reliable/stable estimates across various partitions of the data.
-All these desiderata seem to be fulfilled by the BGAM approach.
-
-[^3]: As in @rousselet_using_2025, we fixed the number of expected change points to two in the binary segmentation algorithm, thus producing always one cluster.
-
-```{r fig-reliability, eval = FALSE, echo = FALSE, cache = FALSE, fig.width = 8, fig.asp = 0.5, fig.cap = "Median error and median absolute deviation of the error for the onset (left) and offset (right) estimates according to each method.
-Methods are ordered from smallest (top) to largest (bottom) median absolute error separately for the onset and offset estimates."}
-# extracting the full data estimates
-full_reliability_results <- reliability_results %>%
-    # filter(split_id == max(split_id) ) %>%
-    filter(split_id == "a001_s051") %>%
-    # keeping only the first and last timestep
-    group_by(split_id, method, onset_offset) %>%
-    summarise(
-        value = if (onset_offset[1] == "onset") min(value) else max(value),
-        .groups = "drop"
-        ) %>%
-    ungroup() %>%
-    data.frame() %>%
-    select(-split_id)
-
-# analysing the split results
-# reliability_results %>%
-#     slice(1:nrow(.)-1) %>%
-#     pivot_longer(cols = -split_id, names_to = "measure", values_to = "value") %>%
-#     data.frame() %>%
-#     left_join(
-#         full_reliability_results,
-#         by = "measure",
-#         suffix = c("_split", "_full")
-#         ) %>%
-#     # mutate(error = abs(value_split - value_full) ) %>%
-#     mutate(error = value_split - value_full) %>%
-#     # mutate(error = (value_split - value_full)^2) %>%
-#     separate(measure, into = c("type", "method"), sep = "_", extra = "merge") %>%
-#     group_by(type, method) %>%
-#     summarise(
-#         MAE = median(error, na.rm = TRUE),
-#         # variance = var(error, na.rm = TRUE),
-#         variance = var(value_split, na.rm = TRUE)
-#         # variance = sum(error, na.rm = TRUE)
-#         # .groups = "drop"
-#         ) %>%
-#     ungroup() %>%
-#     # now plotting
-#     mutate(type = factor(x = type, levels = c("onset", "offset") ) ) %>%
-#     mutate(
-#         method = factor(
-#             x = method,
-#             levels = c(
-#                 "p", "bh", "by", "holm", "cluster_mass",
-#                 "cluster_tfce", "cpt", "brms", "brms_full"
-#                 ),
-#             labels = c(
-#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-#                 "TFCE", "Change point", "brms", "brms_full"
-#                 ),
-#             )
-#         ) %>%
-#     ggplot(
-#         aes(
-#             x = MAE,
-#             y = variance,
-#             colour = method,
-#             fill = method
-#             )
-#         ) +
-#     geom_vline(xintercept = 0, linetype = 2) +
-#     geom_point(
-#         size = 2,
-#         pch = 21,
-#         colour = "white",
-#         show.legend = FALSE
-#         ) +
-#     geom_label_repel(
-#         aes(label = method),
-#         colour = "white",
-#         size = 3,
-#         segment.color = NA,
-#         max.overlaps = 20,
-#         show.legend = FALSE
-#         ) +
-#     scale_y_log10() +
-#     facet_wrap(~type, scales = "free") +
-#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 9) ) +
-#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 9) ) +
-#     labs(
-#         x = "Average (median) difference to full dataset (s)",
-#         y = "Variance across splits"
-#         )
-
-# defining a consistent colour palette using original method levels
-original_methods <- c(
-    "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-    "TFCE", "Change point", "BGAM (our method)"
-    )
-
-color_palette <- setNames(
-    met.brewer(name = "Johnson", n = length(original_methods) ),
-    original_methods
-    )
-
-# sanity check
-# reliability_results %>%
-#     filter(str_detect(string = split_id, pattern = "s051", negate = TRUE) ) %>%
-#     pull(split_id) %>%
-#     n_distinct()
-
-# plotting the error in descending order
-reliability_results %>%
-    filter(str_detect(string = split_id, pattern = "s051", negate = TRUE) ) %>%
-    data.frame() %>%
-    # keeping only the first and last timestep
-    group_by(split_id, method, onset_offset) %>%
-    summarise(
-        value = if (onset_offset[1] == "onset") min(value) else max(value),
-        .groups = "drop"
-        ) %>%
-    ungroup() %>%
-    select(-split_id) %>%
-    left_join(
-        full_reliability_results,
-        # by = "measure",
-        by = c("method", "onset_offset"),
-        suffix = c("_split", "_full")
-        ) %>%
-    group_by(method, onset_offset) %>%
-    # mutate(error = abs(value_split - value_full) ) %>%
-    mutate(error = value_split - value_full) %>%
-    summarise(
-        # mean_error = mean(error, na.rm = TRUE),
-        # variance = var(error, na.rm = TRUE),
-        # variance = var(value_split, na.rm = TRUE)
-        # variance = sd(error, na.rm = TRUE) / sqrt(n() )
-        # variance = sum(error, na.rm = TRUE)
-        mean_error = mean(error, na.rm = TRUE),
-        median_error = median(error, na.rm = TRUE),
-        # variance = IQR(error, na.rm = TRUE)
-        mad = mad(error, na.rm = TRUE),
-        variance = var(error, na.rm = TRUE)
-        # .groups = "drop"
-        ) %>%
-    ungroup() %>%
-    # head()
-    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
-    mutate(
-        method = factor(
-            x = method,
-            levels = c(
-                "raw_p", "pval_bh", "pval_by", "pval_holm", "cluster_mass",
-                "cluster_tfce", "cpt", "brms"
-                ),
-            labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-                "TFCE", "Change point", "BGAM (our method)"
-                ),
-            )
-        ) %>%
-    mutate(
-        method_reordered = tidytext::reorder_within(
-            x = method,
-            by = 1 / abs(median_error),
-            # by = 1 / abs(mean_error),
-            within = onset_offset
-            )
-        ) %>%
-    ggplot(
-        aes(
-            x = median_error,
-            xmin = median_error - mad,
-            xmax = median_error + mad,
-            # x = mean_error,
-            # xmin = mean_error - variance,
-            # xmax = mean_error + variance,
-            y = method_reordered,
-            colour = method,
-            fill = method
-            )
-        ) +
-    geom_vline(xintercept = 0, linetype = 2) +
-    geom_pointrange(
-        size = 0.25,
-        linewidth = 0.75,
-        show.legend = FALSE
-        ) +
-    facet_wrap(~onset_offset, scales = "free") +
-    scale_y_reordered() +
-    scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-    labs(
-        x = "Difference to full dataset (s)",
-        y = "Method"
-        )
-
-# plotting the distribution of results
-# reliability_results %>%
-#     # filter(split_id < max(split_id) ) %>%
-#     filter(str_detect(string = split_id, pattern = "s051", negate = TRUE) ) %>%
-#     data.frame() %>%
-#     # keeping only the first and last timestep
-#     group_by(split_id, method, onset_offset) %>%
-#     summarise(
-#         value = if (onset_offset[1] == "onset") min(value) else max(value),
-#         .groups = "drop"
-#         ) %>%
-#     ungroup() %>%
-#     select(-split_id) %>%
-#     left_join(
-#         full_reliability_results,
-#         # by = "measure",
-#         by = c("method", "onset_offset"),
-#         suffix = c("_split", "_full")
-#         ) %>%
-#     group_by(method, onset_offset) %>%
-#     # mutate(error = abs(value_split - value_full) ) %>%
-#     mutate(error = value_split - value_full) %>%
-#     mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
-#     mutate(
-#         method = factor(
-#             x = method,
-#             levels = c(
-#                 "raw_p", "pval_bh", "pval_by", "pval_holm",
-#                 "cluster_mass", "cluster_tfce", "cpt", "brms"
-#                 ),
-#             labels = c(
-#                 "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-#                 "TFCE", "Change point", "BGAM (our method)"
-#                 ),
-#             )
-#         ) %>%
-#     mutate(
-#         method_reordered = tidytext::reorder_within(
-#             x = method,
-#             by = abs(error),
-#             fun = median,
-#             # by = 1 / abs(median_error),
-#             # by = 1 / abs(mean_error),
-#             within = onset_offset
-#             )
-#         ) %>%
-#     # mutate(
-#     #     method_reordered = fct_reorder(
-#     #         method, error, .fun = ~abs(median(.x) ), .na_rm = TRUE
-#     #         )
-#     #     ) %>%
-#     # arrange(desc(error) ) %>%
-#     # arrange(error) %>%
-#     # head()
-#     ggplot(
-#         aes(
-#             # x = median_error,
-#             x = error,
-#             # xmin = median_error - mad,
-#             # xmax = median_error + mad,
-#             # x = mean_error,
-#             # xmin = mean_error - variance,
-#             # xmax = mean_error + variance,
-#             y = method_reordered,
-#             colour = method,
-#             fill = method
-#             )
-#         ) +
-#     geom_vline(xintercept = 0, linetype = 2) +
-#     # geom_pointrange(
-#     #     size = 0.25,
-#     #     linewidth = 0.75,
-#     #     show.legend = FALSE
-#     #     ) +
-#     stat_slabinterval(
-#         n = 200,
-#         normalize = "groups",
-#         point_interval = "median_qi",
-#         # .width = c(0.1, 0.9),
-#         # .width = 0.2,
-#         .width = 0,
-#         interval_size_range = c(1, 1),
-#         fatten_point = 1.25,
-#         # slab_colour = "white",
-#         # slab_linewidth = 0,
-#         # density = "unbounded",
-#         slab_alpha = 0.5,
-#         show.legend = FALSE
-#         ) +
-#     facet_wrap(~onset_offset, scales = "free") +
-#     scale_y_reordered() +
-#     scale_fill_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-#     scale_colour_manual(values = met.brewer(name = "Johnson", n = 8) ) +
-#     labs(
-#         x = "Estimation error (s)",
-#         y = "Method"
-#         )
-```
-
-```{r tbl-reliability-results, eval = TRUE, echo = FALSE, results = "asis"}
-#| tbl-cap: "Summary statistics of the onset and offset estimates for each method (in ms, ordered by the MAE) according to the reliability simulation study."
-
-# extracting the full data estimates
-reliability_errors <- reliability_results %>%
-    data.frame() %>%
-    # keeping only the first and last timestep
-    group_by(split_id, method, onset_offset) %>%
-    summarise(
-        value = if (onset_offset[1] == "onset") min(value) else max(value),
-        .groups = "drop"
-        ) %>%
-    ungroup() %>%
-    select(-split_id) %>%
-    left_join(
-        full_reliability_results,
-        by = c("method", "onset_offset"),
-        suffix = c("_split", "_full")
-        ) %>%
-    group_by(method, onset_offset) %>%
-    mutate(error = value_split - value_full) %>%
-    ungroup()
-
-# summarising the results
-reliability_summary <- reliability_errors %>%
-    # mutate(
-    #     error = case_when(
-    #         grepl("onset", onset_offset) ~ value - true_onset,
-    #         grepl("offset", onset_offset) ~ value - true_offset
-    #         )
-    #     ) %>%
-    group_by(method, onset_offset) %>%
-    mutate(error = 1e3 * error) %>%
-    summarise(
-        # mean_error = mean(error, na.rm = TRUE),
-        median_error = median(error, na.rm = TRUE),
-        median_abs_error = median(abs(error), na.rm = TRUE),
-        rmse = sqrt(mean(error)^2),
-        variance = var(error, na.rm = TRUE),
-        mad = mad(error, na.rm = TRUE),
-        .groups = "drop"
-        ) %>%
-    ungroup() %>%
-    mutate(
-        method = factor(
-            x = method,
-            levels = c(
-                "raw_p", "pval_bh", "pval_by", "pval_holm",
-                "cluster_mass", "cluster_tfce", "cpt", "brms"
-                ),
-            labels = c(
-                "Raw p-value", "FDR BH95", "FDR BY01", "Holm", "Cluster mass",
-                "TFCE", "Change point", "BGAM (our method)"
-                ),
-            )
-        ) %>%
-    mutate(onset_offset = factor(x = onset_offset, levels = c("onset", "offset") ) ) %>%
-    data.frame()
-
-# creating the table
-reliability_summary %>%
-    mutate(
-        onset_offset = factor(
-            x = onset_offset,
-            levels = c("offset", "onset")
-            )
-        ) %>%
-    # arrange(median_abs_error) %>%
-    # arrange(abs(median_error) ) %>%
-    # arrange(median_abs_error) %>%
-    arrange(abs(median_error) ) %>%
-    group_by(onset_offset) %>%
-    gt() %>%
-    # gt(groupname_col = "onset_offset") %>%
-    fmt_number(
-        columns = everything(),
-        decimals = 2
-        ) %>%
-    cols_align(
-        align = "center",
-        columns = everything()
-        ) %>%
-    cols_align(
-        align = "left",
-        columns = method
-        ) %>%
-    cols_label(
-        method = "",
-        # mean_error = "Bias_mean",
-        median_error = "Bias",
-        median_abs_error = "MAE",
-        rmse = "RMSE",
-        variance = "Variance",
-        mad = "MAD"
-        ) %>%
-    tab_options(
-        table.align = "center",
-        # table.width = "75%",
-        table.font.size = 12
-        ) %>%
-    as_latex()
-```
-
-\newpage
-
--->
-
 # Discussion
 
-In brief, our results show that the model-based approach we introduced outperforms conventional methods such as cluster-based inference in identifying the onset and offset of M/EEG effects.
-Performance was assessed both on simulated data--allowing us to evaluate the method's ability to recover ground-truth onset and offset values--and on actual MEG data--allowing us to assess its sensitivity to realistic data properties (subset similarity).
+In brief, our results show that the model-based approach we introduced outperforms conventional cluster-based methods in identifying the onset and offset of M/EEG effects.
+We first assessed the performance of this approach on simulated data allowing us to evaluate the method's ability to recover ground-truth onset and offset values.
+We then assessed its performance on actual MEG data, allowing us to assess its sensitivity to realistic data properties (subset similarity).
 Together, these results highlight desirable properties for any method aiming to precisely and reliably estimate the onset and offset of M/EEG effects: it should i) recover true onsets and offsets in simulation (good asymptotic behaviour), ii) identify clusters that are interpretable and consistent in empirical data, and iii) show sensitivity to subtle changes in the data.
@@
 
-**I am not sure what to do with this paragraph (not sure how to discuss these results/observations):** Interestingly, the TFCE method performed worse than the traditional cluster-sum approach, consistent with the predictions of @rousselet_using_2025 based on the original findings of @smith2009.
+Interestingly, the TFCE method performed worse than the traditional cluster-sum approach, consistent with the predictions of @rousselet_using_2025 based on the original findings of @smith2009.
 We also found a striking overlap in the clusters identified by the `Holm` and `TFCE` procedures (cf.
@@
 This convergence warrants further methodological work.
+
+<!--
+
 We chose not to include the cluster-depth algorithm [@frossard2022], as prior work already shows its performance to be inferior to cluster-mass approaches in similar contexts [@rousselet_using_2025].
 
-A critical consideration for any model-based approach is that the model must adequately capture the underlying data-generating process.
+To make this paragraph more concrete, can we reflect on our own method in this regard?
+
+-->
+
+A critical requirement for any model-based approach is that the model must adequately capture the underlying data-generating process.
 Misspecified models are likely to produce biased or unreliable onset/offset estimates.
 This underscores the importance of thorough model diagnostics, including posterior predictive checks, fit assessments, and model comparison [@gelman2020].
-As with any inferential modelling framework, transparency in assumptions and rigorous validation procedures are indispensable to ensure replicability and generalisability of the results.
-
-Another important methodological consideration concerns the selection of model hyperparameters, such as the threshold for posterior odds and the number of basis functions.
+An important and related methodological consideration concerns the selection of model hyperparameters, such as the number of basis functions and the threshold for posterior odds.
 Although our simulations suggest that these parameters influence the precision and reliability of onset and offset estimates, optimal values may vary depending on the signal's temporal dynamics and signal-to-noise characteristics.
@@
     quiet = TRUE
-    # progress = FALSE,
-    # errors = "ignored"
     )
@@
 
-```{r fig-sim-timegen, eval = FALSE, echo = FALSE, dpi = 300, out.width = "75%", fig.cap = "Exemplary (simulated) group-level average cross-temporal generalisation matrix of decoding performance (ROC AUC)."}
-# number of participants
-n_participants <- 10
-
-# number of timepoints
-n_timepoints <- 60 # (1.2s at 50Hz)
-timepoints <- seq(from = -0.2, to = 1, length.out = n_timepoints)
-
-# smooth onset and offset parameters
-decode_onset <- 0.2
-decode_offset <- 0.3
-
-# smoothness of onset/offset
-sigma_smooth <- 0.05
-
-# generating participant-specific cross-temporal generalisation matrices
-timegen_data <- data.frame()
-
-# for each participant
-for (p in 1:n_participants) {
-
-    # adding variability in peak amplitude and width per participant
-    peak_variability <- runif(1, 30, 50) # random peak strength (higher = stronger decoding)
-    smooth_variability <- runif(1, 0.04, 0.06) # random smoothness variation
-    onset_variability <- runif(1, -0.05, 0.05) # small jitter in onset time
-    offset_variability <- runif(1, -0.05, 0.05) # small jitter in offset time
-
-    # computing smooth transition function (soft onset and offset with participant variation)
-    decode_window <- outer(timepoints, timepoints, function (t1, t2) {
-
-        # smooth onset
-        onset_factor <- 1 /
-            (1 + exp(-(t1 - (decode_onset + onset_variability) ) / sigma_smooth) )
-
-        # smooth offset
-        offset_factor <- 1 /
-            (1 + exp((t1 - (decode_offset + offset_variability) ) / sigma_smooth) )
-
-        # diagonal Gaussian
-        onset_factor * offset_factor * exp(-((t1 - t2)^2) / (2 * smooth_variability^2) )
-
-    })
-
-    # defining smooth alpha and beta parameters for Beta distribution
-    base_alpha <- 10
-    peak_alpha <- peak_variability
-    alpha_matrix <- base_alpha + peak_alpha * decode_window
-    beta_matrix <- base_alpha
-
-    # generating AUC values using the Beta distribution
-    auc_matrix_beta <- matrix(
-        rbeta(n_timepoints^2, shape1 = alpha_matrix, shape2 = beta_matrix),
-        nrow = n_timepoints, ncol = n_timepoints
-        )
-
-    # converting to long format for plotting
-    temp_data <- expand.grid(
-        train_time = timepoints, test_time = timepoints
-        ) %>%
-        mutate(auc = as.vector(auc_matrix_beta), participant = p)
-
-    # storing results
-    timegen_data <- bind_rows(timegen_data, temp_data)
-
-}
-
-# plotting the cross-temporal generalisation matrix
-timegen_data %>%
-    summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>%
-    ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
-    geom_tile() +
-    geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
-    geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
-    geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
-    scale_x_continuous(expand = c(0, 0) ) +
-    scale_y_continuous(expand = c(0, 0) ) + 
-    coord_fixed() +
-    scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC") +
-    labs(x = "Testing time (s)", y = "Training time (s)")
-```
-
 ```{r mne-timegen, eval = FALSE, echo = FALSE}
@@
 
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt}
-
-To model cross-temporal generalisation matrices of decoding performance (ROC AUC), we extended our initial BGAM to take into account the bivariate temporal distribution of AUC values, thus producing naturally smoothed estimates (timecourses) of AUC values and posterior probabilities.
-This model can be written as follows:
+<!--
 
@@
 where $\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|$ is the Euclidean distance between the covariate $\mathbf{s}_{i}$ and the knot location $\mathbf{k}_{m}$.
-We fitted this model using `brms` and the `t2()` tensor product smooth constructor with full penalties [@pedersen_hierarchical_2019].
-We ran eight MCMCs to approximate the posterior distribution, including each 5000 iterations and a warmup of 2000 iterations, yielding a total of $8 \times (5000-2000) = 24000$ posterior samples to be used for inference.
-
-```{r gam-timegen, eval = TRUE, echo = TRUE, results = "hide"}
+
+$$
+\begin{aligned}
+\text{AUC}_{i} &\sim \mathrm{Beta} \left(\mu_{i}, \phi\right)\\
+\operatorname{logit} \left(\mu_{i} \right) &= \alpha + \sum_{k=1}^K b_{k} B_{k}\left(\text{train}_{i}, \text{test}_{i} \right)
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+\text{AUC}_{i} &\sim \mathrm{Beta} \left(\mu_{i}, \phi\right)\\
+\operatorname{logit} \left(\mu_{i} \right) &= \alpha + f \left(\text{train}_{i}, \text{test}_{i} \right)\\
+f \left(\text{train}_{i}, \text{test}_{i} \right) &\equiv \sum_{k=1}^K b_{k} B_{k}\left(\text{train}_{i}, \text{test}_{i} \right)
+\end{aligned}
+$$
+
+-->
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+To model cross-temporal generalisation matrices of decoding performance, we extended our initial BGAM to take into account the bivariate temporal distribution of AUC values, thus producing naturally smoothed estimates (timecourses) of AUC values and posterior odds.
+This model can be written as follows:
+
+$$
+\begin{aligned}
+\text{AUC}_{i} &\sim \mathrm{Beta} \left(\mu_{i}, \phi\right)\\
+\operatorname{logit} \left(\mu_{i} \right) &= \alpha + f \left(\text{train}_{i}, \text{test}_{i} \right)\\
+f \left(\text{train}_{i}, \text{test}_{i} \right) &\equiv \sum_{k=1}^K b_{k} B_{k}\left(\text{train}_{i}, \text{test}_{i} \right)
+\end{aligned}
+$$
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+where AUC values are assumed to follow a $\mathrm{Beta}$ distribution, parametrised by a mean $\mu_{i}$ and a precision parameter $\phi$.
+The mean $\mu_{i}$ is linked to the predictors through a logit link function.
+The smooth function $f(\text{train}_{i}, \text{test}_{i})$ represents a two-dimensional surface defined over training and testing times, which captures how decoding performance varies across the temporal generalisation matrix.
+This surface is approximated by a linear combination of $K$ basis functions $B_{k}(\cdot, \cdot)$, each weighted by a coefficient $b_{k}$.
+The basis functions are constructed using a tensor product of univariate splines (here thin-plate splines) applied to the training and testing time dimensions [@wood2003; @wood2017].
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt}
+
+We fitted this model using `brms` and the `t2()` tensor product smooth constructor with full penalties [@pedersen_hierarchical_2019; @wood2017].
+We ran eight MCMCs to approximate the posterior distribution, including each 5000 iterations and a warmup of 1000 iterations, yielding a total of $8 \times (5000-1000) = 32000$ posterior samples to be used for inference.
+
+```{r gam-timegen, eval = TRUE, echo = FALSE, results = "hide"}
 # fitting a GAM with two temporal dimensions
@@
 
+```{r gam-timegen-show, eval = FALSE, echo = TRUE, results = "hide"}
+# fitting a GAM with two temporal dimensions
+timegen_gam <- brm(
+    # 2D thin-plate spline (tp) with full penalties
+    auc ~ t2(train_time, test_time, bs = "tp", k = 30, full = TRUE),
+    data = timegen_data,
+    family = Beta(),
+    warmup = 1000,
+    iter = 5000,
+    chains = 8,
+    cores = 8,
+    control = list(adapt_delta = 0.95, max_treedepth = 15)
+    )
+```
+
 ```{r fig-gam-timegen-post-preds, eval = TRUE, echo = FALSE, cache = FALSE, fig.width = 9, fig.asp = 0.5, fig.cap = "Predicted AUC values with threshold (left) and posterior odds of decoding accuracy being above chance (right) according to the bivariate BGAM."}
@@
         ndraws = 1e2,
-        # ndraws = 1e3,
         resolution = n_distinct(data$train_time)
@@
         mutate(prob_ratio = pmax(prob_ratio, 1 / ndraws(model) ) )
-
-    # plotting cross-temporal generalisation matrix
-    # data_plot <- data %>%
-    #     summarise(auc = mean(auc), .by = c(train_time, test_time) ) %>%
-    #     ggplot(aes(x = train_time, y = test_time, fill = auc) ) +
-    #     geom_tile() +
-    #     geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5) +
-    #     geom_hline(yintercept = 0, linetype = 2, linewidth = 0.5) +
-    #     geom_abline(slope = 1, intercept = 0, linetype = 2, linewidth = 0.5) +
-    #     scale_x_continuous(expand = c(0, 0) ) +
-    #     scale_y_continuous(expand = c(0, 0) ) + 
-    #     coord_fixed() +
-    #     scale_fill_scico(palette = "vik", midpoint = 0.5, name = "AUC", limits = fill_limits) +
-    #     labs(x = "Testing time (s)", y = "Training time (s)")
 
@@
 
-# computing the SESOI as the baseline SD
-# baseline_sd <- timegen_data %>%
-#     filter(train_time < 0 & test_time < 0) %>%
-#     summarise(sd(auc) ) %>%
-#     pull()
-
 # computing the SESOI as the baseline upper 99% quantile
@@
     data = timegen_data,
-    # chance_level = 0.5,
-    # sesoi = baseline_sd,
     chance_level = baseline_upper_quantile,
     sesoi = 0,
-    # thresholds = c(20, 50, 100)
     thresholds = 10
@@
 @fig-gam-timegen-post-preds shows the predictions from the model (left) superimposed with the identified cluster as defined by thresholding the posterior odds (right).
-Notably, this model could be extended to a multilevel bivariate GAM via `t2(train_time, test_time, participant, bs = c("tp", "tp", "re"), m = 2, full = TRUE)` and could be generalised to account for both spatial (`x1` and `x2`) and temporal (`time`) dimensions with formulas such as `te(x, y, time, d = c(2, 1) )`.
-The performance of such models should be assessed in future work.
-
-<!--
-
-\newpage
-
-# Alternative to GAMs: Approximate Gaussian Process regression
-
-A Gaussian process (GP) is a stochastic process that defines the distribution over a collection of random variables indexed by a continuous variable, that is $\{f(t): t \in \mathcal{T}\}$ for some index set $\mathcal{T}$ [@riutort-mayol_practical_2023; @rasmussen2005].
-Whereas Bayesian linear regression outputs a distribution over the parameters of some predefined parametric model, the GP approach, in contrast, is a non-parametric approach, in that it finds a distribution over the possible functions that are consistent with the observed data.
-However, "nonparametric" does not mean there aren't parameters, it means that there are infinitely many parameters.
-
-From [brms documentation](https://www.rdocumentation.org/packages/brms/versions/2.22.0/topics/gp): A GP is a stochastic process, which describes the relation between one or more predictors $x=\left(x_{1}, \ldots, x_{d}\right)$ and a response $f(x)$, where $d$ is the number of predictors.
-A GP is the generalization of the multivariate normal distribution to an infinite number of dimensions.
-Thus, it can be interpreted as a prior over functions.
-The values of $f()$ at any finite set of locations are jointly multivariate normal, with a covariance matrix defined by the covariance kernel $k_p\left(x_i, x_j\right)$, where $p$ is the vector of parameters of the GP:
-
-$$
-\left(f\left(x_{1}\right), \ldots f\left(x_{n}\right) \sim \operatorname{MVN}\left(0, \left(k_p\left(x_{i}, x_{j}\right)\right)_{i, j=1}^{n}\right)\right.
-$$
-
-The smoothness and general behaviour of the function $f$ depends only on the choice of covariance kernel, which ensures that values that are close together in the input space will be mapped to similar output values...
-
-From this perspective, $f$ is a realisation of an infinite dimensional normal distribution:
-
-$$
-f \sim \mathrm{Normal}(0, C(\lambda))
-$$
-
-where $C$ is a covariance kernel with hyperparameters $\lambda$ that defines the covariance between two function values $f\left(t_1\right)$ and $f\left(t_2\right)$ for two time points $t_1$ and $t_2$ [@rasmussen2005].
-Similar to the different choices of the basis function for splines, different choices of the covariance kernel lead to different GPs.
-In this article, we consider the squared-exponential (a.k.a.
-radial basis function) kernel, which computes the squared distance between points and converts it into a measure of similarity.
-It is defined as:
-
-$$
-C(\lambda) := C\left(t_1, t_2, \sigma, \gamma\right) := \sigma^2 \exp \left(-\frac{||t_1-t_2||^{2}}{2 \gamma^2}\right)
-$$
-
-with hyperparameters $\lambda = (\sigma, \gamma)$, expressing the overall scale of GP and the length-scale, respectively [@rasmussen2005].
-The advantages of this kernel are that it is computationally efficient and (infinitely) smooth making it a reasonable choice for the purposes of the present article.
-Here again, $\lambda$ hyperparameters are estimated from the data, along with all other model parameters.
-
-Taken from <https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10c-Gaussian-processes.html>: For a given vector $\mathbf{x}$, we can use the kernel to construct finite multi-variate normal distribution associated with it like so:
-
-$$
-\mathbf{x} \mapsto_{G P} \operatorname{MVNormal}(m(\mathbf{x}), k(\mathbf{x}, \mathbf{x}))
-$$
-
-where $m$ is a function that specifies the mean for the distribution associated with $\mathbf{x}$.
-This mapping is essentially the Gaussian process: a systematic association of vectors of arbitrary length with a suitable multi-variate normal distribution.
-
-Low-rank approximate Gaussian processes are of main interest in machine learning and statistics due to the high computational demands of exact Gaussian process models [@riutort-mayol_practical_2023]...
-
-```{r gp, echo = FALSE, eval = FALSE, message = FALSE}
-# See also <https://jtimonen.github.io/lgpr-usage/articles/math.html>...
-# Now we fit the GP regression model...
-# computing eeg_diff and averaging across participants
-# group_df <- raw_df %>%
-#     group_by(participant, time) %>%
-#     pivot_wider(names_from = condition, values_from = eeg, values_fn = mean) %>%
-#     mutate(eeg_diff = cond2 - cond1) %>%
-#     summarise(eeg_diff = mean(eeg_diff) ) %>%
-#     ungroup()
-
-# see https://betanalpha.github.io/assets/case_studies/gp_part3/part3.html
-# and https://betanalpha.github.io/assets/case_studies/gaussian_processes.html
-gp_priors <- c(
-    set_prior("normal(0, 1)", class = "Intercept"),
-    set_prior("normal(0, 1)", class = "b"),
-    set_prior("exponential(1)", class = "sigma"),
-    set_prior("exponential(1)", class = "sdgp")
-    )
-
-gp_model <- brm(
-    # k refers to the number of basis functions for
-    # computing Hilbert-space approximate GPs
-    # if k = NA (default), exact GPs are computed
-    # eeg ~ gp(time, by = condition),
-    # eeg ~ condition + gp(time, by = condition, k = 20, cov = "exp_quad"),
-    eeg ~ condition + gp(time, by = condition, k = 20, cov = "matern32"),
-    data = ppt_df,
-    family = gaussian(),
-    # prior = gp_priors,
-    control = list(adapt_delta = 0.99, max_treedepth = 15),
-    iter = 2000,
-    chains = 4,
-    cores = 4,
-    file = "models/gp.rds"
-    )
-```
-
-```{r meta-gp, echo = FALSE, eval = FALSE, message = FALSE}
-# fitting the GP
-meta_gp <- brm(
-    # using by-participant SD of ERPs across trials
-    eeg_mean | se(eeg_sd) ~
-        condition + gp(time, k = 20, by = condition) +
-        (1 | participant),
-    data = summary_df,
-    family = gaussian(),
-    control = list(adapt_delta = 0.99),
-    iter = 5000,
-    chains = 4,
-    cores = 4,
-    file = "models/meta_gp.rds"
-    )
-```
-
--->
+Notably, this model could be extended to a multilevel bivariate GAM via `t2(train_time, test_time, participant, bs = c("tp", "tp", "re"), m = 2, full = TRUE)` and could be generalised to account for both spatial (`x` and `y`) and temporal (`time`) dimensions with formulas such as `te(x, y, time, d = c(2, 1) )`.
 
@@
 
-# pp_check(gam, ndraws = 100, type = "ecdf_overlay") + 
-#     labs(x = "EEG signal", y = "Cumulative density") + ggtitle(paste("ECDF overlay (k =", k_value, ")") ) + theme(legend.position = "none")
-# pp_check(gam, ndraws = 200, type = "stat_2d", stat = c("median", "mad") ) + 
-#     labs(x = "Mean", y = "SD") + ggtitle(paste("Mean vs. SD (k =", k_value, ")") )
-
 # arranging plots using patchwork
@@
 
-# some posterior predictive checks
+# some more posterior predictive checks
 # pp_check(gam, prefix = "ppd")
@@
     mutate(k = str_remove_all(string = k, pattern = '^"|"$') ) %>%
-    # mutate(across(.cols = where(is.numeric), .fns = ~round(.x, 2) ) ) %>%
     gt() %>%
@@
         ) %>%
-    # cols_label(
-    #     # elpd_diff = "\\$Delta$ diff"
-    #     ) %>%
     tab_options(
@@
 
-\newpage
+\clearpage
+\thispagestyle{empty}
+\null
 
@@
 It is also possible to call functions from the `neurogam R` package (available at <https://github.com/lnalborczyk/neurogam>) which come with sensible defaults.
+
+```{r, eval = TRUE, echo = FALSE}
+library(neurogam)
+```
 

### 2025-08-27 manuscript.tex
pushing final version of manuscript

@@
 
-\hypertarget{toc}{}
-\tableofcontents
-\newpage
-\section[Introduction]{Precise temporal localisation of M/EEG effects
-with Bayesian generalised additive multilevel models}
-
 \setcounter{secnumdepth}{3}
@@
 The probability of obtaining at least one false positive in an ensemble
-(family) of \(m\) tests (i.e., the FWER) is computed as
+(family) of \(m\) independent tests (i.e., the FWER) is computed as
 \(1-\left(1-\alpha\right)^{m}\) (for \(m=10\) independent tests and
@@
 \(\hat{R}\) (\citeproc{ref-gabry2019}{Gabry et al., 2019};
-\citeproc{ref-gelman2020}{Gelman et al., 2020}). The \texttt{brms}
+\citeproc{ref-gelman2020}{Gelman et al., 2020};
+\citeproc{ref-vehtari2021}{Vehtari et al., 2021}). The \texttt{brms}
 package uses the same syntax as the \texttt{R} package \texttt{mgcv} v
@@
 
-However, the previous model only included constant (fixed) effects, thus
-not properly accounting for between-participant variability. We next
-fitted a multilevel version of the BGAM (BGAMM, for an introduction to
-Bayesian multilevel models in \texttt{brms}, see
+However, the model whose predictions are depicted in
+Figure~\ref{fig-plot-post-slope} only included constant (fixed) effects,
+thus not properly accounting for between-participant variability. We
+next fitted a multilevel version of the BGAM (BGAMM, for an introduction
+to Bayesian multilevel models in \texttt{brms}, see
 \citeproc{ref-nalborczyk2019}{Nalborczyk et al., 2019}) including a
@@
 \midrule\addlinespace[2.5pt]
-BGAM (our method) & 0.11 & 2.76 & 0.11 & 28.78 \\ 
+BGAM (our method) & {\bfseries 0.11} & {\bfseries 2.76} & {\bfseries 0.11} & {\bfseries 28.78} \\ 
 Change point & 5.17 & 6.51 & 5.17 & 33.58 \\ 
@@
 \midrule\addlinespace[2.5pt]
-BGAM (our method) & -2.35 & 3.46 & 2.35 & 53.63 \\ 
+BGAM (our method) & {\bfseries -2.35} & {\bfseries 3.46} & {\bfseries 2.35} & 53.63 \\ 
 Cluster mass & -8.64 & 9.44 & 8.64 & 193.63 \\ 
-Change point & -9.28 & 9.82 & 9.28 & 33.03 \\ 
+Change point & -9.28 & 9.82 & 9.28 & {\bfseries 33.03} \\ 
 Holm & -31.86 & 34.03 & 31.86 & 2,764.12 \\ 
@@
 In brief, our results show that the model-based approach we introduced
-outperforms conventional methods such as cluster-based inference in
-identifying the onset and offset of M/EEG effects. Performance was
-assessed both on simulated data--allowing us to evaluate the method's
-ability to recover ground-truth onset and offset values--and on actual
-MEG data--allowing us to assess its sensitivity to realistic data
-properties (subset similarity). Together, these results highlight
-desirable properties for any method aiming to precisely and reliably
-estimate the onset and offset of M/EEG effects: it should i) recover
-true onsets and offsets in simulation (good asymptotic behaviour), ii)
-identify clusters that are interpretable and consistent in empirical
-data, and iii) show sensitivity to subtle changes in the data. Our
-approach meets all three of these desiderata.
+outperforms conventional cluster-based methods in identifying the onset
+and offset of M/EEG effects. We first assessed the performance of this
+approach on simulated data allowing us to evaluate the method's ability
+to recover ground-truth onset and offset values. We then assessed its
+performance on actual MEG data, allowing us to assess its sensitivity to
+realistic data properties (subset similarity). Together, these results
+highlight desirable properties for any method aiming to precisely and
+reliably estimate the onset and offset of M/EEG effects: it should i)
+recover true onsets and offsets in simulation (good asymptotic
+behaviour), ii) identify clusters that are interpretable and consistent
+in empirical data, and iii) show sensitivity to subtle changes in the
+data. Our approach meets all three of these desiderata.
 
@@
 
-\textbf{I am not sure what to do with this paragraph (not sure how to
-discuss these results/observations):} Interestingly, the TFCE method
-performed worse than the traditional cluster-sum approach, consistent
-with the predictions of Rousselet
+Interestingly, the TFCE method performed worse than the traditional
+cluster-sum approach, consistent with the predictions of Rousselet
 (\citeproc{ref-rousselet_using_2025}{2025}) based on the original
@@
 high-intensity points. This convergence warrants further methodological
-work. We chose not to include the cluster-depth algorithm
-(\citeproc{ref-frossard2022}{Frossard \& Renaud, 2022}), as prior work
-already shows its performance to be inferior to cluster-mass approaches
-in similar contexts (\citeproc{ref-rousselet_using_2025}{Rousselet,
-2025}).
-
-A critical consideration for any model-based approach is that the model
+work.
+
+A critical requirement for any model-based approach is that the model
 must adequately capture the underlying data-generating process.
@@
 assessments, and model comparison (\citeproc{ref-gelman2020}{Gelman et
-al., 2020}). As with any inferential modelling framework, transparency
-in assumptions and rigorous validation procedures are indispensable to
-ensure replicability and generalisability of the results.
-
-Another important methodological consideration concerns the selection of
-model hyperparameters, such as the threshold for posterior odds and the
-number of basis functions. Although our simulations suggest that these
-parameters influence the precision and reliability of onset and offset
-estimates, optimal values may vary depending on the signal's temporal
-dynamics and signal-to-noise characteristics. Future work could explore
-principled approaches to hyperparameter tuning, including
-cross-validation or fully Bayesian model selection using tools such as
-leave-one-out cross-validation (LOO-CV) or Bayes factors
-(\citeproc{ref-gelman2020}{Gelman et al., 2020}). We provide initial
-guidance in \hyperref[apx-basis]{Appendix~B} and advocate for future
-development of adaptive heuristics to support flexible yet parsimonious
-model specification.
+al., 2020}). An important and related methodological consideration
+concerns the selection of model hyperparameters, such as the number of
+basis functions and the threshold for posterior odds. Although our
+simulations suggest that these parameters influence the precision and
+reliability of onset and offset estimates, optimal values may vary
+depending on the signal's temporal dynamics and signal-to-noise
+characteristics. Future work could explore principled approaches to
+hyperparameter tuning, including cross-validation or fully Bayesian
+model selection using tools such as leave-one-out cross-validation
+(LOO-CV) or Bayes factors (\citeproc{ref-gelman2020}{Gelman et al.,
+2020}). We provide initial guidance in \hyperref[apx-basis]{Appendix~B}
+and advocate for future development of adaptive heuristics to support
+flexible yet parsimonious model specification.
 
@@
 \citeproc{ref-brms2021}{2021}), doParallel v. 1.0.17
-(\citeproc{ref-doParallel}{Corporation \& Weston, 2022}), easystats v.
-0.7.4 (\citeproc{ref-easystats}{Lüdecke et al., 2022}), foreach v. 1.5.2
-(\citeproc{ref-foreach}{Microsoft \& Weston, 2022}), furrr v. 0.3.1
-(\citeproc{ref-furrr}{Vaughan \& Dancho, 2022}), future v. 1.58.0
+(\citeproc{ref-doParallel}{Corporation \& Weston, 2022}), foreach v.
+1.5.2 (\citeproc{ref-foreach}{Microsoft \& Weston, 2022}), furrr v.
+0.3.1 (\citeproc{ref-furrr}{Vaughan \& Dancho, 2022}), future v. 1.58.0
 (\citeproc{ref-future}{Bengtsson, 2021}), ggpubr v. 0.6.0
@@
 
-\bibitem[\citeproctext]{ref-easystats}
-Lüdecke, D., Ben-Shachar, M. S., Patil, I., Wiernik, B. M., Bacher, E.,
-Thériault, R., \& Makowski, D. (2022). {easystats}: Framework for easy
-statistical modeling, visualization, and reporting. \emph{CRAN}.
-\url{https://doi.org/10.32614/CRAN.package.easystats}
-
 \bibitem[\citeproctext]{ref-maris2011}
@@
 \url{https://CRAN.R-project.org/package=furrr}
+
+\bibitem[\citeproctext]{ref-vehtari2021}
+Vehtari, A., Gelman, A., Simpson, D., Carpenter, B., \& Bürkner, P.-C.
+(2021). Rank-normalization, folding, and localization: An improved Rˆ
+for assessing convergence of MCMC (with discussion). \emph{Bayesian
+Analysis}, \emph{16}(2). \url{https://doi.org/10.1214/20-ba1221}
 
@@
 
-To model cross-temporal generalisation matrices of decoding performance
-(ROC AUC), we extended our initial BGAM to take into account the
-bivariate temporal distribution of AUC values, thus producing naturally
-smoothed estimates (timecourses) of AUC values and posterior
-probabilities. This model can be written as follows:
+To model cross-temporal generalisation matrices of decoding performance,
+we extended our initial BGAM to take into account the bivariate temporal
+distribution of AUC values, thus producing naturally smoothed estimates
+(timecourses) of AUC values and posterior odds. This model can be
+written as follows:
 
@@
 \begin{aligned}
-\text{AUC}_{i} &\sim \mathrm{Beta}(\mu_{i}, \phi)\\
-g(\mu_{i}) &= f \left(\text{train}_{i}, \text{test}_{i} \right)\\
+\text{AUC}_{i} &\sim \mathrm{Beta} \left(\mu_{i}, \phi\right)\\
+\operatorname{logit} \left(\mu_{i} \right) &= \alpha + f \left(\text{train}_{i}, \text{test}_{i} \right)\\
+f \left(\text{train}_{i}, \text{test}_{i} \right) &\equiv \sum_{k=1}^K b_{k} B_{k}\left(\text{train}_{i}, \text{test}_{i} \right)
 \end{aligned}
@@
 
-where we assume that AUC values come from a \(\mathrm{Beta}\)
-distribution with two parameters \(\mu\) and \(\phi\). We can think of
-\(f \left(\text{train}_{i}, \text{test}_{i} \right)\) as a surface (a
-smooth function of two variables) that we can model using a
-2-dimensional splines. Let
-\(\mathbf{s}_{i} = \left(\text{train}_{i}, \text{test}_{i} \right)\) be
-some pair of training and testing samples, and let
-\(\mathbf{k}_{m} = \left(\text{train}_{m}, \text{test}_{m} \right)\)
-denote the \(m^{\text{th}}\) knot in the domain of \(\text{train}_{i}\)
-and \(\text{test}_{i}\). We can then express the smooth function as:
-
-\[
-f \left(\text{train}_{i}, \text{test}_{i} \right) = \alpha + \sum_{m=1}^M \beta_{m} b_{m} \left(\tilde{s}_{i}, \tilde{k}_{m} \right)
-\]
+where AUC values are assumed to follow a \(\mathrm{Beta}\) distribution,
+parametrised by a mean \(\mu_{i}\) and a precision parameter \(\phi\).
+The mean \(\mu_{i}\) is linked to the predictors through a logit link
+function. The smooth function \(f(\text{train}_{i}, \text{test}_{i})\)
+represents a two-dimensional surface defined over training and testing
+times, which captures how decoding performance varies across the
+temporal generalisation matrix. This surface is approximated by a linear
+combination of \(K\) basis functions \(B_{k}(\cdot, \cdot)\), each
+weighted by a coefficient \(b_{k}\). The basis functions are constructed
+using a tensor product of univariate splines (here thin-plate splines)
+applied to the training and testing time dimensions
+(\citeproc{ref-wood2003}{Wood, 2003a}, \citeproc{ref-wood2017}{2017a}).
 
@@
 
-A popular bivariate basis function uses \emph{thin-plate splines}
-(\citeproc{ref-wood2003}{Wood, 2003a}). These splines are designed to
-interpolate and approximate smooth surfaces over two dimensions (hence
-the ``bivariate'' term). For \(d=2\) dimensions and \(l=2\) (smoothness
-penalty involving second order derivative):
-
-\[
-f \left(\tilde{s}_{i} \right) = \alpha + \beta_{1} x_{i} + \beta_{2} z_{i} +\sum_{m=1}^{M} \beta_{2+m} b_m\left(\tilde{s}_i, \tilde{k}_m\right)
-\]
-
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt}
-
-using the the radial basis function given by:
-
-\[
-b_m\left(\tilde{s}_i, \tilde{k}_m\right)=\left\|\tilde{s}_i-\tilde{k}_m\right\|^2 \log \left\|\tilde{s}_i-\tilde{k}_m\right\|
-\]
-
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt}
-
-where \(\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|\) is the Euclidean
-distance between the covariate \(\mathbf{s}_{i}\) and the knot location
-\(\mathbf{k}_{m}\). We fitted this model using \texttt{brms} and the
-\texttt{t2()} tensor product smooth constructor with full penalties
-(\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al.,
-2019}). We ran eight MCMCs to approximate the posterior distribution,
-including each 5000 iterations and a warmup of 2000 iterations, yielding
-a total of \(8 \times (5000-2000) = 24000\) posterior samples to be used
-for inference.
+We fitted this model using \texttt{brms} and the \texttt{t2()} tensor
+product smooth constructor with full penalties
+(\citeproc{ref-pedersen_hierarchical_2019}{E. J. Pedersen et al., 2019};
+\citeproc{ref-wood2017}{Wood, 2017a}). We ran eight MCMCs to approximate
+the posterior distribution, including each 5000 iterations and a warmup
+of 1000 iterations, yielding a total of \(8 \times (5000-1000) = 32000\)
+posterior samples to be used for inference.
 
@@
     \AttributeTok{data =}\NormalTok{ timegen\_data,}
-    \AttributeTok{family =}\NormalTok{ brms}\SpecialCharTok{::}\FunctionTok{Beta}\NormalTok{(),}
+    \AttributeTok{family =} \FunctionTok{Beta}\NormalTok{(),}
     \AttributeTok{warmup =} \DecValTok{1000}\NormalTok{,}
@@
     \AttributeTok{cores =} \DecValTok{8}\NormalTok{,}
-    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{15}\NormalTok{),}
-    \AttributeTok{backend =} \StringTok{"cmdstanr"}\NormalTok{,}
-    \AttributeTok{stan\_model\_args =} \FunctionTok{list}\NormalTok{(}\AttributeTok{stanc\_options =} \FunctionTok{list}\NormalTok{(}\StringTok{"O1"}\NormalTok{) ),}
-    \AttributeTok{file =} \StringTok{"models/timegen\_gam\_meg\_t2\_k30\_full.rds"}
+    \AttributeTok{control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{adapt\_delta =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{max\_treedepth =} \DecValTok{15}\NormalTok{)}
 \NormalTok{    )}
@@
 \texttt{t2(train\_time,\ test\_time,\ participant,\ bs\ =\ c("tp",\ "tp",\ "re"),\ m\ =\ 2,\ full\ =\ TRUE)}
-and could be generalised to account for both spatial (\texttt{x1} and
-\texttt{x2}) and temporal (\texttt{time}) dimensions with formulas such
-as \texttt{te(x,\ y,\ time,\ d\ =\ c(2,\ 1)\ )}. The performance of such
-models should be assessed in future work.
+and could be generalised to account for both spatial (\texttt{x} and
+\texttt{y}) and temporal (\texttt{time}) dimensions with formulas such
+as \texttt{te(x,\ y,\ time,\ d\ =\ c(2,\ 1)\ )}.
 

### 2025-08-27 code/old/analysing_sim_results.R

@@ -1,183 +1,0 @@
-###########################################################
-# Monte-Carlo simulation of onset/offset error properties #
-# Written by Ladislas Nalborczyk                          #
-# Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on February 13, 2025                       #
-###########################################################
-
-library(patchwork)
-library(tidyverse)
-library(ggrepel)
-
-# loading simulation results
-sim_results <- readRDS(file = "results/sim_results.rds")
-
-# identifying the best brms results
-# eps=0 and threshold = 10 or 20 seems pretty good...
-sim_results %>%
-    select(simulation_id:offset_brms) %>%
-    summarise(
-        MAE_onset = median(abs(onset_brms - true_onset) ),
-        variance_onset = var(onset_brms),
-        MAE_offset = median(abs(offset_brms - true_offset) ),
-        variance_offset = var(offset_brms),
-        .by = c(eps, threshold)
-        ) %>%
-    arrange(MAE_onset, MAE_offset) %>%
-    head(10)
-
-# plotting some results
-sim_results %>%
-    summarise(error_brms = mean(abs(onset_brms-true_onset) ), .by = c(eps, threshold) ) %>%
-    ggplot(aes(x = eps, y = threshold, fill = error_brms) ) +
-    geom_tile(show.legend = FALSE) +
-    geom_point(
-        data = sim_results %>%
-            summarise(error_brms = mean(abs(onset_brms-true_onset) ), .by = c(eps, threshold) ) %>%
-            dplyr::filter(error_brms == min(error_brms, na.rm = TRUE) ),
-        aes(x = eps, y = threshold),
-        color = "orangered", size = 2, shape = 4,
-        show.legend = FALSE
-        ) +
-    scale_x_continuous(expand = c(0, 0) ) +
-    scale_y_continuous(expand = c(0, 0) ) +
-    scale_fill_gradientn(colors = rev(met.brewer("Hokusai1") ) ) +
-    labs(
-        title = "Onset error",
-        x = "eps",
-        y = "Threshold",
-        fill = "Error"
-        )
-
-sim_results %>%
-    summarise(error_brms = mean(abs(offset_brms-true_offset) ), .by = c(eps, threshold) ) %>%
-    # arrange(error_brms) %>% head()
-    ggplot(aes(x = eps, y = threshold, fill = error_brms) ) +
-    geom_tile(show.legend = FALSE) +
[…]

### 2025-08-27 code/old/cluster_reticulate_test.R

@@ -1,25 +1,0 @@
-library(reticulate)
-
-# use_virtualenv("./reticulate", required = TRUE)
-py_require(packages = c("numpy", "mne"), python_version = "3.12.0")
-np <- import("numpy")
-mne <- import("mne")
-X <- np$random$randn(as.integer(10), as.integer(50) )
-mne$set_log_level("WARNING") # or "ERROR"
-
-results <- tryCatch({
-    mne$stats$spatio_temporal_cluster_1samp_test(
-        X = X,
-        out_type = "mask",
-        threshold = NULL,
-        n_permutations = as.integer(2^12),
-        n_jobs = as.integer(1),
-        verbose = TRUE
-        )
-    }, error = function(e) {
-    message("Error: ", e$message)
-    reticulate::py_last_error()
-    return(NULL)
-    })
-
-print(results)

### 2025-08-27 code/old/errors_simulation.R

@@ -1,462 +1,0 @@
-###########################################################
-# Monte-Carlo simulation of onset/offset error properties #
-# Written by Ladislas Nalborczyk                          #
-# Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on March 31, 2025                          #
-###########################################################
-
-# importing R packages
-library(changepoint)
-library(reticulate)
-library(tidyverse)
-library(tidybayes)
-library(brms)
-
-# importing Python modules
-# use_condaenv("/Users/ladislas/opt/anaconda3/envs/r-reticulate/bin/python", required = TRUE)
-# use_condaenv("r-reticulate", required = TRUE)
-use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
-np <- import("numpy")
-mne <- import("mne")
-
-# importing R version of Matlab code from Yeung et al. (2004)
-source("code/eeg_noise.R")
-
-# importing the ERP template with true onset = 160 ms, F=81, and max at F=126
-source("code/erp_template.R")
-
-# importing helper functions
-source("code/functions.R")
-
-# to use with the eeg_noise function
-meanpower <- unlist(read.table("code/meanpower.txt") )
-
-# defining the number of simulations to be performed
-nsims <- 100
-
-# defining simulation parameters
-n_trials <- 50 # number of trials
-n_ppt <- 20 # number of participants
-outvar <- 1 # noise variance
-srate <- 500 # sampling rate in Hz
-ronset <- seq(from = 150, to = 170, by = 2) # random onset for each participant
-
-# defining the true onset and offset in seconds
-true_onset <- 0.160
-true_offset <- 0.342
-
-# defining significance (alpha) thresholds
-significance_thresholds <- c(0.05, 0.01, 0.005, 0.001)
-
-# defining a function to generate a dataframe
-generate_data <- function (n_trials, n_ppt, outvar, srate, ronset) {
-    
-    for (P in 1:n_ppt) { # for each participant
-        
-        # get random onset
-        ponset <- sample(x = ronset, size = 1)
-        
-        # find starting point
-        st <- which(Xf == ponset)
-        
-        # pad vector
-        temp2 <- c(rep(0, st - 2), erp, rep(0, Nf - st - length(erp) + 2) )
-        
-        # initialising empty conditions
-        cond1 <- matrix(0, nrow = n_trials, ncol = Nf)
-        cond2 <- matrix(0, nrow = n_trials, ncol = Nf)
-        
[…]

### 2025-08-27 code/old/functions.R

@@ -1,173 +1,0 @@
-######################################################################################
-# Functions from https://github.com/GRousselet/onsetsim/blob/main/code/functions.R
-# Generate permutation distributions of t values for independent groups
-# cond1 and cond2 are Nt trials x Nf time frames matrices
-######################################################################################
-
-# Cluster correction
-# Form clusters using binary vector: pvals < alpha
-# clusters must be at least 2 time frames
-cluster.make <- function (x) {
-    
-    y <- rle(x)
-    cmap <- vector(mode = "numeric", length = 0)
-    nC <- length(y$values) # number of clusters
-    indx <- 0 # cluster counter
-    
-    for (CL in 1:nC) {
-        
-    if(y$values[CL] == 0 || y$lengths[CL] == 1){
-      val <- 0
-    } else {
-      indx <- indx + 1
-      val <- indx
-    }
-    cmap <- c(cmap, rep(val, y$lengths[CL]))
-    }
-    
-    cmap
-    
-}
-
-# ## Save sum for each cluster
-# # values = statistics (t values)
-# # cmap = output from cluster.make
-# cluster.sum <- function(values, cmap){
-#   csum <- vector(mode = "numeric", length = max(cmap))
-#   if(max(cmap)>0){
-#     for(CL in 1:max(cmap)){
-#       csum[CL] <- sum(values[cmap==CL])
-#     }
-#   } else {
-#     csum <- 0
-#   }
-#   csum
-# }
-# 
-# # Cluster test
-# # values = statistics (t values)
-# # cmap = output from cluster.make
-# # boot.th = bootstrap quantile of distribution of max cluster sums
-# cluster.test <- function(values, cmap, boot.th){
-#   csig <- vector(mode = "logical", length = length(cmap))
-#   if(max(cmap)>0){
-#     for(CL in 1:max(cmap)){
-#       csig[cmap==CL] <- sum(values[cmap==CL]) > boot.th
-#     }
-#   }
-#   csig
-# }
-
-sim.counter <- function(S, nsim, inc){
-  if(S == 1){
-    # print(paste(nsim,"iterations:",S))
-    cat(nsim,"iterations:",S)
-    #beep(2)
-  }
-  if(S %% inc == 0){
[…]

### 2025-08-27 code/old/model_examples.R

@@ -1,97 +1,0 @@
-# averaging across participants
-# ppt_df <- raw_df %>%
-#     group_by(participant, condition, time) %>%
-#     summarise(eeg = mean(eeg) ) %>%
-#     ungroup()
-
-# defining a contrast for condition
-# contrasts(ppt_df$condition) <- c(-0.5, 0.5)
-
-# fitting the BGAM
-# gam <- brm(
-#     # thin-plate regression splines with k-1 basis functions
-#     eeg ~ condition + s(time, bs = "tp", k = 20, by = condition),
-#     data = ppt_df,
-#     family = gaussian(),
-#     warmup = 2000,
-#     iter = 5000,
-#     chains = 8,
-#     cores = 8,
-#     file = "models/gam.rds"
-#     )
-# 
-# meta_gam <- brm(
-#     # using by-participant SD of ERPs across trials
-#     eeg_mean | se(eeg_sd) ~
-#         condition + s(time, bs = "tp", k = 20, by = condition) +
-#         (1 + condition | participant),
-#     data = summary_df,
-#     family = gaussian(),
-#     warmup = 2000,
-#     iter = 5000,
-#     chains = 8,
-#     cores = 8,
-#     file = "models/meta_gam.rds"
-#     )
-# 
-# full_meta_gamm <- brm(
-#     # using by-participant SD of ERPs across trials
-#     eeg_mean | se(eeg_sd) ~ 1 + condition +
-#         (1 + condition | participant) +
-#         s(time, by = condition, bs = "tp", k = 20) +
-#         # s(participant, by = condition, bs = "re", xt = list(bs = "tp") ),
-#         s(time, participant, by = condition, bs = "fs", xt = list(bs = "tp") ),
-#     data = summary_df,
-#     family = gaussian(),
-#     warmup = 2000,
-#     iter = 5000,
-#     chains = 8,
-#     cores = 8,
-#     control = list(adapt_delta = 0.95),
-#     file = "models/full_meta_gamm.rds"
-#     )
-
-# errors on the HPC cluster
-# Error in unserialize(socklist[[n]]) : error reading from connection
-# Calls: %dopar% ... recvOneData -> recvOneData.SOCKcluster -> unserialize
-# Execution halted
[…]

### 2025-08-27 code/old/split_half_reliability.R

@@ -1,750 +1,0 @@
-###########################################################
-# Split-half reliability of onset/offset estimates        #
-# Written by Ladislas Nalborczyk                          #
-# Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on April 2, 2025                           #
-###########################################################
-
-# importing R packages
-library(changepoint)
-library(reticulate)
-library(easystats)
-library(tidyverse)
-library(tidybayes)
-library(scales)
-library(brms)
-
-# defining the default ggplot2 theme
-theme_set(theme_light(base_size = 12, base_family = "Open Sans") )
-
-# importing Python modules
-use_condaenv("r-reticulate3", conda = "~/miniforge3/bin/conda", required = TRUE)
-np <- import("numpy")
-mne <- import("mne")
-
-# reshaping the decoding results
-# read.csv(file = "code/decoding_results.csv") %>%
-#     filter(crossval_strategy == "4") %>%
-#     select(participant, cv_fold = fold, time, auc = score) %>%
-#     mutate(cv_fold = cv_fold + 1) %>%
-#     mutate(participant = str_trunc(string = participant, width = 12, ellipsis = "") ) %>%
-#     mutate(participant = str_replace(string = participant, pattern = "tickertape", replacement = "participant") ) %>%
-#     data.frame() %>%
-#     filter(time >= -0.2) %>%
-#     write.csv(file = "code/decoding_results_reshaped.csv", row.names = FALSE)
-
-# importing the reshaped MEG decoding results
-raw_df <- read.csv(file = "code/decoding_results_reshaped.csv") %>%
-    # removing the last participant (to get a round number of participants)
-    mutate(participant_id = cur_group_id(), .by = participant) %>%
-    filter(participant_id < 33) %>%
-    select(-participant_id)
-
-# averaging across participants
-summary_df <- raw_df %>%
-    summarise(
-        auc_mean = mean(auc),
-        auc_sd = sd(auc),
-        auc_se = sd(auc) / sqrt(n() ),
-        .by = c(participant, time)
-        )
-
-# plotting the group-level average decoding performance
-# raw_df %>%
-#     summarise(
-#         auc_mean = mean(auc),
-#         auc_sd = sd(auc),
-#         auc_se = sd(auc) / sqrt(n() ),
-#         .by = time
-#         ) %>%
-#     ggplot(aes(x = time, y = auc_mean) ) +
[…]

### 2025-08-27 code/old/split_half_reliability_cluster.R

@@ -1,502 +1,0 @@
-###########################################################
-# Split-half reliability of onset/offset estimates        #
-# Written by Ladislas Nalborczyk                          #
-# Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on April 14, 2025                          #
-###########################################################
-
-# importing R packages
-library(changepoint)
-library(reticulate)
-library(tidybayes)
-library(dplyr)
-library(tidyr)
-library(brms)
-
-# for setting up the cluster
-library(doParallel)
-library(foreach)
-
-# timing the simulations
-# library(tictoc)
-# tic()
-
-# total number of total CPU cores available on the HPC node
-total_cores <- 64
-
-# number of cores per brms model
-cores_per_model <- 4
-
-# number of parallel simulations to run at once
-n_parallel <- total_cores / cores_per_model
-
-# registering parallel backend
-cl <- makeCluster(n_parallel)
-registerDoParallel(cl)
-
-# importing home-made helper functions
-source("code/utils.R")
-
-# importing the reshaped MEG decoding results
-raw_df <- read.csv(file = "data/decoding_results_reshaped.csv") |>
-    # removing the last participant (to get a round number of participants)
-    mutate(participant_id = cur_group_id(), .by = participant) |>
-    filter(participant_id < 33) |>
-    select(-participant_id)
-
-# averaging across participants
-summary_df <- raw_df |>
-    summarise(
-        auc_mean = mean(auc),
-        auc_sd = sd(auc),
-        auc_se = sd(auc) / sqrt(n() ),
-        .by = c(participant, time)
-        )
-
-# computing the smallest effect size of interest (SESOI)
-# as the SD of decoding performance during the baseline
-full_sesoi <- summary_df %>%
-    filter(time < 0) %>%
-    summarise(auc_mean = mean(auc_mean), .by = time) %>%
-    summarise(sesoi = sd(auc_mean) ) %>%
-    pull(sesoi)
-
-# getting unique participants
-unique_participants <- unique(raw_df$participant)
-
-# getting the number of participants
-n_participants <- as.numeric(length(unique_participants) )
-
-# defining the number of splits
-# n_splits <- 1000
-n_splits <- 32
-
-# number of posterior samples to use to compute the posterior probability ratio
-# n_post_samples <- 1e4
-
-# creating a dataframe where each participant appears in every split
-split_df <- crossing(participant = unique_participants, split_number = 1:n_splits)
-
[…]

### 2025-08-27 manuscript.qmd
pushing first public version of the manuscript

@@
 In brief, our results show that the model-based approach we introduced outperforms conventional cluster-based methods in identifying the onset and offset of M/EEG effects.
-We first assessed the performance of this approach on simulated data allowing us to evaluate the method's ability to recover ground-truth onset and offset values.
+We first assessed the performance of this approach on simulated data, allowing us to evaluate the method's ability to recover ground-truth onset and offset values.
 We then assessed its performance on actual MEG data, allowing us to assess its sensitivity to realistic data properties (subset similarity).
@@
 This convergence warrants further methodological work.
-
-<!--
-
-We chose not to include the cluster-depth algorithm [@frossard2022], as prior work already shows its performance to be inferior to cluster-mass approaches in similar contexts [@rousselet_using_2025].
-
-To make this paragraph more concrete, can we reflect on our own method in this regard?
-
--->
 
@@
 ```
-
-<!--
-
-$$
-\begin{aligned}
-\text{AUC}_{i} &\sim \mathrm{Beta}(\mu_{i}, \phi)\\
-g(\mu_{i}) &= f \left(\text{train}_{i}, \text{test}_{i} \right)\\
-\end{aligned}
-$$
-
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt}
-
-where we assume that AUC values come from a $\mathrm{Beta}$ distribution with two parameters $\mu$ and $\phi$.
-We can think of $f \left(\text{train}_{i}, \text{test}_{i} \right)$ as a surface (a smooth function of two variables) that we can model using a 2-dimensional splines.
-Let $\mathbf{s}_{i} = \left(\text{train}_{i}, \text{test}_{i} \right)$ be some pair of training and testing samples, and let $\mathbf{k}_{m} = \left(\text{train}_{m}, \text{test}_{m} \right)$ denote the $m^{\text{th}}$ knot in the domain of $\text{train}_{i}$ and $\text{test}_{i}$.
-We can then express the smooth function as:
-
-$$
-f \left(\text{train}_{i}, \text{test}_{i} \right) = \alpha + \sum_{m=1}^M \beta_{m} b_{m} \left(\tilde{s}_{i}, \tilde{k}_{m} \right)
-$$
-
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt}
-
-A popular bivariate basis function uses *thin-plate splines* [@wood2003].
-These splines are designed to interpolate and approximate smooth surfaces over two dimensions (hence the "bivariate" term).
-For $d=2$ dimensions and $l=2$ (smoothness penalty involving second order derivative):
-
-$$
-f \left(\tilde{s}_{i} \right) = \alpha + \beta_{1} x_{i} + \beta_{2} z_{i} +\sum_{m=1}^{M} \beta_{2+m} b_m\left(\tilde{s}_i, \tilde{k}_m\right)
-$$
-
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt}
-
-using the the radial basis function given by:
-
-$$
-b_m\left(\tilde{s}_i, \tilde{k}_m\right)=\left\|\tilde{s}_i-\tilde{k}_m\right\|^2 \log \left\|\tilde{s}_i-\tilde{k}_m\right\|
-$$
-
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt}
-
-where $\left\|\mathbf{s}_i-\mathbf{k}_{m}\right\|$ is the Euclidean distance between the covariate $\mathbf{s}_{i}$ and the knot location $\mathbf{k}_{m}$.
-
-$$
-\begin{aligned}
-\text{AUC}_{i} &\sim \mathrm{Beta} \left(\mu_{i}, \phi\right)\\
-\operatorname{logit} \left(\mu_{i} \right) &= \alpha + \sum_{k=1}^K b_{k} B_{k}\left(\text{train}_{i}, \text{test}_{i} \right)
-\end{aligned}
-$$
-
-$$
-\begin{aligned}
-\text{AUC}_{i} &\sim \mathrm{Beta} \left(\mu_{i}, \phi\right)\\
-\operatorname{logit} \left(\mu_{i} \right) &= \alpha + f \left(\text{train}_{i}, \text{test}_{i} \right)\\
-f \left(\text{train}_{i}, \text{test}_{i} \right) &\equiv \sum_{k=1}^K b_{k} B_{k}\left(\text{train}_{i}, \text{test}_{i} \right)
-\end{aligned}
-$$
-
--->
 

### 2025-08-27 manuscript.tex
pushing first public version of the manuscript

@@
 and offset of M/EEG effects. We first assessed the performance of this
-approach on simulated data allowing us to evaluate the method's ability
+approach on simulated data, allowing us to evaluate the method's ability
 to recover ground-truth onset and offset values. We then assessed its

### 2025-08-29 manuscript.qmd
fixing some typos

@@
 Researchers are often interested in determining whether and when these signals differ across experimental conditions or participant groups.
-The conventional approach involves mass-univariate statistical testing across time and space followed by corrections for multiple comparisons or some form of cluster-based inference.
+The conventional approach involves mass univariate statistical testing across time and space followed by corrections for multiple comparisons or some form of cluster-based inference.
 While effective for controlling error rates at the cluster-level, cluster-based inference comes with a significant limitation: by shifting the focus of inference from individual time points to clusters, it prevents drawing conclusions about the precise onset or offset of observed effects.
@@
 Magnetoencephalography and electroencephalography (M/EEG) are two such methods, widely used in cognitive neuroscience for their ability to track brain activity at the millisecond scale.
-These techniques provide rich time series data that reflect how neural responses unfold in response to stimuli or tasks.
+These techniques provide rich timeseries data that reflect how neural responses unfold in response to stimuli or tasks.
 A central goal in many M/EEG studies is to determine whether, when, and where neural responses differ across experimental conditions or groups.
 
-The conventional approach involves mass-univariate statistical testing through time and/or space followed by some form of correction for multiple comparisons with the goal of maintaining the familywise error rate (FWER) or the false discovery rate (FDR) at the nominal level (e.g., 5%).
+The conventional approach involves mass univariate statistical testing through time and/or space followed by some form of correction for multiple comparisons with the goal of maintaining the family-wise error rate (FWER) or the false discovery rate (FDR) at the nominal level (e.g., 5%).
 Cluster-based inference is the most common way of achieving this sort of error control in the M/EEG literature, being the recommended approach in several software programs [e.g., `EEGlab`, @delorme2004; `MNE-Python`, @gramfort2013].
@@
 
-Scalp-recorded M/EEG signals capture neural activity originating from various brain regions and are often contaminated by artifacts unrelated to the cognitive processes under investigation.
+Scalp-recorded M/EEG signals capture neural activity originating from various brain regions and are often contaminated by artefacts unrelated to the cognitive processes under investigation.
 Consequently, analysing M/EEG data necessitates methods that can disentangle task-relevant neural signals from extraneous "noise." A widely adopted technique for this purpose is the estimation of event-related potentials (ERPs), which are stereotyped electrophysiological responses time-locked to specific sensory, cognitive, or motor events.
@@
 One widely used implementation of this approach is provided by the LIMO EEG toolbox [@pernet2011], which follows a multi-stage analysis pipeline.
-First, a separate regression model is fit for each datapoint (e.g., each time point and electrode) at the individual participant level to estimate ERP responses.
+First, a separate regression model is fit for each datapoint (e.g., each time point and electrode) at the individual level to estimate ERP responses.
 This is followed by group-level statistical analyses of the resulting regression coefficients, often accompanied by corrections for multiple comparisons or cluster-based inference [for recent applied examples, see @dunagan2025; @wüllhorst2025].
@@
 This approach allows for the estimation of smooth, data-driven functions that characterise how neural responses evolve over time, offering a flexible alternative to traditional linear models.
-In the following section, we provide a brief introduction to GAMs, highlighting their applicability to M/EEG time series analysis and the advantages they offer over conventional methods.
+In the following section, we provide a brief introduction to GAMs, highlighting their applicability to M/EEG timeseries analysis and the advantages they offer over conventional methods.
 
@@
 A detailed treatment of the technical underpinnings of GAMs is beyond the scope of this article [see reference books such as @hastie2017; @wood2017].
-However, it is worth emphasising that GAMs have been successfully applied to a wide range of time series data across the cognitive sciences, including pupillometry [e.g., @vanrij2019], articulography [e.g., @wieling2018], speech formant dynamics [e.g., @sóskuthy2021], neuroimaging data [e.g., @dinga2021], and event-related potentials [e.g., @abugaber2023; @baayen2018; @meulman2015; @meulman2023].
+However, it is worth emphasising that GAMs have been successfully applied to a wide range of timeseries data across the cognitive sciences, including pupillometry [e.g., @vanrij2019], articulography [e.g., @wieling2018], speech formant dynamics [e.g., @sóskuthy2021], neuroimaging data [e.g., @dinga2021], and event-related potentials [e.g., @abugaber2023; @baayen2018; @meulman2015; @meulman2023].
 Their appeal for modelling M/EEG data lies in their ability to flexibly capture the complex shape of ERP waveforms without overfitting, through the use of smooth functions constrained by penalisation.
@@
 We then compared the ability of the BGAM to accurately estimate the onset and offset of the ERP difference to other widely-used methods.
-First, we conducted mass-univariate t-tests (thus treating each timestep independently) and identified the onset and offset of the ERP difference as the first and last values crossing an arbitrary significance threshold ($\alpha = 0.05$).
+First, we conducted mass univariate t-tests (thus treating each timestep independently) and identified the onset and offset of the ERP difference as the first and last values crossing an arbitrary significance threshold ($\alpha = 0.05$).
 We then followed the same approach but after applying different forms of multiplicity correction to the $p$-values.
 We compared two methods that control the FDR [i.e., `BH95`, @benjamini1995; and `BY01`, @benjamini2001], one method that controls the FWER [i.e., Holm–Bonferroni method, @holm1979], and two cluster-based permutation methods [permutation with a single cluster-forming threshold and threshold-free cluster enhancement, `TFCE`, @smith2009].
-The `BH95`, `BY01`, and `Holm` corrections were applied to the p-values using the `p.adjust()` function in `R`.
+The `BH95`, `BY01`, and `Holm` corrections were applied to the $p$-values using the `p.adjust()` function in `R`.
 The cluster-based inference was implemented using a cluster-sum statistic of squared $t$-values, as implemented in `MNE-Python` [@gramfort2013], called via `r pkrt("reticulate")`.
-We also compared these estimates to the onset and offset as estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared $t$-values [as in @rousselet_using_2025].[^2] @fig-corrections illustrates the onsets and offsets estimated by each method on a single simulated dataset and shows that all methods systematically overestimate the true onset and underestimate the true offset.
+We also compared these estimates to the onset and offset estimated using the binary segmentation algorithm, as implemented in `r pkrt("changepoint")`, and applied directly to the squared $t$-values [as in @rousselet_using_2025].[^2] @fig-corrections illustrates the onsets and offsets estimated by each method on a single simulated dataset and shows that all methods systematically overestimate the true onset and underestimate the true offset.
 In addition, the `Raw p-value`, `FDR BH95`, and `FDR BY01` methods identify clusters well before the true onset and after the true offset.
@@
 
-```{r fig-corrections, echo = FALSE, results = "hide", fig.width = 8, fig.asp = 0.5, fig.cap = "Exemplary timecourse of squared t-values with true onset and offset (vertical black dashed lines) and onsets/offsets identified using the raw p-values, the corrected p-values (BH95, BY01, Holm), the cluster-based methods (Cluster mass, TFCE), or using the binary segmentation method (Change point)."}
+```{r fig-corrections, echo = FALSE, results = "hide", fig.width = 8, fig.asp = 0.5, fig.cap = "Exemplary timecourse of squared t-values with true onset and offset (vertical black dashed lines) and clusters identified using the raw p-values, the corrected p-values (BH95, BY01, Holm), the cluster-based methods (Cluster mass, TFCE), or using the binary segmentation method (Change point)."}
 # defining an arbitrary alpha threshold
@@
 Following @rousselet_using_2025, each participant was assigned a random onset between 150 and 170ms.
-Whereas the present article focuses on one-dimensional signals (e.g., one M/EEG channel), we provide a 2D application in @apx-2D.
+Whereas the present article focuses on one-dimensional signals (e.g., one M/EEG channel), we provide an application to 2D temporal data in @apx-2D.
 
@@
 In this study, the authors conducted time-resolved multivariate pattern analysis (MVPA, also known as decoding) of MEG data recorded in 32 human participants during a reading task.
-As a result, the authors obtained a timecourse of decoding accuray (ROC AUC), bounded between 0 and 1, for each participant.
+As a result, the authors obtained a timecourse of decoding accuracy (ROC AUC), bounded between 0 and 1, for each participant.
 To test whether the group-level average decoding accuracy was above chance (i.e., 0.5) at each timestep, we fitted a BGAM as introduced previously with a basis dimension $k=50$ and retained all timesteps exceeding a posterior odds of 20.
@@
 
-We then assessed the sensitivity of the various methods using a form of permutation-based sensitivity study, which consisted of the following steps.
+We then assessed the sensitivity of all methods using a form of permutation-based sensitivity study, which consisted of the following steps.
 First, we created a large number of split halves of the data, that is, subsets of the dataset containing only 16 out of 32 participants.
-For each possible pair of subsets, we have 16 possible levels of overlap/similarity that can be quantified using the Jaccard index, ranging from 0 (perfectly disjoints subsets) to $\approx0.88$ (identical subsets except one participant).
+For each possible pair of subsets, we have 16 possible levels of overlap that can be quantified using the Jaccard index, ranging from 0 (perfectly disjoints subsets) to $\approx0.88$ (identical subsets except one participant).
 For each of these 16 levels of Jaccard similarity, we created 1,000 pairs of subsets, resulting in 16,000 pairs of subsets in total.
@@
     labs(x = "Testing time (s)", y = "Training time (s)")
-```
-
-```{r mne-timegen, eval = FALSE, echo = FALSE}
-# defining the function in R (it will be executed in Python)
-mne_timegen <- function (X, labels, ncores = 8) {
-
-    # converting R dataframe to NumPy array (reshaping if needed)
-    # X should be a matrix before conversion
-    X_np <- np$array(X)
-
-    if (length(dim(X_np) ) == 2) {
-
-        # adding a second dimension (channels) if missing
-        X_np <- np$expand_dims(X_np, axis = as.integer(1) )
-
-    }
-
-    # defining the classifier
-    clf <- sklearn$linear_model$LogisticRegression(solver = "liblinear")
-
-    # sliding the estimator on all time frames
-    time_gen <- mne$decoding$GeneralizingEstimator(
-        clf,
-        n_jobs = as.integer(ncores),
-        scoring = "roc_auc",
-        verbose = FALSE
-        )
-
-    # or using N-fold cross-validation
-    scores <- mne$decoding$cross_val_multiscore(
-        time_gen,
-        X_np,
-        labels,
-        cv = as.integer(4),
-        n_jobs = as.integer(ncores),
-        verbose = FALSE
-        )
-
-    # returning the scores
-    return (scores)
-
-}
-
-# listing all participants
-participants <- unique(raw_df$participant)
-
-# initialising empty decoding results
-group_timegen_scores <- data.frame()
-
-# running decoding for each participant
-for (ppt in participants) {
-
-    # printing progress
-    print(ppt)
-
-    # retrieve data from one participant
-    ppt_data <- raw_df %>%
-        filter(participant == ppt) %>%
-        select(-participant) %>%
-        pivot_wider(names_from = time, values_from = eeg) %>%
-        select(-condition, -trial)
-
-    # extracting the labels
-    labels <- raw_df %>%
-        filter(participant == ppt) %>%
-        select(-participant) %>%
-        pivot_wider(names_from = time, values_from = eeg) %>%
-        pull(condition) %>%
-        as.numeric()
-
-    # extracting the timesteps
-    timesteps <- raw_df %>%
-        filter(participant == ppt) %>%
-        select(-participant) %>%
-        pull(time) %>%
-        unique()
-
-    # running the decoding
-    timegen_scores <- mne_timegen(
-        X = ppt_data,
-        labels = labels-1
-        )
-
-    # computing the average over CV folds
-    # a <- timegen_scores %>%
-    #     apply(MARGIN = c(2, 3), FUN = mean) %>%
-    #     data.frame()
-    timegen_scores <- data.frame(np$mean(timegen_scores, axis = as.integer(0) ) )
-
-    # sanity check
-    # image(as.matrix(timegen_scores) )
-
-    # appending to previous results
-    group_timegen_scores <- bind_rows(group_timegen_scores, timegen_scores)
-
-}
-
-# saving the scores
-saveRDS(object = group_timegen_scores, file = "results/timegen_scores.rds")
-
-# reshaping the data
-colnames(group_timegen_scores) <- unique(raw_df$time)
-timegen_data <- group_timegen_scores %>%
-    mutate(participant = rep(x = participants, each = n_distinct(raw_df$time) ) ) %>%
-    mutate(train_time = rep(x = unique(raw_df$time), times = length(participants) ) ) %>%
-    pivot_longer(
-        cols = -c(participant, train_time),
-        names_to = "test_time",
-        values_to = "auc"
-        ) %>%
-    mutate(test_time = as.numeric(test_time) )
 ```
@@
     )
+
+# gaussian alternative
+# timegen_gam <- brm(
+#     # 2D thin-plate spline (tp) with full penalties
+#     auc ~ t2(train_time, test_time, bs = "tp", k = 30, full = TRUE),
+#     data = timegen_data,
+#     family = gaussian(),
+#     warmup = 1000,
+#     iter = 5000,
+#     chains = 8,
+#     cores = 8,
+#     control = list(adapt_delta = 0.95, max_treedepth = 15),
+#     backend = "cmdstanr",
+#     stan_model_args = list(stanc_options = list("O1") ),
+#     file = "models/timegen_gam_meg_t2_k30_full_gaussian.rds"
+#     )
 ```
@@
 With the exception of the $k=5$ model, all other fits yield satisfactory PPCs, indicating that the predicted data closely resemble the empirical observations.
-However, model comparison using the leave-one-out information crtierio (LOOIC), as summarised in @tbl-simulation-results, identifies the $k=40$ model as the best-performing one in terms of LOOIC, closely followed by the $k=80$ model.
+However, model comparison using the leave-one-out information criterion (LOOIC), as summarised in @tbl-simulation-results, identifies the $k=40$ model as the best-performing one in terms of LOOIC, closely followed by the $k=80$ model.
 This suggests that the optimal number of basis functions likely lies between these two values.
@@
     # dataframe with M/EEG data in long format
-    data = raw_df,
+    data = df,
     # the *_id arguments are used to specify the relevant columns in data 
@@
 
-```{python, eval = FALSE, echo = TRUE}
-# loading the Python modules
-import rpy2.robjects as robjects
-from rpy2.robjects.packages import importr
-from rpy2.robjects import pandas2ri
-from rpy2.robjects.conversion import localconverter
-
-# importing the "neurogam" R package
-neurogam = importr("neurogam")
-
-# activating automatic pandas-R conversion
-pandas2ri.activate()
-
-# assuming reshaped_df is some M/EEG data reshaped in long format
-with localconverter(robjects.default_converter + pandas2ri.converter):
-
-    reshaped_df_r = robjects.conversion.py2rpy(reshaped_df)
-
-# using the testing_through_time() function from the neurogam R package
-gam_onset_offset = neurogam.testing_through_time(
-    data=reshaped_df_r,
-    threshold=20,
-    multilevel=False
-    )
-
-# displaying the results
-print(list(gam_onset_offset) )
-```
-
 ```{python, eval = FALSE, echo = FALSE}
@@
 ```
+
+```{python, eval = FALSE, echo = TRUE}
+# loading the Python modules
+import rpy2.robjects as robjects
+from rpy2.robjects.packages import importr
+from rpy2.robjects import pandas2ri
+from rpy2.robjects.conversion import localconverter
+
+# importing the "neurogam" R package
+neurogam = importr("neurogam")
+
+# activating automatic pandas-R conversion
+# pandas2ri.activate()
+
+# assuming reshaped_df is some M/EEG data reshaped in long format
+with localconverter(robjects.default_converter + pandas2ri.converter):
+
+    reshaped_df_r = robjects.conversion.py2rpy(reshaped_df)
+
+# using the testing_through_time() function from the neurogam R package
+gam_onset_offset = neurogam.testing_through_time(
+    data=reshaped_df_r,
+    threshold=20,
+    multilevel=False
+    )
+
+# displaying the results
+print(list(gam_onset_offset) )
+```

### 2025-08-29 manuscript.tex
fixing some typos

@@
 signals differ across experimental conditions or participant groups. The
-conventional approach involves mass-univariate statistical testing
+conventional approach involves mass univariate statistical testing
 across time and space followed by corrections for multiple comparisons
@@
 their ability to track brain activity at the millisecond scale. These
-techniques provide rich time series data that reflect how neural
+techniques provide rich timeseries data that reflect how neural
 responses unfold in response to stimuli or tasks. A central goal in many
@@
 
-The conventional approach involves mass-univariate statistical testing
+The conventional approach involves mass univariate statistical testing
 through time and/or space followed by some form of correction for
-multiple comparisons with the goal of maintaining the familywise error
+multiple comparisons with the goal of maintaining the family-wise error
 rate (FWER) or the false discovery rate (FDR) at the nominal level
@@
 Scalp-recorded M/EEG signals capture neural activity originating from
-various brain regions and are often contaminated by artifacts unrelated
+various brain regions and are often contaminated by artefacts unrelated
 to the cognitive processes under investigation. Consequently, analysing
@@
 fit for each datapoint (e.g., each time point and electrode) at the
-individual participant level to estimate ERP responses. This is followed
-by group-level statistical analyses of the resulting regression
+individual level to estimate ERP responses. This is followed by
+group-level statistical analyses of the resulting regression
 coefficients, often accompanied by corrections for multiple comparisons
@@
 traditional linear models. In the following section, we provide a brief
-introduction to GAMs, highlighting their applicability to M/EEG time
-series analysis and the advantages they offer over conventional methods.
+introduction to GAMs, highlighting their applicability to M/EEG
+timeseries analysis and the advantages they offer over conventional
+methods.
 
@@
 \citeproc{ref-wood2017}{Wood, 2017a}). However, it is worth emphasising
-that GAMs have been successfully applied to a wide range of time series
+that GAMs have been successfully applied to a wide range of timeseries
 data across the cognitive sciences, including pupillometry (e.g.,
@@
 onset and offset of the ERP difference to other widely-used methods.
-First, we conducted mass-univariate t-tests (thus treating each timestep
+First, we conducted mass univariate t-tests (thus treating each timestep
 independently) and identified the onset and offset of the ERP difference
@@
 \texttt{BH95}, \texttt{BY01}, and \texttt{Holm} corrections were applied
-to the p-values using the \texttt{p.adjust()} function in \texttt{R}.
-The cluster-based inference was implemented using a cluster-sum
-statistic of squared \(t\)-values, as implemented in \texttt{MNE-Python}
-(\citeproc{ref-gramfort2013}{Gramfort, 2013}), called via the \texttt{R}
-package \texttt{reticulate} v 1.42.0 (\citeproc{ref-reticulate}{Ushey et
-al., 2024}). We also compared these estimates to the onset and offset as
-estimated using the binary segmentation algorithm, as implemented in the
-\texttt{R} package \texttt{changepoint} v 2.3
-(\citeproc{ref-changepoint}{Killick et al., 2022}), and applied directly
-to the squared \(t\)-values (as in
+to the \(p\)-values using the \texttt{p.adjust()} function in
+\texttt{R}. The cluster-based inference was implemented using a
+cluster-sum statistic of squared \(t\)-values, as implemented in
+\texttt{MNE-Python} (\citeproc{ref-gramfort2013}{Gramfort, 2013}),
+called via the \texttt{R} package \texttt{reticulate} v 1.42.0
+(\citeproc{ref-reticulate}{Ushey et al., 2024}). We also compared these
+estimates to the onset and offset estimated using the binary
+segmentation algorithm, as implemented in the \texttt{R} package
+\texttt{changepoint} v 2.3 (\citeproc{ref-changepoint}{Killick et al.,
+2022}), and applied directly to the squared \(t\)-values (as in
 \citeproc{ref-rousselet_using_2025}{Rousselet, 2025}).\footnote{As in
@@
 \caption{\label{fig-corrections}Exemplary timecourse of squared t-values
-with true onset and offset (vertical black dashed lines) and
-onsets/offsets identified using the raw p-values, the corrected p-values
-(BH95, BY01, Holm), the cluster-based methods (Cluster mass, TFCE), or
-using the binary segmentation method (Change point).}
+with true onset and offset (vertical black dashed lines) and clusters
+identified using the raw p-values, the corrected p-values (BH95, BY01,
+Holm), the cluster-based methods (Cluster mass, TFCE), or using the
+binary segmentation method (Change point).}
 
@@
 article focuses on one-dimensional signals (e.g., one M/EEG channel), we
-provide a 2D application in \hyperref[apx-2D]{Appendix~A}.
+provide an application to 2D temporal data in
+\hyperref[apx-2D]{Appendix~A}.
 
@@
 decoding) of MEG data recorded in 32 human participants during a reading
-task. As a result, the authors obtained a timecourse of decoding accuray
-(ROC AUC), bounded between 0 and 1, for each participant. To test
-whether the group-level average decoding accuracy was above chance
+task. As a result, the authors obtained a timecourse of decoding
+accuracy (ROC AUC), bounded between 0 and 1, for each participant. To
+test whether the group-level average decoding accuracy was above chance
 (i.e., 0.5) at each timestep, we fitted a BGAM as introduced previously
@@
 
-We then assessed the sensitivity of the various methods using a form of
+We then assessed the sensitivity of all methods using a form of
 permutation-based sensitivity study, which consisted of the following
@@
 participants. For each possible pair of subsets, we have 16 possible
-levels of overlap/similarity that can be quantified using the Jaccard
-index, ranging from 0 (perfectly disjoints subsets) to \(\approx0.88\)
+levels of overlap that can be quantified using the Jaccard index,
+ranging from 0 (perfectly disjoints subsets) to \(\approx0.88\)
 (identical subsets except one participant). For each of these 16 levels
@@
 the empirical observations. However, model comparison using the
-leave-one-out information crtierio (LOOIC), as summarised in
+leave-one-out information criterion (LOOIC), as summarised in
 Table~\ref{tbl-simulation-results}, identifies the \(k=40\) model as the
@@
     \CommentTok{\# dataframe with M/EEG data in long format}
-    \AttributeTok{data =}\NormalTok{ raw\_df,}
+    \AttributeTok{data =}\NormalTok{ df,}
     \CommentTok{\# the *\_id arguments are used to specify the relevant columns in data }
@@
 \CommentTok{\# activating automatic pandas{-}R conversion}
-\NormalTok{pandas2ri.activate()}
+\CommentTok{\# pandas2ri.activate()}
 

### 2025-08-29 code/edf_estimation.R
cleaning up the repo

@@ -1,113 +1,0 @@
-###########################################################
-# Comparing various EDF estimates                         #
-# Written by Ladislas Nalborczyk                          #
-# Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on May 22, 2025                            #
-###########################################################
-
-library(tidyverse)
-library(bamlss)
-library(brms)
-
-# importing the reshaped MEG decoding results
-decoding_df <- read.csv(file = "data/decoding_results_reshaped.csv") %>%
-    # removing some participants
-    mutate(participant_id = cur_group_id(), .by = participant) %>%
-    filter(participant_id < 33) %>%
-    select(-participant_id)
-
-# averaging across participants
-decoding_summary_df <- decoding_df %>%
-    summarise(
-        auc_mean = mean(auc),
-        auc_sd = sd(auc),
-        .by = c(participant, time)
-        )
-
-# setting possible k values
-k_values <- seq(from = 10, to = 50, by = 5)
-
-# initialising en empty dataframe to store the results
-results <- data.frame()
-
-# fitting the model for each k
-for (k_value in k_values) {
-    
-    # printing progress
-    cat("\nProcessing k =", k_value, "\n")
-    
-    # fitting the model with bamlss
-    b_model <- bamlss(
-        formula = auc_mean ~ s(time, bs = "tp", k = k_value),
-        family = beta_bamlss,
-        data = decoding_summary_df
-        )
-    
-    # computing the EDF
-    ss <- samplestats(samples = b_model)
-    
-    # constructing the smooth term dynamically
-    smooth_term <- glue::glue("s(time, bs = 'tp', k = {k_value})")
-    
-    # full formula
-    formula_str <- glue::glue("auc_mean ~ {smooth_term}")
-    
-    # convert to formula
-    formula_obj <- brms::bf(formula_str)
-    
-    # fitting the Beta GAM
-    meg_decoding_gam <- brm(
-        formula = formula_obj,
-        data = decoding_summary_df,
-        family = brms::Beta(),
-        warmup = 1000,
-        iter = 4000,
-        chains = 4,
-        cores = 4
-        )
-    
-    # computing the p_WAIC
-    pwaic <- waic(meg_decoding_gam)
-    looic <- loo(meg_decoding_gam)
-    
-    # combining the estimates in a dataframe
[…]

### 2025-08-29 code/meta_gam_errors_cluster.R

@@ -1,190 +1,0 @@
-###########################################################
-# Monte-Carlo simulation of onset/offset error properties #
-# Written by Ladislas Nalborczyk                          #
-# Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on April 9, 2025                           #
-###########################################################
-
-# importing R packages
-library(tidybayes)
-library(dplyr)
-library(tidyr)
-library(brms)
-
-# setting up parallel workers
-library(purrr)
-library(furrr)
-
-# timing the simulations
-library(tictoc)
-tic()
-
-# total number of total CPU cores available on the HPC node
-total_cores <- 12
-
-# number of cores per brms model
-cores_per_model <- 4
-
-# number of parallel simulations to run at once
-n_parallel <- total_cores / cores_per_model
-
-# setting up parallel workers
-future::plan(future::multisession(workers = n_parallel) )
-
-# number of simulations to run per combination
-nsims <- 100
-
-# importing the data generation functions 
-source(file = "code/generate_eeg_data.R")
-source(file = "code/utils.R")
-
-# defining a function to run a single iteration of the simulation
-single_iteration <- function (i) {
-    
-    # sanity check
-    system(
-        sprintf(
-            "echo 'Running simulation %d on PID %d with %s cores' >> ./sim_log.txt",
-            i, Sys.getpid(), cores_per_model
-            )
-        )
-    
-    # initialising an empty dataframe
-    temp_results <- data.frame()
-    
-    # number of posterior samples to use to compute the posterior probability ratio
-    # n_post_samples <- 1e4
-    
-    # defining the parameter grid
-    param_grid <- crossing(
-        kvalue = c(5, 10, 15, 20, 25, 30, 35, 40),
-        threshold = seq(from = 1, to = 50, by = 1)
-        )
-    
-    for (j in seq_len(nrow(param_grid) ) ) {
-        
-        # retrieving the current kvalue and threshold
-        kvalue <- param_grid$kvalue[j]
-        threshold <- param_grid$threshold[j]
-        
-        # printing progress
-        message_parallel(sprintf("Running sim %d, k = %d, threshold = %d", i, kvalue, threshold) )
-        
-        # when we assess a new sim_id, we generate new data
-        if (kvalue == 5 & threshold == 1) {
-    
-            # generate some data
-            message_parallel("\nSimulating new data...\n")
-            raw_df <- generate_data(
-                n_trials = n_trials, n_ppt = n_ppt, outvar = outvar,
[…]

### 2025-08-29 code/peaks_across_splits.R

@@ -1,120 +1,0 @@
-###########################################################
-# Defining subsets of participants with uniform Jaccard   #
-# Written by Ladislas Nalborczyk                          #
-# Contact: ladislas.nalborczyk@gmail.com                  #
-# Last updated on July 15, 2025                            #
-###########################################################
-
-# importing R packages
-library(tidyverse)
-library(signal)
-
-# for setting up the cluster
-library(doParallel)
-library(foreach)
-
-# total number of total CPU cores available on my laptop
-total_cores <- 12
-
-# registering parallel backend
-cl <- makeCluster(total_cores)
-registerDoParallel(cl)
-
-# importing home-made helper functions
-source("code/utils.R")
-
-# importing the data splits list
-# split_df <- read.csv(file = "uniform_jaccard_data_splits.csv")
-split_df <- read.csv(file = "uniform_jaccard_data_splits_100pairs.csv")
-
-# importing the reshaped MEG decoding results
-raw_df <- read.csv(file = "data/decoding_results_reshaped.csv") |>
-    # removing the last participant (to get a round number of participants)
-    mutate(participant_id = cur_group_id(), .by = participant) |>
-    dplyr::filter(participant_id < 33) |>
-    select(-participant_id)
-
-# averaging across participants
-summary_df <- raw_df |>
-    summarise(
-        auc_mean = mean(auc),
-        auc_sd = sd(auc),
-        auc_se = sd(auc) / sqrt(n() ),
-        .by = c(participant, time)
-        )
-
-# creating a dataframe with all (split_number, group) combinations to iterate over
-split_group_combinations <- crossing(
-    split_number = split_df$split_id,
-    group = split_df$group
-    ) |>
-    # creating a custom simulation ID per group
-    mutate(split_id = sprintf("s%05d_%s", split_number, group) )
-
-# for each simulation
-reliability_results <- foreach(
-    # i = seq_len(n_splits), .combine = bind_rows,
-    i = 1:nrow(split_group_combinations), .combine = bind_rows,
-    .packages = c("dplyr", "tidyr", "signal")
-    ) %dopar% {
-    
-        # extracting split_number and group for this simulation
-        current_split <- split_group_combinations$split_number[i]
-        current_group <- split_group_combinations$group[i]
-        current_split_id <- split_group_combinations$split_id[i]
-        
-        # printing progress
-        message_parallel(sprintf("\nRunning simulation: %s (%d out of %d)...", current_split_id, i, nrow(split_group_combinations) ) )
-        
[…]
